### 2017-05-03 main.tex
added Seth's notes from our meeting to define the new direction of the paper.

\documentclass[11pt]{article}
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1.7}
\usepackage[utf8]{inputenc}
\usepackage{lineno}

%%%%%%%%%%%%%%%%%%%%%
% LaTeX packages
%%%%%%%%%%%%%%%%%%%%%
\usepackage{siunitx}
\usepackage{graphicx}
% Please be sparing in your use of additional LaTeX packages, and
% upload any required style files to Editorial Manager with the file
% type "LaTeX ancillary files (.sty, .bst)."

%%%%%%%%%%%%%%%%%%%%%
% Line numbering
%%%%%%%%%%%%%%%%%%%%%
\usepackage{lineno}
% Please use line numbering with your initial submission and
% subsequent revisions. After acceptance, please comment out 
% the commands \usepackage{lineno}, \linenumbers{} 
% and \modulolinenumbers[3] below.

%%%%%%%%%%%%%%%%%%%%%
% File path for figures
%%%%%%%%%%%%%%%%%%%%%
% Depends on 'graphicx' package
\graphicspath{ {figures/} }

\title{Ecological convergence via parallel speciation}
%\title{Coevolution of competitors can destabilize food webs}
%\title{Ecological character displacement can destabilize food-web dynamics}

%%%%%%%%%%%%%%%%%%%%%
% Authorship
%%%%%%%%%%%%%%%%%%%%%
% Please remove authorship information while your paper is under review,
% unless you wish to waive your anonymity under double-blind review. 
% Remember to uncomment the information after acceptance.

\author{Seth M. Rudman$^{1,2,\ast}$ \\ 
Matthew A. Barbour$^{2,3}$ \\ 
Julian Heavyside$^{2}$ \\
Dolph Schluter$^{2}$}
%Dolph Schluter$^{2,\ddag}$}

\date{}

\begin{document}

\maketitle

\noindent{}1. Department of Biology, University of Pennsylvania 102 Leidy Laboratories Philadelphia, PA USA 19104;

\noindent{}2. Department of Zoology, University of British Columbia 4200-6270 University Blvd. Vancouver, B.C. Canada V6T1Z4;

\noindent{}3. Department of Evolutionary Biology and Environmental Studies, University of Zurich 190 Winterthurerstrasse 8006 Zurich Switzerland.

\noindent{}$\ast$ Corresponding author; e-mail: rudman@sas.upenn.edu.

%\noindent{}$\ddag$ ORCIDs: Cook, 0000-0002-1234-5678; Collaborator, 0000-0001-5678-9123; Expert, 0000-0003-3456-789X.

\bigskip

\textit{Manuscript elements}: Figure~1, figure~2, table~1, online
appendices~A and B (including figure~A1 and figure~A2). Figure~2 is to
print in color.

\bigskip

\textit{Keywords}: Examples, model, template, guidelines.

\bigskip

\textit{Manuscript type}: Article. 
% Or e-article, note, e-note, natural history miscellany,
% e-natural history miscellany, comment, reply, symposium, or
% countdown to 150.

\bigskip

\noindent{\footnotesize Prepared using the suggested \LaTeX{} 
template for \textit{Am.\ Nat.}}

\linenumbers{}
\modulolinenumbers[3]

\newpage{}

\section*{Abstract}

Parallel speciation, the repeated evolution of the same modes of reproductive isolation, has fascinated evolutionary biologists because it provides strong evidence for the role of natural selection in genesis of new species (Futuyma 1986, Schluter and Nagel 1995, Schluter 2000).  In cases of parallel speciation there is often evidence of repeated evolution of ecologically important traits associated with resource use (e.g. Williams et al. 1972, Losos et al. 1998, Matthews et al. 2010, Elmer et al. 2014).  Convergent evolution in traits associated with resource use could cascade to alter community structure, creating the potential for community convergence.  Here we use the replicated ecological speciation observed in threespine stickleback fish to assess whether parallel speciation leads to community convergence in their prey.  A field study and mesocosm experiment combine to illustrate that stickleback speciation leads to convergence in the zooplankton and aquatic larval insect communities.  For example, zooplankton biomass showed a 13-fold reduction in natural lakes where speciation has occurred.  This convergence is also measurable across the aquatic-terrestrial boundary as we observed a 24\% increase in insect emergence associated with speciation.  This convergence demonstrates the profound impacts that evolutionary change can have on communities in a natural ecosystem and hints at a potential role of eco-evolutionary dynamics in ecological speciation.

The process of ecological speciation, whereby reproductive isolation between populations evolves as a by product of local adaptation, is the primary mechanism by which natural selection generates species diversity.  Yet current models of ecological speciation largely do not account for reciprocal interactions between evolutionary change and ecology, limiting our understanding of the speciation process.  Here we present the results of a two species consumer-resource model that illustrates how evolution in consumer species can drive evolution in resource species that may then feedback to dictate divergence between consumer species.  We pair this model with a large-scale observational field study and an evolutionarily replicated mesocosm experiment to assess whether ecological speciation in consumers alters resource dynamics in nature.  As predicted by the model, we found that divergence between consumer taxa alters community structure and stability in natural ecosystems, even across ecosystem boundaries.  As a whole, our study provides evidence that the process of ecological speciation can have profound effects on natural communities and suggests a currently under appreciated role for eco-evolutionary dynamics in the process of ecological speciation.

notes:
speciation as a driver of diversity and the effects of diversity on ecology
specialization as a process that can reduce stability and potentially

Intro
diversity-stability and effects of speciation on ecology

paragraph on ecological speciation, how it tends to be associated with specialization, and utility of repeated speciation

paragraph stickleback biology, cases of repeated evolution leading to repeated community structure as powerful evidence for determinism in ecology, and suitability for this question

paragraph where we lay out hypotheses using morphology and model. Model for resource dyanmics and stability.

\newpage{}

\section*{Introduction}

% The journal does not have numbered sections in the main portion of
% articles. Please refrain from using section references such as
% section~\ref{section:CountingOwlEggs}, and refer to sections by name
% (e.g. section ``Counting Owl Eggs'').

Community convergence is recognized when multiple independent communities reach the same state, often measured as relative abundances or biomasses of constituent species within a community (Schluter 1986; Grover and Lawton 1994).  Community convergence is one possible outcome of the process of community assembly, which has long been hypothesized to be governed by “assembly rules” (Cody and Diamond 1975; Gotelli and McCabe).  However, convergence is not inevitable or even likely.  There are numerous biotic and abiotic factors that influence ecological communities and even when these factors are largely similar, there is evidence that there can be multiple stable equilibrium states for a given set of ecological conditions (Cody and Diamond 1975; Weiher and Keddy 1995; Chase 2003).  As such, cases of community convergence can be viewed as striking evidence for the presence of deterministic processes in community ecology (Cody and Diamond 1975).

Evolutionary biologists likewise use repeatability to deduce mechanisms from patterns in nature.  Parallel speciation, a special case of parallel evolution where the same mechanisms of reproductive isolation evolve independently, has been recognized as powerful evidence for the role of natural selection in evolution (Williams 1972; Schluter and Nagel 1995; Losos et al. 1998).  The repeated evolution of ecomorphs, species specialized to use particular microhabitats or to feeding on particularly prey species, are amongst the best studied cases of parallel speciation.  Despite a wealth of research documenting striking morphological and genetic parallelism of replicate ecomorphs, there has been no investigation of whether repeatedly evolved forms lead to ecological convergence.  If functional traits of dominant species are a key factor influencing ecological communities (Bolnick et al. 2011; Crutsinger et al. 2014; Levine 2016), repeated evolution of specialized morphology and diet should lead to convergence in species interactions between independent environments .  However, this potential link between ecological and evolutionary convergence has not been explored.

Threespine stickleback (Gasterosteus aculeatus) present an ideal case to test for community convergence stemming from repeated speciation.  Stickleback have undergone parallel speciation in four small (17-45 ha) coastal lakes in Southwestern British Columbia (Gow et al. 2008), with each representing an independent speciation event.  Each of these lakes contain two sympatric species of stickleback, commonly called ‘benthic’ and ‘limnetic’ stickleback (collectively called a ‘species pair’), that differ from each other genetically (Jones et al. 2012), phenotypically (Schluter and McPhail 1992), and ecologically (Schluter and McPhail 1992; Matthews et al. 2010).  Benthic stickleback are deep bodied fish that inhabit the nearshore environment and feed primarily on benthic invertebrates, particularly aquatic larval insects.  Limnetic stickleback mostly forage on zooplankton in the pelagic zone.  The divergence between benthic and limnetic species has evolved due, in part, to ecological character displacement, a process whereby competition for resources causes divergence (Schluter 1995; Rundle et al. 2000; Arnegard et al. 2014).   The vast majority of small coastal lakes in the region contain stickleback populations that have not undergone speciation.  Stickleback in these lakes are intermediate in morphology and diet relative to the benthic and limnetic specialists (Schluter and McPhail 1992; Rudman and Schluter 2016), hence these are referred to as ‘generalist’ populations.  An extensive study on the physical and chemical properties of lakes containing species pairs and generalist populations of stickleback found broad overlap in these features and no significant differences between them (Ormond et al. 2011). The repeated nature of diversification and the presence of similar lakes where speciation has not occurred allows for a replicated assessment of the role of speciation in altering ecology and potentially driving community convergence.

To understand how parallel ecological speciation of threespine stickleback impacts community structure and ecosystem function we conducted both a replicated mesocosm experiment and extensive community sampling of six natural lakes.  The experiment and the lake sampling were designed to assess how speciation alters community structure by comparing three lakes where stickleback have independently undergone speciation with nearby and physically similar lakes that contain only a generalist form of stickleback (Ormond et al. 2011).  The diet of benthic and limnetic ecotypes of stickleback differ substantially, with benthic ecotypes consuming more substrate dwelling invertebrates and limnetics ingesting more zooplankton (Schluter and McPhail 1992; Matthews et al. 2010; Behm et al. 2010).  In comparison, the morphology and diet of stickleback from generalist populations are intermediate compared to benthic and limnetic specialists (Schluter and McPhail 1992; Rudman and Schluter 2016).  The repeated evolution of a zooplankton specialist in the species pairs is predicted to lead to increased trophic impacts on the zooplankton community, including, decreases in zooplankton biomass, average size, and changes in zooplankton movement patterns associated with escaping fish predation.  Furthermore, we predicted that the repeated evolution of a benthic form during stickleback speciation would be associated with increased trophic impact on the aquatic larval insect community.  We anticipated an increase in the number of aquatic larval insects emerging from the aquatic environments where speciation has occurred, as previous studies have shown that increased insect emergence associated with increased benthic foraging likely stemming from facilitation (Rudman and Schluter 2016; Rudman et al. 2016).

% Please note that we prefer (\citealt{Xiao2015}) to \citep{Xiao2015},
% since \citep{} inserts a comma after "et al."

\section*{Methods}

\subsection*{Mathematical Model}

We modeled the dynamics of two consumers indirectly competing through the consumption of two, patchily distributed resources using the following equations:
\[\frac{dR_i}{dt}=rR_i\left ( 1-\frac{R_i}{K} \right )-F_{ii}\left ( a_{ii},a_{ij},h,w,R_i,R_j \right )C_i-F_{ji}\left ( a_{ji},a_{jj},h,w,R_i,R_j \right )C_j\]\[\frac{dC_i}{dt}=eF_{ii}\left ( a_{ii},a_{ij},h,w,R_i,R_j \right )C_i+eF_{ij}\left ( a_{ii},a_{ij},h,w,R_i,R_j \right )C_i-mC_i\]
In the absence of consumers, resource $i$ exhibits logistic population growth, which is determined by its intrinsic growth rate ($r$) and carrying capacity ($K$). In the presence of consumers, the functional response (feeding rate) of each consumer on resource $i$ ($F_{ii}$, $F_{ji}$) negatively affects resource growth rates. Population growth rate of consumer $i$ is positively influenced by its functional response on each resource ($F_{ii}$ and $F_{ji}$) and its efficiency ($e$) at converting resources into new consumers, but negatively affected by a density-independent rate of mortality ($m$).

Our model departs from other models of character displacement that explicitly examine resource dynamics in that our functional response describes a more realistic foraging scenario for consumers (McCann et al. (2005). This functional response assumes that each resource is restricted to a separate habitat patch of equal size. This functional response also assumes that consumers are mobile and rapidly switch between habitat patches depending on the relative densities of the resources. Here, a ‘rapid switch’ is relative to a consumer’s population growth rate, which enables the spatially-implicit model to approximate a spatially-explicit model (McCann et al. 2005). In addition, this spatially-implicit model assumes that a consumer only perceives resources as well mixed at the scale of the habitat patch, resulting in the following functional response equations:

\[F_{ii}=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}hW_{ii}R_i+a_{ij}hW_{ij}R_j}\]

Similar to other multi-species functional response models, we assumed that feeding rate of consumer $i$ on resource $i$ depends on its handling time ($h$), attack rates on both resources ($a_{ii}$, $a_{ij}$). However, we also , and habitat preference ($w$). where is the relative preference for consumer $i$ to feed in the habitat where resource $j$ is found when resources $i$ and $j$ are at equal densities. All parameters in our model were constrained to be positive, except that w was constrained to be between 0 and 1.

\[W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}\]

\subsection*{How does resource competition affect the evolution of consumer attack rates?} 
We used an Adaptive Dynamics approach to analyze the evolution of consumer attack rates. We allowed attack rates to evolve because of their fundamental importance in determining the consumer’s functional response, which is why prior models have also examined character displacement in this key trait. Following the methods outlined in Otto and Day (2007), we augmented the above model to track the dynamics of a mutant attack-rate allele for a given consumer (e.g. for :that appears in a population fixed for the resident allele (e.g. for :. We then identified the conditions that would enable a consumer with the mutant attack-rate allele to have a positive population growth rate, and subsequently invade the system. We assumed that the mutant allele had no effect on resource population dynamics because the mutant was at low density. We also assumed that the mutant allele had only a small effect on attack rate and can be written as aii,m=ii+aii.

\subsection*{How  does the evolution of consumer attack rates affect food-web dynamics? } 
Analytical Solution -- To analyze the ecological dynamics, we identified equilibrium conditions and examined their local stability using standard techniques (Otto and Day 2007). In the supplementary material, we give a detailed analysis of all model equilibria; however, we focus our results in the main text on the equilibria communities where competition is present ($C_i$, $C_j$, $R_i$, $R_j$) and absent ($C_i$, $R_i$, $R_j$). We did this because this comparison enabled us to determine the effects of ecological character displacement on community dynamics.  For the two-consumer, two-resource equilibrium, we used Routh-Hurwitz criteria to examine its local stability properties (Otto and Day 2007). The Routh-Hurwitz criteria specify when equilibria are locally stable without requiring explicit solving of the eigenvalues of the stability matrix, and are therefore useful for complicated stability matrices.

\subsection*{Mesocosm Experiment}

We set up an array of 41 1136L mesocosms, which were filled, seeded with sterilized play sand, zooplankton (community dominated by copepods with some cladocerans (e.g. Bosmina, Daphnia) present), and benthic mud from nearby experiment ponds, and allowed to settle for ~4 weeks.  Mesocosms were then randomly assigned to one of three treatments: species pair, generalist, and a fish-free control. Stickleback for the experiment were wild-caught and held in the lab for at least one week before being placed into mesocosms.  Treatments for each of the three independently evolved species pairs were replicated nine times, for a total of 27 replicates containing benthic and limnetic specialists.  The generalist treatment was composed of fish taken from 3 different lakes replicated 4, 3, and 2 times for a total of 9 mesocosms.  Five tanks were set up as fish-free controls.

 We used minnow traps to collect stickleback from six lakes containing three types of stickleback in early April 2011: benthic and limnetic specialists from Little Quarry, Paxton and Priest lakes with independently derived species pairs and generalists from three neighbouring lakes (Cranby, Kirk, and Trout) containing a single species of stickleback.  Generalist populations were chosen based on three criteria: 1) the fish community had to match that of species pair lakes, which contain only stickleback and cutthroat trout 2) they had to be roughly similar in size 3) they had to be within a 30km radius of the species pair lakes.  Handling of stickleback in this experiment adhered to an approved University of British Columbia animal care protocol (no. 11-0402).  Fish were collected with permission from Fisheries and Oceans Canada (SARA permit no. 197) and Ministry of Forests, Lands, and Natural Resource Operations of British Columbia (permit no. NA/SU10-68002).

Following fish introduction on May 3rd, 2011, mesocosms were surveyed at least once per day for dead or dying stickleback, which were replaced within 12 hours until the conclusion of the experiment on June 30th, 2011.  To match biomass approximately between treatments containing fish, different numbers of stickleback were stocked (generalist\=4, species pair\=2 benthic \& 3 limnetic).  Species pair treatments received more limnetics than benthics to simulate relative abundances in the natural lakes (D. Schluter unpublished data) and as a way to match biomass (4-5g/mesocosm) between treatments, since limnetics are comparatively small-bodied (Schluter and McPhail 1992). We sampled the communities of phytoplankton, zooplankton and benthic invertebrates, and we counted insects that emerged from the mesocosms at the end of the study.  We examined decomposition, photosynthetically active radiation (PAR), dissolved organic carbon (DOC) and gross primary productivity (GPP) to monitor components of ecosystem function.

We treated lake as a fixed effect in our statistical analyses for the mesocosm experiment, recognizing the small number of stickleback species pairs and the single aggregate treatment for generalist lakes. Our goal here is to evaluate the prediction that community convergence is present within this set of lakes. Comparisons used planned contrasts between the means of species pairs and the generalist aggregate, which served as comparison of variation between independent species pair lakes and variation between species pair and generalists lakes.  Following a previously established method (Schluter 1986), we interpreted significant differences as evidence for convergence.

\subsection*{Field Data Collection}

We collected zooplankton samples from the same six lakes that were used in the mesocosm experiment in May, June, July, August, and October 2012, and in February 2013 (three containing species pairs, three containing generalists).  In each lake we collected two five-meter and one ten-meter vertical zooplankton samples using a 30cm diameter zooplankton tow constructed from 80 $\mu$m mesh with a Cod end.  Many zooplankton species vertically migrate below the photic zone when fish are present, likely to reduce the risk of predation by fish (or other visual predators) (Zaret and Suffern 1976).  Samples were collected at five and ten meter depths to get a basic estimate of the proportion of zooplankton biomass below the photic zone.   Vertical migration behavior was assessed by the relative proportion of zooplankton found in the 10-meter sample compared to the two 5-meter samples.   All zooplankton samples were sub-sampled to 1/16th volume, stained with Rose bengal’s solution, then counted with the length of each individual zooplankton measured using an ocular micrometer.  We also collected data on dissolved oxygen, temperature, turbidity, conductivity, and physical size for each lake.

We used floating traps to collect insects as they emerged from the lake environment from two independent watersheds (five total lakes; 3 species pair and 2 generalist) in July and August 2013.  We constructed cone-shaped floating traps (33 cm in diameter) using wire and 400$\mu$m mesh.  We placed two floating traps in the shallows (~1.5m depth) and one at an intermediate depth (~2-3.5m depth).  Traps were emptied the following morning using a modified hand vacuum (BioQuip) and insects were deposited directly into vials containing 95\% ethanol.  Floating traps were placed on paired species pair and generalist lakes on the same nights to reduce any differences caused by nightly weather variation.  Lakes containing species pairs of stickleback were paired with the nearest lake containing generalist stickleback (>1 km and in the same watershed in two cases and ~20km in the third).  Insect emergence data were analyzed as the difference between paired lakes. Each insect was measured and identified to the lowest readily identifiable taxonomic unit.

We used passive echolocation recording equipment (Wildlife Acoustic SM2BAT+ with SMZ-US microphone) to quantify bat activity above each lake during 12 nights in July and August 2013.  We placed recording equipment at the edge of a pair of lakes on the same night (i.e. one recording setup at one lake containing a species pair and one at a lake containing generalist fish) and oriented the recording equipment out over the open water.  Recordings began at 10pm each night and were stopped at 6am and were taken on the same nights as insect emergence samples.  We used callViewer software (Wildlife Acoustics) to manually count and identify calls from a subset of the recordings.  The only genera recorded were Myotis and Eptesicus.  Using the manually counted files as a guide, we used frequency and amplitude information for each recorded call to count the total number of calls from both Myotis (80-40kHz) and Eptesicus (34-25kHz).  We used the ‘seewave’ package in R to transform wave files and perform a fast fourier transformation before automated counting was done in R.

We used repeated measures ANOVAs to test for convergence in the zooplankton community over time (biomass, average size, abundance of large zooplankton, and zooplankton migratory behavior) in lakes with independently evolved species pairs of stickleback relative to neighbouring lakes containing a single generalist stickleback species.  We also carried out community composition analyses using a jaccard dissimilarity index on samples collected each month from each lake.  The effects of stickleback community, month, and the interaction between these factors was assessed using a PERMAnova implement in the ‘vegan’ R package (Anderson 2001).  In addition, we assessed differences in the major groups of zooplankton, cladocerans and copepods, to assess whether We analyzed bat and insect data using a paired design, with two species pair lakes grouped with one generalist lake from the nearest watershed. Insect emergence and bat data were analyzed using paired t-tests between generalist and species pair lakes from the same areas to partially control for location and weather.

\section*{Results}

\subsection*{Coevolution of competitors destabilizes a model food web}
Resource competition resulted in consumers evolving specialized attack rates on different resources \textemdash \ the signature of ecological character displacement (Fig. \ref{Fig:theory}A). In general, consumer $i$ will evolve when $a_{ii,m}wR_i^2 + a_{ij,m}(1-w)R_j^2>a_{ii}wR_i^2+a_{ij}(1-w)R_j^2$. Assuming there is a trade-off in attack rates on different resources (e.g., if $a_{ii,m}>a_{ii}$ then $a_{ij,m}<a_{ij}$), then consumer $i$ will tend to evolve higher attack rates on the most abundant resource or, if resources are equally abundant, on their preferred resource ($R_i$ if $w>0.5$; $R_j$ if $w < 0.5$). However, resource abundances and consumer preference impose conflicting selection pressures on attack rates. This occurs because increasing attack rate on the preferred resource decreases the abundance of this resource, which also decreases the fitness benefit. In the absence of a competitor, these conflicting selection pressures result in consumers evolving a generalist strategy that can more equally exploit both resources. In the presence of a competitor though, the equilibrium density of each resource will be the same (derivation in appendix); therefore, consumers will evolve specialized attack rates on their preferred resource. Ecological character displacement also resulted in consumers evolving a higher effective attack rate, $a_{ii}w+a_{ij}(1-w)$, in general (Fig. \ref{Fig:theory}B).

The evolution of character displacement had a profound effect on ecological dynamics, resulting in both lower resource densities (Fig. \ref{Fig:theory}C) and a less stable food web (Fig. \ref{Fig:theory}D). Lower resource densities occur because character displacement increases the effective attack rate of consumers, which plays a key role in determining the equilibrium density of resources when a competitor is present ($\widehat{R_i}=\frac{m}{(a_{ii}w+a_{ij}(1-w))(e-hm)}$, derivation in appendix). In addition to altering resource densities, the effective attack rate critically determines the local stability of the food web. Specifically, the food web is locally stable if $a_{ii}w + a_{ij}(1-w) < \frac{e+hm}{hK(e-hm)}$ (derivation in appendix; dotted line in Fig. \ref{Fig:theory}D); therefore, the evolution of higher effective attack rates pushes the system toward its boundary of stability (Fig. \ref{Fig:theory}D). Note that even when both food webs are locally stable (region above dotted line in Fig. \ref{Fig:theory}D), it takes longer for the food web with a competing consumer to return to its equilibrium density following a perturbation, as indicated by the consistently lower level of stability ($-1\times \text{Re}(\lambda_\text{max})$). This decreased stability is not simply an ecological effect of the additional consumer, but is magnified by coevolution (figure in appendix).

%In real food webs, which are frequently experiencing perturbations, this should correspond to %something about the stochastic process resulting in higher CV, and reference Gellner and McCann's paper 2017 Theor. Pop. Biol. paper on the Duality of Stability.

% * <matthew.a.barbour@gmail.com> 2017-04-20T06:26:09.012Z:
% 
% > presence of competition
% How do I explain the fact that resource densities decrease when C1 evolves lower effective attack rate in the absence of competition (Fig. 1B,C)...
% 
% ^ <matthew.a.barbour@gmail.com> 2017-04-23T12:22:43.589Z.

% Although we’ve illustrated this result with a linear trade-off, a non-linear trade-off function does not qualitatively alter the fact that character displacement is inevitable. However, if the trade-off is nonlinear and convex, then the magnitude of character displacement will be less, simply because the fitness benefit of evolving a specialist strategy is less.  
% * <matthew.a.barbour@gmail.com> 2017-04-20T05:57:24.382Z:
% 
% > Although we’ve illustrated this result with a linear trade-off, a non-linear trade-off function does not qualitatively alter the fact that character displacement is inevitable. However, if the trade-off is nonlinear and convex, then the magnitude of character displacement will be less, simply because the fitness benefit of evolving a specialist strategy is less.  
% Need to demonstrate this in the appendix.
% 
% ^ <matthew.a.barbour@gmail.com> 2017-04-23T12:23:09.708Z.
% * <matthew.a.barbour@gmail.com> 2017-04-20T05:47:10.727Z:
% 
% > $wR_i^2 +(1-w)R_j^2=0$ 
% Might be more meaningful to also include attack rate details here. This is because it will again emphasize the importance of effective attack rate. It will also keep the equation in its most general form.
% 
% ^ <matthew.a.barbour@gmail.com> 2017-04-23T12:23:16.894Z.

\subsection*{Empirical}

For all measured aspects of the zooplankton and emerging insect communities trends were the same in mesocosms and natural lakes, but were only significant in the lakes were the effects were uniformly larger.  Mesocosms containing species pairs of stickleback showed trends towards reduced zooplankton biomass (Fig. 1a) and reduced average zooplankton size (Fig. 1b) compared to those housing generalist stickleback.  These effects were greatly magnified in natural lakes, where zooplankton biomass was 13-fold lower ($F_{1,5}=42.50, P=0.001$, Fig. 1c) and average size was ~2.5-fold smaller ( F1,5=83.77, p=0.0003, Fig. 1d) in species pair lakes than in generalist lakes.  Lakes without species pairs of stickleback had 8-fold greater biomass of large zooplankton (F1,5=28.46, p<0.00001), which we classified as individuals with a calculated dry weight of >10ug.  The effects of stickleback speciation were visible on the two dominant zooplankton taxa, cyclopoid and calanoid copepods, which were ~2.5 fold and ~1.5 fold smaller respectively in species pair lakes than in generalist lakes.  We also observed differences in the biomass distribution of zooplankton by depth ( F1,5=11.12, p=0.021), with a higher proportion of the zooplankton biomass found at depth in lakes with species pairs of stickleback.  This pattern is consistent with increased vertical migration of zooplankton in lakes where stickleback have undergone speciation; zooplankton vertical migration is associated with avoidance of visual predators, including zooplanktivorous fish (Zaret and Suffern 1976).  There was a significant effect of stickleback speciation on community composition (F1,24=2.87, p=0.007), however speciation did not have an interactive effect on composition with time (F1,24=0.90, p=0.644).  These effects of speciation on community composition were also observed in an increase in the proportion of zooplankton biomass made up of sub-adult copepods (F1,5=9.38, p=0.0022), which fits with expectations of increased zooplanktivory leading towards a shift to more evasive and smaller prey.

Mesocosms containing species pairs of stickleback showed a trend towards increased insect emergence (37\% greater emergence than those containing generalist fish (t=1.00, df=29, p=0.32, Fig. 2a).  These effects were more consistent in the field where lakes containing species pairs of stickleback had an average of 24\% more insects emerging per night in July (t=-8.27, df=2, p=0.014, Fig. 2b), but not in August.

We found that the presence of stickleback species pairs had some effects on ecosystem function, as stickleback speciation was associated with decreases in total nitrogen in the mesocosms (t=-3.12, df=27, p=0.004).  Lakes housing species pair and generalist stickleback were found to broadly overlap in their physical and chemical features (Fig. S1).  The lack of clear differences in any physical or chemical parameters of the lakes themselves suggests that differences in the zooplankton and emerging insect community are unlikely to stem from systematic differences between the species pair and generalist lake environments.

\section*{Discussion}

Parallel evolution has long been of interest to evolutionary biologists because it provides strong evidence for adaptation (Williams 1972).  Here we demonstrate that parallel speciation, a subset of parallel evolution that includes repeated evolution of reproductive barriers, can itself alter ecology in predictable ways.  We found that parallel speciation of threespine stickleback was associated with convergent changes in prey community structure within natural aquatic ecosystems.  These prey community changes, particularly in zooplankton biomass and size, were convergent across cases of parallel speciation compared to nearby lakes that contained a single generalist population of stickleback.  All the lakes in the study contain a simple fish community of stickleback and cutthroat trout and broadly overlap in their physical and chemical features (Fig. S1, Ormond et al. 2011), making the comparison of lakes containing species pair and generalist populations of stickleback a unique opportunity to assess the effects of speciation in a natural environment.  In addition, the community patterns we observed in the field were similar in direction to those documented in a replicated experiment, providing additional evidence that these convergent community changes are resultant from stickleback diversification.  Furthermore, the shifts in biomass, zooplankton size, and vertical migration fit with expectations based on previous trophic cascade studies on the impacts of zooplanktivory on zooplankton communities (Brooks and Dodson 1965; Zaret and Suffern 1976; Henrikson et al. 1980; Post et al. 2008).

The direction of the community convergence observed in the mesocosms and natural lakes was the same, but the difference in magnitude was striking.  A number of factors could be responsible for these stronger field patterns.  One such factor is the inherent limitation of a mesocosm environment, which does not offer the habitat diversity or spatial separation between habitats that is found in a natural lake.  Another plausible cause is a natural systematic difference in density between lakes containing species pair and generalist stickleback.  If ecological character displacement leads to more efficient resource partitioning it may also increase the total carrying capacity, causing a difference in biomass between lakes houses one and two species of stickleback.  Our mesocosm experiment standardized density between treatments, which may have removed part of the effects of ecological speciation.  Age and size structuring may also contribute to the observed differences.  The mesocosm experiment was conducted with adult fish and the natural lakes contain fish of a variety of sizes.  This particularly true in late summer when a large percentage of fish are juveniles; that was also the time period when the most striking convergence was observed.

In addition to the effects we observed on the zooplankton community within the aquatic ecosystem, we also documented the effects of stickleback speciation cascading across the ecosystem boundary.  We observed an increase in emerging insect abundance in July, which was concordant with the direction of the data from the experimental system.  We also explored how far the effects of stickleback speciation extends into the terrestrial system by recording bat foraging above lakes.  Given the increase in insect abundance associated with stickleback speciation and data demonstrating that a significant proportion of prey for some common bat species diet are emerged aquatic insects (LaVal et al. 1977; Clare et al. 2014), we predicted greater bat foraging over lakes where speciation had occurred.  Although we observed a trend in this direction, we didn’t see a significant difference in bat foraging associated with stickleback speciation.  Insect and bat data show a weaker pattern of convergence than was observed in the zooplankton community.  This may be due to an attenuation effect, as emerging aquatic insects have non-aquatic life stages and hence interact less strongly with stickleback than zooplankton.  Overall, these cross-ecosystem effects demonstrate the extent to which speciation within one dominant taxa can influence community dynamics of an ecosystem.

Our observations of community convergence associated with speciation also have implications for our understanding of ecological speciation as eco-evolutionary processes.  Local adaptation to different ecological conditions is the foundation of ecological speciation, whereby barriers to gene flow evolve between populations due to divergent or disruptive selection (Schluter 2000).  In cases of ecological character displacement, the strength of disruptive selection is strongly influenced by the strength of competition, which is in turn influenced by both functional trait values and the distribution of the available resources (Taper and Case 1992; Abrams et al. 2008).  This suggests a feedback where competition promotes trait divergence and trait divergence alters resource availability, which then feeds back to alter the strength of competition.  Here we demonstrate that repeated speciation by the process of ecological character displacement has lead to convergent depletion of prey resources for stickleback.  Alterations of resources may be key to facilitating divergence, as a prey based composed of more evasive prey could favor specialization.  Future directed experiments assessing both the interplay between resource availability, trait change, and reproductive isolation could further illuminate this eco-evolutionary dynamic providing a greater understanding of the processes that shape communities, local adaptation, and ecological speciation.

\section*{Conclusion}

Duis pharetra enim at libero cursus, eu commodo mi vestibulum. Nullam 
eget velit nec lectus viverra sodales. Suspendisse egestas, eros at 
dictum tincidunt, mi orci laoreet libero, eget rutrum sapien arcu 
blandit odio.

%%%%%%%%%%%%%%%%%%%%%
% Acknowledgments
%%%%%%%%%%%%%%%%%%%%%
% You are encouraged to remove the Acknowledgments section while
% your paper is under review (unless you wish to waive your anonymity
% under double-blind review) if the Acknowledgments reveal your
% identity. If you remove this section, you will need to add it back
% in to your final files after acceptance.

\section*{Acknowledgments}

OEC would like to thank the world. GHC is much indebted to 
the solar system. AQE was supported by a generous grant from 
the Milky Way (MW/01010/987654).

%\newpage{}

%\renewcommand{\thesection}{\Alph{section}}

%\section*{Online Appendix A: Supplementary Figures}

% Subsection numbering is permitted (but by no means necessary) in 
% online appendices. Please note that if you have sections (thus Online
% Appendix A, B, and C), these will become three separate online PDFs.
% You may wish to consolidate these into one PDF (hence one section,
% divided into subsections as necessary). Please reset counters for
% each such section.

%\renewcommand{\theequation}{A\arabic{equation}}
% redefine the command that creates the equation number.
%\renewcommand{\thetable}{A\arabic{table}}
%\setcounter{equation}{0}  % reset counter 
%\setcounter{figure}{0}
%\setcounter{table}{0}

%\subsection*{Fox--dog encounters through the ages}

%The quick red fox jumps over the lazy brown dog. The quick red fox has 
%always jumped over the lazy brown dog. The quick red fox began jumping 
%over the lazy brown dog in the 19th century and has never ceased from so 
%jumping, as we shall see in figure~\ref{Fig:Jumps}.

%[Figure A1 goes here.]

%[Figure A2 goes here.]

%\newpage{}

%\section*{Online Appendix B: Additional Methods}

%\renewcommand{\theequation}{B\arabic{equation}}
% redefine the command that creates the equation number.
%\setcounter{equation}{0}  % reset counter 
%\renewcommand{\thetable}{B\arabic{table}}
%\setcounter{figure}{0}
%\setcounter{table}{0}

%To determine how this enhanced resource exploitation affected community dynamics, we first identified the equilibrium density of consumers and resources in the presence of competition (details in Supplement):
%\[\widehat{R_i}=\frac{m}{(a_{ii}w+a_{ij}(1-w))(e-hm)}\]
%\[\widehat{C_i}=\frac{r(K-R_i)(hR_i(a_{ii}w+a_{ij}(1-w))}{K(a_{ii}w+a_{ij}(1-w))}\]
%The term $a_{ii}w + a_{ij}(1-w)$ describes the effective attack rate of consumers on resources (hereafter, $a_{effect}$ and plays a key role in determining both resource and consumer densities. Ecological character displacement results in an increase in this effective attack rate (FIGURE). Looking at the equilibrium equations, it is straightforward to see that increases in the effective attack rate of consumers will decrease the equilibrium density of resources; therefore, ecological character displacement will result in the suppression of resource densities. For consumers, increasing effective attack rate from relatively low values will increase the equilibrium density of the consumer; however, once this effective attack rate becomes relatively high, increasing specialization will result in lower equilibrium density of the consumer.

%To determine how increases in the effective attack rate of consumers affected community stability, we used two approaches. First, we used numerical simulations to identify how increases in effective attack rate altered community resilience, defined by the return time of the system following an infinitesimally small perturbation (details in Methods and Supplement). Second, we used Routh-Hurwitz criteria to identify the conditions under which the system switched qualitatively in its local stability. While increases in effective attack rates can stabilize community dynamics when attack rates are initially low, as resources become suppressed well below their carrying capacity, this destabilizes community dynamics (FIGURE). We found that the consumer-resource dynamics switch from a steady-state equilibrium to a limit cycle when: $a_{ii}w + a_{ij}(1-w) > \frac{e+hm}{hK(e-hm)}$. Biologically, this means that increasing the effective attack rates of the consumers, all else equal, will push the system toward its boundary of stability. Figure 1 illustrates graphically the effect of increasing aii on local stability. If we hold the other parameters constant, increasing $a_{ii}$ shifts the consumers’ zero net growth isocline to the left relative to the position of the resources’ isocline. Biologically, this means that increasing specialization increases the capacity of the consumers to exhibit a positive growth rate even when resources are at low densities. This results in the consumer suppressing the resources well below their carrying capacity. If this suppression is large enough, this causes the consumers and resources to exhibit a limit cycle (Fig. 1).

%\subsection*{Measuring the height of fox jumps without a meterstick}

%Pellentesque ac nibh placerat, luctus lectus non, elementum mauris. 
%Morbi odio velit, eleifend ut hendrerit vitae, consequat sit amet 
%nulla. Pellentesque porttitor vitae nisl quis tempus. Pellentesque 
%habitant morbi tristique senectus et netus et malesuada fames ac 
%turpis egestas. Praesent ut nisi odio. Vivamus vel lorem gravida 
%odio molestie volutpat condimentum et arcu.

%\begin{equation}
%{ \frac{1}{N_k-1} \sum \limits_{t=1}^{N_k} (M_{tjk} - \bar{M}_{jk})^2}
%\end{equation}

%\subsection*{Quantifying the brownness of the dog}

%Pellentesque eu nulla odio (\citealt{Xiao2015,CookEtAl2015}). Nulla 
%aliquam porta metus, quis malesuada orci faucibus quis. Suspendisse nunc 
%magna, tristique sit amet sollicitudin nec, elementum et lacus. Sed 
%vitae elementum mi. In hac habitasse platea dictumst. Etiam eu tortor 
%elit. Sed ac tortor purus. Aliquam volutpat, odio sit amet posuere 
%pretium, dolor ex interdum ante, sed luctus quam eros ac nulla.

%\begin{equation}
%{ (\sum \limits_{p=1}^P {n_{sp}})^{-1}\sum \limits_{p=1}^P {n_{sp}Q_{p}}}
%\end{equation}

\newpage{}

%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%
% You can either type your references following the examples below, or
% compile your BiBTeX database and paste the contents of your .bbl file
% here. The amnatnat.bst style file should work for this---but please
% email the journal office at amnat at uchicago dot edu if you run into
% any hitches with it!
% The list below includes sample journal articles, book chapters, and
% Dryad references.

%\bibliographystyle{amnatnat}
\begin{thebibliography}{}

\bibitem[{Cook et~al.(2015)Cook, Collaborator, and Expert}]{CookEtAl2015}
Cook, O.~E., G.~H. Collaborator, and A.~Q. Expert. 2015.
\newblock Data from: Template and guidelines for using \LaTeX{} 
in \textit{The American Naturalist}.
\newblock American Naturalist, Dryad Digital Repository, 
http://dx.doi.org/10.5061/dryad.XYZAB.

\bibitem[{Davis et~al.(2011)Davis, Brakora, and Lee}]{DavisEtAl2011}
Davis, E.~B., K.~A. Brakora, and A.~H. Lee. 2011.
\newblock Evolution of ruminant headgear: a review.
\newblock Proceedings of the Royal Society B 278:2857--2865.

\bibitem[{Inglis et~al.(2011)Inglis, Roberts, Gardner, and Buckling}]{Ing11}
Inglis, R.~F., P.~G. Roberts, A.~Gardner, and A.~Buckling. 2011.
\newblock Spite and the scale of competition in \textit{Pseudomonas
  aeruginosa}.
\newblock American Naturalist 178:276--285.

\bibitem[{Lemod\`{e}le et~al.(2007)Lemodele, Kapitelschreiber, 
and Exemplar}]{LemKapEx07}
Lemod\`{e}le, P.-Q., A.~B. Kapitelschreiber, and C.~D.~E. Exemplar. 2007.
\newblock An exemplary instance of chapters in books.
\newblock Pages 231--245 \emph{in} J.-P. \'{E}crivain and M.~A. 
Term\'{e}szettud\'{o}s, eds. Inspiring Instances of Brilliant Writing. 
Truth Pudding Press, Fond du Lac, WI.

\bibitem[{Xiao et~al.(2015)Xiao, McGlinn, and White}]{Xiao2015}
Xiao, X., D.~J. McGlinn, and E.~P. White. 2015.
\newblock A strong test of the maximum entropy theory of ecology.
\newblock American Naturalist 185:E705--E80.

\end{thebibliography}

%\newpage{}

%\section*{Tables}
%\renewcommand{\thetable}{\arabic{table}}
%\setcounter{table}{0}

%\begin{table}[h]
%\caption{Animals in various cities with equations}
%\label{Table:Okapi}
%\centering
%\begin{tabular}{llc}\hline
%Animal    & City         & Equation \\ \hline
%Dog       & Springfield  & $x+y=z$ \\
%Fox       & Indianapolis & $2x+2y=2z$ \\
%Okapi$^a$ & Chicago      & $x-y<z$ \\
%Badger    & Madison      & $x+2y>z$ \\ \hline
%\end{tabular}
%\bigskip{}
%\\
%{\footnotesize Note: Table titles should be short. Further details 
%should go in a `notes' area after the tabular environment, as shown 
%here. $^a$ Okapis are not native to Chicago, but they are to be met with 
%in both of the major Chicagoland zoos.}
%\end{table}

\newpage{}

%%%%%%%%%%%%%%%%%%%%%
% Figure legends
%%%%%%%%%%%%%%%%%%%%%
% Please include all figure legends in a separate section at the end of 
% the document. If you use \label{} and \ref{} to refer to your figures,
% these can still work even if you comment out the %includegraphics{}
% line. If you refer to figures as "fig. 1" (etc.) manually, the
% legends can also appear simply as paragraphs.
% For submission, please upload the relevant figure files separately to
% Editorial Manager; Editorial Manager should insert them at the end of
% the PDF automatically.
% Figure legends should be concise, though they can be longer than the
% titles of tables.

\section*{Figure legends}

\begin{figure}[ht]
\includegraphics[scale = 0.6]{fig_theory.pdf}
\caption{Coevolution of competitors destabilizes a model food web. In the presence of a competitor, $C_2$, each consumer specializes on the resource it is pre-adapted to, whereas in the absence of competition, a consumer will evolve a generalist phenotype that can more equally exploit both resources (A). This ecological character displacement causes consumers to evolve a higher effective attack rate (B), which suppresses the density of both resources relative to the food web with no competing consumer (C). This enhanced resource exploitation with coevolving competitors results in a relatively less stable food web (D). The grey line in (D), where $-1\times \text{Re}(\lambda_{max})=0$, corresponds to the local stability boundary of the food web. Non-evolving model parameters for this simulation: $r=1$, $K=4$, $e=0.8$, $h=0.4$, $w=0.6$, $m=1$. Initial evolving parameters and species densities when "Evolutionary time = 0": $a_{ii}=1.93$, $a_{ij}=1.2$; $R_1=R_2=2$, $C_1=C_2=1$.}
\label{Fig:theory}
\end{figure}

\begin{figure}[ht]
\includegraphics[trim=1cm 8cm 1cm 8cm, scale=0.8]{repeat_Fig1_161109.pdf}
\caption{Panels A and B illustrate the effects of stickleback speciation on log zooplankton biomass and average size from the mesocosm experiment.  Small triangles show the mean of each independently evolved species pair of stickleback.  Panels C and D show the effects of stickleback speciation on log zooplankton biomass and average size across the year from the field study.}
\label{Fig:Zoop}
\end{figure}

\newpage

\begin{figure}[ht]
%\includegraphics[scale = 0.3]{meso_EmergedLogAbun_160119.pdf}
%\includegraphics[scale = 0.3]{Field_Insect_Emerge_170403.pdf}
\includegraphics[scale = 0.4]{Field_Bat_170403.pdf}
\caption{Panel A shows the number of insects emerging from the aquatic environment in the mesocosm experiment.  Panel B shows the difference in insect emergence in observational field data between paired lakes, with negative values indicating greater emergence in the species pair lakes.  Panel C shows the difference in microchiropteran foraging activity between paired lakes, with negative values indicating greater foraging above species pair lakes.}
\label{Fig:AnotherFigure}
\end{figure}

%\subsection*{Online figure legends}

%\renewcommand{\thefigure}{A\arabic{figure}}
%\setcounter{figure}{0}

%\begin{figure}[h!]
%\includegraphics{jumps20m}
%\caption{\textit{A}, the quick red fox proceeding to jump 20~m straight 
%into the air over not one, but several lazy dogs. \textit{B}, the quick 
%red fox landing gracefully despite the skepticism of naysayers.}
%\label{Fig:Jumps}
%\end{figure}

%\begin{figure}[h!]
%\includegraphics{jumps-nr-okapi}
%\caption{The quicker the red fox jumps, the likelier it is to land near 
%an okapi. For further details, see \citet{LemKapEx07}.}
%\label{Fig:JumpsOk}
%\end{figure}

%\renewcommand{\thefigure}{B\arabic{figure}}
%\setcounter{figure}{0}

\end{document}

### 2018-03-29 code/03_sim_3sp_C2_evo.R
updates

@@ -1,3 +1,5 @@
+### NEED TO RERUN WITH MUT.SIZE!!! 
+
 #### Description: 3 species model with C2 evolution simulation
 
 ## source required functions ----

### 2018-03-29 parameterized_CR_model.R

@@ -1,0 +1,178 @@
+
+## load required functions ----
+source('code/00_general_parameters.R')
+source('code/00_numerical_fxns.R')
+
+mass_scaling <- function(W, intercept, exponent) {
+  log_rate <- intercept + exponent * log(W)
+  return(exp(log_rate))
+}
+
+## BODY SIZE (g) ----
+
+g_Daphnia_pulex <- 1.6 * 10e-4  # from Table 1 in Fenchel 1974 (they cite Frank et al. 1957)
+g_Chironomid <- 1.6 * 10e-3 # very rough estimate, need to confirm
+
+g_limnetic <- 0.81 # from Schluter 1993, Ecology Table 1
+g_benthic <- 1.48  # from Schluter 1993, Ecology Table 1
+g_solitary <- 1.12 # from Schluter 1993, Ecology Table 1
+
+
+## INTRINSIC GROWTH RATE (r) ----
+# r = birth rate per day
+# proportional to M^(-1/4) from Rip and McCann 2011 (they cite Brown et al. 2004)
+# using relationship for heterotherm metazoa from Fenchel 1974, Oecologia
+
+r_Daphnia_pulex <- 0.3          # from Table 1 in Fenchel 1974 (they cite Frank et al. 1957)
+
+r_zoops_estimate <- mass_scaling(W = g_Daphnia_pulex, intercept = -1.6391, exponent = -0.2738) 
+r_zoops_estimate/r_Daphnia_pulex # much higher than Frank et al. 1957 estimate...I'm going to stick with the mass-based estimates though to get the relative differences accurate between zooplankton and benthic inverts
+
+r_benthos_estimate <- mass_scaling(W = g_Chironomid, intercept = -1.6391, exponent = -0.2738) 
+
+
+#### CARRYING CAPACITY (K) ----
+# proportional to M^(-3/4) from Rip and McCann 2011 (they cite Enquist et al. 1998 and Brown et al. 2004)
+
+K_intercept <- -14.29 # from Brown et al. 2004, Fig. 6
+K_zoops_estimate <- mass_scaling(W = g_Daphnia_pulex, intercept = K_intercept, exponent = -0.77); K_zoops_estimate
+K_benthos_estimate <- mass_scaling(W = g_Chironomid, intercept = K_intercept, exponent = -0.77); K_benthos_estimate
+
+
+#### CONVERSION EFFICIENCY (e) ----
[…]

### 2018-04-01 code/main.tex

--- 2017-05-03 main.tex
@@
 This convergence demonstrates the profound impacts that evolutionary change can have on communities in a natural ecosystem and hints at a potential role of eco-evolutionary dynamics in ecological speciation.
-
-The process of ecological speciation, whereby reproductive isolation between populations evolves as a by product of local adaptation, is the primary mechanism by which natural selection generates species diversity.
-Yet current models of ecological speciation largely do not account for reciprocal interactions between evolutionary change and ecology, limiting our understanding of the speciation process.
-Here we present the results of a two species consumer-resource model that illustrates how evolution in consumer species can drive evolution in resource species that may then feedback to dictate divergence between consumer species.
-We pair this model with a large-scale observational field study and an evolutionarily replicated mesocosm experiment to assess whether ecological speciation in consumers alters resource dynamics in nature.
-As predicted by the model, we found that divergence between consumer taxa alters community structure and stability in natural ecosystems, even across ecosystem boundaries.
-As a whole, our study provides evidence that the process of ecological speciation can have profound effects on natural communities and suggests a currently under appreciated role for eco-evolutionary dynamics in the process of ecological speciation.
-
-notes:
-speciation as a driver of diversity and the effects of diversity on ecology
-specialization as a process that can reduce stability and potentially
-
-Intro
-diversity-stability and effects of speciation on ecology
-
-paragraph on ecological speciation, how it tends to be associated with specialization, and utility of repeated speciation
-
-paragraph stickleback biology, cases of repeated evolution leading to repeated community structure as powerful evidence for determinism in ecology, and suitability for this question
-
-paragraph where we lay out hypotheses using morphology and model. Model for resource dyanmics and stability.
+TEST
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[h!]
 \includegraphics[scale = 0.6]{fig_theory.pdf}
@@
 
-\begin{figure}[ht]
+\begin{figure}[h!]
 \includegraphics[trim=1cm 8cm 1cm 8cm, scale=0.8]{repeat_Fig1_161109.pdf}
@@
 
-\begin{figure}[ht]
-%\includegraphics[scale = 0.3]{meso_EmergedLogAbun_160119.pdf}
-%\includegraphics[scale = 0.3]{Field_Insect_Emerge_170403.pdf}
-\includegraphics[scale = 0.4]{Field_Bat_170403.pdf}
+\begin{figure}[h!]
+%\includegraphics{elegance}
 \caption{Panel A shows the number of insects emerging from the aquatic environment in the mesocosm experiment.

### 2018-04-01 parameterized_CR_model.R
added rfishbase

@@ -3,6 +3,8 @@
 source('code/00_general_parameters.R')
 source('code/00_numerical_fxns.R')
 
+library(rfishbase)
+
 mass_scaling <- function(W, intercept, exponent) {
   log_rate <- intercept + exponent * log(W)
   return(exp(log_rate))

### 2018-05-10 ECD_Supp_Mat.Rmd
updating code

---
title: "ECD_Supp_Mat"
output: html_document
author: Matt Barbour
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How does character displacement affect resource dynamics?

Despite the numerous models that have been developed to examine the ecological conditions resulting in character displacement, none make predictions about how character divergence influences resource abundances or stability (except perhaps Lawlor and Smith 1976). This is a bit surprising, given that explicit modeling of resource dynamics has been shown to give key insight to the process leading to character divergence.

Intuitively, you might expect character divergence to decrease resource abundances. The rationale being that consumers are evolving to be specialists that are more effective at attacking a particular resource. This however is not necessarily true. In fact, the effect of character displacement on resource dynamics depends on both the foraging scenario and trade-off in attack rates. For example, let's start with the simplest model of competition that explicitly models resource dynamics is MacArthur's model of competition (MacArthur 1972):

$$\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-a_{11}R_1C_1-a_{21}R_1C_2$$
$$\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-a_{12}R_2C_1-a_{22}R_2C_2$$
$$\frac{dC_1}{dt}=e_{11}a_{11}R_1C_1+e_{12}a_{12}R_2C_1-m_1C_1$$
$$\frac{dC_2}{dt}=e_{21}a_{21}R_1C_2+e_{22}a_{22}R_2C_2-m_2C_2$$

Now, let's examine the simplest scenario where resources are equivalent ($r=r_i$ and $K=K_i$) and consumers have equivalent conversion efficiencies ($e=e_{ii}=e_{ij}$) except that their attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$). It can be shown that, at the equilibrium where both consumers and resources are present, resource abundances are equivalent ($R_i=R_1=R_2$) and are determined by the following equation:

$$\hat{R_i}=\frac{m}{e(a_{ii}+a_{aij})}$$
Note that a key determinant of resource abundance is the consumer's total attack rate, $a_{ii}+a_{ij}$. An increased in total attack rate results in a decrease in resource abundances. Now let's assume that the evolution of consumer attack rates are constrained by a linear trade-off, i.e. a unit increase in $a_{ii}$ results in a unit decrease in $a_{ij}$. Although such a trade-off results in divergent character displacement (Abrams 1986), there is no net change in resource abundances when consumer's diverge. Although this equilibrium equation for resource abundances was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when there is only one consumer as well. This is likely because a consumer, in the absence of competition, evolves to be a generalist that has equal attack rates on each resources, resulting in equivalent resource abundances.

Now let's consider a different type of trade-off in consumer attack rates. We know from past work that consumers still undergo character divergence regardless of the shape of the trade-off, even though the trade-off does influence the magnitude of displacement (Lawlor and Smith 1976). We specified the following trade-off function (Sargent and Otto, 2006):

$$\Big(\frac{a_{ii}}{A}\Big)^n+\Big(\frac{a_{ij}}{A}\Big)^n=1$$
where $A$ is the total investment in attack rates and $n$ describes the shape of the trade-off. This trade-off has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$). When $n=1$, the trade-off function is linear, and all combinations of $a_ii$ and $a_ij$ have the same total attack rate.

```{r Trade-off Function, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

```

In contrast, the shape of the trade-off qualitatively affects the relationship between divergence and resource abundance. As mentioned previously, a linear trade-off results in no change in resource densities. If the trade-off is convex, then resource densities can actually increase under character displacement! This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{aij}$) and decreases as consumers diverge. Note though that detecting this effect in nature may be difficult, since a convex trade-off results in relatively small displacement relative to other trade-off forms (Lawlor and Smith 1976). When the trade-off is concave, however, we see that character divergence suppresses resource densities.

Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (i.e. resources occur in the same microhabitat). This assumption may be valid for some foraging scenarios, e.g. Darwin's finches foraging for seeds. However, in many situations, consumers forage for resources that occur in distinct microhabitats (McCann et al. 2005). The only chracter displacement model that we are aware of that modeled resources in distinct microhabitat was one examined by Lawlor and Smith (1976). They found that consumers still underwent character divergence, regardless of the foraging scenario. Interestingly, we find that the foraging scenario qualitatively affects the relationship between character divergence and resource abundances. Specifically, if consumers are competing for resources that occur in distinct microhabitats, then consumer divergence always results in resource suppression. This occurs because now the consumers effective attack rate is influenced by its habitat preference. Note also that this resource suppression occurs regardless of the shape of the trade-off, although the shape of the trade-off can amplify the effect of character divergence on resource supression. The type of trade-off does affect the amount of character displacement though, and actually, we see again the a concave trade-off results in less displacement and less of an effect on resource abundances; whereas a convex trade-off results in the largest displacement and the largest effect on resource abundances. This effect of spatial context on the relationship between character displacement and resource abundances appears to be quite general. For example, in the appendix, we analyze a model that was designed to capture the spatial coupling of consumers in many food-web context (McCann 2005). Here, we again find that divergence results in a negative effect on resource abundances.

We would like to note that the scenario where the relationship between character divergence and resource suppression is the strongest is when the trade-off in attack rates is concave and consumers are competiting for resources that occupy distinct habitat types. Given that prior work in threespine stickleback has shown a concave trade-off in feeding rates (Schluter 1993; Arnegard et al. 2014) and that stickleback ecotypes are competing for resources that occur in distinct microhabitats (benthic vs. limnetic), we predict that character divergence in sticklebacks would supress resource densities.

This relationship between character divergence and resource abundances begs the question: how does character divergence affect food-web stability? This question has been addressed empirically by Pimentel as well as with a conceptual model. Lawlor and Smith though did a detailed theoretical analysis of how character divergence affects three different aspects of stability. They looked at stability through the lens of competition. The first aspect was whether character divergence resulted in mutual invasibility of consumers, a common metric of "stable coexistence" in competition models. Another very interesting aspect they looked at was a global metric that examined the persistence of the system to perturbations in resources. Here, they conclude the divergence increases community stability, because it expands the range of resource perturbations that a system can withstand before a consumer going extinct. Note though, that they overlooked that there conclusions do not actually quantify the effects of ecological character displacement on stability because they do not compare scenarios where a consumer is in allopatry and one where consumers are in sympatry. Finally, they also look at the Lyapunov exponent (local stability), and while they find a negative effect of divergence, they suggest that it would likely be overshadowed by the stabilizing effects of divergence on mutual invasibility and global stability.

Here we focus on local stability, which makes the most sense in a food-web context. We find that divergence follows the typical "check-mark" stability pattern that has been described recently in theoretical food-web models.

Importantly, this metric of stability allows us to compare the stability of three-species vs. four-species system, which is key for understanding how ecological character displacement affects food-web dynamics.

## Testing the generality of our predictions

So far, our theoretical predictions of the effects of consumer divergence on resource dynamics has considered a simplistic scenario, where resources are equivalent and consumers have equivalent conversion efficiences, but have mirror image attack rates on each resource. We conjecture that these results are quite general, given that the process of character displacement is quite general. To explore the generality of our predictions, we examined whether our conclusions still held under a foraging scenario that was designed to mimic the stickleback lake food web. To do this, we used empirical relationships between body size and the parameters that govern consumer-resource dynamics to parameterize an eco-evolutionary model. Specifically, we used consumer and resource biomasses

### 2018-05-10 LSC.sym.evo.4sp.R

@@ -1,0 +1,659 @@
+
+#### Description: 4 species model with consumer coevolution simulation,
+####              but this is only for one parameter set where resource
+####              parameters are equivalent and consumers are mirror
+####              images of each other.
+
+## source general parameters and functions ----
+source('code/00_general_parameters.R')
+source('code/00_numerical_fxns.R')
+source('Lawlor_Smith_Coarse.R')
+
+trait.thres <- 1e-10
+j.step <- 100 # for other simulations # had at 1000 for some others...
+
+## general parameters for simulations
+r <- 1
+K <- 4
+e <- 0.8
+#h <- 0.4
+aii <- 1.9 #1.93
+aij <- 1.2
+#a <- 2
+w <- 0.5
+m <- 1
+R <- 2
+C <- 1
+
+
+## 4 species simulation ----
+
+## create initial data
+sym4.init.df <- data.frame(r1 = r, r2 = r, K1 = K, K2 = K,
+                           e11 = e, e12 = e, e21 = e, e22 = e,
+                           #h11 = h, h12 = h, h21 = h, h22 = h,
+                           B11 = aii, B12 = aij, B21 = aij, B22 = aii,
+                           w11 = w, w22 = w, 
+                           m1 = m, m2 = m,
+                           R1 = R, R2 = R, C1 = C, C2 = C)
+
+set.temp <- as.matrix(sym4.init.df[1, ])
+
+set.steady <- safe.runsteady(set.temp[1,c("R1","R2","C1","C2")], func = ECD_model_LSC, parms = set.temp[1,c("r1","r2","K1","K2","e11","e12","e21","e22","B11","B12","B21","B22","w11","w22","m1","m2")])
+
+sym4.eq.jac <- jacobian.full(y = set.steady$y, func = ECD_model_LSC, parms = set.temp[1,c("r1","r2","K1","K2","e11","e12","e21","e22","B11","B12","B21","B22","w11","w22","m1","m2")])
+
+## update with new steady state and eigenvalues after the initial run
+sym4.init.df[ ,c("R1","R2","C1","C2")] <- set.steady$y
+sym4.init.df$max.Re.eigen <- max(Re(eigen(sym4.eq.jac)$values))
+sym4.init.df$max.Im.eigen <- max(Im(eigen(sym4.eq.jac)$values))
+
+
+## setup and run simulation
+df <- sym4.init.df
+sim.list.4 <- list()
+for(i in 1:dim(df)[1]){
+  
+  # choose parameter set to examine and calculate the maximum real eigenvalue
[…]

### 2018-05-10 Lawlor_Smith_Coarse.R

@@ -1,0 +1,304 @@
+######################################################
+## Description: Dynamical model and other useful functions for checking dynamics
+## Code author(s): Matt Barbour
+## Email: barbour@zoology.ubc.ca
+######################################################
+
+
+## Consumer-resource dynamics ----
+# general assumptions:
+# (1) logistic growth for the resource(s)
+# (2) multispecies, spatially-implicit type 2 functional response of consumer 
+# (3) density-independent mortality for consumer 
+# function is in the necessary format "function(Time,State,Pars)" for solving used the ode() function in the R package "deSolve"
+
+## 4 species model: 2 consumers, 2 resources
+ECD_model_LSC <- function(Time,State,Pars) {
+  # t,y,p is the necessary format for solving using the ode() function in R
+  with(as.list(c(State,Pars)), {
+    
+    a11 <- w11 * B11
+    a12 <- w12 * B12
+    a21 <- w21 * B21
+    a22 <- w22 * B22
+    
+    # consumer functional responses
+    C1R1fxn <- a11 * R1 # (W11 * a11 * R1)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
+    C1R2fxn <- a12 * R2 # (W12 * a12 * R2)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
+    C2R1fxn <- a21 * R1 # (W21 * a21 * R1)/(1 + W21 * a21 * h21 * R1 + W22 * a22 * h22 * R2)
+    C2R2fxn <- a22 * R2 # (W22 * a22 * R2)/(1 + W21 * a21 * h21 * R1 + W22 * a22 * h22 * R2)
+    
+    # dynamical equations
+    dR1.dt <- r1 * R1 * (1 - R1 / K1) - C1 * C1R1fxn - C2 * C2R1fxn
[…]

### 2018-05-10 attack_rate_trade_off_function.R

@@ -5,3 +5,16 @@
 a_ij_tradeoff <- function(A, a_ii, n) {
   A * (1 - (a_ii/A)^n)^(1/n)
 }
+a_ij_tradeoff(A=2, a_ii=1.5, n=1)
+
+# from Abrams 1986
+# ki*C1i^2 + (1-ki)*C2i^2 = ki*(1-ki)
+a_ij_Abrams <- function(trait, k, n) {
+  ((k * (1 - k) - k * trait^n)/(1 - k))^(1/n)
+}
+df <- data.frame(B11 = 1)
+a_ij_Abrams(trait = 1, k=0.5, n=1)
+
+a_ii_seq <- seq(0,2,0.01)
+
+plot(a_ij_Abrams(ki=0.7, a_ii=a_ii_seq, n=1) ~ a_ii_seq, type='l')

### 2018-05-10 code/00_general_parameters.R

@@ -9,16 +9,16 @@ library(tidyr) # for tidying data
 library(cowplot) # for better base, ggplot graphics
 
 ## Abundance threshold for feasible equilibriums
-abund.thres <- 1e-2 
+#abund.thres <- 1e-2 
 
 ## Trait change threshold for identifying ESS
-trait.thres <- 1e-4
+#trait.thres <- 1e-4
 
 ## Number of mutation iterations
-mut.reps <- 10000 
+#mut.reps <- 10000 
 
 ## Mutation size
-mut.size <- 0.01
+#mut.size <- 0.01
 
 ## All runsteady simulations go for 1000 time steps with a steady state tolerance of 1e-4. 
 runsteady.ps <- function(y, func = func, parms = parms, times = c(0,1000), stol = 1e-4){

### 2018-05-10 eco_evo_sim_function.R

@@ -1,0 +1,199 @@
+
+#### A general function for an eco-evo simulation using Adaptive Dynamics ####
+
+eco_evo_CR <- function(init_parameters,              # data frame of initial parameter values.
+                       init_states,                  # data frame of initial state variables
+                       eco_CR_model,                 # named function that describes the ecological Consumer-Resource Dynamics
+                       AD_CR_models,                 # vector of named functions describing the Adaptive Dynamics of the Consumer-Resource Model
+                       species_traits,               # vector that matches each trait to the mutant dynamics in AD_CR_models 
+                       eco_param_names,              # vector of evolving parameter names in the ecological model.
+                       eco_tradeoff_names,           # vector of trade-off parameter names in the ecological model. Order must correspond to 'eco_param_names'
+                       evo_param_names,              # vector of mutant parameter names in the Adaptive Dynamics model. Order must correspond to 'eco_param_names'
+                       evo_tradeoff_names,           # vector of mutant trade-off parameter names in the Adaptive Dynamics model. Order must correspon to 'eco_param_names'
+                       tradeoff_functions,           # names function of the trade-off with corresponding parameters
+                       extra_tradeoff_params=c(NOT_USED=1), # vector of extra tradeoff parameters. Default is a nonsense parameter to just make sure the function runs
+                       mut.size=0.001,               # absolute change in evolving parameter. Directionality is chosen at random.
+                       mut.suc.thres=1e-4,           # threshold of mutant success rate for identifying ESS.
+                       j.step=500,                   # number of evolutionary time steps over which to assess evolutionary change.
+                       abund.thres=1e-4,             # abundance threshold for feasible equilibriums
+                       mut.reps=10000) {             # number of mutation iterations to try for simulation. Note that this is the upper limit of the simulation length. Usually the traits will hit their ESS or a constraint on the parameter space before this limit is reached.   
+                            
+  #### SETUP ####
+  
+  ## Get steady-state resource and consumer abundances from initial parameter values and state variables 
[…]

### 2018-05-10 eco_evo_simulations.R

@@ -1,0 +1,592 @@
+
+#### PRELIMINARIES ####
+
+## Source in required functions
+source('code/00_general_parameters.R')
+source('code/00_numerical_fxns.R')
+source('eco_evo_sim_function.R')
+source('Consumer_Resource_Functions.R')
+
+#### GENERAL PARAMETERS AND STATE VARIABLES #### 
+
+## Intrinsic growth rates
+r <- 1
+r1 <- r 
+r2 <- r 
+
+## Carrying capacity
+K <- 4
+K1 <- K
+K2 <- K
+
+## Conversion efficiency
+e <- 0.8
+e11 <- e 
+e12 <- e 
+e21 <- e 
+e22 <- e 
+
+## Mortality rate
+m <- 1
+m1 <- m
+m2 <- m
+
+## General, non-evolving parameter set for 2C,2R model
+general_parameters <- data.frame(r1 = r1, r2 = r2, K1 = K1, K2 = K2, e11 = e11, e12 = e12, e21 = e21, e22 = e22, m1 = m1, m2 = m2)
+
+## Attack rate
+A <- 2 
+aii <- A*0.6
+
+## Habitat preference (for LawlorSmith and McCann models)
+w <- 0.6
+wii <- w
+wij <- (1 - w)
+
+## Handling time (only for McCann model)
+h <- 0.4
+h11 <- h
+h12 <- h
+h21 <- h
+h22 <- h
+
+## State variables
+R <- 2
+R1 <- R
+R2 <- R
+C <- 1
+C1 <- C
+C2 <- C
+
+states_2C_2R <- data.frame(R1 = R1, R2 = R2, C1 = C1, C2 = C2)
+states_1C_2R <- data.frame(R1 = R1, R2 = R2, C1 = C1)
+
+## Trade-offs. Borrowed from Sargent and Otto, 2006 Am Nat
+a12_tradeoff <- expression(A*(1-(a11m/A)^n)^(1/n)) 
+a21_tradeoff <- expression(A*(1-(a22m/A)^n)^(1/n)) 
+
+n_convex <- 1.15
+n_linear <- 1
+n_concave <- 0.85
+
+convex_df <- data.frame(n = n_convex, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_convex)^(1/n_convex))
+linear_df <- data.frame(n = n_linear, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_linear)^(1/n_linear))
[…]

### 2018-05-10 ECD_Supp_Mat.Rmd
updates to supp mat

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(
+	message = TRUE,
+	warning = TRUE,
+	include = FALSE
+)
+source('eco_evo_simulations.R')
 ```
@@
 Although such a trade-off results in divergent character displacement (Abrams 1986), there is no net change in resource abundances when consumer's diverge.
-Although this equilibrium equation for resource abundances was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when there is only one consumer as well.
-This is likely because a consumer, in the absence of competition, evolves to be a generalist that has equal attack rates on each resources, resulting in equivalent resource abundances.
+
+## Trade-offs
 
@@
 ```{r Trade-off Function, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
+bind_rows(convex_df, linear_df, concave_df) %>%
+  ggplot(., aes(x=aii, y=aij, color = factor(n))) + geom_line() + ggtitle("Attack rate trade-offs")
 ```
@@
 When the trade-off is concave, however, we see that character divergence suppresses resource densities.
+Although this equilibrium equation for resource abundances was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when there is only one consumer as well.
+This is likely because a consumer, in the absence of competition, evolves to be a generalist that has equal attack rates on each resources, resulting in equivalent resource abundances.
+
+```{r Effect of trade-off on ECD-Resource relationship, echo=FALSE, message=FALSE, warning=FALSE}
+ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+
+cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+```
+
+## Foraging Scenario
 
@@
 The only chracter displacement model that we are aware of that modeled resources in distinct microhabitat was one examined by Lawlor and Smith (1976).
-They found that consumers still underwent character divergence, regardless of the foraging scenario.
+This model takes the same form as the MacArthur competition model, except now the consumer's functional response takes the form:
+
+$$F_{ij}(R_j)=w_{ij}a_{ij}R_j$$
+
+where $w_{ij}$ represents the proportion of time $C_i$ spends foraging in the habitat where $R_j$ is found. Note that since $w_ij$ is a proportion that $w_{ii}=1-w_{ij}$.
+
+Lawlor and Smith found that consumers still underwent character divergence, regardless of the foraging scenario.
 Interestingly, we find that the foraging scenario qualitatively affects the relationship between character divergence and resource abundances.
-Specifically, if consumers are competing for resources that occur in distinct microhabitats, then consumer divergence always results in resource suppression.
-This occurs because now the consumers effective attack rate is influenced by its habitat preference.
-Note also that this resource suppression occurs regardless of the shape of the trade-off, although the shape of the trade-off can amplify the effect of character divergence on resource supression.
-The type of trade-off does affect the amount of character displacement though, and actually, we see again the a concave trade-off results in less displacement and less of an effect on resource abundances; whereas a convex trade-off results in the largest displacement and the largest effect on resource abundances.
+Note that now, equilibrium resource abundances are determined by the following equation:
+
+$$\hat{R_i}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e}$$
+
+This means that if if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g.
+$w_{ii}>0.5$ and $a_{ii}>a_{ij}$), then resource abundances will always decrease, regardless of the trade-off.
+Thus, divergent character displacement always results in resource suppression if consumers are competing for resources that occur in distinct microhabitats.
+Note that the shape of the trade-off can dampen or amplify the effect of character divergence on resource supression.
+This is because the shape of the trade-off influences the magnitude of character displacement.
+For example, a convex relationship results in less displacement and less of an effect on resource abundances; whereas a concave trade-off results in the largest displacement and the largest effect on resource abundances.
+
+```{r}
+ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+
+cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor),
+             shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+```
+
 This effect of spatial context on the relationship between character displacement and resource abundances appears to be quite general.
-For example, in the appendix, we analyze a model that was designed to capture the spatial coupling of consumers in many food-web context (McCann 2005).
-Here, we again find that divergence results in a negative effect on resource abundances.
+For example, even if we consider a more realistic consumer functional response (derived by McCann et al.
+2005):
 
-We would like to note that the scenario where the relationship between character divergence and resource suppression is the strongest is when the trade-off in attack rates is concave and consumers are competiting for resources that occupy distinct habitat types.
-Given that prior work in threespine stickleback has shown a concave trade-off in feeding rates (Schluter 1993; Arnegard et al.
+$$F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}hW_{ii}R_i+a_{ij}hW_{ij}R_j}$$
+
+where consumer feeding rates on $R_i$ are influenced by both resource densities; saturate as resource densities increase (due to handling time $h$); and consumer habitat preferences are modified by relative resource densities, such that:
+
+$$W_{ii}=\frac{w_{ii}R_i}{w_{ii}R_i+(1-w_{ii})R_j}$$ 
+we still observe the same qualitative relationship between character divergence and resource suppression.
+This is because resource competition always leads to character divergence (see proof in Appendix) and equilibrium resource abundances are governed by similar dynamics:
+
+$$\hat{R_i}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e-hm}$$
+
+## Predictions for threespine stickleback
+
+The scenario where the relationship between character divergence and resource suppression is the strongest is when the trade-off in attack rates is concave and consumers are competiting for resources that occupy distinct habitat types.
+Prior work in threespine stickleback suggests that the evolution of stickleback attack rates are constrained by a concave trade-off (Schluter 1993; Arnegard et al.
 2014) and that stickleback ecotypes are competing for resources that occur in distinct microhabitats (benthic vs.
-limnetic), we predict that character divergence in sticklebacks would supress resource densities.
+limnetic).
+Thus, we predict that character divergence in sticklebacks would supress resource densities.
+
+We note, however, that our analytical insights to the effects of consumer divergence on resource dynamics has considered a simplistic scenario (resources are equivalent and consumers have equivalent conversion efficiences, but have mirror image attack rates on each resource).
+To explore the generality of our predictions, we examined whether our conclusions still held under a foraging scenario that was designed to mimic the stickleback lake food web.
+To do this, we used empirical relationships between body size and the parameters that govern consumer-resource dynamics to parameterize an eco-evolutionary model.
+Specifically, we used consumer and resource biomasses
+
+```{r Allometric Scaling of Stickleback Food Web}
+
+mass_scaling <- function(W, intercept, exponent) {
+  log_rate <- intercept + exponent * log(W)
+  return(exp(log_rate))
+}
+
+## BODY SIZE (g) ----
+
+g_zoop <- 0.004/1000            # covert 0.004 mg to g
+g_benthic_invert <- 0.25/1000   # convert 0.25 mg to g
+
+g_limnetic <- 0.81              # from Schluter 1993, Ecology Table 1
+g_benthic <- 1.48               # from Schluter 1993, Ecology Table 1
+g_solitary <- 1.12              # from Schluter 1993, Ecology Table 1
+
+## INTRINSIC GROWTH RATE (r) ----
+# r = birth rate per day
+# proportional to M^(-1/4) from Rip and McCann 2011 (they cite Brown et al. 2004)
+# alternatively use scaling exponent = -0.27377 for heterotherm metazoa from Fenchel 1974, Oecologia
+r_exponent <- -0.25             # theoretical scaling.
+r_scale <- 0.3                  # estimate of r from Daphnia pulex in Table 1 of Fenchel 1974. Original study is Frank et al. 1957       
+r_intercept <- log(r_scale) - r_exponent * log(g_zoop) # scale based on zooplankton size
+
+r_benthos_estimate <- mass_scaling(W = g_benthic_invert, intercept = r_intercept, exponent = r_exponent)
+
+#### CARRYING CAPACITY (K) ----
+# proportional to M^(-3/4) from Rip and McCann 2011 (they cite Enquist et al. 1998 and Brown et al. 2004)
+# alternatively use empirical exponent = -0.77 from Brown et al. 2004
+K_exponent <- -0.75 
+K_scale <- 5          # admittedly, this is a bit of an arbitrary number. It does correspond to zooplankton biomass density (g/500 L) from fishless ponds
+K_intercept <- log(K_scale) - K_exponent * log(g_zoop) # scale based on zooplankton size
+
+K_benthos_estimate <- mass_scaling(W = g_benthic_invert, intercept = K_intercept, exponent = K_exponent)
+
+#### CONVERSION EFFICIENCY (e) ----
+# from Mittelbach 1981 (see refs to Ware 1975 and Elliot 1976) 
+# independent of body size (Rip and McCann 2011, Yodzis & Innes 1992)
+e <- 0.7
+
+#### HABITAT PREFERENCE (w) ----
+# Based on % littoral Carbon from Matthews et al. 2010
+limnetics_littoral <- c(0.4, 0.1)
+benthics_littoral <- c(0.7, 0.6)
+solitary_littoral <- c(0.5, 0.6, 1.0, 0.1)
+
+w_limnetics_in_open_water <- 1 - mean(limnetics_littoral) # 0.65 is approximated from Fig. 2 in Schluter 1993
+w_benthics_in_benthos <- mean(benthics_littoral) # 0.9 is approximated from Fig. 2 in Schluter 1993
+
+# as 1 - littoral to correspond to parameters in ECD_model.C1.3sp 
+w_solitary <- 1 - mean(solitary_littoral)                 # 0.5
+
+#### MORTALITY RATE (m) ----
+# m = individual deaths per unit time
+# proportional to M^(-1/4) from Brown et al. 2004
+# empirical estimate of exponent = -0.24 from Fig. 4 in Brown et al. 2004
+
+# estimate of annual threespine stickleback mortality from fishbase
+library(rfishbase) 
+threespine_stickleback_M <- filter(popgrowth(species_list = "Gasterosteus aculeatus"), M != "NA")$M
+
+m_exponent <- -0.25                   # theoretical exponent
+m_scale <- threespine_stickleback_M        
+#m_intercept <- log(m_scale) - m_exponent * log(g_solitary) # use solitary stickleback mass for scaling
+
+m_limnetic_estimate <- m_scale #mass_scaling(W = g_limnetic, intercept = m_intercept, exponent = m_exponent); m_limnetic_estimate
+m_benthic_estimate <- m_scale #mass_scaling(W = g_benthic, intercept =  m_intercept, exponent = m_exponent); m_benthic_estimate
+
+#### HANDLING TIME (h) ---
+
+# from Mittelbach 1981, Table 3: Minimum handling time for Daphnia vs. Chironomus. Units = seconds per individual
+min_h_Chironomus <- 9.63  # this can be much higher because stickleback than sunfish
+min_h_Daphnia <- 1.02
+benthos_relative_handling_time <- min_h_Chironomus/min_h_Daphnia
+
+h_scale <- 0.01 # assuming it's independent of consumer body size. This assumption is likely okay, given that prey varied 2-orders of magnitude while stickleback vary much less than 1-order.
+
+#### ATTACK RATE (a) ----
+# estimates from Fig. 4, Schluter 1993
+# although these are estimates of prey volume consumed per strike, 
+# Schluter noted that these relative estimates were virtually identical to intake rate.
+# The intake rate data taken by Schluter  is more in line with per capita attack rate; however,
+# not all of these data were reported. This is why we are using capture success as a proxy.
+# We use these estimates instead of mass-based estimates of attack rate (e.g. Rip and McCann 2011 cite Peters 1983 and Brown et al. 2004),
+# since these traits depend other evolved traits that are not related to body size (e.g. gill rakers, gape size and body shape)
+
+limnetic_capture_success_open_water <- 0.011
+limnetic_capture_success_benthos <- 0.025        # note that limnetics have an overall higher capture succes in benthos vs.
+open water, this doesn't nullify the existence of a trade-off between the two habitats
+benthic_capture_success_open_water <- 0.004 
+benthic_capture_success_benthos <- 0.16
+
+solitary_capture_success_open_water <- 0.009
+solitary_capture_success_benthos <- 0.04
+
+a_scale <- 300                                   # scaled until I was able to get coexistence for species pairs
+
+#### CONSUMER-RESOURCE DYNAMICS ----
+
+## Attack rate
+A <- 5 #2 
+aii <- A*0.6
+aij <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_linear)) # n_concave
+
+m_solitary_McCann <- expression(e11*(((w11m * R1)/(w11m * R1 + w12m * R2)) * a11m * R1)/(1 + ((w11m * R1)/(w11m * R1 + w12m * R2)) * a11m * h11 * R1 + ((w12m * R2)/(w11m * R1 + w12m * R2)) * a12m * h12 * R2) + e12*(((w12m * R2)/(w11m * R1 + w12m * R2)) * a12m * R2)/(1 + ((w11m * R1)/(w11m * R1 + w12 * R2)) * a11m * h11 * R1 + ((w12m * R2)/(w11m * R1 + w12m * R2)) * a12m * h12 * R2) - m1)
+
+w12_tradeoff <- expression(1-w11m)
+
+## Solitary lakes
+solitary.init.states <- c(R1=0.1, R2=0.1, C1=0.01) # note that equilibrium dynamics appear to be independent of initial state variables
+solitary.parms <- data.frame(r1 = r_scale*g_zoop*100000, 
+                             r2 = r_benthos_estimate*g_benthic_invert*100000, 
+                             K1 = K_scale*g_zoop*100000, 
+                             K2 = K_benthos_estimate*g_benthic_invert*100000,
+                             e11 = e, 
+                             e12 = e, 
+                             #e21 = e, 
+                             #e22 = e,
+                             h11 = h_scale, 
+                             h12 = h_scale*benthos_relative_handling_time, 
+                             #h21 = h_scale, 
+                             #h22 = h_scale*benthos_relative_handling_time,
+                             a11 = aii, #a_scale*solitary_capture_success_open_water, 
+                             a12 = aij, #a_scale*solitary_capture_success_benthos, 
+                             #a21 = a_scale*benthic_capture_success_open_water, 
+                             #a22 = a_scale*benthic_capture_success_benthos,
+                             w11 = 0.6, #w_solitary,
+                             w12 = 0.4,
+                             #w22 = w_benthics_in_benthos, 
+                             m1 = m_scale) 
+                             #m2 = m_benthic_estimate)
+
+solitary_McCann_concave <- eco_evo_CR(init_parameters = solitary.parms, 
+                                          init_states = states_1C_2R, #solitary.init.states, 
+                                          eco_CR_model = McCann_1C_2R, 
+                                          AD_CR_models = c(m_solitary_McCann),
+                                          species_traits = c(1,1), 
+                                          eco_param_names = c("a11","w11"), 
+                                          eco_tradeoff_names = c("a12","w12"),  
+                                          evo_param_names = c("a11m","w11m"),  
+                                          evo_tradeoff_names = c("a12m","w12m"), 
+                                          tradeoff_functions = c(a12_tradeoff, w12_tradeoff),
+                                          extra_tradeoff_params = c(A=A, n=n_concave),
+                                      mut.size=0.01)
+solitary_McCann_concave[c(1,nrow(solitary_McCann_concave)), ]
+## Species-pair lakes
+species_pair.init.states <- c(R1=1, R2=1, C1=0.25, C2=0.25) # note that equilibrium dynamics appear to be independent of initial state variables
+species_pair.parms <- data.frame(r1 = r_scale, 
+                                r2 = r_benthos_estimate, 
+                                K1 = K_scale, 
+                                K2 = K_benthos_estimate,
+                                e11 = e, 
+                                e12 = e, 
+                                e21 = e, 
+                                e22 = e,
+                                h11 = h_scale, 
+                                h12 = h_scale*benthos_relative_handling_time, 
+                                h21 = h_scale, 
+                                h22 = h_scale*benthos_relative_handling_time,
+                                a11 = a_scale*limnetic_capture_success_open_water, 
+                                a12 = a_scale*limnetic_capture_success_benthos, 
+                                a21 = a_scale*benthic_capture_success_open_water, 
+                                a22 = a_scale*benthic_capture_success_benthos,
+                                w11 = w_limnetics_in_open_water, 
+                                w22 = w_benthics_in_benthos, 
+                                m1 = m_limnetic_estimate, 
+                                m2 = m_benthic_estimate)
+
+sim_2C_2R_McCann_concave <- eco_evo_CR(init_parameters = species_pair.parms, 
+                                          init_states = species_pair.init.states, 
+                                          eco_CR_model = McCann_2C_2R, 
+                                          AD_CR_models = c(mC1_McCann, mC2_McCann),
+                                          species_traits = c(1,2), 
+                                          eco_param_names = c("a11","a22"), 
+                                          eco_tradeoff_names = c("a12","a21"),  
+                                          evo_param_names = c("a11m","a22m"),  
+                                          evo_tradeoff_names = c("a12m","a21m"), 
+                                          tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                          extra_tradeoff_params = c(A=A, n=n_concave))
+
+```
+
+## Effect of character divergence on food-web stability
 
@@
 four-species system, which is key for understanding how ecological character displacement affects food-web dynamics.
-
-## Testing the generality of our predictions
-
-So far, our theoretical predictions of the effects of consumer divergence on resource dynamics has considered a simplistic scenario, where resources are equivalent and consumers have equivalent conversion efficiences, but have mirror image attack rates on each resource.
-We conjecture that these results are quite general, given that the process of character displacement is quite general.
-To explore the generality of our predictions, we examined whether our conclusions still held under a foraging scenario that was designed to mimic the stickleback lake food web.
-To do this, we used empirical relationships between body size and the parameters that govern consumer-resource dynamics to parameterize an eco-evolutionary model.
-Specifically, we used consumer and resource biomasses

### 2018-05-10 eco_evo_simulations.R

@@ -6,6 +6,7 @@ source('code/00_general_parameters.R')
 source('code/00_numerical_fxns.R')
 source('eco_evo_sim_function.R')
 source('Consumer_Resource_Functions.R')
+source('CR_dynamics.R')
 
 #### GENERAL PARAMETERS AND STATE VARIABLES #### 
 

### 2018-05-10 ECD_Supp_Mat.Rmd
draft of effect of character divergence on resource abundances.

--- 2018-05-10 ECD_Supp_Mat.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(
+	message = TRUE,
+	warning = TRUE,
+	include = FALSE
+)
+source('eco_evo_simulations.R')
 ```
 
-## How does character displacement affect resource dynamics?
+```{r}
+# Numerous mathematical models have been developed to examine how resource competition influences the evolution of consumer foraging traits (reviewed in Taper and Case 1985).
+The general result emerging from these models is that competition for shared (substitutable) resources results in the evolution of consumers that specialize on different resources relative to the generalists that evolve in the absence of competition.
+The only scenarios in which character divergence will not occur is when density dependence in intra- or interspecific interactions due to other factors are sufficiently strong (Abrams 1986), although there remains little empirical evidence of this to date (Schluter 2000).
+Despite the many models that have been developed to examine the ecological scenarios resulting in character displacement, scant attention has been paid to the ecological consequences of such character divergence on resource abundances and food-web stability (but see Lawlor and Smith 1976 for an assessment of stability).
 
-Despite the numerous models that have been developed to examine the ecological conditions resulting in character displacement, none make predictions about how character divergence influences resource abundances or stability (except perhaps Lawlor and Smith 1976).
-This is a bit surprising, given that explicit modeling of resource dynamics has been shown to give key insight to the process leading to character divergence.
+# In terms of resources, perhaps the intuitive expectation is that character divergence would suppress resource abundances.
+The rationale being that character divergence will result in specialized consumers that are more effective at attacking a particular resource.
+Below, we demonstrate that this is not necessarily true.
+In fact, the effect of character displacement on resource abundances depends on both the foraging scenario and tradeoff in consumer feeding rates.
+```
 
-Intuitively, you might expect character divergence to decrease resource abundances.
-The rationale being that consumers are evolving to be specialists that are more effective at attacking a particular resource.
-This however is not necessarily true.
-In fact, the effect of character displacement on resource dynamics depends on both the foraging scenario and trade-off in attack rates.
-For example, let's start with the simplest model of competition that explicitly models resource dynamics is MacArthur's model of competition (MacArthur 1972):
+## How does character displacement affect resource abundances?
 
-$$\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-a_{11}R_1C_1-a_{21}R_1C_2$$
-$$\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-a_{12}R_2C_1-a_{22}R_2C_2$$
-$$\frac{dC_1}{dt}=e_{11}a_{11}R_1C_1+e_{12}a_{12}R_2C_1-m_1C_1$$
-$$\frac{dC_2}{dt}=e_{21}a_{21}R_1C_2+e_{22}a_{22}R_2C_2-m_2C_2$$
+To examine this, we used a continuous-time model of consumer-resource dynamics for two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
 
-Now, let's examine the simplest scenario where resources are equivalent ($r=r_i$ and $K=K_i$) and consumers have equivalent conversion efficiencies ($e=e_{ii}=e_{ij}$) except that their attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
-It can be shown that, at the equilibrium where both consumers and resources are present, resource abundances are equivalent ($R_i=R_1=R_2$) and are determined by the following equation:
+$$\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2$$
+$$\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2$$
+$$\frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1$$
+$$\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2$$
+where $r_i$ represents intrinsic growth rates, $K_i$ represents carrying capacity, $e_{ij}$ represents the conversion efficiency of resources into new consumers, and $m_i$ represents the mortality rate of consumers.
+$F_{ij}(R_j)$ represents the consumer's feeding rate (or functional response) on a specific resource (in this case, $C_i$ on $R_j$).
+This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g.
+zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g.
+seed size; Taper and Case 1985).
+In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate, such that $F_{ij}(R_j)=a_{ij}R_j$, then our model encapsulates MacArthur's model of resource competition (MacArthur 1972).
+The evolution of consumer attack rates in this model, and several extensions, have been analyzed in detail (Lawlor and Smith 1976; Abrams 1986).
+The general result of these analyses is that consumer divergence is an inevitable outcome of competition.
+Here, we examine the effects of this inevitable divergence on resource dynamics and food-web stability.
+We show that the effects of character displacement are not intuitive, and depend critically on the foraging scenario and tradeoff in consumer attack rates.
 
-$$\hat{R_i}=\frac{m}{e(a_{ii}+a_{aij})}$$
+## Consumers forage for resources simultaneously
+
+The first foragine scenario we explore is already assumed in MacArthur's model of resource competition, namely that consumers can forage for both resources simultaneously.
+This also implies that both resources occur within the same microhabitat.
+To gain analytical insight to this scenario, we assume that resources are equivalent ($r=r_i$ and $K=K_i$) as well as consumers ($e=e_{ii}=e_{ij}$; $m=m_i$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
+While this scenario is simple, it isolates the effects of character displacement, rather than any inherent differences in consumers or resources.
+In the appendix, we show that when both consumers and resources are present that the abundance of resources at equilibrium are equivalent ($R_i=R_1=R_2$) and are determined by the following equation:
+
+$$\hat{R_i}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
 Note that a key determinant of resource abundance is the consumer's total attack rate, $a_{ii}+a_{ij}$.
-An increased in total attack rate results in a decrease in resource abundances.
-Now let's assume that the evolution of consumer attack rates are constrained by a linear trade-off, i.e.
-a unit increase in $a_{ii}$ results in a unit decrease in $a_{ij}$.
-Although such a trade-off results in divergent character displacement (Abrams 1986), there is no net change in resource abundances when consumer's diverge.
-Although this equilibrium equation for resource abundances was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when there is only one consumer as well.
-This is likely because a consumer, in the absence of competition, evolves to be a generalist that has equal attack rates on each resources, resulting in equivalent resource abundances.
+Since the total attack rate is in the denominator, this implies that resource abundances will decrease if evolution favors an increase in total attack rate.
+Consumers, however, cannot maximize their attack rate
 
-Now let's consider a different type of trade-off in consumer attack rates.
-We know from past work that consumers still undergo character divergence regardless of the shape of the trade-off, even though the trade-off does influence the magnitude of displacement (Lawlor and Smith 1976).
-We specified the following trade-off function (Sargent and Otto, 2006):
+To understand how a consumer's total attack rate would change with character divergence, we explored the effect of different tradeoffs in consumer attack rates. We used the following general function:
 
 $$\Big(\frac{a_{ii}}{A}\Big)^n+\Big(\frac{a_{ij}}{A}\Big)^n=1$$
-where $A$ is the total investment in attack rates and $n$ describes the shape of the trade-off.
-This trade-off has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$).
-When $n=1$, the trade-off function is linear, and all combinations of $a_ii$ and $a_ij$ have the same total attack rate.
 
-```{r Trade-off Function, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff.
+This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$; Sargent and Otto 2006).
+When $n=1$, the tradeoff function is linear, and all combinations of $a_ii$ and $a_ij$ have the same total attack rate.
 
+```{r Tradeoff function, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+bind_rows(convex_df, linear_df, concave_df) %>%
+  ggplot(., aes(x=aii, y=aij, color = factor(n))) + geom_line() + ggtitle("Tradeoffs in attack rates")
 ```
 
-In contrast, the shape of the trade-off qualitatively affects the relationship between divergence and resource abundance.
-As mentioned previously, a linear trade-off results in no change in resource densities.
-If the trade-off is convex, then resource densities can actually increase under character displacement!
+We know from past work that consumers still undergo character divergence regardless of the shape of the tradeoff (Lawlor and Smith 1976; Abrams 1986).
+However, the shape of the tradeoff qualitatively affects the relationship between character divergence and resource abundance.
+For example, if consumer's are constrained by a linear tradeoff ($n=1$), such that a unit increase in $a_{ii}$ results in a unit decrease in $a_{ij}$, then character divergence has no effect on resource abundances.
+If the tradeoff is concave down ($n>1$), then resource abundances can actually increase under character displacement!
 This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{aij}$) and decreases as consumers diverge.
-Note though that detecting this effect in nature may be difficult, since a convex trade-off results in relatively small displacement relative to other trade-off forms (Lawlor and Smith 1976).
-When the trade-off is concave, however, we see that character divergence suppresses resource densities.
+Note though that detecting this effect in nature will likely be difficult, since a concave down tradeoff results in relatively small displacement relative to other tradeoff forms (Lawlor and Smith 1976).
+When the tradeoff is concave up ($n<1$), however, we see that character divergence suppresses resource densities.
+Although this equilibrium equation for resource abundances was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when there is only one consumer as well (white circle).
+This is because a consumer, in the absence of competition, evolves to be a generalist that has equal attack rates on each resource, resulting in equivalent resource abundances.
 
-Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (i.e.
-resources occur in the same microhabitat).
-This assumption may be valid for some foraging scenarios, e.g.
-Darwin's finches foraging for seeds.
-However, in many situations, consumers forage for resources that occur in distinct microhabitats (McCann et al.
+```{r Same resource habitat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+
+Mac_resources <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+
+plot_grid(Mac_attack, Mac_resources, ncol=2)
+```
+
+## Consumers cannot forage for resources simultaneously
+
+Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (Taper and Case 1985; Abrams 1986).
+This assumption may be valid for some foraging scenarios, such as Darwin's finches foraging for seeds.
+However, in many situations, consumers forage for resources that occur in distinct habitats, and thus cannot forage for both simultaneosly (McCann et al.
 2005).
-The only chracter displacement model that we are aware of that modeled resources in distinct microhabitat was one examined by Lawlor and Smith (1976).
-They found that consumers still underwent character divergence, regardless of the foraging scenario.
+The only chracter displacement model that we are aware of that modeled resources in distinct habitats was one examined by Lawlor and Smith (1976).
+This model takes the same form as the previous model, except now the consumer's functional response takes the form:
+
+$$F_{ij}(R_j)=w_{ij}a_{ij}R_j$$
+
+where $w_{ij}$ represents the proportion of time a consumer ($C_i$) spends foraging in a habitat where a particular resource ($R_j$) is found.
+Note that since $w_ij$ is a proportion that $w_{ii}=1-w_{ij}$.
+
+Lawlor and Smith (1976) analyzed character divergence in models where consumers could and could not forage for resources simultaneously, and they found that consumers still underwent character divergence regardless of the foraging scenario.
 Interestingly, we find that the foraging scenario qualitatively affects the relationship between character divergence and resource abundances.
-Specifically, if consumers are competing for resources that occur in distinct microhabitats, then consumer divergence always results in resource suppression.
-This occurs because now the consumers effective attack rate is influenced by its habitat preference.
-Note also that this resource suppression occurs regardless of the shape of the trade-off, although the shape of the trade-off can amplify the effect of character divergence on resource supression.
-The type of trade-off does affect the amount of character displacement though, and actually, we see again the a concave trade-off results in less displacement and less of an effect on resource abundances; whereas a convex trade-off results in the largest displacement and the largest effect on resource abundances.
-This effect of spatial context on the relationship between character displacement and resource abundances appears to be quite general.
-For example, in the appendix, we analyze a model that was designed to capture the spatial coupling of consumers in many food-web context (McCann 2005).
-Here, we again find that divergence results in a negative effect on resource abundances.
+To demonstrate this, note that now resource abundances at equilibrium are determined by the following equation (see derivation in Appendix):
 
-We would like to note that the scenario where the relationship between character divergence and resource suppression is the strongest is when the trade-off in attack rates is concave and consumers are competiting for resources that occupy distinct habitat types.
-Given that prior work in threespine stickleback has shown a concave trade-off in feeding rates (Schluter 1993; Arnegard et al.
-2014) and that stickleback ecotypes are competing for resources that occur in distinct microhabitats (benthic vs.
-limnetic), we predict that character divergence in sticklebacks would supress resource densities.
+$$\hat{R_i}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e}$$
+
+This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g.
+$w_{ii}>0.5$ and $a_{ii}>a_{ij}$), then resource abundances will always decrease, regardless of the tradeoff.
+Thus, character divergence always results in resource suppression if consumers are competing for resources that occur in distinct habitats.
+Note that the shape of the tradeoff can dampen or amplify the effect of character divergence on resource supression.
+This is because the shape of the tradeoff influences the magnitude of character displacement.
+For example, a concave-down tradeoff results in less displacement and less of an effect on resource abundances; whereas a concave-up tradeoff results in the largest displacement and the largest effect on resource abundances.
+
+```{r Distinct resource habitats, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+
+LS_resources <- cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor),
+             shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+
+plot_grid(LS_attack, LS_resources, ncol=2)
+```
+
+This effect of the foraging scenario on the relationship between character divergence and resource abundances appears to be quite general.
+For example, even if we consider a more realistic consumer functional response (derived by McCann et al.
+2005):
+
+$$F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}hW_{ii}R_i+a_{ij}hW_{ij}R_j}$$
+
+where consumer feeding rates on $R_i$ are influenced by both resource densities; saturate as resource densities increase (due to handling time $h$); and consumer habitat preferences are modified by relative resource densities, such that:
+
+$$W_{ii}=\frac{w_{ii}R_i}{w_{ii}R_i+(1-w_{ii})R_j}$$
+
+we still observe the same qualitative relationship between character divergence and resource suppression.
+This is because resource competition always leads to character divergence (see proof in Appendix) and equilibrium resource abundances are governed by similar dynamics:
+
+$$\hat{R_i}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e-hm}$$
+
+```{r}
+McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+
+McCann_resources <- cbind.data.frame(tradeoff_df, McCann_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  scale_fill_manual(values=c("white","black"))
+
+plot_grid(McCann_attack, McCann_resources, ncol=2)
+```
+
+## Predictions for threespine stickleback
+
+The scenario where the relationship between character divergence and resource suppression is the strongest is when the tradeoff in attack rates is concave-up and consumers are competing for resources that occur in distinct habitats.
+Prior work in threespine stickleback suggests that the evolution of stickleback attack rates are constrained by a concave-up tradeoff (Schluter 1993; Arnegard et al.
+2014) and that stickleback ecotypes are competing for resources that occur in distinct habitats (benthic vs.
+limnetic).
+Thus, we predict that character divergence in sticklebacks would supress resource densities.
+
+## Effect of character divergence on food-web stability
 

### 2018-05-10 maybe_useful_morgue.R

@@ -1,0 +1,192 @@
+We note, however, that our analytical insights to the effects of consumer divergence on resource dynamics has considered a simplistic scenario (resources are equivalent and consumers have equivalent conversion efficiences, but have mirror image attack rates on each resource). To explore the generality of our predictions, we examined whether our conclusions still held under a foraging scenario that was designed to mimic the stickleback lake food web. To do this, we used empirical relationships between body size and the parameters that govern consumer-resource dynamics to parameterize an eco-evolutionary model. Specifically, we used consumer and resource biomasses 
+
+```{r Allometric Scaling of Stickleback Food Web}
+
+mass_scaling <- function(W, intercept, exponent) {
+  log_rate <- intercept + exponent * log(W)
+  return(exp(log_rate))
+}
+
+
+## BODY SIZE (g) ----
+
+g_zoop <- 0.004/1000            # covert 0.004 mg to g
+g_benthic_invert <- 0.25/1000   # convert 0.25 mg to g
+
+g_limnetic <- 0.81              # from Schluter 1993, Ecology Table 1
+g_benthic <- 1.48               # from Schluter 1993, Ecology Table 1
+g_solitary <- 1.12              # from Schluter 1993, Ecology Table 1
+
+
+## INTRINSIC GROWTH RATE (r) ----
+# r = birth rate per day
+# proportional to M^(-1/4) from Rip and McCann 2011 (they cite Brown et al. 2004)
+# alternatively use scaling exponent = -0.27377 for heterotherm metazoa from Fenchel 1974, Oecologia
+r_exponent <- -0.25             # theoretical scaling.
+r_scale <- 0.3                  # estimate of r from Daphnia pulex in Table 1 of Fenchel 1974. Original study is Frank et al. 1957       
+r_intercept <- log(r_scale) - r_exponent * log(g_zoop) # scale based on zooplankton size
+
+r_benthos_estimate <- mass_scaling(W = g_benthic_invert, intercept = r_intercept, exponent = r_exponent) 
+
+
+#### CARRYING CAPACITY (K) ----
+# proportional to M^(-3/4) from Rip and McCann 2011 (they cite Enquist et al. 1998 and Brown et al. 2004)
+# alternatively use empirical exponent = -0.77 from Brown et al. 2004
+K_exponent <- -0.75 
+K_scale <- 5          # admittedly, this is a bit of an arbitrary number. It does correspond to zooplankton biomass density (g/500 L) from fishless ponds
[…]

### 2018-05-11 ECD_Supp_Mat.Rmd
solid updates to supp mat

@@
 ---
-title: "ECD_Supp_Mat"
+title: "How does character divergence affect resource dynamics?"
 output: html_document
@@
 knitr::opts_chunk$set(
-	message = TRUE,
-	warning = TRUE,
+	message = FALSE,
+	warning = FALSE,
 	include = FALSE
@@
 
-```{r}
-# Numerous mathematical models have been developed to examine how resource competition influences the evolution of consumer foraging traits (reviewed in Taper and Case 1985).
-The general result emerging from these models is that competition for shared (substitutable) resources results in the evolution of consumers that specialize on different resources relative to the generalists that evolve in the absence of competition.
-The only scenarios in which character divergence will not occur is when density dependence in intra- or interspecific interactions due to other factors are sufficiently strong (Abrams 1986), although there remains little empirical evidence of this to date (Schluter 2000).
-Despite the many models that have been developed to examine the ecological scenarios resulting in character displacement, scant attention has been paid to the ecological consequences of such character divergence on resource abundances and food-web stability (but see Lawlor and Smith 1976 for an assessment of stability).
+### Model of resource competition
 
-# In terms of resources, perhaps the intuitive expectation is that character divergence would suppress resource abundances.
-The rationale being that character divergence will result in specialized consumers that are more effective at attacking a particular resource.
-Below, we demonstrate that this is not necessarily true.
-In fact, the effect of character displacement on resource abundances depends on both the foraging scenario and tradeoff in consumer feeding rates.
-```
-
-## How does character displacement affect resource abundances?
-
-To examine this, we used a continuous-time model of consumer-resource dynamics for two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
+We analyzed a continuous-time model of consumer-resource dynamics for two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
 
@@
 seed size; Taper and Case 1985).
-In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate, such that $F_{ij}(R_j)=a_{ij}R_j$, then our model encapsulates MacArthur's model of resource competition (MacArthur 1972).
+In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate, such that $F_{ij}(R_j)=a_{ij}R_j$, then our model becomes MacArthur's model of resource competition (MacArthur 1972).
 The evolution of consumer attack rates in this model, and several extensions, have been analyzed in detail (Lawlor and Smith 1976; Abrams 1986).
@@
 Here, we examine the effects of this inevitable divergence on resource dynamics and food-web stability.
-We show that the effects of character displacement are not intuitive, and depend critically on the foraging scenario and tradeoff in consumer attack rates.
+We show that the ecological consequences of character displacement are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
 
-## Consumers forage for resources simultaneously
+### Consumers forage for resources simultaneously
 
@@
 bind_rows(convex_df, linear_df, concave_df) %>%
-  ggplot(., aes(x=aii, y=aij, color = factor(n))) + geom_line() + ggtitle("Tradeoffs in attack rates")
+  ggplot(., aes(x=aii, y=aij, linetype = factor(n))) + 
+  geom_line() + 
+  xlab(expression(a[ii])) +
+  ylab(expression(a[ij])) +
+  scale_linetype_manual(values=1:3, name="Tradeoff (n)") 
 ```
+
+DESCRIBE THE SIMULATIONS THAT ARE USED IN ALL OF THE FOLLOWING ANALYSES.
 
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  ylab(expression(a[ii]+a[ij])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
   scale_fill_manual(values=c("white","black"))
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  ylab(expression(R[i]+R[j])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
   scale_fill_manual(values=c("white","black"))
 
-plot_grid(Mac_attack, Mac_resources, ncol=2)
+Mac_plots <- plot_grid(Mac_attack + theme(legend.position = "none"), Mac_resources + theme(legend.position = "none"), nrow=1)
+
+Mac_legend <- get_legend(Mac_attack)
+
+Mac_plot_legend <- plot_grid(Mac_plots, Mac_legend, ncol=2, rel_widths = c(3,0.3), align='hv')
+
+Mac_title <- ggdraw() + draw_label("Resources in same habitat", fontface='bold')
+plot_grid(Mac_title, Mac_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
 ```
 
-## Consumers cannot forage for resources simultaneously
+### Consumers cannot forage for resources simultaneously
 
@@
 
-Lawlor and Smith (1976) analyzed character divergence in models where consumers could and could not forage for resources simultaneously, and they found that consumers still underwent character divergence regardless of the foraging scenario.
+Lawlor and Smith (1976) found that consumers still underwent character divergence regardless of whether consumers could or could not forage for resources simultaneously.
 Interestingly, we find that the foraging scenario qualitatively affects the relationship between character divergence and resource abundances.
-To demonstrate this, note that now resource abundances at equilibrium are determined by the following equation (see derivation in Appendix):
+To demonstrate this, note that now resource abundances at equilibrium (both consumers and resources present) are determined by the following equation (see derivation in Appendix):
 
@@
 Thus, character divergence always results in resource suppression if consumers are competing for resources that occur in distinct habitats.
+
+ACTUALLY, THE SHAPE OF THE TRADE-OFF DOES DAMPEN OR AMPLIFY BUT NOT BECAUSE OF THE MAGNITUDE OF DISPLACEMENT.
+ITS BECAUSE CONCAVE TRADEOFFS HAVE THE LOWEST ATTACK RATE AT INTERMEDIATE VALUES, BUT HIGHEST AT EXTREME VALUES, RESULTING IN THE LARGEST CHANGE IN RESOURCE ABUNDANCES WITH CHARACTER DIVERGENCE.
+ESSENTIALLY, THERE IS MORE 'ROOM' TO EVOLVE DIFFERENCES.
+SO, THE MAGNITUDE OF DISPLACEMENT IS ONLY A USEFUL PREDICTOR OF AN EFFECT ON RESOURCE ABUNDANCES WHEN COMPARED WITHIN SPECIES, IT CAN'T BE USED TO PREDICT ACROSS SPECIES, WHERE THE NATURE OF THE TRADE-OFFS COULD BE DIFFERENT.
+
+DISCUSS ALSO THE DEVIATIONS FROM THE 2C-2R PREDICTIONS WHEN A COMPETITOR IS ABSENT. THESE DEVIATIONS INCREASE AS WE INCREASE MODEL COMPLEXITY.
+
 Note that the shape of the tradeoff can dampen or amplify the effect of character divergence on resource supression.
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  ylab(expression(w[ii]*a[ii]+(1-w[ii])*a[ij])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
   scale_fill_manual(values=c("white","black"))
@@
   geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor),
-             shape=21, size=10, inherit.aes = FALSE) +
+             shape=21, size=5, inherit.aes = FALSE) +
+  ylab(expression(R[i]+R[j])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
   scale_fill_manual(values=c("white","black"))
 
-plot_grid(LS_attack, LS_resources, ncol=2)
+LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), LS_resources + theme(legend.position = "none"), nrow=1)
+
+LS_legend <- get_legend(LS_attack)
+
+LS_plot_legend <- plot_grid(LS_plots, LS_legend, nrow=1, rel_widths = c(3,0.3))
+
+LS_title <- ggdraw() + draw_label("Resources in different habitats", fontface='bold')
+plot_grid(LS_title, LS_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
 ```
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  ylab(expression(w[ii]*a[ii]+(1-w[ii])*a[ij])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
   scale_fill_manual(values=c("white","black"))
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=10, inherit.aes = FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  ylab(expression(R[i]+R[j])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
   scale_fill_manual(values=c("white","black"))
 
-plot_grid(McCann_attack, McCann_resources, ncol=2)
+McCann_plots <- plot_grid(McCann_attack + theme(legend.position = "none"), McCann_resources + theme(legend.position = "none"), nrow=1)
+
+McCann_legend <- get_legend(McCann_attack)
+
+McCann_plot_legend <- plot_grid(McCann_plots, McCann_legend, nrow=1, rel_widths = c(3,0.3))
+
+McCann_title <- ggdraw() + draw_label("Resources in different habitats & realistic functional response", fontface='bold')
+plot_grid(McCann_title, McCann_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
 ```
 
-## Predictions for threespine stickleback
+### Predictions for threespine stickleback
 
@@
 
-## Effect of character divergence on food-web stability
+### Effect of character divergence on food-web stability
 

### 2018-05-11 eco_evo_simulations.R

@@ -66,18 +66,18 @@ states_1C_2R <- data.frame(R1 = R1, R2 = R2, C1 = C1)
 a12_tradeoff <- expression(A*(1-(a11m/A)^n)^(1/n)) 
 a21_tradeoff <- expression(A*(1-(a22m/A)^n)^(1/n)) 
 
-n_convex <- 1.15
+n_concaveDN <- 1.15
 n_linear <- 1
-n_concave <- 0.85
+n_concaveUP <- 0.85
 
-convex_df <- data.frame(n = n_convex, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_convex)^(1/n_convex))
+concaveDN_df <- data.frame(n = n_concaveDN, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_concaveDN)^(1/n_concaveDN))
 linear_df <- data.frame(n = n_linear, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_linear)^(1/n_linear))
-concave_df <- data.frame(n = n_concave, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_concave)^(1/n_concave))
+concaveUP_df <- data.frame(n = n_concaveUP, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_concaveUP)^(1/n_concaveUP))
 
-bind_rows(convex_df, linear_df, concave_df) %>%
+bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
   ggplot(., aes(x=aii, y=aij, color = factor(n))) + geom_line() + ggtitle("Attack rate trade-offs")
 
-bind_rows(convex_df, linear_df, concave_df) %>%
+bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
   ggplot(., aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + geom_line() + ggtitle("Relationship between specialization and total attack rate")
 
 ######################################### MACARTHUR MODEL #######################################################
@@ -97,17 +97,17 @@ params_2C_2R_MacArthur_linear <- data.frame(general_parameters,
 params_1C_2R_MacArthur_linear <- params_2C_2R_MacArthur_linear[c("r1","r2","K1","K2","e11","e12","a11","a12","m1")]
 
 
-## Convex
[…]

### 2018-05-11 maybe_useful_morgue.R

@@ -1,3 +1,9 @@
+```{r}
+# Numerous mathematical models have been developed to examine how resource competition influences the evolution of consumer foraging traits (reviewed in Taper and Case 1985). The general result emerging from these models is that competition for shared (substitutable) resources results in the evolution of consumers that specialize on different resources relative to the generalists that evolve in the absence of competition. The only scenarios in which character divergence will not occur is when density dependence in intra- or interspecific interactions due to other factors are sufficiently strong (Abrams 1986), although there remains little empirical evidence of this to date (Schluter 2000). Despite the many models that have been developed to examine the ecological scenarios resulting in character displacement, scant attention has been paid to the ecological consequences of such character divergence on resource abundances and food-web stability (but see Lawlor and Smith 1976 for an assessment of stability). 
+
+# In terms of resources, perhaps the intuitive expectation is that character divergence would suppress resource abundances. The rationale being that character divergence will result in specialized consumers that are more effective at attacking a particular resource. Below, we demonstrate that this is not necessarily true. In fact, the effect of character displacement on resource abundances depends on both the foraging scenario and tradeoff in consumer feeding rates.
+```
+
 We note, however, that our analytical insights to the effects of consumer divergence on resource dynamics has considered a simplistic scenario (resources are equivalent and consumers have equivalent conversion efficiences, but have mirror image attack rates on each resource). To explore the generality of our predictions, we examined whether our conclusions still held under a foraging scenario that was designed to mimic the stickleback lake food web. To do this, we used empirical relationships between body size and the parameters that govern consumer-resource dynamics to parameterize an eco-evolutionary model. Specifically, we used consumer and resource biomasses 
 
 ```{r Allometric Scaling of Stickleback Food Web}

### 2018-05-11 ECD_Supp_Mat.Rmd
adding stability results

@@
 output: html_document
+  fig_caption: yes
 author: Matt Barbour
@@
 $$\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2$$
-where $r_i$ represents intrinsic growth rates, $K_i$ represents carrying capacity, $e_{ij}$ represents the conversion efficiency of resources into new consumers, and $m_i$ represents the mortality rate of consumers.
-$F_{ij}(R_j)$ represents the consumer's feeding rate (or functional response) on a specific resource (in this case, $C_i$ on $R_j$).
+Resource growth is determined by is intrinsic growth rate ($r_i$), carrying capacity ($K_i$), and the feeding rate (or functional response) of each consumer ($F_{ii}(R_i)$, $F_{ij}(R_j)$).
+Consumer growth is determined by its feeding rate on each resource as well as its conversion efficiency ($e_{ij}$) and mortality rate ($m_i$).
 This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g.
@@
 seed size; Taper and Case 1985).
-In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate, such that $F_{ij}(R_j)=a_{ij}R_j$, then our model becomes MacArthur's model of resource competition (MacArthur 1972).
+In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate ($a_{ij}$), such that $F_{ij}(R_j)=a_{ij}R_j$, then our model becomes MacArthur's model of resource competition (MacArthur 1972).
 The evolution of consumer attack rates in this model, and several extensions, have been analyzed in detail (Lawlor and Smith 1976; Abrams 1986).
-The general result of these analyses is that consumer divergence is an inevitable outcome of competition.
-Here, we examine the effects of this inevitable divergence on resource dynamics and food-web stability.
+The general result of these analyses is that consumer divergence is an inevitable outcome of resource competition.
+
+Here, we examine the effects of this inevitable divergence on resource dynamics.
+
+GIVE DETAILS ABOUT MY ANALYTICAL ANALYSIS AS WELL AS THE TYPE OF SIMULATIONS I PERFORM.
+
 We show that the ecological consequences of character displacement are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
@@
 
-The first foragine scenario we explore is already assumed in MacArthur's model of resource competition, namely that consumers can forage for both resources simultaneously.
-This also implies that both resources occur within the same microhabitat.
+The first foraging scenario we examine is already assumed in MacArthur's model of resource competition, namely that consumers can forage for both resources simultaneously (e.g.
+they occur in the same habitat).
 To gain analytical insight to this scenario, we assume that resources are equivalent ($r=r_i$ and $K=K_i$) as well as consumers ($e=e_{ii}=e_{ij}$; $m=m_i$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
-While this scenario is simple, it isolates the effects of character displacement, rather than any inherent differences in consumers or resources.
-In the appendix, we show that when both consumers and resources are present that the abundance of resources at equilibrium are equivalent ($R_i=R_1=R_2$) and are determined by the following equation:
+While this scenario is simple, it allow us to isolate the effects of character displacement rather than any inherent differences in consumers or resources.
+In the appendix, we show that when both consumers and resources are present, the abundance of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation:
 
-$$\hat{R_i}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
+$$\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
 Note that a key determinant of resource abundance is the consumer's total attack rate, $a_{ii}+a_{ij}$.
-Since the total attack rate is in the denominator, this implies that resource abundances will decrease if evolution favors an increase in total attack rate.
-Consumers, however, cannot maximize their attack rate
-
-To understand how a consumer's total attack rate would change with character divergence, we explored the effect of different tradeoffs in consumer attack rates. We used the following general function:
-
-$$\Big(\frac{a_{ii}}{A}\Big)^n+\Big(\frac{a_{ij}}{A}\Big)^n=1$$
-
-where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff.
+Since the total attack rate is in the denominator, this implies that resource abundances will decrease if consumers evolve higher total attack rates.
+If there were no constraints on attack rates, each consumer would evolve increasingly higher attack rates on each resource until it goes extinct.
+In consideration of this, we have included a tradeoff between investments in attacking one resource versus the other such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff.
 This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$; Sargent and Otto 2006).
@@
 
-```{r Tradeoff function, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-bind_rows(convex_df, linear_df, concave_df) %>%
+```{r Tradeoff function, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Tradeoffs in consumer attack rates"}
+bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
   ggplot(., aes(x=aii, y=aij, linetype = factor(n))) + 
@@
 ```
-
-DESCRIBE THE SIMULATIONS THAT ARE USED IN ALL OF THE FOLLOWING ANALYSES.
 
@@
 
-```{r Same resource habitat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r Same resource habitat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Effect of character divergence on total attack rates (A) and resource abundances (B) when both resources are found in the same habitat."}
 Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, linetype = factor(n))) + 
@@
 
-Mac_plots <- plot_grid(Mac_attack + theme(legend.position = "none"), Mac_resources + theme(legend.position = "none"), nrow=1)
+Mac_plots <- plot_grid(Mac_attack + theme(legend.position = "none"), 
+                       Mac_resources + theme(legend.position = "none"), 
+                       nrow=1, labels = "AUTO")
 
@@
 
-Mac_plot_legend <- plot_grid(Mac_plots, Mac_legend, ncol=2, rel_widths = c(3,0.3), align='hv')
+#Mac_plot_legend <- 
+plot_grid(Mac_plots, Mac_legend, ncol=2, rel_widths = c(3,0.3), align='hv')
 
-Mac_title <- ggdraw() + draw_label("Resources in same habitat", fontface='bold')
-plot_grid(Mac_title, Mac_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
+#Mac_title <- ggdraw() + draw_label("Resources in same habitat", fontface='bold')
+#plot_grid(Mac_title, Mac_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
 ```
@@
 
-```{r Distinct resource habitats, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r Distinct resource habitats, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Effect of character divergence on total attack rates (A) and resource abundances (B) when each resource is found in a different habitat."}
 LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, linetype = factor(n))) + 
@@
 
-LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), LS_resources + theme(legend.position = "none"), nrow=1)
+LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), 
+                      LS_resources + theme(legend.position = "none"), 
+                      nrow=1, labels = "AUTO")
 
@@
 
-LS_plot_legend <- plot_grid(LS_plots, LS_legend, nrow=1, rel_widths = c(3,0.3))
+#LS_plot_legend <- 
+plot_grid(LS_plots, LS_legend, nrow=1, rel_widths = c(3,0.3))
 
-LS_title <- ggdraw() + draw_label("Resources in different habitats", fontface='bold')
-plot_grid(LS_title, LS_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
+#LS_title <- ggdraw() + draw_label("Resources in different habitats", fontface='bold')
+#plot_grid(LS_title, LS_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
 ```
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Effect of character divergence on total attack rates (A) and resource abundances (B) when consumers have a more realistic functional response.
+Note also that each resource is found in a different habitat."}
 McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, linetype = factor(n))) + 
@@
 
-McCann_plots <- plot_grid(McCann_attack + theme(legend.position = "none"), McCann_resources + theme(legend.position = "none"), nrow=1)
+McCann_plots <- plot_grid(McCann_attack + theme(legend.position = "none"), 
+                          McCann_resources + theme(legend.position = "none"), 
+                          nrow=1, labels = "AUTO")
 
@@
 
-McCann_plot_legend <- plot_grid(McCann_plots, McCann_legend, nrow=1, rel_widths = c(3,0.3))
+#McCann_plot_legend <- 
+plot_grid(McCann_plots, McCann_legend, nrow=1, rel_widths = c(3,0.3))
 
-McCann_title <- ggdraw() + draw_label("Resources in different habitats & realistic functional response", fontface='bold')
-plot_grid(McCann_title, McCann_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
+#McCann_title <- ggdraw() + draw_label("Resources in different habitats & realistic functional response", fontface='bold')
+#plot_grid(McCann_title, McCann_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
 ```
+
+### Effect of character divergence on food-web stability
+
+Of the prior models of character displacement, only Lawlor and Smith (1976) examine the effect of character divergence on the ecological stability of the system.
+They conducted a detailed theoretical analysis of three different metrics of stability: (i) stable coexistence of consumers, defined as the mutual invasibility of each consumer; (ii) global stability of consumers, defined as the continued coexistence of both consumers when resources dynamics are perturbed; and (iii) resilience, defined as the rate at which the four-species community returns to equilibrium after a small perturbation to species abundances.
+Note that the first two metrics correspond to the stable coexistence of consumers, while the third metric measures the stability of the entire system.
+Lawlor and Smith found that character divergence enhanced the stable coexistence of consumers, yet had a small negative effect of on food-web resilience (which they downplay in their interpretation).
+
+Lawlor and Smith's work gave fundamental insight to the effect of character divergence on ecological stability; however, none of their stability analyses compared systems with and without a competitor.
+Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement (Schluter and McPhail 1992; Schluter 2000).
+In fact, the first two metrics of stability can only be assessed when both consumers are present, because it interprets the effect of divergence in the presence --- not absence --- of a competitor.
+Food-web reslience, however, can be analyzed in the presence and absence of a competitor, and thus provides a useful benchmark for exploring the effects of character displacement on food-web stability.
+
+To gain analytical insight, we examined the conditions governing the local stability of the four-species food web when consumers have a more realistic functional response (McCann 2005).
+In the Appendix, we show that the system will transition from being locally stable to a limit cycle under the following conditions:
+
+$$w_{ii}a_{ii}+(1-w_{ii})a_{ij} < \frac{e+hm}{hK(e-hm)}$$
+
+Previously, we showed that character displacement always increases the effective attack rate of consumers, regardless of the the shape of the tradeoff.
+Thus, character divergence will push system toward the boundary of stability where the system transitions from a stable equilibrium to a limit cycle.
+This result supports the general tendency observed by Lawlor and Smith's numerical examination of local stability when consumers exhibited linear functional responses.
+
+Still, this does not compare the resilience of food-webs in the presence and absence of a competitor.
+To explore this, we plot the results from the same simulations for resource abundances, but now stability is on the y-axis
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Effect of character divergence on food-web resilience across different foraging scenarios and tradeoffs in attack rates."}
+Stability_Mac <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
+  scale_fill_manual(values=c("white","black"))
+
+Stability_LS <- cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
+  scale_fill_manual(values=c("white","black"))
+
+Stability_McCann <- cbind.data.frame(tradeoff_df, McCann_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, linetype = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
+  scale_fill_manual(values=c("white","black"))
+
+Stability_plots <- plot_grid(Stability_Mac + theme(legend.position = "none"), 
+                             Stability_LS + theme(legend.position = "none"),
+                             Stability_McCann + theme(legend.position = "none"),
+                             nrow=1, labels = "AUTO")
+
+Stability_legend <- get_legend(Stability_Mac)
+
+#Stability_plot_legend <- 
+plot_grid(Stability_plots, Stability_legend, nrow=1, rel_widths = c(3,0.3))
+```
+
+Here we focus on local stability, which makes the most sense in a food-web context.
+We find that divergence follows the typical "check-mark" stability pattern that has been described recently in theoretical food-web models.
+
+Importantly, this metric of stability allows us to compare the stability of three-species vs.
+four-species system, which is key for understanding how ecological character displacement affects food-web dynamics.
 
@@
 Thus, we predict that character divergence in sticklebacks would supress resource densities.
-
-### Effect of character divergence on food-web stability
-
-This relationship between character divergence and resource abundances begs the question: how does character divergence affect food-web stability?
-This question has been addressed empirically by Pimentel as well as with a conceptual model.
-Lawlor and Smith though did a detailed theoretical analysis of how character divergence affects three different aspects of stability.
-They looked at stability through the lens of competition.
-The first aspect was whether character divergence resulted in mutual invasibility of consumers, a common metric of "stable coexistence" in competition models.
-Another very interesting aspect they looked at was a global metric that examined the persistence of the system to perturbations in resources.
-Here, they conclude the divergence increases community stability, because it expands the range of resource perturbations that a system can withstand before a consumer going extinct.
-Note though, that they overlooked that there conclusions do not actually quantify the effects of ecological character displacement on stability because they do not compare scenarios where a consumer is in allopatry and one where consumers are in sympatry.
-Finally, they also look at the Lyapunov exponent (local stability), and while they find a negative effect of divergence, they suggest that it would likely be overshadowed by the stabilizing effects of divergence on mutual invasibility and global stability.
-
-Here we focus on local stability, which makes the most sense in a food-web context.
-We find that divergence follows the typical "check-mark" stability pattern that has been described recently in theoretical food-web models.
-
-Importantly, this metric of stability allows us to compare the stability of three-species vs.
-four-species system, which is key for understanding how ecological character displacement affects food-web dynamics.

### 2018-05-20 ECD_Supp_Mat.Rmd
updated much of the writing

--- 2018-05-11 ECD_Supp_Mat.Rmd
@@
 zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g.
-seed size; Taper and Case 1985).
+seed size, Taper and Case 1985).
 In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate, such that $F_{ij}(R_j)=a_{ij}R_j$, then our model becomes MacArthur's model of resource competition (MacArthur 1972).
@@
 The general result of these analyses is that consumer divergence is an inevitable outcome of competition.
-Here, we examine the effects of this inevitable divergence on resource dynamics and food-web stability.
-We show that the ecological consequences of character displacement are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
+To be explicit, this means that the specialization of evolved consumers ($\frac{a_{ii}{a_{ii}+a_{ij}}$) is greater in the presence of competitor compared to when no competitor is present.
+
+Here, we examine the effects of this inevitable divergence in attack rates on food-web dynamics.
+In particular, we examine how divergence in attack rates affects resource abundances as well as the resilience of both resources and consumers to perturbations.
+We do this by (i) deriving analytical expressions for the relationship between attack rates and food-web dynamics and (ii) simulate the effects of competition on the eco-evolutionary dynamics of consumers and resources.
+For these simulations, we used an Adaptive Dynamics approach.
+At each evolutionary time step, we created a mutant consumer by randomly choosing a consumer and modifying its attack rate by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability.
+Assuming this mutant consumer was rare, we then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer.
+If the mutant was able to invade, we updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state.
+We then repeated the simulation 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy (ESS, Price & Maynard Smith 1973) or limit (e.g.
+$\frac{a_{ii}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
+Below is a table of the parameters used in these simulations.
+Our mathematical analyses show that the consequences of character displacement on food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
+
+## Effect of character divergence on resource abundances
+
+We start by examing the effects of character displacement on the abundance of resources at equilibrium, which will also set the stage for a later analysis of food-web stability.
+In particular, we examine three different foraging scenarios: (i) consumers forage for resource simultaneously and their feeding rates increase linearly with resource density (MacArthur 1972); (ii) consumers *cannot* forage for resources simultaneously and their feeding rates increase linearly with resource density (Lawlor & Smith 1976); (iii) consumers *cannot* forage for resources simulaneously and their feeding rates on each resource depend on both resource densities and saturate at high densities (McCann 2005).
 
@@
 The first foragine scenario we explore is already assumed in MacArthur's model of resource competition, namely that consumers can forage for both resources simultaneously.
-This also implies that both resources occur within the same microhabitat.
-To gain analytical insight to this scenario, we assume that resources are equivalent ($r=r_i$ and $K=K_i$) as well as consumers ($e=e_{ii}=e_{ij}$; $m=m_i$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
-While this scenario is simple, it isolates the effects of character displacement, rather than any inherent differences in consumers or resources.
-In the appendix, we show that when both consumers and resources are present that the abundance of resources at equilibrium are equivalent ($R_i=R_1=R_2$) and are determined by the following equation:
+In essence, this implies that both resources occur within the same microhabitat.
+To gain analytical insight to this scenario, we assume that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
+While this scenario is simple, it isolates the process of character displacement, rather than whether these effects depend on inherent differences in consumers or resources.
+In the appendix, we show that when both consumers and resources are present that the abundance of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation:
 
-$$\hat{R_i}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
+$$\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
 Note that a key determinant of resource abundance is the consumer's total attack rate, $a_{ii}+a_{ij}$.
 Since the total attack rate is in the denominator, this implies that resource abundances will decrease if evolution favors an increase in total attack rate.
-Consumers, however, cannot maximize their attack rate
-
-To understand how a consumer's total attack rate would change with character divergence, we explored the effect of different tradeoffs in consumer attack rates. We used the following general function:
+We assume that consumers are not 'Darwinian demons' (Law 1979) and thus the evolution of their attack rates are subject to certain constraints.
+We modelled these constraints using the following general function:
 
@@
 ```{r Tradeoff function, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-bind_rows(convex_df, linear_df, concave_df) %>%
+bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
   ggplot(., aes(x=aii, y=aij, linetype = factor(n))) + 
@@
 
-DESCRIBE THE SIMULATIONS THAT ARE USED IN ALL OF THE FOLLOWING ANALYSES.
-
 We know from past work that consumers still undergo character divergence regardless of the shape of the tradeoff (Lawlor and Smith 1976; Abrams 1986).
-However, the shape of the tradeoff qualitatively affects the relationship between character divergence and resource abundance.
+Interestingly, we find that the shape of the tradeoff qualitatively affects the relationship between character divergence and resource abundance.
 For example, if consumer's are constrained by a linear tradeoff ($n=1$), such that a unit increase in $a_{ii}$ results in a unit decrease in $a_{ij}$, then character divergence has no effect on resource abundances.
@@
 When the tradeoff is concave up ($n<1$), however, we see that character divergence suppresses resource densities.
-Although this equilibrium equation for resource abundances was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when there is only one consumer as well (white circle).
+Although this equilibrium equation for resource abundances was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when there is only one consumer (white circle).
 This is because a consumer, in the absence of competition, evolves to be a generalist that has equal attack rates on each resource, resulting in equivalent resource abundances.
@@
 
-### Consumers cannot forage for resources simultaneously
+### Consumers *cannot* forage for resources simultaneously
 
 Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (Taper and Case 1985; Abrams 1986).
-This assumption may be valid for some foraging scenarios, such as Darwin's finches foraging for seeds.
-However, in many situations, consumers forage for resources that occur in distinct habitats, and thus cannot forage for both simultaneosly (McCann et al.
-2005).
+This assumption may be valid for some foraging scenarios (e.g.
+Darwin's finches foraging for seeds); however, in many situations consumers forage for resources that occur in distinct habitats, and thus cannot forage for both simultaneosly.
 The only chracter displacement model that we are aware of that modeled resources in distinct habitats was one examined by Lawlor and Smith (1976).
@@
 
-where $w_{ij}$ represents the proportion of time a consumer ($C_i$) spends foraging in a habitat where a particular resource ($R_j$) is found.
+where $w_{ij}$ represents the proportion of time a consumer ($C_i$) spends foraging in a habitat where a particular resource ($R_j$) is found (i.e.
+habitat preference).
 Note that since $w_ij$ is a proportion that $w_{ii}=1-w_{ij}$.
+As with attack rates, we assume that consumer habitat preferences are mirror images of each other ($w=w_{11}=w_[22]$).
 
 Lawlor and Smith (1976) found that consumers still underwent character divergence regardless of whether consumers could or could not forage for resources simultaneously.
-Interestingly, we find that the foraging scenario qualitatively affects the relationship between character divergence and resource abundances.
+We find that the foraging scenario qualitatively affects the relationship between character divergence and resource abundances.
 To demonstrate this, note that now resource abundances at equilibrium (both consumers and resources present) are determined by the following equation (see derivation in Appendix):
 
-$$\hat{R_i}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e}$$
+$$\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e}$$
 
 This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g.
-$w_{ii}>0.5$ and $a_{ii}>a_{ij}$), then resource abundances will always decrease, regardless of the tradeoff.
+$w>0.5$ and $a_{ii}>a_{ij}$), then resource abundances will always decrease, regardless of the tradeoff.
 Thus, character divergence always results in resource suppression if consumers are competing for resources that occur in distinct habitats.
 
-ACTUALLY, THE SHAPE OF THE TRADE-OFF DOES DAMPEN OR AMPLIFY BUT NOT BECAUSE OF THE MAGNITUDE OF DISPLACEMENT.
-ITS BECAUSE CONCAVE TRADEOFFS HAVE THE LOWEST ATTACK RATE AT INTERMEDIATE VALUES, BUT HIGHEST AT EXTREME VALUES, RESULTING IN THE LARGEST CHANGE IN RESOURCE ABUNDANCES WITH CHARACTER DIVERGENCE.
-ESSENTIALLY, THERE IS MORE 'ROOM' TO EVOLVE DIFFERENCES.
-SO, THE MAGNITUDE OF DISPLACEMENT IS ONLY A USEFUL PREDICTOR OF AN EFFECT ON RESOURCE ABUNDANCES WHEN COMPARED WITHIN SPECIES, IT CAN'T BE USED TO PREDICT ACROSS SPECIES, WHERE THE NATURE OF THE TRADE-OFFS COULD BE DIFFERENT.
+Note that the shape of the tradeoff can dampen or amplify the effect of character displacement.
+This is not because the tradeoff affects the magnitude of displacement, because as you can see, it has little effect (slightly less with concave down tradeoff).
+Its because for the concave up tradeoff, consumers have the lowest attack rate at intermediate values, but highest at extreme values, resulting in the largest change in resource abundances with character divergence.
 
-DISCUSS ALSO THE DEVIATIONS FROM THE 2C-2R PREDICTIONS WHEN A COMPETITOR IS ABSENT. THESE DEVIATIONS INCREASE AS WE INCREASE MODEL COMPLEXITY.
+These results do lead to an interesting prediction that could be tested.
+If consumers are competing for resources that occur in distinct habitats, then greater ecological character displacement corresponds to a greater decrease in resource densities.
+Note that this prediction is only useful for comparing divergence within species.
+It cannot be used to predict across species where the nature of the trade-offs could be different.
 
-Note that the shape of the tradeoff can dampen or amplify the effect of character divergence on resource supression.
-This is because the shape of the tradeoff influences the magnitude of character displacement.
-For example, a concave-down tradeoff results in less displacement and less of an effect on resource abundances; whereas a concave-up tradeoff results in the largest displacement and the largest effect on resource abundances.
+It is important to note that, in this foraging scenario, resource abundances in the absence of competition no longer correspond to the analytical expression we derived when both consumers are present.
+This is because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat.
+Note that our analytical expression consistently underestimates the effect of character displacement on resource dynamics.
+This implies that resource abundances are always higher than our analytical expression suggests when there is only one consumer.
+We will revisit this subtle point when we examine the effects of character displacement on food-web stability.
 
@@
 
-$$W_{ii}=\frac{w_{ii}R_i}{w_{ii}R_i+(1-w_{ii})R_j}$$
+$$W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}$$
 
@@
 
-$$\hat{R_i}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e-hm}$$
+$$\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e-hm}$$
+
+This more realistic consumer functional response does actually make an interesting evolutionary prediction that differs from the scenario where consumer feeding rate scales linearly with resource density.
+Namely, that in the absence of competition, consumers remain slightly more specialized on the resources that occur in their preferred habitats, which actually opposes the prediction from the model where consumer feeding rate scales linearly with resource density.
 
@@
 
-The scenario where the relationship between character divergence and resource suppression is the strongest is when the tradeoff in attack rates is concave-up and consumers are competing for resources that occur in distinct habitats.
+The scenario where the relationship between character divergence and resource suppression is the strongest is when the tradeoff in attack rates is concave up and consumers are competing for resources that occur in distinct habitats.
 Prior work in threespine stickleback suggests that the evolution of stickleback attack rates are constrained by a concave-up tradeoff (Schluter 1993; Arnegard et al.
@@
 
-### Effect of character divergence on food-web stability
+## Effect of character divergence on food-web stability
 
@@
 four-species system, which is key for understanding how ecological character displacement affects food-web dynamics.
+
+### Conclusions
+
+Although we used an Adaptive Dynamics approach here, we expect that our results would apply to different models of the genetic architecture of attack rates (e.g.
+quantitative genetics, Taper & Case 1985), given that character divergence is an inevitable outcome of resource competition, regardless of the model of genetic architecture, when resources dynamics are explicitly modelled (Taper & Case 1985).
+
+These conclusions are only for substitutable resources and that exhibit biological dynamics.

### 2018-05-20 ECD_Supp_Mat.Rmd

--- 2018-05-11 ECD_Supp_Mat.Rmd
@@
 $$\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2$$
-Resource growth is determined by is intrinsic growth rate ($r_i$), carrying capacity ($K_i$), and the feeding rate (or functional response) of each consumer ($F_{ii}(R_i)$, $F_{ij}(R_j)$).
-Consumer growth is determined by its feeding rate on each resource as well as its conversion efficiency ($e_{ij}$) and mortality rate ($m_i$).
+
+where $r_i$ represents intrinsic growth rates, $K_i$ represents carrying capacity, $e_{ij}$ represents the conversion efficiency of resources into new consumers, and $m_i$ represents the mortality rate of consumers.
+$F_{ij}(R_j)$ represents the consumer's feeding rate (or functional response) on a specific resource (in this case, $C_i$ on $R_j$).
 This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g.
 zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g.
-seed size; Taper and Case 1985).
-In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate ($a_{ij}$), such that $F_{ij}(R_j)=a_{ij}R_j$, then our model becomes MacArthur's model of resource competition (MacArthur 1972).
+seed size, Taper and Case 1985).
+In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate, such that $F_{ij}(R_j)=a_{ij}R_j$, then our model becomes MacArthur's model of resource competition (MacArthur 1972).
 The evolution of consumer attack rates in this model, and several extensions, have been analyzed in detail (Lawlor and Smith 1976; Abrams 1986).
-The general result of these analyses is that consumer divergence is an inevitable outcome of resource competition.
+The general result of these analyses is that consumer divergence is an inevitable outcome of competition.
+To be explicit, this means that the specialization of evolved consumers ($\frac{a_{ii}{a_{ii}+a_{ij}}$) is greater in the presence of competitor compared to when no competitor is present.
 
-Here, we examine the effects of this inevitable divergence on resource dynamics.
+Here, we examine the effects of this inevitable divergence in attack rates on food-web dynamics.
+In particular, we examine how divergence in attack rates affects resource abundances as well as the resilience of both resources and consumers to perturbations.
+We do this by (i) deriving analytical expressions for the relationship between attack rates and food-web dynamics and (ii) simulate the effects of competition on the eco-evolutionary dynamics of consumers and resources.
+For these simulations, we used an Adaptive Dynamics approach.
+At each evolutionary time step, we created a mutant consumer by randomly choosing a consumer and modifying its attack rate by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability.
+Assuming this mutant consumer was rare, we then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer.
+If the mutant was able to invade, we updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state.
+We then repeated the simulation 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy (ESS, Price & Maynard Smith 1973) or limit (e.g.
+$\frac{a_{ii}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
+Below is a table of the parameters used in these simulations.
+Our mathematical analyses show that the consequences of character displacement on food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
 
-GIVE DETAILS ABOUT MY ANALYTICAL ANALYSIS AS WELL AS THE TYPE OF SIMULATIONS I PERFORM.
+## Effect of character divergence on resource abundances
 
-We show that the ecological consequences of character displacement are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
+We start by examing the effects of character displacement on the abundance of resources at equilibrium, which will also set the stage for a later analysis of food-web stability.
+In particular, we examine three different foraging scenarios: (i) consumers forage for resource simultaneously and their feeding rates increase linearly with resource density (MacArthur 1972); (ii) consumers *cannot* forage for resources simultaneously and their feeding rates increase linearly with resource density (Lawlor & Smith 1976); (iii) consumers *cannot* forage for resources simulaneously and their feeding rates on each resource depend on both resource densities and saturate at high densities (McCann 2005).
 
@@
 
-The first foraging scenario we examine is already assumed in MacArthur's model of resource competition, namely that consumers can forage for both resources simultaneously (e.g.
-they occur in the same habitat).
-To gain analytical insight to this scenario, we assume that resources are equivalent ($r=r_i$ and $K=K_i$) as well as consumers ($e=e_{ii}=e_{ij}$; $m=m_i$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
-While this scenario is simple, it allow us to isolate the effects of character displacement rather than any inherent differences in consumers or resources.
-In the appendix, we show that when both consumers and resources are present, the abundance of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation:
+The first foragine scenario we explore is already assumed in MacArthur's model of resource competition, namely that consumers can forage for both resources simultaneously.
+In essence, this implies that both resources occur within the same microhabitat.
+To gain analytical insight to this scenario, we assume that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
+While this scenario is simple, it isolates the process of character displacement, rather than whether these effects depend on inherent differences in consumers or resources.
+In the appendix, we show that when both consumers and resources are present that the abundance of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation:
 
@@
 Note that a key determinant of resource abundance is the consumer's total attack rate, $a_{ii}+a_{ij}$.
-Since the total attack rate is in the denominator, this implies that resource abundances will decrease if consumers evolve higher total attack rates.
-If there were no constraints on attack rates, each consumer would evolve increasingly higher attack rates on each resource until it goes extinct.
-In consideration of this, we have included a tradeoff between investments in attacking one resource versus the other such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff.
+Since the total attack rate is in the denominator, this implies that resource abundances will decrease if evolution favors an increase in total attack rate.
+We assume that consumers are not 'Darwinian demons' (Law 1979) and thus the evolution of their attack rates are subject to certain constraints.
+We modelled these constraints such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff.
 This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$; Sargent and Otto 2006).
@@
 We know from past work that consumers still undergo character divergence regardless of the shape of the tradeoff (Lawlor and Smith 1976; Abrams 1986).
-However, the shape of the tradeoff qualitatively affects the relationship between character divergence and resource abundance.
+Interestingly, we find that the shape of the tradeoff qualitatively affects the relationship between character divergence and resource abundance.
 For example, if consumer's are constrained by a linear tradeoff ($n=1$), such that a unit increase in $a_{ii}$ results in a unit decrease in $a_{ij}$, then character divergence has no effect on resource abundances.
@@
 
-Here we focus on local stability, which makes the most sense in a food-web context.
-We find that divergence follows the typical "check-mark" stability pattern that has been described recently in theoretical food-web models.
-
-Importantly, this metric of stability allows us to compare the stability of three-species vs.
-four-species system, which is key for understanding how ecological character displacement affects food-web dynamics.
-
 ### Predictions for threespine stickleback
@@
 Thus, we predict that character divergence in sticklebacks would supress resource densities.
+
+### Other relevant points
+
+Although we used an Adaptive Dynamics approach here, we expect that our results would apply to different models of the genetic architecture of attack rates (e.g.
+quantitative genetics, Taper & Case 1985), given that character divergence is an inevitable outcome of resource competition, regardless of the model of genetic architecture, when resources dynamics are explicitly modelled (Taper & Case 1985).
+
+These conclusions are only for substitutable resources and that exhibit biological dynamics.

### 2018-05-20 ECD_Supp_Mat.Rmd
finalized draft. still trying to git figures to print with doc

--- 2018-05-20 ECD_Supp_Mat.Rmd
@@
 ---
-title: "How does character divergence affect resource dynamics?"
-output: html_document
-author: Matt Barbour
+title: "How does character displacement affect food-web dynamics?"
+author: "Matt Barbour"
+output:
+  html_document: default
+  fig_caption: yes
 ---
@@
 knitr::opts_chunk$set(
-	message = FALSE,
+	message = TRUE,
 	warning = FALSE,
@@
 source('eco_evo_simulations.R')
+
+library(knitr)
 ```
 
-### Model of resource competition
+***
+
+## Model of resource competition
 
@@
 $$\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2$$
+
 where $r_i$ represents intrinsic growth rates, $K_i$ represents carrying capacity, $e_{ij}$ represents the conversion efficiency of resources into new consumers, and $m_i$ represents the mortality rate of consumers.
@@
 zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g.
-seed size, Taper and Case 1985).
+seed size, see Taper and Case 1985 for an example).
 In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate, such that $F_{ij}(R_j)=a_{ij}R_j$, then our model becomes MacArthur's model of resource competition (MacArthur 1972).
@@
 The general result of these analyses is that consumer divergence is an inevitable outcome of competition.
-To be explicit, this means that the specialization of evolved consumers ($\frac{a_{ii}{a_{ii}+a_{ij}}$) is greater in the presence of competitor compared to when no competitor is present.
+In other words, the presence of a competitor causes consumers to evolve more specialized attack rates ($\frac{a_{ii}}{a_{ii}+a_{ij}}$) compared to when no competitor is present.
 
@@
 If the mutant was able to invade, we updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state.
-We then repeated the simulation 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy (ESS, Price & Maynard Smith 1973) or limit (e.g.
-$\frac{a_{ii}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
-Below is a table of the parameters used in these simulations.
+We then repeated the simulation 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy (ESS, Price & Maynard Smith 1973) or an evolutionary limit (e.g.
+$\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
 Our mathematical analyses show that the consequences of character displacement on food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
 
-## Effect of character divergence on resource abundances
+***
+
+## Effect of character displacement on resource abundances
 
 We start by examing the effects of character displacement on the abundance of resources at equilibrium, which will also set the stage for a later analysis of food-web stability.
-In particular, we examine three different foraging scenarios: (i) consumers forage for resource simultaneously and their feeding rates increase linearly with resource density (MacArthur 1972); (ii) consumers *cannot* forage for resources simultaneously and their feeding rates increase linearly with resource density (Lawlor & Smith 1976); (iii) consumers *cannot* forage for resources simulaneously and their feeding rates on each resource depend on both resource densities and saturate at high densities (McCann 2005).
+In particular, we examine two different foraging scenarios: one where consumers forage for resource simultaneously (e.g.
+MacArthur 1972) and one where they cannot because resources occur in distinct habitats (e.g.
+Lawlor and Smith 1972; McCann 2005).
 
@@
 
-The first foragine scenario we explore is already assumed in MacArthur's model of resource competition, namely that consumers can forage for both resources simultaneously.
-In essence, this implies that both resources occur within the same microhabitat.
+MacArthur's model of resource competition assumes that consumers can forage for both resources simultaneously.
+In essence, this implies that both resources occur within the same habitat.
 To gain analytical insight to this scenario, we assume that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
-While this scenario is simple, it isolates the process of character displacement, rather than whether these effects depend on inherent differences in consumers or resources.
+While this scenario is arguably simplistic, it isolates the process of character displacement, rather than any dependence on inherent differences in consumers or resources.
 In the appendix, we show that when both consumers and resources are present that the abundance of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation:
@@
 We assume that consumers are not 'Darwinian demons' (Law 1979) and thus the evolution of their attack rates are subject to certain constraints.
-We modelled these constraints using the following general function:
+We modelled these constraints such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff (Sargent and Otto 2006).
+This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$, dotted line in \ref{fig:Tradeoff}) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$, solid line \ref{fig:Tradeoff}).
+When $n=1$, the tradeoff function is linear, and all combinations of $a_ii$ and $a_ij$ have the same total attack rate (dashed line in \ref{fig:Tradeoff}).
 
-$$\Big(\frac{a_{ii}}{A}\Big)^n+\Big(\frac{a_{ij}}{A}\Big)^n=1$$
-
-where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff.
-This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$; Sargent and Otto 2006).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_ii$ and $a_ij$ have the same total attack rate.
-
-```{r Tradeoff function, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
-  ggplot(., aes(x=aii, y=aij, linetype = factor(n))) + 
+```{r Tradeoff, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Tradeoffs in consumer attack rates"}
+tradeoff_plot <- bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
+  ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
   geom_line() + 
@@
   ylab(expression(a[ij])) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") 
+  scale_color_discrete(name="Tradeoff (n)")
+
+print(tradeoff_plot)
 ```
@@
 We know from past work that consumers still undergo character divergence regardless of the shape of the tradeoff (Lawlor and Smith 1976; Abrams 1986).
-Interestingly, we find that the shape of the tradeoff qualitatively affects the relationship between character divergence and resource abundance.
-For example, if consumer's are constrained by a linear tradeoff ($n=1$), such that a unit increase in $a_{ii}$ results in a unit decrease in $a_{ij}$, then character divergence has no effect on resource abundances.
-If the tradeoff is concave down ($n>1$), then resource abundances can actually increase under character displacement!
+Interestingly, we find that the shape of the tradeoff qualitatively affects the relationship between character divergence and resource abundance (\ref{fig:MacArthur_Resources}).
+For example, if consumer's are constrained by a linear tradeoff ($n=1$), such that a unit increase in $a_{ii}$ results in a unit decrease in $a_{ij}$, then character divergence has no effect on resource abundances (solid line in \ref{fig:MacArthur_Resources}).
+If the tradeoff is concave down ($n>1$), then resource abundances can actually increase under character displacement (dotted line in \ref{fig:MacArthur_Resources}).
 This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{aij}$) and decreases as consumers diverge.
-Note though that detecting this effect in nature will likely be difficult, since a concave down tradeoff results in relatively small displacement relative to other tradeoff forms (Lawlor and Smith 1976).
-When the tradeoff is concave up ($n<1$), however, we see that character divergence suppresses resource densities.
-Although this equilibrium equation for resource abundances was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when there is only one consumer (white circle).
-This is because a consumer, in the absence of competition, evolves to be a generalist that has equal attack rates on each resource, resulting in equivalent resource abundances.
+Note that detecting this effect in nature would likely be difficult, since a concave-down tradeoff results in relatively small displacement relative to other tradeoff shapes (Lawlor and Smith 1976).
+When the tradeoff is concave up ($n<1$), we find that character divergence suppresses resource densities.
+Although our equation for resource abundance was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when a single consumer reaches its ESS (white circle).
+This is because a single consumer evolves to be a generalist that has equal attack rates on each resource, resulting in equivalent resource abundances.
 
-```{r Same resource habitat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, linetype = factor(n))) + 
+```{r MacArthur_Resources, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Effect of character displacement on total attack rates (A) and resource abundances (B) when both resources are found in the same habitat."}
+Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
   ylab(expression(a[ii]+a[ij])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_discrete(name="Tradeoff (n)") +
+  scale_fill_discrete(name="Tradeoff (n)") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
 Mac_resources <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
   ylab(expression(R[i]+R[j])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_discrete(name="Tradeoff (n)") +
+  scale_fill_discrete(name="Tradeoff (n)") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
-Mac_plots <- plot_grid(Mac_attack + theme(legend.position = "none"), Mac_resources + theme(legend.position = "none"), nrow=1)
-
-Mac_legend <- get_legend(Mac_attack)
-
-Mac_plot_legend <- plot_grid(Mac_plots, Mac_legend, ncol=2, rel_widths = c(3,0.3), align='hv')
-
-Mac_title <- ggdraw() + draw_label("Resources in same habitat", fontface='bold')
-plot_grid(Mac_title, Mac_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
+Mac_plots <- plot_grid(Mac_attack + theme(legend.position = "none"), Mac_resources, 
+                       ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
+print(Mac_plots)
 ```
@@
 
-Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (Taper and Case 1985; Abrams 1986).
+Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (e.g.
+Taper and Case 1985; Abrams 1986).
 This assumption may be valid for some foraging scenarios (e.g.
@@
 $w>0.5$ and $a_{ii}>a_{ij}$), then resource abundances will always decrease, regardless of the tradeoff.
-Thus, character divergence always results in resource suppression if consumers are competing for resources that occur in distinct habitats.
+Thus, character displacement always results in resource suppression if consumers are competing for resources that occur in distinct habitats (\ref{fig:LS_Resources}).
+Note that the shape of the tradeoff can dampen or amplify the effect of character displacement.
+This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but it's rather minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached its ESS (white circles in \ref{fig:LS_Resources}).
+In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (black circles in \ref{fig:LS_Resources}), because character displacement tends to reach a constraint of complete specialization.
 
-Note that the shape of the tradeoff can dampen or amplify the effect of character displacement.
-This is not because the tradeoff affects the magnitude of displacement, because as you can see, it has little effect (slightly less with concave down tradeoff).
-Its because for the concave up tradeoff, consumers have the lowest attack rate at intermediate values, but highest at extreme values, resulting in the largest change in resource abundances with character divergence.
+These results do lead to an interesting prediction that could be tested either experimentally or with field data.
+If consumers are competing for resources that occur in distinct habitats, then greater ecological character displacement corresponds to a greater decrease in resource densities.
+This prediction only applies to comparisons of the effect of character displacement within species (assuming that allopatric and sympatric consumers are constrained by the same tradeoff).
 
-These results do lead to an interesting prediction that could be tested.
-If consumers are competing for resources that occur in distinct habitats, then greater ecological character displacement corresponds to a greater decrease in resource densities.
-Note that this prediction is only useful for comparing divergence within species.
-It cannot be used to predict across species where the nature of the trade-offs could be different.
-
-It is important to note that, in this foraging scenario, resource abundances in the absence of competition no longer correspond to the analytical expression we derived when both consumers are present.
-This is because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat.
-Note that our analytical expression consistently underestimates the effect of character displacement on resource dynamics.
-This implies that resource abundances are always higher than our analytical expression suggests when there is only one consumer.
+It is worth noting that in this foraging scenario, resource abundances when a single consumer reaches its ESS no longer correspond to the analytical expression we derived for when both consumers are present (deviation of white circles from lines \ref{fig:LS_Resources}).
+This is likely because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of white circles from 0.5 in \ref{fig:LS_Resources}).
+Also, our analytical expression consistently underestimates the effect of character displacement on resource dynamics.
+This implies that resource abundances are always higher than our analytical expression suggests when a single consumer reaches its ESS.
 We will revisit this subtle point when we examine the effects of character displacement on food-web stability.
 
-```{r Distinct resource habitats, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, linetype = factor(n))) + 
+```{r LS_Resources, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Effect of character divergence on total attack rates (A) and resource abundances (B) when each resource is found in a different habitat."}
+LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
-  ylab(expression(w[ii]*a[ii]+(1-w[ii])*a[ij])) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_discrete(name="Tradeoff (n)") +
+  scale_fill_discrete(name="Tradeoff (n)") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
 LS_resources <- cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color=factor(n))) + 
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor),
-             shape=21, size=5, inherit.aes = FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
   ylab(expression(R[i]+R[j])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_discrete(name="Tradeoff (n)") +
+  scale_fill_discrete(name="Tradeoff (n)") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
-LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), LS_resources + theme(legend.position = "none"), nrow=1)
-
-LS_legend <- get_legend(LS_attack)
-
-LS_plot_legend <- plot_grid(LS_plots, LS_legend, nrow=1, rel_widths = c(3,0.3))
-
-LS_title <- ggdraw() + draw_label("Resources in different habitats", fontface='bold')
-plot_grid(LS_title, LS_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
+LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), LS_resources, 
+                      ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
+print(LS_plots)
 ```
 
-This effect of the foraging scenario on the relationship between character divergence and resource abundances appears to be quite general.
+The effect of the foraging scenario on the relationship between character displacement and resource abundances appears to be quite general.
 For example, even if we consider a more realistic consumer functional response (derived by McCann et al.
@@
 
-we still observe the same qualitative relationship between character divergence and resource suppression.
+we still observe the same qualitative relationship between character displacement and resource suppression (\ref{fig:McCann_Resources}).
 This is because resource competition always leads to character divergence (see proof in Appendix) and equilibrium resource abundances are governed by similar dynamics:
@@
 This more realistic consumer functional response does actually make an interesting evolutionary prediction that differs from the scenario where consumer feeding rate scales linearly with resource density.
-Namely, that in the absence of competition, consumers remain slightly more specialized on the resources that occur in their preferred habitats, which actually opposes the prediction from the model where consumer feeding rate scales linearly with resource density.
+Specifically, a consumer's ESS in the absence of competition will remain slightly more specialized on the resources that occur in their preferred habitats.
+This is in contrast to the prediction when feeding rates scale linearly with resource density, where consumers evolve to be slightly more specialized on resources occurring in their non-preferred habitats.
+An important caveat for these predictions is that our model does not allow habitat preferences to evolve, which would likely alter these predictions.
 
-```{r}
-McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, linetype = factor(n))) + 
+```{r McCann_Resources, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Effect of character divergence on total attack rates (A) and resource abundances (B) when consumers have a more realistic functional response.
+Note also that each resource is found in a different habitat."}
+McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
-  ylab(expression(w[ii]*a[ii]+(1-w[ii])*a[ij])) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_discrete(name="Tradeoff (n)") +
+  scale_fill_discrete(name="Tradeoff (n)") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
 McCann_resources <- cbind.data.frame(tradeoff_df, McCann_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
   ylab(expression(R[i]+R[j])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_discrete(name="Tradeoff (n)") +
+  scale_fill_discrete(name="Tradeoff (n)") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
-McCann_plots <- plot_grid(McCann_attack + theme(legend.position = "none"), McCann_resources + theme(legend.position = "none"), nrow=1)
+McCann_plots <- plot_grid(McCann_attack + theme(legend.position = "none"), McCann_resources, 
+                          ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
+print(McCann_plots)
+```
 
-McCann_legend <- get_legend(McCann_attack)
+***
 
-McCann_plot_legend <- plot_grid(McCann_plots, McCann_legend, nrow=1, rel_widths = c(3,0.3))
+## Effect of character displacement on food-web stability
 
-McCann_title <- ggdraw() + draw_label("Resources in different habitats & realistic functional response", fontface='bold')
-plot_grid(McCann_title, McCann_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
+Of the prior models of character displacement, only Lawlor and Smith (1976) examine the effect of character displacement on the ecological stability of the system.
+They conducted a detailed theoretical analysis of three different metrics of stability: (i) stable coexistence of consumers, defined as the mutual invasibility of each consumer; (ii) global stability of consumers, defined as the continued coexistence of both consumers when resources dynamics are perturbed; and (iii) local stability, defined as the rate at which the four-species community returns to equilibrium after a small perturbation to species abundances.
+Note that the first two metrics correspond to the stable coexistence of consumers, while the third metric measures the stability of the entire system.
+Lawlor and Smith found that character displacement enhanced the stable coexistence of consumers, yet had a small negative effect on the resilience of the system.
+
+Lawlor and Smith's work gave fundamental insight to the effect of character divergence on ecological stability; however, none of their stability analyses compared food webs with and without a competing consumer.
+Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement (Schluter and McPhail 1992; Schluter 2000).
+In fact, the first two metrics of stability can only be assessed when both consumers are present because it interprets the effect of divergence in the presence of a competitor ---not in its absence.
+In contrast, local stability can be compared in food webs with and without a competing consumer, and thus provides a useful benchmark for exploring the effects of character displacement on food-web stability.
+
+Below, we plot the results from the same simulations we did previously except now local stability ($-1\times\operatorname{Re}(\lambda_{max})$) is on the y-axis (\ref{fig:Stability}).
+In general, we find that resource surpression goes hand-in-hand with local stability, a tendency that has been noted by others (Murdoch et al.
+2003; McCann 2012).
+Whenever ecological character displacement results in resource suppression, we observe a decrease in local stability (\ref{fig:Stability}).
+This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (compare initial points (small points) with evolved strategies (large points), \ref{fig:Stability}).
+
+```{r Stability, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Effect of character displacement on local stability across different foraging scenarios and tradeoffs in attack rates.
+Small and large points correspond to the beginning and of the simulation, respectively."}
+Stability_Mac <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend=FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_discrete(name="Tradeoff (n)") +
+  scale_fill_discrete(name="Tradeoff (n)") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_LS <- cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_discrete(name="Tradeoff (n)") +
+  scale_fill_discrete(name="Tradeoff (n)") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_McCann <- cbind.data.frame(tradeoff_df, McCann_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color=factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend = FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_discrete(name="Tradeoff (n)") +
+  scale_fill_discrete(name="Tradeoff (n)") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_plots <- plot_grid(Stability_Mac + theme(legend.position = "none") + ggtitle("Resources in same habitat"), 
+                             Stability_LS + theme(legend.position = "none", axis.title = element_blank(), axis.text.x = element_blank()) + ggtitle("Resources in different habitats"),
+                             NULL, Stability_McCann + theme(legend.position = "none"),
+                             ncol=2, align='hv', labels = c("A","B","","C"))
+
+Stability_legend <- get_legend(Stability_Mac)
+
+print(ggdraw(Stability_plots) + draw_grob(Stability_legend, x=0.2, y=-0.2))
 ```
+
+Our simulations focus on a scenario where the food web remains locally stable ($Re$\lambda_{max}$<0).
+Is it possible that ecological character displacement would ever result an unstable system?
+To gain analytical insight to this question, we examined the conditions governing the local stability of the four-species food web when consumers exhibit a more realistic functional response (McCann 2005).
+In the Appendix, we show that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions:
+
+$$wa_{ii}+(1-w)a_{ij} > \frac{e+hm}{hK(e-hm)}$$
+
+Previously, we showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff.
+Thus, ecological character displacement can push food webs toward the boundary of local stability (i.e.
+where the system transitions from having a stable equilibrium point to a limit cycle).
 
@@
 
-The scenario where the relationship between character divergence and resource suppression is the strongest is when the tradeoff in attack rates is concave up and consumers are competing for resources that occur in distinct habitats.
+The scenario where the relationship between character displacement and resource suppression is the strongest is when the tradeoff in attack rates is concave-up and consumers are competing for resources that occur in distinct habitats (dashed lines in \ref{fig:LS_Resources} and \ref{fig:McCann_Resources}).
 Prior work in threespine stickleback suggests that the evolution of stickleback attack rates are constrained by a concave-up tradeoff (Schluter 1993; Arnegard et al.
 2014) and that stickleback ecotypes are competing for resources that occur in distinct habitats (benthic vs.
-limnetic).
-Thus, we predict that character divergence in sticklebacks would supress resource densities.
+limnetic; Schluter and McPhail 1992).
+Thus, we predict that ecological character displacement in sticklebacks would supress resource densities and decrease local stability.
 
-## Effect of character divergence on food-web stability
+While measuring resource densities is relatively straightforward, measuring stability is notoriously difficult.
+One of the more easy to acquire empirical metrics of stability is the coefficient of variation ($CV=\frac{\sigma}{\mu}$) as this only requires measurements of resource (or consumer) densities over time.
+Interestingly, there is a close correspondence between the coefficient of variation and the local stability of a food web (Gellner et al.
+2016).
+This close correspondence emerges from the fact that stochastic perturbations in resource (or consumer) abundances will determine the variability in abundances.
+And since food webs are constantly buffeted by stochastic processes in the real world, we expect that our predictions using local stability would correspond to the $CV$ of resources (or consumers).
+Thus, we predict that character displacement in threespine stickleback will increase the $CV$ of resource (or consumer) densities.
 
-This relationship between character divergence and resource abundances begs the question: how does character divergence affect food-web stability?
-This question has been addressed empirically by Pimentel as well as with a conceptual model.
-Lawlor and Smith though did a detailed theoretical analysis of how character divergence affects three different aspects of stability.
-They looked at stability through the lens of competition.
-The first aspect was whether character divergence resulted in mutual invasibility of consumers, a common metric of "stable coexistence" in competition models.
-Another very interesting aspect they looked at was a global metric that examined the persistence of the system to perturbations in resources.
-Here, they conclude the divergence increases community stability, because it expands the range of resource perturbations that a system can withstand before a consumer going extinct.
-Note though, that they overlooked that there conclusions do not actually quantify the effects of ecological character displacement on stability because they do not compare scenarios where a consumer is in allopatry and one where consumers are in sympatry.
-Finally, they also look at the Lyapunov exponent (local stability), and while they find a negative effect of divergence, they suggest that it would likely be overshadowed by the stabilizing effects of divergence on mutual invasibility and global stability.
+Consumer and resource densities in our model can be interpreted from either a population or biomass perspective (Yodzis and Innes 1992; Murdoch et al.
+2003; McCann 2012).
+For the stickleback system, we had data on seasonal dynamics of resources in terms of both abundances and biomass.
+Within a season, we do not expect stickleback abundances to be coupled to the abundances of zooplankton and benthic invertebrates, given that sticklebacks reproduce annually whereas zooplankton and benthic invertebrates can reproduces many times within a season.
+Instead, we expect that stickleback biomass will be coupled to the biomass of zooplankton and benthic invertebrates within a season.
+This is because there will be lots of young stickleback at the beginning of the season, and they will eat a lot of resources and grow in size as the season progresses.
+If this is true, we also expect to see the largest negative effect of stickleback on resource densities late in the season.
 
-Here we focus on local stability, which makes the most sense in a food-web context.
-We find that divergence follows the typical "check-mark" stability pattern that has been described recently in theoretical food-web models.
+### Other relevant points
 
-Importantly, this metric of stability allows us to compare the stability of three-species vs.
-four-species system, which is key for understanding how ecological character displacement affects food-web dynamics.
-
-### Conclusions
-
-Although we used an Adaptive Dynamics approach here, we expect that our results would apply to different models of the genetic architecture of attack rates (e.g.
-quantitative genetics, Taper & Case 1985), given that character divergence is an inevitable outcome of resource competition, regardless of the model of genetic architecture, when resources dynamics are explicitly modelled (Taper & Case 1985).
-
-These conclusions are only for substitutable resources and that exhibit biological dynamics.
+Although we used an Adaptive Dynamics approach here, we expect that our results would apply to models that assumed a different genetic architecture of attack rates (e.g.
+quantitative genetics, Taper & Case 1985).
+We expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent charactere displacement, regardless of the genetic architecture of traits (see synthesis by Taper & Case 1985).
+Finally, our conclusions only apply to biological resources that are nutritionally substitutable.
+It would be interesting to extend our current analyses to those situations.

### 2018-05-20 eco_evo_sim_function.R

@@ -16,7 +16,7 @@ eco_evo_CR <- function(init_parameters,              # data frame of initial par
                        mut.suc.thres=1e-4,           # threshold of mutant success rate for identifying ESS.
                        j.step=500,                   # number of evolutionary time steps over which to assess evolutionary change.
                        abund.thres=1e-4,             # abundance threshold for feasible equilibriums
-                       mut.reps=10000) {             # number of mutation iterations to try for simulation. Note that this is the upper limit of the simulation length. Usually the traits will hit their ESS or a constraint on the parameter space before this limit is reached.   
+                       mut.reps=1000) {             # number of mutation iterations to try for simulation. Note that this is the upper limit of the simulation length. Usually the traits will hit their ESS or a constraint on the parameter space before this limit is reached.   
                             
   #### SETUP ####
   

### 2018-05-20 eco_evo_simulations.R

@@ -36,7 +36,7 @@ m2 <- m
 general_parameters <- data.frame(r1 = r1, r2 = r2, K1 = K1, K2 = K2, e11 = e11, e12 = e12, e21 = e21, e22 = e22, m1 = m1, m2 = m2)
 
 ## Attack rate
-A <- 2 
+A <- 2
 aii <- A*0.6
 
 ## Habitat preference (for LawlorSmith and McCann models)
@@ -417,68 +417,68 @@ sim_1C_2R_McCann_concaveUP <- eco_evo_CR(init_parameters = params_1C_2R_McCann_c
 
 ## Gather data. Note different formula for effective attack rate of consumer 1 "a_eff_C1". This accounts for the differences between fine- and coarse-grained environments
 sim_data <- bind_rows(
-  mutate(sim_2C_2R_MacArthur_linear, Model = "MacArthur", Trade_off = "Linear", Competitor = "Yes", a_eff_C1 = a11+a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_MacArthur_linear, Model = "MacArthur", Trade_off = "Linear", Competitor = "No", a_eff_C1 = a11+a12, Ctot = C1),
-  mutate(sim_2C_2R_MacArthur_concaveDN, Model = "MacArthur", Trade_off = "concaveDN", Competitor = "Yes", a_eff_C1 = a11+a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_MacArthur_concaveDN, Model = "MacArthur", Trade_off = "concaveDN", Competitor = "No", a_eff_C1 = a11+a12, Ctot = C1),
-  mutate(sim_2C_2R_MacArthur_concaveUP, Model = "MacArthur", Trade_off = "concaveUP", Competitor = "Yes", a_eff_C1 = a11+a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_MacArthur_concaveUP, Model = "MacArthur", Trade_off = "concaveUP", Competitor = "No", a_eff_C1 = a11+a12, Ctot = C1),
-  mutate(sim_2C_2R_LawlorSmith_linear, Model = "LawlorSmith", Trade_off = "Linear", Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
[…]

### 2018-05-21 ECD_Supp_Mat.Rmd
near final draft of ECD_Supp_Mat

--- 2018-05-20 ECD_Supp_Mat.Rmd
@@
 ---
-title: "How does character divergence affect resource dynamics?"
-output: html_document
-author: Matt Barbour
+title: "Predicting the effects of character displacement on food-web dynamics"
+author: "Matt Barbour"
+output:
+  pdf_document:
+    df_print: paged
+    fig_caption: true
+    citation_package: natbib
+    includes:  
+      in_header: preamble-latex.tex
+bibliography: references_ECD_supp_mat.bib
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	message = FALSE,
-	warning = FALSE,
-	include = FALSE
-)
-source('eco_evo_simulations.R')
+knitr::opts_chunk$set(warning = FALSE)
+
+## Load required libraries
+library(deSolve) # numerical integration library
+library(rootSolve) # for runsteady integration function
+library(pse) # for latin hypercube
+library(plyr) # for ldply
+library(dplyr) # for manipulating data
+library(tidyr) # for tidying data
+library(cowplot) # for better base, ggplot graphics
+
+## color palette
+cbPalette <- c("#D55E00", "#0072B2", "#009E73")
 ```
 
-### Model of resource competition
+```{r Simulations, include=FALSE, messages=TRUE, cache=TRUE}
 
-We analyzed a continuous-time model of consumer-resource dynamics for two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
+#### GET REQUIRED FUNCTIONS ####
+
+source('identify_steady_state.R')
+source('eco_evo_sim_function.R')
+source('Consumer_Resource_Functions.R')
+source('CR_dynamics.R')
+
+#### GENERAL PARAMETERS AND STATE VARIABLES ####
+
+## Intrinsic growth rates
+r <- 1
+r1 <- r 
+r2 <- r
+
+## Carrying capacity
+K <- 4
+K1 <- K
+K2 <- K
+
+## Conversion efficiency
+e <- 0.8
+e11 <- e 
+e12 <- e 
+e21 <- e 
+e22 <- e
+
+## Mortality rate
+m <- 1
+m1 <- m
+m2 <- m
+
+## General, non-evolving parameter set for 2C,2R model
+general_parameters <- data.frame(r1 = r1, r2 = r2, K1 = K1, K2 = K2, e11 = e11, e12 = e12, e21 = e21, e22 = e22, m1 = m1, m2 = m2)
+
+## Attack rate
+A <- 2
+aii <- A*0.6
+
+## Habitat preference (for LawlorSmith and McCann models)
+w <- 0.6
+wii <- w
+wij <- (1 - w)
+
+## Handling time (only for McCann model)
+h <- 0.4
+h11 <- h
+h12 <- h
+h21 <- h
+h22 <- h
+
+## State variables
+R <- 2
+R1 <- R
+R2 <- R
+C <- 1
+C1 <- C
+C2 <- C
+
+states_2C_2R <- data.frame(R1 = R1, R2 = R2, C1 = C1, C2 = C2)
+states_1C_2R <- data.frame(R1 = R1, R2 = R2, C1 = C1)
+
+## Trade-offs. Borrowed from Sargent and Otto, 2006 Am Nat
+a12_tradeoff <- expression(A*(1-(a11m/A)^n)^(1/n)) 
+a21_tradeoff <- expression(A*(1-(a22m/A)^n)^(1/n))
+
+n_concaveDN <- 1.15
+n_linear <- 1
+n_concaveUP <- 0.85
+
+concaveDN_df <- data.frame(n = n_concaveDN, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_concaveDN)^(1/n_concaveDN))
+linear_df <- data.frame(n = n_linear, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_linear)^(1/n_linear))
+concaveUP_df <- data.frame(n = n_concaveUP, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_concaveUP)^(1/n_concaveUP))
+
+bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
+  ggplot(., aes(x=aii, y=aij, color = factor(n))) + geom_line() + ggtitle("Attack rate trade-offs")
+
+bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + geom_line() + ggtitle("Relationship between specialization and total attack rate")
+
+######################################### MACARTHUR MODEL #######################################################
+
+## Mutant growth rates
+mC1_MacArthur <- expression(e11*a11m*R1 + e12*a12m*R2 - m1)
+mC2_MacArthur <- expression(e21*a21m*R1 + e22*a22m*R2 - m2)
+
+#### PARAMETERS ####
+
+## Linear
+params_2C_2R_MacArthur_linear <- data.frame(general_parameters,
+                                            a11 = aii, 
+                                            a12 = eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_linear)),                     
+                                            a21 = eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_linear)), 
+                                            a22 = aii)                      
+params_1C_2R_MacArthur_linear <- params_2C_2R_MacArthur_linear[c("r1","r2","K1","K2","e11","e12","a11","a12","m1")]
+
+## concaveDN
+params_2C_2R_MacArthur_concaveDN <- params_2C_2R_MacArthur_linear
+params_2C_2R_MacArthur_concaveDN["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveDN))
+params_2C_2R_MacArthur_concaveDN["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveDN))
+params_1C_2R_MacArthur_concaveDN <- params_2C_2R_MacArthur_concaveDN[c("r1","r2","K1","K2","e11","e12","a11","a12","m1")]
+
+## concaveUP
+params_2C_2R_MacArthur_concaveUP <- params_2C_2R_MacArthur_linear
+params_2C_2R_MacArthur_concaveUP["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveUP))
+params_2C_2R_MacArthur_concaveUP["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveUP))
+params_1C_2R_MacArthur_concaveUP <- params_2C_2R_MacArthur_concaveUP[c("r1","r2","K1","K2","e11","e12","a11","a12","m1")]
+
+#### SIMULATIONS ####
+
+## Linear
+sim_2C_2R_MacArthur_linear <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_linear, 
+                                init_states = states_2C_2R, 
+                                eco_CR_model = MacArthur_2C_2R, 
+                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_linear))
+sim_1C_2R_MacArthur_linear <- eco_evo_CR(init_parameters = params_1C_2R_MacArthur_linear, 
+                                init_states = states_1C_2R, 
+                                eco_CR_model = MacArthur_1C_2R, 
+                                AD_CR_models = c(mC1_MacArthur),
+                                species_traits = c(1), 
+                                eco_param_names = c("a11"), 
+                                eco_tradeoff_names = c("a12"),  
+                                evo_param_names = c("a11m"),  
+                                evo_tradeoff_names = c("a12m"), 
+                                tradeoff_functions = c(a12_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_linear))
+
+## concaveDN
+sim_2C_2R_MacArthur_concaveDN <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_concaveDN, 
+                                init_states = states_2C_2R, 
+                                eco_CR_model = MacArthur_2C_2R, 
+                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
+sim_1C_2R_MacArthur_concaveDN <- eco_evo_CR(init_parameters = params_1C_2R_MacArthur_concaveDN, 
+                                init_states = states_1C_2R, 
+                                eco_CR_model = MacArthur_1C_2R, 
+                                AD_CR_models = c(mC1_MacArthur),
+                                species_traits = c(1), 
+                                eco_param_names = c("a11"), 
+                                eco_tradeoff_names = c("a12"),  
+                                evo_param_names = c("a11m"),  
+                                evo_tradeoff_names = c("a12m"), 
+                                tradeoff_functions = c(a12_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
+## concaveUP
+sim_2C_2R_MacArthur_concaveUP <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_concaveUP, 
+                                       init_states = states_2C_2R, 
+                                       eco_CR_model = MacArthur_2C_2R, 
+                                       AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
+                                       species_traits = c(1,2), 
+                                       eco_param_names = c("a11","a22"), 
+                                       eco_tradeoff_names = c("a12","a21"),  
+                                       evo_param_names = c("a11m","a22m"),  
+                                       evo_tradeoff_names = c("a12m","a21m"), 
+                                       tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                       extra_tradeoff_params = c(A=A, n=n_concaveUP))
+sim_1C_2R_MacArthur_concaveUP <- eco_evo_CR(init_parameters = params_1C_2R_MacArthur_concaveUP, 
+                                       init_states = states_1C_2R, 
+                                       eco_CR_model = MacArthur_1C_2R, 
+                                       AD_CR_models = c(mC1_MacArthur),
+                                       species_traits = c(1), 
+                                       eco_param_names = c("a11"), 
+                                       eco_tradeoff_names = c("a12"),  
+                                       evo_param_names = c("a11m"),  
+                                       evo_tradeoff_names = c("a12m"), 
+                                       tradeoff_functions = c(a12_tradeoff),
+                                       extra_tradeoff_params = c(A=A, n=n_concaveUP))
+
+######################################### LAWLOR SMITH COARSE-GRAIN MODEL #######################################################
+
+## Mutant growth rates
+mC1_LawlorSmith <- expression(e11*w11*a11m*R1 + e12*w12*a12m*R2 - m1)
+mC2_LawlorSmith <- expression(e21*w21*a21m*R1 + e22*w22*a22m*R2 - m2)
+
+#### PARAMETERS ####
+
+## Linear
+params_2C_2R_LawlorSmith_linear <- data.frame(general_parameters,
+                                              a11 = aii,
+                                              a12 = eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_linear)), 
+                                              a21 = eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_linear)), 
+                                              a22 = aii,
+                                              w11 = wii, 
+                                              w12 = wij, 
+                                              w21 = wij,
+                                              w22 = wii)
+params_1C_2R_LawlorSmith_linear <- params_2C_2R_LawlorSmith_linear[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22")]
+
+## concaveDN
+params_2C_2R_LawlorSmith_concaveDN <- params_2C_2R_LawlorSmith_linear
+params_2C_2R_LawlorSmith_concaveDN["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveDN))
+params_2C_2R_LawlorSmith_concaveDN["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveDN))
+params_1C_2R_LawlorSmith_concaveDN <- params_2C_2R_LawlorSmith_concaveDN[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22")]
+
+## concaveUP
+params_2C_2R_LawlorSmith_concaveUP <- params_2C_2R_LawlorSmith_linear
+params_2C_2R_LawlorSmith_concaveUP["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveUP))
+params_2C_2R_LawlorSmith_concaveUP["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveUP))
+params_1C_2R_LawlorSmith_concaveUP <- params_2C_2R_LawlorSmith_concaveUP[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22")]
+
+#### SIMULATIONS ####
+
+## Linear
+sim_2C_2R_LawlorSmith_linear <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_linear, 
+                                       init_states = states_2C_2R, 
+                                       eco_CR_model = LawlorSmith_2C_2R, 
+                                       AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
+                                       species_traits = c(1,2), 
+                                       eco_param_names = c("a11","a22"), 
+                                       eco_tradeoff_names = c("a12","a21"),  
+                                       evo_param_names = c("a11m","a22m"),  
+                                       evo_tradeoff_names = c("a12m","a21m"), 
+                                       tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                       extra_tradeoff_params = c(A=A, n=n_linear))
+sim_1C_2R_LawlorSmith_linear <- eco_evo_CR(init_parameters = params_1C_2R_LawlorSmith_linear, 
+                                       init_states = states_1C_2R, 
+                                       eco_CR_model = LawlorSmith_1C_2R, 
+                                       AD_CR_models = c(mC1_LawlorSmith),
+                                       species_traits = c(1), 
+                                       eco_param_names = c("a11"), 
+                                       eco_tradeoff_names = c("a12"),  
+                                       evo_param_names = c("a11m"),  
+                                       evo_tradeoff_names = c("a12m"), 
+                                       tradeoff_functions = c(a12_tradeoff),
+                                       extra_tradeoff_params = c(A=A, n=n_linear))
+
+## concaveDN
+# max mut iter reach!
+sim_2C_2R_LawlorSmith_concaveDN <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveDN, 
+                                       init_states = states_2C_2R, 
+                                       eco_CR_model = LawlorSmith_2C_2R, 
+                                       AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
+                                       species_traits = c(1,2), 
+                                       eco_param_names = c("a11","a22"), 
+                                       eco_tradeoff_names = c("a12","a21"),  
+                                       evo_param_names = c("a11m","a22m"),  
+                                       evo_tradeoff_names = c("a12m","a21m"), 
+                                       tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                       extra_tradeoff_params = c(A=A, n=n_concaveDN))
+# note that although the max mutation iterations are reached, the evolving parameters appear to oscillate near high values, i.e. they are effectively at an ESS
+plot(sim_2C_2R_LawlorSmith_concaveDN[,c("a11","a22")]) 
+plot(sim_2C_2R_LawlorSmith_concaveDN[,c("a11","a22")], xlim=c(1.85,1.95))
+
+sim_1C_2R_LawlorSmith_concaveDN <- eco_evo_CR(init_parameters = params_1C_2R_LawlorSmith_concaveDN, 
+                                       init_states = states_1C_2R, 
+                                       eco_CR_model = LawlorSmith_1C_2R, 
+                                       AD_CR_models = c(mC1_LawlorSmith),
+                                       species_traits = c(1), 
+                                       eco_param_names = c("a11"), 
+                                       eco_tradeoff_names = c("a12"),  
+                                       evo_param_names = c("a11m"),  
+                                       evo_tradeoff_names = c("a12m"), 
+                                       tradeoff_functions = c(a12_tradeoff),
+                                       extra_tradeoff_params = c(A=A, n=n_concaveDN))
+## concaveUP
+sim_2C_2R_LawlorSmith_concaveUP <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveUP, 
+                                        init_states = states_2C_2R, 
+                                        eco_CR_model = LawlorSmith_2C_2R, 
+                                        AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
+                                        species_traits = c(1,2), 
+                                        eco_param_names = c("a11","a22"), 
+                                        eco_tradeoff_names = c("a12","a21"),  
+                                        evo_param_names = c("a11m","a22m"),  
+                                        evo_tradeoff_names = c("a12m","a21m"), 
+                                        tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                        extra_tradeoff_params = c(A=A, n=n_concaveUP))
+sim_1C_2R_LawlorSmith_concaveUP <- eco_evo_CR(init_parameters = params_1C_2R_LawlorSmith_concaveUP, 
+                                        init_states = states_1C_2R, 
+                                        eco_CR_model = LawlorSmith_1C_2R, 
+                                        AD_CR_models = c(mC1_LawlorSmith),
+                                        species_traits = c(1), 
+                                        eco_param_names = c("a11"), 
+                                        eco_tradeoff_names = c("a12"),  
+                                        evo_param_names = c("a11m"),  
+                                        evo_tradeoff_names = c("a12m"), 
+                                        tradeoff_functions = c(a12_tradeoff),
+                                        extra_tradeoff_params = c(A=A, n=n_concaveUP))
+
+######################################### MCCANN MODEL #######################################################
+
+## Mutant growth rates
+mC1_McCann <- expression(e11*(((w11 * R1)/(w11 * R1 + w12 * R2)) * a11m * R1)/(1 + ((w11 * R1)/(w11 * R1 + w12 * R2)) * a11m * h11 * R1 + ((w12 * R2)/(w11 * R1 + w12 * R2)) * a12m * h12 * R2) + e12*(((w12 * R2)/(w11 * R1 + w12 * R2)) * a12m * R2)/(1 + ((w11 * R1)/(w11 * R1 + w12 * R2)) * a11m * h11 * R1 + ((w12 * R2)/(w11 * R1 + w12 * R2)) * a12m * h12 * R2) - m1)
+mC2_McCann <- expression(e21*(((w21 * R1)/(w21 * R1 + w22 * R2)) * a21m * R1)/(1 + ((w21 * R1)/(w21 * R1 + w22 * R2)) * a21m * h21 * R1 + ((w22 * R2)/(w21 * R1 + w22 * R2)) * a22m * h22 * R2) + e22*(((w22 * R2)/(w21 * R1 + w22 * R2)) * a22m * R2)/(1 + ((w21 * R1)/(w21 * R1 + w22 * R2)) * a21m * h21 * R1 + ((w22 * R2)/(w21 * R1 + w22 * R2)) * a22m * h22 * R2) - m2)
+
+#### PARAMETERS ####
+
+## Linear
+params_2C_2R_McCann_linear <- data.frame(general_parameters,
+                                         a11 = aii,
+                                         a12 = eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_linear)), 
+                                         a21 = eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_linear)), 
+                                         a22 = aii,
+                                         w11 = wii, 
+                                         w12 = wij, 
+                                         w21 = wij,
+                                         w22 = wii,
+                                         h11 = h11,
+                                         h12 = h12,
+                                         h21 = h21,
+                                         h22 = h22)
+
+params_1C_2R_McCann_linear <- params_2C_2R_McCann_linear[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22","h11","h12","h21","h22")]
+
+## concaveDN
+params_2C_2R_McCann_concaveDN <- params_2C_2R_McCann_linear
+params_2C_2R_McCann_concaveDN["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveDN))
+params_2C_2R_McCann_concaveDN["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveDN))
+params_1C_2R_McCann_concaveDN <- params_2C_2R_McCann_concaveDN[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22","h11","h12","h21","h22")]
+
+## concaveUP
+params_2C_2R_McCann_concaveUP <- params_2C_2R_McCann_linear
+params_2C_2R_McCann_concaveUP["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveUP))
+params_2C_2R_McCann_concaveUP["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveUP))
+params_1C_2R_McCann_concaveUP <- params_2C_2R_McCann_concaveUP[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22","h11","h12","h21","h22")]
+
+#### SIMULATIONS ####
+
+## Linear
+sim_2C_2R_McCann_linear <- eco_evo_CR(init_parameters = params_2C_2R_McCann_linear, 
+                                         init_states = states_2C_2R, 
+                                         eco_CR_model = McCann_2C_2R, 
+                                         AD_CR_models = c(mC1_McCann, mC2_McCann),
+                                         species_traits = c(1,2), 
+                                         eco_param_names = c("a11","a22"), 
+                                         eco_tradeoff_names = c("a12","a21"),  
+                                         evo_param_names = c("a11m","a22m"),  
+                                         evo_tradeoff_names = c("a12m","a21m"), 
+                                         tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                         extra_tradeoff_params = c(A=A, n=n_linear))
+sim_1C_2R_McCann_linear <- eco_evo_CR(init_parameters = params_1C_2R_McCann_linear, 
+                                         init_states = states_1C_2R, 
+                                         eco_CR_model = McCann_1C_2R, 
+                                         AD_CR_models = c(mC1_McCann),
+                                         species_traits = c(1), 
+                                         eco_param_names = c("a11"), 
+                                         eco_tradeoff_names = c("a12"),  
+                                         evo_param_names = c("a11m"),  
+                                         evo_tradeoff_names = c("a12m"), 
+                                         tradeoff_functions = c(a12_tradeoff),
+                                         extra_tradeoff_params = c(A=A, n=n_linear))
+
+## concaveDN
+# max mut iter
+sim_2C_2R_McCann_concaveDN <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveDN, 
+                                         init_states = states_2C_2R, 
+                                         eco_CR_model = McCann_2C_2R, 
+                                         AD_CR_models = c(mC1_McCann, mC2_McCann),
+                                         species_traits = c(1,2), 
+                                         eco_param_names = c("a11","a22"), 
+                                         eco_tradeoff_names = c("a12","a21"),  
+                                         evo_param_names = c("a11m","a22m"),  
+                                         evo_tradeoff_names = c("a12m","a21m"), 
+                                         tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                         extra_tradeoff_params = c(A=A, n=n_concaveDN))
+# note that although the max mutation iterations are reached, the evolving parameters appear to oscillate near high values. i.e. they are effectively at an ESS
+plot(sim_2C_2R_McCann_concaveDN[,c("a11","a22")]) 
+plot(sim_2C_2R_McCann_concaveDN[,c("a11","a22")], xlim=c(1.85,1.95))
+
+sim_1C_2R_McCann_concaveDN <- eco_evo_CR(init_parameters = params_1C_2R_McCann_concaveDN, 
+                                         init_states = states_1C_2R, 
+                                         eco_CR_model = McCann_1C_2R, 
+                                         AD_CR_models = c(mC1_McCann),
+                                         species_traits = c(1), 
+                                         eco_param_names = c("a11"), 
+                                         eco_tradeoff_names = c("a12"),  
+                                         evo_param_names = c("a11m"),  
+                                         evo_tradeoff_names = c("a12m"), 
+                                         tradeoff_functions = c(a12_tradeoff),
+                                         extra_tradeoff_params = c(A=A, n=n_concaveDN))
+## concaveUP
+sim_2C_2R_McCann_concaveUP <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveUP, 
+                                          init_states = states_2C_2R, 
+                                          eco_CR_model = McCann_2C_2R, 
+                                          AD_CR_models = c(mC1_McCann, mC2_McCann),
+                                          species_traits = c(1,2), 
+                                          eco_param_names = c("a11","a22"), 
+                                          eco_tradeoff_names = c("a12","a21"),  
+                                          evo_param_names = c("a11m","a22m"),  
+                                          evo_tradeoff_names = c("a12m","a21m"), 
+                                          tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                          extra_tradeoff_params = c(A=A, n=n_concaveUP))
+sim_1C_2R_McCann_concaveUP <- eco_evo_CR(init_parameters = params_1C_2R_McCann_concaveUP, 
+                                          init_states = states_1C_2R, 
+                                          eco_CR_model = McCann_1C_2R, 
+                                          AD_CR_models = c(mC1_McCann),
+                                          species_traits = c(1), 
+                                          eco_param_names = c("a11"), 
+                                          eco_tradeoff_names = c("a12"),  
+                                          evo_param_names = c("a11m"),  
+                                          evo_tradeoff_names = c("a12m"), 
+                                          tradeoff_functions = c(a12_tradeoff),
+                                          extra_tradeoff_params = c(A=A, n=n_concaveUP))
+```
+
+```{r Organize_Simulation_Data, include=FALSE, cache=TRUE}
+
+## Gather data. Note different formula for effective attack rate of consumer 1 "a_eff_C1". This accounts for the differences between fine- and coarse-grained environments
+sim_data <- bind_rows(
+  mutate(sim_2C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "Yes", a_eff_C1 = a11+a12, Ctot = C1+C2), 
+  mutate(sim_1C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "No", a_eff_C1 = a11+a12, Ctot = C1),
+  mutate(sim_2C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "Yes", a_eff_C1 = a11+a12, Ctot = C1+C2), 
+  mutate(sim_1C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "No", a_eff_C1 = a11+a12, Ctot = C1),
+  mutate(sim_2C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "Yes", a_eff_C1 = a11+a12, Ctot = C1+C2), 
+  mutate(sim_1C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "No", a_eff_C1 = a11+a12, Ctot = C1),
+  mutate(sim_2C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
+  mutate(sim_1C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
+  mutate(sim_2C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
+  mutate(sim_1C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
+  mutate(sim_2C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
+  mutate(sim_1C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
+  mutate(sim_2C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
+  mutate(sim_1C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
+  mutate(sim_2C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
+  mutate(sim_1C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
+  mutate(sim_2C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
+  mutate(sim_1C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1))
+
+#### PREDICTING EFFECTS OF CHARACTER DISPLACEMENT ON C-R DYNAMICS ####
+
+tradeoff_df <- bind_rows(mutate(concaveDN_df, n = n_concaveDN), 
+                         mutate(linear_df, n = n_linear), 
+                         mutate(concaveUP_df, n = n_concaveUP)) %>%
+  data.frame(., general_parameters) %>%
+  mutate(a11 = aii, a12 = aij, a21 = aij, a22 = aii,
+         w11 = wii, w12 = wij, w21 = wij, w22 = wii)
+
+## MacArthur
+MacArthur_displacements <- bind_rows(
+  mutate(sim_2C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_linear)), ], 
+  mutate(sim_1C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_linear)), ],
+  mutate(sim_2C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveDN)), ], 
+  mutate(sim_1C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveDN)), ],
+  mutate(sim_2C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveUP)), ], 
+  mutate(sim_1C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveUP)), ]) 
+MacArthur_displacements$Time <- rep(c("Begin","End"),6)
+
+MacArthur_2C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:a22), init_states = states_2C_2R, eco_CR_model = MacArthur_2C_2R)
+MacArthur_1C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:a12), init_states = states_1C_2R, eco_CR_model = MacArthur_1C_2R)
+
+## Lawlor Smith
+LawlorSmith_displacements <- bind_rows(
+  mutate(sim_2C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_linear)), ], 
+  mutate(sim_1C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_linear)), ],
+  mutate(sim_2C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_concaveDN)), ], 
+  mutate(sim_1C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_concaveDN)), ],
+  mutate(sim_2C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_concaveUP)), ], 
+  mutate(sim_1C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_concaveUP)), ]) 
+LawlorSmith_displacements$Time <- rep(c("Begin","End"),6)
+
+LawlorSmith_2C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:w22), init_states = states_2C_2R, eco_CR_model = LawlorSmith_2C_2R)
+LawlorSmith_1C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:a12, w11, w12), init_states = states_1C_2R, eco_CR_model = LawlorSmith_1C_2R)
+
+## McCann
+McCann_displacements <- bind_rows(
+  mutate(sim_2C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_linear)), ], 
+  mutate(sim_1C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_linear)), ],
+  mutate(sim_2C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_concaveDN)), ], 
+  mutate(sim_1C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_concaveDN)), ],
+  mutate(sim_2C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_concaveUP)), ], 
+  mutate(sim_1C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_concaveUP)), ]) 
+McCann_displacements$Time <- rep(c("Begin","End"),6)
+
+McCann_2C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:w22), init_states = states_2C_2R, eco_CR_model = McCann_2C_2R)
+McCann_1C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:a12, w11, w12), init_states = states_1C_2R, eco_CR_model = McCann_1C_2R)
+
+```
+
+## Model of resource competition
+
+I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
 
@@
 $$\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2$$
-where $r_i$ represents intrinsic growth rates, $K_i$ represents carrying capacity, $e_{ij}$ represents the conversion efficiency of resources into new consumers, and $m_i$ represents the mortality rate of consumers.
-$F_{ij}(R_j)$ represents the consumer's feeding rate (or functional response) on a specific resource (in this case, $C_i$ on $R_j$).
+
+where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ji}$ represents the conversion efficiency of resource $i$ into new consumer $j$, and $m_j$ represents the mortality rate of consumer $j$.
+$F_{ji}(R_i)$ represents consumer $j$'s feeding rate (i.e.
+functional response) on resource $i$.
 This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g.
 zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g.
-seed size, Taper and Case 1985).
-In the scenario where consumer feeding rates increase linearly with resource density as a function of their attack rate, such that $F_{ij}(R_j)=a_{ij}R_j$, then our model becomes MacArthur's model of resource competition (MacArthur 1972).
-The evolution of consumer attack rates in this model, and several extensions, have been analyzed in detail (Lawlor and Smith 1976; Abrams 1986).
-The general result of these analyses is that consumer divergence is an inevitable outcome of competition.
-To be explicit, this means that the specialization of evolved consumers ($\frac{a_{ii}{a_{ii}+a_{ij}}$) is greater in the presence of competitor compared to when no competitor is present.
+seed size, see @Taper1985 for an example).
+In the scenario where consumer feeding rates increase linearly with resource density, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$, then my model becomes MacArthur's model of resource competition (@MacArthur1972).
+The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail (@Lawlor1976; @Abrams1986).
+The general result of these analyses is that divergent character displacement in attack rates is an inevitable outcome of resource competition.
+<!-- In other words, the presence of a competitor causes consumers to evolve more specialized attack rates ($\frac{a_{ii}}{a_{ii}+a_{ij}}$) compared to when no competitor is present.
+-->
 
-Here, we examine the effects of this inevitable divergence in attack rates on food-web dynamics.
-In particular, we examine how divergence in attack rates affects resource abundances as well as the resilience of both resources and consumers to perturbations.
-We do this by (i) deriving analytical expressions for the relationship between attack rates and food-web dynamics and (ii) simulate the effects of competition on the eco-evolutionary dynamics of consumers and resources.
-For these simulations, we used an Adaptive Dynamics approach.
-At each evolutionary time step, we created a mutant consumer by randomly choosing a consumer and modifying its attack rate by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability.
-Assuming this mutant consumer was rare, we then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer.
-If the mutant was able to invade, we updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state.
-We then repeated the simulation 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy (ESS, Price & Maynard Smith 1973) or limit (e.g.
-$\frac{a_{ii}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
-Below is a table of the parameters used in these simulations.
-Our mathematical analyses show that the consequences of character displacement on food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
+Here, I examine the effects of this inevitable divergence in attack rates on food-web dynamics.
+In particular, I examine how divergence in attack rates affects resource abundances as well as the local stability of both resources and consumers to perturbations.
+I do this by (i) deriving analytical expressions for the relationship between attack rates and food-web dynamics in different foraging scenarios and (ii) simulating the effects of competition on the eco-evolutionary dynamics of consumers and resources.
+For these simulations, I used an Adaptive Dynamics approach.
+At each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate by either subtracting or adding a small constant <!-- (0.001 in the following simulations) -->with equal probability.
+Assuming this mutant consumer is rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer.
+If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state.
+I then repeated the simulation 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy (ESS, @Smith1973) or an evolutionary limit (e.g.
+$\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
+I show that the consequences of character displacement on food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
 
-## Effect of character divergence on resource abundances
+## Effect of character displacement on resource abundances
 
-We start by examing the effects of character displacement on the abundance of resources at equilibrium, which will also set the stage for a later analysis of food-web stability.
-In particular, we examine three different foraging scenarios: (i) consumers forage for resource simultaneously and their feeding rates increase linearly with resource density (MacArthur 1972); (ii) consumers *cannot* forage for resources simultaneously and their feeding rates increase linearly with resource density (Lawlor & Smith 1976); (iii) consumers *cannot* forage for resources simulaneously and their feeding rates on each resource depend on both resource densities and saturate at high densities (McCann 2005).
+I start by examing the effects of character displacement on the abundance of resources at equilibrium, which will also set the stage for a later analysis of food-web stability.
+In particular, I examine two different foraging scenarios: one where consumers forage for both resources simultaneously (e.g.
+@MacArthur1972) and one where they cannot because resources occur in distinct habitats (e.g.
+@Lawlor1976; @McCann2005).
+I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, defined as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
 
-### Consumers forage for resources simultaneously
+### Consumers forage for both resources simultaneously
 
-The first foragine scenario we explore is already assumed in MacArthur's model of resource competition, namely that consumers can forage for both resources simultaneously.
-In essence, this implies that both resources occur within the same microhabitat.
-To gain analytical insight to this scenario, we assume that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
-While this scenario is simple, it isolates the process of character displacement, rather than whether these effects depend on inherent differences in consumers or resources.
-In the appendix, we show that when both consumers and resources are present that the abundance of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation:
+An inherent assumption in MacArthur's model of resource competition is that consumers can forage for both resources simultaneously.
+In essence, this implies that both resources occur within the same habitat.
+To gain analytical insight to this scenario, I assume that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
+While this scenario is arguably simplistic, it isolates the effect of character displacement rather than any dependence on inherent differences in consumers or resources.
+I show that when both consumers and resources are present that the abundance of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation (*Mathematica* file with derivation available upon request):
 
@@
 Since the total attack rate is in the denominator, this implies that resource abundances will decrease if evolution favors an increase in total attack rate.
-We assume that consumers are not 'Darwinian demons' (Law 1979) and thus the evolution of their attack rates are subject to certain constraints.
-We modelled these constraints using the following general function:
+I assume that consumers are not 'Darwinian demons' (@Law1979) and thus the evolution of their attack rates are subject to certain constraints.
+I modelled these constraints such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff (@Sargent2006).
+This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$, green line in Fig.
+\ref{fig:Tradeoff}) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$, orange line Fig.
+\ref{fig:Tradeoff}).
+When $n=1$, the tradeoff function is linear, and all combinations of $a_ii$ and $a_ij$ have the same total attack rate (blue line in Fig.
+\ref{fig:Tradeoff}).
 
-$$\Big(\frac{a_{ii}}{A}\Big)^n+\Big(\frac{a_{ij}}{A}\Big)^n=1$$
-
-where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff.
-This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$; Sargent and Otto 2006).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_ii$ and $a_ij$ have the same total attack rate.
-
-```{r Tradeoff function, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
-  ggplot(., aes(x=aii, y=aij, linetype = factor(n))) + 
+```{r Tradeoff, echo=FALSE, fig.cap="\\label{fig:Tradeoff}Tradeoff forms for consumer attack rates."}
+tradeoff_plot <- bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
+  ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
   geom_line() + 
@@
   ylab(expression(a[ij])) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") 
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)
+
+tradeoff_plot
 ```
 
-We know from past work that consumers still undergo character divergence regardless of the shape of the tradeoff (Lawlor and Smith 1976; Abrams 1986).
-Interestingly, we find that the shape of the tradeoff qualitatively affects the relationship between character divergence and resource abundance.
-For example, if consumer's are constrained by a linear tradeoff ($n=1$), such that a unit increase in $a_{ii}$ results in a unit decrease in $a_{ij}$, then character divergence has no effect on resource abundances.
-If the tradeoff is concave down ($n>1$), then resource abundances can actually increase under character displacement!
-This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{aij}$) and decreases as consumers diverge.
-Note though that detecting this effect in nature will likely be difficult, since a concave down tradeoff results in relatively small displacement relative to other tradeoff forms (Lawlor and Smith 1976).
-When the tradeoff is concave up ($n<1$), however, we see that character divergence suppresses resource densities.
-Although this equilibrium equation for resource abundances was derived for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when there is only one consumer (white circle).
-This is because a consumer, in the absence of competition, evolves to be a generalist that has equal attack rates on each resource, resulting in equivalent resource abundances.
+Previous work has shown that consumers still undergo divergent character displacement regardless of the shape of the tradeoff (@Lawlor1976; @Abrams1986).
+Interestingly, I find that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource abundance (Fig.
+\ref{fig:MacArthur_Resources}).
+For example, if consumer's are constrained by a linear tradeoff ($n=1$), then there is no net change in total attack rate (blue points in Fig.
+\ref{fig:MacArthur_Resources}A) and character displacement has no effect on resource abundances (Fig.
+\ref{fig:MacArthur_Resources}B).
+If the tradeoff is concave down ($n>1$), then resource abundances can actually increase under character displacement (green points in Fig.
+\ref{fig:MacArthur_Resources}B).
+This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{aij}$) and decreases as consumers diverge (Fig.
+\ref{fig:MacArthur_Resources}A).
+Note that detecting this effect in nature would likely be difficult, since a concave-down tradeoff results in relatively small displacement relative to other tradeoff shapes (@Lawlor1976; green points in Fig.
+\ref{fig:MacArthur_Resources}A,B).
+When the tradeoff is concave up ($n<1$), character displacement suppresses resource densities due to the increase in total attack rates (orange points in Fig.
+\ref{fig:MacArthur_Resources}A,B).
+Although the equation I derived for resource abundance was for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
+\ref{fig:MacArthur_Resources}B).
+This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in Fig.
+\ref{fig:MacArthur_Resources}A), resulting in equivalent resource abundances.
 
-```{r Same resource habitat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, linetype = factor(n))) + 
+```{r MacArthur_Resources, echo=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers can forage for both resources simultaneously."}
+Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
   ylab(expression(a[ii]+a[ij])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
 Mac_resources <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
   ylab(expression(R[i]+R[j])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
-Mac_plots <- plot_grid(Mac_attack + theme(legend.position = "none"), Mac_resources + theme(legend.position = "none"), nrow=1)
-
-Mac_legend <- get_legend(Mac_attack)
-
-Mac_plot_legend <- plot_grid(Mac_plots, Mac_legend, ncol=2, rel_widths = c(3,0.3), align='hv')
-
-Mac_title <- ggdraw() + draw_label("Resources in same habitat", fontface='bold')
-plot_grid(Mac_title, Mac_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
+Mac_plots <- plot_grid(Mac_attack + theme(legend.position = "none"), Mac_resources, 
+                       ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
+Mac_plots
 ```
@@
 
-Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (Taper and Case 1985; Abrams 1986).
+Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (e.g.
+@Taper1985; @Abrams1986).
 This assumption may be valid for some foraging scenarios (e.g.
-Darwin's finches foraging for seeds); however, in many situations consumers forage for resources that occur in distinct habitats, and thus cannot forage for both simultaneosly.
-The only chracter displacement model that we are aware of that modeled resources in distinct habitats was one examined by Lawlor and Smith (1976).
-This model takes the same form as the previous model, except now the consumer's functional response takes the form:
+Darwin's finches foraging for seeds); however, in many situations consumers forage for resources that occur in different habitats, and thus cannot forage for both simultaneosly.
+The only character displacement model that I am aware of that modeled resources in different habitats was one examined by @Lawlor1976.
+This model takes the same form as the previous model, except now the consumer's feeding rate takes the form:
 
-$$F_{ij}(R_j)=w_{ij}a_{ij}R_j$$
+$$F_{ji}(R_i)=w_{ji}a_{ji}R_i$$
 
-where $w_{ij}$ represents the proportion of time a consumer ($C_i$) spends foraging in a habitat where a particular resource ($R_j$) is found (i.e.
+where $w_{ji}$ represents the proportion of time consumer $j$ spends foraging in a habitat where only resource $j$ is found (i.e.
 habitat preference).
-Note that since $w_ij$ is a proportion that $w_{ii}=1-w_{ij}$.
-As with attack rates, we assume that consumer habitat preferences are mirror images of each other ($w=w_{11}=w_[22]$).
+Note that since $w_ji$ is a proportion that $w_{jj}=1-w_{ji}$.
+As with attack rates, I assume that consumer habitat preferences are mirror images of each other ($w=w_{11}=w_{22}$).
 
-Lawlor and Smith (1976) found that consumers still underwent character divergence regardless of whether consumers could or could not forage for resources simultaneously.
-We find that the foraging scenario qualitatively affects the relationship between character divergence and resource abundances.
-To demonstrate this, note that now resource abundances at equilibrium (both consumers and resources present) are determined by the following equation (see derivation in Appendix):
+@Lawlor1976 found that consumers still underwent divergent character displacement regardless of whether consumers could or could not forage for resources simultaneously.
+I find that the foraging scenario qualitatively affects the relationship between character displacement and resource abundances.
+To demonstrate this, note that again resource abundances are equivalent at equilibrium ($R=R_1=R_2$ when both consumers and resources present), but are now determined by the following equation (*Mathematica* file with derivation available upon request):
 
@@
 This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g.
-$w>0.5$ and $a_{ii}>a_{ij}$), then resource abundances will always decrease, regardless of the tradeoff.
-Thus, character divergence always results in resource suppression if consumers are competing for resources that occur in distinct habitats.
+$w>0.5$ and $a_{ii}>a_{ij}$), then character displacement will always increase the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the tradeoff (Fig.
+\ref{fig:LS_Resources}A).
+Thus, character displacement always results in resource suppression if consumers are competing for resources that occur in different habitats (Fig.
+\ref{fig:LS_Resources}B).
+Note that the shape of the tradeoff can modify the effect of character displacement.
+This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached its ESS (triangles in Fig.
+\ref{fig:LS_Resources}B).
+In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
+\ref{fig:LS_Resources}B), because character displacement tends to reach a constraint of complete specialization.
 
-Note that the shape of the tradeoff can dampen or amplify the effect of character displacement.
-This is not because the tradeoff affects the magnitude of displacement, because as you can see, it has little effect (slightly less with concave down tradeoff).
-Its because for the concave up tradeoff, consumers have the lowest attack rate at intermediate values, but highest at extreme values, resulting in the largest change in resource abundances with character divergence.
+These results do lead to an interesting prediction that could be tested either experimentally or with field data.
+If consumers are competing for resources that occur in different habitats, then greater character displacement corresponds to a greater decrease in resource densities (line trajectories in Fig.
+\ref{fig:LS_Resources}B).
+This prediction likely only applies to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
 
-These results do lead to an interesting prediction that could be tested.
-If consumers are competing for resources that occur in distinct habitats, then greater ecological character displacement corresponds to a greater decrease in resource densities.
-Note that this prediction is only useful for comparing divergence within species.
-It cannot be used to predict across species where the nature of the trade-offs could be different.
+It is worth noting that in this foraging scenario, resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
+\ref{fig:LS_Resources}B).
+This is likely because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in Fig.
+\ref{fig:LS_Resources}B).
+<!-- I will revisit this subtle point when I examine the effects of character displacement on food-web stability.
+-->
 
-It is important to note that, in this foraging scenario, resource abundances in the absence of competition no longer correspond to the analytical expression we derived when both consumers are present.
-This is because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat.
-Note that our analytical expression consistently underestimates the effect of character displacement on resource dynamics.
-This implies that resource abundances are always higher than our analytical expression suggests when there is only one consumer.
-We will revisit this subtle point when we examine the effects of character displacement on food-web stability.
-
-```{r Distinct resource habitats, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, linetype = factor(n))) + 
+```{r LS_Resources, echo=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource densities (B) when consumers cannot forage for resources simultaneously."}
+LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
-  ylab(expression(w[ii]*a[ii]+(1-w[ii])*a[ij])) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
 LS_resources <- cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color=factor(n))) + 
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor),
-             shape=21, size=5, inherit.aes = FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
   ylab(expression(R[i]+R[j])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
-LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), LS_resources + theme(legend.position = "none"), nrow=1)
-
-LS_legend <- get_legend(LS_attack)
-
-LS_plot_legend <- plot_grid(LS_plots, LS_legend, nrow=1, rel_widths = c(3,0.3))
-
-LS_title <- ggdraw() + draw_label("Resources in different habitats", fontface='bold')
-plot_grid(LS_title, LS_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
+LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), LS_resources, 
+                      ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
+LS_plots
 ```
 
-This effect of the foraging scenario on the relationship between character divergence and resource abundances appears to be quite general.
-For example, even if we consider a more realistic consumer functional response (derived by McCann et al.
-2005):
+The effect of the foraging scenario on the relationship between character displacement and resource abundances appears to be quite general.
+For example, even if I consider a more realistic consumer functional response (derived by @McCann2005):
 
@@
 
-where consumer feeding rates on $R_i$ are influenced by both resource densities; saturate as resource densities increase (due to handling time $h$); and consumer habitat preferences are modified by relative resource densities, such that:
+where consumer feeding rates on resource $i$ are influenced by both resource densities; saturate as resource densities increase (due to handling time $h$); and consumer habitat preferences are modified by relative resource densities, such that:
 
@@
 
-we still observe the same qualitative relationship between character divergence and resource suppression.
-This is because resource competition always leads to character divergence (see proof in Appendix) and equilibrium resource abundances are governed by similar dynamics:
+I still observe the same qualitative relationship between character displacement and resource suppression (Fig.
+\ref{fig:McCann_Resources} in Appendix).
+This is because resource competition always leads to character divergence and equilibrium resource abundances are governed by similar dynamics  (*Mathematica* file with derivations available upon request):
 
@@
 
-This more realistic consumer functional response does actually make an interesting evolutionary prediction that differs from the scenario where consumer feeding rate scales linearly with resource density.
-Namely, that in the absence of competition, consumers remain slightly more specialized on the resources that occur in their preferred habitats, which actually opposes the prediction from the model where consumer feeding rate scales linearly with resource density.
+<!-- This more realistic consumer functional response does actually make an interesting evolutionary prediction that differs from the scenario where consumer feeding rate scales linearly with resource density.
+Specifically, a consumer's ESS in the absence of competition will remain slightly more specialized on the resources that occur in their preferred habitats.
+This is in contrast to the prediction when feeding rates scale linearly with resource density, where consumers evolve to be slightly more specialized on resources occurring in their non-preferred habitats.
+An important caveat for these predictions is that my model does not allow habitat preferences to evolve, which would likely alter these predictions.
+-->
 
-```{r}
-McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, linetype = factor(n))) + 
+## Effect of character displacement on food-web stability
+
+Of the prior models of character displacement, only @Lawlor1976 examine the effect of character displacement on the ecological stability of the system.
+They conducted a detailed theoretical analysis of three different metrics of stability: (i) stable coexistence of consumers, defined as the mutual invasibility of each consumer; (ii) global stability of consumers, defined as the continued coexistence of both consumers when resource dynamics are perturbed; and (iii) local stability, defined as the rate at which the four-species food web returned to equilibrium after a small perturbation to species abundances.
+Note that the first two metrics correspond to the stable coexistence of consumers, while the third metric measures the stability of the entire food web.
+@Lawlor1976 found that character displacement enhanced the stable coexistence of consumers, yet had a small negative effect on the local stability of the food web.
+
+@Lawlor1976's work gave fundamental insight to the effect of character displacement on ecological stability; however, none of their stability analyses compared food webs with and without a competing consumer.
+Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement (@Schluter2000; @Schluter1992).
+In fact, the first two metrics of stability can only be assessed when both consumers are present because it examine the effect of divergence on coexistence, which requires both consumers to be present.
+In contrast, local stability can be compared in food webs with and without a competing consumer, and thus provides a useful benchmark for examining the effects of character displacement on food-web stability.
+
+Below, I plot the results from the same simulations I did previously except now local stability ($-1\times\operatorname{Re}(\lambda_{max})$) is on the y-axis (Fig.
+\ref{fig:Stability}).
+In general, I find that resource supression goes hand-in-hand with local stability, a tendency that has been noted by others (@Murdoch2003; @McCann2011).
+Whenever character displacement results in resource suppression, I observe a decrease in local stability (Fig.
+\ref{fig:Stability}).
+This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (compare initial (small points) with evolved strategies (large points), Fig.
+\ref{fig:Stability}).
+Note that I reach the same fundamental conclusions when I incorporated a more realistic functional response for the consumers (Fig.
+\ref{fig:Stability_McCann}).
+
+```{r Stability, echo=FALSE, fig.cap="\\label{fig:Stability}Effect of character displacement on local stability when consumers can (A) and cannot (B) forage for both resources simultaneously.
+Small and large points correspond to initial and evolved strategies, respectively."}
+Stability_Mac <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
-  ylab(expression(w[ii]*a[ii]+(1-w[ii])*a[ij])) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend=FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_linetype_manual(values=1:3, name="Tradeoff (n)") +
-  scale_fill_manual(values=c("white","black"))
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_LS <- cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_plots <- plot_grid(Stability_Mac + theme(legend.position = "none"), Stability_LS + theme( axis.title.y = element_blank()),
+                             ncol = 2, labels = "AUTO", rel_widths = c(0.75,1))
+Stability_plots
+```
+
+I condcuted the previous simulations on a scenario where the food web remains locally stable ($\operatorname{Re}(\lambda_{max})<0$).
+Is it possible that character displacement would ever result in a locally unstable food web?
+To gain analytical insight to this question, I examined the conditions governing the local stability of the four-species food web when consumers exhibit a more realistic functional response (@McCann2005).
+I show that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (*Mathematica* file with derivation available upon request):
+
+$$wa_{ii}+(1-w)a_{ij} > \frac{e+hm}{hK(e-hm)}$$
+
+Previously, I showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff (Fig.
+\ref{fig:LS_Resources}A).
+Thus, character displacement is capable of pushing food webs to a point where they are locally unstable.
+
+## Predictions for threespine stickleback
+
+The scenario where the relationship between character displacement and resource suppression is the strongest is when the tradeoff in attack rates is concave-up and consumers are competing for resources that occur in different habitats (orange points in Fig.
+\ref{fig:LS_Resources}B and Fig.
+\ref{fig:McCann_Resources}B).
+Prior work in threespine stickleback suggests that the evolution of stickleback attack rates are constrained by a concave-up tradeoff (@Schluter1993; @Arnegard2014) and that stickleback ecotypes are competing for resources that occur in different habitats (benthic vs.
+limnetic, @Schluter1992).
+Thus, I predict that ecological character displacement in sticklebacks would suppress resource densities and decrease local stability.
+
+While measuring resource densities is relatively straightforward, measuring stability is notoriously difficult.
+One of the more easy to acquire empirical metrics of stability is the coefficient of variation ($CV=\frac{\sigma}{\mu}$) as this only requires measurements of resource (or consumer) densities over time.
+Interestingly, there is a close correspondence between the coefficient of variation and the local stability of a food web (@Gellner2016).
+<!-- This close correspondence emerges from the fact that stochastic perturbations in resource (or consumer) densities will determine the variability in abundances.
+And since food webs are constantly buffeted by stochastic processes in the real world, I expect that my predictions using local stability would correspond to the $CV$ of resources (or consumers).
+-->Thus, I predict that character displacement in threespine stickleback will increase the $CV$ of resource (or consumer) densities.
+
+Consumer and resource densities in these dynamical model can be interpreted from either a population or biomass perspective (@Yodzis1992; @Murdoch2003; @McCann2011).
+For the stickleback system, you have data on seasonal dynamics of resources in terms of both abundances and biomass.
+Within a season, I do not expect stickleback abundances to be coupled to the abundances of zooplankton and benthic invertebrates, given that stickleback reproduce annually whereas zooplankton and benthic invertebrates can reproduce many times within a season.
+Instead, I expect that stickleback biomass will be coupled to the biomass of zooplankton and benthic invertebrates within a season.
+This is because there will be lots of young stickleback at the beginning of the season, and they will eat a lot of resources and grow in size as the season progresses.
+If this is true, I also expect to see the largest negative effect of stickleback on resource densities late in the season.
+
+### Other relevant points
+
+Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
+quantitative genetics).
+I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits (see synthesis by @Taper1985).
+Finally, my conclusions only apply to biological resources that are nutritionally substitutable.
+It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected (@Abrams1987; @Fox2008).
+
+## Appendix: More realistic functional response
+
+```{r McCann_Resources, echo=FALSE, fig.cap="\\label{fig:McCann_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers exhibit a more realistic functional response."}
+McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
 McCann_resources <- cbind.data.frame(tradeoff_df, McCann_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, linetype = factor(n))) + 
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=Competitor), shape=21, size=5, inherit.aes = FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
   ylab(expression(R[i]+R[j])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_fill_manual(values=c("white","black"))
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
-McCann_plots <- plot_grid(McCann_attack + theme(legend.position = "none"), McCann_resources + theme(legend.position = "none"), nrow=1)
-
-McCann_legend <- get_legend(McCann_attack)
-
-McCann_plot_legend <- plot_grid(McCann_plots, McCann_legend, nrow=1, rel_widths = c(3,0.3))
-
-McCann_title <- ggdraw() + draw_label("Resources in different habitats & realistic functional response", fontface='bold')
-plot_grid(McCann_title, McCann_plot_legend, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
+McCann_plots <- plot_grid(McCann_attack + theme(legend.position = "none"), McCann_resources, 
+                          ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
+McCann_plots
 ```
 
-### Predictions for threespine stickleback
+```{r Stability_McCann, echo=FALSE, fig.cap="\\label{fig:Stability_McCann}Effect of character displacement on local stability when consumers exhibit a more realistic functional response."}
+Stability_McCann <- cbind.data.frame(tradeoff_df, McCann_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color=factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend = FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
-The scenario where the relationship between character divergence and resource suppression is the strongest is when the tradeoff in attack rates is concave up and consumers are competing for resources that occur in distinct habitats.
-Prior work in threespine stickleback suggests that the evolution of stickleback attack rates are constrained by a concave-up tradeoff (Schluter 1993; Arnegard et al.
-2014) and that stickleback ecotypes are competing for resources that occur in distinct habitats (benthic vs.
-limnetic).
-Thus, we predict that character divergence in sticklebacks would supress resource densities.
-
-## Effect of character divergence on food-web stability
-
-This relationship between character divergence and resource abundances begs the question: how does character divergence affect food-web stability?
-This question has been addressed empirically by Pimentel as well as with a conceptual model.
-Lawlor and Smith though did a detailed theoretical analysis of how character divergence affects three different aspects of stability.
-They looked at stability through the lens of competition.
-The first aspect was whether character divergence resulted in mutual invasibility of consumers, a common metric of "stable coexistence" in competition models.
-Another very interesting aspect they looked at was a global metric that examined the persistence of the system to perturbations in resources.
-Here, they conclude the divergence increases community stability, because it expands the range of resource perturbations that a system can withstand before a consumer going extinct.
-Note though, that they overlooked that there conclusions do not actually quantify the effects of ecological character displacement on stability because they do not compare scenarios where a consumer is in allopatry and one where consumers are in sympatry.
-Finally, they also look at the Lyapunov exponent (local stability), and while they find a negative effect of divergence, they suggest that it would likely be overshadowed by the stabilizing effects of divergence on mutual invasibility and global stability.
-
-Here we focus on local stability, which makes the most sense in a food-web context.
-We find that divergence follows the typical "check-mark" stability pattern that has been described recently in theoretical food-web models.
-
-Importantly, this metric of stability allows us to compare the stability of three-species vs.
-four-species system, which is key for understanding how ecological character displacement affects food-web dynamics.
-
-### Conclusions
-
-Although we used an Adaptive Dynamics approach here, we expect that our results would apply to different models of the genetic architecture of attack rates (e.g.
-quantitative genetics, Taper & Case 1985), given that character divergence is an inevitable outcome of resource competition, regardless of the model of genetic architecture, when resources dynamics are explicitly modelled (Taper & Case 1985).
-
-These conclusions are only for substitutable resources and that exhibit biological dynamics.
+Stability_McCann
+```

### 2018-05-21 eco_evo_sim_function.R
increase mut.reps

@@ -16,7 +16,7 @@ eco_evo_CR <- function(init_parameters,              # data frame of initial par
                        mut.suc.thres=1e-4,           # threshold of mutant success rate for identifying ESS.
                        j.step=500,                   # number of evolutionary time steps over which to assess evolutionary change.
                        abund.thres=1e-4,             # abundance threshold for feasible equilibriums
-                       mut.reps=1000) {             # number of mutation iterations to try for simulation. Note that this is the upper limit of the simulation length. Usually the traits will hit their ESS or a constraint on the parameter space before this limit is reached.   
+                       mut.reps=10000) {             # number of mutation iterations to try for simulation. Note that this is the upper limit of the simulation length. Usually the traits will hit their ESS or a constraint on the parameter space before this limit is reached.   
                             
   #### SETUP ####
   

### 2018-05-21 ECD_Supp_Mat.Rmd
final draft that I will send to Seth and Dolph for feedback

@@
 bibliography: references_ECD_supp_mat.bib
+fontsize: 12pt
 ---
@@
 
-where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ji}$ represents the conversion efficiency of resource $i$ into new consumer $j$, and $m_j$ represents the mortality rate of consumer $j$.
+where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ji}$ represents the conversion efficiency of resource $i$ into consumer $j$, and $m_j$ represents the mortality rate of consumer $j$.
 $F_{ji}(R_i)$ represents consumer $j$'s feeding rate (i.e.
@@
 seed size, see @Taper1985 for an example).
-In the scenario where consumer feeding rates increase linearly with resource density, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$, then my model becomes MacArthur's model of resource competition (@MacArthur1972).
+In the scenario where consumer feeding rates increase linearly with resource density, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$, then this model becomes MacArthur's model of resource competition (@MacArthur1972).
 The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail (@Lawlor1976; @Abrams1986).
@@
 Here, I examine the effects of this inevitable divergence in attack rates on food-web dynamics.
-In particular, I examine how divergence in attack rates affects resource abundances as well as the local stability of both resources and consumers to perturbations.
+In particular, I examine how divergence in attack rates affect resource densities as well as the local stability of both resources and consumers.
 I do this by (i) deriving analytical expressions for the relationship between attack rates and food-web dynamics in different foraging scenarios and (ii) simulating the effects of competition on the eco-evolutionary dynamics of consumers and resources.
 For these simulations, I used an Adaptive Dynamics approach.
-At each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate by either subtracting or adding a small constant <!-- (0.001 in the following simulations) -->with equal probability.
+At each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate on one resource by either subtracting or adding a small constant <!-- (0.001 in the following simulations) -->with equal probability.
+The mutant's attack rate on the other resource was determined by the shape of the tradeoff (details below).
 Assuming this mutant consumer is rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer.
 If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state.
-I then repeated the simulation 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy (ESS, @Smith1973) or an evolutionary limit (e.g.
+I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy (ESS, @Smith1973) or an evolutionary limit (e.g.
 $\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
@@
 
-## Effect of character displacement on resource abundances
+## Effect of character displacement on resource densities
 
-I start by examing the effects of character displacement on the abundance of resources at equilibrium, which will also set the stage for a later analysis of food-web stability.
+I start by examing the effects of character displacement on the density of resources at equilibrium, which will also set the stage for a later analysis of food-web stability.
 In particular, I examine two different foraging scenarios: one where consumers forage for both resources simultaneously (e.g.
@@
 While this scenario is arguably simplistic, it isolates the effect of character displacement rather than any dependence on inherent differences in consumers or resources.
-I show that when both consumers and resources are present that the abundance of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation (*Mathematica* file with derivation available upon request):
+I show that when both consumers and resources are present that the density of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation (*Mathematica* file with derivation available upon request):
 
 $$\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
-Note that a key determinant of resource abundance is the consumer's total attack rate, $a_{ii}+a_{ij}$.
-Since the total attack rate is in the denominator, this implies that resource abundances will decrease if evolution favors an increase in total attack rate.
+Note that a key determinant of resource density is the consumer's total attack rate, $a_{ii}+a_{ij}$.
+Since the total attack rate is in the denominator, this implies that resource densities will decrease if evolution favors an increase in total attack rate.
 I assume that consumers are not 'Darwinian demons' (@Law1979) and thus the evolution of their attack rates are subject to certain constraints.
@@
 \ref{fig:Tradeoff}).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_ii$ and $a_ij$ have the same total attack rate (blue line in Fig.
+When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig.
 \ref{fig:Tradeoff}).
@@
 
-Previous work has shown that consumers still undergo divergent character displacement regardless of the shape of the tradeoff (@Lawlor1976; @Abrams1986).
-Interestingly, I find that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource abundance (Fig.
+Previous work has shown that consumers undergo divergent character displacement regardless of the shape of the tradeoff (@Lawlor1976; @Abrams1986).
+Interestingly, I find that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource density (Fig.
 \ref{fig:MacArthur_Resources}).
-For example, if consumer's are constrained by a linear tradeoff ($n=1$), then there is no net change in total attack rate (blue points in Fig.
-\ref{fig:MacArthur_Resources}A) and character displacement has no effect on resource abundances (Fig.
+For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (Fig.
+\ref{fig:MacArthur_Resources}A) and character displacement has no effect on resource densities (Fig.
 \ref{fig:MacArthur_Resources}B).
-If the tradeoff is concave down ($n>1$), then resource abundances can actually increase under character displacement (green points in Fig.
+If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (Fig.
 \ref{fig:MacArthur_Resources}B).
-This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{aij}$) and decreases as consumers diverge (Fig.
+This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (Fig.
 \ref{fig:MacArthur_Resources}A).
@@
 \ref{fig:MacArthur_Resources}A,B).
-When the tradeoff is concave up ($n<1$), character displacement suppresses resource densities due to the increase in total attack rates (orange points in Fig.
+When the tradeoff is concave up (orange lines), character displacement suppresses resource densities due to the increase in total attack rates (Fig.
 \ref{fig:MacArthur_Resources}A,B).
-Although the equation I derived for resource abundance was for the scenario where both consumers and both resources are present, it predicts well the abundance of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
+Although the equation I derived for resource density was for the scenario where both consumers and both resources are present, it predicts well the density of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
 \ref{fig:MacArthur_Resources}B).
 This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in Fig.
-\ref{fig:MacArthur_Resources}A), resulting in equivalent resource abundances.
+\ref{fig:MacArthur_Resources}A), resulting in equivalent resource densities.
 
@@
 habitat preference).
-Note that since $w_ji$ is a proportion that $w_{jj}=1-w_{ji}$.
+Note that since $w_{ji}$ is a proportion that $w_{jj}=1-w_{ji}$.
 As with attack rates, I assume that consumer habitat preferences are mirror images of each other ($w=w_{11}=w_{22}$).
@@
 @Lawlor1976 found that consumers still underwent divergent character displacement regardless of whether consumers could or could not forage for resources simultaneously.
-I find that the foraging scenario qualitatively affects the relationship between character displacement and resource abundances.
-To demonstrate this, note that again resource abundances are equivalent at equilibrium ($R=R_1=R_2$ when both consumers and resources present), but are now determined by the following equation (*Mathematica* file with derivation available upon request):
+I find that the foraging scenario qualitatively affects the relationship between character displacement and resource densities.
+To demonstrate this, note that again resource densities are equivalent at equilibrium ($R=R_1=R_2$ when both consumers and resources present), but are now determined by the following equation (*Mathematica* file with derivation available upon request):
 
@@
 \ref{fig:LS_Resources}B).
-In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
+In contrast, resource densities reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
 \ref{fig:LS_Resources}B), because character displacement tends to reach a constraint of complete specialization.
@@
 
-It is worth noting that in this foraging scenario, resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
+It is worth noting that in this foraging scenario, resource densities are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
 \ref{fig:LS_Resources}B).
@@
 
-The effect of the foraging scenario on the relationship between character displacement and resource abundances appears to be quite general.
+The effect of the foraging scenario on the relationship between character displacement and resource densities appears to be quite general.
 For example, even if I consider a more realistic consumer functional response (derived by @McCann2005):
@@
 \ref{fig:McCann_Resources} in Appendix).
-This is because resource competition always leads to character divergence and equilibrium resource abundances are governed by similar dynamics  (*Mathematica* file with derivations available upon request):
+This is because resource competition always leads to character divergence and equilibrium resource densities are governed by similar dynamics  (*Mathematica* file with derivations available upon request):
 
@@
 Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement (@Schluter2000; @Schluter1992).
-In fact, the first two metrics of stability can only be assessed when both consumers are present because it examine the effect of divergence on coexistence, which requires both consumers to be present.
+In fact, the first two metrics of stability can only be assessed when both consumers are present because it examines the effect of divergence on coexistence, which requires both consumers to be present.
 In contrast, local stability can be compared in food webs with and without a competing consumer, and thus provides a useful benchmark for examining the effects of character displacement on food-web stability.
@@
 Note that I reach the same fundamental conclusions when I incorporated a more realistic functional response for the consumers (Fig.
-\ref{fig:Stability_McCann}).
+\ref{fig:Stability_McCann} in Appendix).
 
@@
 
-I condcuted the previous simulations on a scenario where the food web remains locally stable ($\operatorname{Re}(\lambda_{max})<0$).
+I conducted the previous simulations on a scenario where the food web remains locally stable ($\operatorname{Re}(\lambda_{max})<0$).
 Is it possible that character displacement would ever result in a locally unstable food web?
@@
 
-Consumer and resource densities in these dynamical model can be interpreted from either a population or biomass perspective (@Yodzis1992; @Murdoch2003; @McCann2011).
+Consumer and resource densities in these dynamical models can be interpreted from either a population or biomass perspective (@Yodzis1992; @Murdoch2003; @McCann2011).
 For the stickleback system, you have data on seasonal dynamics of resources in terms of both abundances and biomass.

### 2018-08-12 ECD_Supp_Mat.Rmd
adding intro from some old written text.

@@
 ---
+
+Keywords: competition; eco-evolutionary dynamics; ecological speciation; three-spine stickleback; zooplankton
+
+## Abstract
+
+Ecological character displacement is a key driver of diversification and adaptive radiation; however, there has been little work examining the consequences of character displacement on the dynamics of ecological communities.
+Given the growing recognition for the role of evolution in structuring ecological communities, understanding how major evolutionary processes impact ecological dynamics is increasingly important.
+Here, we build a mathematical model to predict how exploitative competition between two consumers for two resources affects the evolution of consumer attack rates, and, in turn, consumer-resource population dynamics.
+We then compared these predictions to results from a natural system, where a fish consumer (three-spine stickleback) has become locally adapted to different aquatic habitats with different resources (benthic vs.
+limnetic zones).
+Our model suggested that increasing ecological character displacement decreases the equilibrium density of resources, and with sufficient divergence, results in more oscillatory (destabilizing) consumer-resource dynamics.
+In support of this prediction, we found that the biomass density of zooplankton (resource) in lakes with two species of stickleback was only a fraction of that found in lakes with one species, varying from 4 to 41%.
+Moreover, we found that the biomass density of zooplankton was 2-fold more variable (CV) in lakes with two vs.
+one species of stickleback.
+These results suggest that ecological character displacement can eventually have a negative feedback on food-web persistence in the absence of other processes.
+
+## Introduction
+
+Ecological character displacement (ECD) is thought to be a key evolutionary process in the diversification of species and consequently the generation of biodiversity (Schluter 2000; Pfennig and Pfennig 2010; Stuart and Losos 2013).
+ECD describes the “process of phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” (Schluter 2000).
+Over the past 40 years, a large body of theoretical (e.g.
+Lawlor and Smith 1976; Abrams 1986; Doebeli 1996; Taper and Chase 1985) and empirical (reviewed in: Schluter 2000; Dayan and Simberloff 2005; Stuart and Losos 2013) work has been generated to understand the scenarios under which exploitative competition for resources leads to the divergence of consumer foraging traits.
+Yet, although the process of ecological character displacement is driven by the interplay between competition and trait change, and hence is strongly dependent on resource availability, little of the work to date has explicitly considered how character displacement impacts other species within the community.
+Similarly, research on food-web dynamics has largely ignored the potential role of evolution in altering interactions and stability (Fussmann et al.
+2007).
+Given the emerging evidence that evolutionary processes can drive ecological dynamics (Hairston et al.
+2005, Post et al.
+2008, Harmon et al.
+2009, Schoener 2011, Rudman et al.
+2015), understanding the ecological consequences of character displacement is a natural first step in investigating how evolutionary processes affect food-web dynamics.
+
+Ecological character displacement has typically been considered to be a stabilizing factor in communities (Lawlor and Smith 1976).
+This is because divergence in foraging traits is thought to reduce the strength of interspecific competition between consumers.
+Reducing the strength of interspecific competition relative to intraspecific competition is a stabilizing force in consumer-resource systems.
+Consumer specialization on distinct resources suggests that there are niche differences between consumers.
+Mathematically, this is also often assumed to be the case, because using techniques such as evolutionary invasion analysis describe the ability for a consumer with a mutant allele to increase in abundance when rare, a demographic signature of niche differentiation.
+Therefore, if both consumers are able to invade when rare, then this suggests that they can coexist.
+On the other hand, relatively specialized consumers have also been shown to be a strongly stabilizing force in food webs (McCann et al.
+2005).
+This is because these more specialized consumers may generate asynchrony in the dynamics of patchily distributed resources.
+This asynchrony in resources then permits consumers to move to habitat patches with higher resource densities after they have depleted the resources in a different patch.
+If, for example, the consumers were not specialized, then the resources would exhibit the same dynamics, so when the resources at low densities, consumers would be unable to switch to alternative resources, which would cause the consumer populations to crash, and oscillate more with their resources (McCann et al.
+2005).
+Consequently, whether ecological character displacement will tend to stabilize (via asynchronous resource dynamics) or destabilize (via increased interaction strength) food webs is unclear.
+
+On the other hand, ecological character displacement often results in the evolution of characters that enhance a consumer’s ability to capture resources, which tends to destabilize consumer-resource interactions (Murdoch et al.
+2003; McCann 2011).
+
+This occurs because increases in a consumer’s attack rate results in the resource being suppressed well below its carrying capacity, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics (Rosenzweig 1971; Murdoch et al.
+2003).
+Experimental decreases in interaction strength between consumers and resources has been shown to promote coexistence between consumers and resources (Luckinbill 1973).
+Moreover, there is broad support that specialized consumers, since they are tightly coupled to their resources, tend to exhibit consumer-resource cycles, whereas this is not the case for generalists (Murdoch et al.
+2002).
+Therefore, although evolutionary responses to exploitative competition may reduce the strength of interactions between consumers, they ultimately increase the strength of interactions with their resources which can destabilize the direct consumer-resource interaction (rather than the indirect consumer-consumer interaction).
+While prior models of character displacement have included more realistic functional responses and spatial relationships between consumers and resources, they have focused more on whether these factors alter patterns of character displacement rather than the effects of character displacement on the stability of the ecological community.
+As a consequence, it is currently unclear whether ecological character displacement has a net stabilizing or destabilizing effect on ecological communities.
+
+Here, we address this knowledge gap by building a mathematical model to examine how ecological character displacement affects consumer-resource dynamics.
+We then compared the predictions from our model to field data from one of the most well-studied examples of ecological character displacement in a natural system, threespine stickleback in lakes of southern British Columbia.
+We specifically sought to address two questions: (1) How does exploitative competition between two consumers for two resources affect the evolution of consumer attack rates?
+(2) How does the evolution of consumer attack rates affect community stability?
 

### 2019-01-23 ECD_Supp_Mat.tex
update repo with many changes

\documentclass[11pt,]{article}
\usepackage{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Matthew A. Barbour\^{}\{1,2,\textbackslash{}ast\}},
            pdftitle={Predicting the effects of character displacement on food-web dynamics},
            colorlinks=true,
            citecolor=blue,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Predicting the effects of character displacement on food-web dynamics}
    \author{Matthew A. Barbour\(^{1,2,\ast}\)}
    \date{}

%%% From AmNat_MS_template.tex
%\documentclass[11pt]{article}
%\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
%\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1.7}
%\usepackage[utf8]{inputenc}
\usepackage{lineno}

%%%%%%% Sections not currently working properly
%\usepackage{titlesec}
%\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
%\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
%\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
%\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}

%%% My added packages
%\usepackage{amsmath} % for matrices
%\usepackage{graphicx} % for organizing figures
%\usepackage{lmodern} % for tilde
%\usepackage[T1]{fontenc} % for tilde

%%%%%%%%% Previous - likely useful though
%\usepackage{setspace}\doublespacing
%\usepackage{float}
%\let\origfigure\figure
%\let\endorigfigure\endfigure
%\renewenvironment{figure}[1][2] {
%    \expandafter\origfigure\expandafter[H]
%} {
%    \endorigfigure
%}

\begin{document}

\maketitle

\noindent 1. University of British Columbia, Department of Zoology,
Vancouver, BC V6T 1Z4, Canada;

\noindent 2. University of Zurich, Department of Evolutionary Biology
and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich,
Switzerland;

\(^\ast\) Corresponding author; e-mail:
\href{mailto:matthew.barbour@ieu.uzh.ch}{\nolinkurl{matthew.barbour@ieu.uzh.ch}}

\bigskip

\emph{Manuscript elements}: Figure 1, figure 2, figure 3, figure 4. All
figures should be printed in color.

\bigskip

\emph{Keywords}: competition; eco-evolutionary dynamics; ecological
speciation; three-spine stickleback; zooplankton

\bigskip

\emph{Manuscript type}: Note.

\bigskip

\footnotesize Prepared using an \emph{Am. Nat.} inspired \LaTeX{}
template for Rmarkdown. \normalsize

\linenumbers{} \modulolinenumbers[3]

\newpage

\section{Abstract}\label{abstract}

Ecological character displacement is a key driver of diversification and
adaptive radiation; however, there has been little work examining the
consequences of character displacement on the dynamics of ecological
communities. Given the growing recognition for the role of evolution in
structuring ecological communities, understanding how major evolutionary
processes impact ecological dynamics is increasingly important. Here, we
build a mathematical model to predict how exploitative competition
between two consumers for two resources affects the evolution of
consumer attack rates, and, in turn, consumer-resource population
dynamics. Our model suggested that increasing ecological character
displacement decreases the equilibrium density of resources, and with
sufficient divergence, results in more oscillatory (destabilizing)
consumer-resource dynamics. These results suggest that ecological
character displacement can eventually have a negative feedback on
food-web persistence in the absence of other processes.

\newpage

\section{Introduction}\label{introduction}

Ecological character displacement is thought to be a key evolutionary
process generating biodiversity (Dolph Schluter 2000; D. W. Pfennig and
Pfennig 2010; but see Stuart and Losos 2013). Ecological character
displacement describes the ``process of phenotypic evolution in a
species generated or maintained by {[}exploitative{]} resource
competition with one or more coexisting species'' (Dolph Schluter 2000).
Over the past 40 years, a large body of theoretical (e.g. Lawlor and
Smith 1976; P A Abrams 1986; Doebeli 1996; Taper and Chase 1985) and
empirical (reviewed in: Dolph Schluter 2000; Dayan and Simberloff 2005;
Stuart and Losos 2013) work has been generated to understand the
scenarios under which exploitative competition for resources leads to
the divergence of consumer foraging traits. The general result emerging
from these studies is that, if resources are nutritionally substitutable
(Peter A Abrams 1987; Fox and Vasseur 2008) and there is no other strong
source of density dependence acting on consumers (P A Abrams 1986), then
resource competition drives the divergence of consumer foraging traits
{[}Lawlor and Smith (1976); Taper and Chase (1985)). This process is not
simply driven by ecological differences, but creates an eco-evolutionary
feedback that drives further differentiation. This insight was clearly
by theoretical models that highlighted the importance of explicitly
including resource dynamics as a mediator of competition in driving
evolutionary change (Lawlor and Smith 1976; P A Abrams 1986; Taper and
Chase 1985).

Although character displacement is the result of an eco-evolutionary
feedback, the ecological consequences of this process have received
surprisingly little attention (Germain et al. 2018). Perhaps this is
because early theoretical work suggested that character displacement was
a potent stabilizing force in ecological communities (Lawlor and Smith
1976). This ecological outcome is also an intuitive one ---character
divergence reduces interspecific competition between consumers,
promoting coexistence. However, inferences about the effect of character
displacement, like the process itself, can only be made when comparing
communities in the presence and absence of competition (criteria \#3 for
character displacement, D Schluter and McPhail 1992; Dolph Schluter
2000). Current inferences about the stabilizing effects of consumer
divergence are only comparisons about communities where there are two
consumers, and even though, the focus is only on the stable coexistence
of consumer species.

The consequence of this eco-evolutionary feedback for consumer
coexistence has been well characterized (Lawlor and Smith 1976;
Germain's work). Ecological character displacement has typically been
considered to be a stabilizing factor in communities (Lawlor and Smith
1976). This is because divergence in foraging traits is thought to
reduce the strength of interspecific competition between consumers.
Reducing the strength of interspecific competition relative to
intraspecific competition is a stabilizing force in consumer-resource
systems. Consumer specialization on distinct resources suggests that
there are niche differences between consumers. Mathematically, this is
also often assumed to be the case, because using techniques such as
evolutionary invasion analysis describe the ability for a consumer with
a mutant allele to increase in abundance when rare, a demographic
signature of niche differentiation. Therefore, if both consumers are
able to invade when rare, then this suggests that they can coexist.

Character displacement involves evolutionary shifts in character states
in response to competition (D Schluter and McPhail 1992; Dolph Schluter
2000). Theoretically, this is determined by comparing character changes
in the presence and absence of a competitor. Empirically, this is
determined by meeting several criteria that rule out other possible
explanations (D Schluter and McPhail 1992; Dolph Schluter 2000).
Importantly, current inferences about the effect of character
displacement on community stability do not meet this critera. For
example, the inference that character displacement enhances species
coexistence is only made in a scenario where consumers are competing.
This type of stability is important, but it does not enable us to
examine the effect of character displacement on community stability. As
a consequence, I argue that the effect of character displacement on
community stability remains an open question.

Indeed, ecological character displacement often results in the evolution
of characters that enhance a consumer's ability to capture resources,
which tends to destabilize consumer-resource interactions (Murdoch et
al. 2003; McCann 2011). This occurs because increases in a consumer's
attack rate results in the resource being suppressed well below its
carrying capacity, which if sufficient enough, can generate
oscillations, and hence less stable, consumer-resource dynamics
(Rosenzweig 1971; Murdoch et al. 2003). Experimental decreases in
interaction strength between consumers and resources has been shown to
promote coexistence between consumers and resources (Luckinbill 1973).
Moreover, there is broad support that specialized consumers, since they
are tightly coupled to their resources, tend to exhibit
consumer-resource cycles, whereas this is not the case for generalists
(Murdoch et al. 2002). Therefore, although evolutionary responses to
exploitative competition may reduce the strength of interactions between
consumers, they ultimately increase the strength of interactions with
their resources which can destabilize the direct consumer-resource
interaction (rather than the indirect consumer-consumer interaction).
While prior models of character displacement have included more
realistic functional responses and spatial relationships between
consumers and resources, they have focused more on whether these factors
alter patterns of character displacement rather than the effects of
character displacement on the stability of the ecological community. As
a consequence, it is currently unclear whether ecological character
displacement has a net stabilizing or destabilizing effect on ecological
communities.

Here, I address this knowledge gap by building a mathematical model to
examine how ecological character displacement affects consumer-resource
dynamics in a food web context. I address two question: (1) How does
ecological character displacement affect resource abundances? (2) How
does character displacement affect food-web stability? I show that the
consequences of character displacement on food-web dynamics are not
intuitive and depend critically on the foraging scenario and tradeoff in
consumer attack rates.

\section{Methods and Results}\label{methods-and-results}

\subsection{Model of resource
competition}\label{model-of-resource-competition}

To examine how ecological character displacement affects
consumer-resource interactions and food-web dynamics, I analyzed a
continuous-time model of two consumers (\(C_1,C_2\)) competing for two
shared resources (\(R_1,R_2\)):

\[\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2\]
\[\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2\]
\[\frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1\]
\[\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2\]

where \(r_i\) represents the intrinsic growth rate of resource \(i\),
\(K_i\) represents the carrying capacity of resource \(i\), \(e_{ji}\)
represents the conversion efficiency of resource \(i\) into consumer
\(j\), and \(m_j\) represents the mortality rate of consumer \(j\).
\(F_{ji}(R_i)\) represents consumer \(j\)'s feeding rate
(i.e.~functional response) on resource \(i\). This model is a useful
characterization of a scenario where consumers compete for two distinct
resources (e.g.~zooplankton and benthic invertebrates) rather than if
resources are better characterized by a continuous trait distribution
(e.g.~seed size, see Taper and Chase (1985) for an example). In the
scenario where consumer feeding rates increase linearly with resource
density, such that \(F_{ji}(R_i)=a_{ji}R_i\) where \(a_{ji}\) is the
attack rate of consumer \(j\) on resource \(i\), then this model becomes
MacArthur's model of resource competition (MacArthur (1972)). The
evolution of consumer attack rates in this model (and several
extensions) have been analyzed in detail (Lawlor and Smith (1976); P A
Abrams (1986)). The general result of these analyses is that divergent
character displacement in attack rates is an inevitable outcome of
resource competition.

Here, I examine the effects of this inevitable divergence in attack
rates on food-web dynamics. In particular, I examine how divergence in
attack rates affect resource densities as well as the local stability of
both resources and consumers. I do this by (i) deriving analytical
expressions for the relationship between attack rates and food-web
dynamics in different foraging scenarios and (ii) simulating the effects
of competition on the eco-evolutionary dynamics of consumers and
resources. For these simulations, I used an Adaptive Dynamics approach.
At each evolutionary time step, I created a mutant consumer by randomly
choosing a consumer and modifying its attack rate on one resource by
either subtracting or adding a small constant with equal probability.
The mutant's attack rate on the other resource was determined by the
shape of the tradeoff (details below). Assuming this mutant consumer is
rare, I then determined whether the mutant had higher relative fitness
than the resident consumer, and thus could invade and replace the
resident consumer. If the mutant was able to invade, I updated the
attack rate of the resident consumer to the mutant attack rate and
allowed consumer and resource dynamics to reach a steady state. I then
repeated the simulation up to 10,000 times, which was sufficient for
consumers to either reach an evolutionary stable strategy (ESS, Smith
and Price (1973)) or an evolutionary limit (e.g.
\(\frac{a_{ii}}{a_{ii}+a_{ij}}\) is constrained to a maximum of 1 and
minimum of 0).

\subsection{Resource densities}\label{resource-densities}

I start by examing the effects of character displacement on the density
of resources at equilibrium, which will set the stage for a later
analysis of food-web stability. In particular, I examine two different
foraging scenarios: one where consumers forage for both resources
simultaneously (e.g. MacArthur (1972)) and one where they cannot because
resources occur in distinct habitats (e.g. Lawlor and Smith (1976); K S
McCann, Rasmussen, and Umbanhowar (2005)). I say consumers have
undergone divergent character displacement if there evolved attack rates
are more specialized, defined as \(|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|\),
when evolving with a competing consumer.

\textbf{Consumers forage for both resources simultaneously}: An inherent
assumption in MacArthur's model of resource competition is that
consumers can forage for both resources simultaneously. In essence, this
implies that both resources occur within the same habitat. To gain
analytical insight to this scenario, I assume that resources are
equivalent (\(r=r_1=r_2\) and \(K=K_1=K_2\)) as well as consumers
(\(e=e_{11}=e_{12}=e_{21}=e_{22}\); \(m=m_1=m_2\)), except that consumer
attack rates are mirror images of each other (\(a_{ii}=a_{11}=a_{22}\);
\(a_{ij}=a_{12}=a_{21}\)). While this scenario is arguably simplistic,
it isolates the effect of character displacement rather than any
dependence on inherent differences in consumers or resources. I show
that when both consumers and resources are present that the density of
resources at equilibrium are equivalent (\(R=R_1=R_2\)) and are
determined by the following equation (\emph{Mathematica} file with
derivation available upon request):

\[\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}\] Note that a key
determinant of resource density is the consumer's total attack rate,
\(a_{ii}+a_{ij}\). Since the total attack rate is in the denominator,
this implies that resource densities will decrease if evolution favors
an increase in total attack rate. I assume that consumers are not
`Darwinian demons' (Law (1979)) and thus the evolution of their attack
rates are subject to certain constraints. I modelled these constraints
such that \(\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1\), where
\(A\) is the total investment in attack rates and \(n\) describes the
shape of the tradeoff (Sargent and Otto (2006)). This function has the
useful property that it differentiates between cases where intermediate
combinations of \(a_{ii}\) and \(a_{ij}\) are higher, on average, than
the extremes (when \(n>1\), green line in Fig. \ref{fig:Tradeoff}) or,
conversely, where the two extremes are higher, on average, than
intermediate investments (when \(n<1\), orange line Fig.
\ref{fig:Tradeoff}). When \(n=1\), the tradeoff function is linear, and
all combinations of \(a_{ii}\) and \(a_{ij}\) have the same total attack
rate (blue line in Fig. \ref{fig:Tradeoff}).

\begin{figure}
\centering
\includegraphics{ECD_Supp_Mat_files/figure-latex/Tradeoff-1.pdf}
\caption{\label{fig:Tradeoff}Tradeoff forms for consumer attack rates.}
\end{figure}

Previous work has shown that consumers undergo divergent character
displacement regardless of the shape of the tradeoff (Lawlor and Smith
(1976); P A Abrams (1986)). Interestingly, I find that the shape of the
tradeoff qualitatively affects the relationship between character
displacement and resource density (Fig. \ref{fig:MacArthur_Resources}).
For example, if consumer's are constrained by a linear tradeoff (blue
lines), then there is no net change in total attack rate (Fig.
\ref{fig:MacArthur_Resources}A) and character displacement has no effect
on resource densities (Fig. \ref{fig:MacArthur_Resources}B). If the
tradeoff is concave down (green lines), then resource abundances can
actually increase under character displacement (Fig.
\ref{fig:MacArthur_Resources}B). This is because the total attack rate
of consumers is maximized at intermediate values (\(a_{ii}=a_{ij}\)) and
decreases as consumers diverge (Fig. \ref{fig:MacArthur_Resources}A).
Note that detecting this effect in nature would likely be difficult,
since a concave-down tradeoff results in relatively small displacement
relative to other tradeoff shapes (Lawlor and Smith (1976); green points
in Fig. \ref{fig:MacArthur_Resources}A,B). When the tradeoff is concave
up (orange lines), character displacement suppresses resource densities
due to the increase in total attack rates (Fig.
\ref{fig:MacArthur_Resources}A,B). Although the equation I derived for
resource density was for the scenario where both consumers and both
resources are present, it predicts well the density of resources when a
single consumer reaches its ESS (triangles on respective colored lines
in Fig. \ref{fig:MacArthur_Resources}B). This is because a single
consumer evolves to be a generalist that has equal attack rates on each
resource (triangles at 0.5 along x-axis in Fig.
\ref{fig:MacArthur_Resources}A), resulting in equivalent resource
densities.

\begin{figure}
\centering
\includegraphics{ECD_Supp_Mat_files/figure-latex/MacArthur_Resources-1.pdf}
\caption{\label{fig:MacArthur_Resources}Effect of character displacement
on total attack rates (A) and resource densities (B) when consumers can
forage for both resources simultaneously.}
\end{figure}

\textbf{Consumers \emph{cannot} forage for both resources
simultaneously}: Most models of character displacement have assumed a
scenario where consumers can forage for both resources simultaneously
(e.g. Taper and Chase (1985); P A Abrams (1986)). This assumption may be
valid for some foraging scenarios (e.g.~Darwin's finches foraging for
seeds); however, in many situations consumers forage for resources that
occur in different habitats, and thus cannot forage for both
simultaneosly. The only character displacement model that I am aware of
that modeled resources in different habitats was one examined by Lawlor
and Smith (1976). This model takes the same form as the previous model,
except now the consumer's feeding rate takes the form:

\[F_{ji}(R_i)=w_{ji}a_{ji}R_i\]

where \(w_{ji}\) represents the proportion of time consumer \(j\) spends
foraging in a habitat where only resource \(j\) is found (i.e.~habitat
preference). Note that since \(w_{ji}\) is a proportion that
\(w_{jj}=1-w_{ji}\). As with attack rates, I assume that consumer
habitat preferences are mirror images of each other
(\(w=w_{11}=w_{22}\)).

Lawlor and Smith (1976) found that consumers still underwent divergent
character displacement regardless of whether consumers could or could
not forage for resources simultaneously. I find that the foraging
scenario qualitatively affects the relationship between character
displacement and resource densities. To demonstrate this, note that
again resource densities are equivalent at equilibrium (\(R=R_1=R_2\)
when both consumers and resources present), but are now determined by
the following equation (\emph{Mathematica} file with derivation
available upon request):

\[\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e}\]

This equation implies that if consumers evolve to become specialists on
resources that occur in their preferred habitat (e.g. \(w>0.5\) and
\(a_{ii}>a_{ij}\)), then character displacement will always increase the
effective attack rate of consumers (\(wa_{ii}+(1-w)a_{ij}\)), regardless
of the tradeoff (Fig. \ref{fig:LS_Resources}A). Thus, character
displacement always results in resource suppression if consumers are
competing for resources that occur in different habitats (Fig.
\ref{fig:LS_Resources}B). Note that the shape of the tradeoff can modify
the effect of character displacement. This is not so much due to the
tradeoff affecting the magnitude of displacement (it does, but the
effect is minor), but because the form of the tradeoff affects resource
abundances when a single consumer has reached its ESS (triangles in Fig.
\ref{fig:LS_Resources}B). In contrast, resource densities reach a
similar value when consumers evolve in the presence of a competitor
(circles in Fig. \ref{fig:LS_Resources}B), because character
displacement tends to reach a constraint of complete specialization.

These results do lead to an interesting prediction that could be tested
either experimentally or with field data. If consumers are competing for
resources that occur in different habitats, then greater character
displacement corresponds to a greater decrease in resource densities
(line trajectories in Fig. \ref{fig:LS_Resources}B). This prediction
likely only applies to comparisons within species where there is likely
little difference in the shape of the tradeoff among populations.

It is worth noting that in this foraging scenario, resource densities
are consistently higher at the single consumer ESS compared to the
predictions I derived for when both consumers are present (deviation of
triangles from respective colored lines in Fig.
\ref{fig:LS_Resources}B). This is likely because consumers actually
evolve to be slightly specialized on the resources that occur in their
non-preferred habitat (deviation of triangles from 0.5 along x-axis in
Fig. \ref{fig:LS_Resources}B).

\begin{figure}
\centering
\includegraphics{ECD_Supp_Mat_files/figure-latex/LS_Resources-1.pdf}
\caption{\label{fig:LS_Resources}Effect of character displacement on
effective attack rates (A) and resource densities (B) when consumers
cannot forage for resources simultaneously.}
\end{figure}

The effect of the foraging scenario on the relationship between
character displacement and resource densities appears to be quite
general. For example, even if I consider a more realistic consumer
functional response (derived by K S McCann, Rasmussen, and Umbanhowar
(2005)):

\[F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}hW_{ii}R_i+a_{ij}hW_{ij}R_j}\]

where consumer feeding rates on resource \(i\) are influenced by both
resource densities; saturate as resource densities increase (due to
handling time \(h\)); and consumer habitat preferences are modified by
relative resource densities, such that:

\[W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}\]

I still observe the same qualitative relationship between character
displacement and resource suppression (Fig. \ref{fig:McCann_Resources}
in Appendix). This is because resource competition always leads to
character divergence and equilibrium resource densities are governed by
similar dynamics (\emph{Mathematica} file with derivations available
upon request):

\[\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e-hm}\]

\subsection{Food-web stability}\label{food-web-stability}

Of the prior models of character displacement, only Lawlor and Smith
(1976) examine the effect of character displacement on the ecological
stability of the system. They conducted a detailed theoretical analysis
of three different metrics of stability: (i) stable coexistence of
consumers, defined as the mutual invasibility of each consumer; (ii)
global stability of consumers, defined as the continued coexistence of
both consumers when resource dynamics are perturbed; and (iii) local
stability, defined as the rate at which the four-species food web
returned to equilibrium after a small perturbation to species
abundances. Note that the first two metrics correspond to the stable
coexistence of consumers, while the third metric measures the stability
of the entire food web. Lawlor and Smith (1976) found that character
displacement enhanced the stable coexistence of consumers, yet had a
small negative effect on the local stability of the food web.

Lawlor and Smith (1976)'s work gave fundamental insight to the effect of
character displacement on ecological stability; however, none of their
stability analyses compared food webs with and without a competing
consumer. Such comparisons are necessary for inferring the effects of
ecological character displacement, a point that has been made clear in
the criteria to demonstrate character displacement (Dolph Schluter
(2000); D Schluter and McPhail (1992)). In fact, the first two metrics
of stability can only be assessed when both consumers are present
because it examines the effect of divergence on coexistence, which
requires both consumers to be present. In contrast, local stability can
be compared in food webs with and without a competing consumer, and thus
provides a useful benchmark for examining the effects of character
displacement on food-web stability.

Below, I plot the results from the same simulations I did previously
except now local stability
(\(-1\times\operatorname{Re}(\lambda_{max})\)) is on the y-axis (Fig.
\ref{fig:Stability}). In general, I find that resource supression goes
hand-in-hand with local stability, a tendency that has been noted by
others (Murdoch, Briggs, and Nisbet (2003); Kevin S McCann (2011)).
Whenever character displacement results in resource suppression, I
observe a decrease in local stability (Fig. \ref{fig:Stability}). This
is not simply a consequence of having an additional consumer in the
system, but emerges from the eco-evolutionary feedback between character
displacement and resource suppression (compare initial (small points)
with evolved strategies (large points), Fig. \ref{fig:Stability}). Note
that I reach the same fundamental conclusions when I incorporated a more
realistic functional response for the consumers (Fig.
\ref{fig:Stability_McCann} in Appendix).

\begin{figure}
\centering
\includegraphics{ECD_Supp_Mat_files/figure-latex/Stability-1.pdf}
\caption{\label{fig:Stability}Effect of character displacement on local
stability when consumers can (A) and cannot (B) forage for both
resources simultaneously. Small and large points correspond to initial
and evolved strategies, respectively.}
\end{figure}

I conducted the previous simulations on a scenario where the food web
remains locally stable (\(\operatorname{Re}(\lambda_{max})<0\)). Is it
possible that character displacement would ever result in a locally
unstable food web? To gain analytical insight to this question, I
examined the conditions governing the local stability of the
four-species food web when consumers exhibit a more realistic functional
response (K S McCann, Rasmussen, and Umbanhowar (2005)). I show that the
four-species food web will transition from having a locally stable
equilibrium to a limit cycle under the following conditions
(\emph{Mathematica} file with derivation available upon request):

\[wa_{ii}+(1-w)a_{ij} > \frac{e+hm}{hK(e-hm)}\]

Previously, I showed that character displacement always increases the
effective attack rate of consumers (\(wa_{ii}+(1-w)a_{ij}\)), regardless
of the the shape of the tradeoff (Fig. \ref{fig:LS_Resources}A). Thus,
character displacement is capable of pushing food webs to a point where
they are locally unstable.

\section{Discussion}\label{discussion}

Important point: Criteria 5 in D Schluter and McPhail (1992) does not
make sense since character displacement usually alters resource
distributions. Think about this more.

\subsection{Predictions for threespine
stickleback}\label{predictions-for-threespine-stickleback}

The scenario where the relationship between character displacement and
resource suppression is the strongest is when the tradeoff in attack
rates is concave-up and consumers are competing for resources that occur
in different habitats (orange points in Fig. \ref{fig:LS_Resources}B and
Fig. \ref{fig:McCann_Resources}B). Prior work in threespine stickleback
suggests that the evolution of stickleback attack rates are constrained
by a concave-up tradeoff (Dolph Schluter (1993); Arnegard et al. (2014))
and that stickleback ecotypes are competing for resources that occur in
different habitats (benthic vs.~limnetic, D Schluter and McPhail
(1992)). Thus, I predict that ecological character displacement in
sticklebacks would suppress resource densities and decrease local
stability.

While measuring resource densities is relatively straightforward,
measuring stability is notoriously difficult. One of the more easy to
acquire empirical metrics of stability is the coefficient of variation
(\(CV=\frac{\sigma}{\mu}\)) as this only requires measurements of
resource (or consumer) densities over time. Interestingly, there is a
close correspondence between the coefficient of variation and the local
stability of a food web (Gellner, McCann, and Hastings (2016)). Thus, I
predict that character displacement in threespine stickleback will
increase the \(CV\) of resource (or consumer) densities.

Consumer and resource densities in these dynamical models can be
interpreted from either a population or biomass perspective (Yodzis and
Innes (1992); Murdoch, Briggs, and Nisbet (2003); Kevin S McCann
(2011)). For the stickleback system, you have data on seasonal dynamics
of resources in terms of both abundances and biomass. Within a season, I
do not expect stickleback abundances to be coupled to the abundances of
zooplankton and benthic invertebrates, given that stickleback reproduce
annually whereas zooplankton and benthic invertebrates can reproduce
many times within a season. Instead, I expect that stickleback biomass
will be coupled to the biomass of zooplankton and benthic invertebrates
within a season. This is because there will be lots of young stickleback
at the beginning of the season, and they will eat a lot of resources and
grow in size as the season progresses. If this is true, I also expect to
see the largest negative effect of stickleback on resource densities
late in the season.

\subsubsection{Other relevant points}\label{other-relevant-points}

Although I used an Adaptive Dynamics approach here, I expect that these
theoretical predictions would apply to models that assumed a different
genetic architecture of attack rates (e.g.~quantitative genetics). I
expect this because, time and time again, models that explicitly include
resource dynamics inevitably show that resource competition results in
divergent character displacement, regardless of the genetic architecture
of traits (see synthesis by Taper and Chase (1985)). Finally, my
conclusions only apply to biological resources that are nutritionally
substitutable. It would be interesting to extend these current analyses
to non-substitutable resources where convergent character displacement
is expected (Peter A Abrams (1987); Fox and Vasseur (2008)).

\subsection{Acknowledgement}\label{acknowledgement}

The content of this paper was greatly enhanced by numerous conversations
with Seth Rudman. Sally Otto provided much help in early analyses of the
mathematical model and I would not have been able to take the theory as
far as I did without her guidance and her pushing me.

\subsection{Appendix: More realistic functional
response}\label{appendix-more-realistic-functional-response}

\begin{figure}
\centering
\includegraphics{ECD_Supp_Mat_files/figure-latex/McCann_Resources-1.pdf}
\caption{\label{fig:McCann_Resources}Effect of character displacement on
total attack rates (A) and resource densities (B) when consumers exhibit
a more realistic functional response.}
\end{figure}

\begin{figure}
\centering
\includegraphics{ECD_Supp_Mat_files/figure-latex/Stability_McCann-1.pdf}
\caption{\label{fig:Stability_McCann}Effect of character displacement on
local stability when consumers exhibit a more realistic functional
response.}
\end{figure}

\hypertarget{refs}{}
\hypertarget{ref-Abrams1986}{}
Abrams, P A. 1986. ``Character Displacement and Niche Shift Analyzed
Using Consumer-Resource Models of Competition.'' \emph{Theor. Popul.
Biol.} 29 (1): 107--60.

\hypertarget{ref-Abrams1987}{}
Abrams, Peter A. 1987. ``Alternative Models of Character Displacement
and Niche Shift. I. Adaptive Shifts in Resource Use When There Is
Competition for Nutritionally Nonsubstitutable Resources.''
\emph{Evolution} 41 (3). Wiley Online Library: 651--61.

\hypertarget{ref-Arnegard2014}{}
Arnegard, Matthew E, Matthew D McGee, Blake Matthews, Kerry B Marchinko,
Gina L Conte, Sahriar Kabir, Nicole Bedford, et al. 2014. ``Genetics of
Ecological Divergence During Speciation.'' \emph{Nature} 511 (7509):
307--11.

\hypertarget{ref-Dayan2005}{}
Dayan, Tamar, and Daniel Simberloff. 2005. ``Ecological and
Community-Wide Character Displacement: The Next Generation.''
\emph{Ecol. Lett.} 8 (8). Blackwell Science Ltd: 875--94.

\hypertarget{ref-Doebeli1996}{}
Doebeli, Michael. 1996. ``An Explicit Genetic Model for Ecological
Character Displacement.'' \emph{Ecology} 77 (2). Ecological Society of
America: 510--20.

\hypertarget{ref-Fox2008}{}
Fox, Jeremy W, and David A Vasseur. 2008. ``Character Convergence Under
Competition for Nutritionally Essential Resources.'' \emph{The American
Naturalist} 172 (5). The University of Chicago Press: 667--80.

\hypertarget{ref-Gellner2016}{}
Gellner, Gabriel, Kevin S McCann, and Alan Hastings. 2016. ``The Duality
of Stability: Towards a Stochastic Theory of Species Interactions.''
\emph{Theor. Ecol.} 9 (4): 477--85.

\hypertarget{ref-Germain2018}{}
Germain, Rachel M, Jennifer L Williams, Dolph Schluter, and Amy L
Angert. 2018. ``Moving Character Displacement Beyond Characters Using
Contemporary Coexistence Theory.'' \emph{Trends Ecol. Evol.} 33 (2):
74--84.

\hypertarget{ref-Law1979}{}
Law, Richard. 1979. ``Optimal Life Histories Under Age-Specific
Predation.'' \emph{Am. Nat.} 114 (3): 399--417.

\hypertarget{ref-Lawlor1976}{}
Lawlor, Lawrence R, and J Maynard Smith. 1976. ``The Coevolution and
Stability of Competing Species.'' \emph{Am. Nat.} 110 (971).
{[}University of Chicago Press, American Society of Naturalists{]}:
79--99.

\hypertarget{ref-MacArthur1972}{}
MacArthur, R H. 1972. \emph{Geographical Ecology: Patterns in the
Distribution of Species}. Biology / {[}Princeton University Press{]}.
Princeton University Press.

\hypertarget{ref-McCann2005}{}
McCann, K S, J B Rasmussen, and J Umbanhowar. 2005. ``The Dynamics of
Spatially Coupled Food Webs.'' \emph{Ecol. Lett.} 8 (5). Blackwell
Science Ltd: 513--23.

\hypertarget{ref-McCann2011}{}
McCann, Kevin S. 2011. \emph{Food Webs (Mpb-50)}. Princeton University
Press.

\hypertarget{ref-Murdoch2003}{}
Murdoch, W W, C J Briggs, and R M Nisbet. 2003. \emph{Consumer-Resource
Dynamics}. Monographs in Population Biology. Princeton University Press.

\hypertarget{ref-Pfennig2010}{}
Pfennig, David W, and Karin S Pfennig. 2010. ``Character Displacement
and the Origins of Diversity.'' \emph{Am. Nat.} 176 Suppl 1 (December):
S26--44.

\hypertarget{ref-Sargent2006}{}
Sargent, Risa D, and Sarah P Otto. 2006. ``The Role of Local Species
Abundance in the Evolution of Pollinator Attraction in Flowering
Plants.'' \emph{Am. Nat.} 167 (1): 67--80.

\hypertarget{ref-Schluter1992}{}
Schluter, D, and J D McPhail. 1992. ``Ecological Character Displacement
and Speciation in Sticklebacks.'' \emph{Am. Nat.} 140 (1): 85--108.

\hypertarget{ref-Schluter1993}{}
Schluter, Dolph. 1993. ``Adaptive Radiation in Sticklebacks: Size,
Shape, and Habitat Use Efficiency.'' \emph{Ecology} 74 (3): 699--709.

\hypertarget{ref-Schluter2000}{}
---------. 2000. ``Ecological Character Displacement in Adaptive
Radiation.'' \emph{Am. Nat.} 156 (S4): S4--S16.

\hypertarget{ref-Smith1973}{}
Smith, J Maynard, and G R Price. 1973. ``The Logic of Animal Conflict.''
\emph{Nature} 246 (November). Nature Publishing Group: 15.

\hypertarget{ref-Stuart2013}{}
Stuart, Yoel E, and Jonathan B Losos. 2013. ``Ecological Character
Displacement: Glass Half Full or Half Empty?'' \emph{Trends Ecol. Evol.}
28 (7): 402--8.

\hypertarget{ref-Taper1985}{}
Taper, Mark L, and Ted J Chase. 1985. ``Quantitative Genetic Models for
the Coevolution of Character Displacement.'' \emph{Ecology} 66 (2).
Ecological Society of America: 355--71.

\hypertarget{ref-Yodzis1992}{}
Yodzis, P, and S Innes. 1992. ``Body Size and Consumer-Resource
Dynamics.'' \emph{Am. Nat.} 139 (6). {[}University of Chicago Press,
American Society of Naturalists{]}: 1151--75.

\end{document}

### 2019-01-23 manuscript_ECD_model.Rmd
update repo with many changes

--- 2018-05-21 ECD_Supp_Mat.Rmd
@@
 ---
-title: "Predicting the effects of character displacement on food-web dynamics"
-author: "Matt Barbour"
+title: Predicting the effects of character displacement on food-web dynamics
+author: 
+- Matthew A. Barbour$^{1,2,\ast}$
+# potentially use for 'date' for affiliations, as recommended by https://github.com/rstudio/bookdown/issues/348 for cross-format compatibility, with | to create new sections.
+#date: | 
+#    | 1. Affiliation
 output:
   pdf_document:
+    keep_tex: true
+    toc: false
+    number_sections: false
     df_print: paged
     fig_caption: true
-    citation_package: natbib
+    #citation_package: natbib
+    template: Rmd_AmNat_MS_template.tex
     includes:  
-      in_header: preamble-latex.tex
+      in_header: Am-Nat-preamble-latex.tex
 bibliography: references_ECD_supp_mat.bib
-fontsize: 12pt
+biblio-style: amnat.bst
+fontsize: 11pt
+fontfamily: mathpazo
+documentclass: article
+linkcolor: black
+urlcolor: black
 ---
+
+<!-- Begin title page formatting -->
+
+\noindent 1. University of British Columbia, Department of Zoology, Vancouver, BC V6T 1Z4, Canada;
+
+\noindent 2. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;
+
+$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch
+
+\bigskip
+
+*Manuscript elements*: Figure 1, figure 2, figure 3, figure 4. All figures should be printed in color.
+
+\bigskip
+
+*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptive radiation; ecological community
+
+\bigskip
+
+*Manuscript type*: Note.
+
+\bigskip
+
+\footnotesize Prepared using an *Am. Nat.* inspired \LaTeX{} template for Rmarkdown. \normalsize
+
+<!-- End title page formatting -->
+
+\linenumbers{}
+\modulolinenumbers[3]
+
+\newpage
+
+<!-- Outline
+
+Introduction
+
+1. ECD is an important process, and current empirical and theoretical work suggests it often results in divergence of consumer foraging traits.
+
+2.
+Ecological consequences of character displacement have received surprisingly little attention.
+In particular, prior work doesn't allow a proper comparision of the effects of character displacement on community stability.
+
+-> Another paragraph illustrating how the direct and indirect effects of character displacement likely have constrasting effects on stability.
+
+3.
+Here we fill this knowledge gap.
+We focus on the ecological side of these dynamics, because the evolutionary aspects have been well characterized, although we show in the appendices that the evolutionary dynamics are as expected (especially for McCann model).
+
+Methods and Results
+
+- Underlying model of consumer-resource dynamics
+  1. Explanation of model parameters
+  2. Methods for assessing resource abundances and stability.
+  3. Trade-off function and definition of character displacement.
+  4. Analytical and simulation methods. Briefly mention the different scenarios I'll explore.
+
+- Consumers can forage for resources simultaneously
+  1. Functional response equation and illustration of assumptions
+  2. Effect on attack rates, resource densities, and stability
+
+- Consumers cannot foraging for resource simultaneously
+  1. Functional response equation and illustration of assumptions
+  2. Effect on attack rates, resource densities, and stability
+
+- McCann's most realistic functional response
+  1. Functional response equation and illustration of assumptions
+  2. Effect on attack rates, resource densities, and stability
+
+Discussion
+
+-->
 
@@
 
-## Model of resource competition
+# Abstract
 
-I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
+<!--
+Ecological character displacement is a key driver of diversification and adaptive radiation; however, there has been little work examining the consequences of character displacement on the dynamics of ecological communities.
+Given the growing recognition for the role of evolution in structuring ecological communities, understanding how major evolutionary processes impact ecological dynamics is increasingly important.
+Here, we build a mathematical model to predict how exploitative competition between two consumers for two resources affects the evolution of consumer attack rates, and, in turn, consumer-resource population dynamics.
+Our model suggested that increasing ecological character displacement decreases the equilibrium density of resources, and with sufficient divergence, results in more oscillatory (destabilizing) consumer-resource dynamics.
+These results suggest that ecological character displacement can eventually have a negative feedback on food-web persistence in the absence of other processes.
+-->
 
-$$\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2$$
-$$\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2$$
-$$\frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1$$
-$$\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2$$
+\newpage
+
+# Introduction
+
+<!-- ECD is an important process, and current empirical and theoretical work suggests it often results in divergence of consumer foraging traits.-->
+
+Ecological character displacement is thought to be a key evolutionary process generating biodiversity [@Schluter2000; @Pfennig2010; but see @Stuart2013].
+Ecological character displacement describes the “process of phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000].
+Over the past 40 years, a large body of theoretical [e.g.
+@Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has been generated to understand the scenarios under which exploitative competition for resources leads to the divergence of consumer foraging traits.
+The general result emerging from these studies is that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the divergence of consumer foraging traits [@Lawlor1976; @Taper1985).
+This process is not simply driven by ecological differences, but creates an eco-evolutionary feedback that drives further differentiation.
+This key insight was made by theoretical models that explicitly included resource dynamics as a mediator of competition in driving evolutionary change [@Lawlor1976; @Abrams1986; @Taper1985].
+
+<!-- Ecological consequences of character displacement have received surprisingly little attention.
+In particular, prior work doesn't allow a proper comparision of the effects of character displacement on community stability.-->
+
+Although explicit models of resources gave key insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics have received surprisingly little attention.
+This is likely because character displacement has been primarily studied through the lens of competition and coexistence theory [Lawlor1976; Germain2018].
+For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition.
+On the other hand, specialization often increases a consumer’s ability to capture a specific resource.
+A well known consequence of increased foraging efficiency is resource suppression, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics [Rosenzweig 1971; Murdoch et al.
+2003; McCann 2011].
+NEED A LAST SENTENCE TO WRAP IT UP, HIGHLIGHTING HOW A FOOD-WEB PERSPECTIVE MAY YIELD NEW ECOLOGICAL INSIGHTS AND GIVE A CONTRASTING PERSPECTIVE TO COEXISTENCE THEORY.
+PERHAPS A I NEED TO INCLUDE A SENTENCE MID-WAY THROUGH SUGGESTING THAT A FOOD-WEB PERSPECTIVE MAY GIVE QUALITATIVELY DIFFERENT INSIGHTS.
+IT MAY BE NICE TO BRING IN THE SPATIAL COUPLING WORK OF MCCANN SUGGESTING THAT SPECIALIZATION CAN COUPLE FOOD-WEBS, BUT THIS WAS ONLY WITH ONE TOP PREDATOR (I.E.
+NO CONSUMERS...LOOK AT OLD TEXT FOR WORDING SUGGESTIONS.)
+
+<!-- older idea...Write 3rd paragraph of intro, highlighting how the direct effects of character displacement, surprisingly, have been ignored.
+We are currently lacking insight to the relative importance of these direct and indirect effects of character displacement on community dynamics.
+
+<!--
+The role of character displacement in governing food-web dynamics remains unclear A food-web perspective, however, has been lacking from theoretical work on the ecological effects of character displacement.
+, thus the consequences of this process for food-web stabilit
+
+<!--
+Despite the key insights made by explicit models of resource dynamics,
+
+gave key insights to the evolutionary dynamics of character displacement, the ecological effects on resources dynamics have received surprisingly little attention.
+Perhaps this is because the ecological dynamics have been studied primarily through the lens of competition and coexistence thoery
+
+This oversight is unfortunate, given that the stability of consumer-resources is often closely tied with resource abundances.
+
+Although character displacement is the result of an eco-evolutionary feedback, the ecological consequences of this process have received surprisingly little attention [@Germain2018].
+Perhaps this is because early theoretical work suggested that character displacement was a potent stabilizing force in ecological communities [@Lawlor1976].
+This ecological outcome is intuitive in light of competition theory ---character divergence reduces interspecific competition between consumers, promoting coexistence.
+On the other hand, specialization often enhances a consumer’s ability to capture a specific resource.
+A well known consequence of increased foraging efficiency (all else equal) is resource suppression, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics (Rosenzweig 1971; Murdoch et al.
+2003; McCann 2011).
+
+Moreover, past inferences about the stabilizing effects of character displacement have only been made in communities where both consumers are present.
+However, I argue that inferences about the effect of character displacement, like the process itself, should only be made when comparing communities in the presence and absence of competition [criteria #3 for character displacement, @Schluter1992; @Schluter2000].
+Therefore, the role of ecological character displacement in determining community stability remains unclear.
+
+<!--
+However, inferences about the effect of character displacement, like the process itself, can only be made when comparing communities in the presence and absence of competition [criteria #3 for character displacement, @Schluter1992; @Schluter2000].
+Current inferences about the stabilizing effects of consumer divergence are only comparisons about communities where there are two consumers, and even though, the focus is only on the stable coexistence of consumer species.
+
+<!--
+The consequence of this eco-evolutionary feedback for consumer coexistence has been well characterized (Lawlor and Smith 1976; Germain's work).
+Ecological character displacement has typically been considered to be a stabilizing factor in communities (Lawlor and Smith 1976).
+This is because divergence in foraging traits is thought to reduce the strength of interspecific competition between consumers.
+Reducing the strength of interspecific competition relative to intraspecific competition is a stabilizing force in consumer-resource systems.
+Consumer specialization on distinct resources suggests that there are niche differences between consumers.
+Mathematically, this is also often assumed to be the case, because using techniques such as evolutionary invasion analysis describe the ability for a consumer with a mutant allele to increase in abundance when rare, a demographic signature of niche differentiation.
+Therefore, if both consumers are able to invade when rare, then this suggests that they can coexist.
+
+Character displacement involves evolutionary shifts in character states in response to competition [@Schluter1992; @Schluter2000].
+Theoretically, this is determined by comparing character changes in the presence and absence of a competitor.
+Empirically, this is determined by meeting several criteria that rule out other possible explanations [@Schluter1992; @Schluter2000].
+Importantly, current inferences about the effect of character displacement on community stability do not meet this critera.
+For example, the inference that character displacement enhances species coexistence is only made in a scenario where consumers are competing.
+This type of stability is important, but it does not enable us to examine the effect of character displacement on community stability.
+As a consequence, I argue that the effect of character displacement on community stability remains an open question.
+
+Of the prior models of character displacement, only @Lawlor1976 examine the effect of character displacement on the ecological stability of the system.
+They conducted a detailed theoretical analysis of three different metrics of stability: (i) stable coexistence of consumers, defined as the mutual invasibility of each consumer; (ii) global stability of consumers, defined as the continued coexistence of both consumers when resource dynamics are perturbed; and (iii) local stability, defined as the rate at which the four-species food web returned to equilibrium after a small perturbation to species abundances.
+Note that the first two metrics correspond to the stable coexistence of consumers, while the third metric measures the stability of the entire food web.
+@Lawlor1976 found that character displacement enhanced the stable coexistence of consumers, yet had a small negative effect on the local stability of the food web.
+@Lawlor1976's work gave fundamental insight to the effect of character displacement on ecological stability; however, none of their stability analyses compared food webs with and without a competing consumer.
+Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement (@Schluter2000; @Schluter1992).
+In fact, the first two metrics of stability can only be assessed when both consumers are present because it examines the effect of divergence on coexistence, which requires both consumers to be present.
+In contrast, local stability can be compared in food webs with and without a competing consumer, and thus provides a useful benchmark for examining the effects of character displacement on food-web stability.
+
+<!-- -> Another paragraph illustrating how the direct and indirect effects of character displacement likely have constrasting effects on stability. -->
+
+<!--
+Indeed, ecological character displacement often results in the evolution of characters that enhance a consumer’s ability to capture resources, which tends to destabilize consumer-resource interactions (Murdoch et al.
+2003; McCann 2011).
+This occurs because increases in a consumer’s attack rate results in the resource being suppressed well below its carrying capacity, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics (Rosenzweig 1971; Murdoch et al.
+2003).
+Experimental decreases in interaction strength between consumers and resources has been shown to promote coexistence between consumers and resources (Luckinbill 1973).
+Moreover, there is broad support that specialized consumers, since they are tightly coupled to their resources, tend to exhibit consumer-resource cycles, whereas this is not the case for generalists (Murdoch et al.
+2002).
+Therefore, although evolutionary responses to exploitative competition may reduce the strength of interactions between consumers, they ultimately increase the strength of interactions with their resources which can destabilize the direct consumer-resource interaction (rather than the indirect consumer-consumer interaction).
+While prior models of character displacement have included more realistic functional responses and spatial relationships between consumers and resources, they have focused more on whether these factors alter patterns of character displacement rather than the effects of character displacement on the stability of the ecological community.
+As a consequence, it is currently unclear whether ecological character displacement has a net stabilizing or destabilizing effect on ecological communities.
+
+<!-- Here we fill this knowledge gap.
+We focus on the ecological side of these dynamics, because the evolutionary aspects have been well characterized, although we show in the appendices that the evolutionary dynamics are as expected (especially for McCann model).
+-->
+
+Here, I address this knowledge gap by building a mathematical model to examine how ecological character displacement affects consumer-resource dynamics in a food-web context.
+I address two questions: (1) How does ecological character displacement affect resource abundances?
+(2) How does character displacement affect food-web stability?
+To test the generality of these ecological effects, I examined three different foraging scenarios: (i) resources occur in the same habitat and consumers exhibit a linear functional response (e.g.
+@MacArthur1972; @Lawlor1976); (ii) resources occur in different habitats and consumers exhibit a linear functional response (e.g.
+@Lawlor1976); (iii) resources occur in distinct habitats and consumers exhibit a more realistic functional response [@McCann2005].
+I also explore the contingency of these effects to different assumptions about evolutionary trade-offs in consumer foraging traits.
+I focus on the ecological side of these dynamics because the evolutionary aspects have been well characterized [@Lawlor1976; @Taper1985; @Abrams1986].
+I show that the consequences of character displacement on food-web dynamics are not always intuitive and depend on the foraging scenario and tradeoff in consumer traits.
+
+# Methods and Results
+
+## Underlying Food-web Dynamics
+
+<!-- Explanation of model parameters -->
+
+To examine how ecological character displacement affects consumer-resource interactions and food-web dynamics, I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
+
+$$\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2,$$
+$$\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2,$$
+$$\frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1,$$
+$$\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2,$$
 
@@
 In the scenario where consumer feeding rates increase linearly with resource density, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$, then this model becomes MacArthur's model of resource competition (@MacArthur1972).
-The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail (@Lawlor1976; @Abrams1986).
-The general result of these analyses is that divergent character displacement in attack rates is an inevitable outcome of resource competition.
-<!-- In other words, the presence of a competitor causes consumers to evolve more specialized attack rates ($\frac{a_{ii}}{a_{ii}+a_{ij}}$) compared to when no competitor is present.
--->
+The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail (@Lawlor1976; @Abrams1986), with the general result being divergent character displacement.
+I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, defined as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
+I focus here on the ecological consequences of this adaptive divergence.
 
-Here, I examine the effects of this inevitable divergence in attack rates on food-web dynamics.
-In particular, I examine how divergence in attack rates affect resource densities as well as the local stability of both resources and consumers.
-I do this by (i) deriving analytical expressions for the relationship between attack rates and food-web dynamics in different foraging scenarios and (ii) simulating the effects of competition on the eco-evolutionary dynamics of consumers and resources.
-For these simulations, I used an Adaptive Dynamics approach.
-At each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate on one resource by either subtracting or adding a small constant <!-- (0.001 in the following simulations) -->with equal probability.
-The mutant's attack rate on the other resource was determined by the shape of the tradeoff (details below).
-Assuming this mutant consumer is rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer.
-If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state.
-I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy (ESS, @Smith1973) or an evolutionary limit (e.g.
-$\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
-I show that the consequences of character displacement on food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
+<!-- Methods for assessing resource abundances and stability. -->
 
-## Effect of character displacement on resource densities
+To investigate the ecological consequences, I studied the effect of adaptive divergence in consumers on resource densities and the stability of the entire food web at equilibrium.
+An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium densities for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$).
+At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource densities and performed a linear stability analysis [@Otto2007].
+This stability analysis derives the dominant (largest in absolute value) eigenvalue, $\lambda$, which determines whether (and how readily) the food web, when perturbed a small amount, will return to equilibrium (see supplementary material for derivation).
+Together, these two measures indicate how ecological character displacement affects resource densities and food-web stability.
 
-I start by examing the effects of character displacement on the density of resources at equilibrium, which will also set the stage for a later analysis of food-web stability.
-In particular, I examine two different foraging scenarios: one where consumers forage for both resources simultaneously (e.g.
-@MacArthur1972) and one where they cannot because resources occur in distinct habitats (e.g.
-@Lawlor1976; @McCann2005).
-I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, defined as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
+<!-- Explain analytical methods -->
 
-### Consumers forage for both resources simultaneously
+I examine the effects of this inevitable divergence in attack rates on food-web dynamics by (i) deriving analytical expressions for the relationship between attack rates and food-web dynamics and (ii) simulating the effects of competition on the eco-evolutionary dynamics of consumers and resources.
+To gain analytical insight, I assume that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
+While this scenario is arguably simplistic, it allows me to isolate the effects of ecological character displacement.
+<!--, rather than any dependence on inherent asymmetries between consumers or resources.-->
 
-An inherent assumption in MacArthur's model of resource competition is that consumers can forage for both resources simultaneously.
-In essence, this implies that both resources occur within the same habitat.
-To gain analytical insight to this scenario, I assume that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
-While this scenario is arguably simplistic, it isolates the effect of character displacement rather than any dependence on inherent differences in consumers or resources.
-I show that when both consumers and resources are present that the density of resources at equilibrium are equivalent ($R=R_1=R_2$) and are determined by the following equation (*Mathematica* file with derivation available upon request):
+<!-- Explain simulations -->
 
-$$\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
-Note that a key determinant of resource density is the consumer's total attack rate, $a_{ii}+a_{ij}$.
-Since the total attack rate is in the denominator, this implies that resource densities will decrease if evolution favors an increase in total attack rate.
-I assume that consumers are not 'Darwinian demons' (@Law1979) and thus the evolution of their attack rates are subject to certain constraints.
-I modelled these constraints such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff (@Sargent2006).
+To simulate eco-evolutionary dynamics, I used an Adaptive Dynamics approach.
+At each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate on one resource by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability.
+The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006].
 This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$, green line in Fig.
@@
 \ref{fig:Tradeoff}).
+Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer.
+If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state.
+I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g.
+$\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
 
-```{r Tradeoff, echo=FALSE, fig.cap="\\label{fig:Tradeoff}Tradeoff forms for consumer attack rates."}
+<!-- Redundant, removing for now
+In what follows, I examine the ecological consequences of character displacement under three different attack-rate tradeoffs (n>1, n=1, n<1) and in three different foraging scenarios: (i) resources occur in the same habitat and consumers exhibit a linear functional response (e.g.
+@MacArthur1972); (ii) resources occur in different habitats and consumers exhibit a linear functional response (e.g.
+@Lawlor1976); (iii) resources occur in distinct habitats and consumers exhibit a more realistic functional response [@McCann2005].
+I show that the consequences of character displacement for food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
+-->
+
+```{r Tradeoff, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:Tradeoff}Tradeoff forms for consumer attack rates."}
 tradeoff_plot <- bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
@@
 
-Previous work has shown that consumers undergo divergent character displacement regardless of the shape of the tradeoff (@Lawlor1976; @Abrams1986).
-Interestingly, I find that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource density (Fig.
+## Consumers can forage for both resources simultaneously
+
+<!-- Functional response equation and illustration of assumptions already given in previous section -->
+
+<!-- Redundant, removing for now
+The evolution of consumer attack rates in this model has been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. -->
+When both consumers and resources are present, the density of resources at equilibrium are equivalent ($\hat R = \hat R_1 = \hat R_2$) and are determined by the following equation (derivation in supplementary material):
+
+$$\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
+A key determinant of resource density in this model is the consumer's total attack rate, $a_{ii}+a_{ij}$.
+Therefore, the ecological consequences of character displacement depend on how the tradeoff influences the evolution of consumer attack rates.
+
+My simulations show that the shape of the attack-rate tradeoff qualitatively affects the relationship between character displacement and resource density (Fig.
 \ref{fig:MacArthur_Resources}).
@@
 \ref{fig:MacArthur_Resources}A).
-Note that detecting this effect in nature would likely be difficult, since a concave-down tradeoff results in relatively small displacement relative to other tradeoff shapes (@Lawlor1976; green points in Fig.
-\ref{fig:MacArthur_Resources}A,B).
 When the tradeoff is concave up (orange lines), character displacement suppresses resource densities due to the increase in total attack rates (Fig.
 \ref{fig:MacArthur_Resources}A,B).
-Although the equation I derived for resource density was for the scenario where both consumers and both resources are present, it predicts well the density of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
+Although the equation I derived for resource density was for the scenario where both consumers and both resources are present, it accurately predicts the density of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
 \ref{fig:MacArthur_Resources}B).
@@
 
-```{r MacArthur_Resources, echo=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers can forage for both resources simultaneously."}
+The effect of character displacement on resource density closely matches its impact on food-web stability (Fig.
+\ref{fig:Stability}).
+For example, when character displacement results in resource suppression (orange), there is a corresponding decrease in stability.
+Similarly, if character displacement does not influence resource densities (blue), then there is no corresponding effect on stability.
+Curiously, this match between impact on resources and stability breakdowns with the concave down tradeoff (green).
+<!--This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (Fig.
+\ref{fig:Stability}).
+The effect of this eco-evolutionary feedback can be seen by comparing initial (small points) with evolved strategies (large points).
+For example, when the tradeoff is concave up (orange points), the three-species system (small triangle) is initially more stable than the four-species system (small circle), but this relationship reverses by the time consumers reach their evolutionary stable strategy.-->
+
+```{r MacArthur_Resources, include=FALSE, echo=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers can forage for both resources simultaneously."}
 Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) + size=5,
   #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
@@
 
-Mac_resources <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
+Mac_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
   ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) + size=5,
   #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
@@
 
-### Consumers *cannot* forage for resources simultaneously
+```{r Stability, include=FALSE, echo=FALSE, fig.cap="\\label{fig:Stability}Effect of character displacement on local stability when consumers can (A) and cannot (B) forage for both resources simultaneously.
+Small and large points correspond to initial and evolved strategies, respectively."}
+Stability_Mac <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="black", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
 
-Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (e.g.
-@Taper1985; @Abrams1986).
-This assumption may be valid for some foraging scenarios (e.g.
-Darwin's finches foraging for seeds); however, in many situations consumers forage for resources that occur in different habitats, and thus cannot forage for both simultaneosly.
+Stability_LS <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression("Stability ("-lambda)) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_plots <- plot_grid(Stability_Mac + theme(legend.position = "none"), Stability_LS + theme( axis.title.y = element_blank()),
+                             ncol = 2, labels = "AUTO", rel_widths = c(0.75,1))
+Stability_plots
+```
+
+```{r MacArthur Model, echo=FALSE, fig.cap="New MacArthur model figure"}
+plot_macarthur_model <- plot_grid(
+  tradeoff_plot + theme(legend.position = "none"), 
+  Mac_attack  + theme(legend.position = "none"), 
+  Mac_resources  + theme(legend.position = "none"),
+  Stability_Mac  + theme(legend.position = "none"), ncol=2, align='hv')
+
+save_plot(filename="Figure_ECD_MacArthur.pdf", plot=plot_macarthur_model, base_height = 6, base_width = 6)
+```
+
+```{r MacArthur_Plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}Fill in."}
+knitr::include_graphics("Figure_ECD_MacArthur.pdf")
+```
+
+## Consumers *cannot* forage for both resources simulanteously
+
+<!-- Functional response equation and illustration of assumptions -->
+
 The only character displacement model that I am aware of that modeled resources in different habitats was one examined by @Lawlor1976.
@@
 As with attack rates, I assume that consumer habitat preferences are mirror images of each other ($w=w_{11}=w_{22}$).
+In @Lawlor1976's detailed evolutionary analysis of this model, they again always observe divergent character displacement.
 
-@Lawlor1976 found that consumers still underwent divergent character displacement regardless of whether consumers could or could not forage for resources simultaneously.
-I find that the foraging scenario qualitatively affects the relationship between character displacement and resource densities.
-To demonstrate this, note that again resource densities are equivalent at equilibrium ($R=R_1=R_2$ when both consumers and resources present), but are now determined by the following equation (*Mathematica* file with derivation available upon request):
+<!-- Effect on attack rates, resource densities, and stability -->
+
+To examine the ecological consequences of this model, I again derive the equilibrium solution when both consumers and both resources are present.
+Again, resource densities are equivalent at this equilibrium ($\hat R = \hat R_1 = \hat R_2$ when both consumers and resources present), but are now determined by the following equation (derivation in supplementary material):
 
@@
 This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g.
-$w>0.5$ and $a_{ii}>a_{ij}$), then character displacement will always increase the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the tradeoff (Fig.
+$w>0.5$ and $a_{ii}>a_{ij}$), then the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase, regardless of the tradeoff (Fig.
 \ref{fig:LS_Resources}A).
@@
 \ref{fig:LS_Resources}B), because character displacement tends to reach a constraint of complete specialization.
-
-These results do lead to an interesting prediction that could be tested either experimentally or with field data.
-If consumers are competing for resources that occur in different habitats, then greater character displacement corresponds to a greater decrease in resource densities (line trajectories in Fig.
-\ref{fig:LS_Resources}B).
-This prediction likely only applies to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
-
 It is worth noting that in this foraging scenario, resource densities are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
@@
 -->
+
+The effect of character displacement on resource densities qualitatively matches its negative effect of food-web stability (Fig.
+\ref{fig:Stability}).
+This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (compare initial (small points) with evolved strategies (large points), Fig.
+\ref{fig:Stability}).
+For example, when the tradeoff is concave up (orange), then the four-species (small circle) food web actually starts off as less stable than the three-species one (small triangle), but the eco-evolutionary dynamics cause a switch in the pattern once they reach their evolutionary stable strategies (large points).
 
@@
 
-LS_resources <- cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+LS_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
   ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color=factor(n))) + 
@@
 
-The effect of the foraging scenario on the relationship between character displacement and resource densities appears to be quite general.
+```{r LS Model, echo=FALSE, fig.cap="LS model figure"}
+plot_LS_model <- plot_grid(
+  #tradeoff_plot + theme(legend.position = "none"), 
+  LS_attack  + theme(legend.position = "none"), 
+  LS_resources  + theme(legend.position = "none"),
+  Stability_LS  + theme(legend.position = "none"), ncol=2, align='hv')
+
+save_plot(filename="Figure_ECD_LS.pdf", plot=plot_LS_model, base_height = 6, base_width = 6)
+```
+
+```{r LS Plot, echo=FALSE}
+knitr::include_graphics("Figure_ECD_LS.pdf")
+```
+
+## Adding a more realistic functional response
+
+<!-- Functional response equation and illustration of assumptions -->
+
+Thus far I have consider a simple type I functional response of consumers foraging for resources in either the same or distinct habitats.  
 For example, even if I consider a more realistic consumer functional response (derived by @McCann2005):
@@
 
+<!-- Effect on attack rates, resource densities, and stability -->
+
 I still observe the same qualitative relationship between character displacement and resource suppression (Fig.
@@
 $$\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e-hm}$$
-
-<!-- This more realistic consumer functional response does actually make an interesting evolutionary prediction that differs from the scenario where consumer feeding rate scales linearly with resource density.
-Specifically, a consumer's ESS in the absence of competition will remain slightly more specialized on the resources that occur in their preferred habitats.
-This is in contrast to the prediction when feeding rates scale linearly with resource density, where consumers evolve to be slightly more specialized on resources occurring in their non-preferred habitats.
-An important caveat for these predictions is that my model does not allow habitat preferences to evolve, which would likely alter these predictions.
--->
-
-## Effect of character displacement on food-web stability
-
-Of the prior models of character displacement, only @Lawlor1976 examine the effect of character displacement on the ecological stability of the system.
-They conducted a detailed theoretical analysis of three different metrics of stability: (i) stable coexistence of consumers, defined as the mutual invasibility of each consumer; (ii) global stability of consumers, defined as the continued coexistence of both consumers when resource dynamics are perturbed; and (iii) local stability, defined as the rate at which the four-species food web returned to equilibrium after a small perturbation to species abundances.
-Note that the first two metrics correspond to the stable coexistence of consumers, while the third metric measures the stability of the entire food web.
-@Lawlor1976 found that character displacement enhanced the stable coexistence of consumers, yet had a small negative effect on the local stability of the food web.
-
-@Lawlor1976's work gave fundamental insight to the effect of character displacement on ecological stability; however, none of their stability analyses compared food webs with and without a competing consumer.
-Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement (@Schluter2000; @Schluter1992).
-In fact, the first two metrics of stability can only be assessed when both consumers are present because it examines the effect of divergence on coexistence, which requires both consumers to be present.
-In contrast, local stability can be compared in food webs with and without a competing consumer, and thus provides a useful benchmark for examining the effects of character displacement on food-web stability.
-
-Below, I plot the results from the same simulations I did previously except now local stability ($-1\times\operatorname{Re}(\lambda_{max})$) is on the y-axis (Fig.
-\ref{fig:Stability}).
-In general, I find that resource supression goes hand-in-hand with local stability, a tendency that has been noted by others (@Murdoch2003; @McCann2011).
-Whenever character displacement results in resource suppression, I observe a decrease in local stability (Fig.
-\ref{fig:Stability}).
-This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (compare initial (small points) with evolved strategies (large points), Fig.
-\ref{fig:Stability}).
-Note that I reach the same fundamental conclusions when I incorporated a more realistic functional response for the consumers (Fig.
-\ref{fig:Stability_McCann} in Appendix).
-
-```{r Stability, echo=FALSE, fig.cap="\\label{fig:Stability}Effect of character displacement on local stability when consumers can (A) and cannot (B) forage for both resources simultaneously.
-Small and large points correspond to initial and evolved strategies, respectively."}
-Stability_Mac <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend=FALSE) +
-  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression(-1%*%"Re"(lambda[max]))) +
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Stability_LS <- cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend=FALSE) +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression(-1%*%"Re"(lambda[max]))) +
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Stability_plots <- plot_grid(Stability_Mac + theme(legend.position = "none"), Stability_LS + theme( axis.title.y = element_blank()),
-                             ncol = 2, labels = "AUTO", rel_widths = c(0.75,1))
-Stability_plots
-```
 
@@
 Thus, character displacement is capable of pushing food webs to a point where they are locally unstable.
+(it simple depends on the choice of parameter, I could put a situation where evolution pushes it toward local instability, or even the hopf bifurcation.)
 
-## Predictions for threespine stickleback
+# Discussion
+
+I find that the foraging scenario qualitatively affects the relationship between character displacement and resource densities.
+
+Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously [e.g.
+@Taper1985; @Abrams1986].
+This assumption may be valid for some foraging scenarios (e.g.
+Darwin's finches foraging for seeds); however, in many situations consumers forage for resources that occur in different habitats, and thus cannot forage for both simultaneously [@McCann2005].
+
+The effect of the foraging scenario on the relationship between character displacement and resource densities appears to be quite general.
+
+Note that detecting this effect in nature would likely be difficult, since a concave-down tradeoff results in relatively small displacement relative to other tradeoff shapes (@Lawlor1976; green points in Fig.
+\ref{fig:MacArthur_Resources}A,B).
+
+Previous work has shown that consumers undergo divergent character displacement regardless of the shape of the tradeoff (@Lawlor1976; @Abrams1986).
+
+Important point: Criteria 5 in @Schluter1992 does not make sense since character displacement usually alters resource distributions. Think about this more.
+
+<!-- This more realistic consumer functional response does actually make an interesting evolutionary prediction that differs from the scenario where consumer feeding rate scales linearly with resource density.
+Specifically, a consumer's ESS in the absence of competition will remain slightly more specialized on the resources that occur in their preferred habitats.
+This is in contrast to the prediction when feeding rates scale linearly with resource density, where consumers evolve to be slightly more specialized on resources occurring in their non-preferred habitats.
+An important caveat for these predictions is that my model does not allow habitat preferences to evolve, which would likely alter these predictions.
+-->
+
+@Lawlor1976 found that consumers still underwent divergent character displacement regardless of whether consumers could or could not forage for resources simultaneously.
+
+These results do lead to an interesting prediction that could be tested either experimentally or with field data.
+If consumers are competing for resources that occur in different habitats, then greater character displacement corresponds to a greater decrease in resource densities (line trajectories in Fig.
+\ref{fig:LS_Resources}B).
+This prediction likely only applies to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
+
+In general, I find that resource supression goes hand-in-hand with local stability, a tendency that has been noted by others (@Murdoch2003; @McCann2011).
+
+<!-- Looking for this effect in nature -->
 
@@
 
-### Other relevant points
+<!-- Caveats -->
 
@@
 
-## Appendix: More realistic functional response
+# Acknowledgements
+
+The content of this paper was greatly enhanced by numerous conversations with Seth Rudman.
+Sally Otto provided much help in early analyses of the mathematical model and I would not have been able to take the theory as far as I did without her guidance and her pushing me.
+
+Thank Ben Gilbert for his comment during a discussion of the model that pushed me to take a comparative approach that looked across different model structures to examine the generality of these results.
+
+# References
+
+<div id="refs"></div> <!-- this code inserts references wherever you want rather than at the end of the document -->
+
+\newpage
+
+# Appendix: More realistic functional response
 
@@
 
-McCann_resources <- cbind.data.frame(tradeoff_df, McCann_2C_2R_tradeoff_dynamics) %>%
+McCann_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
   ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
@@
 ```{r Stability_McCann, echo=FALSE, fig.cap="\\label{fig:Stability_McCann}Effect of character displacement on local stability when consumers exhibit a more realistic functional response."}
-Stability_McCann <- cbind.data.frame(tradeoff_df, McCann_2C_2R_tradeoff_dynamics) %>%
+Stability_McCann <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
   ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color=factor(n))) + 

### 2019-01-23 manuscript_ECD_model.tex
update repo with many changes

--- 2019-01-23 ECD_Supp_Mat.tex
@@
 
-\emph{Keywords}: competition; eco-evolutionary dynamics; ecological
-speciation; three-spine stickleback; zooplankton
+\emph{Keywords}: competition; eco-evolutionary dynamics;
+consumer-resource interactions; adaptive radiation; ecological community
 
@@
 \section{Abstract}\label{abstract}
-
-Ecological character displacement is a key driver of diversification and
-adaptive radiation; however, there has been little work examining the
-consequences of character displacement on the dynamics of ecological
-communities. Given the growing recognition for the role of evolution in
-structuring ecological communities, understanding how major evolutionary
-processes impact ecological dynamics is increasingly important. Here, we
-build a mathematical model to predict how exploitative competition
-between two consumers for two resources affects the evolution of
-consumer attack rates, and, in turn, consumer-resource population
-dynamics. Our model suggested that increasing ecological character
-displacement decreases the equilibrium density of resources, and with
-sufficient divergence, results in more oscillatory (destabilizing)
-consumer-resource dynamics. These results suggest that ecological
-character displacement can eventually have a negative feedback on
-food-web persistence in the absence of other processes.
 
@@
 simply driven by ecological differences, but creates an eco-evolutionary
-feedback that drives further differentiation. This insight was clearly
-by theoretical models that highlighted the importance of explicitly
-including resource dynamics as a mediator of competition in driving
-evolutionary change (Lawlor and Smith 1976; P A Abrams 1986; Taper and
-Chase 1985).
+feedback that drives further differentiation. This key insight was made
+by theoretical models that explicitly included resource dynamics as a
+mediator of competition in driving evolutionary change (Lawlor and Smith
+1976; P A Abrams 1986; Taper and Chase 1985).
 
-Although character displacement is the result of an eco-evolutionary
-feedback, the ecological consequences of this process have received
-surprisingly little attention (Germain et al. 2018). Perhaps this is
-because early theoretical work suggested that character displacement was
-a potent stabilizing force in ecological communities (Lawlor and Smith
-1976). This ecological outcome is also an intuitive one ---character
-divergence reduces interspecific competition between consumers,
-promoting coexistence. However, inferences about the effect of character
-displacement, like the process itself, can only be made when comparing
-communities in the presence and absence of competition (criteria \#3 for
-character displacement, D Schluter and McPhail 1992; Dolph Schluter
-2000). Current inferences about the stabilizing effects of consumer
-divergence are only comparisons about communities where there are two
-consumers, and even though, the focus is only on the stable coexistence
-of consumer species.
-
-The consequence of this eco-evolutionary feedback for consumer
-coexistence has been well characterized (Lawlor and Smith 1976;
-Germain's work). Ecological character displacement has typically been
-considered to be a stabilizing factor in communities (Lawlor and Smith
-1976). This is because divergence in foraging traits is thought to
-reduce the strength of interspecific competition between consumers.
-Reducing the strength of interspecific competition relative to
-intraspecific competition is a stabilizing force in consumer-resource
-systems. Consumer specialization on distinct resources suggests that
-there are niche differences between consumers. Mathematically, this is
-also often assumed to be the case, because using techniques such as
-evolutionary invasion analysis describe the ability for a consumer with
-a mutant allele to increase in abundance when rare, a demographic
-signature of niche differentiation. Therefore, if both consumers are
-able to invade when rare, then this suggests that they can coexist.
-
-Character displacement involves evolutionary shifts in character states
-in response to competition (D Schluter and McPhail 1992; Dolph Schluter
-2000). Theoretically, this is determined by comparing character changes
-in the presence and absence of a competitor. Empirically, this is
-determined by meeting several criteria that rule out other possible
-explanations (D Schluter and McPhail 1992; Dolph Schluter 2000).
-Importantly, current inferences about the effect of character
-displacement on community stability do not meet this critera. For
-example, the inference that character displacement enhances species
-coexistence is only made in a scenario where consumers are competing.
-This type of stability is important, but it does not enable us to
-examine the effect of character displacement on community stability. As
-a consequence, I argue that the effect of character displacement on
-community stability remains an open question.
-
-Indeed, ecological character displacement often results in the evolution
-of characters that enhance a consumer's ability to capture resources,
-which tends to destabilize consumer-resource interactions (Murdoch et
-al. 2003; McCann 2011). This occurs because increases in a consumer's
-attack rate results in the resource being suppressed well below its
-carrying capacity, which if sufficient enough, can generate
-oscillations, and hence less stable, consumer-resource dynamics
-(Rosenzweig 1971; Murdoch et al. 2003). Experimental decreases in
-interaction strength between consumers and resources has been shown to
-promote coexistence between consumers and resources (Luckinbill 1973).
-Moreover, there is broad support that specialized consumers, since they
-are tightly coupled to their resources, tend to exhibit
-consumer-resource cycles, whereas this is not the case for generalists
-(Murdoch et al. 2002). Therefore, although evolutionary responses to
-exploitative competition may reduce the strength of interactions between
-consumers, they ultimately increase the strength of interactions with
-their resources which can destabilize the direct consumer-resource
-interaction (rather than the indirect consumer-consumer interaction).
-While prior models of character displacement have included more
-realistic functional responses and spatial relationships between
-consumers and resources, they have focused more on whether these factors
-alter patterns of character displacement rather than the effects of
-character displacement on the stability of the ecological community. As
-a consequence, it is currently unclear whether ecological character
-displacement has a net stabilizing or destabilizing effect on ecological
-communities.
+Although explicit models of resources gave key insights to the evolution
+of character displacement, the ecological feedback onto
+consumer-resource dynamics have received surprisingly little attention.
+This is likely because character displacement has been primarily studied
+through the lens of competition and coexistence theory {[}Lawlor1976;
+Germain2018{]}. For example, early theoretical work by Lawlor and Smith
+(1976) showed that ecological character displacement promotes
+coexistence by favoring specialized consumers that experience reduced
+interspecific competition. On the other hand, specialization often
+increases a consumer's ability to capture a specific resource. A well
+known consequence of increased foraging efficiency is resource
+suppression, which if sufficient enough, can generate oscillations, and
+hence less stable, consumer-resource dynamics {[}Rosenzweig 1971;
+Murdoch et al. 2003; McCann 2011{]}. NEED A LAST SENTENCE TO WRAP IT UP,
+HIGHLIGHTING HOW A FOOD-WEB PERSPECTIVE MAY YIELD NEW ECOLOGICAL
+INSIGHTS AND GIVE A CONTRASTING PERSPECTIVE TO COEXISTENCE THEORY.
+PERHAPS A I NEED TO INCLUDE A SENTENCE MID-WAY THROUGH SUGGESTING THAT A
+FOOD-WEB PERSPECTIVE MAY GIVE QUALITATIVELY DIFFERENT INSIGHTS. IT MAY
+BE NICE TO BRING IN THE SPATIAL COUPLING WORK OF MCCANN SUGGESTING THAT
+SPECIALIZATION CAN COUPLE FOOD-WEBS, BUT THIS WAS ONLY WITH ONE TOP
+PREDATOR (I.E. NO CONSUMERS\ldots{}LOOK AT OLD TEXT FOR WORDING
+SUGGESTIONS.)
 
@@
 examine how ecological character displacement affects consumer-resource
-dynamics in a food web context. I address two question: (1) How does
+dynamics in a food-web context. I address two questions: (1) How does
 ecological character displacement affect resource abundances? (2) How
-does character displacement affect food-web stability? I show that the
+does character displacement affect food-web stability? To test the
+generality of these ecological effects, I examined three different
+foraging scenarios: (i) resources occur in the same habitat and
+consumers exhibit a linear functional response (e.g. MacArthur (1972);
+Lawlor and Smith (1976)); (ii) resources occur in different habitats and
+consumers exhibit a linear functional response (e.g. Lawlor and Smith
+(1976)); (iii) resources occur in distinct habitats and consumers
+exhibit a more realistic functional response (K S McCann, Rasmussen, and
+Umbanhowar 2005). I also explore the contingency of these effects to
+different assumptions about evolutionary trade-offs in consumer foraging
+traits. I focus on the ecological side of these dynamics because the
+evolutionary aspects have been well characterized (Lawlor and Smith
+1976; Taper and Chase 1985; P A Abrams 1986). I show that the
 consequences of character displacement on food-web dynamics are not
-intuitive and depend critically on the foraging scenario and tradeoff in
-consumer attack rates.
+always intuitive and depend on the foraging scenario and tradeoff in
+consumer traits.
 
@@
 
-\subsection{Model of resource
-competition}\label{model-of-resource-competition}
+\subsection{Underlying Food-web
+Dynamics}\label{underlying-food-web-dynamics}
 
@@
 
-\[\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2\]
-\[\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2\]
-\[\frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1\]
-\[\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2\]
+\[\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2,\]
+\[\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2,\]
+\[\frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1,\]
+\[\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2,\]
 
@@
 extensions) have been analyzed in detail (Lawlor and Smith (1976); P A
-Abrams (1986)). The general result of these analyses is that divergent
-character displacement in attack rates is an inevitable outcome of
-resource competition.
+Abrams (1986)), with the general result being divergent character
+displacement. I say consumers have undergone divergent character
+displacement if there evolved attack rates are more specialized, defined
+as \(|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|\), when evolving with a
+competing consumer. I focus here on the ecological consequences of this
+adaptive divergence.
 
-Here, I examine the effects of this inevitable divergence in attack
-rates on food-web dynamics. In particular, I examine how divergence in
-attack rates affect resource densities as well as the local stability of
-both resources and consumers. I do this by (i) deriving analytical
-expressions for the relationship between attack rates and food-web
-dynamics in different foraging scenarios and (ii) simulating the effects
-of competition on the eco-evolutionary dynamics of consumers and
-resources. For these simulations, I used an Adaptive Dynamics approach.
-At each evolutionary time step, I created a mutant consumer by randomly
-choosing a consumer and modifying its attack rate on one resource by
-either subtracting or adding a small constant with equal probability.
-The mutant's attack rate on the other resource was determined by the
-shape of the tradeoff (details below). Assuming this mutant consumer is
-rare, I then determined whether the mutant had higher relative fitness
-than the resident consumer, and thus could invade and replace the
-resident consumer. If the mutant was able to invade, I updated the
-attack rate of the resident consumer to the mutant attack rate and
-allowed consumer and resource dynamics to reach a steady state. I then
-repeated the simulation up to 10,000 times, which was sufficient for
-consumers to either reach an evolutionary stable strategy (ESS, Smith
-and Price (1973)) or an evolutionary limit (e.g.
-\(\frac{a_{ii}}{a_{ii}+a_{ij}}\) is constrained to a maximum of 1 and
-minimum of 0).
+To investigate the ecological consequences, I studied the effect of
+adaptive divergence in consumers on resource densities and the stability
+of the entire food web at equilibrium. An equilibrium is reached when
+the four rates of change in the above equations are 0, and solving the
+system at this point gives equilibrium densities for each resource
+(\(\hat R_1,\hat R_2\)) and consumer (\(\hat C_1,\hat C_2\)). At the
+coexistence equilibrium (i.e., where
+\(\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0\)), I calculated resource
+densities and performed a linear stability analysis (Otto and Day 2007).
+This stability analysis derives the dominant (largest in absolute value)
+eigenvalue, \(\lambda\), which determines whether (and how readily) the
+food web, when perturbed a small amount, will return to equilibrium (see
+supplementary material for derivation). Together, these two measures
+indicate how ecological character displacement affects resource
+densities and food-web stability.
 
-\subsection{Resource densities}\label{resource-densities}
+I examine the effects of this inevitable divergence in attack rates on
+food-web dynamics by (i) deriving analytical expressions for the
+relationship between attack rates and food-web dynamics and (ii)
+simulating the effects of competition on the eco-evolutionary dynamics
+of consumers and resources. To gain analytical insight, I assume that
+resources are equivalent (\(r=r_1=r_2\) and \(K=K_1=K_2\)) as well as
+consumers (\(e=e_{11}=e_{12}=e_{21}=e_{22}\); \(m=m_1=m_2\)), except
+that consumer attack rates are mirror images of each other
+(\(a_{ii}=a_{11}=a_{22}\); \(a_{ij}=a_{12}=a_{21}\)). While this
+scenario is arguably simplistic, it allows me to isolate the effects of
+ecological character displacement.
 
-I start by examing the effects of character displacement on the density
-of resources at equilibrium, which will set the stage for a later
-analysis of food-web stability. In particular, I examine two different
-foraging scenarios: one where consumers forage for both resources
-simultaneously (e.g. MacArthur (1972)) and one where they cannot because
-resources occur in distinct habitats (e.g. Lawlor and Smith (1976); K S
-McCann, Rasmussen, and Umbanhowar (2005)). I say consumers have
-undergone divergent character displacement if there evolved attack rates
-are more specialized, defined as \(|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|\),
-when evolving with a competing consumer.
-
-\textbf{Consumers forage for both resources simultaneously}: An inherent
-assumption in MacArthur's model of resource competition is that
-consumers can forage for both resources simultaneously. In essence, this
-implies that both resources occur within the same habitat. To gain
-analytical insight to this scenario, I assume that resources are
-equivalent (\(r=r_1=r_2\) and \(K=K_1=K_2\)) as well as consumers
-(\(e=e_{11}=e_{12}=e_{21}=e_{22}\); \(m=m_1=m_2\)), except that consumer
-attack rates are mirror images of each other (\(a_{ii}=a_{11}=a_{22}\);
-\(a_{ij}=a_{12}=a_{21}\)). While this scenario is arguably simplistic,
-it isolates the effect of character displacement rather than any
-dependence on inherent differences in consumers or resources. I show
-that when both consumers and resources are present that the density of
-resources at equilibrium are equivalent (\(R=R_1=R_2\)) and are
-determined by the following equation (\emph{Mathematica} file with
-derivation available upon request):
-
-\[\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}\] Note that a key
-determinant of resource density is the consumer's total attack rate,
-\(a_{ii}+a_{ij}\). Since the total attack rate is in the denominator,
-this implies that resource densities will decrease if evolution favors
-an increase in total attack rate. I assume that consumers are not
-`Darwinian demons' (Law (1979)) and thus the evolution of their attack
-rates are subject to certain constraints. I modelled these constraints
-such that \(\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1\), where
-\(A\) is the total investment in attack rates and \(n\) describes the
-shape of the tradeoff (Sargent and Otto (2006)). This function has the
-useful property that it differentiates between cases where intermediate
-combinations of \(a_{ii}\) and \(a_{ij}\) are higher, on average, than
-the extremes (when \(n>1\), green line in Fig. \ref{fig:Tradeoff}) or,
-conversely, where the two extremes are higher, on average, than
-intermediate investments (when \(n<1\), orange line Fig.
-\ref{fig:Tradeoff}). When \(n=1\), the tradeoff function is linear, and
-all combinations of \(a_{ii}\) and \(a_{ij}\) have the same total attack
-rate (blue line in Fig. \ref{fig:Tradeoff}).
+To simulate eco-evolutionary dynamics, I used an Adaptive Dynamics
+approach. At each evolutionary time step, I created a mutant consumer by
+randomly choosing a consumer and modifying its attack rate on one
+resource by either subtracting or adding a small constant (0.001 in the
+following simulations) with equal probability. The mutant's attack rate
+on the other resource was determined by a tradeoff, such that
+\(\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1\), where \(A\) is the
+total investment in attack rates and \(n\) describes the shape of the
+tradeoff (Sargent and Otto 2006). This function has the useful property
+that it differentiates between cases where intermediate combinations of
+\(a_{ii}\) and \(a_{ij}\) are higher, on average, than the extremes
+(when \(n>1\), green line in Fig. \ref{fig:Tradeoff}) or, conversely,
+where the two extremes are higher, on average, than intermediate
+investments (when \(n<1\), orange line Fig. \ref{fig:Tradeoff}). When
+\(n=1\), the tradeoff function is linear, and all combinations of
+\(a_{ii}\) and \(a_{ij}\) have the same total attack rate (blue line in
+Fig. \ref{fig:Tradeoff}). Assuming the mutant consumer was rare, I then
+determined whether the mutant had higher relative fitness than the
+resident consumer, and thus could invade and replace the resident
+consumer. If the mutant was able to invade, I updated the attack rate of
+the resident consumer to the mutant attack rate and allowed consumer and
+resource dynamics to reach a steady state. I then repeated the
+simulation up to 10,000 times, which was sufficient for consumers to
+either reach an evolutionary stable strategy (ESS, Smith and Price 1973)
+or an evolutionary limit (e.g. \(\frac{a_{ii}}{a_{ii}+a_{ij}}\) is
+constrained to a maximum of 1 and minimum of 0).
 
@@
 \centering
-\includegraphics{ECD_Supp_Mat_files/figure-latex/Tradeoff-1.pdf}
+\includegraphics{manuscript_ECD_model_files/figure-latex/Tradeoff-1.pdf}
 \caption{\label{fig:Tradeoff}Tradeoff forms for consumer attack rates.}
@@
 
-Previous work has shown that consumers undergo divergent character
-displacement regardless of the shape of the tradeoff (Lawlor and Smith
-(1976); P A Abrams (1986)). Interestingly, I find that the shape of the
-tradeoff qualitatively affects the relationship between character
-displacement and resource density (Fig. \ref{fig:MacArthur_Resources}).
-For example, if consumer's are constrained by a linear tradeoff (blue
-lines), then there is no net change in total attack rate (Fig.
+\subsection{Consumers Forage For Both Resources
+Simultaneously}\label{consumers-forage-for-both-resources-simultaneously}
+
+When both consumers and resources are present, the density of resources
+at equilibrium are equivalent (\(\hat R = \hat R_1 = \hat R_2\)) and are
+determined by the following equation (derivation in supplementary
+material):
+
+\[\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}\] A key determinant of
+resource density in this model is the consumer's total attack rate,
+\(a_{ii}+a_{ij}\). Therefore, the ecological consequences of character
+displacement depend on how the tradeoff influences the evolution of
+consumer attack rates.
+
+My simulations show that the shape of the attack-rate tradeoff
+qualitatively affects the relationship between character displacement
+and resource density (Fig. \ref{fig:MacArthur_Resources}). For example,
+if consumer's are constrained by a linear tradeoff (blue lines), then
+there is no net change in total attack rate (Fig.
 \ref{fig:MacArthur_Resources}A) and character displacement has no effect
@@
 decreases as consumers diverge (Fig. \ref{fig:MacArthur_Resources}A).
-Note that detecting this effect in nature would likely be difficult,
-since a concave-down tradeoff results in relatively small displacement
-relative to other tradeoff shapes (Lawlor and Smith (1976); green points
-in Fig. \ref{fig:MacArthur_Resources}A,B). When the tradeoff is concave
-up (orange lines), character displacement suppresses resource densities
-due to the increase in total attack rates (Fig.
-\ref{fig:MacArthur_Resources}A,B). Although the equation I derived for
-resource density was for the scenario where both consumers and both
-resources are present, it predicts well the density of resources when a
-single consumer reaches its ESS (triangles on respective colored lines
-in Fig. \ref{fig:MacArthur_Resources}B). This is because a single
+When the tradeoff is concave up (orange lines), character displacement
+suppresses resource densities due to the increase in total attack rates
+(Fig. \ref{fig:MacArthur_Resources}A,B). Although the equation I derived
+for resource density was for the scenario where both consumers and both
+resources are present, it accurately predicts the density of resources
+when a single consumer reaches its ESS (triangles on respective colored
+lines in Fig. \ref{fig:MacArthur_Resources}B). This is because a single
 consumer evolves to be a generalist that has equal attack rates on each
@@
 
+The effect of character displacement on resource density closely matches
+its impact on food-web stability (Fig. \ref{fig:Stability}). For
+example, when character displacement results in resource suppression
+(orange), there is a corresponding decrease in stability. Similarly, if
+character displacement does not influence resource densities (blue),
+then there is no corresponding effect on stability. Curiously, this
+match between impact on resources and stability breakdowns with the
+concave down tradeoff (green).
+
 \begin{figure}
 \centering
-\includegraphics{ECD_Supp_Mat_files/figure-latex/MacArthur_Resources-1.pdf}
+\includegraphics{manuscript_ECD_model_files/figure-latex/MacArthur_Resources-1.pdf}
 \caption{\label{fig:MacArthur_Resources}Effect of character displacement
@@
 
-\textbf{Consumers \emph{cannot} forage for both resources
-simultaneously}: Most models of character displacement have assumed a
-scenario where consumers can forage for both resources simultaneously
-(e.g. Taper and Chase (1985); P A Abrams (1986)). This assumption may be
-valid for some foraging scenarios (e.g.~Darwin's finches foraging for
-seeds); however, in many situations consumers forage for resources that
-occur in different habitats, and thus cannot forage for both
-simultaneosly. The only character displacement model that I am aware of
-that modeled resources in different habitats was one examined by Lawlor
-and Smith (1976). This model takes the same form as the previous model,
-except now the consumer's feeding rate takes the form:
+\includegraphics{Figure_ECD_MacArthur.pdf}
+
+\subsection{\texorpdfstring{Consumers \emph{cannot} forage for both
+resources
+simulanteously}{Consumers cannot forage for both resources simulanteously}}\label{consumers-cannot-forage-for-both-resources-simulanteously}
+
+The only character displacement model that I am aware of that modeled
+resources in different habitats was one examined by Lawlor and Smith
+(1976). This model takes the same form as the previous model, except now
+the consumer's feeding rate takes the form:
 
@@
 habitat preferences are mirror images of each other
-(\(w=w_{11}=w_{22}\)).
+(\(w=w_{11}=w_{22}\)). In Lawlor and Smith (1976)'s detailed
+evolutionary analysis of this model, they again always observe divergent
+character displacement.
 
-Lawlor and Smith (1976) found that consumers still underwent divergent
-character displacement regardless of whether consumers could or could
-not forage for resources simultaneously. I find that the foraging
-scenario qualitatively affects the relationship between character
-displacement and resource densities. To demonstrate this, note that
-again resource densities are equivalent at equilibrium (\(R=R_1=R_2\)
-when both consumers and resources present), but are now determined by
-the following equation (\emph{Mathematica} file with derivation
-available upon request):
+To examine the ecological consequences of this model, I again derive the
+equilibrium solution when both consumers and both resources are present.
+Again, resource densities are equivalent at this equilibrium
+(\(\hat R = \hat R_1 = \hat R_2\) when both consumers and resources
+present), but are now determined by the following equation (derivation
+in supplementary material):
 
@@
 resources that occur in their preferred habitat (e.g. \(w>0.5\) and
-\(a_{ii}>a_{ij}\)), then character displacement will always increase the
-effective attack rate of consumers (\(wa_{ii}+(1-w)a_{ij}\)), regardless
-of the tradeoff (Fig. \ref{fig:LS_Resources}A). Thus, character
-displacement always results in resource suppression if consumers are
-competing for resources that occur in different habitats (Fig.
+\(a_{ii}>a_{ij}\)), then the effective attack rate of consumers
+(\(wa_{ii}+(1-w)a_{ij}\)) will always increase, regardless of the
+tradeoff (Fig. \ref{fig:LS_Resources}A). Thus, character displacement
+always results in resource suppression if consumers are competing for
+resources that occur in different habitats (Fig.
 \ref{fig:LS_Resources}B). Note that the shape of the tradeoff can modify
@@
 (circles in Fig. \ref{fig:LS_Resources}B), because character
-displacement tends to reach a constraint of complete specialization.
-
-These results do lead to an interesting prediction that could be tested
-either experimentally or with field data. If consumers are competing for
-resources that occur in different habitats, then greater character
-displacement corresponds to a greater decrease in resource densities
-(line trajectories in Fig. \ref{fig:LS_Resources}B). This prediction
-likely only applies to comparisons within species where there is likely
-little difference in the shape of the tradeoff among populations.
-
-It is worth noting that in this foraging scenario, resource densities
-are consistently higher at the single consumer ESS compared to the
+displacement tends to reach a constraint of complete specialization. It
+is worth noting that in this foraging scenario, resource densities are
+consistently higher at the single consumer ESS compared to the
 predictions I derived for when both consumers are present (deviation of
@@
 
+The effect of character displacement on resource densities qualitatively
+matches its negative effect of food-web stability (Fig.
+\ref{fig:Stability}). This is not simply a consequence of having an
+additional consumer in the system, but emerges from the eco-evolutionary
+feedback between character displacement and resource suppression
+(compare initial (small points) with evolved strategies (large points),
+Fig. \ref{fig:Stability}). For example, when the tradeoff is concave up
+(orange), then the four-species (small circle) food web actually starts
+off as less stable than the three-species one (small triangle), but the
+eco-evolutionary dynamics cause a switch in the pattern once they reach
+their evolutionary stable strategies (large points).
+
 \begin{figure}
 \centering
-\includegraphics{ECD_Supp_Mat_files/figure-latex/LS_Resources-1.pdf}
+\includegraphics{manuscript_ECD_model_files/figure-latex/LS_Resources-1.pdf}
 \caption{\label{fig:LS_Resources}Effect of character displacement on
@@
 
-The effect of the foraging scenario on the relationship between
-character displacement and resource densities appears to be quite
-general. For example, even if I consider a more realistic consumer
-functional response (derived by K S McCann, Rasmussen, and Umbanhowar
-(2005)):
+\includegraphics{Figure_ECD_LS.pdf}
+
+\subsection{Adding A More Realistic Functional
+Response}\label{adding-a-more-realistic-functional-response}
+
+Thus far I have consider a simple type I functional response of
+consumers foraging for resources in either the same or distinct
+habitats.\\
+For example, even if I consider a more realistic consumer functional
+response (derived by K S McCann, Rasmussen, and Umbanhowar (2005)):
 
@@
 
-\subsection{Food-web stability}\label{food-web-stability}
-
-Of the prior models of character displacement, only Lawlor and Smith
-(1976) examine the effect of character displacement on the ecological
-stability of the system. They conducted a detailed theoretical analysis
-of three different metrics of stability: (i) stable coexistence of
-consumers, defined as the mutual invasibility of each consumer; (ii)
-global stability of consumers, defined as the continued coexistence of
-both consumers when resource dynamics are perturbed; and (iii) local
-stability, defined as the rate at which the four-species food web
-returned to equilibrium after a small perturbation to species
-abundances. Note that the first two metrics correspond to the stable
-coexistence of consumers, while the third metric measures the stability
-of the entire food web. Lawlor and Smith (1976) found that character
-displacement enhanced the stable coexistence of consumers, yet had a
-small negative effect on the local stability of the food web.
-
-Lawlor and Smith (1976)'s work gave fundamental insight to the effect of
-character displacement on ecological stability; however, none of their
-stability analyses compared food webs with and without a competing
-consumer. Such comparisons are necessary for inferring the effects of
-ecological character displacement, a point that has been made clear in
-the criteria to demonstrate character displacement (Dolph Schluter
-(2000); D Schluter and McPhail (1992)). In fact, the first two metrics
-of stability can only be assessed when both consumers are present
-because it examines the effect of divergence on coexistence, which
-requires both consumers to be present. In contrast, local stability can
-be compared in food webs with and without a competing consumer, and thus
-provides a useful benchmark for examining the effects of character
-displacement on food-web stability.
-
-Below, I plot the results from the same simulations I did previously
-except now local stability
-(\(-1\times\operatorname{Re}(\lambda_{max})\)) is on the y-axis (Fig.
-\ref{fig:Stability}). In general, I find that resource supression goes
-hand-in-hand with local stability, a tendency that has been noted by
-others (Murdoch, Briggs, and Nisbet (2003); Kevin S McCann (2011)).
-Whenever character displacement results in resource suppression, I
-observe a decrease in local stability (Fig. \ref{fig:Stability}). This
-is not simply a consequence of having an additional consumer in the
-system, but emerges from the eco-evolutionary feedback between character
-displacement and resource suppression (compare initial (small points)
-with evolved strategies (large points), Fig. \ref{fig:Stability}). Note
-that I reach the same fundamental conclusions when I incorporated a more
-realistic functional response for the consumers (Fig.
-\ref{fig:Stability_McCann} in Appendix).
-
-\begin{figure}
-\centering
-\includegraphics{ECD_Supp_Mat_files/figure-latex/Stability-1.pdf}
-\caption{\label{fig:Stability}Effect of character displacement on local
-stability when consumers can (A) and cannot (B) forage for both
-resources simultaneously. Small and large points correspond to initial
-and evolved strategies, respectively.}
-\end{figure}
-
 I conducted the previous simulations on a scenario where the food web
@@
 character displacement is capable of pushing food webs to a point where
-they are locally unstable.
+they are locally unstable. (it simple depends on the choice of
+parameter, I could put a situation where evolution pushes it toward
+local instability, or even the hopf bifurcation.)
 
 \section{Discussion}\label{discussion}
+
+I find that the foraging scenario qualitatively affects the relationship
+between character displacement and resource densities.
+
+Most models of character displacement have assumed a scenario where
+consumers can forage for both resources simultaneously (e.g. Taper and
+Chase 1985; P A Abrams 1986). This assumption may be valid for some
+foraging scenarios (e.g.~Darwin's finches foraging for seeds); however,
+in many situations consumers forage for resources that occur in
+different habitats, and thus cannot forage for both simultaneously (K S
+McCann, Rasmussen, and Umbanhowar 2005).
+
+The effect of the foraging scenario on the relationship between
+character displacement and resource densities appears to be quite
+general.
+
+Note that detecting this effect in nature would likely be difficult,
+since a concave-down tradeoff results in relatively small displacement
+relative to other tradeoff shapes (Lawlor and Smith (1976); green points
+in Fig. \ref{fig:MacArthur_Resources}A,B).
+
+Previous work has shown that consumers undergo divergent character
+displacement regardless of the shape of the tradeoff (Lawlor and Smith
+(1976); P A Abrams (1986)).
 
@@
 
-\subsection{Predictions for threespine
-stickleback}\label{predictions-for-threespine-stickleback}
+Lawlor and Smith (1976) found that consumers still underwent divergent
+character displacement regardless of whether consumers could or could
+not forage for resources simultaneously.
+
+These results do lead to an interesting prediction that could be tested
+either experimentally or with field data. If consumers are competing for
+resources that occur in different habitats, then greater character
+displacement corresponds to a greater decrease in resource densities
+(line trajectories in Fig. \ref{fig:LS_Resources}B). This prediction
+likely only applies to comparisons within species where there is likely
+little difference in the shape of the tradeoff among populations.
+
+In general, I find that resource supression goes hand-in-hand with local
+stability, a tendency that has been noted by others (Murdoch, Briggs,
+and Nisbet (2003); Kevin S McCann (2011)).
 
@@
 
-\subsubsection{Other relevant points}\label{other-relevant-points}
-
 Although I used an Adaptive Dynamics approach here, I expect that these
@@
 
-\subsection{Acknowledgement}\label{acknowledgement}
+\section{Acknowledgements}\label{acknowledgements}
 
@@
 
-\subsection{Appendix: More realistic functional
-response}\label{appendix-more-realistic-functional-response}
+Thank Ben Gilbert for his comment during a discussion of the model that
+pushed me to take a comparative approach that looked across different
+model structures to examine the generality of these results.
 
-\begin{figure}
-\centering
-\includegraphics{ECD_Supp_Mat_files/figure-latex/McCann_Resources-1.pdf}
-\caption{\label{fig:McCann_Resources}Effect of character displacement on
-total attack rates (A) and resource densities (B) when consumers exhibit
-a more realistic functional response.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{ECD_Supp_Mat_files/figure-latex/Stability_McCann-1.pdf}
-\caption{\label{fig:Stability_McCann}Effect of character displacement on
-local stability when consumers exhibit a more realistic functional
-response.}
-\end{figure}
+\section{References}\label{references}
 
@@
 
-\hypertarget{ref-Germain2018}{}
-Germain, Rachel M, Jennifer L Williams, Dolph Schluter, and Amy L
-Angert. 2018. ``Moving Character Displacement Beyond Characters Using
-Contemporary Coexistence Theory.'' \emph{Trends Ecol. Evol.} 33 (2):
-74--84.
-
-\hypertarget{ref-Law1979}{}
-Law, Richard. 1979. ``Optimal Life Histories Under Age-Specific
-Predation.'' \emph{Am. Nat.} 114 (3): 399--417.
-
 \hypertarget{ref-Lawlor1976}{}
@@
 Dynamics}. Monographs in Population Biology. Princeton University Press.
+
+\hypertarget{ref-Otto2007}{}
+Otto, Sarah P, and Troy Day. 2007. \emph{A Biologist's Guide to
+Mathematical Modeling in Ecology and Evolution}. Princeton University
+Press.
 
@@
 
+\newpage
+
+\section{Appendix: More realistic functional
+response}\label{appendix-more-realistic-functional-response}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_ECD_model_files/figure-latex/McCann_Resources-1.pdf}
+\caption{\label{fig:McCann_Resources}Effect of character displacement on
+total attack rates (A) and resource densities (B) when consumers exhibit
+a more realistic functional response.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_ECD_model_files/figure-latex/Stability_McCann-1.pdf}
+\caption{\label{fig:Stability_McCann}Effect of character displacement on
+local stability when consumers exhibit a more realistic functional
+response.}
+\end{figure}
+
 \end{document}

### 2019-02-03 manuscript_ECD_model.Rmd
many changes plus working on Discussion content.

@@
     fig_caption: true
-    #citation_package: natbib
+    #citation_package: natbib # why isn't this working?
     template: Rmd_AmNat_MS_template.tex
@@
       in_header: Am-Nat-preamble-latex.tex
-bibliography: references_ECD_supp_mat.bib
+bibliography: references.bib
 biblio-style: amnat.bst
@@
 urlcolor: black
+citecolor: black
 ---
@@
 
-\noindent 1. University of British Columbia, Department of Zoology, Vancouver, BC V6T 1Z4, Canada;
-
-\noindent 2. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;
+\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;
+
+\noindent 2. University of British Columbia, Department of Zoology, Vancouver, BC V6T 1Z4, Canada;
 
@@
 
-*Manuscript elements*: Figure 1, figure 2, figure 3, figure 4. All figures should be printed in color.
+*Manuscript elements*: Figure 1, figure 2, and figure 3. All figures should be printed in color.
 
@@
 
-Although explicit models of resources gave key insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics have received surprisingly little attention.
-This is likely because character displacement has been primarily studied through the lens of competition and coexistence theory [Lawlor1976; Germain2018].
+Although explicit models of resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics have received surprisingly little attention.
+This is likely because the ecological feedback has been primarily studied through the lens of competition and coexistence theory [Lawlor1976; Germain2018; Bassar; McPeek].
 For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition.
@@
 (2) How does character displacement affect food-web stability?
-To test the generality of these ecological effects, I examined three different foraging scenarios: (i) resources occur in the same habitat and consumers exhibit a linear functional response (e.g.
+To test the generality of these ecological effects, I examined three different foraging scenarios: *CONDENSE THIS.
+NO NEED TO GIVE ALL THE DETAILS YET.
+RIGHT NOW IT IS JUST CONFUSING.* (i) resources occur in the same habitat and consumers exhibit a linear functional response (e.g.
 @MacArthur1972; @Lawlor1976); (ii) resources occur in different habitats and consumers exhibit a linear functional response (e.g.
@@
 I also explore the contingency of these effects to different assumptions about evolutionary trade-offs in consumer foraging traits.
-I focus on the ecological side of these dynamics because the evolutionary aspects have been well characterized [@Lawlor1976; @Taper1985; @Abrams1986].
+I focus on the ecological side of these dynamics because the evolutionary aspects have been well characterized [@Lawlor1976; @Taper1985; @Abrams1986; McPeekRECENT].
 I show that the consequences of character displacement on food-web dynamics are not always intuitive and depend on the foraging scenario and tradeoff in consumer traits.
@@
 
-To examine how ecological character displacement affects consumer-resource interactions and food-web dynamics, I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
+To examine how ecological character displacement affects resource abundances and food-web stability<!--consumer-resource interactions and food-web dynamics-->, I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
 
@@
 seed size, see @Taper1985 for an example).
-In the scenario where consumer feeding rates increase linearly with resource density, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$, then this model becomes MacArthur's model of resource competition (@MacArthur1972).
-The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail (@Lawlor1976; @Abrams1986), with the general result being divergent character displacement.
-I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, defined as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
+In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource density, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972].
+The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement.
+I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
 I focus here on the ecological consequences of this adaptive divergence.
@@
 
-I examine the effects of this inevitable divergence in attack rates on food-web dynamics by (i) deriving analytical expressions for the relationship between attack rates and food-web dynamics and (ii) simulating the effects of competition on the eco-evolutionary dynamics of consumers and resources.
+I examine the effects of character displacement on food-web dynamics by (i) deriving analytical expressions for the relationship between consumer attack rates and food-web dynamics and (ii) simulating the effects of competition on the eco-evolutionary dynamics of consumers and resources.
 To gain analytical insight, I assume that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
@@
 This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$, green line in Fig.
-\ref{fig:Tradeoff}) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$, orange line Fig.
-\ref{fig:Tradeoff}).
+\ref{fig:MacArthur_Plot}A) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$, orange line Fig.
+\ref{fig:MacArthur_Plot}A).
 When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig.
-\ref{fig:Tradeoff}).
+\ref{fig:MacArthur_Plot}A).
 Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer.
@@
 
-```{r Tradeoff, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:Tradeoff}Tradeoff forms for consumer attack rates."}
+```{r Tradeoff, include=FALSE, fig.pos="H", fig.cap="\\label{fig:Tradeoff}Tradeoff forms for consumer attack rates."}
+initial_points <- bind_rows(params_2C_2R_MacArthur_linear, params_2C_2R_MacArthur_concaveDN, params_2C_2R_MacArthur_concaveUP) %>% mutate(n = c(n_linear, n_concaveDN, n_concaveUP))
+
 tradeoff_plot <- bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
@@
   geom_line() + 
+  geom_point(data = initial_points, aes(x=a11, y=a12)) +
   xlab(expression(a[ii])) +
@@
 My simulations show that the shape of the attack-rate tradeoff qualitatively affects the relationship between character displacement and resource density (Fig.
-\ref{fig:MacArthur_Resources}).
+\ref{fig:MacArthur_Plot}).
 For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (Fig.
-\ref{fig:MacArthur_Resources}A) and character displacement has no effect on resource densities (Fig.
-\ref{fig:MacArthur_Resources}B).
+\ref{fig:MacArthur_Plot}B) and character displacement has no effect on resource densities (Fig.
+\ref{fig:MacArthur_Plot}C).
 If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (Fig.
-\ref{fig:MacArthur_Resources}B).
+\ref{fig:MacArthur_Plot}C).
 This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (Fig.
-\ref{fig:MacArthur_Resources}A).
+\ref{fig:MacArthur_Plot}B).
 When the tradeoff is concave up (orange lines), character displacement suppresses resource densities due to the increase in total attack rates (Fig.
-\ref{fig:MacArthur_Resources}A,B).
+\ref{fig:MacArthur_Plot}B,C).
 Although the equation I derived for resource density was for the scenario where both consumers and both resources are present, it accurately predicts the density of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
-\ref{fig:MacArthur_Resources}B).
+\ref{fig:MacArthur_Plot}C).
 This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in Fig.
-\ref{fig:MacArthur_Resources}A), resulting in equivalent resource densities.
+\ref{fig:MacArthur_Plot}B), resulting in equivalent resource densities.
 
 The effect of character displacement on resource density closely matches its impact on food-web stability (Fig.
-\ref{fig:Stability}).
+\ref{fig:MacArthur_Plot}D).
 For example, when character displacement results in resource suppression (orange), there is a corresponding decrease in stability.
@@
 
-```{r MacArthur_Resources, include=FALSE, echo=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers can forage for both resources simultaneously."}
+```{r MacArthur_Resources, include=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers can forage for both resources simultaneously."}
 Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
@@
 
-```{r Stability, include=FALSE, echo=FALSE, fig.cap="\\label{fig:Stability}Effect of character displacement on local stability when consumers can (A) and cannot (B) forage for both resources simultaneously.
+```{r Stability, include=FALSE, fig.cap="\\label{fig:Stability}Effect of character displacement on local stability when consumers can (A) and cannot (B) forage for both resources simultaneously.
 Small and large points correspond to initial and evolved strategies, respectively."}
@@
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression(-1%*%"Re"(lambda[max]))) +
+  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
   scale_shape_manual(values=c(24,21))  +
@@
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability ("-lambda)) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
+  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
   scale_shape_manual(values=c(24,21))  +
@@
 
-```{r MacArthur Model, echo=FALSE, fig.cap="New MacArthur model figure"}
+```{r MacArthur Model, include=FALSE}
 plot_macarthur_model <- plot_grid(
@@
   Mac_resources  + theme(legend.position = "none"),
-  Stability_Mac  + theme(legend.position = "none"), ncol=2, align='hv')
+  Stability_Mac  + theme(legend.position = "none"), ncol=2, align='hv', labels = "AUTO")
 
@@
 
-```{r MacArthur_Plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}Fill in."}
+```{r MacArthur_Plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}Effect of character displacement when consumers can forage for both resources simultaneously.
+Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Panel (B) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation.
+Panels (C) and (D) show how different tradeoffs indirectly affect total resource densities (C) and food-web stability (D).
+Our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
 knitr::include_graphics("Figure_ECD_MacArthur.pdf")
@@
 
+The previous model, although the starting point for many models of competition, does not reflect many food webs where consumers are more mobile than their resources.
+In this scenario, consumer foraging behavior can link resources that occur in distinct habitats.
 The only character displacement model that I am aware of that modeled resources in different habitats was one examined by @Lawlor1976.
@@
 As with attack rates, I assume that consumer habitat preferences are mirror images of each other ($w=w_{11}=w_{22}$).
-In @Lawlor1976's detailed evolutionary analysis of this model, they again always observe divergent character displacement.
+In @Lawlor1976's detailed evolutionary analysis of this model, they always observed divergent character displacement.
 
@@
 
-To examine the ecological consequences of this model, I again derive the equilibrium solution when both consumers and both resources are present.
-Again, resource densities are equivalent at this equilibrium ($\hat R = \hat R_1 = \hat R_2$ when both consumers and resources present), but are now determined by the following equation (derivation in supplementary material):
+To examine the ecological consequences of this divergence, I again derive the equilibrium solution when both consumers and both resources are present.
+As before, resource densities are equivalent at this equilibrium ($\hat R = \hat R_1 = \hat R_2$ when both consumers and resources present), but are now determined by the following equation (derivation in supplementary material):
 
@@
 $w>0.5$ and $a_{ii}>a_{ij}$), then the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase, regardless of the tradeoff (Fig.
-\ref{fig:LS_Resources}A).
+\ref{fig:LS_Plot}A).
 Thus, character displacement always results in resource suppression if consumers are competing for resources that occur in different habitats (Fig.
-\ref{fig:LS_Resources}B).
+\ref{fig:LS_Plot}B).
 Note that the shape of the tradeoff can modify the effect of character displacement.
 This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached its ESS (triangles in Fig.
-\ref{fig:LS_Resources}B).
+\ref{fig:LS_Plot}B).
 In contrast, resource densities reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
-\ref{fig:LS_Resources}B), because character displacement tends to reach a constraint of complete specialization.
+\ref{fig:LS_Plot}B), because character displacement tends to reach a constraint of complete specialization.
 It is worth noting that in this foraging scenario, resource densities are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
-\ref{fig:LS_Resources}B).
+\ref{fig:LS_Plot}B).
 This is likely because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in Fig.
-\ref{fig:LS_Resources}B).
+\ref{fig:LS_Plot}B).
 <!-- I will revisit this subtle point when I examine the effects of character displacement on food-web stability.
@@
 
-The effect of character displacement on resource densities qualitatively matches its negative effect of food-web stability (Fig.
-\ref{fig:Stability}).
-This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (compare initial (small points) with evolved strategies (large points), Fig.
-\ref{fig:Stability}).
-For example, when the tradeoff is concave up (orange), then the four-species (small circle) food web actually starts off as less stable than the three-species one (small triangle), but the eco-evolutionary dynamics cause a switch in the pattern once they reach their evolutionary stable strategies (large points).
-
-```{r LS_Resources, echo=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource densities (B) when consumers cannot forage for resources simultaneously."}
+As we saw previously, the effect of character displacement on resource densities qualitatively matches its negative effect on food-web stability (Fig.
+\ref{fig:LS_Plot}C).
+This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (Fig.
+\ref{fig:LS_Plot}C).
+For example, when the tradeoff is concave up (orange), then the four-species (small circle) food web actually starts off as less stable than the three-species one (small triangle); however, the eco-evolutionary dynamics cause a switch in this pattern once they reach their evolutionary stable strategies (large points).
+
+```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource densities (B) when consumers cannot forage for resources simultaneously."}
 LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
@@
 
-```{r LS Model, echo=FALSE, fig.cap="LS model figure"}
+```{r LS Model, include=FALSE}
 plot_LS_model <- plot_grid(
@@
   LS_resources  + theme(legend.position = "none"),
-  Stability_LS  + theme(legend.position = "none"), ncol=2, align='hv')
+  Stability_LS  + theme(legend.position = "none"), ncol=2, align='hv', labels = "AUTO")
 
@@
 
-```{r LS Plot, echo=FALSE}
+```{r LS_Plot, echo=FALSE, fig.cap="\\label{fig:LS_Plot}Effect of character displacement when consumers *cannot* forage for both resources simultaneously.
+Different line colors correspond to the tradeoff values shown in fig.
+1A.
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
+Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
+This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
+As in fig.
+1, our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
 knitr::include_graphics("Figure_ECD_LS.pdf")
@@
 
-Thus far I have consider a simple type I functional response of consumers foraging for resources in either the same or distinct habitats.  
-For example, even if I consider a more realistic consumer functional response (derived by @McCann2005):
+So far I have only considered a simple functional response, where consumer feeding rates increase linearly with resource densities when foraging in either the same or distinct habitats.
+We know, however, that consumer feeding rates often saturate at high resource densities [@Holling1959; @Murdoch2003; @McCann2011].
+We also know that consumers do not usually spend a fixed proportion of time in a particular habitat, but their habitat preferences may be modified by resource densities [@McCann2005].
+To account for these biological realities, I consider a more realistic functional response [derived by @McCann2005]:
 
@@
 
-I still observe the same qualitative relationship between character displacement and resource suppression (Fig.
-\ref{fig:McCann_Resources} in Appendix).
-This is because resource competition always leads to character divergence and equilibrium resource densities are governed by similar dynamics  (*Mathematica* file with derivations available upon request):
+Even after adding more biological realism, I still observe the same qualitative relationship between character displacement and resource suppression (Fig.
+\ref{fig:McCann_Plot} in Appendix).
+This is because resource competition always leads to character divergence and equilibrium resource densities are governed by similar dynamics  (derivation in supplementary material):
 
@@
 
-I conducted the previous simulations on a scenario where the food web remains locally stable ($\operatorname{Re}(\lambda_{max})<0$).
+In the previous models, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$).
+In this more realistic model, however, whether the food web is locally stable depends on the parameters of both consumers and resources.
+Specifically, <!--I conducted the previous simulations on a scenario where the food web remains locally stable ($\operatorname{Re}(\lambda_{max})<0$).
 Is it possible that character displacement would ever result in a locally unstable food web?
 To gain analytical insight to this question, I examined the conditions governing the local stability of the four-species food web when consumers exhibit a more realistic functional response (@McCann2005).
-I show that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (*Mathematica* file with derivation available upon request):
+I show that -->the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derivation in supplementary material):
 
@@
 Previously, I showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff (Fig.
-\ref{fig:LS_Resources}A).
-Thus, character displacement is capable of pushing food webs to a point where they are locally unstable.
-(it simple depends on the choice of parameter, I could put a situation where evolution pushes it toward local instability, or even the hopf bifurcation.)
-
-# Discussion
-
-I find that the foraging scenario qualitatively affects the relationship between character displacement and resource densities.
-
-Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously [e.g.
-@Taper1985; @Abrams1986].
-This assumption may be valid for some foraging scenarios (e.g.
-Darwin's finches foraging for seeds); however, in many situations consumers forage for resources that occur in different habitats, and thus cannot forage for both simultaneously [@McCann2005].
-
-The effect of the foraging scenario on the relationship between character displacement and resource densities appears to be quite general.
-
-Note that detecting this effect in nature would likely be difficult, since a concave-down tradeoff results in relatively small displacement relative to other tradeoff shapes (@Lawlor1976; green points in Fig.
-\ref{fig:MacArthur_Resources}A,B).
-
-Previous work has shown that consumers undergo divergent character displacement regardless of the shape of the tradeoff (@Lawlor1976; @Abrams1986).
-
-Important point: Criteria 5 in @Schluter1992 does not make sense since character displacement usually alters resource distributions. Think about this more.
-
-<!-- This more realistic consumer functional response does actually make an interesting evolutionary prediction that differs from the scenario where consumer feeding rate scales linearly with resource density.
-Specifically, a consumer's ESS in the absence of competition will remain slightly more specialized on the resources that occur in their preferred habitats.
-This is in contrast to the prediction when feeding rates scale linearly with resource density, where consumers evolve to be slightly more specialized on resources occurring in their non-preferred habitats.
-An important caveat for these predictions is that my model does not allow habitat preferences to evolve, which would likely alter these predictions.
--->
-
-@Lawlor1976 found that consumers still underwent divergent character displacement regardless of whether consumers could or could not forage for resources simultaneously.
-
-These results do lead to an interesting prediction that could be tested either experimentally or with field data.
-If consumers are competing for resources that occur in different habitats, then greater character displacement corresponds to a greater decrease in resource densities (line trajectories in Fig.
-\ref{fig:LS_Resources}B).
-This prediction likely only applies to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
-
-In general, I find that resource supression goes hand-in-hand with local stability, a tendency that has been noted by others (@Murdoch2003; @McCann2011).
-
-<!-- Looking for this effect in nature -->
-
-The scenario where the relationship between character displacement and resource suppression is the strongest is when the tradeoff in attack rates is concave-up and consumers are competing for resources that occur in different habitats (orange points in Fig.
-\ref{fig:LS_Resources}B and Fig.
-\ref{fig:McCann_Resources}B).
-Prior work in threespine stickleback suggests that the evolution of stickleback attack rates are constrained by a concave-up tradeoff (@Schluter1993; @Arnegard2014) and that stickleback ecotypes are competing for resources that occur in different habitats (benthic vs.
-limnetic, @Schluter1992).
-Thus, I predict that ecological character displacement in sticklebacks would suppress resource densities and decrease local stability.
-
-While measuring resource densities is relatively straightforward, measuring stability is notoriously difficult.
-One of the more easy to acquire empirical metrics of stability is the coefficient of variation ($CV=\frac{\sigma}{\mu}$) as this only requires measurements of resource (or consumer) densities over time.
-Interestingly, there is a close correspondence between the coefficient of variation and the local stability of a food web (@Gellner2016).
-<!-- This close correspondence emerges from the fact that stochastic perturbations in resource (or consumer) densities will determine the variability in abundances.
-And since food webs are constantly buffeted by stochastic processes in the real world, I expect that my predictions using local stability would correspond to the $CV$ of resources (or consumers).
--->Thus, I predict that character displacement in threespine stickleback will increase the $CV$ of resource (or consumer) densities.
-
-Consumer and resource densities in these dynamical models can be interpreted from either a population or biomass perspective (@Yodzis1992; @Murdoch2003; @McCann2011).
-For the stickleback system, you have data on seasonal dynamics of resources in terms of both abundances and biomass.
-Within a season, I do not expect stickleback abundances to be coupled to the abundances of zooplankton and benthic invertebrates, given that stickleback reproduce annually whereas zooplankton and benthic invertebrates can reproduce many times within a season.
-Instead, I expect that stickleback biomass will be coupled to the biomass of zooplankton and benthic invertebrates within a season.
-This is because there will be lots of young stickleback at the beginning of the season, and they will eat a lot of resources and grow in size as the season progresses.
-If this is true, I also expect to see the largest negative effect of stickleback on resource densities late in the season.
-
-<!-- Caveats -->
-
-Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
-quantitative genetics).
-I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits (see synthesis by @Taper1985).
-Finally, my conclusions only apply to biological resources that are nutritionally substitutable.
-It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected (@Abrams1987; @Fox2008).
-
-# Acknowledgements
-
-The content of this paper was greatly enhanced by numerous conversations with Seth Rudman.
-Sally Otto provided much help in early analyses of the mathematical model and I would not have been able to take the theory as far as I did without her guidance and her pushing me.
-
-Thank Ben Gilbert for his comment during a discussion of the model that pushed me to take a comparative approach that looked across different model structures to examine the generality of these results.
-
-# References
-
-<div id="refs"></div> <!-- this code inserts references wherever you want rather than at the end of the document -->
-
-\newpage
-
-# Appendix: More realistic functional response
-
-```{r McCann_Resources, echo=FALSE, fig.cap="\\label{fig:McCann_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers exhibit a more realistic functional response."}
+\ref{fig:LS_Plot}A).
+Thus, character displacement always pushes the food web toward the boundary of local stability under this more realistic foraging scenario.
+<!--Whether or not the food web becomes locally unstable, however, depends on the choice of initial parameters.-->
+
+```{r McCann_Resources, include=FALSE, fig.cap="\\label{fig:McCann_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers exhibit a more realistic functional response."}
 McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
@@
 
-```{r Stability_McCann, echo=FALSE, fig.cap="\\label{fig:Stability_McCann}Effect of character displacement on local stability when consumers exhibit a more realistic functional response."}
+```{r Stability_McCann, include=FALSE, fig.cap="\\label{fig:Stability_McCann}Effect of character displacement on local stability when consumers exhibit a more realistic functional response."}
 Stability_McCann <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
@@
 ```
+```{r McCann Model, include=FALSE}
+plot_McCann_model <- plot_grid(
+  #tradeoff_plot + theme(legend.position = "none"), 
+  McCann_attack  + theme(legend.position = "none"), 
+  McCann_resources  + theme(legend.position = "none"),
+  Stability_McCann  + theme(legend.position = "none"), ncol=2, align='hv', labels = "AUTO")
+
+save_plot(filename="Figure_ECD_McCann.pdf", plot=plot_McCann_model, base_height = 6, base_width = 6)
+```
+
+```{r McCann_Plot, echo=FALSE, fig.cap="\\label{fig:McCann_Plot}Effect of character displacement when consumers exhibit a more realistic functional response.
+Different line colors correspond to the tradeoff values shown in fig.
+1A.
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
+Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
+This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
+As in fig.
+1, our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
+knitr::include_graphics("Figure_ECD_McCann.pdf")
+```
+
+# Discussion
+
+<!-- Summarize key findings -->
+
+Our results suggest that qualitative predictions about the effect of character displacement on food-web dynamics can be made from knowledge of evolutionary constraints/tradeoffs in attack rates as well as consumer foraging behavior.
+For all of the tradeoffs and foraging scenarios we examined, competition drives divergence in consumer foraging traits.
+Yet, these different ecological and evolutionary constraints sometimes lead to qualitatively different effects on resources densities and food-web stability. 
+Specifically, the shape of the tradeoff resulted in qualitatively different relationships between character displacement and food-web dynamics in the foraging scenario most frequently examined in past models of the evolution of character displacement.
+However, whenever consumers were mobile and competed for resources that occured in distinct habitats, character displacement suppressed resource densities and resulted in less stable food webs. 
+Taken together, our results illustrate how a food-web perspective can give novel insight to an eco-evolutionary process that appears common in nature [@Schluter2000; @Pfennig2010; but see @Stuart2013].
+
+<!-- Influence of tradeoff shape and application to natural systems -->
+
+Although the shape of the tradeoff could qualitatively change the relationships between character displacement and food-web dynamics, I do not expect this to be common in real ecosystems. 
+This is because these distinct effects only occur in a foraging scenario that I expect is uncommon in nature (linear functional response of consumers competiting for resources in the same habitat). 
+It was still important to examine this scenario, however, as it is the basis of most theoretical models that explicitly model resource dynamics in competition [@MacArthur1972; @Lawlor1976; @Abrams1986].
+Even if we suspend our disbelief (for the moment) about the reality of the foraging situation, I expect that we would only be able to detect two effects of character displacement on food-web dynamics: no change in resource densities or stability; or decrease in resource densities and stability.
+I expect it would be unlikely to ever observe an increase in resource densities under character displacement, although its theoretical plausible (fig. \ref{fig:MacArthur_Plot}A). 
+This is because the concave down tradeoff does not favor strong divergence, and thus there is little potential for an increase in resource densities.
+It is unclear to me which tradeoff shape is most common in nature.
+Of the two examples I found, there appears to be a concave up tradeoff in consumer foraging traits for two classic examples of character displacement: threespine stickleback in coastal lakes of British Columbia [@Arnegard2014] and Darwin's finches on the Galapagos islands [@Schluter1985].
+A concave up tradeoff favors greater divergence, and thus is also the most likely evolutionary scenario to result in detectable character displacment.
+The shape of the tradeoff becomes evident from examining the form of natural selection on consumer traits in the absence of interspecific competition and when resources were limiting. 
+Specifically, in both systems there is evidence for disruptive selection (favoring extreme phenotypes) in environments with a broad spectrum of resources present, but where resources are ultimately limiting.
+Note that while this gives insight to the shape of the tradeoff, it does not give insight to the evolutionary dynamics of the system. 
+This is because, in the absence of interspecific competition, resource densities are the key determinant of consumer evolution (see supplementary material). 
+This is because even if specialization on one resource is favored in the short-term, the other resource will increase in density, favoring consumer to evolve traits that better exploit the other resource.
+Thus, there is an eco-evolutionary feedback that eventually favors an intermediate phenotype that is able to best take advantage of both resources (fig. \ref{fig:MacArthur_Plot}C).
+Based on the shape of their tradeoffs alone, I would predict character displacement in both of these systems to decrease total resource densities.
+I expect this prediction to hold particularly well for threespine stickleback, where the food-web dynamics more closely mimic this model (competition for two distinct resources, rather than competition along a continuous trait distribution, seeds for Darwin's finches).
+Whether this translates into a detectable effect on food-web stability is difficult, but if does occur, I would expect it to be more likely in threespine stickleback. 
+This is because there tends to be less of a stoichiometric mismatch between consumers and their resources in aquatic vs. terrestrial ecosystems. 
+This means that energy transfer is more efficient to upper trophic levels, which creates more top-heavy food webs (cite McCann's work) that are less stable.
+
+<!-- Influence of foraging scenario and application to natural systems 
+- Divergent character displacement may be quite common in systems with animal consumers (and thus biological resources).
+This means there are many situations where consumers couple resources in distinct habitats.
+Draw on sticklebacks as an example.
+
+- Overrepresentation of carnivores (and even including herbivores) in this Schluter 2000.
+
+- Important point: Criteria 5 in @Schluter1992 does not make sense since character displacement usually alters resource distributions. Think about this more.
+-->
+
+<!-- Food-web stability
+- Our most striking result is that character displacement, under foraging scenarios that capture the biology of many consumers, could result in unstable food webs.
+This contrasts the notion from coexistence theory that character displacement contributes to species persistence.
+My results do not contradict these past inferences, as they were based on different measures of stability.
+However, its clear that stability measures do not always go hand-in-hand and that a pluralistic view of different aspects of stability is needed for understanding the eco-evolutionary effect on food-web dynamics.
+- In line with past theory, we found that food-web stability tended to go hand-in-hand with resource densities. 
+- Drawing on the theory I developed, I can point to situations where we would expect character displacement to results in less stable food webs. 
+    - Consumers foraging in distinct habitats, concave up tradeoff in attack rates, and in aquatic systems.
+Aquatic systems are more likely because its been shown that attack rates are often much stronger in these systems.
+This lower baseline, means that it is easier for character displacement to push these systems toward the boundary of local stability.
+    - Interestingly, the study system that remains the hallmark of character displacement, threespine stickleback, meets all of these criteria.
+Thus, I predict that in the coastal lakes of British Columbia, where stickleback have undergone character displacement, for these food webs to exhibit less stable dynamics and more variable population abundances.
+    - Although stability is a traditionally thorny issue to measure.
+--
+
+<!-- Caveats -->
+
+Our model assumes symmetry in displacement which is not uncommon in empirical data [@Schluter2000].
+
+<!-- older cut and pasted -->
+
+I find that the foraging scenario qualitatively affects the relationship between character displacement and resource densities.
+
+Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously [e.g.
+@Taper1985; @Abrams1986].
+This assumption may be valid for some foraging scenarios (e.g.
+Darwin's finches foraging for seeds); however, in many situations consumers forage for resources that occur in different habitats, and thus cannot forage for both simultaneously [@McCann2005].
+
+The effect of the foraging scenario on the relationship between character displacement and resource densities appears to be quite general.
+
+Note that detecting this effect in nature would likely be difficult, since a concave-down tradeoff results in relatively small displacement relative to other tradeoff shapes (@Lawlor1976; green points in Fig.
+\ref{fig:MacArthur_Resources}A,B).
+
+Previous work has shown that consumers undergo divergent character displacement regardless of the shape of the tradeoff (@Lawlor1976; @Abrams1986).
+
+Important point: Criteria 5 in @Schluter1992 does not make sense since character displacement usually alters resource distributions. Think about this more.
+
+<!-- This more realistic consumer functional response does actually make an interesting evolutionary prediction that differs from the scenario where consumer feeding rate scales linearly with resource density.
+Specifically, a consumer's ESS in the absence of competition will remain slightly more specialized on the resources that occur in their preferred habitats.
+This is in contrast to the prediction when feeding rates scale linearly with resource density, where consumers evolve to be slightly more specialized on resources occurring in their non-preferred habitats.
+An important caveat for these predictions is that my model does not allow habitat preferences to evolve, which would likely alter these predictions.
+-->
+
+@Lawlor1976 found that consumers still underwent divergent character displacement regardless of whether consumers could or could not forage for resources simultaneously.
+
+These results do lead to an interesting prediction that could be tested either experimentally or with field data.
+If consumers are competing for resources that occur in different habitats, then greater character displacement corresponds to a greater decrease in resource densities (line trajectories in Fig.
+\ref{fig:LS_Resources}B).
+This prediction likely only applies to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
+
+In general, I find that resource supression goes hand-in-hand with local stability, a tendency that has been noted by others (@Murdoch2003; @McCann2011).
+
+<!-- Looking for this effect in nature -->
+
+The scenario where the relationship between character displacement and resource suppression is the strongest is when the tradeoff in attack rates is concave-up and consumers are competing for resources that occur in different habitats (orange points in Fig.
+\ref{fig:LS_Resources}B and Fig.
+\ref{fig:McCann_Resources}B).
+Prior work in threespine stickleback suggests that the evolution of stickleback attack rates are constrained by a concave-up tradeoff (@Schluter1993; @Arnegard2014) and that stickleback ecotypes are competing for resources that occur in different habitats (benthic vs.
+limnetic, @Schluter1992).
+Thus, I predict that ecological character displacement in sticklebacks would suppress resource densities and decrease local stability.
+
+While measuring resource densities is relatively straightforward, measuring stability is notoriously difficult.
+One of the more easy to acquire empirical metrics of stability is the coefficient of variation ($CV=\frac{\sigma}{\mu}$) as this only requires measurements of resource (or consumer) densities over time.
+Interestingly, there is a close correspondence between the coefficient of variation and the local stability of a food web (@Gellner2016).
+<!-- This close correspondence emerges from the fact that stochastic perturbations in resource (or consumer) densities will determine the variability in abundances.
+And since food webs are constantly buffeted by stochastic processes in the real world, I expect that my predictions using local stability would correspond to the $CV$ of resources (or consumers).
+-->Thus, I predict that character displacement in threespine stickleback will increase the $CV$ of resource (or consumer) densities.
+
+Consumer and resource densities in these dynamical models can be interpreted from either a population or biomass perspective (@Yodzis1992; @Murdoch2003; @McCann2011).
+For the stickleback system, you have data on seasonal dynamics of resources in terms of both abundances and biomass.
+Within a season, I do not expect stickleback abundances to be coupled to the abundances of zooplankton and benthic invertebrates, given that stickleback reproduce annually whereas zooplankton and benthic invertebrates can reproduce many times within a season.
+Instead, I expect that stickleback biomass will be coupled to the biomass of zooplankton and benthic invertebrates within a season.
+This is because there will be lots of young stickleback at the beginning of the season, and they will eat a lot of resources and grow in size as the season progresses.
+If this is true, I also expect to see the largest negative effect of stickleback on resource densities late in the season.
+
+<!-- Caveats -->
+
+Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
+quantitative genetics).
+I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits (see synthesis by @Taper1985).
+Finally, my conclusions only apply to biological resources that are nutritionally substitutable.
+It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected (@Abrams1987; @Fox2008).
+
+# Acknowledgements
+
+The content of this paper was greatly enhanced by numerous conversations with Seth Rudman.
+Sally Otto provided much help in early analyses of the mathematical model and I would not have been able to take the theory as far as I did without her guidance and encouragement.
+A discussion with Ben Gilbert pushed me to take a comparative approach that looked across different foraging scenarios to examine the generality of my conclusions.

### 2019-02-06 manuscript_ECD_model.Rmd
first draft of Discussion finished.

@@
 Specifically, the shape of the tradeoff resulted in qualitatively different relationships between character displacement and food-web dynamics in the foraging scenario most frequently examined in past models of the evolution of character displacement.
-However, whenever consumers were mobile and competed for resources that occured in distinct habitats, character displacement suppressed resource densities and resulted in less stable food webs. 
+However, in a foraging scenario that is likely more common in nature, where mobile consumers compete for resources that occured in distinct habitats, character displacement suppressed resource densities and resulted in less stable food webs.
 Taken together, our results illustrate how a food-web perspective can give novel insight to an eco-evolutionary process that appears common in nature [@Schluter2000; @Pfennig2010; but see @Stuart2013].
@@
 
-<!-- Influence of foraging scenario and application to natural systems 
-- Divergent character displacement may be quite common in systems with animal consumers (and thus biological resources).
-This means there are many situations where consumers couple resources in distinct habitats.
-Draw on sticklebacks as an example.
-
-- Overrepresentation of carnivores (and even including herbivores) in this Schluter 2000.
-
-- Important point: Criteria 5 in @Schluter1992 does not make sense since character displacement usually alters resource distributions. Think about this more.
--->
-
-<!-- Food-web stability
-- Our most striking result is that character displacement, under foraging scenarios that capture the biology of many consumers, could result in unstable food webs.
-This contrasts the notion from coexistence theory that character displacement contributes to species persistence.
-My results do not contradict these past inferences, as they were based on different measures of stability.
-However, its clear that stability measures do not always go hand-in-hand and that a pluralistic view of different aspects of stability is needed for understanding the eco-evolutionary effect on food-web dynamics.
-- In line with past theory, we found that food-web stability tended to go hand-in-hand with resource densities. 
-- Drawing on the theory I developed, I can point to situations where we would expect character displacement to results in less stable food webs. 
-    - Consumers foraging in distinct habitats, concave up tradeoff in attack rates, and in aquatic systems.
-Aquatic systems are more likely because its been shown that attack rates are often much stronger in these systems.
+<!-- Influence of foraging scenario and application to natural systems -->
+A general result that emerged from our analysis was that when consumers were competing for resources in distinct habitats, character displacement always surpressed resource densities, regardless of the functional form of the consumer's feeding rate or shape of the tradeoff.
+This is a very common foraging scenario in nature, as consumers tend to be more mobile than their resources, often coupling their dynamics across traditional habitat boundaries [@McCann2005].
+This may be especially so for character displacement, where a disproportionate number of the documented cases come from carnivores [@Schluter2000]. 
+Perhaps the clearest example of this would be the threespine stickleback in coastal lakes of British Columbia. 
+Here, the diet of stickleback is composed of two main resource types, zooplankton and benthic invertebrates. 
+These resource types occur in distinct zones of the lake (pelagic and littoral, respectively), and sticklebacks move between these zones when foraging for resources. 
+I predict that in species-pair lakes, where stickleback have undergone character displacement, that total resource densities will be lower compared to lakes with a single, generalist species.
+
+These results do lead to an interesting prediction that could be tested either experimentally or with field data. 
+If consumers are competing for resources that occur in different habitats, then greater character displacement corresponds to a greater decrease in resource densities (line trajectories in fig.
+\ref{fig:LS_Plot}B, \ref{fig:McCann_Plot}B).
+This prediction likely only applies, however, to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
+
+<!-- Consider moving this to the opening of the discussion or the opening of the conclusion section -->
+Our work also points to a potential revision for one of the criteria used to demonstrate character displacement. 
+Specifically, criteria 5 in @Schluter1992 states: "Sites of sympatry and allopatry should not differ greatly in food, ...". 
+Our results demonstrate that character displacement can alter resource distributions in a predictable way, usually resulting in lower resource availability where competition has resulted in character displacement.
+
+<!-- Food-web stability -->
+Our most stiking result was that character displacement, when examined through the lens of food-web theory, usually resulted in less stable food webs. 
+In fact, under a foraging scenario that incorporates more biological realism, character displacement may even result in an unstable food web. 
+This result aligns well with the general result from food-web theory that food-web stability tends to go hand-in-hand with resource densities [@May1973; @Murdoch2003; @McCann2011]. 
+However, this result contrasts the notion from competition theory that character displacement contributes to species coexistence. 
+My results do not contradict these past inferences, as they were based on a different measure of stability that considers the ability of a consumer with a different phenotype to invade when rare (invasion criteria).
+However, its clear that stability measures do not always go hand-in-hand and that a pluralistic view of different aspects of stability is needed [@Donohue2016] for understanding eco-evolutionary feedbacks on food-web dynamics.
+
+<!-- Ghost of competition and stability past --->
+A single species of threespine stickleback lives in hundreds of lakes of British Columbia, but the species pairs, where character displacement has happened, are only known from four lakes one of which has undergone reverse speciation [@Rudman2016].
+Our results hint at an interesting hypothesis for the apparent rarity, namely founded on ecological stability [@Borrelli2015]. 
+Specifically, the commonness of the lakes with only a single species is due to the higher ecological stability of this food-web structure, compared to the one with a species pair. 
+Testing this hypothesis may be quite difficult, but it could be possible with recent advances in genomics. 
+For example, the species-pair lake that has undergone reverse speciation has a different ecological impact than other generalist lakes in that it is much more like the benthic-ecotype in its morphology.
+It is unclear whether this is a transient or persistent evolutionary state.
+But advances in genomics may be able to tease apart whether a currently generalist species had a genomic imprint of being more benthic or limnetic.
+
+<!-- Predictions for a natural system -->
+Drawing on the theory I developed, I can point to systems where we would expect character displacement to result in less stable food webs. 
+My theory suggests that this may be more likely to occur when character displacement occurs between: (i) consumers foraging in distinct habitats; (ii) exhibit concave-up tradeoff in attack rates; and (iii) in aquatic systems.
+Aquatic systems are more likely because its been shown that attack rates are often much stronger in these systems [@Rip2011]. 
 This lower baseline, means that it is easier for character displacement to push these systems toward the boundary of local stability. 
-    - Interestingly, the study system that remains the hallmark of character displacement, threespine stickleback, meets all of these criteria.
+Interestingly, one of the hallmark examples of character displacement, threespine stickleback, meets these criteria.
 Thus, I predict that in the coastal lakes of British Columbia, where stickleback have undergone character displacement, for these food webs to exhibit less stable dynamics and more variable population abundances.
-    - Although stability is a traditionally thorny issue to measure.
---
+Although stability is a traditionally thorny issue to measure, there are now clear theoretical links between the stability measure we used here ($-\lambda_{\text{max}}$) and the coefficient of variation (CV), which simply measures the variability in a dynamic time series relative to the mean.
+Also, there are now methods for assessing the dynamics of natural communities that link to these theoretical stability measures [@Ives2003]. Now it just requires the effort to collect this data. 
+Also, in applying this theory to a natural system, an assessment of stability for a specific system requires a realistic model of the population and community dynamics.
+e.g.
+stage-structure,...Still, our results point toward a general tendency toward instability.
+
+It's entirely plausible though that the some aspects of the system, e.g.
+stage structure could prevent this instability.
+This could be quite interesting, as this may point to systems where we are more likely to observe character displacement, because it enhances their stability.
 
 <!-- Caveats -->
-
-Our model assumes symmetry in displacement which is not uncommon in empirical data [@Schluter2000].
-
-<!-- older cut and pasted -->
+Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
+quantitative genetics).
+I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits (see synthesis by @Taper1985).
+Since these models are focused primarily on trait evolution, it is not clear that divergence in these models would have different ecological effects.
+In fact, recent work by @McPeek2019 has shown that genetic divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
+Finally, my conclusions may only apply to biological resources that are nutritionally substitutable. 
+It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected (@Abrams1987; @Fox2008).
+
+Throughout, I have also setup a special case scenario where consumers are mirror images of each other.
+This resulted in symmetric character displacement, which is quite common in empirical data [@Schluter2000].
+Again though, I do not expect that adding mild asymmetries (ones that still would permit coexistence) would qualitatively change our results.
+This is because, again, divergent character displacement is still evident in these models.
+Also, we conducted simulations to test whether our results were robust to this assumption and they were (ADD SUPPLEMENTARY MATERIAL).
+
+<!-- older cut and pasted
 
@@
 An important caveat for these predictions is that my model does not allow habitat preferences to evolve, which would likely alter these predictions.
--->
 
@@
 
-These results do lead to an interesting prediction that could be tested either experimentally or with field data.
-If consumers are competing for resources that occur in different habitats, then greater character displacement corresponds to a greater decrease in resource densities (line trajectories in Fig.
-\ref{fig:LS_Resources}B).
-This prediction likely only applies to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
-
 In general, I find that resource supression goes hand-in-hand with local stability, a tendency that has been noted by others (@Murdoch2003; @McCann2011).
 
-<!-- Looking for this effect in nature -->
+<!-- Looking for this effect in nature
 
@@
 And since food webs are constantly buffeted by stochastic processes in the real world, I expect that my predictions using local stability would correspond to the $CV$ of resources (or consumers).
--->Thus, I predict that character displacement in threespine stickleback will increase the $CV$ of resource (or consumer) densities.
+Thus, I predict that character displacement in threespine stickleback will increase the $CV$ of resource (or consumer) densities.
 
@@
 If this is true, I also expect to see the largest negative effect of stickleback on resource densities late in the season.
-
-<!-- Caveats -->
-
-Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
-quantitative genetics).
-I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits (see synthesis by @Taper1985).
-Finally, my conclusions only apply to biological resources that are nutritionally substitutable.
-It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected (@Abrams1987; @Fox2008).
+-->
 
@@
 A discussion with Ben Gilbert pushed me to take a comparative approach that looked across different foraging scenarios to examine the generality of my conclusions.
+Discussions with Dolph Schluter pushed me to also to take a comparative approach that better set my results in the context of past theory on character displacement.

### 2019-02-06 manuscript_ECD_model.Rmd
edited end of 2nd paragraph in intro to better tie in rationale for a food-web perspective.

@@
 For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition.
+<!--
 On the other hand, specialization often increases a consumer’s ability to capture a specific resource.
-A well known consequence of increased foraging efficiency is resource suppression, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics [Rosenzweig 1971; Murdoch et al.
+From food-web theory, we know that increased feeding rates often lead to resource suppression, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics [Rosenzweig 1971; Murdoch et al.
 2003; McCann 2011].
-NEED A LAST SENTENCE TO WRAP IT UP, HIGHLIGHTING HOW A FOOD-WEB PERSPECTIVE MAY YIELD NEW ECOLOGICAL INSIGHTS AND GIVE A CONTRASTING PERSPECTIVE TO COEXISTENCE THEORY.
+-->
+Yet, this reduction in interspecific competition may, at the same time, increase the interspecific interaction between a consumer and the resource it specializes on.  
+Both food-web theory and empirical studies have shown that increasing the strength of consumer-resource interactions often leads to resource suppression, which if sufficient enough, can generate oscillations, and hence less stable consumer-resource dynamics [@Rosenzweig1971; @Luckinbill1973; @Murdoch2002; @Murdoch2003; @McCann2011].
+Thus, a food-web perspective, which accounts for both the direct and indirect effects of consumer-resource interactions, may yield new insight to the ecological consequences of character displacement.
+
+<!--Thus, a consumer-resource perspective may yield new insight to the ecological feedback of character displacement on food-web dynamics. 
+<!--Given that character displacement is mediated by a consumer-resource interaction, a food-web perspective may yield   NEED A LAST SENTENCE TO WRAP IT UP, HIGHLIGHTING HOW A FOOD-WEB PERSPECTIVE MAY YIELD NEW ECOLOGICAL INSIGHTS AND GIVE A CONTRASTING PERSPECTIVE TO COEXISTENCE THEORY.
 PERHAPS A I NEED TO INCLUDE A SENTENCE MID-WAY THROUGH SUGGESTING THAT A FOOD-WEB PERSPECTIVE MAY GIVE QUALITATIVELY DIFFERENT INSIGHTS.
 IT MAY BE NICE TO BRING IN THE SPATIAL COUPLING WORK OF MCCANN SUGGESTING THAT SPECIALIZATION CAN COUPLE FOOD-WEBS, BUT THIS WAS ONLY WITH ONE TOP PREDATOR (I.E.
-NO CONSUMERS...LOOK AT OLD TEXT FOR WORDING SUGGESTIONS.)
+NO CONSUMERS...LOOK AT OLD TEXT FOR WORDING SUGGESTIONS.)-->
 

### 2019-02-07 manuscript_ECD_model.Rmd
fixed ref formatting. adding in initial asymmetric sims

@@
     fig_caption: true
-    #citation_package: natbib # why isn't this working?
-    template: Rmd_AmNat_MS_template.tex
+    citation_package: natbib 
     includes:  
       in_header: Am-Nat-preamble-latex.tex
-bibliography: references.bib
-biblio-style: amnat.bst
+bibliography: references
+biblio-style: amnat
 fontsize: 11pt
@@
                                 extra_tradeoff_params = c(A=A, n=n_linear))
+# asymmetric starting points
+params_2C_2R_MacArthur_linear_asymm <- params_2C_2R_MacArthur_linear
+params_2C_2R_MacArthur_linear_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_linear[nrow(sim_1C_2R_MacArthur_linear), c("a11","a12")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
+This will create an asymmetry in the attack rates and follows the logic (and our results) that a likely scenerio with asymmetry is a specialist invading a generalist.
+# just altering the parameters, everything else stays the same
+sim_2C_2R_MacArthur_linear_asymm <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_linear_asymm,
+                                init_states = states_2C_2R, 
+                                eco_CR_model = MacArthur_2C_2R, 
+                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_linear))
 
@@
                                 extra_tradeoff_params = c(A=A, n=n_concaveDN))
+# asymmetric starting points
+params_2C_2R_MacArthur_concaveDN_asymm <- params_2C_2R_MacArthur_concaveDN
+params_2C_2R_MacArthur_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveDN[nrow(sim_1C_2R_MacArthur_concaveDN), c("a11","a12")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
+This will create an asymmetry in the attack rates and follows the logic (and our results) that a likely scenerio with asymmetry is a specialist invading a generalist.
+# just altering the parameters, everything else stays the same
+sim_2C_2R_MacArthur_concaveDN_asymm <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_concaveDN_asymm,
+                                init_states = states_2C_2R, 
+                                eco_CR_model = MacArthur_2C_2R, 
+                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
+
 ## concaveUP
@@
                                        tradeoff_functions = c(a12_tradeoff),
+                                       extra_tradeoff_params = c(A=A, n=n_concaveUP))
+# asymmetric starting points
+params_2C_2R_MacArthur_concaveUP_asymm <- params_2C_2R_MacArthur_concaveUP
+params_2C_2R_MacArthur_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveUP[nrow(sim_1C_2R_MacArthur_concaveUP), c("a11","a12")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
+This will create an asymmetry in the attack rates and follows the logic (and our results) that a likely scenerio with asymmetry is a specialist invading a generalist.
+# just altering the parameters, everything else stays the same
+
+# max iterations reached!
+sim_2C_2R_MacArthur_concaveUP_asymm <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_concaveUP_asymm,
+                                init_states = states_2C_2R, 
+                                eco_CR_model = MacArthur_2C_2R, 
+                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
                                 extra_tradeoff_params = c(A=A, n=n_concaveUP))
@@
 
+## MacArthur asymmetry
+#MacArthur_displacements_asymm <- mutate(sim_2C_2R_MacArthur_linear_asymm, Model = "MacArthur", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_linear_asymm)), ]
+#MacArthur_displacements_asymm$Time <- rep(c("Begin","End"),1)
+
+MacArthur_displacements_asymm <- bind_rows(
+  mutate(sim_2C_2R_MacArthur_linear_asymm, Model = "MacArthur", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_linear_asymm)), ], 
+  #mutate(sim_1C_2R_MacArthur_linear_asymm, Model = "MacArthur", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_linear_asymm)), ],
+  mutate(sim_2C_2R_MacArthur_concaveDN_asymm, Model = "MacArthur", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveDN_asymm)), ], 
+  #mutate(sim_1C_2R_MacArthur_concaveDN_asymm, Model = "MacArthur", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveDN_asymm)), ],
+  mutate(sim_2C_2R_MacArthur_concaveUP_asymm, Model = "MacArthur", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveUP_asymm)), ])#, 
+  #mutate(sim_1C_2R_MacArthur_concaveUP_asymm, Model = "MacArthur", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveUP_asymm)), ]) 
+MacArthur_displacements_asymm$Time <- rep(c("Begin","End"),3)
+
 ## Lawlor Smith
@@
 
+```{r MacArthur_asymm}
+# replaced MacArthur_displacements with MacArthur_displacements_asymm
+
+Mac_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  ylab(expression(a[ii]+a[ij])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Mac_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  ylab(expression(R[i]+R[j])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_Mac_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="black", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  geom_hline(yintercept=0, linetype="dotted", color="grey") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Mac_plots_asymm <- plot_grid(Mac_attack_asymm + theme(legend.position = "none"), 
+                             Mac_resources_asymm + theme(legend.position = "none"), 
+                             Stability_Mac_asymm + theme(legend.position = "none"), 
+                       ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
+Mac_plots_asymm
+```
+
 ```{r Stability, include=FALSE, fig.cap="\\label{fig:Stability}Effect of character displacement on local stability when consumers can (A) and cannot (B) forage for both resources simultaneously.
@@
   scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  geom_hline(yintercept=0, linetype="dotted", color="grey") +
   guides(fill = guide_legend(override.aes=list(shape=21)))
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="black", show.legend=FALSE) +
   geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) +
   #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) +
   #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) +
   #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=5, color="black") + #, show.legend=FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) +
   #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=5, color="black", show.legend = FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="black", show.legend = FALSE) +
   geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
@@
 ```
+
 ```{r McCann Model, include=FALSE}

### 2019-02-08 manuscript_ECD_model.Rmd
implemented asymmetric models. I now identified follow-up scenarios that I have to explore (different pre-adaptations of single consumer)

@@
                                        extra_tradeoff_params = c(A=A, n=n_linear))
+# asymmetric starting points
+params_2C_2R_LawlorSmith_linear_asymm <- params_2C_2R_LawlorSmith_linear
+params_2C_2R_LawlorSmith_linear_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_linear[nrow(sim_1C_2R_LawlorSmith_linear), c("a11","a12")] 
+sim_2C_2R_LawlorSmith_linear_asymm <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_linear_asymm,
+                                init_states = states_2C_2R, 
+                                eco_CR_model = LawlorSmith_2C_2R, 
+                                AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_linear))
 
@@
                                        extra_tradeoff_params = c(A=A, n=n_concaveDN))
+# asymmetric starting points
+params_2C_2R_LawlorSmith_concaveDN_asymm <- params_2C_2R_LawlorSmith_concaveDN
+params_2C_2R_LawlorSmith_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_concaveDN[nrow(sim_1C_2R_LawlorSmith_concaveDN), c("a11","a12")] 
+sim_2C_2R_LawlorSmith_concaveDN_asymm <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveDN_asymm,
+                                init_states = states_2C_2R, 
+                                eco_CR_model = LawlorSmith_2C_2R, 
+                                AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
+
 ## concaveUP
@@
                                         tradeoff_functions = c(a12_tradeoff),
+                                        extra_tradeoff_params = c(A=A, n=n_concaveUP))
+# asymmetric starting points
+params_2C_2R_LawlorSmith_concaveUP_asymm <- params_2C_2R_LawlorSmith_concaveUP
+params_2C_2R_LawlorSmith_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_concaveUP[nrow(sim_1C_2R_LawlorSmith_concaveUP), c("a11","a12")] 
+sim_2C_2R_LawlorSmith_concaveUP_asymm <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveUP_asymm,
+                                init_states = states_2C_2R, 
+                                eco_CR_model = LawlorSmith_2C_2R, 
+                                AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
                                 extra_tradeoff_params = c(A=A, n=n_concaveUP))
@@
                                          extra_tradeoff_params = c(A=A, n=n_linear))
+# asymmetric starting points
+params_2C_2R_McCann_linear_asymm <- params_2C_2R_McCann_linear
+params_2C_2R_McCann_linear_asymm[c("a21","a22")] <- sim_1C_2R_McCann_linear[nrow(sim_1C_2R_McCann_linear), c("a11","a12")] 
+sim_2C_2R_McCann_linear_asymm <- eco_evo_CR(init_parameters = params_2C_2R_McCann_linear_asymm,
+                                init_states = states_2C_2R, 
+                                eco_CR_model = McCann_2C_2R, 
+                                AD_CR_models = c(mC1_McCann, mC2_McCann),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_linear))
+
+# asymmetry with C1 starting with a different habitat preference (w)
+params_1C_2R_McCann_linear_alt_w <- params_1C_2R_McCann_linear # NEED TO FIGURE OUT WHAT TO DO.
+i SHOULD PROBABLY MAKE SURE ATTACK RATES AND PREFERENCE ARE ALIGNED, BUT I SHOULD JUST DO THAT FOR THE ALTERNATIVE RESOURCE.
+E.G.
+C1 SPECIALIZING ON R2 INSTEAD OF R1 IN THE C1 ONLY SIMS.
+sim_1C_2R_McCann_linear <- eco_evo_CR(init_parameters = params_1C_2R_McCann_linear_alt_w, 
+                                         init_states = states_1C_2R, 
+                                         eco_CR_model = McCann_1C_2R, 
+                                         AD_CR_models = c(mC1_McCann),
+                                         species_traits = c(1), 
+                                         eco_param_names = c("a11"), 
+                                         eco_tradeoff_names = c("a12"),  
+                                         evo_param_names = c("a11m"),  
+                                         evo_tradeoff_names = c("a12m"), 
+                                         tradeoff_functions = c(a12_tradeoff),
+                                         extra_tradeoff_params = c(A=A, n=n_linear))
 
@@
                                          extra_tradeoff_params = c(A=A, n=n_concaveDN))
+# asymmetric starting points
+params_2C_2R_McCann_concaveDN_asymm <- params_2C_2R_McCann_concaveDN
+params_2C_2R_McCann_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_McCann_concaveDN[nrow(sim_1C_2R_McCann_concaveDN), c("a11","a12")] 
+sim_2C_2R_McCann_concaveDN_asymm <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveDN_asymm,
+                                init_states = states_2C_2R, 
+                                eco_CR_model = McCann_2C_2R, 
+                                AD_CR_models = c(mC1_McCann, mC2_McCann),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
+
 ## concaveUP
@@
                                           tradeoff_functions = c(a12_tradeoff),
+                                          extra_tradeoff_params = c(A=A, n=n_concaveUP))
+# asymmetric starting points
+params_2C_2R_McCann_concaveUP_asymm <- params_2C_2R_McCann_concaveUP
+params_2C_2R_McCann_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_McCann_concaveUP[nrow(sim_1C_2R_McCann_concaveUP), c("a11","a12")] 
+sim_2C_2R_McCann_concaveUP_asymm <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveUP_asymm,
+                                init_states = states_2C_2R, 
+                                eco_CR_model = McCann_2C_2R, 
+                                AD_CR_models = c(mC1_McCann, mC2_McCann),
+                                species_traits = c(1,2), 
+                                eco_param_names = c("a11","a22"), 
+                                eco_tradeoff_names = c("a12","a21"),  
+                                evo_param_names = c("a11m","a22m"),  
+                                evo_tradeoff_names = c("a12m","a21m"), 
+                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
                                 extra_tradeoff_params = c(A=A, n=n_concaveUP))
@@
 
-## MacArthur asymmetry
-#MacArthur_displacements_asymm <- mutate(sim_2C_2R_MacArthur_linear_asymm, Model = "MacArthur", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_linear_asymm)), ]
-#MacArthur_displacements_asymm$Time <- rep(c("Begin","End"),1)
-
+## MacArthur asymmetry. Note that all 1C_2R models are from the original simulations, as these were the starting points for the asymmetric sims (and therefore the appropriate comparison)
 MacArthur_displacements_asymm <- bind_rows(
   mutate(sim_2C_2R_MacArthur_linear_asymm, Model = "MacArthur", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_linear_asymm)), ], 
-  #mutate(sim_1C_2R_MacArthur_linear_asymm, Model = "MacArthur", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_linear_asymm)), ],
+  mutate(sim_1C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_linear)), ],
   mutate(sim_2C_2R_MacArthur_concaveDN_asymm, Model = "MacArthur", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveDN_asymm)), ], 
-  #mutate(sim_1C_2R_MacArthur_concaveDN_asymm, Model = "MacArthur", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveDN_asymm)), ],
-  mutate(sim_2C_2R_MacArthur_concaveUP_asymm, Model = "MacArthur", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveUP_asymm)), ])#, 
-  #mutate(sim_1C_2R_MacArthur_concaveUP_asymm, Model = "MacArthur", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveUP_asymm)), ]) 
-MacArthur_displacements_asymm$Time <- rep(c("Begin","End"),3)
+  mutate(sim_1C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveDN)), ],
+  mutate(sim_2C_2R_MacArthur_concaveUP_asymm, Model = "MacArthur", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveUP_asymm)), ], 
+  mutate(sim_1C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveUP)), ]) 
+MacArthur_displacements_asymm$Time <- rep(c("Begin","End"),6)
 
@@
 
+# asymmetry
+LawlorSmith_displacements_asymm <- bind_rows(
+  mutate(sim_2C_2R_LawlorSmith_linear_asymm, Model = "LawlorSmith", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_linear_asymm)), ], 
+  mutate(sim_1C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_linear)), ],
+  mutate(sim_2C_2R_LawlorSmith_concaveDN_asymm, Model = "LawlorSmith", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_concaveDN_asymm)), ], 
+  mutate(sim_1C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_concaveDN)), ],
+  mutate(sim_2C_2R_LawlorSmith_concaveUP_asymm, Model = "LawlorSmith", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_concaveUP_asymm)), ], 
+  mutate(sim_1C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_concaveUP)), ]) 
+LawlorSmith_displacements_asymm$Time <- rep(c("Begin","End"),6)
+
 ## McCann
@@
 
+# asymmetry
+McCann_displacements_asymm <- bind_rows(
+  mutate(sim_2C_2R_McCann_linear_asymm, Model = "McCann", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_linear_asymm)), ], 
+  mutate(sim_1C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_linear)), ],
+  mutate(sim_2C_2R_McCann_concaveDN_asymm, Model = "McCann", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_concaveDN_asymm)), ], 
+  mutate(sim_1C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_concaveDN)), ],
+  mutate(sim_2C_2R_McCann_concaveUP_asymm, Model = "McCann", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_concaveUP_asymm)), ], 
+  mutate(sim_1C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_concaveUP)), ]) 
+McCann_displacements_asymm$Time <- rep(c("Begin","End"),6)
+
+# write out data for exploring later and avoiding rerunning the simulations
+write.csv(bind_rows(MacArthur_displacements, LawlorSmith_displacements, McCann_displacements), file = "symmetric_displacement_data.csv")
+write.csv(bind_rows(MacArthur_displacements_asymm, LawlorSmith_displacements_asymm, McCann_displacements_asymm), file = "asymmetric_displacement_data.csv")
 ```
@@
 2003).
-Experimental decreases in interaction strength between consumers and resources has been shown to promote coexistence between consumers and resources (Luckinbill 1973).
+Experimental decreases in interaction strength between consumers and resources has been shown to promote coexistence between consumers and resources (Luckinbill 1973).
 Moreover, there is broad support that specialized consumers, since they are tightly coupled to their resources, tend to exhibit consumer-resource cycles, whereas this is not the case for generalists (Murdoch et al.
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) + size=5,
-  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(a[ii]+a[ij])) +
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) + size=5,
-  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
 Mac_plots
-```
-
-```{r MacArthur_asymm}
-# replaced MacArthur_displacements with MacArthur_displacements_asymm
-
-Mac_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) + size=5,
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
-  ylab(expression(a[ii]+a[ij])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Mac_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) + size=5,
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
-  ylab(expression(R[i]+R[j])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Stability_Mac_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="black", show.legend=FALSE) + #size=5,
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
-  scale_shape_manual(values=c(21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  geom_hline(yintercept=0, linetype="dotted", color="grey") +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Mac_plots_asymm <- plot_grid(Mac_attack_asymm + theme(legend.position = "none"), 
-                             Mac_resources_asymm + theme(legend.position = "none"), 
-                             Stability_Mac_asymm + theme(legend.position = "none"), 
-                       ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
-Mac_plots_asymm
 ```
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="black", show.legend=FALSE) + #size=5,
-  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="black", show.legend=FALSE) +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) +
-  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(w*a[ii]+(1-w)*a[ij])) +
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) +
-  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) +
-  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(w*a[ii]+(1-w)*a[ij])) +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="black") + #, show.legend=FALSE) +
-  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="black", show.legend = FALSE) +
-  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="black") +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend = FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
 Discussions with Dolph Schluter pushed me to also to take a comparative approach that better set my results in the context of past theory on character displacement.
+
+# Supplementary Material
+
+```{r MacArthur_asymm, include=FALSE}
+# replaced MacArthur_displacements with MacArthur_displacements_asymm
+
+Mac_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(a[ii]+a[ij])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Mac_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(R[i]+R[j])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_Mac_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  geom_hline(yintercept=0, linetype="dotted", color="grey") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Mac_plots_asymm <- plot_grid(Mac_attack_asymm + theme(legend.position = "none"), 
+                             Mac_resources_asymm + theme(legend.position = "none"), 
+                             Stability_Mac_asymm + theme(legend.position = "none"), 
+                       ncol=2, labels = "AUTO", align = 'hv')
+Mac_plots_asymm
+
+save_plot(filename="Figure_ECD_MacArthur_asymm.pdf", plot=Mac_plots_asymm, base_height = 6, base_width = 6)
+```
+
+```{r MacArthur_Plot_asymm, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}Effect of character displacement when consumers can forage for both resources simultaneously.
+Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Panel (B) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation.
+Panels (C) and (D) show how different tradeoffs indirectly affect total resource densities (C) and food-web stability (D).
+Our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
+knitr::include_graphics("Figure_ECD_MacArthur_asymm.pdf")
+```
+
+```{r LawlorSmith_asymm, include=FALSE}
+# replaced LawlorSmith_displacements with LawlorSmith_displacements_asymm
+
+LS_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(a[ii]+a[ij])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+LS_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(R[i]+R[j])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_LS_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  geom_hline(yintercept=0, linetype="dotted", color="grey") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+LS_plots_asymm <- plot_grid(LS_attack_asymm + theme(legend.position = "none"), 
+                             LS_resources_asymm + theme(legend.position = "none"), 
+                             Stability_LS_asymm + theme(legend.position = "none"), 
+                       ncol=2, labels = "AUTO", align='hv')
+LS_plots_asymm
+
+save_plot(filename="Figure_ECD_LS_asymm.pdf", plot=LS_plots_asymm, base_height = 6, base_width = 6)
+```
+
+```{r LS_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:LS_Plot_asymm}Effect of character displacement when consumers *cannot* forage for both resources simultaneously.
+Different line colors correspond to the tradeoff values shown in fig.
+1A.
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
+Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
+This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
+As in fig.
+1, our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
+knitr::include_graphics("Figure_ECD_LS_asymm.pdf")
+```
+
+```{r McCann_asymm, include=FALSE}
+# replaced McCann_displacements with McCann_displacements_asymm
+
+McCann_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(a[ii]+a[ij])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+McCann_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(R[i]+R[j])) +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_McCann_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  geom_hline(yintercept=0, linetype="dotted", color="grey") +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+McCann_plots_asymm <- plot_grid(McCann_attack_asymm + theme(legend.position = "none"), 
+                             McCann_resources_asymm + theme(legend.position = "none"), 
+                             Stability_McCann_asymm + theme(legend.position = "none"), 
+                       ncol=2, labels = "AUTO", align = 'hv')
+McCann_plots_asymm
+
+save_plot(filename="Figure_ECD_McCann_asymm.pdf", plot=McCann_plots_asymm, base_height = 6, base_width = 6)
+```
+
+```{r McCann_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:McCann_Plot_asymm}Effect of character displacement when consumers exhibit a more realistic functional response.
+Different line colors correspond to the tradeoff values shown in fig.
+1A.
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
+Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
+This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
+As in fig.
+1, our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
+knitr::include_graphics("Figure_ECD_McCann_asymm.pdf")
+```

### 2019-02-09 manuscript_ECD_model.Rmd
fixed an important error in previous asymmetric sims. Now a_ii and a_ij are matched up properly.

@@
 params_2C_2R_MacArthur_linear_asymm <- params_2C_2R_MacArthur_linear
-params_2C_2R_MacArthur_linear_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_linear[nrow(sim_1C_2R_MacArthur_linear), c("a11","a12")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
+params_2C_2R_MacArthur_linear_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_linear[nrow(sim_1C_2R_MacArthur_linear), c("a12","a11")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
 This will create an asymmetry in the attack rates and follows the logic (and our results) that a likely scenerio with asymmetry is a specialist invading a generalist.
@@
 params_2C_2R_MacArthur_concaveDN_asymm <- params_2C_2R_MacArthur_concaveDN
-params_2C_2R_MacArthur_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveDN[nrow(sim_1C_2R_MacArthur_concaveDN), c("a11","a12")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
+params_2C_2R_MacArthur_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveDN[nrow(sim_1C_2R_MacArthur_concaveDN), c("a12","a11")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
 This will create an asymmetry in the attack rates and follows the logic (and our results) that a likely scenerio with asymmetry is a specialist invading a generalist.
@@
 params_2C_2R_MacArthur_concaveUP_asymm <- params_2C_2R_MacArthur_concaveUP
-params_2C_2R_MacArthur_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveUP[nrow(sim_1C_2R_MacArthur_concaveUP), c("a11","a12")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
+params_2C_2R_MacArthur_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveUP[nrow(sim_1C_2R_MacArthur_concaveUP), c("a12","a11")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
 This will create an asymmetry in the attack rates and follows the logic (and our results) that a likely scenerio with asymmetry is a specialist invading a generalist.
@@
 params_2C_2R_LawlorSmith_linear_asymm <- params_2C_2R_LawlorSmith_linear
-params_2C_2R_LawlorSmith_linear_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_linear[nrow(sim_1C_2R_LawlorSmith_linear), c("a11","a12")] 
+params_2C_2R_LawlorSmith_linear_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_linear[nrow(sim_1C_2R_LawlorSmith_linear), c("a12","a11")] 
 sim_2C_2R_LawlorSmith_linear_asymm <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_linear_asymm,
@@
 params_2C_2R_LawlorSmith_concaveDN_asymm <- params_2C_2R_LawlorSmith_concaveDN
-params_2C_2R_LawlorSmith_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_concaveDN[nrow(sim_1C_2R_LawlorSmith_concaveDN), c("a11","a12")] 
+params_2C_2R_LawlorSmith_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_concaveDN[nrow(sim_1C_2R_LawlorSmith_concaveDN), c("a12","a11")] 
 sim_2C_2R_LawlorSmith_concaveDN_asymm <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveDN_asymm,
@@
 params_2C_2R_LawlorSmith_concaveUP_asymm <- params_2C_2R_LawlorSmith_concaveUP
-params_2C_2R_LawlorSmith_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_concaveUP[nrow(sim_1C_2R_LawlorSmith_concaveUP), c("a11","a12")] 
+params_2C_2R_LawlorSmith_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_concaveUP[nrow(sim_1C_2R_LawlorSmith_concaveUP), c("a12","a11")] 
 sim_2C_2R_LawlorSmith_concaveUP_asymm <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveUP_asymm,
@@
                                          extra_tradeoff_params = c(A=A, n=n_linear))
+# I'm thinking these initial asymmetric simulations don't quite make sense (at least for LawlorSmith and McCann).
+This is because I didn't switch the pre-adapted habitat preferences as well...This is important because C1 in the first simulation becomes C2 in the 2C-2R sim.
+This isn't an issue in the symmetry model, because I assume they are pre-adapted on different resources.
 # asymmetric starting points
 params_2C_2R_McCann_linear_asymm <- params_2C_2R_McCann_linear
-params_2C_2R_McCann_linear_asymm[c("a21","a22")] <- sim_1C_2R_McCann_linear[nrow(sim_1C_2R_McCann_linear), c("a11","a12")] 
+params_2C_2R_McCann_linear_asymm[c("a21","a22")] <- sim_1C_2R_McCann_linear[nrow(sim_1C_2R_McCann_linear), c("a12","a11")] # make sure a_ij and a_ii are in correct order.
+Note that habitat preference (w) is already set in the appropriate direction
 sim_2C_2R_McCann_linear_asymm <- eco_evo_CR(init_parameters = params_2C_2R_McCann_linear_asymm,
@@
 # asymmetry with C1 starting with a different habitat preference (w)
-params_1C_2R_McCann_linear_alt_w <- params_1C_2R_McCann_linear # NEED TO FIGURE OUT WHAT TO DO.
+#params_1C_2R_McCann_linear_switch <- params_1C_2R_McCann_linear # NEED TO FIGURE OUT WHAT TO DO.
 i SHOULD PROBABLY MAKE SURE ATTACK RATES AND PREFERENCE ARE ALIGNED, BUT I SHOULD JUST DO THAT FOR THE ALTERNATIVE RESOURCE.
@@
 C1 SPECIALIZING ON R2 INSTEAD OF R1 IN THE C1 ONLY SIMS.
-sim_1C_2R_McCann_linear <- eco_evo_CR(init_parameters = params_1C_2R_McCann_linear_alt_w, 
-                                         init_states = states_1C_2R, 
-                                         eco_CR_model = McCann_1C_2R, 
-                                         AD_CR_models = c(mC1_McCann),
-                                         species_traits = c(1), 
-                                         eco_param_names = c("a11"), 
-                                         eco_tradeoff_names = c("a12"),  
-                                         evo_param_names = c("a11m"),  
-                                         evo_tradeoff_names = c("a12m"), 
-                                         tradeoff_functions = c(a12_tradeoff),
-                                         extra_tradeoff_params = c(A=A, n=n_linear))
+# switching direction of pre-adapted attack rates and habitat preference
+#params_1C_2R_McCann_linear_switch["a11"] <- params_1C_2R_McCann_linear["a12"] 
+#params_1C_2R_McCann_linear_switch["a12"] <- params_1C_2R_McCann_linear["a11"] 
+#params_1C_2R_McCann_linear_switch["w11"] <- params_1C_2R_McCann_linear["w12"] 
+#params_1C_2R_McCann_linear_switch["w12"] <- params_1C_2R_McCann_linear["w11"]
+#sim_1C_2R_McCann_linear_switch <- eco_evo_CR(init_parameters = params_1C_2R_McCann_linear_switch, 
+#                                             init_states = states_1C_2R, 
+#                                             eco_CR_model = McCann_1C_2R, 
+#                                             AD_CR_models = c(mC1_McCann),
+#                                             species_traits = c(1), 
+#                                             eco_param_names = c("a11"), 
+#                                             eco_tradeoff_names = c("a12"),  
+#                                             evo_param_names = c("a11m"),  
+#                                             evo_tradeoff_names = c("a12m"), 
+#                                             tradeoff_functions = c(a12_tradeoff),
+#                                             extra_tradeoff_params = c(A=A, n=n_linear))
+#params_2C_2R_McCann_linear_asymm_switch <- params_2C_2R_McCann_linear
+#params_2C_2R_McCann_linear_asymm_switch[c("a21","a22")] <- sim_1C_2R_McCann_linear_switch[nrow(sim_1C_2R_McCann_linear_switch), c("a11","a12")] 
+#sim_2C_2R_McCann_linear_asymm_switch <- eco_evo_CR(init_parameters = params_2C_2R_McCann_linear_asymm_switch,
+#                                                    init_states = states_2C_2R, 
+#                                                    eco_CR_model = McCann_2C_2R, 
+#                                                    AD_CR_models = c(mC1_McCann, mC2_McCann),
+#                                                    species_traits = c(1,2), 
+#                                                    eco_param_names = c("a11","a22"), 
+#                                                    eco_tradeoff_names = c("a12","a21"),  
+#                                                    evo_param_names = c("a11m","a22m"),  
+#                                                    evo_tradeoff_names = c("a12m","a21m"), 
+#                                                    tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+#                                                    extra_tradeoff_params = c(A=A, n=n_linear))
 
@@
 params_2C_2R_McCann_concaveDN_asymm <- params_2C_2R_McCann_concaveDN
-params_2C_2R_McCann_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_McCann_concaveDN[nrow(sim_1C_2R_McCann_concaveDN), c("a11","a12")] 
+params_2C_2R_McCann_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_McCann_concaveDN[nrow(sim_1C_2R_McCann_concaveDN), c("a12","a11")] 
 sim_2C_2R_McCann_concaveDN_asymm <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveDN_asymm,
@@
                                 extra_tradeoff_params = c(A=A, n=n_concaveDN))
+# asymmetric starting points | switched
 
@@
 params_2C_2R_McCann_concaveUP_asymm <- params_2C_2R_McCann_concaveUP
-params_2C_2R_McCann_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_McCann_concaveUP[nrow(sim_1C_2R_McCann_concaveUP), c("a11","a12")] 
+params_2C_2R_McCann_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_McCann_concaveUP[nrow(sim_1C_2R_McCann_concaveUP), c("a12","a11")] 
 sim_2C_2R_McCann_concaveUP_asymm <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveUP_asymm,

### 2019-02-09 manuscript_ECD_model.Rmd
edits to figure legends.

@@
 ```{r LS_Plot, echo=FALSE, fig.cap="\\label{fig:LS_Plot}Effect of character displacement when consumers *cannot* forage for both resources simultaneously.
-Different line colors correspond to the tradeoff values shown in fig.
-1A.
+Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
@@
 ```{r McCann_Plot, echo=FALSE, fig.cap="\\label{fig:McCann_Plot}Effect of character displacement when consumers exhibit a more realistic functional response.
-Different line colors correspond to the tradeoff values shown in fig.
-1A.
+Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
@@
 
-```{r MacArthur_Plot_asymm, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}Effect of character displacement when consumers can forage for both resources simultaneously.
-Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Panel (B) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation.
-Panels (C) and (D) show how different tradeoffs indirectly affect total resource densities (C) and food-web stability (D).
+```{r MacArthur_Plot_asymm, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}Effect of character displacement when consumers can forage for both resources simultaneously and have an asymmetry in initial attack rates.
+Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
+Panel (A) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
+Panels (B) and (C) show how different tradeoffs indirectly affect total resource densities (B) and food-web stability (C).
 Our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
@@
 
-```{r LS_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:LS_Plot_asymm}Effect of character displacement when consumers *cannot* forage for both resources simultaneously.
-Different line colors correspond to the tradeoff values shown in fig.
-1A.
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
+```{r LS_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:LS_Plot_asymm}Effect of character displacement when consumers *cannot* forage for both resources simultaneously and have an asymmetry in initial attack rates.
+Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
 Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
@@
 
-```{r McCann_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:McCann_Plot_asymm}Effect of character displacement when consumers exhibit a more realistic functional response.
-Different line colors correspond to the tradeoff values shown in fig.
-1A.
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
+```{r McCann_Plot_asymm, echo=FALSE, fig.cap="**\\label{fig:McCann_Plot_asymm} Consumer asymmetry and Realistic Functional Response.** Effect of character displacement when consumers exhibit a more realistic functional response and have an asymmetry in initial attack rates.
+Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
 Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.

### 2019-02-09 manuscript_ECD_model.Rmd
add Borrelli cites
edits to in text citations

@@
 @Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has been generated to understand the scenarios under which exploitative competition for resources leads to the divergence of consumer foraging traits.
-The general result emerging from these studies is that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the divergence of consumer foraging traits [@Lawlor1976; @Taper1985).
+The general result emerging from these studies is that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the divergence of consumer foraging traits [@Lawlor1976; @Taper1985].
 This process is not simply driven by ecological differences, but creates an eco-evolutionary feedback that drives further differentiation.
@@
 Although explicit models of resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics have received surprisingly little attention. 
-This is likely because the ecological feedback has been primarily studied through the lens of competition and coexistence theory [Lawlor1976; Germain2018; Bassar; McPeek]. 
+This is likely because the ecological feedback has been primarily studied through the lens of competition and coexistence theory [@Lawlor1976; @Germain2018; @Bassar; @McPeek2019]. 
 For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition.
@@
 NO NEED TO GIVE ALL THE DETAILS YET.
-RIGHT NOW IT IS JUST CONFUSING.* (i) resources occur in the same habitat and consumers exhibit a linear functional response (e.g.
-@MacArthur1972; @Lawlor1976); (ii) resources occur in different habitats and consumers exhibit a linear functional response (e.g.
-@Lawlor1976); (iii) resources occur in distinct habitats and consumers exhibit a more realistic functional response [@McCann2005].
+RIGHT NOW IT IS JUST CONFUSING.* (i) resources occur in the same habitat and consumers exhibit a linear functional response [e.g.
+@MacArthur1972; @Lawlor1976]; (ii) resources occur in different habitats and consumers exhibit a linear functional response [e.g.
+@Lawlor1976]; (iii) resources occur in distinct habitats and consumers exhibit a more realistic functional response [@McCann2005].
 I also explore the contingency of these effects to different assumptions about evolutionary trade-offs in consumer foraging traits.
@@
 My results do not contradict these past inferences, as they were based on a different measure of stability that considers the ability of a consumer with a different phenotype to invade when rare (invasion criteria).
-However, its clear that stability measures do not always go hand-in-hand and that a pluralistic view of different aspects of stability is needed [@Donohue2016] for understanding eco-evolutionary feedbacks on food-web dynamics.
+However, its clear that stability measures do not always go hand-in-hand and that a pluralistic view of different aspects of stability is needed [@Donohue2013; @Donohue2016] for understanding eco-evolutionary feedbacks on food-web dynamics.
 
@@
 A single species of threespine stickleback lives in hundreds of lakes of British Columbia, but the species pairs, where character displacement has happened, are only known from four lakes one of which has undergone reverse speciation [@Rudman2016].
-Our results hint at an interesting hypothesis for the apparent rarity, namely founded on ecological stability [@Borrelli2015]. 
+Our results hint at an interesting hypothesis for the apparent rarity, namely founded on ecological stability [@Borrelli2015a; @Borrelli2015b]. 
 Specifically, the commonness of the lakes with only a single species is due to the higher ecological stability of this food-web structure, compared to the one with a species pair. 
@@
 quantitative genetics).
-I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits (see synthesis by @Taper1985).
+I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits [see synthesis by @Taper1985].
 Since these models are focused primarily on trait evolution, it is not clear that divergence in these models would have different ecological effects.
@@
 Finally, my conclusions may only apply to biological resources that are nutritionally substitutable. 
-It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected (@Abrams1987; @Fox2008).
+It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
 

### 2019-02-09 manuscript_ECD_model.Rmd
updates to in-text citations

@@
 Although explicit models of resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics have received surprisingly little attention. 
-This is likely because the ecological feedback has been primarily studied through the lens of competition and coexistence theory [@Lawlor1976; @Germain2018; @Bassar; @McPeek2019]. 
+This is likely because the ecological feedback has been primarily studied through the lens of competition and coexistence theory [@Lawlor1976; @Germain2018; @Bassar2017; @McPeek2019]. 
 For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition.
@@
 I also explore the contingency of these effects to different assumptions about evolutionary trade-offs in consumer foraging traits.
-I focus on the ecological side of these dynamics because the evolutionary aspects have been well characterized [@Lawlor1976; @Taper1985; @Abrams1986; McPeekRECENT].
+I focus on the ecological side of these dynamics because the evolutionary aspects have been well characterized [@Lawlor1976; @Taper1985; @Abrams1986; @McPeek2019].
 I show that the consequences of character displacement on food-web dynamics are not always intuitive and depend on the foraging scenario and tradeoff in consumer traits.

### 2019-02-13 manuscript_ECD_model.Rmd
updates to asymm sims; figure legends; discussion of McPeek 2019; and adding abstract.

@@
 ---
-title: Predicting the effects of character displacement on food-web dynamics
-author: 
-- Matthew A. Barbour$^{1,2,\ast}$
+#title: |
+#   | Predicting the effects of character displacement on food-web dynamics
+#   | Food-web dynamics under ecological character displacement
+#author: 
+#- Matthew A. Barbour$^{1,2,\ast}$
 # potentially use for 'date' for affiliations, as recommended by https://github.com/rstudio/bookdown/issues/348 for cross-format compatibility, with | to create new sections.
@@
 
+\vspace*{0.1cm} 
+\begin{center} \LARGE Food-web dynamics under ecological character displacement \end{center}
+
+\bigskip
+
+<!-- Remove author info for double-blind peer review; Keep for bioRxiv submission -->
+\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}
+
+\bigskip
+
 \noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;
@@
 A <- 2
-aii <- A*0.6
+aii <- A*0.61
 
@@
 params_2C_2R_MacArthur_concaveDN_asymm <- params_2C_2R_MacArthur_concaveDN
-params_2C_2R_MacArthur_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveDN[nrow(sim_1C_2R_MacArthur_concaveDN), c("a12","a11")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
-This will create an asymmetry in the attack rates and follows the logic (and our results) that a likely scenerio with asymmetry is a specialist invading a generalist.
-# just altering the parameters, everything else stays the same
+params_2C_2R_MacArthur_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveDN[nrow(sim_1C_2R_MacArthur_concaveDN), c("a12","a11")] 
 sim_2C_2R_MacArthur_concaveDN_asymm <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_concaveDN_asymm,
@@
 params_2C_2R_MacArthur_concaveUP_asymm <- params_2C_2R_MacArthur_concaveUP
-params_2C_2R_MacArthur_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveUP[nrow(sim_1C_2R_MacArthur_concaveUP), c("a12","a11")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
-This will create an asymmetry in the attack rates and follows the logic (and our results) that a likely scenerio with asymmetry is a specialist invading a generalist.
-# just altering the parameters, everything else stays the same
-
+params_2C_2R_MacArthur_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveUP[nrow(sim_1C_2R_MacArthur_concaveUP), c("a12","a11")]
 # max iterations reached!
@@
 
-# asymmetry with C1 starting with a different habitat preference (w)
-#params_1C_2R_McCann_linear_switch <- params_1C_2R_McCann_linear # NEED TO FIGURE OUT WHAT TO DO.
-i SHOULD PROBABLY MAKE SURE ATTACK RATES AND PREFERENCE ARE ALIGNED, BUT I SHOULD JUST DO THAT FOR THE ALTERNATIVE RESOURCE.
-E.G.
-C1 SPECIALIZING ON R2 INSTEAD OF R1 IN THE C1 ONLY SIMS.
-# switching direction of pre-adapted attack rates and habitat preference
-#params_1C_2R_McCann_linear_switch["a11"] <- params_1C_2R_McCann_linear["a12"] 
-#params_1C_2R_McCann_linear_switch["a12"] <- params_1C_2R_McCann_linear["a11"] 
-#params_1C_2R_McCann_linear_switch["w11"] <- params_1C_2R_McCann_linear["w12"] 
-#params_1C_2R_McCann_linear_switch["w12"] <- params_1C_2R_McCann_linear["w11"]
-#sim_1C_2R_McCann_linear_switch <- eco_evo_CR(init_parameters = params_1C_2R_McCann_linear_switch, 
-#                                             init_states = states_1C_2R, 
-#                                             eco_CR_model = McCann_1C_2R, 
-#                                             AD_CR_models = c(mC1_McCann),
-#                                             species_traits = c(1), 
-#                                             eco_param_names = c("a11"), 
-#                                             eco_tradeoff_names = c("a12"),  
-#                                             evo_param_names = c("a11m"),  
-#                                             evo_tradeoff_names = c("a12m"), 
-#                                             tradeoff_functions = c(a12_tradeoff),
-#                                             extra_tradeoff_params = c(A=A, n=n_linear))
-#params_2C_2R_McCann_linear_asymm_switch <- params_2C_2R_McCann_linear
-#params_2C_2R_McCann_linear_asymm_switch[c("a21","a22")] <- sim_1C_2R_McCann_linear_switch[nrow(sim_1C_2R_McCann_linear_switch), c("a11","a12")] 
-#sim_2C_2R_McCann_linear_asymm_switch <- eco_evo_CR(init_parameters = params_2C_2R_McCann_linear_asymm_switch,
-#                                                    init_states = states_2C_2R, 
-#                                                    eco_CR_model = McCann_2C_2R, 
-#                                                    AD_CR_models = c(mC1_McCann, mC2_McCann),
-#                                                    species_traits = c(1,2), 
-#                                                    eco_param_names = c("a11","a22"), 
-#                                                    eco_tradeoff_names = c("a12","a21"),  
-#                                                    evo_param_names = c("a11m","a22m"),  
-#                                                    evo_tradeoff_names = c("a12m","a21m"), 
-#                                                    tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-#                                                    extra_tradeoff_params = c(A=A, n=n_linear))
-
 ## concaveDN
@@
                                 extra_tradeoff_params = c(A=A, n=n_concaveDN))
-# asymmetric starting points | switched
 
@@
 
-<!--
-Ecological character displacement is a key driver of diversification and adaptive radiation; however, there has been little work examining the consequences of character displacement on the dynamics of ecological communities.
-Given the growing recognition for the role of evolution in structuring ecological communities, understanding how major evolutionary processes impact ecological dynamics is increasingly important.
-Here, we build a mathematical model to predict how exploitative competition between two consumers for two resources affects the evolution of consumer attack rates, and, in turn, consumer-resource population dynamics.
-Our model suggested that increasing ecological character displacement decreases the equilibrium density of resources, and with sufficient divergence, results in more oscillatory (destabilizing) consumer-resource dynamics.
-These results suggest that ecological character displacement can eventually have a negative feedback on food-web persistence in the absence of other processes.
--->
+Ecological character displacement is an adaptive process, often enhancing coexistence between consumers by reducing competition for shared resources. 
+This focus on competition has undermined the fact that consumer-resource interactions ultimately mediate the process of character displacement. 
+As a consequence, the effects of character displacement on consumer-resource interactions and food-web dynamics have not been critically examined.
+Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumers influences resource densities and the resilience of food webs to perturbations.
+To test the generality of these effects, I examined them under different assumptions of evolutionary tradeoffs, consumer foraging behavior, and resource distributions.
+My general finding was that, whenever character displacement increased the strength of consumer-resource interactions, it resulted in lower resource densities and less resilient food webs. 
+Character displacement always strengthened consumer-resource interactions whenever consumers were competing for resources that occured in distinct habitats. 
+This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
+While character displacement did not always increase the strength of consumer-resource interactions, these contingent effects only occurred under the simplest and least realistic foraging scenario. 
+<!--In the simple ecological scenario that has been the focus of past models of character displacement (type 1 functional response and competing for resources in the same habitat), I find that the shape of the tradeoff can qualitatively affect the relationship between character displacement, resource densities, and food-web stability.
+Extensions of this model, ones that include other foraging scenarios with more realistic functional responses (e.g.
+saturating) and that reflect the biological reality that consumers are often more mobile than their resources, I find predictable effects of character displacement on resource densities and food-web stability.
+Competition theory has generally shown that character displacement enhances consumer coexistence.
+I show that this adaptive process comes at the ecological cost of making food webs less resilient to perturbations.  -->
+Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations. 
+They also suggest that food-web and competition theory can give complimentary insights to the ecological consequences of adaptation. <!-- eco-evolutionary feedbacks -->
 
@@
 
-```{r MacArthur_Plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}Effect of character displacement when consumers can forage for both resources simultaneously.
-Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r MacArthur_Plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}**Effect of character displacement when consumers can forage for both resources simultaneously.** Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Panel (B) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
@@
 Panels (C) and (D) show how different tradeoffs indirectly affect total resource densities (C) and food-web stability (D).
-Our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
 We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
@@
 
-```{r LS_Plot, echo=FALSE, fig.cap="\\label{fig:LS_Plot}Effect of character displacement when consumers *cannot* forage for both resources simultaneously.
-Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r LS_Plot, echo=FALSE, fig.cap="\\label{fig:LS_Plot}**Effect of character displacement when consumers *cannot* forage for both resources simultaneously.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
@@
 This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
-As in fig.
-1, our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6; *w~ij~*=0.4.
 We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
@@
 So far I have only considered a simple functional response, where consumer feeding rates increase linearly with resource densities when foraging in either the same or distinct habitats.
-We know, however, that consumer feeding rates often saturate at high resource densities [@Holling1959; @Murdoch2003; @McCann2011].
+We know, however, that consumer feeding rates often saturate at high resource densities [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011].
 We also know that consumers do not usually spend a fixed proportion of time in a particular habitat, but their habitat preferences may be modified by resource densities [@McCann2005].
@@
 
-```{r McCann_Plot, echo=FALSE, fig.cap="\\label{fig:McCann_Plot}Effect of character displacement when consumers exhibit a more realistic functional response.
-Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r McCann_Plot, echo=FALSE, fig.cap="\\label{fig:McCann_Plot}**Effect of character displacement when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
@@
 This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
-As in fig.
-1, our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.
 We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
@@
 ```
+
+## Robustness to consumer asymmetry
+
+The previous analytical results and simulations make a strong assumption that competing consumers start off as perfect mirror images of each other in their attack rates (i.e.
+symmetry).
+Yet, the theory presented here and in past work indicates that there would likely be a predictable asymmetry between consumers.
+This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, reaches an ESS (as a generalist), followed by the invasion of a second, more specialized, consumer.
+This theoretical scenario has also been hypothesized as the sequence of events leading to character displacement in threespine stickleback inhabiting lakes of coastal British Columbia, Canada [@Schluter1992; @Schluter2000].
+
+To test whether our results were robust to this likely asymmetry, we used the evolved attack rates at the end of the one consumer simulations as the starting values for the competitor in the two consumer simulations.
+We did this for all foraging scenarios and tradeoffs previously examined.
+<!--As expected, consumer asymmetry resulted in different ecological conditions (\ref{fig:MacArthur_Plot_asymm}B,C, \ref{fig:LS_Plot_asymm}B,C, \ref{fig:McCann_Plot_asymm}B,C).
+This is evident by seeing the deviations in the small circles (initial two consumers) from the lines that map the resource and stability dynamics of a symmetric two consumer system.
+Despite these initial differences, the system-->
+We found that our previous inferences were robust to including consumer asymmetry across different tradeoffs and foraging scenarios (\ref{fig:MacArthur_Plot_asymm}, \ref{fig:LS_Plot_asymm}, \ref{fig:McCann_Plot_asymm}).
 
@@
 
+It may be quite common for character displacement to result in less stable food webs.
+Recent theory has shown that consumer divergence and specialization on resources happens more readily when either: higher productivity, low structural complexity of habitat, and weak influences of stressors [@McPeek2019].
+In our model, this corresponds to higher values of *K*, higher values of *A*, and lower values of *m*.
+In other words, this recent theory shows that the magnitude of consumer divergence happens in highly productive habitats, where consumers can easily find and consume their prey, and there is low consumer mortality.
+The first situation can lead to the "paradox of enrichment" [@Rosenzweig1971], where higher productivity, all else equal, can decrease food-web stability due to consumers suppressing resources more below their carrying capacity.
+The other factors, maximum attack rate and consumer mortality, also have predictable effects on the stability of the system [@Rip2011; @McCann2011].
+For example, increases in attack rate, as we observed in our model, decreases food-web stability.
+Similarly, lower mortality increases enables consumers to have a higher impact on their resources.
+All of these result from the general tendency of for any biological mechanism that increases the strength of interspecific interactions relative to the strength of intraspecific interactions acts to destabilize consumer-resource dynamics [@Chesson2008; @Rip2011; @McCann2011].
+Mathematically, this happens when their is an imaginary eigenvalue of the system, which puts some "spin" on the system.
+Otherwise, biological mechanisms that enhance interaction strength can be initially stabilizing, if the consumer is poorly adapted to foraging on the available resources (see first parts of stability figures).
+
 <!-- Caveats -->
+An advantage of the Adaptive Dynamics approach is that we could examine the ecological dynamics of the system.
+Current consumer-resource models based on quantitative genetics become quickly intractable to gain analytical insight from, even for simple linear functional responses of consumers [@McPeek2019].
+Those approaches have clear advantages in that they explicitly model a quantitative trait, which then influences the ecological rates (*r, K, A, e, m*, etc.).
+Our approach assumes that there is a one-to-one correspondence between trait evolution and the ecological rate.
+This simplifying assumption, however, allowed us to explore a much more realistic biological scenario and still gain analytical insight.
+This necessary tradeoff was worth it, we feel, because we were interested in the ecological consequences of character displacement, rather than in identifying the conditions resulting in character displacement and that alter the magnitude, as excellent work has already been done on this front [e.g.
+@Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019].
+
 Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
@@
 
+While the underlying genetics of trait evolution may influence the amount of divergence in character displacement, this should not influence the effect of that trait evolution on the consumer-resource dynamics.
+Another way our results may quantitatively differ from a quantitative genetic model is that there is no intraspecific variation in our approach.
+While this may certainly have quantitative effects on the relationship between character displacement and food-web dynamics, or more likely, influence the magnitude of displacement [@McPeek2019], it is unclear whether this would qualitatively influence our inferences.
+
 Throughout, I have also setup a special case scenario where consumers are mirror images of each other.
@@
 
-```{r MacArthur_Plot_asymm, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}Effect of character displacement when consumers can forage for both resources simultaneously and have an asymmetry in initial attack rates.
-Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r MacArthur_Plot_asymm, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}**Effect of character displacement when consumers can forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
@@
 Panels (B) and (C) show how different tradeoffs indirectly affect total resource densities (B) and food-web stability (C).
-Our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
 We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation.
 We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
@@
 
-LS_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(a[ii]+a[ij])) +
+LS_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
 
-```{r LS_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:LS_Plot_asymm}Effect of character displacement when consumers *cannot* forage for both resources simultaneously and have an asymmetry in initial attack rates.
-Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r LS_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:LS_Plot_asymm}**Effect of character displacement when consumers *cannot* forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
@@
 This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
-As in fig.
-1, our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.
 We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation.
 We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
@@
 
-McCann_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
-  geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(a[ii]+a[ij])) +
+McCann_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
 
-```{r McCann_Plot_asymm, echo=FALSE, fig.cap="**\\label{fig:McCann_Plot_asymm} Consumer asymmetry and Realistic Functional Response.** Effect of character displacement when consumers exhibit a more realistic functional response and have an asymmetry in initial attack rates.
-Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r McCann_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:McCann_Plot_asymm}**Effect of character displacement when consumers exhibit a more realistic functional response and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
-Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
+Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of consumers.
 This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
-As in fig.
-1, our simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
+These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.
 We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation.
 We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}

### 2019-02-13 manuscript_ECD_model.Rmd
edited summary paragraph and simplest foraging scenario paragraphs of the Discussion.

@@
 
+<!--
 Our results suggest that qualitative predictions about the effect of character displacement on food-web dynamics can be made from knowledge of evolutionary constraints/tradeoffs in attack rates as well as consumer foraging behavior.
@@
 Taken together, our results illustrate how a food-web perspective can give novel insight to an eco-evolutionary process that appears common in nature [@Schluter2000; @Pfennig2010; but see @Stuart2013].
+-->
+
+My general finding was that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource densities and less resilient food webs. 
+Character displacement always increased attack rates whenever consumers were competing for resources that occured in distinct habitats. 
+This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
+Although character displacement did not always increase effective attack rates, these contingent effects only occurred under the simplest and least realistic foraging scenario. 
+My study takes advantage of past work showing that adaptive divergence in consumer attack rates is always favored under the scenarios examined here [@LawlorSmith1976; @Abrams1986].
+Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations.
 
 <!-- Influence of tradeoff shape and application to natural systems -->
+
+The simplest foraging scenario examined here (linear functional response on resources occuring in the same habitat) is the foundation of mechanistic models of resource competition [@MacArthur1972; @Lawlor1976; @Abrams1986]<!--substitutable resources-->.
+<!--Under this scenario, character displacement either results in no change in food-web dynamics or a decrease in both resource densities and stability.-->
+Under this scenario, it is theoretically possible that character displacement results in increased resource densities and food-web stability [green lines in \ref{fig:MacArthur_Plot}].
+This only occurs, however, under an evolutionary tradeoff that constrains the magnitude of character displacement.
+This suggests that it would be unlikely to ever observe divergent character displacement and increases in resource densities in a real system. 
+In contrast, the evolutionary tradeoff that favors character displacement the most decreased resource densities and food-web stability. 
+Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the foraging rate of consumers (orange lines in \ref{fig:MacArthur_Plot}A,B).
+This is evident by the disruptive selection observed for beak size/shape in Darwin's finches [@Schluter1985] and benthic/limnetic ecotypes for stickleback [@Arnegard2014] in resource-limited environments.
+Taken together, my model predicts that even in the simplest foraging scenario, the most likely relationship we would observe in nature is for food webs where consumers have undergone character displacement to have lower total resource densites and be less resilient to perturbations.
+
+<!--
+In this scenario, the shape of the evolutionary tradeoff in attack rates qualitatively affected the relationship between character displacement and food-web dynamics. 
+In theory, resource densities and food-web stability can even be higher under character displacement if intermediate combinations of attack rates are higher than the extremes [green lines in \ref{fig:MacArthur_Plot}].
+But the magnitude of character displacement in this scenario is quite small, resulting in negligible increases in resource densities and food-web stability.
+Therefore, my models suggest that it would be highly unlikely to ever observe an increase in resource densities in natural settings, although zero or negligible change is possible (under certain constraints).
+
+, although it is a simplistic representation of consumer-resource interactions [@Rosenzweig1963; @McCann2005; @McCann2011]
 
@@
 This means that energy transfer is more efficient to upper trophic levels, which creates more top-heavy food webs (cite McCann's work) that are less stable. 
+-->
 

### 2019-02-14 manuscript_ECD_model.Rmd
edited discussion of results when resources occur in distinct habitats and the first paragraph of the stability results.

@@
 
-My general finding was that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource densities and less resilient food webs. 
+My results suggest that a revision is needed for one of the criteria used to demonstrate character displacement. 
+Criteria #5 in @Schluter1992 states: "Sites of sympatry and allopatry should not differ greatly in food, ...". 
+In contrast, my results show that character displacement leads to predictable differences in food (resource) availability and even food-web stability.
+Specifically, I found<!--My general finding was--> that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource densities and less resilient food webs.
 Character displacement always increased attack rates whenever consumers were competing for resources that occured in distinct habitats. 
 This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
-Although character displacement did not always increase effective attack rates, these contingent effects only occurred under the simplest and least realistic foraging scenario. 
+Although character displacement did not always increase effective attack rates, these contingent effects only occurred under the simplest and (arguably) least realistic foraging scenario. 
 My study takes advantage of past work showing that adaptive divergence in consumer attack rates is always favored under the scenarios examined here [@LawlorSmith1976; @Abrams1986].
@@
 <!--Under this scenario, character displacement either results in no change in food-web dynamics or a decrease in both resource densities and stability.-->
-Under this scenario, it is theoretically possible that character displacement results in increased resource densities and food-web stability [green lines in \ref{fig:MacArthur_Plot}].
+Under this scenario, it is theoretically possible that character displacement results in increased resource densities<!-- and food-web stability--> [green lines in \ref{fig:MacArthur_Plot}].
 This only occurs, however, under an evolutionary tradeoff that constrains the magnitude of character displacement.
 This suggests that it would be unlikely to ever observe divergent character displacement and increases in resource densities in a real system. 
-In contrast, the evolutionary tradeoff that favors character displacement the most decreased resource densities and food-web stability. 
+In contrast, the evolutionary tradeoff that favors character displacement the most decreased resource densities<!-- and food-web stability-->. 
 Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the foraging rate of consumers (orange lines in \ref{fig:MacArthur_Plot}A,B).
 This is evident by the disruptive selection observed for beak size/shape in Darwin's finches [@Schluter1985] and benthic/limnetic ecotypes for stickleback [@Arnegard2014] in resource-limited environments.
-Taken together, my model predicts that even in the simplest foraging scenario, the most likely relationship we would observe in nature is for food webs where consumers have undergone character displacement to have lower total resource densites and be less resilient to perturbations.
+Taken together, my model predicts that even in the simplest foraging scenario, the most likely relationship we would observe in nature is for food webs where consumers have undergone character displacement to have lower total resource densites<!-- and be less resilient to perturbations-->.
 
@@
 <!-- Influence of foraging scenario and application to natural systems -->
-A general result that emerged from our analysis was that when consumers were competing for resources in distinct habitats, character displacement always surpressed resource densities, regardless of the functional form of the consumer's feeding rate or shape of the tradeoff.
-This is a very common foraging scenario in nature, as consumers tend to be more mobile than their resources, often coupling their dynamics across traditional habitat boundaries [@McCann2005].
-This may be especially so for character displacement, where a disproportionate number of the documented cases come from carnivores [@Schluter2000]. 
-Perhaps the clearest example of this would be the threespine stickleback in coastal lakes of British Columbia. 
-Here, the diet of stickleback is composed of two main resource types, zooplankton and benthic invertebrates. 
-These resource types occur in distinct zones of the lake (pelagic and littoral, respectively), and sticklebacks move between these zones when foraging for resources. 
-I predict that in species-pair lakes, where stickleback have undergone character displacement, that total resource densities will be lower compared to lakes with a single, generalist species.
-
-These results do lead to an interesting prediction that could be tested either experimentally or with field data. 
-If consumers are competing for resources that occur in different habitats, then greater character displacement corresponds to a greater decrease in resource densities (line trajectories in fig.
-\ref{fig:LS_Plot}B, \ref{fig:McCann_Plot}B).
-This prediction likely only applies, however, to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
-
-<!-- Consider moving this to the opening of the discussion or the opening of the conclusion section -->
-Our work also points to a potential revision for one of the criteria used to demonstrate character displacement. 
-Specifically, criteria 5 in @Schluter1992 states: "Sites of sympatry and allopatry should not differ greatly in food, ...". 
-Our results demonstrate that character displacement can alter resource distributions in a predictable way, usually resulting in lower resource availability where competition has resulted in character displacement.
+A general result emerging from our analysis was that, when consumers were competing for resources in distinct habitats, character displacement always supressed resource densities.
+<!--, regardless of the functional form of the consumer's feeding rate or shape of the tradeoff.-->
+Consumers are generally larger, and as a result, more mobile than their resources; thus, their foraging behavior acts to couple resource dynamics across traditional habitat boundaries [@McCann2005].
+<!--A disproporationate number of the documented cases of character displacement involve carnivores [@Schluter2000] that are larger, and as a result, often more mobile than their resources [@McCann2005].-->
+An example of character displacement that matches this foraging scenario is threespine stickleback in postglacial lakes of British Columbia, Canada [@Schluter1992; @Schluter2000].
+In these lakes, stickleback have undergone character displacement into limnetic and benthic species that have evolved specialized traits to forage on zooplankton and benthic invertebrates, respectively.
+These resources occur in distinct zones of the lake (pelagic and littoral) and stickleback move between these zones when foraging for resources. 
+The theory developed here predicts that total resource densities will be lower in lakes where stickleback have undergone character displacement compared to nearby lakes with only a single, generalist species.
+A disproporationate number of the documented cases of character displacement involve carnivores that are larger, and likely more mobile, than their resources [@Schluter2000],
+Therefore, my predictions may apply to many cases of ecological character displacement.
+
+Character displacement is far from an all-or-nothing process, but may vary in its magnitude [@Schluter2000]. 
+My results suggest that the magnitude of character displacement will be negatively correlated with total resource densities, at least when resources occur in distinct habitats (line trajectories in \ref{fig:LS_Plot}B, \ref{fig:McCann_Plot}B).
+This prediction could be test either experimentally, by putting populations that vary in divergence in a common resource setting, or with field data that statistically controls for other factors affecting resources.
+Note though, this prediction will only apply to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
 
 <!-- Food-web stability -->
-Our most stiking result was that character displacement, when examined through the lens of food-web theory, usually resulted in less stable food webs. 
+My most striking result was that character displacement, when examined through the lens of food-web theory, generally resulted in less stable food webs. 
 In fact, under a foraging scenario that incorporates more biological realism, character displacement may even result in an unstable food web. 
-This result aligns well with the general result from food-web theory that food-web stability tends to go hand-in-hand with resource densities [@May1973; @Murdoch2003; @McCann2011]. 
-However, this result contrasts the notion from competition theory that character displacement contributes to species coexistence. 
-My results do not contradict these past inferences, as they were based on a different measure of stability that considers the ability of a consumer with a different phenotype to invade when rare (invasion criteria).
-However, its clear that stability measures do not always go hand-in-hand and that a pluralistic view of different aspects of stability is needed [@Donohue2013; @Donohue2016] for understanding eco-evolutionary feedbacks on food-web dynamics.
+This result aligns well with the general finding from food-web theory that stability tends to go hand-in-hand with resource abundances [@May1973; @Murdoch2003; @McCann2011]. 
+Yet, this result contrasts the notion from competition theory that character displacement contributes to species coexistence. 
+My results do not necessarily contradict these past inferences, as competition theory uses a different measure of stability that considers the ability of a consumer with a different phenotype to invade when rare [invasion criteria, @Chesson2000].
+Instead, my results highlight that stability measures do not always go hand-in-hand [@Donohue2013; @Donohue2016].
+A more pluralistic view of stability is likely needed for understanding the consequences of eco-evolutionary feedbacks in a food-web context.
+
+<!-- But talk about the competition inference, and how it may not allow a proper view of how ecological character displacement affects stability, because it also does not compare the scenarios.
+For example, convergence is clearly favored in the absence of a competitor.
+It also doesn't account for the stability of the fully coupled food web that includes resources, which we know is a key contribute to the dynamics of ecological character displacement -->
 

### 2019-02-28 manuscript_ECD_model.Rmd
editing stability discussion

@@
 In fact, under a foraging scenario that incorporates more biological realism, character displacement may even result in an unstable food web. 
-This result aligns well with the general finding from food-web theory that stability tends to go hand-in-hand with resource abundances [@May1973; @Murdoch2003; @McCann2011]. 
-Yet, this result contrasts the notion from competition theory that character displacement contributes to species coexistence. 
-My results do not necessarily contradict these past inferences, as competition theory uses a different measure of stability that considers the ability of a consumer with a different phenotype to invade when rare [invasion criteria, @Chesson2000].
+<!--This result aligns well with the general finding from food-web theory that stability tends to go hand-in-hand with resource abundances [@May1973; @Murdoch2003; @McCann2011]. -->
+The mechanism underlying this destabilization is quite general, resulting from the strength of interspecific interactions increasing relative to intraspecific interactions [@Chesson2008; @Rip2011; @McCann2011].
+The strength of interspecific interactions increases whenever consumers evolved higher net attack rates, but does not change intraspecific interactions.
+This relative increase in interspecific interactions, combined with the natural oscillatory tendency of consumer-resource interactions [@Murdoch2003], creates a food-web structure that is less resilient to perturbations.
+<!-- creates top-heavy food webs (higher $(C_i+C_j)$:$(R_i+R_j)$) that are -->
+
+My results, combined with recent theory, suggest that the very conditions that favor phenotypic diversification also make that diversity more susceptible to local extinction.
+<!--Recent theory suggests that the conditions that favor ecological character displacement are the very ones that would enhance the destabilizing effects of this eco-evolutionary process.-->
+Specifically, work by @McPeek2019 showed that character displacement is favored in environments that are either highly productive, easy to find and capture resources, or under weak abiotic stress. 
+In my model, this corresponds to higher values of $K$ (productivity) or $a_ij$s (finding and capturing resources), or lower values of $m$ (abiotic stress).
+Each of these corresponding changes create food webs that have a lower baseline of stability, as they increase the strength of consumer-resource interactions relative to intraspecific density dependence.
+For example, increasing productivity reduces intraspecific competition in resource populations while increasing the flux of energy to consumers, resulting in the "paradox of enrichment" [@Rosenzweig1971].
+<!--where higher productivity, all else equal, can decrease food-web stability due to consumers suppressing resources more below their carrying capacity. -->
+Similarly, increased capacity to find and capture resources (as we observed in our model) or lower consumer mortality, both increase the relative strength of consumer-resource interactions, which predictably destabilize consumer-resource interactions [@Rip2011; @McCann2011].
+<!--This suggests that the very conditions that favor phenotypic diversification also make that diversity more susceptible to local extinction.-->
+<!--Perhaps nonadaptive selection on ecological stability [@Borrelli2015a; @Borrelli2015b], -->
+
+<!-- Empirical support for notion of instability: Ghost of competition and stability past --->
+Empirical evidence supports the theoretical notion that food webs where consumers have undergone character displacement are less resilient to perturbations. <!--is not far-fetched.-->
+<!--Threespine stickleback in postglacial lakes of British Columbia, one of the most well-studied examples of character displacement, are known to have undergone character displacement in six different -->
+For example, <!--It is worth noting that -->a single species of threespine stickleback lives in hundreds of postglacial lakes of British Columbia, but the species pairs, where character displacement has happened, are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
+Perhaps this rarity of species pairs is a result of nonadaptive selection for more ecologically stable food-web topologies [@Borrelli2015a; @Borrelli2015b].
+Indeed, it is clear that these species pair are vulnerable to perturbations, given that the species pairs have went extinct in two lakes following the introduction of nonnative species [@Hatfield2001; @Taylor2006; @Rudman2016].
+The vulnerability of the stickleback system also aligns well with evidence that aquatic food webs have several properties that make them less resilient to perturbations, including higher productivity and more efficient transfer of energy to consumers [@Rip2011].
+Testing this hypothesis may be quite difficult, but it could be possible with recent advances in genomics.
+For example, @Feulner2018 detected genomic signatures of hybridization in sympatric whitefish species following periods of eutrophication.
+Perhaps stickleback in many other postglacial lakes in British Columbia retain genomic signatures of having been a habitat specialist in the past.
+<!-- could look at this in lakes that don't have a species pair, but are in nearby watersheds -->
+
+<!-- contrasts with competition theory -->
+My stability results contrast, but do not necessarily contradict, the notion from competition theory that character displacement contributes to species coexistence [@Lawlor1976]. 
+Competition theory typically uses a different measure of stability that considers the mutual ability of consumers with different phenotypes to invade when rare [mutual invasibility, @Chesson2000].
+A shortcoming of this mutual invasibility measure, however, is that it does not allow a comparison between food webs with and without a competiting consumer. 
+Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement [@Schluter1992; @Schluter2000].
+Although the addition of consumers can decrease food-web resilience in the absence of evolution [@May1973], our results are primarily driven by an eco-evolutionary feedback between consumer evolution and resource abundances.
+
+<!--
+Typically, they assume that resource distributions are stable, although resource instability can enhance coexistence if consumers contain different forms of their functional responses [@Armstrong1976a; @Armstrong1976b; @Abrams2002].
 Instead, my results highlight that stability measures do not always go hand-in-hand [@Donohue2013; @Donohue2016].
@@
 
-<!-- But talk about the competition inference, and how it may not allow a proper view of how ecological character displacement affects stability, because it also does not compare the scenarios.
-For example, convergence is clearly favored in the absence of a competitor.
-It also doesn't account for the stability of the fully coupled food web that includes resources, which we know is a key contribute to the dynamics of ecological character displacement -->
-
-<!-- Ghost of competition and stability past --->
-A single species of threespine stickleback lives in hundreds of lakes of British Columbia, but the species pairs, where character displacement has happened, are only known from four lakes one of which has undergone reverse speciation [@Rudman2016].
-Our results hint at an interesting hypothesis for the apparent rarity, namely founded on ecological stability [@Borrelli2015a; @Borrelli2015b]. 
-Specifically, the commonness of the lakes with only a single species is due to the higher ecological stability of this food-web structure, compared to the one with a species pair. 
-Testing this hypothesis may be quite difficult, but it could be possible with recent advances in genomics. 
-For example, the species-pair lake that has undergone reverse speciation has a different ecological impact than other generalist lakes in that it is much more like the benthic-ecotype in its morphology.
-It is unclear whether this is a transient or persistent evolutionary state.
-But advances in genomics may be able to tease apart whether a currently generalist species had a genomic imprint of being more benthic or limnetic.
-
-<!-- Predictions for a natural system -->
-Drawing on the theory I developed, I can point to systems where we would expect character displacement to result in less stable food webs. 
-My theory suggests that this may be more likely to occur when character displacement occurs between: (i) consumers foraging in distinct habitats; (ii) exhibit concave-up tradeoff in attack rates; and (iii) in aquatic systems.
-Aquatic systems are more likely because its been shown that attack rates are often much stronger in these systems [@Rip2011]. 
-This lower baseline, means that it is easier for character displacement to push these systems toward the boundary of local stability. 
-Interestingly, one of the hallmark examples of character displacement, threespine stickleback, meets these criteria.
-Thus, I predict that in the coastal lakes of British Columbia, where stickleback have undergone character displacement, for these food webs to exhibit less stable dynamics and more variable population abundances.
-Although stability is a traditionally thorny issue to measure, there are now clear theoretical links between the stability measure we used here ($-\lambda_{\text{max}}$) and the coefficient of variation (CV), which simply measures the variability in a dynamic time series relative to the mean.
-Also, there are now methods for assessing the dynamics of natural communities that link to these theoretical stability measures [@Ives2003]. Now it just requires the effort to collect this data. 
-Also, in applying this theory to a natural system, an assessment of stability for a specific system requires a realistic model of the population and community dynamics.
-e.g.
-stage-structure,...Still, our results point toward a general tendency toward instability.
-
-It's entirely plausible though that the some aspects of the system, e.g.
-stage structure could prevent this instability.
-This could be quite interesting, as this may point to systems where we are more likely to observe character displacement, because it enhances their stability.
-
+Of the prior models of ecological character displacement, only @Lawlor1976 examine the effect of character displacement on the ecological stability of the system.
+They conducted a detailed theoretical analysis of three different metrics of stability: (i) stable coexistence of consumers, defined as the mutual invasibility of each consumer; (ii) global stability of consumers, defined as the continued coexistence of both consumers when resource dynamics are perturbed; and (iii) local stability, defined as the rate at which the four-species food web returned to equilibrium after a small perturbation to species abundances (the measure we also used).
+Note that the first two metrics correspond to the stable coexistence of consumers, while the third metric measures the stability of the entire food web.
+@Lawlor1976 found that character displacement enhanced the stable coexistence of consumers, yet had a small negative effect on the local stability of the food web.
+@Lawlor1976's work gave fundamental insight to the effect of character displacement on ecological stability; however, none of their stability analyses compared food webs with and without a competing consumer.
+Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement (@Schluter2000; @Schluter1992).
+In fact, the first two metrics of stability can only be assessed when both consumers are present because it examines the effect of divergence on coexistence, which requires both consumers to be present.
+In contrast, local stability can be compared in food webs with and without a competing consumer, and thus provides a useful benchmark for examining the effects of character displacement on food-web stability.
+-->
+
+<!-- Caveats -->
+An advantage of the Adaptive Dynamics approach is that we could examine the ecological dynamics of the system.
+Current consumer-resource models based on quantitative genetics become quickly intractable to gain analytical insight from, even for simple linear functional responses of consumers [@McPeek2019].
+Those approaches have clear advantages in that they explicitly model a quantitative trait, which then influences the ecological rates (*r, K, A, e, m*, etc.).
+Our approach assumes that there is a one-to-one correspondence between trait evolution and the ecological rate.
+This simplifying assumption, however, allowed us to explore a much more realistic biological scenario and still gain analytical insight.
+This necessary tradeoff was worth it, we feel, because we were interested in the ecological consequences of character displacement, rather than in identifying the conditions resulting in character displacement and that alter the magnitude, as excellent work has already been done on this front [e.g.
+@Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019].
+
+Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
+quantitative genetics).
+I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits [see synthesis by @Taper1985].
+Since these models are focused primarily on trait evolution, it is not clear that divergence in these models would have different ecological effects.
+In fact, recent work by @McPeek2019 has shown that genetic divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
+Finally, my conclusions may only apply to biological resources that are nutritionally substitutable. 
+It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
+
+While the underlying genetics of trait evolution may influence the amount of divergence in character displacement, this should not influence the effect of that trait evolution on the consumer-resource dynamics.
+Another way our results may quantitatively differ from a quantitative genetic model is that there is no intraspecific variation in our approach.
+While this may certainly have quantitative effects on the relationship between character displacement and food-web dynamics, or more likely, influence the magnitude of displacement [@McPeek2019], it is unclear whether this would qualitatively influence our inferences.
+
+Throughout, I have also setup a special case scenario where consumers are mirror images of each other.
+This resulted in symmetric character displacement, which is quite common in empirical data [@Schluter2000].
+Again though, I do not expect that adding mild asymmetries (ones that still would permit coexistence) would qualitatively change our results.
+This is because, again, divergent character displacement is still evident in these models.
+Also, we conducted simulations to test whether our results were robust to this assumption and they were (ADD SUPPLEMENTARY MATERIAL).
+
+<!--
+All of these result from the general principle that any biological mechanism that increases the strength of interspecific interactions relative to the strength of intraspecific interactions acts to destabilize consumer-resource dynamics [@Chesson2008; @Rip2011; @McCann2011].
+Mathematically, this happens when the equilibrium point of the system is associated with an imaginary eigenvalue.
+This causes the system to exhibit damped oscilattions creates a tendency for the system to oscillate in the presence of any lag in consumer-resource dynamics., which creates a puts some "spin" on the system.
+Otherwise, biological mechanisms that enhance interaction strength can be initially stabilizing, if the consumer is poorly adapted to foraging on the available resources (see first parts of stability figures).
+
+is the imaginary eigenvalue associated with the lag itself? A presence of a lag in the system?
+-->
+
+<!-- environmental gradients and character displacement 
 It may be quite common for character displacement to result in less stable food webs. 
@@
 Otherwise, biological mechanisms that enhance interaction strength can be initially stabilizing, if the consumer is poorly adapted to foraging on the available resources (see first parts of stability figures).
-
-<!-- Caveats -->
-An advantage of the Adaptive Dynamics approach is that we could examine the ecological dynamics of the system.
-Current consumer-resource models based on quantitative genetics become quickly intractable to gain analytical insight from, even for simple linear functional responses of consumers [@McPeek2019].
-Those approaches have clear advantages in that they explicitly model a quantitative trait, which then influences the ecological rates (*r, K, A, e, m*, etc.).
-Our approach assumes that there is a one-to-one correspondence between trait evolution and the ecological rate.
-This simplifying assumption, however, allowed us to explore a much more realistic biological scenario and still gain analytical insight.
-This necessary tradeoff was worth it, we feel, because we were interested in the ecological consequences of character displacement, rather than in identifying the conditions resulting in character displacement and that alter the magnitude, as excellent work has already been done on this front [e.g.
-@Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019].
-
-Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
-quantitative genetics).
-I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits [see synthesis by @Taper1985].
-Since these models are focused primarily on trait evolution, it is not clear that divergence in these models would have different ecological effects.
-In fact, recent work by @McPeek2019 has shown that genetic divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
-Finally, my conclusions may only apply to biological resources that are nutritionally substitutable. 
-It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
-
-While the underlying genetics of trait evolution may influence the amount of divergence in character displacement, this should not influence the effect of that trait evolution on the consumer-resource dynamics.
-Another way our results may quantitatively differ from a quantitative genetic model is that there is no intraspecific variation in our approach.
-While this may certainly have quantitative effects on the relationship between character displacement and food-web dynamics, or more likely, influence the magnitude of displacement [@McPeek2019], it is unclear whether this would qualitatively influence our inferences.
-
-Throughout, I have also setup a special case scenario where consumers are mirror images of each other.
-This resulted in symmetric character displacement, which is quite common in empirical data [@Schluter2000].
-Again though, I do not expect that adding mild asymmetries (ones that still would permit coexistence) would qualitatively change our results.
-This is because, again, divergent character displacement is still evident in these models.
-Also, we conducted simulations to test whether our results were robust to this assumption and they were (ADD SUPPLEMENTARY MATERIAL).
+-->
+
+<!-- But talk about the competition inference, and how it may not allow a proper view of how ecological character displacement affects stability, because it also does not compare the scenarios.
+For example, convergence is clearly favored in the absence of a competitor.
+It also doesn't account for the stability of the fully coupled food web that includes resources, which we know is a key contribute to the dynamics of ecological character displacement -->
+
+<!-- Environmental gradients 
+My theoretical results also give insight to where we might expect to observe the destabilizing effects of character displacement in nature. 
+Specifically, environments that increases the flux of energy to consumers relative to resources generally decrease stability [@Rip2011; @McCann2011], which would amplify the destabilizing effects of character displacement.
+For example, aquatic food webs have a number of attributes that make them less stable than terrestrial food webs [@Rip2011].
+-->
+<!-- excerpts from Rip and McCann 2011
+aquatic primary producers possess less structural material
+(Elser et al. 2000) and contain more photosynthetic tissue (Elser et al.
+2000). These differences result in higher growth rates in aquatic
+primary producers (Cebrian 1999; Elser et al. 2000).
+
+Many differences in consumption rates exist across ecosystem types.
+These differences arise from a number of attributes differing across
+ecosystems. In aquatic ecosystems, the body size ratios of herbivores
+to primary producers are larger (Hairston & Hairston 1993; Chase
+2000) and primary producers are more nutritious to herbivores (Polis
+& Strong 1996; Cebrian 1999; Elser et al. 2000). This results in greater
+herbivory in aquatic ecosystems (Cyr & Pace 1993; Hairston &
+Hairston 1993; Cebrian 1999). The higher consumption rates in
+aquatic ecosystems should have a large impact on parameters used in
+C-R theory. Higher consumption rates may manifest in higher
+maximum consumption rates and higher conversion efficiencies (i.e. a
+or aMax and e). As seen above, an increase in either or both of these
+parameters would result in an increase in energy flow, more inverted
+biomass pyramids, and a decrease in stability. Therefore we predict
+that aquatic ecosystems should be less stable than terrestrial
+ecosystems.
+
+. However, the amount of mass-specific structural material
+(e.g. skeletal material, woody tissue) does positively scale to body
+mass: scaling exponents vary between 0.4 and 0.12 suggesting that
+efficiency should scale across ecosystem types (Schmidt-Nielson
+1984). As conversion efficiency is dimensionless, these positive
+exponents become negative when we divide through by biomass to
+remove the dimensions (exponents become )0.88 to )0.40). Thus, as
+others have argued (Strong 1992, Polis et al. 2000), aquatic pathways
+ought to tend to have far less structural material and promote high
+conversion efficiencies.
+
+In summary, the relationships suggest that the isocline should shift
+left relative to the resource isocline with reduced body size. Moreover,
+these arguments resonate with previous arguments. Strong (1992) and
+Polis et al. (2000) argued cogently that aquatic systems were more
+homogenised reducing prey refugia, more edible and tended to
+experience higher turnover rates than terrestrial ecosystems. All these
+effectively argue for higher energy flux through aquatic pathways as
+our allometric reasoning also suggests. Thus, differences in the
+attributes of aquatic and terrestrial ecosystems relate to differences in
+energy flow within ecosystems. Comparing cross-ecosystem differences
+to parameters used in C-R theory, we predict aquatic ecosystems
+should have high growth rates of primary producers and higher
+numerical responses by consumers. All of these predictions taken
+together suggest that energy flow in aquatic ecosystems should be
+higher, biomass pyramids more inverted and stability decreased
+relative to terrestrial ecosystems.
+
+Strong, D.R. (1992). Are trophic cascades all wet? Differentiation and donorcontrol
+in speciose ecosystems. Ecology. 73:747–754.
+
+Polis, D.A., Sear, A., Huxel, G.R., Strong, D.R. & Maron, J. (2000). When is a
+trophic cascade a trophic cascade? Trends Ecol. Evol., 11, 473–475.
+-->
+
+<!-- Predictions for a natural system 
+My results give insight to where we would expect character displacement to result in less stable food webs in nature: 
+(i) consumers compete for resources in distinct habitats; (ii) exhibit concave-up tradeoff in attack rates; (iii) high productivity (*K*); (iv) low structural complexity (*a_ii_*,*a_ij*); (v) weak stressors (low *m*).
+
+Aquatic ecosystems have a number of attributes that make them generally less stable than terrestrial ecosystems [@Rip2011]. For example,
+
+Aquatic systems are more likely because its been shown that attack rates are often much stronger in these systems [@Rip2011]. 
+This lower baseline, means that it is easier for character displacement to push these systems toward the boundary of local stability. 
+Interestingly, one of the hallmark examples of character displacement, threespine stickleback, meets these criteria.
+Thus, I predict that in the coastal lakes of British Columbia, where stickleback have undergone character displacement, for these food webs to exhibit less stable dynamics and more variable population abundances.
+Although stability is a traditionally thorny issue to measure, there are now clear theoretical links between the stability measure we used here ($-\lambda_{\text{max}}$) and the coefficient of variation (CV), which simply measures the variability in a dynamic time series relative to the mean.
+Also, there are now methods for assessing the dynamics of natural communities that link to these theoretical stability measures [@Ives2003]. Now it just requires the effort to collect this data. 
+Also, in applying this theory to a natural system, an assessment of stability for a specific system requires a realistic model of the population and community dynamics.
+e.g.
+stage-structure,...Still, our results point toward a general tendency toward instability.
+
+It's entirely plausible though that the some aspects of the system, e.g.
+stage structure could prevent this instability.
+This could be quite interesting, as this may point to systems where we are more likely to observe character displacement, because it enhances their stability.
+-->
 

### 2019-03-01 manuscript_ECD_model.Rmd
edit caveat paragraph in Discussion

@@
 <!-- Caveats -->
+I expect that my theoretical predictions are robust to different assumptions about the underlying evolutionary dynamics of consumer foraging traits.
+This is because models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of whether a quantitative genetic or adaptive dynamics model is used for consumer traits [@Lawlor1976; @Taper1985].
+A quantitative genetic model may certainly show differences in the magnitude of character displacement, but this should not affect the qualtiative relationship I observed.  
+In fact, recent work by @McPeek2017 has shown that divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
+I want to note though, that my conclusions may only apply to food webs with biotic resources that are nutritionally substitutable. 
+It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
+
+<!-- Conclusions -->
+
+<!--
+Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
+quantitative genetics).
+I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits [see synthesis by @Taper1985].
+Since these models are focused primarily on trait evolution, it is not clear that divergence in these models would have different ecological effects.
+In fact, recent work by @McPeek2019 has shown that genetic divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
+Finally, my conclusions may only apply to biological resources that are nutritionally substitutable. 
+It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
+
 An advantage of the Adaptive Dynamics approach is that we could examine the ecological dynamics of the system.
@@
 @Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019].
-
-Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
-quantitative genetics).
-I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits [see synthesis by @Taper1985].
-Since these models are focused primarily on trait evolution, it is not clear that divergence in these models would have different ecological effects.
-In fact, recent work by @McPeek2019 has shown that genetic divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
-Finally, my conclusions may only apply to biological resources that are nutritionally substitutable. 
-It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
 

### 2019-03-07 manuscript_ECD_model.Rmd
added conclusion pargraph

@@
 <!-- Conclusions -->
+Competition theory suggested that character displacement was a stabilizing force in ecological communities [@Lawlor1976]. 
+My results show the exact opposite, when a food-web perspective is taken.
+Importantly, a food-web approach permits comparisons between communities in the presence of absence of a competitor, allowing a fair comparison of the effects of character displacement on food-web dynamics.
+These results appear to be quite general, as they characterize the general foraging scenario where consumers are more mobile than their resources. 
+These results also suggest that the very evolutionary process that generates biodiversity, creates an ecological scenario that makes that diversity fragile to loss. 
+These results highlight how evolution, when considered in a food-web context, may actually be a destabilizing force in ecological communities. 
+This may generally be true of consumer evolution, as was pointed out decades ago [@Rosenzweig1963].
+These results also highlight a tension between adaptive process (evolution of consumers to increase feeding efficiency) and a non-adaptive process (food-web stability).
+These results suggest that in systems with evidence of ecological character displacement, why do they persist? 
+These results open up a new avenue for research on ecological character displacement, one that explores in more detail the dynamical consequences of consumer evolution, rather than just static effects that assume resources are persistent and stable.
+It also suggest that actually observing evidence of ecological character displacement may be difficult ...
+It also highlights how evolution of key species in a food web may make these systems more susceptible to environmental change. 
+It's been argued that "inferences about the evolution of community structure cannot be based on ecological arguments about community stability or likely coexistence" [@McPeek2019].
+In contrast, our results suggest that current patterns in food-web structure may be the result of non-adaptive processes selecting for more stable food-web configurations.
+Our results also call for a revision of one of the criteria of character displacement, since it is clear that this process can result in different resource dynamics.
+In fact, this resource similarity hypothesis has not been tested in natural systems, although it has been qualitatively inferred, and some quantitative estimates based on physical properties of systems have been measured to try and say these habitats are similar.
+(NEED TO CHECK SCHLUTER 2000 BOOK, THE AM NAT STUDY SUGGESTS THAT IT IS A COMMONLY FILLED CRITERIA.)
+More generally, our results highlight how variation in food-web structure can alter eco-evolutionary feedbacks within communities [@McPeek2017], that also have consequences for their resilience to environmental disturbances.
+This calls for food-web ecologists to more criticaly consider the role of evolutionary processes in shaping food-web structure and dynamics.
 

### 2019-03-07 manuscript_ECD_model.Rmd
delete old text I was keeping as notes

@@
 
-<!-- Outline
-
-Introduction
-
-1. ECD is an important process, and current empirical and theoretical work suggests it often results in divergence of consumer foraging traits.
-
-2.
-Ecological consequences of character displacement have received surprisingly little attention.
-In particular, prior work doesn't allow a proper comparision of the effects of character displacement on community stability.
-
--> Another paragraph illustrating how the direct and indirect effects of character displacement likely have constrasting effects on stability.
-
-3.
-Here we fill this knowledge gap.
-We focus on the ecological side of these dynamics, because the evolutionary aspects have been well characterized, although we show in the appendices that the evolutionary dynamics are as expected (especially for McCann model).
-
-Methods and Results
-
-- Underlying model of consumer-resource dynamics
-  1. Explanation of model parameters
-  2. Methods for assessing resource abundances and stability.
-  3. Trade-off function and definition of character displacement.
-  4. Analytical and simulation methods. Briefly mention the different scenarios I'll explore.
-
-- Consumers can forage for resources simultaneously
-  1. Functional response equation and illustration of assumptions
-  2. Effect on attack rates, resource densities, and stability
-
-- Consumers cannot foraging for resource simultaneously
-  1. Functional response equation and illustration of assumptions
-  2. Effect on attack rates, resource densities, and stability
-
-- McCann's most realistic functional response
-  1. Functional response equation and illustration of assumptions
-  2. Effect on attack rates, resource densities, and stability
-
-Discussion
-
--->
-
 ```{r setup, include=FALSE}
@@
 While character displacement did not always increase the strength of consumer-resource interactions, these contingent effects only occurred under the simplest and least realistic foraging scenario. 
-<!--In the simple ecological scenario that has been the focus of past models of character displacement (type 1 functional response and competing for resources in the same habitat), I find that the shape of the tradeoff can qualitatively affect the relationship between character displacement, resource densities, and food-web stability.
-Extensions of this model, ones that include other foraging scenarios with more realistic functional responses (e.g.
-saturating) and that reflect the biological reality that consumers are often more mobile than their resources, I find predictable effects of character displacement on resource densities and food-web stability.
-Competition theory has generally shown that character displacement enhances consumer coexistence.
-I show that this adaptive process comes at the ecological cost of making food webs less resilient to perturbations.  -->
 Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations. 
-They also suggest that food-web and competition theory can give complimentary insights to the ecological consequences of adaptation. <!-- eco-evolutionary feedbacks -->
+They also suggest that food-web and competition theory can give complimentary insights to the ecological consequences of adaptation.
 
@@
 # Introduction
-
-<!-- ECD is an important process, and current empirical and theoretical work suggests it often results in divergence of consumer foraging traits.-->
 
@@
 
-<!-- Ecological consequences of character displacement have received surprisingly little attention.
-In particular, prior work doesn't allow a proper comparision of the effects of character displacement on community stability.-->
-
 Although explicit models of resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics have received surprisingly little attention. 
@@
 For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition.
-<!--
-On the other hand, specialization often increases a consumer’s ability to capture a specific resource.
-From food-web theory, we know that increased feeding rates often lead to resource suppression, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics [Rosenzweig 1971; Murdoch et al.
-2003; McCann 2011].
--->
 Yet, this reduction in interspecific competition may, at the same time, increase the interspecific interaction between a consumer and the resource it specializes on.  
@@
 Thus, a food-web perspective, which accounts for both the direct and indirect effects of consumer-resource interactions, may yield new insight to the ecological consequences of character displacement.
-
-<!--Thus, a consumer-resource perspective may yield new insight to the ecological feedback of character displacement on food-web dynamics. 
-<!--Given that character displacement is mediated by a consumer-resource interaction, a food-web perspective may yield   NEED A LAST SENTENCE TO WRAP IT UP, HIGHLIGHTING HOW A FOOD-WEB PERSPECTIVE MAY YIELD NEW ECOLOGICAL INSIGHTS AND GIVE A CONTRASTING PERSPECTIVE TO COEXISTENCE THEORY.
-PERHAPS A I NEED TO INCLUDE A SENTENCE MID-WAY THROUGH SUGGESTING THAT A FOOD-WEB PERSPECTIVE MAY GIVE QUALITATIVELY DIFFERENT INSIGHTS.
-IT MAY BE NICE TO BRING IN THE SPATIAL COUPLING WORK OF MCCANN SUGGESTING THAT SPECIALIZATION CAN COUPLE FOOD-WEBS, BUT THIS WAS ONLY WITH ONE TOP PREDATOR (I.E.
-NO CONSUMERS...LOOK AT OLD TEXT FOR WORDING SUGGESTIONS.)-->
-
-<!-- older idea...Write 3rd paragraph of intro, highlighting how the direct effects of character displacement, surprisingly, have been ignored.
-We are currently lacking insight to the relative importance of these direct and indirect effects of character displacement on community dynamics.
-
-<!--
-The role of character displacement in governing food-web dynamics remains unclear A food-web perspective, however, has been lacking from theoretical work on the ecological effects of character displacement.
-, thus the consequences of this process for food-web stabilit
-
-<!--
-Despite the key insights made by explicit models of resource dynamics,
-
-gave key insights to the evolutionary dynamics of character displacement, the ecological effects on resources dynamics have received surprisingly little attention.
-Perhaps this is because the ecological dynamics have been studied primarily through the lens of competition and coexistence thoery
-
-This oversight is unfortunate, given that the stability of consumer-resources is often closely tied with resource abundances.
-
-Although character displacement is the result of an eco-evolutionary feedback, the ecological consequences of this process have received surprisingly little attention [@Germain2018].
-Perhaps this is because early theoretical work suggested that character displacement was a potent stabilizing force in ecological communities [@Lawlor1976].
-This ecological outcome is intuitive in light of competition theory ---character divergence reduces interspecific competition between consumers, promoting coexistence.
-On the other hand, specialization often enhances a consumer’s ability to capture a specific resource.
-A well known consequence of increased foraging efficiency (all else equal) is resource suppression, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics (Rosenzweig 1971; Murdoch et al.
-2003; McCann 2011).
-
-Moreover, past inferences about the stabilizing effects of character displacement have only been made in communities where both consumers are present.
-However, I argue that inferences about the effect of character displacement, like the process itself, should only be made when comparing communities in the presence and absence of competition [criteria #3 for character displacement, @Schluter1992; @Schluter2000].
-Therefore, the role of ecological character displacement in determining community stability remains unclear.
-
-<!--
-However, inferences about the effect of character displacement, like the process itself, can only be made when comparing communities in the presence and absence of competition [criteria #3 for character displacement, @Schluter1992; @Schluter2000].
-Current inferences about the stabilizing effects of consumer divergence are only comparisons about communities where there are two consumers, and even though, the focus is only on the stable coexistence of consumer species.
-
-<!--
-The consequence of this eco-evolutionary feedback for consumer coexistence has been well characterized (Lawlor and Smith 1976; Germain's work).
-Ecological character displacement has typically been considered to be a stabilizing factor in communities (Lawlor and Smith 1976).
-This is because divergence in foraging traits is thought to reduce the strength of interspecific competition between consumers.
-Reducing the strength of interspecific competition relative to intraspecific competition is a stabilizing force in consumer-resource systems.
-Consumer specialization on distinct resources suggests that there are niche differences between consumers.
-Mathematically, this is also often assumed to be the case, because using techniques such as evolutionary invasion analysis describe the ability for a consumer with a mutant allele to increase in abundance when rare, a demographic signature of niche differentiation.
-Therefore, if both consumers are able to invade when rare, then this suggests that they can coexist.
-
-Character displacement involves evolutionary shifts in character states in response to competition [@Schluter1992; @Schluter2000].
-Theoretically, this is determined by comparing character changes in the presence and absence of a competitor.
-Empirically, this is determined by meeting several criteria that rule out other possible explanations [@Schluter1992; @Schluter2000].
-Importantly, current inferences about the effect of character displacement on community stability do not meet this critera.
-For example, the inference that character displacement enhances species coexistence is only made in a scenario where consumers are competing.
-This type of stability is important, but it does not enable us to examine the effect of character displacement on community stability.
-As a consequence, I argue that the effect of character displacement on community stability remains an open question.
-
-Of the prior models of character displacement, only @Lawlor1976 examine the effect of character displacement on the ecological stability of the system.
-They conducted a detailed theoretical analysis of three different metrics of stability: (i) stable coexistence of consumers, defined as the mutual invasibility of each consumer; (ii) global stability of consumers, defined as the continued coexistence of both consumers when resource dynamics are perturbed; and (iii) local stability, defined as the rate at which the four-species food web returned to equilibrium after a small perturbation to species abundances.
-Note that the first two metrics correspond to the stable coexistence of consumers, while the third metric measures the stability of the entire food web.
-@Lawlor1976 found that character displacement enhanced the stable coexistence of consumers, yet had a small negative effect on the local stability of the food web.
-@Lawlor1976's work gave fundamental insight to the effect of character displacement on ecological stability; however, none of their stability analyses compared food webs with and without a competing consumer.
-Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement (@Schluter2000; @Schluter1992).
-In fact, the first two metrics of stability can only be assessed when both consumers are present because it examines the effect of divergence on coexistence, which requires both consumers to be present.
-In contrast, local stability can be compared in food webs with and without a competing consumer, and thus provides a useful benchmark for examining the effects of character displacement on food-web stability.
-
-<!-- -> Another paragraph illustrating how the direct and indirect effects of character displacement likely have constrasting effects on stability. -->
-
-<!--
-Indeed, ecological character displacement often results in the evolution of characters that enhance a consumer’s ability to capture resources, which tends to destabilize consumer-resource interactions (Murdoch et al.
-2003; McCann 2011).
-This occurs because increases in a consumer’s attack rate results in the resource being suppressed well below its carrying capacity, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics (Rosenzweig 1971; Murdoch et al.
-2003).
-Experimental decreases in interaction strength between consumers and resources has been shown to promote coexistence between consumers and resources (Luckinbill 1973).
-Moreover, there is broad support that specialized consumers, since they are tightly coupled to their resources, tend to exhibit consumer-resource cycles, whereas this is not the case for generalists (Murdoch et al.
-2002).
-Therefore, although evolutionary responses to exploitative competition may reduce the strength of interactions between consumers, they ultimately increase the strength of interactions with their resources which can destabilize the direct consumer-resource interaction (rather than the indirect consumer-consumer interaction).
-While prior models of character displacement have included more realistic functional responses and spatial relationships between consumers and resources, they have focused more on whether these factors alter patterns of character displacement rather than the effects of character displacement on the stability of the ecological community.
-As a consequence, it is currently unclear whether ecological character displacement has a net stabilizing or destabilizing effect on ecological communities.
-
-<!-- Here we fill this knowledge gap.
-We focus on the ecological side of these dynamics, because the evolutionary aspects have been well characterized, although we show in the appendices that the evolutionary dynamics are as expected (especially for McCann model).
--->
 
@@
 
-<!-- Explanation of model parameters -->
-
-To examine how ecological character displacement affects resource abundances and food-web stability<!--consumer-resource interactions and food-web dynamics-->, I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
+To examine how ecological character displacement affects resource abundances and food-web stability, I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
 
@@
 
-<!-- Methods for assessing resource abundances and stability. -->
-
 To investigate the ecological consequences, I studied the effect of adaptive divergence in consumers on resource densities and the stability of the entire food web at equilibrium.
@@
 
-<!-- Explain analytical methods -->
-
 I examine the effects of character displacement on food-web dynamics by (i) deriving analytical expressions for the relationship between consumer attack rates and food-web dynamics and (ii) simulating the effects of competition on the eco-evolutionary dynamics of consumers and resources.
@@
 While this scenario is arguably simplistic, it allows me to isolate the effects of ecological character displacement.
-<!--, rather than any dependence on inherent asymmetries between consumers or resources.-->
-
-<!-- Explain simulations -->
 
@@
 
-<!-- Redundant, removing for now
-In what follows, I examine the ecological consequences of character displacement under three different attack-rate tradeoffs (n>1, n=1, n<1) and in three different foraging scenarios: (i) resources occur in the same habitat and consumers exhibit a linear functional response (e.g.
-@MacArthur1972); (ii) resources occur in different habitats and consumers exhibit a linear functional response (e.g.
-@Lawlor1976); (iii) resources occur in distinct habitats and consumers exhibit a more realistic functional response [@McCann2005].
-I show that the consequences of character displacement for food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
--->
-
 ```{r Tradeoff, include=FALSE, fig.pos="H", fig.cap="\\label{fig:Tradeoff}Tradeoff forms for consumer attack rates."}
@@
 
-<!-- Functional response equation and illustration of assumptions already given in previous section -->
-
-<!-- Redundant, removing for now
-The evolution of consumer attack rates in this model has been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. -->
 When both consumers and resources are present, the density of resources at equilibrium are equivalent ($\hat R = \hat R_1 = \hat R_2$) and are determined by the following equation (derivation in supplementary material):
@@
 Curiously, this match between impact on resources and stability breakdowns with the concave down tradeoff (green).
-<!--This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (Fig.
-\ref{fig:Stability}).
-The effect of this eco-evolutionary feedback can be seen by comparing initial (small points) with evolved strategies (large points).
-For example, when the tradeoff is concave up (orange points), the three-species system (small triangle) is initially more stable than the four-species system (small circle), but this relationship reverses by the time consumers reach their evolutionary stable strategy.-->
 
@@
 
-<!-- Functional response equation and illustration of assumptions -->
-
 The previous model, although the starting point for many models of competition, does not reflect many food webs where consumers are more mobile than their resources.
@@
 In @Lawlor1976's detailed evolutionary analysis of this model, they always observed divergent character displacement.
-
-<!-- Effect on attack rates, resource densities, and stability -->
 
@@
 \ref{fig:LS_Plot}B).
-<!-- I will revisit this subtle point when I examine the effects of character displacement on food-web stability.
--->
 
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
   geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
   geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
 plot_LS_model <- plot_grid(
-  #tradeoff_plot + theme(legend.position = "none"), 
   LS_attack  + theme(legend.position = "none"), 
@@
 
-<!-- Functional response equation and illustration of assumptions -->
-
 So far I have only considered a simple functional response, where consumer feeding rates increase linearly with resource densities when foraging in either the same or distinct habitats.
@@
 
-<!-- Effect on attack rates, resource densities, and stability -->
-
 Even after adding more biological realism, I still observe the same qualitative relationship between character displacement and resource suppression (Fig.
@@
 In this more realistic model, however, whether the food web is locally stable depends on the parameters of both consumers and resources.
-Specifically, <!--I conducted the previous simulations on a scenario where the food web remains locally stable ($\operatorname{Re}(\lambda_{max})<0$).
-Is it possible that character displacement would ever result in a locally unstable food web?
-To gain analytical insight to this question, I examined the conditions governing the local stability of the four-species food web when consumers exhibit a more realistic functional response (@McCann2005).
-I show that -->the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derivation in supplementary material):
+Specifically, the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derivation in supplementary material):
 
@@
 Thus, character displacement always pushes the food web toward the boundary of local stability under this more realistic foraging scenario.
-<!--Whether or not the food web becomes locally unstable, however, depends on the choice of initial parameters.-->
 
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
   geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) +
+  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
   geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
+  ylab(expression("Stability "(-lambda[max]))) + 
   scale_shape_manual(values=c(24,21))  +
@@
 plot_McCann_model <- plot_grid(
-  #tradeoff_plot + theme(legend.position = "none"), 
   McCann_attack  + theme(legend.position = "none"), 
@@
 We did this for all foraging scenarios and tradeoffs previously examined. 
-<!--As expected, consumer asymmetry resulted in different ecological conditions (\ref{fig:MacArthur_Plot_asymm}B,C, \ref{fig:LS_Plot_asymm}B,C, \ref{fig:McCann_Plot_asymm}B,C).
-This is evident by seeing the deviations in the small circles (initial two consumers) from the lines that map the resource and stability dynamics of a symmetric two consumer system.
-Despite these initial differences, the system-->
 We found that our previous inferences were robust to including consumer asymmetry across different tradeoffs and foraging scenarios (\ref{fig:MacArthur_Plot_asymm}, \ref{fig:LS_Plot_asymm}, \ref{fig:McCann_Plot_asymm}).
@@
 # Discussion
-
-<!-- Summarize key findings -->
-
-<!--
-Our results suggest that qualitative predictions about the effect of character displacement on food-web dynamics can be made from knowledge of evolutionary constraints/tradeoffs in attack rates as well as consumer foraging behavior.
-For all of the tradeoffs and foraging scenarios we examined, competition drives divergence in consumer foraging traits.
-Yet, these different ecological and evolutionary constraints sometimes lead to qualitatively different effects on resources densities and food-web stability. 
-Specifically, the shape of the tradeoff resulted in qualitatively different relationships between character displacement and food-web dynamics in the foraging scenario most frequently examined in past models of the evolution of character displacement.
-However, in a foraging scenario that is likely more common in nature, where mobile consumers compete for resources that occured in distinct habitats, character displacement suppressed resource densities and resulted in less stable food webs.
-Taken together, our results illustrate how a food-web perspective can give novel insight to an eco-evolutionary process that appears common in nature [@Schluter2000; @Pfennig2010; but see @Stuart2013].
--->
 
@@
 In contrast, my results show that character displacement leads to predictable differences in food (resource) availability and even food-web stability.
-Specifically, I found<!--My general finding was--> that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource densities and less resilient food webs.
+Specifically, I found that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource densities and less resilient food webs. 
 Character displacement always increased attack rates whenever consumers were competing for resources that occured in distinct habitats. 
@@
 
-<!-- Influence of tradeoff shape and application to natural systems -->
-
-The simplest foraging scenario examined here (linear functional response on resources occuring in the same habitat) is the foundation of mechanistic models of resource competition [@MacArthur1972; @Lawlor1976; @Abrams1986]<!--substitutable resources-->.
-<!--Under this scenario, character displacement either results in no change in food-web dynamics or a decrease in both resource densities and stability.-->
-Under this scenario, it is theoretically possible that character displacement results in increased resource densities<!-- and food-web stability--> [green lines in \ref{fig:MacArthur_Plot}].
+The simplest foraging scenario examined here (linear functional response on resources occuring in the same habitat) is the foundation of mechanistic models of resource competition [@MacArthur1972; @Lawlor1976; @Abrams1986].
+Under this scenario, it is theoretically possible that character displacement results in increased resource densities [green lines in \ref{fig:MacArthur_Plot}].
 This only occurs, however, under an evolutionary tradeoff that constrains the magnitude of character displacement.
 This suggests that it would be unlikely to ever observe divergent character displacement and increases in resource densities in a real system. 
-In contrast, the evolutionary tradeoff that favors character displacement the most decreased resource densities<!-- and food-web stability-->. 
+In contrast, the evolutionary tradeoff that favors character displacement the most decreased resource densities. 
 Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the foraging rate of consumers (orange lines in \ref{fig:MacArthur_Plot}A,B).
 This is evident by the disruptive selection observed for beak size/shape in Darwin's finches [@Schluter1985] and benthic/limnetic ecotypes for stickleback [@Arnegard2014] in resource-limited environments.
-Taken together, my model predicts that even in the simplest foraging scenario, the most likely relationship we would observe in nature is for food webs where consumers have undergone character displacement to have lower total resource densites<!-- and be less resilient to perturbations-->.
-
-<!--
-In this scenario, the shape of the evolutionary tradeoff in attack rates qualitatively affected the relationship between character displacement and food-web dynamics. 
-In theory, resource densities and food-web stability can even be higher under character displacement if intermediate combinations of attack rates are higher than the extremes [green lines in \ref{fig:MacArthur_Plot}].
-But the magnitude of character displacement in this scenario is quite small, resulting in negligible increases in resource densities and food-web stability.
-Therefore, my models suggest that it would be highly unlikely to ever observe an increase in resource densities in natural settings, although zero or negligible change is possible (under certain constraints).
-
-, although it is a simplistic representation of consumer-resource interactions [@Rosenzweig1963; @McCann2005; @McCann2011]
-
-Although the shape of the tradeoff could qualitatively change the relationships between character displacement and food-web dynamics, I do not expect this to be common in real ecosystems. 
-This is because these distinct effects only occur in a foraging scenario that I expect is uncommon in nature (linear functional response of consumers competiting for resources in the same habitat). 
-It was still important to examine this scenario, however, as it is the basis of most theoretical models that explicitly model resource dynamics in competition [@MacArthur1972; @Lawlor1976; @Abrams1986].
-Even if we suspend our disbelief (for the moment) about the reality of the foraging situation, I expect that we would only be able to detect two effects of character displacement on food-web dynamics: no change in resource densities or stability; or decrease in resource densities and stability.
-I expect it would be unlikely to ever observe an increase in resource densities under character displacement, although its theoretical plausible (fig. \ref{fig:MacArthur_Plot}A). 
-This is because the concave down tradeoff does not favor strong divergence, and thus there is little potential for an increase in resource densities.
-It is unclear to me which tradeoff shape is most common in nature.
-Of the two examples I found, there appears to be a concave up tradeoff in consumer foraging traits for two classic examples of character displacement: threespine stickleback in coastal lakes of British Columbia [@Arnegard2014] and Darwin's finches on the Galapagos islands [@Schluter1985].
-A concave up tradeoff favors greater divergence, and thus is also the most likely evolutionary scenario to result in detectable character displacment.
-The shape of the tradeoff becomes evident from examining the form of natural selection on consumer traits in the absence of interspecific competition and when resources were limiting. 
-Specifically, in both systems there is evidence for disruptive selection (favoring extreme phenotypes) in environments with a broad spectrum of resources present, but where resources are ultimately limiting.
-Note that while this gives insight to the shape of the tradeoff, it does not give insight to the evolutionary dynamics of the system. 
-This is because, in the absence of interspecific competition, resource densities are the key determinant of consumer evolution (see supplementary material). 
-This is because even if specialization on one resource is favored in the short-term, the other resource will increase in density, favoring consumer to evolve traits that better exploit the other resource.
-Thus, there is an eco-evolutionary feedback that eventually favors an intermediate phenotype that is able to best take advantage of both resources (fig. \ref{fig:MacArthur_Plot}C).
-Based on the shape of their tradeoffs alone, I would predict character displacement in both of these systems to decrease total resource densities.
-I expect this prediction to hold particularly well for threespine stickleback, where the food-web dynamics more closely mimic this model (competition for two distinct resources, rather than competition along a continuous trait distribution, seeds for Darwin's finches).
-Whether this translates into a detectable effect on food-web stability is difficult, but if does occur, I would expect it to be more likely in threespine stickleback. 
-This is because there tends to be less of a stoichiometric mismatch between consumers and their resources in aquatic vs. terrestrial ecosystems. 
-This means that energy transfer is more efficient to upper trophic levels, which creates more top-heavy food webs (cite McCann's work) that are less stable. 
--->
-
-<!-- Influence of foraging scenario and application to natural systems -->
+Taken together, my model predicts that even in the simplest foraging scenario, the most likely relationship we would observe in nature is for food webs where consumers have undergone character displacement to have lower total resource densites.
+
 A general result emerging from our analysis was that, when consumers were competing for resources in distinct habitats, character displacement always supressed resource densities. 
-<!--, regardless of the functional form of the consumer's feeding rate or shape of the tradeoff.-->
 Consumers are generally larger, and as a result, more mobile than their resources; thus, their foraging behavior acts to couple resource dynamics across traditional habitat boundaries [@McCann2005].
-<!--A disproporationate number of the documented cases of character displacement involve carnivores [@Schluter2000] that are larger, and as a result, often more mobile than their resources [@McCann2005].-->
 An example of character displacement that matches this foraging scenario is threespine stickleback in postglacial lakes of British Columbia, Canada [@Schluter1992; @Schluter2000].
@@
 
-<!-- Food-web stability -->
 My most striking result was that character displacement, when examined through the lens of food-web theory, generally resulted in less stable food webs. 
 In fact, under a foraging scenario that incorporates more biological realism, character displacement may even result in an unstable food web. 
-<!--This result aligns well with the general finding from food-web theory that stability tends to go hand-in-hand with resource abundances [@May1973; @Murdoch2003; @McCann2011]. -->
 The mechanism underlying this destabilization is quite general, resulting from the strength of interspecific interactions increasing relative to intraspecific interactions [@Chesson2008; @Rip2011; @McCann2011].
@@
 This relative increase in interspecific interactions, combined with the natural oscillatory tendency of consumer-resource interactions [@Murdoch2003], creates a food-web structure that is less resilient to perturbations.
-<!-- creates top-heavy food webs (higher $(C_i+C_j)$:$(R_i+R_j)$) that are -->
 
 My results, combined with recent theory, suggest that the very conditions that favor phenotypic diversification also make that diversity more susceptible to local extinction.
-<!--Recent theory suggests that the conditions that favor ecological character displacement are the very ones that would enhance the destabilizing effects of this eco-evolutionary process.-->
 Specifically, work by @McPeek2019 showed that character displacement is favored in environments that are either highly productive, easy to find and capture resources, or under weak abiotic stress. 
@@
 For example, increasing productivity reduces intraspecific competition in resource populations while increasing the flux of energy to consumers, resulting in the "paradox of enrichment" [@Rosenzweig1971].
-<!--where higher productivity, all else equal, can decrease food-web stability due to consumers suppressing resources more below their carrying capacity. -->
 Similarly, increased capacity to find and capture resources (as we observed in our model) or lower consumer mortality, both increase the relative strength of consumer-resource interactions, which predictably destabilize consumer-resource interactions [@Rip2011; @McCann2011].
-<!--This suggests that the very conditions that favor phenotypic diversification also make that diversity more susceptible to local extinction.-->
-<!--Perhaps nonadaptive selection on ecological stability [@Borrelli2015a; @Borrelli2015b], -->
-
-<!-- Empirical support for notion of instability: Ghost of competition and stability past --->
-Empirical evidence supports the theoretical notion that food webs where consumers have undergone character displacement are less resilient to perturbations. <!--is not far-fetched.-->
-<!--Threespine stickleback in postglacial lakes of British Columbia, one of the most well-studied examples of character displacement, are known to have undergone character displacement in six different -->
-For example, <!--It is worth noting that -->a single species of threespine stickleback lives in hundreds of postglacial lakes of British Columbia, but the species pairs, where character displacement has happened, are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
+
+Empirical evidence supports the theoretical notion that food webs where consumers have undergone character displacement are less resilient to perturbations. 
+For example, a single species of threespine stickleback lives in hundreds of postglacial lakes of British Columbia, but the species pairs, where character displacement has happened, are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
 Perhaps this rarity of species pairs is a result of nonadaptive selection for more ecologically stable food-web topologies [@Borrelli2015a; @Borrelli2015b].
@@
 Perhaps stickleback in many other postglacial lakes in British Columbia retain genomic signatures of having been a habitat specialist in the past.
-<!-- could look at this in lakes that don't have a species pair, but are in nearby watersheds -->
-
-<!-- contrasts with competition theory -->
+
 My stability results contrast, but do not necessarily contradict, the notion from competition theory that character displacement contributes to species coexistence [@Lawlor1976]. 
@@
 
-<!--
-Typically, they assume that resource distributions are stable, although resource instability can enhance coexistence if consumers contain different forms of their functional responses [@Armstrong1976a; @Armstrong1976b; @Abrams2002].
-Instead, my results highlight that stability measures do not always go hand-in-hand [@Donohue2013; @Donohue2016].
-A more pluralistic view of stability is likely needed for understanding the consequences of eco-evolutionary feedbacks in a food-web context.
-
-Of the prior models of ecological character displacement, only @Lawlor1976 examine the effect of character displacement on the ecological stability of the system.
-They conducted a detailed theoretical analysis of three different metrics of stability: (i) stable coexistence of consumers, defined as the mutual invasibility of each consumer; (ii) global stability of consumers, defined as the continued coexistence of both consumers when resource dynamics are perturbed; and (iii) local stability, defined as the rate at which the four-species food web returned to equilibrium after a small perturbation to species abundances (the measure we also used).
-Note that the first two metrics correspond to the stable coexistence of consumers, while the third metric measures the stability of the entire food web.
-@Lawlor1976 found that character displacement enhanced the stable coexistence of consumers, yet had a small negative effect on the local stability of the food web.
-@Lawlor1976's work gave fundamental insight to the effect of character displacement on ecological stability; however, none of their stability analyses compared food webs with and without a competing consumer.
-Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement (@Schluter2000; @Schluter1992).
-In fact, the first two metrics of stability can only be assessed when both consumers are present because it examines the effect of divergence on coexistence, which requires both consumers to be present.
-In contrast, local stability can be compared in food webs with and without a competing consumer, and thus provides a useful benchmark for examining the effects of character displacement on food-web stability.
--->
-
-<!-- Caveats -->
 I expect that my theoretical predictions are robust to different assumptions about the underlying evolutionary dynamics of consumer foraging traits.
@@
 
-<!-- Conclusions -->
 Competition theory suggested that character displacement was a stabilizing force in ecological communities [@Lawlor1976]. 
@@
 
-<!--
-Although I used an Adaptive Dynamics approach here, I expect that these theoretical predictions would apply to models that assumed a different genetic architecture of attack rates (e.g.
-quantitative genetics).
-I expect this because, time and time again, models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of the genetic architecture of traits [see synthesis by @Taper1985].
-Since these models are focused primarily on trait evolution, it is not clear that divergence in these models would have different ecological effects.
-In fact, recent work by @McPeek2019 has shown that genetic divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
-Finally, my conclusions may only apply to biological resources that are nutritionally substitutable. 
-It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
-
-An advantage of the Adaptive Dynamics approach is that we could examine the ecological dynamics of the system.
-Current consumer-resource models based on quantitative genetics become quickly intractable to gain analytical insight from, even for simple linear functional responses of consumers [@McPeek2019].
-Those approaches have clear advantages in that they explicitly model a quantitative trait, which then influences the ecological rates (*r, K, A, e, m*, etc.).
-Our approach assumes that there is a one-to-one correspondence between trait evolution and the ecological rate.
-This simplifying assumption, however, allowed us to explore a much more realistic biological scenario and still gain analytical insight.
-This necessary tradeoff was worth it, we feel, because we were interested in the ecological consequences of character displacement, rather than in identifying the conditions resulting in character displacement and that alter the magnitude, as excellent work has already been done on this front [e.g.
-@Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019].
-
-While the underlying genetics of trait evolution may influence the amount of divergence in character displacement, this should not influence the effect of that trait evolution on the consumer-resource dynamics.
-Another way our results may quantitatively differ from a quantitative genetic model is that there is no intraspecific variation in our approach.
-While this may certainly have quantitative effects on the relationship between character displacement and food-web dynamics, or more likely, influence the magnitude of displacement [@McPeek2019], it is unclear whether this would qualitatively influence our inferences.
-
-Throughout, I have also setup a special case scenario where consumers are mirror images of each other.
-This resulted in symmetric character displacement, which is quite common in empirical data [@Schluter2000].
-Again though, I do not expect that adding mild asymmetries (ones that still would permit coexistence) would qualitatively change our results.
-This is because, again, divergent character displacement is still evident in these models.
-Also, we conducted simulations to test whether our results were robust to this assumption and they were (ADD SUPPLEMENTARY MATERIAL).
-
-<!--
-All of these result from the general principle that any biological mechanism that increases the strength of interspecific interactions relative to the strength of intraspecific interactions acts to destabilize consumer-resource dynamics [@Chesson2008; @Rip2011; @McCann2011].
-Mathematically, this happens when the equilibrium point of the system is associated with an imaginary eigenvalue.
-This causes the system to exhibit damped oscilattions creates a tendency for the system to oscillate in the presence of any lag in consumer-resource dynamics., which creates a puts some "spin" on the system.
-Otherwise, biological mechanisms that enhance interaction strength can be initially stabilizing, if the consumer is poorly adapted to foraging on the available resources (see first parts of stability figures).
-
-is the imaginary eigenvalue associated with the lag itself? A presence of a lag in the system?
--->
-
-<!-- environmental gradients and character displacement 
-It may be quite common for character displacement to result in less stable food webs. 
-Recent theory has shown that consumer divergence and specialization on resources happens more readily when either: higher productivity, low structural complexity of habitat, and weak influences of stressors [@McPeek2019].
-In our model, this corresponds to higher values of *K*, higher values of *A*, and lower values of *m*. 
-In other words, this recent theory shows that the magnitude of consumer divergence happens in highly productive habitats, where consumers can easily find and consume their prey, and there is low consumer mortality.
-The first situation can lead to the "paradox of enrichment" [@Rosenzweig1971], where higher productivity, all else equal, can decrease food-web stability due to consumers suppressing resources more below their carrying capacity.
-The other factors, maximum attack rate and consumer mortality, also have predictable effects on the stability of the system [@Rip2011; @McCann2011]. 
-For example, increases in attack rate, as we observed in our model, decreases food-web stability. 
-Similarly, lower mortality increases enables consumers to have a higher impact on their resources. 
-All of these result from the general tendency of for any biological mechanism that increases the strength of interspecific interactions relative to the strength of intraspecific interactions acts to destabilize consumer-resource dynamics [@Chesson2008; @Rip2011; @McCann2011].
-Mathematically, this happens when their is an imaginary eigenvalue of the system, which puts some "spin" on the system.
-Otherwise, biological mechanisms that enhance interaction strength can be initially stabilizing, if the consumer is poorly adapted to foraging on the available resources (see first parts of stability figures).
--->
-
-<!-- But talk about the competition inference, and how it may not allow a proper view of how ecological character displacement affects stability, because it also does not compare the scenarios.
-For example, convergence is clearly favored in the absence of a competitor.
-It also doesn't account for the stability of the fully coupled food web that includes resources, which we know is a key contribute to the dynamics of ecological character displacement -->
-
-<!-- Environmental gradients 
-My theoretical results also give insight to where we might expect to observe the destabilizing effects of character displacement in nature. 
-Specifically, environments that increases the flux of energy to consumers relative to resources generally decrease stability [@Rip2011; @McCann2011], which would amplify the destabilizing effects of character displacement.
-For example, aquatic food webs have a number of attributes that make them less stable than terrestrial food webs [@Rip2011].
--->
-<!-- excerpts from Rip and McCann 2011
-aquatic primary producers possess less structural material
-(Elser et al. 2000) and contain more photosynthetic tissue (Elser et al.
-2000). These differences result in higher growth rates in aquatic
-primary producers (Cebrian 1999; Elser et al. 2000).
-
-Many differences in consumption rates exist across ecosystem types.
-These differences arise from a number of attributes differing across
-ecosystems. In aquatic ecosystems, the body size ratios of herbivores
-to primary producers are larger (Hairston & Hairston 1993; Chase
-2000) and primary producers are more nutritious to herbivores (Polis
-& Strong 1996; Cebrian 1999; Elser et al. 2000). This results in greater
-herbivory in aquatic ecosystems (Cyr & Pace 1993; Hairston &
-Hairston 1993; Cebrian 1999). The higher consumption rates in
-aquatic ecosystems should have a large impact on parameters used in
-C-R theory. Higher consumption rates may manifest in higher
-maximum consumption rates and higher conversion efficiencies (i.e. a
-or aMax and e). As seen above, an increase in either or both of these
-parameters would result in an increase in energy flow, more inverted
-biomass pyramids, and a decrease in stability. Therefore we predict
-that aquatic ecosystems should be less stable than terrestrial
-ecosystems.
-
-. However, the amount of mass-specific structural material
-(e.g. skeletal material, woody tissue) does positively scale to body
-mass: scaling exponents vary between 0.4 and 0.12 suggesting that
-efficiency should scale across ecosystem types (Schmidt-Nielson
-1984). As conversion efficiency is dimensionless, these positive
-exponents become negative when we divide through by biomass to
-remove the dimensions (exponents become )0.88 to )0.40). Thus, as
-others have argued (Strong 1992, Polis et al. 2000), aquatic pathways
-ought to tend to have far less structural material and promote high
-conversion efficiencies.
-
-In summary, the relationships suggest that the isocline should shift
-left relative to the resource isocline with reduced body size. Moreover,
-these arguments resonate with previous arguments. Strong (1992) and
-Polis et al. (2000) argued cogently that aquatic systems were more
-homogenised reducing prey refugia, more edible and tended to
-experience higher turnover rates than terrestrial ecosystems. All these
-effectively argue for higher energy flux through aquatic pathways as
-our allometric reasoning also suggests. Thus, differences in the
-attributes of aquatic and terrestrial ecosystems relate to differences in
-energy flow within ecosystems. Comparing cross-ecosystem differences
-to parameters used in C-R theory, we predict aquatic ecosystems
-should have high growth rates of primary producers and higher
-numerical responses by consumers. All of these predictions taken
-together suggest that energy flow in aquatic ecosystems should be
-higher, biomass pyramids more inverted and stability decreased
-relative to terrestrial ecosystems.
-
-Strong, D.R. (1992). Are trophic cascades all wet? Differentiation and donorcontrol
-in speciose ecosystems. Ecology. 73:747–754.
-
-Polis, D.A., Sear, A., Huxel, G.R., Strong, D.R. & Maron, J. (2000). When is a
-trophic cascade a trophic cascade? Trends Ecol. Evol., 11, 473–475.
--->
-
-<!-- Predictions for a natural system 
-My results give insight to where we would expect character displacement to result in less stable food webs in nature: 
-(i) consumers compete for resources in distinct habitats; (ii) exhibit concave-up tradeoff in attack rates; (iii) high productivity (*K*); (iv) low structural complexity (*a_ii_*,*a_ij*); (v) weak stressors (low *m*).
-
-Aquatic ecosystems have a number of attributes that make them generally less stable than terrestrial ecosystems [@Rip2011]. For example,
-
-Aquatic systems are more likely because its been shown that attack rates are often much stronger in these systems [@Rip2011]. 
-This lower baseline, means that it is easier for character displacement to push these systems toward the boundary of local stability. 
-Interestingly, one of the hallmark examples of character displacement, threespine stickleback, meets these criteria.
-Thus, I predict that in the coastal lakes of British Columbia, where stickleback have undergone character displacement, for these food webs to exhibit less stable dynamics and more variable population abundances.
-Although stability is a traditionally thorny issue to measure, there are now clear theoretical links between the stability measure we used here ($-\lambda_{\text{max}}$) and the coefficient of variation (CV), which simply measures the variability in a dynamic time series relative to the mean.
-Also, there are now methods for assessing the dynamics of natural communities that link to these theoretical stability measures [@Ives2003]. Now it just requires the effort to collect this data. 
-Also, in applying this theory to a natural system, an assessment of stability for a specific system requires a realistic model of the population and community dynamics.
-e.g.
-stage-structure,...Still, our results point toward a general tendency toward instability.
-
-It's entirely plausible though that the some aspects of the system, e.g.
-stage structure could prevent this instability.
-This could be quite interesting, as this may point to systems where we are more likely to observe character displacement, because it enhances their stability.
--->
-
-<!-- older cut and pasted
-
-I find that the foraging scenario qualitatively affects the relationship between character displacement and resource densities.
-
-Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously [e.g.
-@Taper1985; @Abrams1986].
-This assumption may be valid for some foraging scenarios (e.g.
-Darwin's finches foraging for seeds); however, in many situations consumers forage for resources that occur in different habitats, and thus cannot forage for both simultaneously [@McCann2005].
-
-The effect of the foraging scenario on the relationship between character displacement and resource densities appears to be quite general.
-
-Note that detecting this effect in nature would likely be difficult, since a concave-down tradeoff results in relatively small displacement relative to other tradeoff shapes (@Lawlor1976; green points in Fig.
-\ref{fig:MacArthur_Resources}A,B).
-
-Previous work has shown that consumers undergo divergent character displacement regardless of the shape of the tradeoff (@Lawlor1976; @Abrams1986).
-
-Important point: Criteria 5 in @Schluter1992 does not make sense since character displacement usually alters resource distributions. Think about this more.
-
-<!-- This more realistic consumer functional response does actually make an interesting evolutionary prediction that differs from the scenario where consumer feeding rate scales linearly with resource density.
-Specifically, a consumer's ESS in the absence of competition will remain slightly more specialized on the resources that occur in their preferred habitats.
-This is in contrast to the prediction when feeding rates scale linearly with resource density, where consumers evolve to be slightly more specialized on resources occurring in their non-preferred habitats.
-An important caveat for these predictions is that my model does not allow habitat preferences to evolve, which would likely alter these predictions.
-
-@Lawlor1976 found that consumers still underwent divergent character displacement regardless of whether consumers could or could not forage for resources simultaneously.
-
-In general, I find that resource supression goes hand-in-hand with local stability, a tendency that has been noted by others (@Murdoch2003; @McCann2011).
-
-<!-- Looking for this effect in nature
-
-The scenario where the relationship between character displacement and resource suppression is the strongest is when the tradeoff in attack rates is concave-up and consumers are competing for resources that occur in different habitats (orange points in Fig.
-\ref{fig:LS_Resources}B and Fig.
-\ref{fig:McCann_Resources}B).
-Prior work in threespine stickleback suggests that the evolution of stickleback attack rates are constrained by a concave-up tradeoff (@Schluter1993; @Arnegard2014) and that stickleback ecotypes are competing for resources that occur in different habitats (benthic vs.
-limnetic, @Schluter1992).
-Thus, I predict that ecological character displacement in sticklebacks would suppress resource densities and decrease local stability.
-
-While measuring resource densities is relatively straightforward, measuring stability is notoriously difficult.
-One of the more easy to acquire empirical metrics of stability is the coefficient of variation ($CV=\frac{\sigma}{\mu}$) as this only requires measurements of resource (or consumer) densities over time.
-Interestingly, there is a close correspondence between the coefficient of variation and the local stability of a food web (@Gellner2016).
-<!-- This close correspondence emerges from the fact that stochastic perturbations in resource (or consumer) densities will determine the variability in abundances.
-And since food webs are constantly buffeted by stochastic processes in the real world, I expect that my predictions using local stability would correspond to the $CV$ of resources (or consumers).
-Thus, I predict that character displacement in threespine stickleback will increase the $CV$ of resource (or consumer) densities.
-
-Consumer and resource densities in these dynamical models can be interpreted from either a population or biomass perspective (@Yodzis1992; @Murdoch2003; @McCann2011).
-For the stickleback system, you have data on seasonal dynamics of resources in terms of both abundances and biomass.
-Within a season, I do not expect stickleback abundances to be coupled to the abundances of zooplankton and benthic invertebrates, given that stickleback reproduce annually whereas zooplankton and benthic invertebrates can reproduce many times within a season.
-Instead, I expect that stickleback biomass will be coupled to the biomass of zooplankton and benthic invertebrates within a season.
-This is because there will be lots of young stickleback at the beginning of the season, and they will eat a lot of resources and grow in size as the season progresses.
-If this is true, I also expect to see the largest negative effect of stickleback on resource densities late in the season.
--->
-
 # Acknowledgements
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
   geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
   geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
   geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
   geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
   geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + 
   geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
+  ylab(expression("Stability "(-lambda[max]))) + 
   scale_shape_manual(values=c(24,21))  +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
   geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
+  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
   geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +

### 2019-03-07 manuscript_ECD_model.Rmd
edit title page and abstract

@@
 \vspace*{0.1cm} 
-\begin{center} \LARGE Food-web dynamics under ecological character displacement \end{center}
+\begin{center} \LARGE Ecological character displacement destabilizes food webs \end{center}
+<!--\begin{center} \LARGE Food-web dynamics under ecological character displacement \end{center}-->
 
@@
 
-\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;
+\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, 8057 Zurich, Switzerland;
 
@@
 
-*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptive radiation; ecological community
+*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptation; ecological community; stability
 
@@
 
-*Manuscript type*: Note.
+*Manuscript type*: Note. <!-- consider e-Article, depending on length -->
 
@@
 Ecological character displacement is an adaptive process, often enhancing coexistence between consumers by reducing competition for shared resources. 
-This focus on competition has undermined the fact that consumer-resource interactions ultimately mediate the process of character displacement. 
-As a consequence, the effects of character displacement on consumer-resource interactions and food-web dynamics have not been critically examined.
-Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumers influences resource densities and the resilience of food webs to perturbations.
-To test the generality of these effects, I examined them under different assumptions of evolutionary tradeoffs, consumer foraging behavior, and resource distributions.
-My general finding was that, whenever character displacement increased the strength of consumer-resource interactions, it resulted in lower resource densities and less resilient food webs. 
-Character displacement always strengthened consumer-resource interactions whenever consumers were competing for resources that occured in distinct habitats. 
+This competition perspective undermine the fact that consumer-resource interactions ultimately mediate the process of character displacement. 
+As a consequence, the effects of character displacement on consumer-resource interactions and food-web dynamics have been ignored.
+Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumers affects resource densities and the resilience of food webs to perturbations.
+I studied these effects under different assumptions of evolutionary tradeoffs, consumer foraging behavior, and resource distributions.
+I found that character displacement always strengthened consumer-resource interactions whenever consumers competed for resources that occured in distinct habitats.
+This increase in interaction strength resulted in lower resource abundances and less resilient food webs.
 This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
 While character displacement did not always increase the strength of consumer-resource interactions, these contingent effects only occurred under the simplest and least realistic foraging scenario. 
-Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations. 
-They also suggest that food-web and competition theory can give complimentary insights to the ecological consequences of adaptation.
+Taken together, my results show that the adaptive process of character displacement often comes at the ecological cost of decreasing food-web resilience. 
+They also highlight that food-web and competition theory can give complimentary insights to the ecological consequences of adaptation.
 

### 2019-03-07 manuscript_ECD_model.Rmd
add edits to Methods

@@
 
-Ecological character displacement is thought to be a key evolutionary process generating biodiversity [@Schluter2000; @Pfennig2010; but see @Stuart2013].
-Ecological character displacement describes the “process of phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000].
+Ecological character displacement is an important adaptive process in generating biodiversity [@Schluter2000; @Pfennig2010; but see @Stuart2013]. 
+This process results from "phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000]. 
 Over the past 40 years, a large body of theoretical [e.g.
-@Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has been generated to understand the scenarios under which exploitative competition for resources leads to the divergence of consumer foraging traits.
-The general result emerging from these studies is that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the divergence of consumer foraging traits [@Lawlor1976; @Taper1985].
-This process is not simply driven by ecological differences, but creates an eco-evolutionary feedback that drives further differentiation.
-This key insight was made by theoretical models that explicitly included resource dynamics as a mediator of competition in driving evolutionary change [@Lawlor1976; @Abrams1986; @Taper1985].
-
-Although explicit models of resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics have received surprisingly little attention. 
-This is likely because the ecological feedback has been primarily studied through the lens of competition and coexistence theory [@Lawlor1976; @Germain2018; @Bassar2017; @McPeek2019]. 
+@Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has examined which scenarios lead to phenotypic divergence or convergence of competing consumers.
+The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985].
+This adaptive process is not simply a response to statitc differences in resource distributions, but creates an eco-evolutionary feedback that drives further differentiation. 
+This crucial insight was made by theoretical models that explicitly included resource dynamics as a mediator of competition in driving evolutionary change [@Lawlor1976; @Abrams1986; @Taper1985].
+
+Although models that included resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics has received surprisingly little attention. 
+This is likely because the ecological feedback has been primarily studied through the lens of competition theory [@Lawlor1976; @Germain2018; @Bassar2017; @McPeek2019]. 
 For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition.
-Yet, this reduction in interspecific competition may, at the same time, increase the interspecific interaction between a consumer and the resource it specializes on.  
-Both food-web theory and empirical studies have shown that increasing the strength of consumer-resource interactions often leads to resource suppression, which if sufficient enough, can generate oscillations, and hence less stable consumer-resource dynamics [@Rosenzweig1971; @Luckinbill1973; @Murdoch2002; @Murdoch2003; @McCann2011].
+Yet, this reduction in interspecific competition may, at the same time, increase interspecific interactions between specialized consumers and their resources.  
+Both food-web theory and empirical studies have shown that increasing the strength of consumer-resource interactions often suppresses the abundance of resources, which if sufficient enough, can generate oscillations and less stable consumer-resource dynamics [@Rosenzweig1971; @Luckinbill1973; @Murdoch2002; @Murdoch2003; @McCann2011].
 Thus, a food-web perspective, which accounts for both the direct and indirect effects of consumer-resource interactions, may yield new insight to the ecological consequences of character displacement.
 
-Here, I address this knowledge gap by building a mathematical model to examine how ecological character displacement affects consumer-resource dynamics in a food-web context.
+Here, I address this knowledge gap by studying a mathematical model that examines how ecological character displacement affects consumer-resource dynamics in a food-web context.
 I address two questions: (1) How does ecological character displacement affect resource abundances?
 (2) How does character displacement affect food-web stability?
-To test the generality of these ecological effects, I examined three different foraging scenarios: *CONDENSE THIS.
-NO NEED TO GIVE ALL THE DETAILS YET.
-RIGHT NOW IT IS JUST CONFUSING.* (i) resources occur in the same habitat and consumers exhibit a linear functional response [e.g.
-@MacArthur1972; @Lawlor1976]; (ii) resources occur in different habitats and consumers exhibit a linear functional response [e.g.
-@Lawlor1976]; (iii) resources occur in distinct habitats and consumers exhibit a more realistic functional response [@McCann2005].
-I also explore the contingency of these effects to different assumptions about evolutionary trade-offs in consumer foraging traits.
-I focus on the ecological side of these dynamics because the evolutionary aspects have been well characterized [@Lawlor1976; @Taper1985; @Abrams1986; @McPeek2019].
-I show that the consequences of character displacement on food-web dynamics are not always intuitive and depend on the foraging scenario and tradeoff in consumer traits.
-
-# Methods and Results
-
-## Underlying Food-web Dynamics
+To test the generality of these effects, I explored different assumptions of ecological foraging scenarios and evolutionary trade-offs in consumer foraging traits.
+
+# Methods
 
@@
 I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
-I focus here on the ecological consequences of this adaptive divergence.
-
-To investigate the ecological consequences, I studied the effect of adaptive divergence in consumers on resource densities and the stability of the entire food web at equilibrium.
+
+To investigate the ecological consequences of divergent character displacement, I analyzed differences in resource abundances and food-web stability at equilibrium. 
 An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium densities for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$).
 At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource densities and performed a linear stability analysis [@Otto2007]. 
-This stability analysis derives the dominant (largest in absolute value) eigenvalue, $\lambda$, which determines whether (and how readily) the food web, when perturbed a small amount, will return to equilibrium (see supplementary material for derivation).
-Together, these two measures indicate how ecological character displacement affects resource densities and food-web stability.
-
-I examine the effects of character displacement on food-web dynamics by (i) deriving analytical expressions for the relationship between consumer attack rates and food-web dynamics and (ii) simulating the effects of competition on the eco-evolutionary dynamics of consumers and resources.
-To gain analytical insight, I assume that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
-While this scenario is arguably simplistic, it allows me to isolate the effects of ecological character displacement.
-
-To simulate eco-evolutionary dynamics, I used an Adaptive Dynamics approach.
+This stability analysis derives the dominant eigenvalue, $\lambda$, which determines whether (and how readily) the food web will return to equilibrium after a small perturbation.
+
+When possible, I derived analytical expressions for the relationship between consumer attack rates and food-web dynamics. 
+To do this, I simplified the model by assuming that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
+Controlling for other sources of variability allowed me to isolate the general effects of divergent character displacement on food-web dynamics.
+
+Although divergent character displacement occurs in all scenarios we examined, the magnitude of this divergence may be context dependent.
+Therefore, I simulated the eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
 At each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate on one resource by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability.
@@
 \ref{fig:MacArthur_Plot}A).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig.
-\ref{fig:MacArthur_Plot}A).
+When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig. \ref{fig:MacArthur_Plot}A). 
 Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 

### 2019-03-07 manuscript_ECD_model.Rmd
add edits to results

@@
 As a consequence, the effects of character displacement on consumer-resource interactions and food-web dynamics have been ignored.
-Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumers affects resource densities and the resilience of food webs to perturbations.
+Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumers affects resource abundances and the resilience of food webs to perturbations.
 I studied these effects under different assumptions of evolutionary tradeoffs, consumer foraging behavior, and resource distributions.
@@
 @Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has examined which scenarios lead to phenotypic divergence or convergence of competing consumers.
-The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985].
+The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of abundance dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985].
 This adaptive process is not simply a response to statitc differences in resource distributions, but creates an eco-evolutionary feedback that drives further differentiation. 
@@
 seed size, see @Taper1985 for an example).
-In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource density, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972].
+In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972].
 The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
@@
 To investigate the ecological consequences of divergent character displacement, I analyzed differences in resource abundances and food-web stability at equilibrium. 
-An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium densities for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$).
-At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource densities and performed a linear stability analysis [@Otto2007]. 
+An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$).
+At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource abundances and performed a linear stability analysis [@Otto2007]. 
 This stability analysis derives the dominant eigenvalue, $\lambda$, which determines whether (and how readily) the food web will return to equilibrium after a small perturbation.
@@
 
+# Results
+
 ## Consumers can forage for both resources simultaneously
 
-When both consumers and resources are present, the density of resources at equilibrium are equivalent ($\hat R = \hat R_1 = \hat R_2$) and are determined by the following equation (derivation in supplementary material):
+<!-- I haven't made clear that this explores a specific scenario, MacArthur's model -->
+
+When both consumers and resources are present, the abundance of resources at equilibrium are equivalent ($\hat R = \hat R_1 = \hat R_2$) and are determined by the following equation (derivation in supplementary material):
 
 $$\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
-A key determinant of resource density in this model is the consumer's total attack rate, $a_{ii}+a_{ij}$.
+A key determinant of resource abundance in this model is the consumer's total attack rate, $a_{ii}+a_{ij}$.
 Therefore, the ecological consequences of character displacement depend on how the tradeoff influences the evolution of consumer attack rates.
 
-My simulations show that the shape of the attack-rate tradeoff qualitatively affects the relationship between character displacement and resource density (Fig.
-\ref{fig:MacArthur_Plot}).
+My simulations show that the shape of the attack-rate tradeoff qualitatively affects the relationship between character displacement and resource abundance (Fig. \ref{fig:MacArthur_Plot}). 
 For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (Fig.
-\ref{fig:MacArthur_Plot}B) and character displacement has no effect on resource densities (Fig.
+\ref{fig:MacArthur_Plot}B) and character displacement has no effect on resource abundances (Fig.
 \ref{fig:MacArthur_Plot}C).
-If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (Fig.
-\ref{fig:MacArthur_Plot}C).
-This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (Fig.
-\ref{fig:MacArthur_Plot}B).
-When the tradeoff is concave up (orange lines), character displacement suppresses resource densities due to the increase in total attack rates (Fig.
-\ref{fig:MacArthur_Plot}B,C).
-Although the equation I derived for resource density was for the scenario where both consumers and both resources are present, it accurately predicts the density of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
+If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (Fig. \ref{fig:MacArthur_Plot}C). 
+This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (Fig. \ref{fig:MacArthur_Plot}B). 
+When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (Fig. \ref{fig:MacArthur_Plot}B,C). 
+Although the equation I derived for resource abundances was for the scenario where both consumers and both resources were present, it accurately predicts the abundance of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
 \ref{fig:MacArthur_Plot}C).
 This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in Fig.
-\ref{fig:MacArthur_Plot}B), resulting in equivalent resource densities.
-
-The effect of character displacement on resource density closely matches its impact on food-web stability (Fig.
-\ref{fig:MacArthur_Plot}D).
-For example, when character displacement results in resource suppression (orange), there is a corresponding decrease in stability.
-Similarly, if character displacement does not influence resource densities (blue), then there is no corresponding effect on stability.
-Curiously, this match between impact on resources and stability breakdowns with the concave down tradeoff (green).
-
-```{r MacArthur_Resources, include=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers can forage for both resources simultaneously."}
+\ref{fig:MacArthur_Plot}B), resulting in equivalent resource abundances.
+
+The effect of character displacement on resources corresponds to its impact on food-web stability. 
+For example, when character displacement decreases resource abundances (orange points in \ref{fig:MacArthur_Plot}C), there is a decrease in food-web stability (\ref{fig:MacArthur_Plot}D). 
+Similarly, if character displacement does not alter resource abundances (blue and green points in \ref{fig:MacArthur_Plot}C), then there is no change in food-web stability (\ref{fig:MacArthur_Plot}D).
+
+```{r MacArthur_Resources, include=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers can forage for both resources simultaneously."}
 Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
@@
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation.
-Panels (C) and (D) show how different tradeoffs indirectly affect total resource densities (C) and food-web stability (D).
+Panels (C) and (D) show how different tradeoffs indirectly affect total resource abundances (C) and food-web stability (D).
 These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
@@
 
-The previous model, although the starting point for many models of competition, does not reflect many food webs where consumers are more mobile than their resources.
-In this scenario, consumer foraging behavior can link resources that occur in distinct habitats.
-The only character displacement model that I am aware of that modeled resources in different habitats was one examined by @Lawlor1976.
-This model takes the same form as the previous model, except now the consumer's feeding rate takes the form:
+The previous model, although the starting point for many models of competition, does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in distinct habitats [@McCann2005].
+The only character displacement model that I am aware of that examined the effect of resources in different habitats was studied by @Lawlor1976.
+This model is the same as before, except now the consumer's feeding rate takes the form:
 
@@
 To examine the ecological consequences of this divergence, I again derive the equilibrium solution when both consumers and both resources are present.
-As before, resource densities are equivalent at this equilibrium ($\hat R = \hat R_1 = \hat R_2$ when both consumers and resources present), but are now determined by the following equation (derivation in supplementary material):
+As before, resource abundances are equivalent at this equilibrium ($\hat R = \hat R_1 = \hat R_2$ when both consumers and resources present), but are now determined by the following equation (derivation in supplementary material):
 
@@
 \ref{fig:LS_Plot}A).
-Thus, character displacement always results in resource suppression if consumers are competing for resources that occur in different habitats (Fig.
-\ref{fig:LS_Plot}B).
+Thus, divergent character displacement always results in resource suppression (Fig. \ref{fig:LS_Plot}B). 
 Note that the shape of the tradeoff can modify the effect of character displacement. 
@@
 \ref{fig:LS_Plot}B).
-In contrast, resource densities reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
+In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
 \ref{fig:LS_Plot}B), because character displacement tends to reach a constraint of complete specialization.
-It is worth noting that in this foraging scenario, resource densities are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
+It is worth noting that resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
 \ref{fig:LS_Plot}B).
-This is likely because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in Fig.
+This is because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in Fig.
 \ref{fig:LS_Plot}B).
 
-As we saw previously, the effect of character displacement on resource densities qualitatively matches its negative effect on food-web stability (Fig.
-\ref{fig:LS_Plot}C).
+As we saw previously, the effect of character displacement on resource abundances qualitatively corresponds to its effect on food-web stability (Fig. \ref{fig:LS_Plot}C). 
+Specifically, divergent character displacement decreases food-web stability, regardless of the tradeoff in attack rates.
 This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (Fig.
 \ref{fig:LS_Plot}C).
-For example, when the tradeoff is concave up (orange), then the four-species (small circle) food web actually starts off as less stable than the three-species one (small triangle); however, the eco-evolutionary dynamics cause a switch in this pattern once they reach their evolutionary stable strategies (large points).
-
-```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource densities (B) when consumers cannot forage for resources simultaneously."}
+For example, when the tradeoff is concave up (orange), then the four-species (small circle) food web actually starts off as more stable than the three-species one (small triangle); however, this pattern switches once they reach their evolutionary stable strategies (large points).
+
+```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource abundances (B) when consumers cannot forage for resources simultaneously."}
 LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
@@
 Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
-This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
+This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
 These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6; *w~ij~*=0.4.
@@
 
-So far I have only considered a simple functional response, where consumer feeding rates increase linearly with resource densities when foraging in either the same or distinct habitats.
-We know, however, that consumer feeding rates often saturate at high resource densities [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011].
-We also know that consumers do not usually spend a fixed proportion of time in a particular habitat, but their habitat preferences may be modified by resource densities [@McCann2005].
+So far I have only considered a simple functional response, where consumer feeding rates increase linearly with resource abundances when foraging in either the same or distinct habitats. 
+We know, however, that consumer feeding rates often saturate at high resource abundances [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011]. 
+We also know that consumers do not usually spend a fixed proportion of time in a particular habitat, but their habitat preferences may be modified by resource abundances [@McCann2005]. 
 To account for these biological realities, I consider a more realistic functional response [derived by @McCann2005]:
@@
 
-where consumer feeding rates on resource $i$ are influenced by both resource densities; saturate as resource densities increase (due to handling time $h$); and consumer habitat preferences are modified by relative resource densities, such that:
+where consumer $i$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling time $h$); and consumer habitat preferences are modified by the relative abundance of resources, such that:
 
 $$W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}$$
+
+<!-- Need to work on the flow here and mention the evolutionary results. Perhaps it would be best to parallel my writing style as in the previous sections. -->
 
@@
 \ref{fig:McCann_Plot} in Appendix).
-This is because resource competition always leads to character divergence and equilibrium resource densities are governed by similar dynamics  (derivation in supplementary material):
+This is because resource competition always leads to character divergence and equilibrium resource abundances are governed by similar dynamics  (derivation in supplementary material):
 
@@
 
-```{r McCann_Resources, include=FALSE, fig.cap="\\label{fig:McCann_Resources}Effect of character displacement on total attack rates (A) and resource densities (B) when consumers exhibit a more realistic functional response."}
+```{r McCann_Resources, include=FALSE, fig.cap="\\label{fig:McCann_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers exhibit a more realistic functional response."}
 McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
@@
 Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
-This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
+This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
 These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.
@@
 
-The previous analytical results and simulations make a strong assumption that competing consumers start off as perfect mirror images of each other in their attack rates (i.e.
-symmetry).
-Yet, the theory presented here and in past work indicates that there would likely be a predictable asymmetry between consumers.
-This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, reaches an ESS (as a generalist), followed by the invasion of a second, more specialized, consumer.
-This theoretical scenario has also been hypothesized as the sequence of events leading to character displacement in threespine stickleback inhabiting lakes of coastal British Columbia, Canada [@Schluter1992; @Schluter2000].
-
-To test whether our results were robust to this likely asymmetry, we used the evolved attack rates at the end of the one consumer simulations as the starting values for the competitor in the two consumer simulations.
+The previous analytical results and simulations make a strong assumption that competing consumers start off as perfect mirror images of each other in their attack rates (i.e. symmetry). 
+Yet, the theory presented here and in past work indicates a predictable asymmetry between initial consumer attack rates. 
+This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, reaches an ESS with a generalist phenotype, followed by the invasion of a second, more specialized, consumer.
+This theoretical scenario has been hypothesized as the sequence of events leading to character displacement in threespine stickleback in postglacial lakes of British Columbia, Canada [@Schluter1992; @Schluter2000].
+
+To test whether our results were robust to this asymmetry, we used the evolved attack rates at the end of the single consumer simulations as the starting values for the competitor in the two consumer simulations.
 We did this for all foraging scenarios and tradeoffs previously examined. 
-We found that our previous inferences were robust to including consumer asymmetry across different tradeoffs and foraging scenarios (\ref{fig:MacArthur_Plot_asymm}, \ref{fig:LS_Plot_asymm}, \ref{fig:McCann_Plot_asymm}).
+We found that our previous inferences were robust to including consumer asymmetry across different tradeoffs and foraging scenarios (Appendix A\ref{fig:MacArthur_Plot_asymm}, A\ref{fig:LS_Plot_asymm}, A\ref{fig:McCann_Plot_asymm}).
 
@@
 In contrast, my results show that character displacement leads to predictable differences in food (resource) availability and even food-web stability.
-Specifically, I found that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource densities and less resilient food webs. 
+Specifically, I found that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource abundances and less resilient food webs. 
 Character displacement always increased attack rates whenever consumers were competing for resources that occured in distinct habitats. 
@@
 The simplest foraging scenario examined here (linear functional response on resources occuring in the same habitat) is the foundation of mechanistic models of resource competition [@MacArthur1972; @Lawlor1976; @Abrams1986].
-Under this scenario, it is theoretically possible that character displacement results in increased resource densities [green lines in \ref{fig:MacArthur_Plot}].
+Under this scenario, it is theoretically possible that character displacement results in increased resource abundances [green lines in \ref{fig:MacArthur_Plot}].
 This only occurs, however, under an evolutionary tradeoff that constrains the magnitude of character displacement.
-This suggests that it would be unlikely to ever observe divergent character displacement and increases in resource densities in a real system. 
-In contrast, the evolutionary tradeoff that favors character displacement the most decreased resource densities. 
+This suggests that it would be unlikely to ever observe divergent character displacement and increases in resource abundances in a real system. 
+In contrast, the evolutionary tradeoff that favors character displacement the most decreased resource abundances. 
 Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the foraging rate of consumers (orange lines in \ref{fig:MacArthur_Plot}A,B).
@@
 
-A general result emerging from our analysis was that, when consumers were competing for resources in distinct habitats, character displacement always supressed resource densities. 
+A general result emerging from our analysis was that, when consumers were competing for resources in distinct habitats, character displacement always supressed resource abundances. 
 Consumers are generally larger, and as a result, more mobile than their resources; thus, their foraging behavior acts to couple resource dynamics across traditional habitat boundaries [@McCann2005].
@@
 These resources occur in distinct zones of the lake (pelagic and littoral) and stickleback move between these zones when foraging for resources. 
-The theory developed here predicts that total resource densities will be lower in lakes where stickleback have undergone character displacement compared to nearby lakes with only a single, generalist species.
+The theory developed here predicts that total resource abundances will be lower in lakes where stickleback have undergone character displacement compared to nearby lakes with only a single, generalist species.
 A disproporationate number of the documented cases of character displacement involve carnivores that are larger, and likely more mobile, than their resources [@Schluter2000],
@@
 Character displacement is far from an all-or-nothing process, but may vary in its magnitude [@Schluter2000]. 
-My results suggest that the magnitude of character displacement will be negatively correlated with total resource densities, at least when resources occur in distinct habitats (line trajectories in \ref{fig:LS_Plot}B, \ref{fig:McCann_Plot}B).
+My results suggest that the magnitude of character displacement will be negatively correlated with total resource abundances, at least when resources occur in distinct habitats (line trajectories in \ref{fig:LS_Plot}B, \ref{fig:McCann_Plot}B).
 This prediction could be test either experimentally, by putting populations that vary in divergence in a common resource setting, or with field data that statistically controls for other factors affecting resources.
@@
 Panel (A) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
-Panels (B) and (C) show how different tradeoffs indirectly affect total resource densities (B) and food-web stability (C).
+Panels (B) and (C) show how different tradeoffs indirectly affect total resource abundances (B) and food-web stability (C).
 These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
@@
 Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of consumers.
-This always resulted in a suppression of resource densities (B) and concomitant decrease in food-web stability (C).
+This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
 These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.

### 2019-03-07 manuscript_ECD_model.Rmd
add edits to Discussion and Acknowledgements.

@@
 
-My results suggest that a revision is needed for one of the criteria used to demonstrate character displacement. 
-Criteria #5 in @Schluter1992 states: "Sites of sympatry and allopatry should not differ greatly in food, ...". 
-In contrast, my results show that character displacement leads to predictable differences in food (resource) availability and even food-web stability.
+My results show that character displacement leads to predictable differences in food (resource) availability and even food-web stability.
 Specifically, I found that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource abundances and less resilient food webs. 
@@
 Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations.
+
+The ecological and evolutionary scenarios that generated the greatest divergent character displacement always resulted in decreases in resource densities.
+<!-- build paragraph around this topic sentence -->
 
@@
 This is evident by the disruptive selection observed for beak size/shape in Darwin's finches [@Schluter1985] and benthic/limnetic ecotypes for stickleback [@Arnegard2014] in resource-limited environments.
+<!-- key point of the paragraph -->
 Taken together, my model predicts that even in the simplest foraging scenario, the most likely relationship we would observe in nature is for food webs where consumers have undergone character displacement to have lower total resource densites.
 
-A general result emerging from our analysis was that, when consumers were competing for resources in distinct habitats, character displacement always supressed resource abundances. 
+A general result emerging from my analysis was that, when consumers were competing for resources in distinct habitats, character displacement always supressed resource abundances. 
+<!-- show that this is common scenario -->
 Consumers are generally larger, and as a result, more mobile than their resources; thus, their foraging behavior acts to couple resource dynamics across traditional habitat boundaries [@McCann2005].
@@
 These resources occur in distinct zones of the lake (pelagic and littoral) and stickleback move between these zones when foraging for resources. 
-The theory developed here predicts that total resource abundances will be lower in lakes where stickleback have undergone character displacement compared to nearby lakes with only a single, generalist species.
+<!-- good closing sentences -->
 A disproporationate number of the documented cases of character displacement involve carnivores that are larger, and likely more mobile, than their resources [@Schluter2000],
@@
 
+My results suggest that a revision is needed for one of the criteria used to demonstrate character displacement. 
+Criteria #5 in @Schluter1992 states: "Sites of sympatry and allopatry should not differ greatly in food, ...".
+
+<!-- consider removing entirely. This is a prediction I could test with Seth in another paper...
 Character displacement is far from an all-or-nothing process, but may vary in its magnitude [@Schluter2000]. 
 My results suggest that the magnitude of character displacement will be negatively correlated with total resource abundances, at least when resources occur in distinct habitats (line trajectories in \ref{fig:LS_Plot}B, \ref{fig:McCann_Plot}B).
-This prediction could be test either experimentally, by putting populations that vary in divergence in a common resource setting, or with field data that statistically controls for other factors affecting resources.
+This prediction could be tested either experimentally, by putting populations that vary in divergence in a common resource setting, or with field data that statistically controls for other factors affecting resources.
 Note though, this prediction will only apply to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
-
-My most striking result was that character displacement, when examined through the lens of food-web theory, generally resulted in less stable food webs. 
+-->
+
+My most striking result was that divergent character displacement, when examined through the lens of food-web theory, generally destabilized food-web dynamics. 
 In fact, under a foraging scenario that incorporates more biological realism, character displacement may even result in an unstable food web. 
@@
 
+<!-- good, but tighten up -->
 My results, combined with recent theory, suggest that the very conditions that favor phenotypic diversification also make that diversity more susceptible to local extinction.
 Specifically, work by @McPeek2019 showed that character displacement is favored in environments that are either highly productive, easy to find and capture resources, or under weak abiotic stress. 
-In my model, this corresponds to higher values of $K$ (productivity) or $a_ij$s (finding and capturing resources), or lower values of $m$ (abiotic stress).
+In my model, this corresponds to higher values of $K$ (productivity) or $a_{ij}$ (finding and capturing resources), or lower values of $m$ (abiotic stress).
 Each of these corresponding changes create food webs that have a lower baseline of stability, as they increase the strength of consumer-resource interactions relative to intraspecific density dependence.
 For example, increasing productivity reduces intraspecific competition in resource populations while increasing the flux of energy to consumers, resulting in the "paradox of enrichment" [@Rosenzweig1971].
-Similarly, increased capacity to find and capture resources (as we observed in our model) or lower consumer mortality, both increase the relative strength of consumer-resource interactions, which predictably destabilize consumer-resource interactions [@Rip2011; @McCann2011].
-
+Similarly, increased capacity to find and capture resources (as observed in my model) or lower consumer mortality, both increase the relative strength of consumer-resource interactions, which predictably destabilize consumer-resource interactions [@Rip2011; @McCann2011].
+
+<!-- interesting, but tighten up -->
 Empirical evidence supports the theoretical notion that food webs where consumers have undergone character displacement are less resilient to perturbations. 
@@
 
+<!-- good, but tighten up -->
 My stability results contrast, but do not necessarily contradict, the notion from competition theory that character displacement contributes to species coexistence [@Lawlor1976]. 
@@
 In fact, recent work by @McPeek2017 has shown that divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
+<!-- important caveat -->
 I want to note though, that my conclusions may only apply to food webs with biotic resources that are nutritionally substitutable. 
 It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
+
+<!--
+## Conclusions
+- needs revision
+- do this after I rewrite/edit the rest of the Discussion
 
@@
 This calls for food-web ecologists to more criticaly consider the role of evolutionary processes in shaping food-web structure and dynamics.
+-->
 
@@
 
-The content of this paper was greatly enhanced by numerous conversations with Seth Rudman.
-Sally Otto provided much help in early analyses of the mathematical model and I would not have been able to take the theory as far as I did without her guidance and encouragement.
-A discussion with Ben Gilbert pushed me to take a comparative approach that looked across different foraging scenarios to examine the generality of my conclusions.
-Discussions with Dolph Schluter pushed me to also to take a comparative approach that better set my results in the context of past theory on character displacement.
+This work was inspired by discussions with Seth Rudman, Dolph Schluter, Ben Gilbert, and Kevin McCann.
+Sally Otto provided much help in early analyses of the mathematical model. 
+I would not have been able to take the theory as far as I did without her guidance and encouragement.
+For funding support, I thank the University of British Columbia (Four-Year Fellowship to M.A.
+Barbour), NSERC (Discovery grant to Greg Crutsinger), and the Swiss National Science Foundation (grant 31003A_160671 to Jordi Bascompte).
 

### 2019-03-09 manuscript_ECD_model.Rmd
revised and tightened up Discussion

@@
 Ecological character displacement is an adaptive process, often enhancing coexistence between consumers by reducing competition for shared resources. 
-This competition perspective undermine the fact that consumer-resource interactions ultimately mediate the process of character displacement. 
+This competition perspective undermines the fact that consumer-resource interactions ultimately mediate the process of character displacement. 
 As a consequence, the effects of character displacement on consumer-resource interactions and food-web dynamics have been ignored.
@@
 
-My results show that character displacement leads to predictable differences in food (resource) availability and even food-web stability.
-Specifically, I found that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource abundances and less resilient food webs. 
-Character displacement always increased attack rates whenever consumers were competing for resources that occured in distinct habitats. 
-This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
-Although character displacement did not always increase effective attack rates, these contingent effects only occurred under the simplest and (arguably) least realistic foraging scenario. 
-My study takes advantage of past work showing that adaptive divergence in consumer attack rates is always favored under the scenarios examined here [@LawlorSmith1976; @Abrams1986].
-Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations.
-
-The ecological and evolutionary scenarios that generated the greatest divergent character displacement always resulted in decreases in resource densities.
-<!-- build paragraph around this topic sentence -->
-
+One of the criteria used to demonstrate character displacement is that "sites of sympatry [two consumers] and allopatry [one consumer] should not differ greatly in food, ..." [@Schluter1992].
+In contrast, my results suggest that ecological character displacement causes predictable differences in resource abundances.
+In fact, the ecological and evolutionary scenarios that favored the greatest displacement always decreased the abundance of resources. <!-- build paragraph around this topic sentence -->
+For example, the magnitude of character displacement was highest when mobile consumers competed for resources that occur in different habitats compared to within the same area.
+An empirical example of this is threespine stickleback that have diverged into limnetic and benthic species with specialized traits to forage on zooplankton and benthic invertebrates, respectively [@Schluter1992; @Schluter2000].
+These resources occur in distinct zones of the lake (pelagic and littoral), therefore stickleback must move between these zones and compete when foraging for resources. 
+Interestingly, a disproporationate number of the documented cases of character displacement involve carnivores [@Schluter2000] that are larger, and likely more mobile, than their resources [@McCann2005], suggesting my predictions may apply to many cases of ecological character displacement.
+Similarly, the evolutionary tradeoff that favored character displacement (concave up; orange lines) resulted in lower resource densities across all foraging scenarios.
+Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the foraging rate of consumers [@Schluter1985; @Arnegard2014].
+While it is theoretically possible that character displacement does not alter resource abundances, this was limited to the simplest, and arguably least realistic, foraging scenario and under tradeoffs that did not favor large displacements, and thus less likely to detect in nature.
+Taken together, my results call for empirical work to test these clear theoretical predictions and suggest a revision is needed for one of the criteria used to demonstrate character displacement.
+
+<!--
 The simplest foraging scenario examined here (linear functional response on resources occuring in the same habitat) is the foundation of mechanistic models of resource competition [@MacArthur1972; @Lawlor1976; @Abrams1986].
@@
 This is evident by the disruptive selection observed for beak size/shape in Darwin's finches [@Schluter1985] and benthic/limnetic ecotypes for stickleback [@Arnegard2014] in resource-limited environments.
-<!-- key point of the paragraph -->
+<!-- key point of the paragraph 
 Taken together, my model predicts that even in the simplest foraging scenario, the most likely relationship we would observe in nature is for food webs where consumers have undergone character displacement to have lower total resource densites.
@@
 A general result emerging from my analysis was that, when consumers were competing for resources in distinct habitats, character displacement always supressed resource abundances. 
-<!-- show that this is common scenario -->
+<!-- show that this is common scenario 
 Consumers are generally larger, and as a result, more mobile than their resources; thus, their foraging behavior acts to couple resource dynamics across traditional habitat boundaries [@McCann2005].
@@
 These resources occur in distinct zones of the lake (pelagic and littoral) and stickleback move between these zones when foraging for resources. 
-<!-- good closing sentences -->
+<!-- good closing sentences --
 A disproporationate number of the documented cases of character displacement involve carnivores that are larger, and likely more mobile, than their resources [@Schluter2000],
 Therefore, my predictions may apply to many cases of ecological character displacement.
-
-My results suggest that a revision is needed for one of the criteria used to demonstrate character displacement. 
-Criteria #5 in @Schluter1992 states: "Sites of sympatry and allopatry should not differ greatly in food, ...".
+-->
 
@@
 
-My most striking result was that divergent character displacement, when examined through the lens of food-web theory, generally destabilized food-web dynamics. 
-In fact, under a foraging scenario that incorporates more biological realism, character displacement may even result in an unstable food web. 
-The mechanism underlying this destabilization is quite general, resulting from the strength of interspecific interactions increasing relative to intraspecific interactions [@Chesson2008; @Rip2011; @McCann2011].
-The strength of interspecific interactions increases whenever consumers evolved higher net attack rates, but does not change intraspecific interactions.
-This relative increase in interspecific interactions, combined with the natural oscillatory tendency of consumer-resource interactions [@Murdoch2003], creates a food-web structure that is less resilient to perturbations.
-
-<!-- good, but tighten up -->
-My results, combined with recent theory, suggest that the very conditions that favor phenotypic diversification also make that diversity more susceptible to local extinction.
-Specifically, work by @McPeek2019 showed that character displacement is favored in environments that are either highly productive, easy to find and capture resources, or under weak abiotic stress. 
-In my model, this corresponds to higher values of $K$ (productivity) or $a_{ij}$ (finding and capturing resources), or lower values of $m$ (abiotic stress).
-Each of these corresponding changes create food webs that have a lower baseline of stability, as they increase the strength of consumer-resource interactions relative to intraspecific density dependence.
-For example, increasing productivity reduces intraspecific competition in resource populations while increasing the flux of energy to consumers, resulting in the "paradox of enrichment" [@Rosenzweig1971].
-Similarly, increased capacity to find and capture resources (as observed in my model) or lower consumer mortality, both increase the relative strength of consumer-resource interactions, which predictably destabilize consumer-resource interactions [@Rip2011; @McCann2011].
-
-<!-- interesting, but tighten up -->
-Empirical evidence supports the theoretical notion that food webs where consumers have undergone character displacement are less resilient to perturbations. 
-For example, a single species of threespine stickleback lives in hundreds of postglacial lakes of British Columbia, but the species pairs, where character displacement has happened, are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
-Perhaps this rarity of species pairs is a result of nonadaptive selection for more ecologically stable food-web topologies [@Borrelli2015a; @Borrelli2015b].
-Indeed, it is clear that these species pair are vulnerable to perturbations, given that the species pairs have went extinct in two lakes following the introduction of nonnative species [@Hatfield2001; @Taylor2006; @Rudman2016].
-The vulnerability of the stickleback system also aligns well with evidence that aquatic food webs have several properties that make them less resilient to perturbations, including higher productivity and more efficient transfer of energy to consumers [@Rip2011].
-Testing this hypothesis may be quite difficult, but it could be possible with recent advances in genomics.
+My most striking result was that divergent character displacement, when examined through the lens of food-web theory, made food webs less resilient to perturbations.  
+In fact, under the most realistic foraging scenario, character displacement can even result in an unstable food web. 
+The mechanism underlying this destabilization is quite general.
+Character displacement generally increases the strength of consumer-resource interactions, but does not alter the strength of intraspecific interactions.
+This relative increase in interspecific interactions, combined with the natural oscillatory tendency of consumer-resource interactions [@Murdoch2003], creates a food-web structure that is less resilient to perturbations [@Chesson2008; @Rip2011; @McCann2011].
+
+Interestingly, the ecological conditions that favor divergent character displacement are those that are already the least resilient to perturbations.
+For example, @McPeek2019 showed that character displacement is favored in food webs that are either highly productive, easy to find and capture resources, or under weak abiotic stress.
+In my model, this corresponds to higher values of $K$ (productivity) or $A$ (capacity to find and capturing resources), or lower values of $m$ (abiotic stress).
+Each of these corresponding changes decrease food-web resilience, as they increase the strength of consumer-resource interactions relative to intraspecific interactions.
+For example, increasing productivity reduces intraspecific competition in resource populations while increasing the flux of energy to consumers, resulting in the paradox of enrichment [@Rosenzweig1971].
+Similarly, higher feeding rates or lower consumer mortality both increase the relative strength of consumer-resource interactions, which predictably destabilizes food webs [@Rip2011; @McCann2011].
+This suggests that the most dramatic examples of character displacement will not only occur in, but also cause, the least stable food-web structures.
+
+A handful of empirical patterns support the hypothesis that character displacement decreases food-web resilience. 
+As an example, a single species of threespine stickleback lives in hundreds of postglacial lakes of British Columbia, but the species pair <!--, where character displacement has happened,-->are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
+Perhaps many lakes had a species pair in the past, but have lost a species due to a less resilient food-web structure [@Borrelli2015a; @Borrelli2015b].
+The species pair are known to be vulnerable to perturbations, as they have gone extinct in two of the six lakes after the introduction of nonnative species [@Hatfield2001; @Taylor2006; @Rudman2016].
+The vulnerability of the stickleback system also corresponds with the fact that aquatic food webs have several properties that make them less resilient to perturbations, such as higher productivity and more efficient energy transfer to consumers [@Rip2011].
+Detecting the ghost of competition past [@Connell1980] may be quite difficult, but it could be possible with recent advances in genomics.
+<!--Testing this hypothesis ---that some lakes with a single species had a species pair in the past --- may be quite difficult, but it could be possible with recent advances in genomics.-->
 For example, @Feulner2018 detected genomic signatures of hybridization in sympatric whitefish species following periods of eutrophication.
-Perhaps stickleback in many other postglacial lakes in British Columbia retain genomic signatures of having been a habitat specialist in the past.
-
-<!-- good, but tighten up -->
-My stability results contrast, but do not necessarily contradict, the notion from competition theory that character displacement contributes to species coexistence [@Lawlor1976]. 
-Competition theory typically uses a different measure of stability that considers the mutual ability of consumers with different phenotypes to invade when rare [mutual invasibility, @Chesson2000].
-A shortcoming of this mutual invasibility measure, however, is that it does not allow a comparison between food webs with and without a competiting consumer. 
-Such comparisons are necessary for inferring the effects of ecological character displacement, a point that has been made clear in the criteria to demonstrate character displacement [@Schluter1992; @Schluter2000].
-Although the addition of consumers can decrease food-web resilience in the absence of evolution [@May1973], our results are primarily driven by an eco-evolutionary feedback between consumer evolution and resource abundances.
-
-I expect that my theoretical predictions are robust to different assumptions about the underlying evolutionary dynamics of consumer foraging traits.
-This is because models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of whether a quantitative genetic or adaptive dynamics model is used for consumer traits [@Lawlor1976; @Taper1985].
-A quantitative genetic model may certainly show differences in the magnitude of character displacement, but this should not affect the qualtiative relationship I observed.  
-In fact, recent work by @McPeek2017 has shown that divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
-<!-- important caveat -->
-I want to note though, that my conclusions may only apply to food webs with biotic resources that are nutritionally substitutable. 
+Perhaps solitary stickleback in some lakes retain genomic signatures of having been a habitat specialist in the past.
+
+<!-- interesting, but tighten up 
+A plausible explanation for this rarity is that the food-web structure makes the species pair more vulnerable to extinction, and many lakes that used to have a species pair have  single species of stickleback are more stable
+A plausible explanation for this rarity is that nonadaptive selection results in a more ecologically stable food-web topology [@Borrelli2015a; @Borrelli2015b].
+Testing this hypothesis ---some lakes with a single species had two in the past --- may be quite difficult, but it may be feasible with recent advances in genomics.
+For example, @Feulner2018 detected genomic signatures of hybridization in sympatric whitefish species following periods of eutrophication.
+Perhaps stickleback in some lakes retain genomic signatures of having been a habitat specialist in the past.
+Other evidence also suggests that the species pair of stickleback are vulnerable to perturbations.
+For example, the introduction of nonnative species into two of the six lakes have resulted in the extinction of the species pair [@Hatfield2001; @Taylor2006; @Rudman2016].
+-->
+
+My results contrast, but do not necessarily contradict, the notion from competition theory that character displacement contributes to species coexistence [@Lawlor1976]. 
+Rather than studying resilience, competition theory usually studies the mutual ability of consumers with different phenotypes to invade when rare [mutual invasibility, @Chesson2000].
+In the context of character displacement, a shortcoming of this mutual invasibility measure is that it does not allow a comparison between food webs with and without a competiting consumer. 
+Such comparisons are necessary for inferring the effects of character displacement, a point that has been made clear in the criteria to demonstrate character displacement [@Schluter1992; @Schluter2000].
+Although the addition of a consumer to a food web can decrease its resilience in the absence of evolution [@May1973], my results are primarily driven by an eco-evolutionary feedback between consumer evolution and resource abundances (compare initial and ending points in Fig.
+\ref{fig:MacArthur_Plot}C, \ref{fig:LS_Plot}C, \ref{fig:McCann_Plot}C).
+
+The Adaptive Dynamics approach I used here has its strengths and weaknesses.
+A strength of this approach is that it enabled me to gain analytical insight to the effects of character displacement on food-web dynamics in a more realistic foraging scenario.
+This is much less tractable in quantitative genetic [@Taper1985; @McPeek2017] or explicit genetic [@Doebeli1996] models of character displacement, which is why the foraging scenarios previously examined have been limited [but see @McPeek2017].
+A weakness, however, is that I assume a separation of time scales between ecological and evolutionary dynamics, an assumption that is becoming less tenable [@Hairston2005; @Hendry2016].
+I also do not explicitly model an underlying phenotypic trait for consumer attack rates nor do I allow for intraspecific variation. 
+That being said, my theoretical predictions are likely robust to these assumptions. 
+This is because models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
+A quantitative genetic model may certainly show differences in the magnitude of character displacement, but this should not affect the general relationships I observed.  
+<!-- In fact, recent work by @McPeek2017 has shown that divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics. -->
+I want to note though, that my conclusions only apply to food webs with biotic resources that are nutritionally substitutable. 
 It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
+
+<!-- Omitting conclusions section for now. I will wait for feedback from friendly reviews
+
+## Conclusions
+
+<!-- originally opening paragraph of discussion; I think it provides too good of a summary though to put at the beginning. It makes subsequent discussion sound redundant
+My results show that character displacement leads to predictable differences in resource abundances and the resilience of food webs to perturbations.
+Specifically, I found that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource abundances and less resilient food webs. 
+Character displacement always increased attack rates whenever consumers were competing for resources that occured in distinct habitats. 
+This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
+Although character displacement did not always increase effective attack rates, these contingent effects only occurred under the simplest and (arguably) least realistic foraging scenario. 
+My study takes advantage of past work showing that adaptive divergence in consumer attack rates is always favored under the scenarios examined here [@LawlorSmith1976; @Abrams1986].
+Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations.
 

### 2019-03-11 manuscript_ECD_model.Rmd
add evolution results in McCann model and remove small points from all panels except for C (stability)

@@
 
+*Number of words*: `r wordcountaddin::word_count("manuscript_ECD_model.Rmd")`
+
+\bigskip
+
 *Manuscript type*: Note. <!-- consider e-Article, depending on length -->
@@
 seed size, see @Taper1985 for an example).
+<!--
 In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972].
@@
 I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
-
+-->
 To investigate the ecological consequences of divergent character displacement, I analyzed differences in resource abundances and food-web stability at equilibrium. 
@@
 
+In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972].
+The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
+I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
+
 When both consumers and resources are present, the abundance of resources at equilibrium are equivalent ($\hat R = \hat R_1 = \hat R_2$) and are determined by the following equation (derivation in supplementary material):
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
-  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
+  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(a[ii]+a[ij])) +
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
-  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
+  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
   geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(w*a[ii]+(1-w)*a[ij])) +
@@
   geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
 
+The evolutionary dynamics of consumer attack rates have not been examined in this model, but as with the other models [@LawlorSmith1976], I find that divergent character displacement is inevitable. 
+This is because, in the absence of competition, consumer evolution is determined by both resource abundances and their habitat preferences, and reach an evolutionary stable strategy (ESS) when:
+
+$$w_{ii}R_i^2\delta_{ii}=w_{ij}R_j^2\delta_{ij}$$ 
+where $\delta_{ii}$ represents a small mutation in $a_{ii}$ and we assume a negative tradeoff in attack rates, such that $\delta_{ij}=-f(\delta_{ii})$. 
+The square-terms for resource abundances indicate that they largely govern the evolutionary dynamics. 
+This scenario favors the evolution of a generalist phenotype, as resource specialization will decrease that resource's abundance, thus favoring evolution toward the other resource until an ESS is reached.
+In contrast, in the presence of a competitor, resource abundances will be equivalent at equilibrium (derivation in Appendix S2) and consumers reach an ESS when:
+
+$$w_{ii}\delta_{ii}=w_{ij}\delta_{ij}$$
+
+Assuming consumers have a preferred resource (i.e. $w_{ii}>0.5$), then consumers will only reach an ESS when $|\delta_{ii}|<|\delta_{ij}|$ (e.g. green line, Fig. \ref{MacArthur_Plot}A). 
+Otherwise, they will evolve to be complete specialists on their preferred resource.
+
+Divergent character displacement in this foraging scenario has the same general effect on food-web dynamics as the previous scenario (resources in different habitats, but linear functional response). 
+<!--In terms of food-web dynamics, I observe the same general effect of ecological character displacement as before when resources occurred in distinct habitats, but consumers had a simpler functional response.
+-->
+This is because resource abundances at equilibrium are governed by a similar dynamic (derivation in Appendix S2):
+<!--
 Even after adding more biological realism, I still observe the same qualitative relationship between character displacement and resource suppression (Fig.
@@
 This is because resource competition always leads to character divergence and equilibrium resource abundances are governed by similar dynamics  (derivation in supplementary material):
+-->
 
@@
 
-In the previous models, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$).
+And since evolution favors consumer divergence onto their preferred resources, the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase (Fig.
+\ref{fig:McCann_Plot}A), resulting in lower resource abundances (Fig.
+\ref{fig:McCann_Plot}B) and decreased food-web stability (Fig.
+\ref{fig:McCann_Plot}C).
+
+In the previous models, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$, derivation in Appendix S1 and S2).
 In this more realistic model, however, whether the food web is locally stable depends on the parameters of both consumers and resources. 
-Specifically, the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derivation in supplementary material):
+Using Routh-Hurwitz criteria, I found that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derivation in Appendix S2):
 
@@
 
-Previously, I showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff (Fig.
-\ref{fig:LS_Plot}A).
-Thus, character displacement always pushes the food web toward the boundary of local stability under this more realistic foraging scenario.
+<!--Previously, I showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff (Fig.
+\ref{fig:LS_Plot}A).-->
+Thus, character displacement always pushes the food web toward an unstable stucture in this more realistic foraging scenario.
 
@@
   geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(w*a[ii]+(1-w)*a[ij])) +
@@
   geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
   geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(a[ii]+a[ij])) +
@@
   geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + 
   geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
@@
   geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(w*a[ii]+(1-w)*a[ij])) +
@@
   geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
   geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(w*a[ii]+(1-w)*a[ij])) +
@@
   geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  #geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
   geom_line() +
-  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + 
   geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +

### 2019-03-14 eco_evo_sim_function.R
adjust mutation size to 0.01 instead of 0.001

@@ -12,7 +12,7 @@ eco_evo_CR <- function(init_parameters,              # data frame of initial par
                        evo_tradeoff_names,           # vector of mutant trade-off parameter names in the Adaptive Dynamics model. Order must correspon to 'eco_param_names'
                        tradeoff_functions,           # names function of the trade-off with corresponding parameters
                        extra_tradeoff_params=c(NOT_USED=1), # vector of extra tradeoff parameters. Default is a nonsense parameter to just make sure the function runs
-                       mut.size=0.001,               # absolute change in evolving parameter. Directionality is chosen at random.
+                       mut.size=0.01,                # absolute change in evolving parameter. Directionality is chosen at random.
                        mut.suc.thres=1e-4,           # threshold of mutant success rate for identifying ESS.
                        j.step=500,                   # number of evolutionary time steps over which to assess evolutionary change.
                        abund.thres=1e-4,             # abundance threshold for feasible equilibriums

### 2019-03-18 manuscript_ECD_model.Rmd
format title page for ecology letters

@@
 
-$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch
+$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch, phone: +41 44 635 57 11
 
@@
 
-*Manuscript elements*: Figure 1, figure 2, and figure 3. All figures should be printed in color.
+*Running title*: Character divergence destabilizes food webs
 
@@
 
-*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptation; ecological community; stability
+*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptation; community stability
 
@@
 
-*Number of words*: `r wordcountaddin::word_count("manuscript_ECD_model.Rmd")`
+*Manuscript type*: Letter. <!-- consider e-Article, depending on length -->
 
@@
 
-*Manuscript type*: Note. <!-- consider e-Article, depending on length -->
-
-\bigskip
-
-\footnotesize Prepared using an *Am. Nat.* inspired \LaTeX{} template for Rmarkdown. \normalsize
+*Number of words in abstract*:
+
+*Number of words in main text*:
+
+*Number of references*:
+
+*Number of figures*:
 
@@
 @Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has examined which scenarios lead to phenotypic divergence or convergence of competing consumers.
-The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of abundance dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985].
+The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of abundance dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985; \ref{fig:fig_1}].
 This adaptive process is not simply a response to statitc differences in resource distributions, but creates an eco-evolutionary feedback that drives further differentiation. 
@@
 
-# Methods
+```{r fig_1, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:fig_1}**Divergent character displacement.** A textbook example of divergent character displacement is threespine stickleback in small coastal lakes of British Columbia.
+Sticklback consume two main resources, zooplankton and benthic invertebrates.
+In isolation, stickleback evolve a generalist phenotype that can take advantage of both resources.
+In the presence of a competitor, however, stickleback evolve specialized phenotypes that enable them to capture different resources more efficiently.
+A large body of theoretical work has shown that resource competition generally drives the adaptive divergence of consumers onto different resources.
+Currently though, we lack theoretical predictions for how this adaptive divergence affects consumer-resource interactions and food-web dynamics."}
+knitr::include_graphics("ECD_model_fig_1.pdf")
+```
+
+# Material and methods
+
+## Underlying Consumer-Resource Dynamics
 
@@
 seed size, see @Taper1985 for an example).
-<!--
-In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972].
-The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
+
+## Foraging Scenarios
+
+I studied three different foraging scenarios. 
+In the first, I assume that consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that:
+
+$$F_{ji}(R_i)=a_{ji}R_i$$
+
+where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972]. 
+This first scenario is the starting point for many models of resource competition; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005].
+A simple extension of the above scenario takes the form:
+
+$$F_{ji}(R_i)=w_{ji}a_{ji}R_i$$
+
+where $w_{ji}$ represents the proportion of time consumer $j$ spends foraging in a habitat where only resource $j$ is found (i.e. habitat preference). 
+Note that since $w_{ji}$ is a proportion that $w_{jj}=1-w_{ji}$. 
+Finally, it is well known that consumer feeding rates often saturate at high resource abundances [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011] and they do not usually spend a fixed proportion of time in a particular habitat [@McCann2005].
+To account for these biological realities, I consider a more realistic functional response [derived by @McCann2005]:
+
+$$F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}hW_{ii}R_i+a_{ij}hW_{ij}R_j}$$
+
+where consumer $i$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling time $h$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}$.
+
+Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach [@Otto2007], with the general result being divergent character displacement [@Lawlor1976; @Abrams1986].
 I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
--->
-To investigate the ecological consequences of divergent character displacement, I analyzed differences in resource abundances and food-web stability at equilibrium. 
+I also used an Adaptive Dynamics approach to analyze character displacement in the third foraging scenario, and I too observed divergent character displacement (detailed analysis in Appendix S1 of Supplementary Information).
+
+```{r fig_2, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:fig_2}**Foraging scenarios and evolutionary tradeoffs.** I examined the effect of ecological character displacement on food-web dynamics under different assumptions of ecological foraging scenarios and evolutionary trade-offs in consumer foraging traits.
+Specifically, I examined whether these effects depended on whether resources occurred in the same habitat (top left) or in different habitats (top right).
+I also included a more realistic functional responses.
+In each of these foraging scenarios, I examined the effect of different evolutionary tradeoffs in attack rates."}
+knitr::include_graphics("ECD_model_fig_2.pdf")
+```
+
+## Food-web Dynamics
+
+Given that divergent character displacement was an expected result, I focus here on its consequences for food-web dynamics across these foraging scenarios.
+To do this, I analyzed differences in resource abundances and food-web stability at equilibrium. 
 An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$).
@@
 When possible, I derived analytical expressions for the relationship between consumer attack rates and food-web dynamics. 
-To do this, I simplified the model by assuming that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$).
+To do this, I simplified the model by assuming that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $h=h_{11}=h_{12}=h_{21}=h_{22}$; $m=m_1=m_2$), except that consumer attack rates and their habitat preferences (if present) are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$; $w_{ii}=w_{11}=w_{22}$).
 Controlling for other sources of variability allowed me to isolate the general effects of divergent character displacement on food-web dynamics.
 
+To gain insight to the eco-evolutionary feedback generated by divergent character displacement, I conducted simulations using an Adaptive Dynamics approach. 
+<!--
+The evolutionary dynamics of character displacement have been well characterized in some of the ecological scenarios examined here [@Lawlor1976; @Abrams1986] and are not the focus of this work. 
+Still, 
 Although divergent character displacement occurs in all scenarios we examined, the magnitude of this divergence may be context dependent.
 Therefore, I simulated the eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
-At each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate on one resource by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability.
+-->
+Specifically, at each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate on one resource by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability.
 The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006].
@@
   ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
-  geom_line() + 
-  geom_point(data = initial_points, aes(x=a11, y=a12)) +
-  xlab(expression(a[ii])) +
-  ylab(expression(a[ij])) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)
-
-tradeoff_plot
+  geom_line(size=1.5) + 
+  geom_point(data = initial_points, aes(x=a11, y=a12), size=5, show.legend=FALSE) +
+  xlab(expression(bold(a[ii]))) +
+  ylab(expression(bold(a[ij]))) +
+  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette) +
+  theme_cowplot(font_size = 30, line_size=2)# +
+  #theme(axis.title = element_text(size = 30),
+  #      axis.text = element_text(size = 20),
+  #      legend.title = element_text(size = 25),
+  #      legend.text = element_text(size = 20),
+  #      legend.spacing = unit(1.0,'cm'))
+
+save_plot(filename = "attack_tradeoffs.pdf", plot = tradeoff_plot, base_height = 6, base_width = 9)
 ```
@@
 
-<!-- I haven't made clear that this explores a specific scenario, MacArthur's model -->
+<!-- I haven't made clear that this explores a specific scenario, MacArthur's model
 
@@
 I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
-
-When both consumers and resources are present, the abundance of resources at equilibrium are equivalent ($\hat R = \hat R_1 = \hat R_2$) and are determined by the following equation (derivation in supplementary material):
+-->
+
+In this first scenario, the abundance of resources at equilibrium are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), and are determined by the following equation (derived in Appendix S2):
 
@@
   ylab(expression(a[ii]+a[ij])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
   scale_shape_manual(values=c(24,21)) +
@@
   ylab(expression(R[i]+R[j])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
   scale_shape_manual(values=c(24,21)) +
@@
   geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+ #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
   ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
@@
   geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
   ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
@@
 
-```{r MacArthur_Plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}**Effect of character displacement when consumers can forage for both resources simultaneously.** Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r MacArthur_Plot, include=FALSE, eval=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}**Effect of character displacement when consumers can forage for both resources simultaneously.** Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Panel (B) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
@@
 ## Consumers *cannot* forage for both resources simulanteously
-
+<!--
 The previous model, although the starting point for many models of competition, does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in distinct habitats [@McCann2005].
@@
 As before, resource abundances are equivalent at this equilibrium ($\hat R = \hat R_1 = \hat R_2$ when both consumers and resources present), but are now determined by the following equation (derivation in supplementary material):
+-->
+In this second scenario, I again see that resource abundances are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), but are now determined by the following equation (derived in Appendix S3):
 
@@
   ylab(expression(w*a[ii]+(1-w)*a[ij])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
   scale_shape_manual(values=c(24,21)) +
@@
 
-```{r LS_Plot, echo=FALSE, fig.cap="\\label{fig:LS_Plot}**Effect of character displacement when consumers *cannot* forage for both resources simultaneously.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r LS_Plot, include=FALSE, eval=FALSE, fig.cap="\\label{fig:LS_Plot}**Effect of character displacement when consumers *cannot* forage for both resources simultaneously.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
@@
 
+```{r Fig 3, include=FALSE}
+# plot.margin = unit(c(0, 0, 0, 0), "cm") to shrink space between plots
+
+Macs <- plot_grid(
+  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[ii]+a[ij]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none",  axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0)), #  + scale_y_continuous(breaks=c(1.8,2.0,2.2))
+  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text.x = element_blank(), axis.title.y=element_text(vjust=0)), # + scale_y_continuous(breaks=c(1.0,1.2,1.4), limits=c(1.0,1.45))
+  Stability_Mac + xlab("")  +  ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + theme(legend.position = "none", axis.title.y=element_text(vjust=0)), # + scale_y_continuous(breaks=c(0,0.1))
+  ncol=1, align='v'#, labels = "AUTO"
+) # c(0,0.02,0.04,0.06,0.08,0.1)
+
+LSs <- plot_grid(
+  LS_attack  + xlab("") + ylab(bquote(w*a[ii]+"("*1-w*")"*a[ij])) + ggtitle("Resources: Different Habitats\n") + theme(legend.position = "none", axis.text.x = element_blank(), plot.title = element_text(size = 10)), #+ scale_y_continuous(breaks=c(0.8,1.0,1.2))
+  LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text.x = element_blank()), #+ scale_y_continuous(breaks=c(2.0,2.5,3.0), limits=c(2.0,3.2))  
+  Stability_LS + xlab("") + ylab("") + theme(legend.position = "none"), # + scale_y_continuous(breaks=c(0,0.1,0.2))
+  ncol=1, align='v'
+)
+
+plot_fig3 <- plot_grid(Macs, LSs, ncol=2, rel_widths = c(1,0.9)) 
+plot_fig3_sub <- add_sub(plot_fig3, expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")"))), vjust = 0)
+
+save_plot(filename="Figure_3.pdf", plot=plot_fig3_sub, base_height = 8, base_width = 6)
+```
+
+```{r plot_fig3, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:plot_fig3}**Effect of character displacement on food-web dynamics.** Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85; as in Fig.
+1).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
+Different panels show the effect of character displacement on effective attack rates (A), resource abundances (B), and food-web stability (C) when resources occur in the same vs.
+different habitats.
+Note that the equation for effective attack rate depends on the foraging scenario.
+These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6; *w~ij~*=0.4.
+I set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
+I set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
+knitr::include_graphics("Figure_3.pdf")
+```
+
 ## Adding a more realistic functional response
 
+<!--
 So far I have only considered a simple functional response, where consumer feeding rates increase linearly with resource abundances when foraging in either the same or distinct habitats. 
@@
 $$W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}$$
-
-<!-- Need to work on the flow here and mention the evolutionary results. Perhaps it would be best to parallel my writing style as in the previous sections. -->
 
@@
 Otherwise, they will evolve to be complete specialists on their preferred resource.
-
-Divergent character displacement in this foraging scenario has the same general effect on food-web dynamics as the previous scenario (resources in different habitats, but linear functional response). 
+-->
+
+In the third foraging scenario, I observed the same general effect of divergent character displacement as the previous scenario (resources in different habitats, but linear functional response).
+<!--Divergent character displacement in this foraging scenario has the same general effect on food-web dynamics as the previous scenario (resources in different habitats, but linear functional response).
 <!--In terms of food-web dynamics, I observe the same general effect of ecological character displacement as before when resources occurred in distinct habitats, but consumers had a simpler functional response.
 -->
-This is because resource abundances at equilibrium are governed by a similar dynamic (derivation in Appendix S2):
+This is because resource abundances at equilibrium are governed by a similar dynamic (derived in Appendix S1):
 <!--
@@
 
-And since evolution favors consumer divergence onto their preferred resources, the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase (Fig.
+And since evolution favors consumer divergence onto their preferred resources (see Appendix S1), the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase (Fig.
 \ref{fig:McCann_Plot}A), resulting in lower resource abundances (Fig.
@@
 
-In the previous models, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$, derivation in Appendix S1 and S2).
-In this more realistic model, however, whether the food web is locally stable depends on the parameters of both consumers and resources. 
-Using Routh-Hurwitz criteria, I found that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derivation in Appendix S2):
+In the previous foraging scenarios, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$, see Appendix S1 and S2).
+In this more realistic model, however, whether the food web is locally stable depends on consumer and resource parameters. 
+Specifically, I found that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derived using Routh-Hurwitz criteria in Appendix S1):
 
@@
   #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
+  ylab(bquote(atop("Effective attack rate","("*w*a[ii]+"("*1-w*")"*a[ij]*")"))) +
+  #ylab(expression(w*a[ii]+(1-w)*a[ij])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
   #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(R[i]+R[j])) +
+  ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) +
+  #ylab(expression(R[i]+R[j])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability "(-lambda[max]))) + 
+  ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) +
+  #ylab(expression("Stability "(-lambda[max]))) + 
   scale_shape_manual(values=c(24,21))  +
@@
 save_plot(filename="Figure_ECD_McCann.pdf", plot=plot_McCann_model, base_height = 6, base_width = 6)
-```
-
-```{r McCann_Plot, echo=FALSE, fig.cap="\\label{fig:McCann_Plot}**Effect of character displacement when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
-Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
-This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.
-We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
-knitr::include_graphics("Figure_ECD_McCann.pdf")
+# moved to supplementary information
+```
+
+```{r bifurcation, echo=FALSE, cache=TRUE}
+
+## set state variables; same as previous models but in vector form
+i.state.2C_2R <- c(R1 = R1, R2 = R2, C1 = C1, C2 = C2) 
+i.state.1C_2R <- c(R1 = R1, R2 = R2, C1 = C1)
+
+## set duration of simulations
+Time <- 1000
+
+# when r = 2.5, A = 3.3 and concaveUp=0.85, we get consmers evolving toward instability; we don't get this when n=n_linear but we do if we increase mut.size=0.01 (order of magnitude) and even when r=1!
+
+## IDEAL PARAMETERS RIGHT NOW: A=3.3, r=1, mut.size=0.01, n=n_linear (excellent looking figure with bifurcation just after0.8)
+
+## set up general code for bifurcation plot; increase to get bifurcation
+# when A = 2.5 get strong divergence
+# at 2.9, consumer doesn't divergence, C1 becomes generalist though
+# at 2.8, there is less divergence...
+A <- 3.3
+# K <- K+0.5 # increasing K to 4.5 creates a bifurcation at extreme specialization, but the consumer no longer evolves...
+#r <- r #2.5 # r=1.5, now the consumer is still diverging when A = 2.9... at 3.3 though, there is no longer divergence in two consumer situation; get divergence again when r=2
+#aii <- 2.4
+#aij <- 2.0
+
+# assuming linear trade-off in attack rates for simulation; sometimes, the consumer doesn't evolve toward asymmetry though...
+#linear_df <- data.frame(n = n_linear, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_linear)^(1/n_linear))
+aii_seq <- seq(A/2, A, by = 0.01)#[1:75] # seq(aii, A, by = 0.01)[1:75]
+aij_seq <- (A*(1-(aii_seq/A)^n_linear)^(1/n_linear)) # n_linear #[1:75] #seq(A/2, 0, by = -0.01)[1:75]
+
+## Bifurcation plot code
+# this code loops over the sequence of different parameter values
+out <- list()
+for (i in 1:length(aii_seq)) {
+
+  r1 <- r2 <- r 
+  K1 <- K2 <- K 
+  e11 <- e12 <- e21 <- e22 <- e 
+  h11 <- h12 <- h21 <- h22 <- h 
+  m1 <- m2 <- m 
+  w11 <- w22 <- wii
+  w12 <- w21 <- wij
+
+  # set params 
+  param.loop <- c(r1 = r1, r2 = r2, K1 = K1, K2 = K2,
+                  e11 = e11, e12 = e12, e21 = e21, e22 = e22,
+                  h11 = h11, h12 = h12, h21 = h21, h22 = h22,
+                  a11 = aii_seq[i], a12 = aij_seq[i], a21 = aij_seq[i], a22 = aii_seq[i],
+                  m1 = m1, m2 = m2, w11 = w11, w22 = w22, w12 = w12, w21 = w21)
+
+  # Run the experiment.
+  init <- ode(i.state.2C_2R, 
+              1:Time, McCann_2C_2R, param.loop)
+  # Rerun the experiment with the state variables at the end of the initial simulation. This will enable me to determine the range of state variables around the equilibrium.
+  temp <- ode(init[Time,-1], 1:Time, McCann_2C_2R, param.loop)[,-1]
+  out[[i]] <- cbind(temp, Rt = temp[,"R1"] + temp[,"R2"], Ct = temp[,"C1"] + temp[ ,"C2"])
+}
+
+sim.list <- list()
+for (i in 1:length(aii_seq)) {
+  sim.list[[i]] <- data.frame(rep = i,
+                              R_max = max(out[[i]][900:1000,"Rt"]),
+                              R_min = min(out[[i]][900:1000,"Rt"]),
+                              C_max = max(out[[i]][900:1000,"Ct"]),
+                              C_min = min(out[[i]][900:1000,"Ct"]))
+}
+
+sim.df <- ldply(sim.list) %>%
+  mutate(a11 = aii_seq,
+         a12 = aij_seq) %>%
+  gather(state, value, -rep, -a11, -a12) %>%
+  separate(state, into = c("species","max_min"))
+
+## Simulate 3 species food web ----
+
+out_3 <- list()
+for (i in 1:length(aii_seq)) {
+
+  r1 <- r2 <- r
+  K1 <- K2 <- K
+  e11 <- e12 <- e21 <- e22 <- e
+  h11 <- h12 <- h21 <- h22 <- h
+  m1 <- m2 <- m
+
+  # set params 
+  param.loop <- c(r1 = r1, r2 = r2, K1 = K1, K2 = K2,
+                  e11 = e11, e12 = e12, 
+                  h11 = h11, h12 = h12,
+                  a11 = aii_seq[i], a12 = aij_seq[i], 
+                  m1 = m1, w11 = w11, w12 = w12)
+
+  # Run the experiment.
+  init <- ode(i.state.1C_2R, 
+              1:Time, McCann_1C_2R, param.loop)
+  # Rerun the experiment with the state variables at the end of the initial simulation; this will enable me to determine the range of state variables around the equilibrium.
+  temp <- ode(init[Time,-1], 1:Time, McCann_1C_2R, param.loop)[,-1]   
+  out_3[[i]] <- cbind(temp, Rt = temp[,"R1"] + temp[,"R2"])
+}
+
+sim.list_3 <- list()
+for (i in 1:length(aii_seq)) {
+  sim.list_3[[i]] <- data.frame(rep = i,
+                                R_max = max(out_3[[i]][900:1000,"Rt"]),
+                                R_min = min(out_3[[i]][900:1000,"Rt"]),
+                                C_max = max(out_3[[i]][900:1000,"C1"]),
+                                C_min = min(out_3[[i]][900:1000,"C1"]))
+}
+
+sim.df_3 <- ldply(sim.list_3) %>%
+  mutate(a11 = aii_seq,
+         a12 = aij_seq) %>% 
+  gather(state, value, -rep, -a11, -a12) %>%
+  separate(state, into = c("species","max_min"))
+```
+
+```{r bifurcation simulation, include=FALSE, cache=TRUE}
+## eco-evo sims
+
+## respecify initial attack rates based on new A-value
+aii <- A*0.61
+
+# params
+params_2C_2R_McCann_linear_bifur <- data.frame(general_parameters,
+                                               a11 = aii,
+                                               a12 = eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_linear)), 
+                                               a21 = eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_linear)), 
+                                               a22 = aii,
+                                               w11 = wii,
+                                               w12 = wij,
+                                               w21 = wij,
+                                               w22 = wii,
+                                               h11 = h,
+                                               h12 = h,
+                                               h21 = h,
+                                               h22 = h)
+
+params_1C_2R_McCann_linear_bifur <- params_2C_2R_McCann_linear_bifur[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","h11","h12")]
+
+# eco-evo simulations
+sim_2C_2R_McCann_linear_bifur <- eco_evo_CR(init_parameters = params_2C_2R_McCann_linear_bifur, 
+                                             init_states = states_2C_2R, # same as previous models
+                                             eco_CR_model = McCann_2C_2R, 
+                                             AD_CR_models = c(mC1_McCann, mC2_McCann),
+                                             species_traits = c(1,2), 
+                                             eco_param_names = c("a11","a22"), 
+                                             eco_tradeoff_names = c("a12","a21"),  
+                                             evo_param_names = c("a11m","a22m"),  
+                                             evo_tradeoff_names = c("a12m","a21m"), 
+                                             tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
+                                             extra_tradeoff_params = c(A=A, n=n_linear)) 
+sim_1C_2R_McCann_linear_bifur <- eco_evo_CR(init_parameters = params_1C_2R_McCann_linear_bifur, 
+                                             init_states = states_1C_2R, # same as previous models
+                                             eco_CR_model = McCann_1C_2R, 
+                                             AD_CR_models = c(mC1_McCann),
+                                             species_traits = c(1), 
+                                             eco_param_names = c("a11"), 
+                                             eco_tradeoff_names = c("a12"),  
+                                             evo_param_names = c("a11m"),  
+                                             evo_tradeoff_names = c("a12m"), 
+                                             tradeoff_functions = c(a12_tradeoff),
+                                             extra_tradeoff_params = c(A=A, n=n_linear))
+
+# save simulation output 
+write.csv(sim_2C_2R_McCann_linear_bifur, "sim_2C_2R_McCann_linear_bifur.csv")
+write.csv(sim_1C_2R_McCann_linear_bifur, "sim_1C_2R_McCann_linear_bifur.csv")
+```
+
+```{r bifur evo plot, echo=FALSE}
+McCann_displacements_bifur <- bind_rows(
+  mutate(sim_2C_2R_McCann_linear_bifur, Model = "McCann", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_linear_bifur)), ], 
+  mutate(sim_1C_2R_McCann_linear_bifur, Model = "McCann", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_linear_bifur)), ]) 
+McCann_displacements_bifur$Time <- rep(c("Begin","End"),2)
+
+## put it all together
+
+# consumer abundances
+#bind_rows(mutate(sim.df, size = 4), mutate(sim.df_3, size = 3)) %>%
+#  filter(species == "C") %>%
+#  ggplot(., aes(x = a11/(a11+a12), y = value, group=interaction(max_min, size), color=as.factor(size))) +
+#  geom_line() +
+#  ylab("Consumer abundance (max and min)")
+
+#get_stability_boundary <- bind_rows(mutate(sim.df, Competitor = "Yes"), mutate(sim.df_3, Competitor = "No")) %>%
+#  filter(species == "R") %>%
+#  spread(max_min, value)
+
+# with eco-evo begin and end points
+bifurcation_plot <- bind_rows(mutate(sim.df, Competitor = "Yes"), mutate(sim.df_3, Competitor = "No")) %>%
+  filter(species == "R") %>%
+  ggplot(., aes(x = a11/(a11+a12), y = value, group=interaction(max_min, Competitor))) +
+  geom_line(aes(linetype=Competitor), color=cbPalette[2]) +
+  geom_point(data = filter(McCann_displacements_bifur, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor), color="white", fill=cbPalette[2], size=6, inherit.aes = FALSE) +
+  geom_point(data = filter(McCann_displacements_bifur, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor), color="white", fill=cbPalette[2], size=3, inherit.aes = FALSE) +
+  scale_shape_manual(values=c(24,21))  +
+  scale_linetype_manual(values=c("dotted","solid")) +
+  ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*": max and min)"))) +
+  xlab(expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")")))) +
+  guides(fill = guide_legend(override.aes=list(shape=21))) +
+  geom_segment(x=0.815, xend=0.815, y=0, yend=2.65, arrow = arrow(length = unit(0.03, "npc")), color="grey", fill="grey") +
+  theme_cowplot(font_size = 18)
+  #geom_vline(xintercept = 0.82, linetype="dotted", color="grey")
+
+save_plot(filename = "bifurcation_plot.pdf", bifurcation_plot, base_width = 8.5, base_height = 6)
+```
+
+```{r bifur_plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Divergent character displacement creates unstable food webs**.
+I simulated the eco-evolutionary dynamics of character displacement assuming a more realistic functional response for consumers.
+The results are qualtitatively similar for consumer dynamics (not shown).
+Note that the lines illustrating the bifurcation assume perfect symmetry in consumer attack rates, while the points are the results of an eco-evolutionary simulation.
+In this simulation, this symmetry in attack rates may be broken, which is why the point of instability does not exactly match the lines.
+I stopped the simulations once consumers reached an evolutionary stable strategy, a constraint of maximum specialization, or created an unstable food web.
+I do not project beyoned this, because the dynamics of mutant invaders becomes unclear when resource distributions are unstable.
+The grey arrow denotes the boundary of local stability."}
+knitr::include_graphics("bifurcation_plot.pdf")
 ```
@@
 Yet, the theory presented here and in past work indicates a predictable asymmetry between initial consumer attack rates. 
-This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, reaches an ESS with a generalist phenotype, followed by the invasion of a second, more specialized, consumer.
+This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, reaches an evolutionary stable strategy (ESS) with a generalist phenotype, followed by the invasion of a second, more specialized, consumer.
 This theoretical scenario has been hypothesized as the sequence of events leading to character displacement in threespine stickleback in postglacial lakes of British Columbia, Canada [@Schluter1992; @Schluter2000].
 
-To test whether our results were robust to this asymmetry, we used the evolved attack rates at the end of the single consumer simulations as the starting values for the competitor in the two consumer simulations.
-We did this for all foraging scenarios and tradeoffs previously examined. 
-We found that our previous inferences were robust to including consumer asymmetry across different tradeoffs and foraging scenarios (Appendix A\ref{fig:MacArthur_Plot_asymm}, A\ref{fig:LS_Plot_asymm}, A\ref{fig:McCann_Plot_asymm}).
+To test whether my results were robust to this asymmetry, I used the evolved attack rates at the end of the single consumer simulations as the starting values for the competitor in the two consumer simulations.
+I did this for all foraging scenarios and tradeoffs previously examined. 
+I found that my previous inferences are robust to including consumer asymmetry across different tradeoffs and foraging scenarios (see Appendix S4).
+<!--A\ref{fig:MacArthur_Plot_asymm}, A\ref{fig:LS_Plot_asymm}, A\ref{fig:McCann_Plot_asymm}).-->
 
@@
 
-```{r MacArthur_Plot_asymm, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}**Effect of character displacement when consumers can forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r MacArthur_Plot_asymm, include=FALSE, eval=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}**Effect of character displacement when consumers can forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
@@
 save_plot(filename="Figure_ECD_LS_asymm.pdf", plot=LS_plots_asymm, base_height = 6, base_width = 6)
-```
-
-```{r LS_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:LS_Plot_asymm}**Effect of character displacement when consumers *cannot* forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
-Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
-This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.
-We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation.
-We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
-knitr::include_graphics("Figure_ECD_LS_asymm.pdf")
+# moved to supplementary information
 ```
@@
 save_plot(filename="Figure_ECD_McCann_asymm.pdf", plot=McCann_plots_asymm, base_height = 6, base_width = 6)
-```
-
-```{r McCann_Plot_asymm, echo=FALSE, fig.cap="\\label{fig:McCann_Plot_asymm}**Effect of character displacement when consumers exhibit a more realistic functional response and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
-Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of consumers.
-This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.
-We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation.
-We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
-knitr::include_graphics("Figure_ECD_McCann_asymm.pdf")
-```
+# moved to supplementary information
+```

### 2019-03-18 manuscript_ECD_model.Rmd
edit Abstract. its less than 150 words now.

@@
 
-Ecological character displacement is an adaptive process, often enhancing coexistence between consumers by reducing competition for shared resources. 
-This competition perspective undermines the fact that consumer-resource interactions ultimately mediate the process of character displacement. 
-As a consequence, the effects of character displacement on consumer-resource interactions and food-web dynamics have been ignored.
+Ecological character displacement is an adaptive process, frequently leading to phenotypic diversification. 
+Despite the fact that this diversification is due to an eco-evolutionary feedback between consumers competing for shared resources, its consequences for food-web dynamics have not been examined.
 Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumers affects resource abundances and the resilience of food webs to perturbations.
-I studied these effects under different assumptions of evolutionary tradeoffs, consumer foraging behavior, and resource distributions.
-I found that character displacement always strengthened consumer-resource interactions whenever consumers competed for resources that occured in distinct habitats.
+I found that character displacement always strengthened consumer-resource interactions whenever consumers competed for resources that occurred in different habitats.
 This increase in interaction strength resulted in lower resource abundances and less resilient food webs.
 This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
-While character displacement did not always increase the strength of consumer-resource interactions, these contingent effects only occurred under the simplest and least realistic foraging scenario. 
-Taken together, my results show that the adaptive process of character displacement often comes at the ecological cost of decreasing food-web resilience. 
-They also highlight that food-web and competition theory can give complimentary insights to the ecological consequences of adaptation.
+Taken together, my results show that the adaptive process of character displacement often comes with the ecological cost of decreasing food-web resilience.
 

### 2019-03-18 manuscript_ECD_model.Rmd
edited intro and added brief statement summarising my results at the end to align with Ecology Letters style

@@
 
-Ecological character displacement is an important adaptive process in generating biodiversity [@Schluter2000; @Pfennig2010; but see @Stuart2013]. 
+Ecological character displacement is an important adaptive process in generating biodiversity [@Schluter2000; @Pfennig2010]. 
 This process results from "phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000]. 
@@
 @Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has examined which scenarios lead to phenotypic divergence or convergence of competing consumers.
-The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of abundance dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985; \ref{fig:fig_1}].
-This adaptive process is not simply a response to statitc differences in resource distributions, but creates an eco-evolutionary feedback that drives further differentiation. 
+The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985].
+This adaptive process is not simply a response to static differences in resource distributions, but creates an eco-evolutionary feedback that drives further differentiation. 
 This crucial insight was made by theoretical models that explicitly included resource dynamics as a mediator of competition in driving evolutionary change [@Lawlor1976; @Abrams1986; @Taper1985].
@@
 Although models that included resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics has received surprisingly little attention. 
-This is likely because the ecological feedback has been primarily studied through the lens of competition theory [@Lawlor1976; @Germain2018; @Bassar2017; @McPeek2019]. 
+This is likely because the ecological feedback has been primarily studied through the lens of coexistence theory [@Lawlor1976; @Germain2018; @Bassar2017; @McPeek2019]. 
 For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition.
@@
 Here, I address this knowledge gap by studying a mathematical model that examines how ecological character displacement affects consumer-resource dynamics in a food-web context. 
-I address two questions: (1) How does ecological character displacement affect resource abundances?
-(2) How does character displacement affect food-web stability?
-To test the generality of these effects, I explored different assumptions of ecological foraging scenarios and evolutionary trade-offs in consumer foraging traits.
+I address two questions: (1) How does ecological character displacement affect resource abundances? (2) How does character displacement affect food-web stability? 
+To test the generality of these effects, I explored different assumptions of ecological foraging scenarios and evolutionary trade-offs in consumer foraging traits (Fig. \ref{fig:fig_1}). 
+I found that the adaptive process of character displacement often comes with an ecological cost; resulting in food webs with lower resource availability and that are less resilient to perturbations.
 

### 2019-03-18 manuscript_ECD_model.Rmd
add many changes to methods, results, and discussion. added new, perhaps controversial, conclusion paragraph.

@@
 
-Ecological character displacement is an adaptive process, frequently leading to phenotypic diversification. 
+Ecological character displacement is an adaptive process that generally increases phenotypic diversity. 
 Despite the fact that this diversification is due to an eco-evolutionary feedback between consumers competing for shared resources, its consequences for food-web dynamics have not been examined.
@@
 
-where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ji}$ represents the conversion efficiency of resource $i$ into consumer $j$, and $m_j$ represents the mortality rate of consumer $j$.
-$F_{ji}(R_i)$ represents consumer $j$'s feeding rate (i.e.
-functional response) on resource $i$.
+where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ij}$ represents the conversion efficiency of resource $j$ into consumer $i$, and $m_i$ represents the mortality rate of consumer $i$.
+$F_{ij}(R_j)$ represents consumer $i$'s feeding rate on resource $j$ (i.e.
+functional response).
 This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g.
@@
 
-$$F_{ji}(R_i)=a_{ji}R_i$$
-
-where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972]. 
-This first scenario is the starting point for many models of resource competition; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005].
+$$F_{ii}(R_i)=a_{ii}R_i$$
+
+where $a_{ii}$ is the attack rate of consumer $i$ on resource $i$. 
+This first scenario is the starting point for many models of resource competition [@MacArthur1972]; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005].
 A simple extension of the above scenario takes the form:
 
-$$F_{ji}(R_i)=w_{ji}a_{ji}R_i$$
-
-where $w_{ji}$ represents the proportion of time consumer $j$ spends foraging in a habitat where only resource $j$ is found (i.e. habitat preference). 
-Note that since $w_{ji}$ is a proportion that $w_{jj}=1-w_{ji}$. 
-Finally, it is well known that consumer feeding rates often saturate at high resource abundances [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011] and they do not usually spend a fixed proportion of time in a particular habitat [@McCann2005].
+$$F_{ii}(R_i)=w_{ii}a_{ii}R_i$$
+
+where $w_{ii}$ represents the proportion of time consumer $i$ spends foraging in a habitat where only resource $i$ is found (i.e. habitat preference). 
+Note that since $w_{ii}$ is a proportion that $w_{ij}=1-w_{ii}$. 
+Finally, it is well known that consumer feeding rates often saturate at high resource abundances [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011] and that consumers do not usually spend a fixed proportion of time in a particular habitat [@McCann2005].
 To account for these biological realities, I consider a more realistic functional response [derived by @McCann2005]:
 
-$$F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}hW_{ii}R_i+a_{ij}hW_{ij}R_j}$$
-
-where consumer $i$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling time $h$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}$.
-
-Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach [@Otto2007], with the general result being divergent character displacement [@Lawlor1976; @Abrams1986].
+$$F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}h_{ii}W_{ii}R_i+a_{ij}h_{ij}W_{ij}R_j}$$
+
+where consumer $i$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling times $h_{ii}$, $h_{ij}$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ii}=\frac{w_{ii}R_i}{w_{ii}R_i+(1-w_{ij})R_j}$.
+
+Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach, with the general result being divergent character displacement [@Lawlor1976; @Abrams1986].
 I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
@@
 To do this, I analyzed differences in resource abundances and food-web stability at equilibrium. 
-An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$).
-At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource abundances and performed a linear stability analysis [@Otto2007]. 
+An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and each consumer ($\hat C_1,\hat C_2$).
+At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource abundances and performed a local stability analysis [@Otto2007]. 
 This stability analysis derives the dominant eigenvalue, $\lambda$, which determines whether (and how readily) the food web will return to equilibrium after a small perturbation.
@@
 To do this, I simplified the model by assuming that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $h=h_{11}=h_{12}=h_{21}=h_{22}$; $m=m_1=m_2$), except that consumer attack rates and their habitat preferences (if present) are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$; $w_{ii}=w_{11}=w_{22}$).
-Controlling for other sources of variability allowed me to isolate the general effects of divergent character displacement on food-web dynamics.
-
-To gain insight to the eco-evolutionary feedback generated by divergent character displacement, I conducted simulations using an Adaptive Dynamics approach. 
+Controlling for other sources of variability allowed me to isolate the general effects of divergent character displacement.
+All mathematical derivations were conducted in @Mathematica.
+
+To gain insight to the eco-evolutionary feedback generated by character displacement, I conducted simulations using an Adaptive Dynamics approach. 
 <!--
@@
 -->
-Specifically, at each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate on one resource by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability.
+Specifically, at each evolutionary time step, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability.
 The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006].
-This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$, green line in Fig.
-\ref{fig:MacArthur_Plot}A) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$, orange line Fig.
-\ref{fig:MacArthur_Plot}A).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig. \ref{fig:MacArthur_Plot}A). 
+This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher than the extremes (when $n>1$, green line in Fig.
+\ref{fig:fig_1}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line Fig.
+\ref{fig:fig_1}).
+When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig. \ref{fig:fig_1}). 
 Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
@@
 $\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
+Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and *w_{ii}* = 0.6. 
+I set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*. 
+I set initial consumer and resource abundances as: *R~1~* = *R~2~* = 2; *C~1~* = *C~2~* = 1.
+All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model).
 
@@
 
-## Consumers can forage for both resources simultaneously
+## Resources occur in same habitat
 
@@
 
-In this first scenario, the abundance of resources at equilibrium are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), and are determined by the following equation (derived in Appendix S2):
+In this first scenario, the abundance of resources at equilibrium are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), and are determined by the following equation (derived in Appendix S**?**):
 
@@
 A key determinant of resource abundance in this model is the consumer's total attack rate, $a_{ii}+a_{ij}$.
-Therefore, the ecological consequences of character displacement depend on how the tradeoff influences the evolution of consumer attack rates.
-
-My simulations show that the shape of the attack-rate tradeoff qualitatively affects the relationship between character displacement and resource abundance (Fig. \ref{fig:MacArthur_Plot}). 
+Therefore, the effect of character displacement on food-web dynamics depends on how the tradeoff influences the evolution of consumer attack rates.
+
+My simulations show that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource abundances (Fig. \ref{fig:plot_fig3}). 
 For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (Fig.
-\ref{fig:MacArthur_Plot}B) and character displacement has no effect on resource abundances (Fig.
-\ref{fig:MacArthur_Plot}C).
-If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (Fig. \ref{fig:MacArthur_Plot}C). 
-This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (Fig. \ref{fig:MacArthur_Plot}B). 
-When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (Fig. \ref{fig:MacArthur_Plot}B,C). 
+\ref{fig:plot_fig3}) and character displacement has no effect on resource abundances (Fig.
+\ref{fig:plot_fig3}).
+If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (Fig. \ref{fig:plot_fig3}). 
+This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (Fig. \ref{fig:plot_fig3}). 
+When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (Fig. \ref{fig:plot_fig3}). 
 Although the equation I derived for resource abundances was for the scenario where both consumers and both resources were present, it accurately predicts the abundance of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
-\ref{fig:MacArthur_Plot}C).
+\ref{fig:plot_fig3}).
 This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in Fig.
-\ref{fig:MacArthur_Plot}B), resulting in equivalent resource abundances.
+\ref{fig:MacArthur_Plot}), resulting in equivalent resource abundances.
 
 The effect of character displacement on resources corresponds to its impact on food-web stability. 
-For example, when character displacement decreases resource abundances (orange points in \ref{fig:MacArthur_Plot}C), there is a decrease in food-web stability (\ref{fig:MacArthur_Plot}D). 
-Similarly, if character displacement does not alter resource abundances (blue and green points in \ref{fig:MacArthur_Plot}C), then there is no change in food-web stability (\ref{fig:MacArthur_Plot}D).
+For example, when character displacement decreases resource abundances (orange points in \ref{fig:plot_fig3}), there is a decrease in food-web stability (\ref{fig:plot_fig3}). 
+Similarly, if character displacement does not alter resource abundances (blue and green points in \ref{fig:plot_fig3}), then there is no change in food-web stability (\ref{fig:plot_fig3}).
 
@@
 
-## Consumers *cannot* forage for both resources simulanteously
+## Resources occur in different habitats
 <!--
@@
 -->
-In this second scenario, I again see that resource abundances are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), but are now determined by the following equation (derived in Appendix S3):
-
-$$\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e}$$
+In this second scenario, I again see that resource abundances are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), but are now determined by the following equation (derived in Appendix S**?**):
+
+$$\hat{R}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e}$$
 
 This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g.
-$w>0.5$ and $a_{ii}>a_{ij}$), then the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase, regardless of the tradeoff (Fig.
-\ref{fig:LS_Plot}A).
-Thus, divergent character displacement always results in resource suppression (Fig. \ref{fig:LS_Plot}B). 
+$w_{ii}>0.5$ and $a_{ii}>a_{ij}$), then the effective attack rate of consumers ($wa_{ii}+(1-w_{ii})a_{ij}$) will always increase, regardless of the tradeoff (Fig.
+\ref{fig:plot_fig3}).
+Thus, divergent character displacement always results in resource suppression (Fig. \ref{fig:plot_fig3}). 
 Note that the shape of the tradeoff can modify the effect of character displacement. 
-This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached its ESS (triangles in Fig.
-\ref{fig:LS_Plot}B).
+This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached an evolutionary stable strategy (ESS, triangles in Fig.
+\ref{fig:plot_fig3}).
 In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
-\ref{fig:LS_Plot}B), because character displacement tends to reach a constraint of complete specialization.
+\ref{fig:plot_fig3}), because character displacement tends to reach a constraint of complete specialization.
 It is worth noting that resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
-\ref{fig:LS_Plot}B).
+\ref{fig:plot_fig3}).
 This is because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in Fig.
-\ref{fig:LS_Plot}B).
-
-As we saw previously, the effect of character displacement on resource abundances qualitatively corresponds to its effect on food-web stability (Fig. \ref{fig:LS_Plot}C). 
+\ref{fig:plot_fig3}).
+
+As seen previously, the effect of character displacement on resource abundances qualitatively corresponds to its effect on food-web stability (Fig. \ref{fig:plot_fig3}C). 
 Specifically, divergent character displacement decreases food-web stability, regardless of the tradeoff in attack rates.
 This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (Fig.
-\ref{fig:LS_Plot}C).
-For example, when the tradeoff is concave up (orange), then the four-species (small circle) food web actually starts off as more stable than the three-species one (small triangle); however, this pattern switches once they reach their evolutionary stable strategies (large points).
+\ref{fig:plot_fig3}C).
+For example, when the tradeoff is concave up (orange), then the four-species (small circle) food web actually starts off as more stable than the three-species one (small triangle); however, this pattern switches by the end of the eco-evolutionary simulation (large points).
 
@@
 -->
-This is because resource abundances at equilibrium are governed by a similar dynamic (derived in Appendix S1):
+This is because resource abundances at equilibrium are governed by a similar dynamic (derived in Appendix S**?**):
 <!--
@@
 And since evolution favors consumer divergence onto their preferred resources (see Appendix S1), the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase (Fig.
-\ref{fig:McCann_Plot}A), resulting in lower resource abundances (Fig.
-\ref{fig:McCann_Plot}B) and decreased food-web stability (Fig.
-\ref{fig:McCann_Plot}C).
+\ref{fig:McCann_Plot}A), resulting in lower resource abundances and decreased food-web stability (Appendix S?).
 
@@
 In this more realistic model, however, whether the food web is locally stable depends on consumer and resource parameters. 
-Specifically, I found that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derived using Routh-Hurwitz criteria in Appendix S1):
-
-$$wa_{ii}+(1-w)a_{ij} > \frac{e+hm}{hK(e-hm)}$$
+Specifically, I found that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derived using Routh-Hurwitz criteria in Appendix S**?**):
+
+$$w_{ii}a_{ii}+(1-w_{ii})a_{ij} > \frac{e+hm}{hK(e-hm)}$$
 
@@
 \ref{fig:LS_Plot}A).-->
-Thus, character displacement always pushes the food web toward an unstable stucture in this more realistic foraging scenario.
+Thus, character displacement always pushes the food web toward an unstable stucture in this more realistic foraging scenario (Fig. \ref{fig:bifur_plot}).
 
@@
 The results are qualtitatively similar for consumer dynamics (not shown).
-Note that the lines illustrating the bifurcation assume perfect symmetry in consumer attack rates, while the points are the results of an eco-evolutionary simulation.
+Note that the lines illustrate the ecological consequences across the range of consumer specialization, while the points are the results of an eco-evolutionary simulation.
 In this simulation, this symmetry in attack rates may be broken, which is why the point of instability does not exactly match the lines.
@@
 The previous analytical results and simulations make a strong assumption that competing consumers start off as perfect mirror images of each other in their attack rates (i.e. symmetry). 
-Yet, the theory presented here and in past work indicates a predictable asymmetry between initial consumer attack rates. 
-This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, reaches an evolutionary stable strategy (ESS) with a generalist phenotype, followed by the invasion of a second, more specialized, consumer.
-This theoretical scenario has been hypothesized as the sequence of events leading to character displacement in threespine stickleback in postglacial lakes of British Columbia, Canada [@Schluter1992; @Schluter2000].
+Yet, theory indicates a predictable asymmetry between initial consumer attack rates. 
+This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, reaches an ESS as a generalist, followed by the invasion of a second, more specialized, consumer.
+This theoretical scenario has been hypothesized as the sequence of events leading to character displacement in threespine stickleback in small coastal lakes of British Columbia [@Schluter1992; @Schluter2000].
 
@@
 
+## Resource abundances
+
 One of the criteria used to demonstrate character displacement is that "sites of sympatry [two consumers] and allopatry [one consumer] should not differ greatly in food, ..." [@Schluter1992].
 In contrast, my results suggest that ecological character displacement causes predictable differences in resource abundances.
-In fact, the ecological and evolutionary scenarios that favored the greatest displacement always decreased the abundance of resources. <!-- build paragraph around this topic sentence -->
+In fact, the ecological and evolutionary scenarios that favored the greatest character divergence always decreased the abundance of resources. <!-- build paragraph around this topic sentence -->
 For example, the magnitude of character displacement was highest when mobile consumers competed for resources that occur in different habitats compared to within the same area.
@@
 Interestingly, a disproporationate number of the documented cases of character displacement involve carnivores [@Schluter2000] that are larger, and likely more mobile, than their resources [@McCann2005], suggesting my predictions may apply to many cases of ecological character displacement.
-Similarly, the evolutionary tradeoff that favored character displacement (concave up; orange lines) resulted in lower resource densities across all foraging scenarios.
-Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the foraging rate of consumers [@Schluter1985; @Arnegard2014].
+
+Similarly, the evolutionary tradeoff that favored character divergence resulted in lower resource densities across all foraging scenarios.
+Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the net foraging rate of consumers [@Schluter1985; @Arnegard2014].
 While it is theoretically possible that character displacement does not alter resource abundances, this was limited to the simplest, and arguably least realistic, foraging scenario and under tradeoffs that did not favor large displacements, and thus less likely to detect in nature.
@@
 
-My most striking result was that divergent character displacement, when examined through the lens of food-web theory, made food webs less resilient to perturbations.  
-In fact, under the most realistic foraging scenario, character displacement can even result in an unstable food web. 
+## Food-web Stability
+
+My most striking result was that character divergence, when examined through the lens of food-web theory, made food webs less resilient to perturbations. 
+In fact, under the most realistic foraging scenario, character divergence can even result in an unstable food web. 
 The mechanism underlying this destabilization is quite general.
-Character displacement generally increases the strength of consumer-resource interactions, but does not alter the strength of intraspecific interactions.
-This relative increase in interspecific interactions, combined with the natural oscillatory tendency of consumer-resource interactions [@Murdoch2003], creates a food-web structure that is less resilient to perturbations [@Chesson2008; @Rip2011; @McCann2011].
-
-Interestingly, the ecological conditions that favor divergent character displacement are those that are already the least resilient to perturbations.
-For example, @McPeek2019 showed that character displacement is favored in food webs that are either highly productive, easy to find and capture resources, or under weak abiotic stress.
-In my model, this corresponds to higher values of $K$ (productivity) or $A$ (capacity to find and capturing resources), or lower values of $m$ (abiotic stress).
+Character divergence generally increases the strength of consumer-resource interactions, but does not alter the strength of intraspecific interactions.
+This relative increase in interspecific interactions, combined with the natural oscillatory tendency of consumer-resource dynamics [@Lotka1925; @Volterra1926], creates a food-web structure that is less resilient to perturbations [@Chesson2008; @Rip2011; @McCann2011].
+
+Interestingly, the ecological conditions that favor character divergence are those that are already the least resilient to perturbations.
+For example, @McPeek2019 showed that character divergence is favored in food webs that are either highly productive, easy to find and capture resources, or under weak abiotic stress.
+This corresponds to higher values of $K$ (productivity) or $A$ (capacity to find and capturing resources), or lower values of $m$ (abiotic stress).
 Each of these corresponding changes decrease food-web resilience, as they increase the strength of consumer-resource interactions relative to intraspecific interactions.
@@
 Similarly, higher feeding rates or lower consumer mortality both increase the relative strength of consumer-resource interactions, which predictably destabilizes food webs [@Rip2011; @McCann2011].
-This suggests that the most dramatic examples of character displacement will not only occur in, but also cause, the least stable food-web structures.
-
-A handful of empirical patterns support the hypothesis that character displacement decreases food-web resilience. 
-As an example, a single species of threespine stickleback lives in hundreds of postglacial lakes of British Columbia, but the species pair <!--, where character displacement has happened,-->are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
+This suggests that the most dramatic examples of character divergence will not only occur in, but also cause, the least stable food-web structures.
+
+A handful of empirical patterns support the hypothesis that character divergence decreases food-web resilience. 
+For example, a single species of threespine stickleback lives in hundreds of small coastal lakes of British Columbia, but the species pair, where character divergence has resulted in specialized limnetic and benthic species, are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
 Perhaps many lakes had a species pair in the past, but have lost a species due to a less resilient food-web structure [@Borrelli2015a; @Borrelli2015b].
@@
 
-My results contrast, but do not necessarily contradict, the notion from competition theory that character displacement contributes to species coexistence [@Lawlor1976]. 
-Rather than studying resilience, competition theory usually studies the mutual ability of consumers with different phenotypes to invade when rare [mutual invasibility, @Chesson2000].
+My results contrast, but do not necessarily contradict, the notion from coexistence theory that character displacement contributes to species coexistence [@Lawlor1976]. 
+Rather than studying resilience, coexistence theory usually studies the mutual ability of consumers with different phenotypes to invade when rare [mutual invasibility, @Chesson2000].
 In the context of character displacement, a shortcoming of this mutual invasibility measure is that it does not allow a comparison between food webs with and without a competiting consumer. 
@@
 
-The Adaptive Dynamics approach I used here has its strengths and weaknesses.
-A strength of this approach is that it enabled me to gain analytical insight to the effects of character displacement on food-web dynamics in a more realistic foraging scenario.
+## Caveats
+
+I studied this eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
+A strength of this approach is that it enabled me to gain analytical insight to the effects of character divergence on food-web dynamics in a more realistic foraging scenario.
 This is much less tractable in quantitative genetic [@Taper1985; @McPeek2017] or explicit genetic [@Doebeli1996] models of character displacement, which is why the foraging scenarios previously examined have been limited [but see @McPeek2017].
@@
 That being said, my theoretical predictions are likely robust to these assumptions. 
-This is because models that explicitly include resource dynamics inevitably show that resource competition results in divergent character displacement, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
-A quantitative genetic model may certainly show differences in the magnitude of character displacement, but this should not affect the general relationships I observed.  
-<!-- In fact, recent work by @McPeek2017 has shown that divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics. -->
-I want to note though, that my conclusions only apply to food webs with biotic resources that are nutritionally substitutable. 
-It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
+This is because models that explicitly include resource dynamics inevitably show that resource competition results in character divergence, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
+A quantitative genetic model may certainly show differences in the magnitude of character divergence, but this should not affect the consequences for food-web dynamics.
+<!-- In fact, recent work by @McPeek2017 has shown that divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
+-->It is important to note that my conclusions only apply to food webs with biotic resources that are nutritionally substitutable.
+It would be interesting to extend these current analyses to non-substitutable resources where chracter convergence is expected [@Abrams1987; @Fox2008].
+
+## Conclusions
+
+Here, I show that an adaptive process that generates phenotypic diversity generally makes that diversity more susceptible to future extinctions.
+This destabilizing effect emerges from an eco-evolutionary feedback involving direct and indirect interactions between species in a food-web context.
+<!--This highlights how the process of adaptation, when considered in a food-web context, may generate feedbacks that destabilize the interconnected parts of the system.
+This suggests that observed patterns of biodiversity may not simply be a result of evolutionary constraints, but also ecological constraints
+To what extent are current patterns in biodiversity a result of ecological constraints on the persistence of evolved phenotypes? -->
+This result contrasts with the current notion that patterns of phenotypic diversity are solely the result of evolutionary constraints imposed by mutation, natural selection, gene flow, and genetic drift.
+In particular, my result supports the recent suggestion that food-web stability can impose an ecological constraint on phenotypic diversity that is agnostic to these evolutionary processes [@Borrelli2015b].
+I expect that identifying when and where this ecological constraint arises will yield novel insight to the patterns of biodiversity we see in nature.
+<!--Identifying when and where we expect this ecological constraint to arise may yield novel insight to the patterns of biodiversity we see in nature. -->
 

### 2019-03-18 manuscript_ECD_model.Rmd
added edits to main text and supplementary information.

@@
 
-Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach, with the general result being divergent character displacement [@Lawlor1976; @Abrams1986].
-I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
-I also used an Adaptive Dynamics approach to analyze character displacement in the third foraging scenario, and I too observed divergent character displacement (detailed analysis in Appendix S1 of Supplementary Information).
+Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach, with the general result being character divergence [@Lawlor1976; @Abrams1986].
+I say consumers have undergone character divergence if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
+I also used an Adaptive Dynamics approach to analyze character displacement in the third foraging scenario, and I too observed character divergence (detailed analysis in Appendix S1 of Supplementary Information).
 
@@
 
-Given that divergent character displacement was an expected result, I focus here on its consequences for food-web dynamics across these foraging scenarios.
+Given that character divergence was an expected result, I focus here on its consequences for food-web dynamics across these foraging scenarios.
 To do this, I analyzed differences in resource abundances and food-web stability at equilibrium. 
@@
 To do this, I simplified the model by assuming that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $h=h_{11}=h_{12}=h_{21}=h_{22}$; $m=m_1=m_2$), except that consumer attack rates and their habitat preferences (if present) are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$; $w_{ii}=w_{11}=w_{22}$).
-Controlling for other sources of variability allowed me to isolate the general effects of divergent character displacement.
+Controlling for other sources of variability allowed me to isolate the general effects of character divergence.
 All mathematical derivations were conducted in @Mathematica.
@@
 \ref{fig:plot_fig3}).
-Thus, divergent character displacement always results in resource suppression (Fig. \ref{fig:plot_fig3}). 
+Thus, character divergence always results in resource suppression (Fig. \ref{fig:plot_fig3}). 
 Note that the shape of the tradeoff can modify the effect of character displacement. 
@@
 As seen previously, the effect of character displacement on resource abundances qualitatively corresponds to its effect on food-web stability (Fig. \ref{fig:plot_fig3}C). 
-Specifically, divergent character displacement decreases food-web stability, regardless of the tradeoff in attack rates.
+Specifically, character divergence decreases food-web stability, regardless of the tradeoff in attack rates.
 This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (Fig.
@@
 
-In the third foraging scenario, I observed the same general effect of divergent character displacement as the previous scenario (resources in different habitats, but linear functional response).
+In the third foraging scenario, I observed the same general effect of character divergence as the previous scenario (resources in different habitats, but linear functional response).
 <!--Divergent character displacement in this foraging scenario has the same general effect on food-web dynamics as the previous scenario (resources in different habitats, but linear functional response).
@@
 Stability_McCann <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color=factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend = FALSE) +
-  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen)) + 
+  geom_line(aes(color=factor(n))) +
+  #geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend = FALSE) +
+  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  geom_point(data = McCann_displacements, aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor, size=Time), color="white", inherit.aes = FALSE) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
   scale_shape_manual(values=c(24,21))  +
+  scale_size_manual(values=c(1.5,3)) +
   scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
@@
 ```{r McCann Model, include=FALSE}
+McCann_legend <- get_legend(Stability_McCann)
+
 plot_McCann_model <- plot_grid(
-  McCann_attack    + theme(legend.position = "none"), 
-  McCann_resources    + theme(legend.position = "none"),
-  Stability_McCann    + theme(legend.position = "none"), ncol=2, align='hv', labels = "AUTO")
+  McCann_attack    + theme(legend.position = "none") + xlab(""), 
+  McCann_resources    + theme(legend.position = "none") + xlab(""),
+  Stability_McCann    + theme(legend.position = "none") + xlab(""), ncol=2, align='hv') # , labels = "AUTO"
+
+plot_McCann_model <- add_sub(plot_McCann_model, label = expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")"))))
+
+#plot_McCann_model <- ggdraw(plot_McCann_model) + draw_grob(McCann_legend, x = 0.7, y=-0.2)
 
@@
 
-```{r bifur_plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Divergent character displacement creates unstable food webs**.
+```{r bifur_plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Character divergence creates unstable food webs**.
 I simulated the eco-evolutionary dynamics of character displacement assuming a more realistic functional response for consumers.
@@
 ```
+
+```{r all asymmetry}
+
+all_asymm_plots <- plot_grid(
+  Mac_attack_asymm + ggtitle("Scenario I\n") + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[ii]+a[ij]*")"))), #, plot.title=element_text(size=8)
+  LS_attack_asymm  + ggtitle("Scenario II\n") + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(bquote(w[ii]*a[ii]+"("*1-w[ii]*")"*a[ij]*")")), #, plot.title=element_text(size=8)
+  McCann_attack_asymm + ggtitle("Scenario III")  + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(bquote(w[ii]*a[ii]+"("*1-w[ii]*")"*a[ij]*")")), #, plot.title=element_text(size=8), plot.subtitle=element_text(size=8)
+  Mac_resources_asymm   + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))),
+  LS_resources_asymm   + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(""),
+  McCann_resources_asymm   + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(""),
+  Stability_Mac_asymm   + theme(legend.position = "none") + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))),
+  Stability_LS_asymm   + theme(legend.position = "none") + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(""),
+  Stability_McCann_asymm   + theme(legend.position = "none") + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(""),
+  ncol=3, align='v'#, rel_widths = c(1,0.9,0.9)
+)
+
+all_asymm_plots <- add_sub(all_asymm_plots, label = expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")"))))
+
+save_plot(filename="asymm_figs.pdf", plot=all_asymm_plots, base_height = 11, base_width = 8.5)
+```

### 2019-03-18 supplementary_ECD_model.Rmd
added edits to main text and supplementary information.

---
#title: |
#   | Predicting the effects of character displacement on food-web dynamics
#   | Food-web dynamics under ecological character displacement
#author: 
#- Matthew A. Barbour$^{1,2,\ast}$
# potentially use for 'date' for affiliations, as recommended by https://github.com/rstudio/bookdown/issues/348 for cross-format compatibility, with | to create new sections.
#date: | 
#    | 1. Affiliation
output:
  pdf_document:
    keep_tex: true
    toc: false
    number_sections: false
    df_print: paged
    fig_caption: true
    citation_package: natbib 
    includes:  
      in_header: Am-Nat-preamble-latex.tex
bibliography: references
biblio-style: amnat
fontsize: 11pt
fontfamily: mathpazo
documentclass: article
linkcolor: black
urlcolor: black
citecolor: black
---

<!-- Begin title page formatting -->

\vspace*{0.1cm} 
\begin{center} \LARGE Appendix S**?**: \\ Ecological character displacement destabilizes food webs \end{center}
<!--\begin{center} \LARGE Food-web dynamics under ecological character displacement \end{center}-->

\bigskip

<!-- Remove author info for double-blind peer review; Keep for bioRxiv submission -->
\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}

\bigskip

\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, 8057 Zurich, Switzerland;

\noindent 2. University of British Columbia, Department of Zoology, Vancouver, BC V6T 1Z4, Canada;

$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch

\bigskip

*Elements*: Figure S1 and figure S2.

<!-- End title page formatting -->

\linenumbers{}
\modulolinenumbers[3]

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

.

```{r McCann_Plot, echo=FALSE, fig.pos="H", fig.cap="**Effect of character displacement when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85). Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in stability panel). This figure illustrates how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community. This always resulted in a suppression of resource abundances and concomitant decrease in food-web stability. Initial parameter and state variables are given in the main text."}
knitr::include_graphics("Figure_ECD_McCann.pdf")
```

\newpage

```{r All asymmetric figs, echo=FALSE, fig.pos="H", fig.cap=""}
knitr::include_graphics("asymm_figs.pdf")
```

Figure 2: **Robustness to consumer asymmetry**. Scenarios correspond to: (I) resources occur in the same habitat and consumer(s) exhibits a linear functional response; (II) resources occur in different habitats and consumer(s) exhibits a linear functional response; (III) resources occur in different habitats and consumer(s) exhibits a more realistic functional response (details in main text). Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85). Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in stability panels). The similarity between this figure and fig. S1 and fig. 2 (main text) show that adding an asymmetry in consumer attack rates does not alter the conclusions reported in the main text. If anything, it is more likely an asymmetry will push the system toward the boundary of stability (note position of large circles in stability panel of Scenario I). Details of the consumer asymmetry simulation are given in the main text.

```{r MacArthur_Plot_asymm, echo=FALSE, eval=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}**Effect of character displacement when consumers can forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85). Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points. Panel (A) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community. Panels (B) and (C) show how different tradeoffs indirectly affect total resource abundances (B) and food-web stability (C). These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2. We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*. To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation. We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
knitr::include_graphics("Figure_ECD_MacArthur_asymm.pdf")
```

```{r LS_Plot_asymm, eval=FALSE, echo=FALSE, fig.cap="\\label{fig:LS_Plot_asymm}**Effect of character displacement when consumers *cannot* forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85). Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points. Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community. This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C). These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4. We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*. To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation. We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
knitr::include_graphics("Figure_ECD_LS_asymm.pdf")
```

```{r McCann_Plot_asymm, eval=FALSE, echo=FALSE, fig.cap="\\label{fig:McCann_Plot_asymm}**Effect of character displacement when consumers exhibit a more realistic functional response and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85). Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points. Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of consumers. This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C). These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4. We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*. To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation. We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
knitr::include_graphics("Figure_ECD_McCann_asymm.pdf")
```

### 2019-03-20 manuscript_ECD_model.Rmd
removed stickleback figures. added separate figs for foraging scenarios and tradeoffs.

@@
 I address two questions: (1) How does ecological character displacement affect resource abundances? (2) How does character displacement affect food-web stability? 
-To test the generality of these effects, I explored different assumptions of ecological foraging scenarios and evolutionary trade-offs in consumer foraging traits (Fig. \ref{fig:fig_1}). 
-I found that the adaptive process of character displacement often comes with an ecological cost; resulting in food webs with lower resource availability and that are less resilient to perturbations.
-
-```{r fig_1, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:fig_1}**Divergent character displacement.** A textbook example of divergent character displacement is threespine stickleback in small coastal lakes of British Columbia.
-Sticklback consume two main resources, zooplankton and benthic invertebrates.
-In isolation, stickleback evolve a generalist phenotype that can take advantage of both resources.
-In the presence of a competitor, however, stickleback evolve specialized phenotypes that enable them to capture different resources more efficiently.
-A large body of theoretical work has shown that resource competition generally drives the adaptive divergence of consumers onto different resources.
-Currently though, we lack theoretical predictions for how this adaptive divergence affects consumer-resource interactions and food-web dynamics."}
-knitr::include_graphics("ECD_model_fig_1.pdf")
-```
+To test the generality of these effects, I explored different assumptions of ecological foraging scenarios (Fig.
+\ref{fig:foraging_scenarios}) and evolutionary trade-offs in consumer foraging traits (Fig.
+\ref{fig:tradeoff}).
+I found that the adaptive process of character displacement often comes with an ecological cost; resulting in food webs with lower resource availability and less resilient to perturbations.
 
@@
 seed size, see @Taper1985 for an example).
+Importantly, inferences about ecological character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992]; therefore, we arbitrarily removed $C_2$ from the system to compare and infer the effects of character displacement on food-web dynamics.
 
@@
 
-I studied three different foraging scenarios. 
+I studied three different foraging scenarios (Fig. \ref{fig:foraging_scenarios}). 
 In the first, I assume that consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that:
@@
 
-```{r fig_2, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:fig_2}**Foraging scenarios and evolutionary tradeoffs.** I examined the effect of ecological character displacement on food-web dynamics under different assumptions of ecological foraging scenarios and evolutionary trade-offs in consumer foraging traits.
-Specifically, I examined whether these effects depended on whether resources occurred in the same habitat (top left) or in different habitats (top right).
-I also included a more realistic functional responses.
-In each of these foraging scenarios, I examined the effect of different evolutionary tradeoffs in attack rates."}
-knitr::include_graphics("ECD_model_fig_2.pdf")
+```{r foraging_scenarios, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:foraging_scenarios}**Ecological foraging scenarios.** I examined whether the effect of ecological character displacement on food-web dynamics depended on whether consumers competed for resources that occurred in the same vs.
+different habitats.
+In both scenarios, I assumed consumer feeding rate increased linearly with resource density.
+Later, I relax this assumption and consider a more realistic functional response when resources occurred in different habitats."}
+knitr::include_graphics("ECD_model_foraging_scenarios")
 ```
@@
 Controlling for other sources of variability allowed me to isolate the general effects of character divergence.
-All mathematical derivations were conducted in @Mathematica.
+All mathematical derivations were conducted in @Mathematica and are provided in the Supplementary Information (Appendix S1-3).
 
@@
 This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher than the extremes (when $n>1$, green line in Fig.
-\ref{fig:fig_1}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line Fig.
-\ref{fig:fig_1}).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig. \ref{fig:fig_1}). 
+\ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line Fig.
+\ref{fig:tradeoff}).
+When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig. \ref{fig:tradeoff}). 
 Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
@@
 
-```{r Tradeoff, include=FALSE, fig.pos="H", fig.cap="\\label{fig:Tradeoff}Tradeoff forms for consumer attack rates."}
+```{r plot tradeoff, include=FALSE}
 initial_points <- bind_rows(params_2C_2R_MacArthur_linear, params_2C_2R_MacArthur_concaveDN, params_2C_2R_MacArthur_concaveUP) %>% mutate(n = c(n_linear, n_concaveDN, n_concaveUP))
@@
   ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
-  geom_line(size=1.5) + 
-  geom_point(data = initial_points, aes(x=a11, y=a12), size=5, show.legend=FALSE) +
+  geom_line() + # size=1.5
+  geom_point(data = initial_points, aes(x=a11, y=a12), size=3, show.legend=FALSE) + # size=5
   xlab(expression(bold(a[ii]))) +
   ylab(expression(bold(a[ij]))) +
-  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette) +
-  theme_cowplot(font_size = 30, line_size=2)# +
+  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette) #+
+  #theme_cowplot(font_size = 30, line_size=2)# +
   #theme(axis.title = element_text(size = 30),
@@
 
-save_plot(filename = "attack_tradeoffs.pdf", plot = tradeoff_plot, base_height = 6, base_width = 9)
+save_plot(filename = "attack_tradeoffs.pdf", plot = tradeoff_plot, 
+          base_height = 5, # 6
+          base_width = 8) #9
+```
+
+```{r tradeoff, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: (i) intermediate combinations of attack rates (a~ii~, a~ij~) are higher than the extremes (green line, n > 1); (ii) extreme combinations of attack rates are higher than intermediate investments (orange line, n < 1); and (iii) all combinations of attack rates have the same total attack rate (blue line, n = 1).
+Points corresponding to attack rates at the beginning of the simulation.
+Note that C~i~ was always pre-adapted to R~i~ (a~ii~>a~ij~)."}
+knitr::include_graphics("attack_tradeoffs.pdf")
 ```
@@
 
-Similarly, the evolutionary tradeoff that favored character divergence resulted in lower resource densities across all foraging scenarios.
+Similarly, the evolutionary tradeoff that favored character divergence resulted in lower resources across all foraging scenarios.
 Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the net foraging rate of consumers [@Schluter1985; @Arnegard2014].

### 2019-03-20 manuscript_ECD_model.Rmd
edited figure legends. added stability boundary text to bifurcation plot

@@
 In both scenarios, I assumed consumer feeding rate increased linearly with resource density.
-Later, I relax this assumption and consider a more realistic functional response when resources occurred in different habitats."}
+Later, I relax this assumption and consider a more realistic functional response when resources occurred in different habitats.
+Note that C~i~ was always pre-adapted to R~i~ (a~ii~ > a~ij~)."}
 knitr::include_graphics("ECD_model_foraging_scenarios")
@@
 
-```{r tradeoff, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: (i) intermediate combinations of attack rates (a~ii~, a~ij~) are higher than the extremes (green line, n > 1); (ii) extreme combinations of attack rates are higher than intermediate investments (orange line, n < 1); and (iii) all combinations of attack rates have the same total attack rate (blue line, n = 1).
+```{r tradeoff, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: intermediate combinations of attack rates (a~ii~, a~ij~) are higher than the extremes (green line, n > 1); extreme combinations of attack rates are higher than intermediate investments (orange line, n < 1); and all combinations of attack rates have the same total attack rate (blue line, n = 1).
 Points corresponding to attack rates at the beginning of the simulation.
-Note that C~i~ was always pre-adapted to R~i~ (a~ii~>a~ij~)."}
+Note that C~i~ was always pre-adapted to R~i~ (a~ii~ > a~ij~)."}
 knitr::include_graphics("attack_tradeoffs.pdf")
@@
   scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  geom_hline(yintercept=0, linetype="dotted", color="grey") +
+  geom_hline(yintercept=0, color="grey") +
   guides(fill = guide_legend(override.aes=list(shape=21)))
@@
   geom_line(aes(linetype=Competitor), color=cbPalette[2]) +
-  geom_point(data = filter(McCann_displacements_bifur, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor), color="white", fill=cbPalette[2], size=6, inherit.aes = FALSE) +
-  geom_point(data = filter(McCann_displacements_bifur, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor), color="white", fill=cbPalette[2], size=3, inherit.aes = FALSE) +
+  geom_point(data = McCann_displacements_bifur, aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor, size=Time), color=cbPalette[2], fill=cbPalette[2], inherit.aes = FALSE) +
+  #geom_point(data = filter(McCann_displacements_bifur, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor), color="white", fill=cbPalette[2], size=6, inherit.aes = FALSE) +
+  #geom_point(data = filter(McCann_displacements_bifur, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor), color="white", fill=cbPalette[2], size=3, inherit.aes = FALSE) +
   scale_shape_manual(values=c(24,21))  +
+  scale_size_manual(values=c(3,6)) +
   scale_linetype_manual(values=c("dotted","solid")) +
@@
   xlab(expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")")))) +
-  guides(fill = guide_legend(override.aes=list(shape=21))) +
+  #guides(fill = guide_legend(override.aes=list(shape=21))) +
   geom_segment(x=0.815, xend=0.815, y=0, yend=2.65, arrow = arrow(length = unit(0.03, "npc")), color="grey", fill="grey") +
+  geom_text(x=0.77, y=2, label="Stability\nboundary", color="grey") +
   theme_cowplot(font_size = 18)

### 2019-03-22 manuscript_ECD_model.Rmd
added final edits and sent to Seth, JP, and Matt Osmond for friendly reviews.

@@
 
-$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch, phone: +41 44 635 57 11
+$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch<!--, phone: +41 44 635 57 11-->
 
@@
 
+<!--
 \bigskip
 
-*Manuscript type*: Letter. <!-- consider e-Article, depending on length -->
+*Manuscript type*: Letter.
 
@@
 
-<!-- End title page formatting -->
+End title page formatting -->
 
@@
 Despite the fact that this diversification is due to an eco-evolutionary feedback between consumers competing for shared resources, its consequences for food-web dynamics have not been examined.
-Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumers affects resource abundances and the resilience of food webs to perturbations.
+Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumer attack rates affects resource abundances and the resilience of food webs to perturbations.
 I found that character displacement always strengthened consumer-resource interactions whenever consumers competed for resources that occurred in different habitats.
@@
 Ecological character displacement is an important adaptive process in generating biodiversity [@Schluter2000; @Pfennig2010]. 
-This process results from "phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000]. 
+This process is due to "phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000]. 
 Over the past 40 years, a large body of theoretical [e.g.
@@
 I address two questions: (1) How does ecological character displacement affect resource abundances? (2) How does character displacement affect food-web stability? 
-To test the generality of these effects, I explored different assumptions of ecological foraging scenarios (Fig.
-\ref{fig:foraging_scenarios}) and evolutionary trade-offs in consumer foraging traits (Fig.
+To test the generality of these effects, I explored different ecological foraging scenarios (fig.
+\ref{fig:foraging_scenarios}) and evolutionary tradeoffs in consumer attack rates (fig.
 \ref{fig:tradeoff}).
-I found that the adaptive process of character displacement often comes with an ecological cost; resulting in food webs with lower resource availability and less resilient to perturbations.
+I found that the adaptive process of character displacement often comes with an ecological cost; resulting in food webs with lower resource availability and that are less resilient to perturbations.
 
@@
 
-$$\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2,$$
-$$\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2,$$
-$$\frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1,$$
-$$\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2,$$
+\begin{equation} \label{eq:1}
+  \begin{split}
+     & \frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2 \\
+     & \frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2 \\
+     & \frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1 \\
+     & \frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2 \\
+  \end{split}
+\end{equation}
 
@@
 This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g.
-zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g.
+zooplankton and benthic invertebrates in lakes) rather than if resources are better characterized by a continuous trait distribution (e.g.
 seed size, see @Taper1985 for an example).
-Importantly, inferences about ecological character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992]; therefore, we arbitrarily removed $C_2$ from the system to compare and infer the effects of character displacement on food-web dynamics.
 
@@
 
-I studied three different foraging scenarios (Fig. \ref{fig:foraging_scenarios}). 
+I studied three different foraging scenarios (fig. \ref{fig:foraging_scenarios}). 
 In the first, I assume that consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that:
 
-$$F_{ii}(R_i)=a_{ii}R_i$$
+\begin{equation} \label{eq:2}
+  F_{ii}(R_i)=a_{ii}R_i
+\end{equation}
 
@@
 This first scenario is the starting point for many models of resource competition [@MacArthur1972]; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005].
-A simple extension of the above scenario takes the form:
-
-$$F_{ii}(R_i)=w_{ii}a_{ii}R_i$$
+A simple extension that accounts for this spatial context takes the form:
+
+\begin{equation} \label{eq:3}
+  F_{ii}(R_i)=w_{ii}a_{ii}R_i
+\end{equation}
 
@@
 To account for these biological realities, I consider a more realistic functional response [derived by @McCann2005]:     
-
-$$F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}h_{ii}W_{ii}R_i+a_{ij}h_{ij}W_{ij}R_j}$$
+\begin{equation} \label{eq:4}
+  F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}h_{ii}W_{ii}R_i+a_{ij}h_{ij}W_{ij}R_j}
+\end{equation}
 
@@
 different habitats.
-In both scenarios, I assumed consumer feeding rate increased linearly with resource density.
-Later, I relax this assumption and consider a more realistic functional response when resources occurred in different habitats.
-Note that C~i~ was always pre-adapted to R~i~ (a~ii~ > a~ij~)."}
+The width of each arrow corresponds to initial attack rates (a~ii~, a~ij~) of consumers.
+Note that C~i~ was always pre-adapted to R~i~ (a~ii~ > a~ij~).
+In each scenario, I assumed consumer feeding rate increased linearly with resource abundance.
+I also relax this assumption and consider a more realistic functional response when resources occurred in different habitats."}
 knitr::include_graphics("ECD_model_foraging_scenarios")
@@
 
-Given that character divergence was an expected result, I focus here on its consequences for food-web dynamics across these foraging scenarios.
+Given that character divergence occurred across these foraging scenarios, I focus here on its consequences for food-web dynamics.
 To do this, I analyzed differences in resource abundances and food-web stability at equilibrium. 
-An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and each consumer ($\hat C_1,\hat C_2$).
-At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource abundances and performed a local stability analysis [@Otto2007]. 
+An equilibrium is reached when the rates of change in equation \ref{eq:1} are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and each consumer ($\hat C_1,\hat C_2$).
+Importantly, inferences about ecological character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992].
+Therefore, I compared resource abundances between food webs with ($\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$) and without ($\hat R_1,\hat R_2, \hat C_1 > 0$) a competing consumer.
+I also compared the local stability of these food webs using standard methods [@Otto2007]. 
 This stability analysis derives the dominant eigenvalue, $\lambda$, which determines whether (and how readily) the food web will return to equilibrium after a small perturbation.
@@
 Controlling for other sources of variability allowed me to isolate the general effects of character divergence.
-All mathematical derivations were conducted in @Mathematica and are provided in the Supplementary Information (Appendix S1-3).
+All mathematical derivations were conducted in Mathematica [@Mathematica] and are provided in the Supplementary Information (Appendix S1-3).
 
@@
 The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006].
-This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher than the extremes (when $n>1$, green line in Fig.
-\ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line Fig.
+This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher than the extremes (when $n>1$, green line in fig.
+\ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line fig.
 \ref{fig:tradeoff}).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig. \ref{fig:tradeoff}). 
+When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in fig. \ref{fig:tradeoff}). 
 Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
@@
 $\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
-Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and *w_{ii}* = 0.6. 
+Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and *w~ii~* = 0.6. 
 I set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*. 
-I set initial consumer and resource abundances as: *R~1~* = *R~2~* = 2; *C~1~* = *C~2~* = 1.
+I set initial consumer and resource abundances to: *R~1~* = *R~2~* = 2; *C~1~* = *C~2~* = 1.
 All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model).
@@
           base_height = 5, # 6
-          base_width = 8) #9
-```
-
-```{r tradeoff, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: intermediate combinations of attack rates (a~ii~, a~ij~) are higher than the extremes (green line, n > 1); extreme combinations of attack rates are higher than intermediate investments (orange line, n < 1); and all combinations of attack rates have the same total attack rate (blue line, n = 1).
+          base_width = 7) #9
+```
+
+```{r tradeoff, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: intermediate combinations of attack rates (a~ii~, a~ij~) are higher than the extremes (green line, n > 1); extreme combinations of attack rates are higher than intermediate investments (orange line, n < 1); and all combinations of attack rates have the same total attack rate (blue line, n = 1).
 Points corresponding to attack rates at the beginning of the simulation.
@@
 
-In this first scenario, the abundance of resources at equilibrium are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), and are determined by the following equation (derived in Appendix S**?**):
-
-$$\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}$$
+In this first scenario (equation \ref{eq:2}), the abundance of resources at equilibrium are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), and are determined by the following equation (derived in Appendix S2):
+
+\begin{equation}
+  \hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}
+\end{equation}
+
 A key determinant of resource abundance in this model is the consumer's total attack rate, $a_{ii}+a_{ij}$.
@@
 
-My simulations show that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource abundances (Fig. \ref{fig:plot_fig3}). 
-For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (Fig.
-\ref{fig:plot_fig3}) and character displacement has no effect on resource abundances (Fig.
+My simulations show that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource abundances (fig. \ref{fig:plot_fig3}). 
+For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (fig.
+\ref{fig:plot_fig3}) and character displacement has no effect on resource abundances (fig.
 \ref{fig:plot_fig3}).
-If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (Fig. \ref{fig:plot_fig3}). 
-This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (Fig. \ref{fig:plot_fig3}). 
-When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (Fig. \ref{fig:plot_fig3}). 
-Although the equation I derived for resource abundances was for the scenario where both consumers and both resources were present, it accurately predicts the abundance of resources when a single consumer reaches its ESS (triangles on respective colored lines in Fig.
+If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (fig. \ref{fig:plot_fig3}). 
+This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (fig. \ref{fig:plot_fig3}). 
+When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (fig. \ref{fig:plot_fig3}). 
+Although the equation I derived for resource abundances was for the scenario where both consumers and both resources were present, it accurately predicts the abundance of resources when a single consumer reaches its ESS (triangles on respective colored lines in fig.
 \ref{fig:plot_fig3}).
-This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in Fig.
-\ref{fig:MacArthur_Plot}), resulting in equivalent resource abundances.
+This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in fig.
+\ref{fig:plot_fig3}), resulting in equivalent resource abundances.
 
 The effect of character displacement on resources corresponds to its impact on food-web stability. 
-For example, when character displacement decreases resource abundances (orange points in \ref{fig:plot_fig3}), there is a decrease in food-web stability (\ref{fig:plot_fig3}). 
-Similarly, if character displacement does not alter resource abundances (blue and green points in \ref{fig:plot_fig3}), then there is no change in food-web stability (\ref{fig:plot_fig3}).
+For example, when character displacement decreases resource abundances (orange points in fig. \ref{fig:plot_fig3}), there is a decrease in food-web stability (fig. \ref{fig:plot_fig3}). 
+Character divergence may not affect or even increase food-web stability (blue and green lines in fig.
+\ref{fig:plot_fig3}); however, evolution does not favor strong divergence in these scenarios (blue and green points in fig.
+\ref{fig:plot_fig3}), which dampens these contingent effects.
 
@@
   scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  geom_hline(yintercept=0, color="grey") +
+  #geom_hline(yintercept=0, color="grey") +
   guides(fill = guide_legend(override.aes=list(shape=21)))
@@
 -->
-In this second scenario, I again see that resource abundances are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), but are now determined by the following equation (derived in Appendix S**?**):
-
-$$\hat{R}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e}$$
+In the second foraging scenario (equation \ref{eq:3}), I again see that resource abundances are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), but are now determined by the following equation (derived in Appendix S3):
+
+\begin{equation}
+  \hat{R}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e}
+\end{equation}
 
 This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g.
-$w_{ii}>0.5$ and $a_{ii}>a_{ij}$), then the effective attack rate of consumers ($wa_{ii}+(1-w_{ii})a_{ij}$) will always increase, regardless of the tradeoff (Fig.
+$w_{ii}>0.5$ and $a_{ii}>a_{ij}$), then the effective attack rate of consumers ($wa_{ii}+(1-w_{ii})a_{ij}$) will always increase, regardless of the tradeoff (fig.
 \ref{fig:plot_fig3}).
-Thus, character divergence always results in resource suppression (Fig. \ref{fig:plot_fig3}). 
+Thus, character divergence always results in resource suppression (fig. \ref{fig:plot_fig3}). 
 Note that the shape of the tradeoff can modify the effect of character displacement. 
-This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached an evolutionary stable strategy (ESS, triangles in Fig.
+This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached an evolutionary stable strategy (ESS, triangles in fig.
 \ref{fig:plot_fig3}).
-In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
+In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in fig.
 \ref{fig:plot_fig3}), because character displacement tends to reach a constraint of complete specialization.
-It is worth noting that resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
+It is worth noting that resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in fig.
 \ref{fig:plot_fig3}).
-This is because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in Fig.
+This is because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in fig.
 \ref{fig:plot_fig3}).
 
-As seen previously, the effect of character displacement on resource abundances qualitatively corresponds to its effect on food-web stability (Fig. \ref{fig:plot_fig3}C). 
+As seen previously, the effect of character displacement on resource abundances qualitatively corresponds to its effect on food-web stability (fig. \ref{fig:plot_fig3}). 
 Specifically, character divergence decreases food-web stability, regardless of the tradeoff in attack rates.
-This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (Fig.
-\ref{fig:plot_fig3}C).
-For example, when the tradeoff is concave up (orange), then the four-species (small circle) food web actually starts off as more stable than the three-species one (small triangle); however, this pattern switches by the end of the eco-evolutionary simulation (large points).
+This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (fig.
+\ref{fig:plot_fig3}).
+For example, when the tradeoff is concave up (orange), the initial two-consumer food web (small circle) is more stable than when there is only one consumer (small triangle); however, this pattern switches by the end of the eco-evolutionary simulation (large points).
 
@@
 Macs <- plot_grid(
-  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[ii]+a[ij]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none",  axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0)), #  + scale_y_continuous(breaks=c(1.8,2.0,2.2))
-  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text.x = element_blank(), axis.title.y=element_text(vjust=0)), # + scale_y_continuous(breaks=c(1.0,1.2,1.4), limits=c(1.0,1.45))
-  Stability_Mac + xlab("")  +  ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + theme(legend.position = "none", axis.title.y=element_text(vjust=0)), # + scale_y_continuous(breaks=c(0,0.1))
+  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[ii]+a[ij]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), #  + scale_y_continuous(breaks=c(1.8,2.0,2.2))
+  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)), # + scale_y_continuous(breaks=c(1.0,1.2,1.4), limits=c(1.0,1.45))
+  Stability_Mac + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)), # + scale_y_continuous(breaks=c(0,0.1))
   ncol=1, align='v'#, labels = "AUTO"
@@
 LSs <- plot_grid(
-  LS_attack  + xlab("") + ylab(bquote(w*a[ii]+"("*1-w*")"*a[ij])) + ggtitle("Resources: Different Habitats\n") + theme(legend.position = "none", axis.text.x = element_blank(), plot.title = element_text(size = 10)), #+ scale_y_continuous(breaks=c(0.8,1.0,1.2))
-  LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text.x = element_blank()), #+ scale_y_continuous(breaks=c(2.0,2.5,3.0), limits=c(2.0,3.2))  
-  Stability_LS + xlab("") + ylab("") + theme(legend.position = "none"), # + scale_y_continuous(breaks=c(0,0.1,0.2))
+  LS_attack  + xlab("") + ylab(bquote(w*a[ii]+"("*1-w*")"*a[ij])) + ggtitle("Resources: Different Habitats\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), #+ scale_y_continuous(breaks=c(0.8,1.0,1.2))
+  LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()), #+ scale_y_continuous(breaks=c(2.0,2.5,3.0), limits=c(2.0,3.2))
+  Stability_LS + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10)), # + scale_y_continuous(breaks=c(0,0.1,0.2))
   ncol=1, align='v'
@@
 plot_fig3 <- plot_grid(Macs, LSs, ncol=2, rel_widths = c(1,0.9)) 
-plot_fig3_sub <- add_sub(plot_fig3, expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")"))), vjust = 0)
-
-save_plot(filename="Figure_3.pdf", plot=plot_fig3_sub, base_height = 8, base_width = 6)
-```
-
-```{r plot_fig3, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:plot_fig3}**Effect of character displacement on food-web dynamics.** Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85; as in Fig.
-1).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
-Different panels show the effect of character displacement on effective attack rates (A), resource abundances (B), and food-web stability (C) when resources occur in the same vs.
-different habitats.
-Note that the equation for effective attack rate depends on the foraging scenario.
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6; *w~ij~*=0.4.
-I set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-I set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
+plot_fig3_sub <- add_sub(plot_fig3, expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")"))), vjust = 0, size=12)
+
+save_plot(filename="Figure_3.pdf", plot=plot_fig3_sub, base_height = 6.5, base_width = 6)
+```
+
+```{r plot_fig3, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:plot_fig3}**Effect of character displacement on food-web dynamics under different evolutionary tradeoffs and foraging scenarios.** Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in stability panels).
+In both foraging scenarios, feeding rates increase linearly with resource abundance, but the equation for effective attack rate is different."}
 knitr::include_graphics("Figure_3.pdf")
@@
 
-Assuming consumers have a preferred resource (i.e. $w_{ii}>0.5$), then consumers will only reach an ESS when $|\delta_{ii}|<|\delta_{ij}|$ (e.g. green line, Fig. \ref{MacArthur_Plot}A). 
+Assuming consumers have a preferred resource (i.e. $w_{ii}>0.5$), then consumers will only reach an ESS when $|\delta_{ii}|<|\delta_{ij}|$ (e.g. green line, fig. \ref{MacArthur_Plot}A). 
 Otherwise, they will evolve to be complete specialists on their preferred resource.
@@
 
-In the third foraging scenario, I observed the same general effect of character divergence as the previous scenario (resources in different habitats, but linear functional response).
+In the third foraging scenario (equation \ref{eq:4}), I observed the same general effect of character divergence as the previous scenario (resources in different habitats, but linear functional response).
 <!--Divergent character displacement in this foraging scenario has the same general effect on food-web dynamics as the previous scenario (resources in different habitats, but linear functional response).
@@
 -->
-This is because resource abundances at equilibrium are governed by a similar dynamic (derived in Appendix S**?**):
+This is because resource abundances at equilibrium are governed by a similar dynamic (derived in Appendix S1):
 <!--
-Even after adding more biological realism, I still observe the same qualitative relationship between character displacement and resource suppression (Fig.
+Even after adding more biological realism, I still observe the same qualitative relationship between character displacement and resource suppression (fig.
 \ref{fig:McCann_Plot} in Appendix).
@@
 
-$$\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e-hm}$$
-
-And since evolution favors consumer divergence onto their preferred resources (see Appendix S1), the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase (Fig.
-\ref{fig:McCann_Plot}A), resulting in lower resource abundances and decreased food-web stability (Appendix S?).
-
-In the previous foraging scenarios, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$, see Appendix S1 and S2).
+\begin{equation}
+  \hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e-hm}
+\end{equation}
+
+And since evolution favors consumer divergence onto their preferred resources (see Appendix S1), the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase, resulting in lower resource abundances and decreased food-web stability (Appendix S4).
+
+In firs two foraging scenarios, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$, see Appendix S2-3). 
 In this more realistic model, however, whether the food web is locally stable depends on consumer and resource parameters. 
-Specifically, I found that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derived using Routh-Hurwitz criteria in Appendix S**?**):
-
-$$w_{ii}a_{ii}+(1-w_{ii})a_{ij} > \frac{e+hm}{hK(e-hm)}$$
-
-<!--Previously, I showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff (Fig.
+Specifically, I found that the two-consumer food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derived using Routh-Hurwitz criteria in Appendix S1):
+
+\begin{equation}
+  w_{ii}a_{ii}+(1-w_{ii})a_{ij} > \frac{e+hm}{hK(e-hm)}
+\end{equation}
+
+<!--Previously, I showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff (fig.
 \ref{fig:LS_Plot}A).-->
-Thus, character displacement always pushes the food web toward an unstable stucture in this more realistic foraging scenario (Fig. \ref{fig:bifur_plot}).
+Thus, character displacement always pushes the food web toward an unstable stucture in this more realistic foraging scenario (fig.
+\ref{fig:bifur_plot}).
+Note that I stopped the simulation in the four-species food web once it became locally unstable.
+I do not simulate beyond this point as this would require making assumptions about the dynamics of mutant consumers in variable environments, which is beyond the scope of this work.
 
@@
 
-```{r bifur_plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Character divergence creates unstable food webs**.
-I simulated the eco-evolutionary dynamics of character displacement assuming a more realistic functional response for consumers.
-The results are qualtitatively similar for consumer dynamics (not shown).
-Note that the lines illustrate the ecological consequences across the range of consumer specialization, while the points are the results of an eco-evolutionary simulation.
-In this simulation, this symmetry in attack rates may be broken, which is why the point of instability does not exactly match the lines.
-I stopped the simulations once consumers reached an evolutionary stable strategy, a constraint of maximum specialization, or created an unstable food web.
-I do not project beyoned this, because the dynamics of mutant invaders becomes unclear when resource distributions are unstable.
-The grey arrow denotes the boundary of local stability."}
+```{r bifur_plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Character divergence creates an unstable food web**.
+Lines illustrate the effects of character divergence across the range of specialization, while the points are the results of an eco-evolutionary simulation.
+Note that I increased the total investment in attack rates (*A* = 3.3) to create a scenario that could result in an unstable food web."}
 knitr::include_graphics("bifurcation_plot.pdf")
@@
 
-The previous analytical results and simulations make a strong assumption that competing consumers start off as perfect mirror images of each other in their attack rates (i.e. symmetry). 
+The previous analytical results and simulations make a strong assumption that competing consumers start off as perfect mirror images of each other (i.e. symmetry). 
 Yet, theory indicates a predictable asymmetry between initial consumer attack rates. 
@@
 
-To test whether my results were robust to this asymmetry, I used the evolved attack rates at the end of the single consumer simulations as the starting values for the competitor in the two consumer simulations.
+To test whether my results were robust to this asymmetry, I used the evolved attack rates at the end of the simulations with one consumer as the starting values for one of the two consumers. 
 I did this for all foraging scenarios and tradeoffs previously examined. 
-I found that my previous inferences are robust to including consumer asymmetry across different tradeoffs and foraging scenarios (see Appendix S4).
+I found that my previous inferences are robust to including consumer asymmetry across different foraging scenarios and tradeoffs (Appendix S4).
 <!--A\ref{fig:MacArthur_Plot_asymm}, A\ref{fig:LS_Plot_asymm}, A\ref{fig:McCann_Plot_asymm}).-->
@@
 In fact, the ecological and evolutionary scenarios that favored the greatest character divergence always decreased the abundance of resources. <!-- build paragraph around this topic sentence -->
-For example, the magnitude of character displacement was highest when mobile consumers competed for resources that occur in different habitats compared to within the same area.
+For example, the magnitude of character displacement was highest when mobile consumers competed for resources that occur in different habitats.
 An empirical example of this is threespine stickleback that have diverged into limnetic and benthic species with specialized traits to forage on zooplankton and benthic invertebrates, respectively [@Schluter1992; @Schluter2000].
-These resources occur in distinct zones of the lake (pelagic and littoral), therefore stickleback must move between these zones and compete when foraging for resources. 
-Interestingly, a disproporationate number of the documented cases of character displacement involve carnivores [@Schluter2000] that are larger, and likely more mobile, than their resources [@McCann2005], suggesting my predictions may apply to many cases of ecological character displacement.
+These two resource types occur in distinct zones of the lake (pelagic and littoral), therefore stickleback must move between these zones when foraging for resources. 
+Interestingly, a disproportionate number of the documented cases of character displacement involve carnivores [@Schluter2000] that are larger, and likely more mobile, than their resources [@McCann2005], suggesting many cases of ecological character displacement may result in lower resource availability.
 
@@
 Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the net foraging rate of consumers [@Schluter1985; @Arnegard2014].
-While it is theoretically possible that character displacement does not alter resource abundances, this was limited to the simplest, and arguably least realistic, foraging scenario and under tradeoffs that did not favor large displacements, and thus less likely to detect in nature.
+While it is theoretically possible that character displacement does not alter (or even increase) resource abundances, this was limited to the simplest, and arguably least realistic, foraging scenario and under tradeoffs that did not favor large displacements, and thus less likely to detect in nature.
 Taken together, my results call for empirical work to test these clear theoretical predictions and suggest a revision is needed for one of the criteria used to demonstrate character displacement.
@@
 
-My most striking result was that character divergence, when examined through the lens of food-web theory, made food webs less resilient to perturbations. 
+My most striking result was that character divergence made food webs less resilient to perturbations. 
 In fact, under the most realistic foraging scenario, character divergence can even result in an unstable food web. 
@@
 For example, @McPeek2019 showed that character divergence is favored in food webs that are either highly productive, easy to find and capture resources, or under weak abiotic stress.
-This corresponds to higher values of $K$ (productivity) or $A$ (capacity to find and capturing resources), or lower values of $m$ (abiotic stress).
+This corresponds to higher values of $K$ (productivity) or $A$ (investment in attack rates), or lower values of $m$ (abiotic stress).
 Each of these corresponding changes decrease food-web resilience, as they increase the strength of consumer-resource interactions relative to intraspecific interactions.
@@
 Such comparisons are necessary for inferring the effects of character displacement, a point that has been made clear in the criteria to demonstrate character displacement [@Schluter1992; @Schluter2000].
-Although the addition of a consumer to a food web can decrease its resilience in the absence of evolution [@May1973], my results are primarily driven by an eco-evolutionary feedback between consumer evolution and resource abundances (compare initial and ending points in Fig.
-\ref{fig:MacArthur_Plot}C, \ref{fig:LS_Plot}C, \ref{fig:McCann_Plot}C).
+Although the addition of a consumer to a food web can decrease its resilience in the absence of evolution [@May1973], my results are primarily driven by an eco-evolutionary feedback between consumer evolution and resource abundances.
 
@@
 I studied this eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
-A strength of this approach is that it enabled me to gain analytical insight to the effects of character divergence on food-web dynamics in a more realistic foraging scenario.
+A strength of this approach is that it enabled me to gain analytical insight to the effects of character divergence in a more realistic foraging scenario.
 This is much less tractable in quantitative genetic [@Taper1985; @McPeek2017] or explicit genetic [@Doebeli1996] models of character displacement, which is why the foraging scenarios previously examined have been limited [but see @McPeek2017].
@@
 This is because models that explicitly include resource dynamics inevitably show that resource competition results in character divergence, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
-A quantitative genetic model may certainly show differences in the magnitude of character divergence, but this should not affect the consequences for food-web dynamics.
+A quantitative genetic model may certainly show differences in the magnitude of character divergence, but this should not qualtitatively change its effect on food-web dynamics.
 <!-- In fact, recent work by @McPeek2017 has shown that divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
@@
 
-```{r all asymmetry}
+```{r all asymmetry, echo=FALSE}
 

### 2019-03-22 supplementary_ECD_model.Rmd
added final edits and sent to Seth, JP, and Matt Osmond for friendly reviews.

@@
     citation_package: natbib 
-    includes:  
-      in_header: Am-Nat-preamble-latex.tex
+#    includes:  
+#      in_header: Am-Nat-preamble-latex.tex
+header-includes:
+- \usepackage{caption}
+- \captionsetup[figure]{labelformat=empty}
+- \usepackage{fullpage}
+- \linespread{1.7}
 bibliography: references
@@
 \vspace*{0.1cm} 
-\begin{center} \LARGE Appendix S**?**: \\ Ecological character displacement destabilizes food webs \end{center}
+\begin{center} \LARGE Appendix S4: \\ Ecological character displacement destabilizes food webs \end{center}
 <!--\begin{center} \LARGE Food-web dynamics under ecological character displacement \end{center}-->
@@
 
-*Elements*: Figure S1 and figure S2.
+*Elements*: Figures S1 and S2.
 
-<!-- End title page formatting -->
+<!-- End title page formatting
 
@@
 \modulolinenumbers[3]
+-->
 
@@
 
-.
+(ignore this blank page for the moment. Rmarkdown can be a pain in the ass.)
 
-```{r McCann_Plot, echo=FALSE, fig.pos="H", fig.cap="**Effect of character displacement when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+```{r McCann_Plot, echo=FALSE, fig.pos="H", fig.cap="Figure S1: **Effect of character displacement when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in stability panel).
-This figure illustrates how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
+This figure illustrates how, regardless of the tradeoff, character displacement increases the effective attack rate of consumers.
 This always resulted in a suppression of resource abundances and concomitant decrease in food-web stability.
@@
 
-Figure 2: **Robustness to consumer asymmetry**.
+Figure S2: **Robustness to consumer asymmetry**.
 Scenarios correspond to: (I) resources occur in the same habitat and consumer(s) exhibits a linear functional response; (II) resources occur in different habitats and consumer(s) exhibits a linear functional response; (III) resources occur in different habitats and consumer(s) exhibits a more realistic functional response (details in main text).
@@
 Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in stability panels).
-The similarity between this figure and fig.
+The similarity with fig.
 S1 and fig.
 2 (main text) show that adding an asymmetry in consumer attack rates does not alter the conclusions reported in the main text.
-If anything, it is more likely an asymmetry will push the system toward the boundary of stability (note position of large circles in stability panel of Scenario I).
+If anything, it is more likely that an asymmetry will push the system toward the boundary of stability (note position of large circles in stability panel of Scenario I).
 Details of the consumer asymmetry simulation are given in the main text.

### 2019-04-10 manuscript_ECD_model.Rmd
incorporated minor comments from JP, still need to work on notation.

@@
 This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
-Taken together, my results show that the adaptive process of character displacement often comes with the ecological cost of decreasing food-web resilience.
+Taken together, my results show that the adaptive process of character displacement may come with the ecological cost of decreasing food-web resilience.
 
@@
 This process is due to "phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000]. 
-Over the past 40 years, a large body of theoretical [e.g.
+A large body of theoretical [e.g.
 @Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has examined which scenarios lead to phenotypic divergence or convergence of competing consumers.
-The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985].
+The general conclusion has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985].
 This adaptive process is not simply a response to static differences in resource distributions, but creates an eco-evolutionary feedback that drives further differentiation. 
@@
 
-Although models that included resources gave insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics has received surprisingly little attention. 
+Although models that included resources led to insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics has received surprisingly little attention. 
 This is likely because the ecological feedback has been primarily studied through the lens of coexistence theory [@Lawlor1976; @Germain2018; @Bassar2017; @McPeek2019]. 
-For example, early theoretical work by @Lawlor1976 showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition.
+For example, early theoretical work showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition [@Lawlor1976].
 Yet, this reduction in interspecific competition may, at the same time, increase interspecific interactions between specialized consumers and their resources. 
@@
 I address two questions: (1) How does ecological character displacement affect resource abundances? (2) How does character displacement affect food-web stability? 
-To test the generality of these effects, I explored different ecological foraging scenarios (fig.
-\ref{fig:foraging_scenarios}) and evolutionary tradeoffs in consumer attack rates (fig.
-\ref{fig:tradeoff}).
+To test the generality of these effects, I explored different ecological foraging scenarios and evolutionary tradeoffs in consumer attack rates. 
 I found that the adaptive process of character displacement often comes with an ecological cost; resulting in food webs with lower resource availability and that are less resilient to perturbations.
@@
   \begin{split}
-     & \frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2 \\
-     & \frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2 \\
+     & \frac{dR_1}{dt}=r_1R_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2 \\
+     & \frac{dR_2}{dt}=r_2R_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2 \\
      & \frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1 \\
@@
 This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g.
-zooplankton and benthic invertebrates in lakes) rather than if resources are better characterized by a continuous trait distribution (e.g.
+zooplankton and benthic invertebrates in lakes) rather than a scenario where resources are better characterized by a continuous trait distribution (e.g.
 seed size, see @Taper1985 for an example).
@@
 This first scenario is the starting point for many models of resource competition [@MacArthur1972]; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005].
-A simple extension that accounts for this spatial context takes the form:
+The second scenario accounts for this spatial context and takes the form:
 
@@
 Finally, it is well known that consumer feeding rates often saturate at high resource abundances [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011] and that consumers do not usually spend a fixed proportion of time in a particular habitat [@McCann2005].
-To account for these biological realities, I consider a more realistic functional response [derived by @McCann2005]:     
+The third scenario accounts for these biological realities and takes the form [derived by @McCann2005]:     
 \begin{equation} \label{eq:4}
@@
 I also compared the local stability of these food webs using standard methods [@Otto2007]. 
-This stability analysis derives the dominant eigenvalue, $\lambda$, which determines whether (and how readily) the food web will return to equilibrium after a small perturbation.
+This stability analysis derives the dominant eigenvalue, $\lambda$, of the matrix of partial derivatives of all differential equations with respect to all variables evaluated at equilibrium. 
+If the dominant eigenvalue is negative, then the food web will return to equilibrium after a small perturbation (i.e. locally stable), with more negative values indicating a faster return time.
+If the dominant eigenvalue is positive, then the food web is not locally stable. 
+<!--This dominant eigenvalue determines whether (and how readily) the food web will return to equilibrium after a small perturbation.-->
 
@@
 -->
-Specifically, at each evolutionary time step, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability.
+Specifically, after letting consumer and resource abundances reach a steady state, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability.
 The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006].
@@
 Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
-If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state. 
+If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource abundances to reach a steady state. 
 I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g.
@@
 For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (fig.
-\ref{fig:plot_fig3}) and character displacement has no effect on resource abundances (fig.
-\ref{fig:plot_fig3}).
-If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (fig. \ref{fig:plot_fig3}). 
-This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (fig. \ref{fig:plot_fig3}). 
-When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (fig. \ref{fig:plot_fig3}). 
-Although the equation I derived for resource abundances was for the scenario where both consumers and both resources were present, it accurately predicts the abundance of resources when a single consumer reaches its ESS (triangles on respective colored lines in fig.
-\ref{fig:plot_fig3}).
+\ref{fig:plot_fig3} top left) and character displacement has no effect on resource abundances (fig.
+\ref{fig:plot_fig3} middle left).
+If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (fig. \ref{fig:plot_fig3} middle left). 
+This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (fig. \ref{fig:plot_fig3} top left). 
+When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (fig. \ref{fig:plot_fig3} top left). 
+Although the equation I derived for resource abundances was for the scenario where both consumers and both resources were present, it accurately predicts the abundance of resources when a single consumer reaches its evolutionary stable strategy (ESS; triangles on respective colored lines in fig.
+\ref{fig:plot_fig3} middle left).
 This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in fig.
-\ref{fig:plot_fig3}), resulting in equivalent resource abundances.
+\ref{fig:plot_fig3} left panels), resulting in equivalent resource abundances.
 
 The effect of character displacement on resources corresponds to its impact on food-web stability. 
-For example, when character displacement decreases resource abundances (orange points in fig. \ref{fig:plot_fig3}), there is a decrease in food-web stability (fig. \ref{fig:plot_fig3}). 
+For example, when character displacement decreases resource abundances (orange points in fig.
+\ref{fig:plot_fig3} middle left), there is a decrease in food-web stability (fig.
+\ref{fig:plot_fig3} bottom left).
 Character divergence may not affect or even increase food-web stability (blue and green lines in fig.
-\ref{fig:plot_fig3}); however, evolution does not favor strong divergence in these scenarios (blue and green points in fig.
+\ref{fig:plot_fig3} bottom left); however, evolution does not favor strong divergence in these scenarios (blue and green points in fig.
 \ref{fig:plot_fig3}), which dampens these contingent effects.
@@
 $w_{ii}>0.5$ and $a_{ii}>a_{ij}$), then the effective attack rate of consumers ($wa_{ii}+(1-w_{ii})a_{ij}$) will always increase, regardless of the tradeoff (fig.
-\ref{fig:plot_fig3}).
-Thus, character divergence always results in resource suppression (fig. \ref{fig:plot_fig3}). 
+\ref{fig:plot_fig3} top right).
+Thus, character divergence always results in resource suppression (fig. \ref{fig:plot_fig3} middle right). 
 Note that the shape of the tradeoff can modify the effect of character displacement. 
-This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached an evolutionary stable strategy (ESS, triangles in fig.
-\ref{fig:plot_fig3}).
+This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached an ESS (triangles in fig.
+\ref{fig:plot_fig3} middle right).
 In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in fig.
-\ref{fig:plot_fig3}), because character displacement tends to reach a constraint of complete specialization.
+\ref{fig:plot_fig3} middle right), because character displacement tends to reach a constraint of complete specialization.
 It is worth noting that resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in fig.
-\ref{fig:plot_fig3}).
+\ref{fig:plot_fig3} middle right).
 This is because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in fig.
-\ref{fig:plot_fig3}).
+\ref{fig:plot_fig3} right panels).
 
@@
 
+Although I model the indirect effects of coevolution between consumers, I do not account for potential coevolution between consumers and resources. 
+In the context of my model, I would expect prey to evolve traits that reduce consumer attack rates. 
+Thus, prey evolution would act to counter the effects of character displacement on resource abundance and food-web stability. 
+Note that this does not negate my general conclusion that ecological character displacement decreases resource abundances and stability; however, this process may itself create another eco-evolutionary feedback between consumers and resources.
+Thus, not coevolution itself, but the prey's evolutionary response would act to stabilize food webs.
+This may actually help maintain fantastic examples of character divergence and prevent them from destabilizing systems, basically because the consumers trait becomes decoupled with their attack rate.
+Examining this role would be ideal in a quantitative genetic setting that can explicitly track trait dynamics, but it would not fundamentally change the conclusions presented here.
+
+Another potential caveat is that I explored my model in a setting that makes many assumptions about resource and consumer symmetry (but see consumer asymmetry section). 
+Prior work has shown that allowing for resource asymmetry, for example, may decrease the magnitude of character displacement. 
+This is because this asymmetry creates an asymmetry in resource abundances, which dampens the effects of character divergence.
+Again though, this may dampen the effects of divergence, but should not change the qualitative relationship we observed of ecological character displacement decreasing resource abundances and food-web stability.
+
 I studied this eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.

### 2019-04-25 manuscript_ECD_model.Rmd
adjusted notation so that i only refers to resources and j only refers to consumers throughout the text and figures.

@@
 
-where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ij}$ represents the conversion efficiency of resource $j$ into consumer $i$, and $m_i$ represents the mortality rate of consumer $i$.
-$F_{ij}(R_j)$ represents consumer $i$'s feeding rate on resource $j$ (i.e.
+where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ij}$ represents the conversion efficiency of resource $i$ into consumer $j$, and $m_j$ represents the mortality rate of consumer $j$.
+$F_{ij}(R_i)$ represents consumer $j$'s feeding rate on resource $i$ (i.e.
 functional response).
@@
 
-I studied three different foraging scenarios (fig. \ref{fig:foraging_scenarios}). 
-In the first, I assume that consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that:
+I studied three different foraging scenarios. 
+In the first, I assume that consumers can forage for both resources simultaneously (fig. \ref{fig:foraging_scenarios} top) and their feeding rates increase linearly with resource abundance, such that:
 
 \begin{equation} \label{eq:2}
-  F_{ii}(R_i)=a_{ii}R_i
+  F_{ij}(R_i)=a_{ij}R_i
 \end{equation}
 
-where $a_{ii}$ is the attack rate of consumer $i$ on resource $i$. 
+where $a_{ij}$ is the attack rate of consumer $j$ on resource $i$. 
 This first scenario is the starting point for many models of resource competition [@MacArthur1972]; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005].
-The second scenario accounts for this spatial context and takes the form:
+The second scenario accounts for this spatial context (fig. \ref{fig:foraging_scenarios} bottom) and takes the form:
 
 \begin{equation} \label{eq:3}
-  F_{ii}(R_i)=w_{ii}a_{ii}R_i
+  F_{ij}(R_i)=w_{ij}a_{ij}R_i
 \end{equation}
 
-where $w_{ii}$ represents the proportion of time consumer $i$ spends foraging in a habitat where only resource $i$ is found (i.e. habitat preference). 
-Note that since $w_{ii}$ is a proportion that $w_{ij}=1-w_{ii}$. 
+where $w_{ij}$ represents the proportion of time consumer $j$ spends foraging in a habitat where only resource $i$ is found (i.e. habitat preference). 
+Note that since $w_{ij}$ is a proportion that $w_{1,j}=1-w_{2,j}$. <!-- $\sum_jw_{ij}=1$.<!-- $w_{ij}=1-w_{ii}$. -->
 Finally, it is well known that consumer feeding rates often saturate at high resource abundances [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011] and that consumers do not usually spend a fixed proportion of time in a particular habitat [@McCann2005].
@@
 \begin{equation} \label{eq:4}
-  F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}h_{ii}W_{ii}R_i+a_{ij}h_{ij}W_{ij}R_j}
+  F_{ij}(R_i)=\frac{a_{ij}W_{ij}R_i}{1+a_{1,j}h_{1,j}W_{1,j}R_1+a_{2,j}h_{2,j}W_{2,j}R_2}
 \end{equation}
 
-where consumer $i$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling times $h_{ii}$, $h_{ij}$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ii}=\frac{w_{ii}R_i}{w_{ii}R_i+(1-w_{ij})R_j}$.
+where consumer $j$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling times $h_{ij}$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ij}=\frac{w_{ij}R_i}{w_{1,j}R_1+w_{2,j}R_2}$.
 
 Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach, with the general result being character divergence [@Lawlor1976; @Abrams1986].
-I say consumers have undergone character divergence if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
-I also used an Adaptive Dynamics approach to analyze character displacement in the third foraging scenario, and I too observed character divergence (detailed analysis in Appendix S1 of Supplementary Information).
+I say consumers have undergone character divergence if there evolved attack rates are more specialized, measured as $|\frac{a_{ij}}{a_{1,j}+a_{2,j}-0.5|$, when evolving with a competing consumer.
+I also used an Adaptive Dynamics approah to analyze character displacement in the third foraging scenario, and I too observed character divergence (detailed analysis in Appendix S1 of Supplementary Information).
 
@@
 different habitats.
-The width of each arrow corresponds to initial attack rates (a~ii~, a~ij~) of consumers.
-Note that C~i~ was always pre-adapted to R~i~ (a~ii~ > a~ij~).
-In each scenario, I assumed consumer feeding rate increased linearly with resource abundance.
+The width of each arrow corresponds to initial attack rates (*a*~ij~) of consumer *j* on resource *i*.
+Note that *C*~1~ was pre-adapted to *R*~1~ (*a*~11~ > *a*~21~), while *C*~2~ was a mirror image, being pre-adapted to *R~2~* (*a*~22~ > *a*~12~).
+In each scenario, I assumed consumer feeding rates increased linearly with resource abundance.
 I also relax this assumption and consider a more realistic functional response when resources occurred in different habitats."}
@@
 I also compared the local stability of these food webs using standard methods [@Otto2007]. 
-This stability analysis derives the dominant eigenvalue, $\lambda$, of the matrix of partial derivatives of all differential equations with respect to all variables evaluated at equilibrium. 
-If the dominant eigenvalue is negative, then the food web will return to equilibrium after a small perturbation (i.e. locally stable), with more negative values indicating a faster return time.
-If the dominant eigenvalue is positive, then the food web is not locally stable. 
+This stability analysis derives the dominant eigenvalue, $\lambda_{max}$, of the matrix of partial derivatives of all differential equations with respect to all variables evaluated at equilibrium. 
+If $-\lambda_{max}>0$, then the food web will return to equilibrium after a small perturbation (i.e. locally stable), with more positive values indicating a faster return time.
+If $-\lambda_{max}<0$, then the food web is not locally stable. 
 <!--This dominant eigenvalue determines whether (and how readily) the food web will return to equilibrium after a small perturbation.-->
@@
 When possible, I derived analytical expressions for the relationship between consumer attack rates and food-web dynamics. 
-To do this, I simplified the model by assuming that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $h=h_{11}=h_{12}=h_{21}=h_{22}$; $m=m_1=m_2$), except that consumer attack rates and their habitat preferences (if present) are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$; $w_{ii}=w_{11}=w_{22}$).
+To do this, I simplified the model by assuming that resources are equivalent ($r=r_i$ and $K=K_i$) as well as consumers ($e=e_{ij}$; $h=h_{ij}$; $m=m_j$), except that consumer attack rates and their habitat preferences (if present) are mirror images of each other ($a_{11}=a_{22}$; $a_{12}=a_{21}$; $w_{11}=w_{22}$).
+Note that I arbitrarily set $C_1$ as being pre-adapted to $R_1$ ($a_{11} > a_{21}$; $w_{11} > 0.5$), and therefore $C_2$ was pre-adapted to $R_2$.
 Controlling for other sources of variability allowed me to isolate the general effects of character divergence.
@@
 Specifically, after letting consumer and resource abundances reach a steady state, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability.
-The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006].
-This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher than the extremes (when $n>1$, green line in fig.
+The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006].
+This function has the useful property that it differentiates between cases where intermediate combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ are higher than the extremes (when $n>1$, green line in fig.
 \ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line fig.
 \ref{fig:tradeoff}).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in fig. \ref{fig:tradeoff}). 
+When $n=1$, the tradeoff function is linear, and all combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ have the same total attack rate (blue line in fig. \ref{fig:tradeoff}). 
 Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
@@
 I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g.
-$\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
-Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and *w~ii~* = 0.6. 
-I set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*. 
-I set initial consumer and resource abundances to: *R~1~* = *R~2~* = 2; *C~1~* = *C~2~* = 1.
+$\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0).
+Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and $w_{11} = w_{22} = 0.6$. 
+I set an initial value of $a_{11} = a_{22} = 1.2$, while $a_{12}$ and $a_{21}$ depended on the value of *n*. 
+I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
 All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model).
@@
   geom_point(data = initial_points, aes(x=a11, y=a12), size=3, show.legend=FALSE) + # size=5
-  xlab(expression(bold(a[ii]))) +
-  ylab(expression(bold(a[ij]))) +
+  xlab(expression(bold(a[11]))) + # a[ii]
+  ylab(expression(bold(a[21]))) + # a[ij]
   scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette) #+
@@
 
-```{r tradeoff, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: intermediate combinations of attack rates (a~ii~, a~ij~) are higher than the extremes (green line, n > 1); extreme combinations of attack rates are higher than intermediate investments (orange line, n < 1); and all combinations of attack rates have the same total attack rate (blue line, n = 1).
-Points corresponding to attack rates at the beginning of the simulation.
-Note that C~i~ was always pre-adapted to R~i~ (a~ii~ > a~ij~)."}
+```{r tradeoff, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: intermediate combinations of attack rates (*a*~1,*j*~, *a*~2,*j*~) are higher than the extremes (green line, *n* > 1); extreme combinations of attack rates are higher than intermediate investments (orange line, *n* < 1); and all combinations of attack rates have the same total attack rate (blue line, *n* = 1).
+Points corresponding to attack rates at the beginning of the simulation for *C*~1~, which was pre-adapted to *R*~1~ (*a*~11~ > *a*~12~)."}
 knitr::include_graphics("attack_tradeoffs.pdf")
@@
 \begin{equation}
-  \hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}
+  \hat{R}=\frac{1}{a_{1,j}+a_{2,j}}\cdot\frac{m}{e}
 \end{equation}
 
-A key determinant of resource abundance in this model is the consumer's total attack rate, $a_{ii}+a_{ij}$.
-Therefore, the effect of character displacement on food-web dynamics depends on how the tradeoff influences the evolution of consumer attack rates.
-
-My simulations show that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource abundances (fig. \ref{fig:plot_fig3}). 
+A key determinant of resource abundance in this scenario is the consumer's total attack rate, $a_{1,j}+a_{2,j}$. 
+Therefore, the effect of character displacement on food-web dynamics depends on how the shape of the tradeoff function influences the evolution of consumer attack rates.
+
+<!-- My simulations show that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource abundances (fig. \ref{fig:plot_fig3}).-->
+I found that the shape of the tradeoff function qualitatively affects the relationship between character displacement and resource abundances in this scenario (fig. \ref{fig:plot_fig3} middle left).
 For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (fig.
@@
 If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (fig. \ref{fig:plot_fig3} middle left). 
-This is because the total attack rate of consumers is maximized at intermediate values ($a_{ii}=a_{ij}$) and decreases as consumers diverge (fig. \ref{fig:plot_fig3} top left). 
+This is because the total attack rate of consumers is maximized at intermediate values ($a_{1,j}=a_{2,j}$) and decreases as consumers diverge (fig. \ref{fig:plot_fig3} top left). 
 When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (fig. \ref{fig:plot_fig3} top left). 
@@
 \ref{fig:plot_fig3}), which dampens these contingent effects.
+Note that the dip in stability occurs when both consumers evolve to be generalists, a situation that is not favored in any of the foraging scenarios we examined (fig. \ref{fig:plot_fig3} bottom).
 
@@
 We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
+We set initial consumer and resource abundances as: *R~i~* = 2; *C~j~* = 1."}
 knitr::include_graphics("Figure_ECD_MacArthur.pdf")
@@
 \begin{equation}
-  \hat{R}=\frac{1}{w_{ii}a_{ii}+(1-w_{ii})a_{ij}}\cdot\frac{m}{e}
+  \hat{R}=\frac{1}{w_{1,j}a_{1,j}+w_{2,j}a_{2,j}}\cdot\frac{m}{e}
 \end{equation}
@@
 This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g.
-$w_{ii}>0.5$ and $a_{ii}>a_{ij}$), then the effective attack rate of consumers ($wa_{ii}+(1-w_{ii})a_{ij}$) will always increase, regardless of the tradeoff (fig.
+$w_{1,j}>0.5$ and $a_{1,j}>a_{2,j}$), then the effective attack rate of consumers ($w_{1,j}a_{1,j}+w_{2,j}a_{2,j}$) will always increase, regardless of the tradeoff (fig.
 \ref{fig:plot_fig3} top right).
@@
 Macs <- plot_grid(
-  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[ii]+a[ij]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), #  + scale_y_continuous(breaks=c(1.8,2.0,2.2))
+  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1,j]+a[2,j]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), #  + scale_y_continuous(breaks=c(1.8,2.0,2.2))
   Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)), # + scale_y_continuous(breaks=c(1.0,1.2,1.4), limits=c(1.0,1.45))
@@
 LSs <- plot_grid(
-  LS_attack  + xlab("") + ylab(bquote(w*a[ii]+"("*1-w*")"*a[ij])) + ggtitle("Resources: Different Habitats\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), #+ scale_y_continuous(breaks=c(0.8,1.0,1.2))
+  LS_attack  + xlab("") + ylab(bquote(w[1,j]*a[1,j]+w[2,j]*a[2,j])) + ggtitle("Resources: Different Habitats\n") +  theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), #+ scale_y_continuous(breaks=c(0.8,1.0,1.2))  # "("*1-w[1,j]*")"
   LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()), #+ scale_y_continuous(breaks=c(2.0,2.5,3.0), limits=c(2.0,3.2))
@@
 plot_fig3 <- plot_grid(Macs, LSs, ncol=2, rel_widths = c(1,0.9)) 
-plot_fig3_sub <- add_sub(plot_fig3, expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")"))), vjust = 0, size=12)
+plot_fig3_sub <- add_sub(plot_fig3, expression(paste("Specialization ", bgroup("(",frac(a[ii],a[1,j]+a[2,j]),")"))), vjust = 0, size=12)
 
@@
 
-```{r plot_fig3, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:plot_fig3}**Effect of character displacement on food-web dynamics under different evolutionary tradeoffs and foraging scenarios.** Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in stability panels).
+```{r plot_fig3, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:plot_fig3}**Effect of character displacement on food-web dynamics under different evolutionary tradeoffs and foraging scenarios.** Lines show predicted values when both consumers and resources are present.
+Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~, whereas as small shapes correspond to the starting points (only in stability panels).
 In both foraging scenarios, feeding rates increase linearly with resource abundance, but the equation for effective attack rate is different."}
@@
 \begin{equation}
-  \hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e-hm}
+  \hat{R}=\frac{1}{w_{1,j}a_{1,j}+w_{2,j}a_{2,j}}\cdot\frac{m}{e-hm}
 \end{equation}
 
-And since evolution favors consumer divergence onto their preferred resources (see Appendix S1), the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase, resulting in lower resource abundances and decreased food-web stability (Appendix S4).
-
-In firs two foraging scenarios, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$, see Appendix S2-3). 
+And since evolution favors consumer divergence onto their preferred resources (see Appendix S1), the effective attack rate of consumers ($w_{1,j}a_{1,j}+w_{2,j}a_{2,j}$) will always increase, resulting in lower resource abundances and decreased food-web stability (Appendix S4).
+
+In first two foraging scenarios, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $-\lambda_{max}>0$, see Appendix S2-3).
 In this more realistic model, however, whether the food web is locally stable depends on consumer and resource parameters. 
@@
 \begin{equation}
-  w_{ii}a_{ii}+(1-w_{ii})a_{ij} > \frac{e+hm}{hK(e-hm)}
+  w_{1,j}a_{1,j}+w_{2,j}a_{2,j} > \frac{e+hm}{hK(e-hm)}
 \end{equation}
@@
   ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*": max and min)"))) +
-  xlab(expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")")))) +
+  xlab(expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1,j]+a[2,j]),")")))) +
   #guides(fill = guide_legend(override.aes=list(shape=21))) +

### 2019-10-30 morgue/ECD_Supp_Mat.Rmd
committing a bunch of changes that mainly organize the repo and delete old material.

--- 2018-08-12 ECD_Supp_Mat.Rmd
@@
 ---
-title: "Predicting the effects of character displacement on food-web dynamics"
-author: "Matt Barbour"
+title: Predicting the effects of character displacement on food-web dynamics
+#author: "Matt Barbour"
+#output:
+#  pdf_document:
+#    df_print: paged
+#    fig_caption: true
+   # citation_package: natbib
+#    includes:  
+#      in_header: preamble-latex.tex
+#bibliography: references_ECD_supp_mat.bib
+#fontsize: 12pt
+
+#title: Phenotypic evolution is more constrained in simple food webs
+#  | Food-web complexity simplifies the adaptive landscape of an insect herbivore
+#  | Phenotypic evolution is less constrained in complex food webs
+#  | 
+#  | Food-web simplification constrains phenotypic evolution
+#  | Food-web complexity smooths the adaptive landscape of an insect herbivore
+author: 
+- Matthew A. Barbour$^{1,2,\ast}$
+# potentially use for 'date' for affiliations, as recommended by https://github.com/rstudio/bookdown/issues/348 for cross-format compatibility, with | to create new sections.
+#date: | 
+#    | 1. Affiliation
 output:
   pdf_document:
+    keep_tex: true
+    toc: false
+    number_sections: false
     df_print: paged
     fig_caption: true
-    citation_package: natbib
+    #citation_package: natbib
+    template: Rmd_AmNat_MS_template.tex
     includes:  
-      in_header: preamble-latex.tex
+      in_header: Am-Nat-preamble-latex.tex
 bibliography: references_ECD_supp_mat.bib
-fontsize: 12pt
+biblio-style: amnat.bst
+fontsize: 11pt
+fontfamily: mathpazo
+documentclass: article
+linkcolor: black
+urlcolor: black
 ---
 
-Keywords: competition; eco-evolutionary dynamics; ecological speciation; three-spine stickleback; zooplankton
-
-## Abstract
-
-Ecological character displacement is a key driver of diversification and adaptive radiation; however, there has been little work examining the consequences of character displacement on the dynamics of ecological communities.
-Given the growing recognition for the role of evolution in structuring ecological communities, understanding how major evolutionary processes impact ecological dynamics is increasingly important.
-Here, we build a mathematical model to predict how exploitative competition between two consumers for two resources affects the evolution of consumer attack rates, and, in turn, consumer-resource population dynamics.
-We then compared these predictions to results from a natural system, where a fish consumer (three-spine stickleback) has become locally adapted to different aquatic habitats with different resources (benthic vs.
-limnetic zones).
-Our model suggested that increasing ecological character displacement decreases the equilibrium density of resources, and with sufficient divergence, results in more oscillatory (destabilizing) consumer-resource dynamics.
-In support of this prediction, we found that the biomass density of zooplankton (resource) in lakes with two species of stickleback was only a fraction of that found in lakes with one species, varying from 4 to 41%.
-Moreover, we found that the biomass density of zooplankton was 2-fold more variable (CV) in lakes with two vs.
-one species of stickleback.
-These results suggest that ecological character displacement can eventually have a negative feedback on food-web persistence in the absence of other processes.
-
-## Introduction
-
-Ecological character displacement (ECD) is thought to be a key evolutionary process in the diversification of species and consequently the generation of biodiversity (Schluter 2000; Pfennig and Pfennig 2010; Stuart and Losos 2013).
-ECD describes the “process of phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” (Schluter 2000).
-Over the past 40 years, a large body of theoretical (e.g.
-Lawlor and Smith 1976; Abrams 1986; Doebeli 1996; Taper and Chase 1985) and empirical (reviewed in: Schluter 2000; Dayan and Simberloff 2005; Stuart and Losos 2013) work has been generated to understand the scenarios under which exploitative competition for resources leads to the divergence of consumer foraging traits.
-Yet, although the process of ecological character displacement is driven by the interplay between competition and trait change, and hence is strongly dependent on resource availability, little of the work to date has explicitly considered how character displacement impacts other species within the community.
-Similarly, research on food-web dynamics has largely ignored the potential role of evolution in altering interactions and stability (Fussmann et al.
-2007).
-Given the emerging evidence that evolutionary processes can drive ecological dynamics (Hairston et al.
-2005, Post et al.
-2008, Harmon et al.
-2009, Schoener 2011, Rudman et al.
-2015), understanding the ecological consequences of character displacement is a natural first step in investigating how evolutionary processes affect food-web dynamics.
-
-Ecological character displacement has typically been considered to be a stabilizing factor in communities (Lawlor and Smith 1976).
-This is because divergence in foraging traits is thought to reduce the strength of interspecific competition between consumers.
-Reducing the strength of interspecific competition relative to intraspecific competition is a stabilizing force in consumer-resource systems.
-Consumer specialization on distinct resources suggests that there are niche differences between consumers.
-Mathematically, this is also often assumed to be the case, because using techniques such as evolutionary invasion analysis describe the ability for a consumer with a mutant allele to increase in abundance when rare, a demographic signature of niche differentiation.
-Therefore, if both consumers are able to invade when rare, then this suggests that they can coexist.
-On the other hand, relatively specialized consumers have also been shown to be a strongly stabilizing force in food webs (McCann et al.
-2005).
-This is because these more specialized consumers may generate asynchrony in the dynamics of patchily distributed resources.
-This asynchrony in resources then permits consumers to move to habitat patches with higher resource densities after they have depleted the resources in a different patch.
-If, for example, the consumers were not specialized, then the resources would exhibit the same dynamics, so when the resources at low densities, consumers would be unable to switch to alternative resources, which would cause the consumer populations to crash, and oscillate more with their resources (McCann et al.
-2005).
-Consequently, whether ecological character displacement will tend to stabilize (via asynchronous resource dynamics) or destabilize (via increased interaction strength) food webs is unclear.
-
-On the other hand, ecological character displacement often results in the evolution of characters that enhance a consumer’s ability to capture resources, which tends to destabilize consumer-resource interactions (Murdoch et al.
-2003; McCann 2011).
-
-This occurs because increases in a consumer’s attack rate results in the resource being suppressed well below its carrying capacity, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics (Rosenzweig 1971; Murdoch et al.
-2003).
-Experimental decreases in interaction strength between consumers and resources has been shown to promote coexistence between consumers and resources (Luckinbill 1973).
-Moreover, there is broad support that specialized consumers, since they are tightly coupled to their resources, tend to exhibit consumer-resource cycles, whereas this is not the case for generalists (Murdoch et al.
-2002).
-Therefore, although evolutionary responses to exploitative competition may reduce the strength of interactions between consumers, they ultimately increase the strength of interactions with their resources which can destabilize the direct consumer-resource interaction (rather than the indirect consumer-consumer interaction).
-While prior models of character displacement have included more realistic functional responses and spatial relationships between consumers and resources, they have focused more on whether these factors alter patterns of character displacement rather than the effects of character displacement on the stability of the ecological community.
-As a consequence, it is currently unclear whether ecological character displacement has a net stabilizing or destabilizing effect on ecological communities.
-
-Here, we address this knowledge gap by building a mathematical model to examine how ecological character displacement affects consumer-resource dynamics.
-We then compared the predictions from our model to field data from one of the most well-studied examples of ecological character displacement in a natural system, threespine stickleback in lakes of southern British Columbia.
-We specifically sought to address two questions: (1) How does exploitative competition between two consumers for two resources affect the evolution of consumer attack rates?
-(2) How does the evolution of consumer attack rates affect community stability?
+<!-- Begin title page formatting -->
+
+\noindent 1. University of British Columbia, Department of Zoology, Vancouver, BC V6T 1Z4, Canada;
+
+\noindent 2. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;
+
+$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch
+
+\bigskip
+
+*Manuscript elements*: Figure 1, figure 2, figure 3, figure 4. All figures should be printed in color.
+
+\bigskip
+
+*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptive radiation; ecological community
+
+\bigskip
+
+*Manuscript type*: Note.
+
+\bigskip
+
+\footnotesize Prepared using an *Am. Nat.* inspired \LaTeX{} template for Rmarkdown. \normalsize
+
+<!-- End title page formatting -->
+
+\linenumbers{}
+\modulolinenumbers[3]
+
+\newpage
+
+<!-- Outline
+
+Introduction
+
+1. ECD is an important process, and current empirical and theoretical work suggests it often results in divergence of consumer foraging traits.
+
+2.
+Ecological consequences of character displacement have received surprisingly little attention.
+In particular, prior work doesn't allow a proper comparision of the effects of character displacement on community stability.
+
+-> Another paragraph illustrating how the direct and indirect effects of character displacement likely have constrasting effects on stability.
+
+3.
+Here we fill this knowledge gap.
+We focus on the ecological side of these dynamics, because the evolutionary aspects have been well characterized, although we show in the appendices that the evolutionary dynamics are as expected (especially for McCann model).
+
+Methods and Results
+
+- Underlying model of consumer-resource dynamics
+  1. Explanation of model parameters
+  2. Methods for assessing resource abundances and stability.
+  3. Trade-off function and definition of character displacement.
+  4. Analytical and simulation methods. Briefly mention the different scenarios I'll explore.
+
+- Consumers can forage for resources simultaneously
+  1. Functional response equation and illustration of assumptions
+  2. Effect on attack rates, resource densities, and stability
+
+- Consumers cannot foraging for resource simultaneously
+  1. Functional response equation and illustration of assumptions
+  2. Effect on attack rates, resource densities, and stability
+
+- McCann's most realistic functional response
+  1. Functional response equation and illustration of assumptions
+  2. Effect on attack rates, resource densities, and stability
+
+Discussion
+
+-->
 
@@
 
+# Abstract
+
+<!--
+Ecological character displacement is a key driver of diversification and adaptive radiation; however, there has been little work examining the consequences of character displacement on the dynamics of ecological communities.
+Given the growing recognition for the role of evolution in structuring ecological communities, understanding how major evolutionary processes impact ecological dynamics is increasingly important.
+Here, we build a mathematical model to predict how exploitative competition between two consumers for two resources affects the evolution of consumer attack rates, and, in turn, consumer-resource population dynamics.
+Our model suggested that increasing ecological character displacement decreases the equilibrium density of resources, and with sufficient divergence, results in more oscillatory (destabilizing) consumer-resource dynamics.
+These results suggest that ecological character displacement can eventually have a negative feedback on food-web persistence in the absence of other processes.
+-->
+
+\newpage
+
+# Introduction
+
+<!-- ECD as an important eco-evolutionary feedback -->
+
+Ecological character displacement is thought to be a key evolutionary process generating biodiversity [@Schluter2000; @Pfennig2010; but see @Stuart2013].
+Ecological character displacement describes the “process of phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000].
+Over the past 40 years, a large body of theoretical [e.g.
+@Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has been generated to understand the scenarios under which exploitative competition for resources leads to the divergence of consumer foraging traits.
+The general result emerging from these studies is that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the divergence of consumer foraging traits [@Lawlor1976; @Taper1985).
+This process is not simply driven by ecological differences, but creates an eco-evolutionary feedback that drives further differentiation.
+This insight was clearly by theoretical models that highlighted the importance of explicitly including resource dynamics as a mediator of competition in driving evolutionary change [@Lawlor1976; @Abrams1986; @Taper1985].
+
+<!--Yet, although resources mediate this eco-evolutionary feedback between consumers, the consequences of character displacement on consumer-resource interactions and food-web dynamics has received, surprisingly, little attention.-->
+Although character displacement is the result of an eco-evolutionary feedback, the ecological consequences of this process have received surprisingly little attention [@Germain2018].
+Perhaps this is because early theoretical work suggested that character displacement was a potent stabilizing force in ecological communities [@Lawlor1976].
+This ecological outcome is also an intuitive one ---character divergence reduces interspecific competition between consumers, promoting coexistence.
+However, inferences about the effect of character displacement, like the process itself, can only be made when comparing communities in the presence and absence of competition [criteria #3 for character displacement, @Schluter1992; @Schluter2000].
+Current inferences about the stabilizing effects of consumer divergence are only comparisons about communities where there are two consumers, and even though, the focus is only on the stable coexistence of consumer species.
+
+The consequence of this eco-evolutionary feedback for consumer coexistence has been well characterized (Lawlor and Smith 1976; Germain's work).
+Ecological character displacement has typically been considered to be a stabilizing factor in communities (Lawlor and Smith 1976).
+This is because divergence in foraging traits is thought to reduce the strength of interspecific competition between consumers.
+Reducing the strength of interspecific competition relative to intraspecific competition is a stabilizing force in consumer-resource systems.
+Consumer specialization on distinct resources suggests that there are niche differences between consumers.
+Mathematically, this is also often assumed to be the case, because using techniques such as evolutionary invasion analysis describe the ability for a consumer with a mutant allele to increase in abundance when rare, a demographic signature of niche differentiation.
+Therefore, if both consumers are able to invade when rare, then this suggests that they can coexist.
+
+Character displacement involves evolutionary shifts in character states in response to competition [@Schluter1992; @Schluter2000].
+Theoretically, this is determined by comparing character changes in the presence and absence of a competitor.
+Empirically, this is determined by meeting several criteria that rule out other possible explanations [@Schluter1992; @Schluter2000].
+Importantly, current inferences about the effect of character displacement on community stability do not meet this critera.
+For example, the inference that character displacement enhances species coexistence is only made in a scenario where consumers are competing.
+This type of stability is important, but it does not enable us to examine the effect of character displacement on community stability.
+As a consequence, I argue that the effect of character displacement on community stability remains an open question.
+
+Indeed, ecological character displacement often results in the evolution of characters that enhance a consumer’s ability to capture resources, which tends to destabilize consumer-resource interactions (Murdoch et al.
+2003; McCann 2011).
+This occurs because increases in a consumer’s attack rate results in the resource being suppressed well below its carrying capacity, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics (Rosenzweig 1971; Murdoch et al.
+2003).
+Experimental decreases in interaction strength between consumers and resources has been shown to promote coexistence between consumers and resources (Luckinbill 1973).
+Moreover, there is broad support that specialized consumers, since they are tightly coupled to their resources, tend to exhibit consumer-resource cycles, whereas this is not the case for generalists (Murdoch et al.
+2002).
+Therefore, although evolutionary responses to exploitative competition may reduce the strength of interactions between consumers, they ultimately increase the strength of interactions with their resources which can destabilize the direct consumer-resource interaction (rather than the indirect consumer-consumer interaction).
+While prior models of character displacement have included more realistic functional responses and spatial relationships between consumers and resources, they have focused more on whether these factors alter patterns of character displacement rather than the effects of character displacement on the stability of the ecological community.
+As a consequence, it is currently unclear whether ecological character displacement has a net stabilizing or destabilizing effect on ecological communities.
+
+Here, I address this knowledge gap by building a mathematical model to examine how ecological character displacement affects consumer-resource dynamics in a food web context.
+I address two question: (1) How does ecological character displacement affect resource abundances?
+(2) How does character displacement affect food-web stability?
+I show that the consequences of character displacement on food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
+
+<!--
+Another key insight emerging from past work is that character displacement is not simply driven by ecological differences, but also drives further ecological differentiation (cite?).
+This insight emerged from theoretical work that explicitly accounted for resource dynamics when modeling character displacement.
+If, for example, resources are treated as constants, the optimal strategy for each consumer species is unaffected by the presence or absence of a competitor.
+
+This is because character divergence reduces interspecific competition relative to intraspecific competition, facilitating consumer coexistence.
+
+In other words, ecological character displacement is the result of an eco-evolutionary feedback between consumers. -->
+
+<!-- ECD on consumer coexistence -->
+
+<!-- ECD effect on consumer-resource interactions -->
+
+<!-- Ecological character displacement is not simply driven by ecological differences but also drives further ecological differentiation.
+In other words, ecological character displacement embodies an eco-evolutionary feedback between consumers.
+Character divergence in foraging traits also has ecological consequences, in that it promotes stable coexistence between competing consumers.
+
+Yet, although the process of ecological character displacement is driven by consumers competing for shared resources (i.e.
+exploitative competition), the consequences of character displacement on consumer-resource interactions and food-web dynamics has received, suprisingly, little attention.
+-->
+
+<!-- the interplay between competition and trait change, and hence is strongly dependent on resource availability, little of the work to date has explicitly considered how character displacement impacts other species within the community.
+Similarly, research on food-web dynamics has largely ignored the potential role of evolution in altering interactions and stability (Fussmann et al.
+2007).
+Given the emerging evidence that evolutionary processes can drive ecological dynamics (Hairston et al.
+2005, Post et al.
+2008, Harmon et al.
+2009, Schoener 2011, Rudman et al.
+2015), understanding the ecological consequences of character displacement is an important step in investigating how evolutionary processes affect food-web dynamics.
+-->
+
+<!-- Ecological character displacement has typically been considered to be a stabilizing factor in communities (Lawlor and Smith 1976).
+This is because divergence in foraging traits is thought to reduce the strength of interspecific competition between consumers.
+Reducing the strength of interspecific competition relative to intraspecific competition is a stabilizing force in consumer-resource systems.
+Consumer specialization on distinct resources suggests that there are niche differences between consumers.
+Mathematically, this is also often assumed to be the case, because using techniques such as evolutionary invasion analysis describe the ability for a consumer with a mutant allele to increase in abundance when rare, a demographic signature of niche differentiation.
+Therefore, if both consumers are able to invade when rare, then this suggests that they can coexist.
+On the other hand, relatively specialized consumers have also been shown to be a strongly stabilizing force in food webs (McCann et al.
+2005).
+This is because these more specialized consumers may generate asynchrony in the dynamics of patchily distributed resources.
+This asynchrony in resources then permits consumers to move to habitat patches with higher resource densities after they have depleted the resources in a different patch.
+If, for example, the consumers were not specialized, then the resources would exhibit the same dynamics, so when the resources at low densities, consumers would be unable to switch to alternative resources, which would cause the consumer populations to crash, and oscillate more with their resources (McCann et al.
+2005).
+Consequently, whether ecological character displacement will tend to stabilize (via asynchronous resource dynamics) or destabilize (via increased interaction strength) food webs is unclear.
+
+On the other hand, ecological character displacement often results in the evolution of characters that enhance a consumer’s ability to capture resources, which tends to destabilize consumer-resource interactions (Murdoch et al.
+2003; McCann 2011).
+This occurs because increases in a consumer’s attack rate results in the resource being suppressed well below its carrying capacity, which if sufficient enough, can generate oscillations, and hence less stable, consumer-resource dynamics (Rosenzweig 1971; Murdoch et al.
+2003).
+Experimental decreases in interaction strength between consumers and resources has been shown to promote coexistence between consumers and resources (Luckinbill 1973).
+Moreover, there is broad support that specialized consumers, since they are tightly coupled to their resources, tend to exhibit consumer-resource cycles, whereas this is not the case for generalists (Murdoch et al.
+2002).
+Therefore, although evolutionary responses to exploitative competition may reduce the strength of interactions between consumers, they ultimately increase the strength of interactions with their resources which can destabilize the direct consumer-resource interaction (rather than the indirect consumer-consumer interaction).
+While prior models of character displacement have included more realistic functional responses and spatial relationships between consumers and resources, they have focused more on whether these factors alter patterns of character displacement rather than the effects of character displacement on the stability of the ecological community.
+As a consequence, it is currently unclear whether ecological character displacement has a net stabilizing or destabilizing effect on ecological communities.
+
+Here, we address this knowledge gap by building a mathematical model to examine how ecological character displacement affects consumer-resource dynamics.
+We then compared the predictions from our model to field data from one of the most well-studied examples of ecological character displacement in a natural system, threespine stickleback in lakes of southern British Columbia.
+We specifically sought to address two questions: (1) How does exploitative competition between two consumers for two resources affect the evolution of consumer attack rates?
+(2) How does the evolution of consumer attack rates affect community stability?
+-->
+
+# Methods and Results
+
 ## Model of resource competition
 
-I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
+To examine how ecological character displacement affects consumer-resource interactions and food-web dynamics, I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
 
@@
 $\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).
-I show that the consequences of character displacement on food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.
-
-## Effect of character displacement on resource densities
-
-I start by examing the effects of character displacement on the density of resources at equilibrium, which will also set the stage for a later analysis of food-web stability.
+<!--I show that the consequences of character displacement on food-web dynamics are not intuitive and depend critically on the foraging scenario and tradeoff in consumer attack rates.-->
+
+## Resource densities
+
+I start by examing the effects of character displacement on the density of resources at equilibrium, which will set the stage for a later analysis of food-web stability.
 In particular, I examine two different foraging scenarios: one where consumers forage for both resources simultaneously (e.g.
@@
 @Lawlor1976; @McCann2005).
-I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, defined as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
-
-### Consumers forage for both resources simultaneously
-
-An inherent assumption in MacArthur's model of resource competition is that consumers can forage for both resources simultaneously.
+I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, defined as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|>0$, when evolving with a competing consumer.
+
+**Consumers forage for both resources simultaneously**: An inherent assumption in MacArthur's model of resource competition is that consumers can forage for both resources simultaneously.
 In essence, this implies that both resources occur within the same habitat.
@@
 
-Mac_resources <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
+Mac_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
   ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
@@
 
-### Consumers *cannot* forage for resources simultaneously
-
-Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (e.g.
+**Consumers *cannot* forage for both resources simultaneously**: Most models of character displacement have assumed a scenario where consumers can forage for both resources simultaneously (e.g.
 @Taper1985; @Abrams1986).
@@
 
-## Effect of character displacement on food-web stability
+## Food-web stability
 
@@
 Small and large points correspond to initial and evolved strategies, respectively."}
-Stability_Mac <- cbind.data.frame(tradeoff_df, MacArthur_2C_2R_tradeoff_dynamics) %>%
+Stability_Mac <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
   ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
@@
 
-Stability_LS <- cbind.data.frame(tradeoff_df, LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+Stability_LS <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
   ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
@@
 
+# Discussion
+
+Important point: Criteria 5 in @Schluter1992 does not make sense since character displacement usually alters resource distributions. Think about this more.
+
 ## Predictions for threespine stickleback
@@
 It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected (@Abrams1987; @Fox2008).
+
+## Acknowledgement
+
+The content of this paper was greatly enhanced by numerous conversations with Seth Rudman.
+Sally Otto provided much help in early analyses of the mathematical model and I would not have been able to take the theory as far as I did without her guidance and her pushing me.
+
+Thank Ben Gilbert for his comment during a discussion of the model that pushed me to take a comparative approach that looked across different model structures to examine the generality of these results.
 

### 2019-10-30 confirm_asymmetry_effects.R

@@ -1,0 +1,9 @@
+test <- read.csv("asymmetric_displacement_data.csv")
+
+# inspecting the data shows that the starting values are being appropriately switched. Also, we see the same qualitative pattern of displacement in the species that starts as a generalist (becomes a specialist).
+select(test, Model, Competitor, sequence, Time, n, a11:a22, w11:w22, R1:C2, max.Re.eigen, max.Im.eigen) %>%
+  mutate(a11=round(a11,3),
+         a12=round(a12,3),
+         a21=round(a21,3),
+         a22=round(a22,3)) %>%
+  arrange(Model, n, Competitor, Time)

### 2019-10-30 inspect_model_effects.R

@@ -1,0 +1,17 @@
+library(tidyverse)
+test <- read.csv("symmetric_displacement_data.csv")
+
+select(test, Model, Competitor, Time, sequence, feasibility, n, a11:a22, w11:w22, R1:C2) %>%
+  mutate(a11=round(a11,3),
+         a12=round(a12,3),
+         a21=round(a21,3),
+         a22=round(a22,3),
+         R1=round(R1,2),
+         R2=round(R2,2),
+         C1=round(C1,2),
+         C2=round(C2,2)) %>%
+  arrange(Model, n, Competitor, Time)
+
+filter(test, Model=="MacArthur")
+filter(test, Model=="LawlorSmith")
+filter(test, Model=="McCann")

### 2019-10-30 manuscript_ECD_model.Rmd
final edits to main text for bioRxiv submission

@@
 
-To examine how ecological character displacement affects resource abundances and food-web stability, I analyzed a continuous-time model of two consumers ($C_1,C_2$) competing for two shared resources ($R_1,R_2$):
+To examine how ecological character displacement affects resource abundances and food-web stability, I analyzed a continuous-time model of two consumers ($C_{j=1,2}$) competing for two shared resources ($R_{i=1,2}$):
 
@@
   \begin{split}
-     & \frac{dR_1}{dt}=r_1R_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2 \\
-     & \frac{dR_2}{dt}=r_2R_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2 \\
-     & \frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1 \\
-     & \frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2 \\
+     & \frac{dR_1}{dt}=r_1R_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{12}(R_1)C_2 \\
+     & \frac{dR_2}{dt}=r_2R_2(1-\frac{R_2}{K_2})-F_{21}(R_2)C_1-F_{22}(R_2)C_2 \\
+     & \frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{21}F_{21}(R_2)C_1-m_1C_1 \\
+     & \frac{dC_2}{dt}=e_{12}F_{12}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2 \\
   \end{split}
@@
 seed size, see @Taper1985 for an example).
+Importantly, inferences about ecological character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992].
+Therefore, I arbitrarily set $C_2=0$ to create a food-web without a competing consumer for these comparisons.
 
@@
 Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach, with the general result being character divergence [@Lawlor1976; @Abrams1986].
-I say consumers have undergone character divergence if there evolved attack rates are more specialized, measured as $|\frac{a_{ij}}{a_{1,j}+a_{2,j}-0.5|$, when evolving with a competing consumer.
-I also used an Adaptive Dynamics approah to analyze character displacement in the third foraging scenario, and I too observed character divergence (detailed analysis in Appendix S1 of Supplementary Information).
+I also used an Adaptive Dynamics approach to analyze character displacement in the third foraging scenario, and I too observed character divergence (detailed analysis in Appendix S1 of Supplementary Information).
+I say consumers have undergone character divergence if their evolved attack rates are more specialized when evolving with vs. without a competing consumer. 
+Specialization of consumer $j$ on resource $1$ is measured as $\frac{a_{1,j}}{a_{1,j}+a_{2,j}}$, where a value of 0.5 is a complete generalist ($a_{1,j}=a_{2,j}$), and a value of 1 is a complete specialist ($a_{2,j}=0$).
+<!--on resource 1, and a value of 0 is a complete specialist on resource 2.-->
+Values less than 0.5 indicate specialization on the other resource.
 
@@
 To do this, I analyzed differences in resource abundances and food-web stability at equilibrium. 
-An equilibrium is reached when the rates of change in equation \ref{eq:1} are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and each consumer ($\hat C_1,\hat C_2$).
-Importantly, inferences about ecological character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992].
-Therefore, I compared resource abundances between food webs with ($\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$) and without ($\hat R_1,\hat R_2, \hat C_1 > 0$) a competing consumer.
+An equilibrium is reached when the rates of change in equation \ref{eq:1} are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_i$) and each consumer ($\hat C_j$).
+<!--Importantly, inferences about ecological character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992].
+Therefore, I compared resource abundances between food webs with ($\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$) and without ($\hat R_1,\hat R_2, \hat C_1 > 0$) a competing consumer.-->
 I also compared the local stability of these food webs using standard methods [@Otto2007]. 
-This stability analysis derives the dominant eigenvalue, $\lambda_{max}$, of the matrix of partial derivatives of all differential equations with respect to all variables evaluated at equilibrium. 
+This stability analysis derives the dominant eigenvalue, $\lambda_{\text{max}}$, of the matrix of partial derivatives of all differential equations with respect to all variables evaluated at equilibrium.
 If $-\lambda_{max}>0$, then the food web will return to equilibrium after a small perturbation (i.e. locally stable), with more positive values indicating a faster return time.
@@
 save_plot(filename = "attack_tradeoffs.pdf", plot = tradeoff_plot, 
-          base_height = 5, # 6
-          base_width = 7) #9
+          base_height = 4, # 7
+          base_width = 6) # 7
 ```
@@
 
-\begin{equation}
+\begin{equation} \label{eq:5}
   \hat{R}=\frac{1}{a_{1,j}+a_{2,j}}\cdot\frac{m}{e}
@@
 ## Resources occur in different habitats
+
 <!--
@@
 -->
+
 In the second foraging scenario (equation \ref{eq:3}), I again see that resource abundances are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), but are now determined by the following equation (derived in Appendix S3):
 
-\begin{equation}
+\begin{equation} \label{eq:6}
   \hat{R}=\frac{1}{w_{1,j}a_{1,j}+w_{2,j}a_{2,j}}\cdot\frac{m}{e}
@@
 \ref{fig:plot_fig3} right panels).
+<!--This is because the ESS for a single consumer is to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in fig.
+\ref{fig:plot_fig3} right panels), resulting in higher resource abundances compared to if it evolved to be a complete generalist.-->
 
@@
 Macs <- plot_grid(
-  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1,j]+a[2,j]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), #  + scale_y_continuous(breaks=c(1.8,2.0,2.2))
+  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), #  + scale_y_continuous(breaks=c(1.8,2.0,2.2))
   Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)), # + scale_y_continuous(breaks=c(1.0,1.2,1.4), limits=c(1.0,1.45))
@@
 LSs <- plot_grid(
-  LS_attack  + xlab("") + ylab(bquote(w[1,j]*a[1,j]+w[2,j]*a[2,j])) + ggtitle("Resources: Different Habitats\n") +  theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), #+ scale_y_continuous(breaks=c(0.8,1.0,1.2))  # "("*1-w[1,j]*")"
+  LS_attack  + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])) + ggtitle("Resources: Different Habitats\n") +  theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), #+ scale_y_continuous(breaks=c(0.8,1.0,1.2))  # "("*1-w[1,j]*")"
   LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()), #+ scale_y_continuous(breaks=c(2.0,2.5,3.0), limits=c(2.0,3.2))
@@
 plot_fig3 <- plot_grid(Macs, LSs, ncol=2, rel_widths = c(1,0.9)) 
-plot_fig3_sub <- add_sub(plot_fig3, expression(paste("Specialization ", bgroup("(",frac(a[ii],a[1,j]+a[2,j]),")"))), vjust = 0, size=12)
+plot_fig3_sub <- add_sub(plot_fig3, expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","* j]),")"))), vjust = 0, size=12)
 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:7}
   \hat{R}=\frac{1}{w_{1,j}a_{1,j}+w_{2,j}a_{2,j}}\cdot\frac{m}{e-hm}
@@
 
-In first two foraging scenarios, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $-\lambda_{max}>0$, see Appendix S2-3).
+In the first two foraging scenarios, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equilibrium (because $-\lambda_{max}>0$, see Appendix S2-3).
 In this more realistic model, however, whether the food web is locally stable depends on consumer and resource parameters. 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:8}
   w_{1,j}a_{1,j}+w_{2,j}a_{2,j} > \frac{e+hm}{hK(e-hm)}
@@
 \ref{fig:LS_Plot}A).-->
-Thus, character displacement always pushes the food web toward an unstable stucture in this more realistic foraging scenario (fig.
+Thus, character displacement always pushes the food web toward an unstable structure in this more realistic foraging scenario (fig.
 \ref{fig:bifur_plot}).
@@
   #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(bquote(atop("Effective attack rate","("*w*a[ii]+"("*1-w*")"*a[ij]*")"))) +
+  ylab(bquote(atop("Effective attack rate","("*w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j]*")"))) +
   #ylab(expression(w*a[ii]+(1-w)*a[ij])) +
@@
   geom_point(data = McCann_displacements, aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor, size=Time), color="white", inherit.aes = FALSE) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  xlab(expression(frac(a[ij],a[1*","*j]+a[2*","*j]))) +
   ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) +
@@
 
-plot_McCann_model <- add_sub(plot_McCann_model, label = expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")"))))
+plot_McCann_model <- add_sub(plot_McCann_model, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
 
@@
   ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*": max and min)"))) +
-  xlab(expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1,j]+a[2,j]),")")))) +
+  xlab(expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")")))) +
   #guides(fill = guide_legend(override.aes=list(shape=21))) +
@@
 
-## Resource abundances
+## Resource Abundances
 
@@
 Rather than studying resilience, coexistence theory usually studies the mutual ability of consumers with different phenotypes to invade when rare [mutual invasibility, @Chesson2000].
-In the context of character displacement, a shortcoming of this mutual invasibility measure is that it does not allow a comparison between food webs with and without a competiting consumer. 
+In the context of character displacement, a shortcoming of this mutual invasibility measure is that it does not allow a comparison between food webs with and without a competing consumer. 
 Such comparisons are necessary for inferring the effects of character displacement, a point that has been made clear in the criteria to demonstrate character displacement [@Schluter1992; @Schluter2000].
@@
 Note that this does not negate my general conclusion that ecological character displacement decreases resource abundances and stability; however, this process may itself create another eco-evolutionary feedback between consumers and resources.
-Thus, not coevolution itself, but the prey's evolutionary response would act to stabilize food webs.
-This may actually help maintain fantastic examples of character divergence and prevent them from destabilizing systems, basically because the consumers trait becomes decoupled with their attack rate.
-Examining this role would be ideal in a quantitative genetic setting that can explicitly track trait dynamics, but it would not fundamentally change the conclusions presented here.
+<!--Therefore, the prey's evolutionary response would act to stabilize food webs, rather than coevolution *per se*.-->
+This may actually help maintain dramatic examples of character divergence and prevent them from destabilizing systems, because it allows consumer traits to become decoupled from their attack rate.
+Examining this decoupling would be ideal in a quantitative genetic model that can explicitly track trait dynamics, but it would not fundamentally change the conclusions presented here.
 
 Another potential caveat is that I explored my model in a setting that makes many assumptions about resource and consumer symmetry (but see consumer asymmetry section). 
-Prior work has shown that allowing for resource asymmetry, for example, may decrease the magnitude of character displacement. 
+Prior work has shown that allowing for resource asymmetry, for example, may decrease the magnitude of character displacement [@Abrams1986]. 
 This is because this asymmetry creates an asymmetry in resource abundances, which dampens the effects of character divergence.
@@
 This is because models that explicitly include resource dynamics inevitably show that resource competition results in character divergence, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
-A quantitative genetic model may certainly show differences in the magnitude of character divergence, but this should not qualtitatively change its effect on food-web dynamics.
+A quantitative genetic model may certainly show differences in the <!--magnitude-->pace of character divergence, but this should not qualitatively change its effect on food-web dynamics.
 <!-- In fact, recent work by @McPeek2017 has shown that divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
 -->It is important to note that my conclusions only apply to food webs with biotic resources that are nutritionally substitutable.
-It would be interesting to extend these current analyses to non-substitutable resources where chracter convergence is expected [@Abrams1987; @Fox2008].
+It would be interesting to extend these current analyses to non-substitutable resources where character convergence is expected [@Abrams1987; @Fox2008].
 
@@
 
-# Supplementary Material
-
 ```{r MacArthur_asymm, include=FALSE}
@@
   #geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
+  ylab(bquote(atop("Effective attack rate","("*w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j]*")"))) +
+  xlab(expression(frac(a[ii],a[1*","*j]+a[2*","*j]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+McCann_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
+  #geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
 
-McCann_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(R[i]+R[j])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
 Stability_McCann_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
@@
   geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
+  xlab(expression(frac(a[ij],a[1*","*j]+a[2*","*j]))) +
+  ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
   scale_shape_manual(values=c(24,21))  +
@@
 
-McCann_plots_asymm <- plot_grid(McCann_attack_asymm   + theme(legend.position = "none"), 
-                             McCann_resources_asymm   + theme(legend.position = "none"), 
-                             Stability_McCann_asymm   + theme(legend.position = "none"), 
-                       ncol=2, labels = "AUTO", align = 'hv')
+McCann_plots_asymm <- plot_grid(McCann_attack_asymm   + theme(legend.position = "none") + xlab(""), 
+                             McCann_resources_asymm   + theme(legend.position = "none") + xlab(""), 
+                             Stability_McCann_asymm   + theme(legend.position = "none") + xlab(""), 
+                       ncol=2, align = 'hv')
 McCann_plots_asymm
 
-save_plot(filename="Figure_ECD_McCann_asymm.pdf", plot=McCann_plots_asymm, base_height = 6, base_width = 6)
+plot_McCann_asymm <- add_sub(McCann_plots_asymm, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
+
+save_plot(filename="Figure_ECD_McCann_asymm.pdf", plot=plot_McCann_asymm, base_height = 6, base_width = 6)
 # moved to supplementary information
@@
 
-all_asymm_plots <- plot_grid(
-  Mac_attack_asymm + ggtitle("Scenario I\n") + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[ii]+a[ij]*")"))), #, plot.title=element_text(size=8)
-  LS_attack_asymm  + ggtitle("Scenario II\n") + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(bquote(w[ii]*a[ii]+"("*1-w[ii]*")"*a[ij]*")")), #, plot.title=element_text(size=8)
-  McCann_attack_asymm + ggtitle("Scenario III")  + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(bquote(w[ii]*a[ii]+"("*1-w[ii]*")"*a[ij]*")")), #, plot.title=element_text(size=8), plot.subtitle=element_text(size=8)
-  Mac_resources_asymm   + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))),
-  LS_resources_asymm   + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(""),
-  McCann_resources_asymm   + theme(legend.position = "none", axis.text.x=element_blank()) + xlab("") + ylab(""),
-  Stability_Mac_asymm   + theme(legend.position = "none") + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))),
-  Stability_LS_asymm   + theme(legend.position = "none") + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(""),
-  Stability_McCann_asymm   + theme(legend.position = "none") + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(""),
-  ncol=3, align='v'#, rel_widths = c(1,0.9,0.9)
+Macs_asymm <- plot_grid(
+  Mac_attack_asymm + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))),
+  Mac_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))),
+  Stability_Mac_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))),
+  ncol=1, align='v'#, rel_widths = c(1,0.9,0.9)
 )
 
-all_asymm_plots <- add_sub(all_asymm_plots, label = expression(paste("Specialization ", bgroup("(",frac(a[ii],a[ii]+a[ij]),")"))))
-
-save_plot(filename="asymm_figs.pdf", plot=all_asymm_plots, base_height = 11, base_width = 8.5)
-```
+LSs_asymm <- plot_grid(
+  LS_attack_asymm  + ggtitle("Resources: Different Habitats\n") +  theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)) + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])),
+  LS_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x=element_blank()) + xlab("") + ylab(""),
+  Stability_LS_asymm   + theme(legend.position = "none", axis.text = element_text(size=10)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(""),
+  ncol=1, align='v'#, rel_widths = c(1,0.9,0.9)
+)
+
+all_asymm_plots <- plot_grid(Macs_asymm, LSs_asymm, ncol=2, rel_widths = c(1,0.9)) 
+all_asymm_plots <- add_sub(all_asymm_plots, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
+
+save_plot(filename="asymm_figs.pdf", plot=all_asymm_plots, base_height = 6.5, base_width = 6)
+```

### 2019-10-30 manuscript_ECD_model.tex
final edits to main text for bioRxiv submission

@@
 \documentclass[11pt,]{article}
-\usepackage{mathpazo}
+\usepackage[]{mathpazo}
 \usepackage{amssymb,amsmath}
@@
     \usepackage{mathspec}
-    \usepackage{xltxtra,xunicode}
   \else
@@
   \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
 \fi
@@
 }{}
-\usepackage{graphicx}
+\usepackage{hyperref}
+\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
+\hypersetup{unicode=true,
+            colorlinks=true,
+            linkcolor=black,
+            citecolor=black,
+            urlcolor=black,
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{natbib}
+\bibliographystyle{amnat}
+\usepackage{graphicx,grffile}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={Matthew A. Barbour\^{}\{1,2,\textbackslash{}ast\}},
-            pdftitle={Predicting the effects of character displacement on food-web dynamics},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=black,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 \setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
@@
 
-  \title{Predicting the effects of character displacement on food-web dynamics}
-    \author{Matthew A. Barbour\(^{1,2,\ast}\)}
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{}
+    \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+    \author{}
+    \preauthor{}\postauthor{}
     \date{}
+    \predate{}\postdate{}
 
@@
 %\usepackage[authoryear,sectionbib,sort]{natbib}
-\linespread{1.7}
+\linespread{1.5} % 1.7
 %\usepackage[utf8]{inputenc}
 \usepackage{lineno}
+
+% controlling text size in figure captions
+\usepackage{caption}
+\captionsetup[figure]{font=small}
 
@@
 
-\maketitle
+\vspace*{0.1cm}
 
-\noindent 1. University of British Columbia, Department of Zoology,
+\begin{center} \LARGE Ecological character displacement destabilizes food webs \end{center}
+
+\bigskip
+
+\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}
+
+\bigskip
+
+\noindent 1. University of Zurich, Department of Evolutionary Biology
+and Environmental Studies, 8057 Zurich, Switzerland;
+
+\noindent 2. University of British Columbia, Department of Zoology,
 Vancouver, BC V6T 1Z4, Canada;
-
-\noindent 2. University of Zurich, Department of Evolutionary Biology
-and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich,
-Switzerland;
 
@@
 
-\emph{Manuscript elements}: Figure 1, figure 2, figure 3, figure 4. All
-figures should be printed in color.
+\emph{Running title}: Character divergence destabilizes food webs
 
@@
 \emph{Keywords}: competition; eco-evolutionary dynamics;
-consumer-resource interactions; adaptive radiation; ecological community
-
-\bigskip
-
-\emph{Manuscript type}: Note.
-
-\bigskip
-
-\footnotesize Prepared using an \emph{Am. Nat.} inspired \LaTeX{}
-template for Rmarkdown. \normalsize
+consumer-resource interactions; adaptation; community stability
 
@@
 
+Ecological character displacement is an adaptive process that generally
+increases phenotypic diversity. Despite the fact that this
+diversification is due to an eco-evolutionary feedback between consumers
+competing for shared resources, its consequences for food-web dynamics
+have not been examined. Here, I study a model of two consumers competing
+for two shared resources to examine how character displacement in
+consumer attack rates affects resource abundances and the resilience of
+food webs to perturbations. I found that character displacement always
+strengthened consumer-resource interactions whenever consumers competed
+for resources that occurred in different habitats. This increase in
+interaction strength resulted in lower resource abundances and less
+resilient food webs. This occurred under different evolutionary
+tradeoffs and in both simple and more realistic foraging scenarios.
+Taken together, my results show that the adaptive process of character
+displacement may come with the ecological cost of decreasing food-web
+resilience.
+
 \newpage
@@
 
-Ecological character displacement is thought to be a key evolutionary
-process generating biodiversity (Dolph Schluter 2000; D. W. Pfennig and
-Pfennig 2010; but see Stuart and Losos 2013). Ecological character
-displacement describes the ``process of phenotypic evolution in a
-species generated or maintained by {[}exploitative{]} resource
-competition with one or more coexisting species'' (Dolph Schluter 2000).
-Over the past 40 years, a large body of theoretical (e.g. Lawlor and
-Smith 1976; P A Abrams 1986; Doebeli 1996; Taper and Chase 1985) and
-empirical (reviewed in: Dolph Schluter 2000; Dayan and Simberloff 2005;
-Stuart and Losos 2013) work has been generated to understand the
-scenarios under which exploitative competition for resources leads to
-the divergence of consumer foraging traits. The general result emerging
-from these studies is that, if resources are nutritionally substitutable
-(Peter A Abrams 1987; Fox and Vasseur 2008) and there is no other strong
-source of density dependence acting on consumers (P A Abrams 1986), then
-resource competition drives the divergence of consumer foraging traits
-{[}Lawlor and Smith (1976); Taper and Chase (1985)). This process is not
-simply driven by ecological differences, but creates an eco-evolutionary
-feedback that drives further differentiation. This key insight was made
-by theoretical models that explicitly included resource dynamics as a
-mediator of competition in driving evolutionary change (Lawlor and Smith
-1976; P A Abrams 1986; Taper and Chase 1985).
+Ecological character displacement is an important adaptive process in
+generating biodiversity \citep{Schluter2000, Pfennig2010}. This process
+is due to ``phenotypic evolution in a species generated or maintained by
+{[}exploitative{]} resource competition with one or more coexisting
+species'' \citep{Schluter2000}. A large body of theoretical
+\citep[e.g.][]{Lawlor1976, Abrams1986, Doebeli1996, Taper1985, McPeek2019}
+and empirical \citep[reviewed
+in:][]{Schluter2000, Dayan2005, Stuart2013} work has examined which
+scenarios lead to phenotypic divergence or convergence of competing
+consumers. The general conclusion has been that, if resources are
+nutritionally substitutable \citep{Abrams1987, Fox2008} and there is no
+other strong source of density dependence acting on consumers
+\citep{Abrams1986}, then resource competition drives the adaptive
+divergence of competitors \citep{Lawlor1976, Taper1985}. This adaptive
+process is not simply a response to static differences in resource
+distributions, but creates an eco-evolutionary feedback that drives
+further differentiation. This crucial insight was made by theoretical
+models that explicitly included resource dynamics as a mediator of
+competition in driving evolutionary change
+\citep{Lawlor1976, Abrams1986, Taper1985}.
 
-Although explicit models of resources gave key insights to the evolution
+Although models that included resources led to insights to the evolution
 of character displacement, the ecological feedback onto
-consumer-resource dynamics have received surprisingly little attention.
-This is likely because character displacement has been primarily studied
-through the lens of competition and coexistence theory {[}Lawlor1976;
-Germain2018{]}. For example, early theoretical work by Lawlor and Smith
-(1976) showed that ecological character displacement promotes
-coexistence by favoring specialized consumers that experience reduced
-interspecific competition. On the other hand, specialization often
-increases a consumer's ability to capture a specific resource. A well
-known consequence of increased foraging efficiency is resource
-suppression, which if sufficient enough, can generate oscillations, and
-hence less stable, consumer-resource dynamics {[}Rosenzweig 1971;
-Murdoch et al. 2003; McCann 2011{]}. NEED A LAST SENTENCE TO WRAP IT UP,
-HIGHLIGHTING HOW A FOOD-WEB PERSPECTIVE MAY YIELD NEW ECOLOGICAL
-INSIGHTS AND GIVE A CONTRASTING PERSPECTIVE TO COEXISTENCE THEORY.
-PERHAPS A I NEED TO INCLUDE A SENTENCE MID-WAY THROUGH SUGGESTING THAT A
-FOOD-WEB PERSPECTIVE MAY GIVE QUALITATIVELY DIFFERENT INSIGHTS. IT MAY
-BE NICE TO BRING IN THE SPATIAL COUPLING WORK OF MCCANN SUGGESTING THAT
-SPECIALIZATION CAN COUPLE FOOD-WEBS, BUT THIS WAS ONLY WITH ONE TOP
-PREDATOR (I.E. NO CONSUMERS\ldots{}LOOK AT OLD TEXT FOR WORDING
-SUGGESTIONS.)
+consumer-resource dynamics has received surprisingly little attention.
+This is likely because the ecological feedback has been primarily
+studied through the lens of coexistence theory
+\citep{Lawlor1976, Germain2018, Bassar2017, McPeek2019}. For example,
+early theoretical work showed that ecological character displacement
+promotes coexistence by favoring specialized consumers that experience
+reduced interspecific competition \citep{Lawlor1976}. Yet, this
+reduction in interspecific competition may, at the same time, increase
+interspecific interactions between specialized consumers and their
+resources. Both food-web theory and empirical studies have shown that
+increasing the strength of consumer-resource interactions often
+suppresses the abundance of resources, which if sufficient enough, can
+generate oscillations and less stable consumer-resource dynamics
+\citep{Rosenzweig1971, Luckinbill1973, Murdoch2002, Murdoch2003, McCann2011}.
+Thus, a food-web perspective, which accounts for both the direct and
+indirect effects of consumer-resource interactions, may yield new
+insight to the ecological consequences of character displacement.
 
-Here, I address this knowledge gap by building a mathematical model to
-examine how ecological character displacement affects consumer-resource
+Here, I address this knowledge gap by studying a mathematical model that
+examines how ecological character displacement affects consumer-resource
 dynamics in a food-web context. I address two questions: (1) How does
@@
 does character displacement affect food-web stability? To test the
-generality of these ecological effects, I examined three different
-foraging scenarios: (i) resources occur in the same habitat and
-consumers exhibit a linear functional response (e.g. MacArthur (1972);
-Lawlor and Smith (1976)); (ii) resources occur in different habitats and
-consumers exhibit a linear functional response (e.g. Lawlor and Smith
-(1976)); (iii) resources occur in distinct habitats and consumers
-exhibit a more realistic functional response (K S McCann, Rasmussen, and
-Umbanhowar 2005). I also explore the contingency of these effects to
-different assumptions about evolutionary trade-offs in consumer foraging
-traits. I focus on the ecological side of these dynamics because the
-evolutionary aspects have been well characterized (Lawlor and Smith
-1976; Taper and Chase 1985; P A Abrams 1986). I show that the
-consequences of character displacement on food-web dynamics are not
-always intuitive and depend on the foraging scenario and tradeoff in
-consumer traits.
+generality of these effects, I explored different ecological foraging
+scenarios and evolutionary tradeoffs in consumer attack rates. I found
+that the adaptive process of character displacement often comes with an
+ecological cost; resulting in food webs with lower resource availability
+and that are less resilient to perturbations.
 
-\section{Methods and Results}\label{methods-and-results}
+\section{Material and methods}\label{material-and-methods}
 
-\subsection{Underlying Food-web
-Dynamics}\label{underlying-food-web-dynamics}
+\subsection{Underlying Consumer-Resource
+Dynamics}\label{underlying-consumer-resource-dynamics}
 
-To examine how ecological character displacement affects
-consumer-resource interactions and food-web dynamics, I analyzed a
-continuous-time model of two consumers (\(C_1,C_2\)) competing for two
-shared resources (\(R_1,R_2\)):
+To examine how ecological character displacement affects resource
+abundances and food-web stability, I analyzed a continuous-time model of
+two consumers (\(C_{j=1,2}\)) competing for two shared resources
+(\(R_{i=1,2}\)):
 
-\[\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2,\]
-\[\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2,\]
-\[\frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1,\]
-\[\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2,\]
+\begin{equation} \label{eq:1}
+  \begin{split}
+     & \frac{dR_1}{dt}=r_1R_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{12}(R_1)C_2 \\
+     & \frac{dR_2}{dt}=r_2R_2(1-\frac{R_2}{K_2})-F_{21}(R_2)C_1-F_{22}(R_2)C_2 \\
+     & \frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{21}F_{21}(R_2)C_1-m_1C_1 \\
+     & \frac{dC_2}{dt}=e_{12}F_{12}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2 \\
+  \end{split}
+\end{equation}
 
 where \(r_i\) represents the intrinsic growth rate of resource \(i\),
-\(K_i\) represents the carrying capacity of resource \(i\), \(e_{ji}\)
+\(K_i\) represents the carrying capacity of resource \(i\), \(e_{ij}\)
 represents the conversion efficiency of resource \(i\) into consumer
 \(j\), and \(m_j\) represents the mortality rate of consumer \(j\).
-\(F_{ji}(R_i)\) represents consumer \(j\)'s feeding rate
-(i.e.~functional response) on resource \(i\). This model is a useful
+\(F_{ij}(R_i)\) represents consumer \(j\)'s feeding rate on resource
+\(i\) (i.e.~functional response). This model is a useful
 characterization of a scenario where consumers compete for two distinct
-resources (e.g.~zooplankton and benthic invertebrates) rather than if
-resources are better characterized by a continuous trait distribution
-(e.g.~seed size, see Taper and Chase (1985) for an example). In the
-scenario where consumer feeding rates increase linearly with resource
-density, such that \(F_{ji}(R_i)=a_{ji}R_i\) where \(a_{ji}\) is the
-attack rate of consumer \(j\) on resource \(i\), then this model becomes
-MacArthur's model of resource competition (MacArthur (1972)). The
-evolution of consumer attack rates in this model (and several
-extensions) have been analyzed in detail (Lawlor and Smith (1976); P A
-Abrams (1986)), with the general result being divergent character
-displacement. I say consumers have undergone divergent character
-displacement if there evolved attack rates are more specialized, defined
-as \(|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|\), when evolving with a
-competing consumer. I focus here on the ecological consequences of this
-adaptive divergence.
+resources (e.g.~zooplankton and benthic invertebrates in lakes) rather
+than a scenario where resources are better characterized by a continuous
+trait distribution (e.g.~seed size, see \citet{Taper1985} for an
+example). Importantly, inferences about ecological character
+displacement can only be made by comparing food webs with and without a
+competing consumer \citep{Schluter1992}. Therefore, I arbitrarily set
+\(C_2=0\) to create a food-web without a competing consumer for these
+comparisons.
 
-To investigate the ecological consequences, I studied the effect of
-adaptive divergence in consumers on resource densities and the stability
-of the entire food web at equilibrium. An equilibrium is reached when
-the four rates of change in the above equations are 0, and solving the
-system at this point gives equilibrium densities for each resource
-(\(\hat R_1,\hat R_2\)) and consumer (\(\hat C_1,\hat C_2\)). At the
-coexistence equilibrium (i.e., where
-\(\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0\)), I calculated resource
-densities and performed a linear stability analysis (Otto and Day 2007).
-This stability analysis derives the dominant (largest in absolute value)
-eigenvalue, \(\lambda\), which determines whether (and how readily) the
-food web, when perturbed a small amount, will return to equilibrium (see
-supplementary material for derivation). Together, these two measures
-indicate how ecological character displacement affects resource
-densities and food-web stability.
+\subsection{Foraging Scenarios}\label{foraging-scenarios}
 
-I examine the effects of this inevitable divergence in attack rates on
-food-web dynamics by (i) deriving analytical expressions for the
-relationship between attack rates and food-web dynamics and (ii)
-simulating the effects of competition on the eco-evolutionary dynamics
-of consumers and resources. To gain analytical insight, I assume that
-resources are equivalent (\(r=r_1=r_2\) and \(K=K_1=K_2\)) as well as
-consumers (\(e=e_{11}=e_{12}=e_{21}=e_{22}\); \(m=m_1=m_2\)), except
-that consumer attack rates are mirror images of each other
-(\(a_{ii}=a_{11}=a_{22}\); \(a_{ij}=a_{12}=a_{21}\)). While this
-scenario is arguably simplistic, it allows me to isolate the effects of
-ecological character displacement.
+I studied three different foraging scenarios. In the first, I assume
+that consumers can forage for both resources simultaneously (fig.
+\ref{fig:foraging_scenarios} top) and their feeding rates increase
+linearly with resource abundance, such that:
 
-To simulate eco-evolutionary dynamics, I used an Adaptive Dynamics
-approach. At each evolutionary time step, I created a mutant consumer by
-randomly choosing a consumer and modifying its attack rate on one
-resource by either subtracting or adding a small constant (0.001 in the
-following simulations) with equal probability. The mutant's attack rate
-on the other resource was determined by a tradeoff, such that
-\(\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1\), where \(A\) is the
-total investment in attack rates and \(n\) describes the shape of the
-tradeoff (Sargent and Otto 2006). This function has the useful property
-that it differentiates between cases where intermediate combinations of
-\(a_{ii}\) and \(a_{ij}\) are higher, on average, than the extremes
-(when \(n>1\), green line in Fig. \ref{fig:Tradeoff}) or, conversely,
-where the two extremes are higher, on average, than intermediate
-investments (when \(n<1\), orange line Fig. \ref{fig:Tradeoff}). When
-\(n=1\), the tradeoff function is linear, and all combinations of
-\(a_{ii}\) and \(a_{ij}\) have the same total attack rate (blue line in
-Fig. \ref{fig:Tradeoff}). Assuming the mutant consumer was rare, I then
-determined whether the mutant had higher relative fitness than the
-resident consumer, and thus could invade and replace the resident
-consumer. If the mutant was able to invade, I updated the attack rate of
-the resident consumer to the mutant attack rate and allowed consumer and
-resource dynamics to reach a steady state. I then repeated the
-simulation up to 10,000 times, which was sufficient for consumers to
-either reach an evolutionary stable strategy (ESS, Smith and Price 1973)
-or an evolutionary limit (e.g. \(\frac{a_{ii}}{a_{ii}+a_{ij}}\) is
-constrained to a maximum of 1 and minimum of 0).
+\begin{equation} \label{eq:2}
+  F_{ij}(R_i)=a_{ij}R_i
+\end{equation}
+
+where \(a_{ij}\) is the attack rate of consumer \(j\) on resource \(i\).
+This first scenario is the starting point for many models of resource
+competition \citep{MacArthur1972}; however, it does not reflect many
+food webs where consumers are mobile and their foraging behavior links
+resources that occur in different habitats \citep{McCann2005}. The
+second scenario accounts for this spatial context (fig.
+\ref{fig:foraging_scenarios} bottom) and takes the form:
+
+\begin{equation} \label{eq:3}
+  F_{ij}(R_i)=w_{ij}a_{ij}R_i
+\end{equation}
+
+where \(w_{ij}\) represents the proportion of time consumer \(j\) spends
+foraging in a habitat where only resource \(i\) is found (i.e.~habitat
+preference). Note that since \(w_{ij}\) is a proportion that
+\(w_{1,j}=1-w_{2,j}\). Finally, it is well known that consumer feeding
+rates often saturate at high resource abundances
+\citep{Holling1959, Rosenzweig1963, Murdoch2003, McCann2011} and that
+consumers do not usually spend a fixed proportion of time in a
+particular habitat \citep{McCann2005}. The third scenario accounts for
+these biological realities and takes the form \citep[derived
+by][]{McCann2005}:\\
+
+\begin{equation} \label{eq:4}
+  F_{ij}(R_i)=\frac{a_{ij}W_{ij}R_i}{1+a_{1,j}h_{1,j}W_{1,j}R_1+a_{2,j}h_{2,j}W_{2,j}R_2}
+\end{equation}
+
+where consumer \(j\)'s feeding rate on resource \(i\) is influenced by
+the abundance of each resource; saturates as resource abundances
+increase (due to handling times \(h_{ij}\)); and consumer habitat
+preferences are modified by the relative abundance of resources, such
+that: \(W_{ij}=\frac{w_{ij}R_i}{w_{1,j}R_1+w_{2,j}R_2}\).
+
+Previous studies have analyzed the evolution of consumer attack rates in
+the first two foraging scenarios using an Adaptive Dynamics approach,
+with the general result being character divergence
+\citep{Lawlor1976, Abrams1986}. I also used an Adaptive Dynamics
+approach to analyze character displacement in the third foraging
+scenario, and I too observed character divergence (detailed analysis in
+Appendix S1 of Supplementary Information). I say consumers have
+undergone character divergence if their evolved attack rates are more
+specialized when evolving with vs.~without a competing consumer.
+Specialization of consumer \(j\) on resource \(1\) is measured as
+\(\frac{a_{1,j}}{a_{1,j}+a_{2,j}}\), where a value of 0.5 is a complete
+generalist (\(a_{1,j}=a_{2,j}\)), and a value of 1 is a complete
+specialist (\(a_{2,j}=0\)). Values less than 0.5 indicate specialization
+on the other resource.
 
@@
 \centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/Tradeoff-1.pdf}
-\caption{\label{fig:Tradeoff}Tradeoff forms for consumer attack rates.}
+\includegraphics{ECD_model_foraging_scenarios}
+\caption{\label{fig:foraging_scenarios}\textbf{Ecological foraging
+scenarios.} I examined whether the effect of ecological character
+displacement on food-web dynamics depended on whether consumers competed
+for resources that occurred in the same vs.~different habitats. The
+width of each arrow corresponds to initial attack rates
+(\emph{a}\textsubscript{ij}) of consumer \emph{j} on resource \emph{i}.
+Note that \emph{C}\textsubscript{1} was pre-adapted to
+\emph{R}\textsubscript{1} (\emph{a}\textsubscript{11} \textgreater{}
+\emph{a}\textsubscript{21}), while \emph{C}\textsubscript{2} was a
+mirror image, being pre-adapted to \emph{R\textsubscript{2}}
+(\emph{a}\textsubscript{22} \textgreater{} \emph{a}\textsubscript{12}).
+In each scenario, I assumed consumer feeding rates increased linearly
+with resource abundance. I also relax this assumption and consider a
+more realistic functional response when resources occurred in different
+habitats.}
 \end{figure}
 
-\subsection{Consumers Forage For Both Resources
-Simultaneously}\label{consumers-forage-for-both-resources-simultaneously}
+\subsection{Food-web Dynamics}\label{food-web-dynamics}
 
-When both consumers and resources are present, the density of resources
-at equilibrium are equivalent (\(\hat R = \hat R_1 = \hat R_2\)) and are
-determined by the following equation (derivation in supplementary
-material):
+Given that character divergence occurred across these foraging
+scenarios, I focus here on its consequences for food-web dynamics. To do
+this, I analyzed differences in resource abundances and food-web
+stability at equilibrium. An equilibrium is reached when the rates of
+change in equation \ref{eq:1} are 0, and solving the system at this
+point gives equilibrium abundances for each resource (\(\hat R_i\)) and
+each consumer (\(\hat C_j\)). I also compared the local stability of
+these food webs using standard methods \citep{Otto2007}. This stability
+analysis derives the dominant eigenvalue, \(\lambda_{\text{max}}\), of
+the matrix of partial derivatives of all differential equations with
+respect to all variables evaluated at equilibrium. If
+\(-\lambda_{max}>0\), then the food web will return to equilibrium after
+a small perturbation (i.e.~locally stable), with more positive values
+indicating a faster return time. If \(-\lambda_{max}<0\), then the food
+web is not locally stable.
 
-\[\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}\] A key determinant of
-resource density in this model is the consumer's total attack rate,
-\(a_{ii}+a_{ij}\). Therefore, the ecological consequences of character
-displacement depend on how the tradeoff influences the evolution of
-consumer attack rates.
+When possible, I derived analytical expressions for the relationship
+between consumer attack rates and food-web dynamics. To do this, I
+simplified the model by assuming that resources are equivalent
+(\(r=r_i\) and \(K=K_i\)) as well as consumers (\(e=e_{ij}\);
+\(h=h_{ij}\); \(m=m_j\)), except that consumer attack rates and their
+habitat preferences (if present) are mirror images of each other
+(\(a_{11}=a_{22}\); \(a_{12}=a_{21}\); \(w_{11}=w_{22}\)). Note that I
+arbitrarily set \(C_1\) as being pre-adapted to \(R_1\)
+(\(a_{11} > a_{21}\); \(w_{11} > 0.5\)), and therefore \(C_2\) was
+pre-adapted to \(R_2\). Controlling for other sources of variability
+allowed me to isolate the general effects of character divergence. All
+mathematical derivations were conducted in Mathematica
+\citep{Mathematica} and are provided in the Supplementary Information
+(Appendix S1-3).
 
-My simulations show that the shape of the attack-rate tradeoff
-qualitatively affects the relationship between character displacement
-and resource density (Fig. \ref{fig:MacArthur_Resources}). For example,
-if consumer's are constrained by a linear tradeoff (blue lines), then
-there is no net change in total attack rate (Fig.
-\ref{fig:MacArthur_Resources}A) and character displacement has no effect
-on resource densities (Fig. \ref{fig:MacArthur_Resources}B). If the
-tradeoff is concave down (green lines), then resource abundances can
-actually increase under character displacement (Fig.
-\ref{fig:MacArthur_Resources}B). This is because the total attack rate
-of consumers is maximized at intermediate values (\(a_{ii}=a_{ij}\)) and
-decreases as consumers diverge (Fig. \ref{fig:MacArthur_Resources}A).
-When the tradeoff is concave up (orange lines), character displacement
-suppresses resource densities due to the increase in total attack rates
-(Fig. \ref{fig:MacArthur_Resources}A,B). Although the equation I derived
-for resource density was for the scenario where both consumers and both
-resources are present, it accurately predicts the density of resources
-when a single consumer reaches its ESS (triangles on respective colored
-lines in Fig. \ref{fig:MacArthur_Resources}B). This is because a single
-consumer evolves to be a generalist that has equal attack rates on each
-resource (triangles at 0.5 along x-axis in Fig.
-\ref{fig:MacArthur_Resources}A), resulting in equivalent resource
-densities.
-
-The effect of character displacement on resource density closely matches
-its impact on food-web stability (Fig. \ref{fig:Stability}). For
-example, when character displacement results in resource suppression
-(orange), there is a corresponding decrease in stability. Similarly, if
-character displacement does not influence resource densities (blue),
-then there is no corresponding effect on stability. Curiously, this
-match between impact on resources and stability breakdowns with the
-concave down tradeoff (green).
+To gain insight to the eco-evolutionary feedback generated by character
+displacement, I conducted simulations using an Adaptive Dynamics
+approach. Specifically, after letting consumer and resource abundances
+reach a steady state, I created a mutant consumer by randomly choosing
+one and modifying its attack rate on one resource by either subtracting
+or adding a small constant (0.01 in the following simulations) with
+equal probability. The mutant's attack rate on the other resource was
+determined by a tradeoff, such that
+\(\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1\),
+where \(A\) is the total investment in attack rates and \(n\) describes
+the shape of the tradeoff \citep{Sargent2006}. This function has the
+useful property that it differentiates between cases where intermediate
+combinations of \(a_{\text{1},j}\) and \(a_{\text{2},j}\) are higher
+than the extremes (when \(n>1\), green line in fig. \ref{fig:tradeoff})
+or, conversely, where the two extremes are higher than intermediate
+investments (when \(n<1\), orange line fig. \ref{fig:tradeoff}). When
+\(n=1\), the tradeoff function is linear, and all combinations of
+\(a_{\text{1},j}\) and \(a_{\text{2},j}\) have the same total attack
+rate (blue line in fig. \ref{fig:tradeoff}). Assuming the mutant
+consumer was rare, I then determined whether the mutant had higher
+relative fitness than the resident consumer, and thus could invade and
+replace the resident consumer. If the mutant was able to invade, I
+updated the attack rate of the resident consumer to the mutant attack
+rate and allowed consumer and resource abundances to reach a steady
+state. I then repeated the simulation up to 10,000 times, which was
+sufficient for consumers to either reach an evolutionary stable strategy
+\citep[ESS,][]{Smith1973} or an evolutionary limit (e.g.
+\(\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}\) is constrained to a
+maximum of 1 and minimum of 0). Unless otherwise noted, I conducted
+simulations with the following parameter values: \emph{r} = 1; \emph{K}
+= 4; \emph{e} = 0.8; \emph{m} = 1; \emph{A} = 2; \emph{h} = 0.4; and
+\(w_{11} = w_{22} = 0.6\). I set an initial value of
+\(a_{11} = a_{22} = 1.2\), while \(a_{12}\) and \(a_{21}\) depended on
+the value of \emph{n}. I set initial consumer and resource abundances
+to: \(R_1 = R_2 = 2\); \(C_1 = C_2 = 1\). All simulations were conducted
+in R \citep{R} and the code to reproduce these simulations is publically
+available on GitHub (\url{https://github.com/mabarbour/ECD_model}).
 
@@
 \centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/MacArthur_Resources-1.pdf}
-\caption{\label{fig:MacArthur_Resources}Effect of character displacement
-on total attack rates (A) and resource densities (B) when consumers can
-forage for both resources simultaneously.}
+\includegraphics{attack_tradeoffs.pdf}
+\caption{\label{fig:tradeoff}\textbf{Evolutionary tradeoffs in consumer
+attack rates.} In each foraging scenario, I explored the effects of
+three different tradeoffs: intermediate combinations of attack rates
+(\emph{a}\textsubscript{1,\emph{j}}, \emph{a}\textsubscript{2,\emph{j}})
+are higher than the extremes (green line, \emph{n} \textgreater{} 1);
+extreme combinations of attack rates are higher than intermediate
+investments (orange line, \emph{n} \textless{} 1); and all combinations
+of attack rates have the same total attack rate (blue line, \emph{n} =
+1). Points corresponding to attack rates at the beginning of the
+simulation for \emph{C}\textsubscript{1}, which was pre-adapted to
+\emph{R}\textsubscript{1} (\emph{a}\textsubscript{11} \textgreater{}
+\emph{a}\textsubscript{12}).}
 \end{figure}
 
-\includegraphics{Figure_ECD_MacArthur.pdf}
+\section{Results}\label{results}
 
-\subsection{\texorpdfstring{Consumers \emph{cannot} forage for both
-resources
-simulanteously}{Consumers cannot forage for both resources simulanteously}}\label{consumers-cannot-forage-for-both-resources-simulanteously}
+\subsection{Resources occur in same
+habitat}\label{resources-occur-in-same-habitat}
 
-The only character displacement model that I am aware of that modeled
-resources in different habitats was one examined by Lawlor and Smith
-(1976). This model takes the same form as the previous model, except now
-the consumer's feeding rate takes the form:
+In this first scenario (equation \ref{eq:2}), the abundance of resources
+at equilibrium are equivalent when both consumers and resources are
+present (\(\hat R = \hat R_1 = \hat R_2\)), and are determined by the
+following equation (derived in Appendix S2):
 
-\[F_{ji}(R_i)=w_{ji}a_{ji}R_i\]
+\begin{equation} \label{eq:5}
+  \hat{R}=\frac{1}{a_{1,j}+a_{2,j}}\cdot\frac{m}{e}
+\end{equation}
 
-where \(w_{ji}\) represents the proportion of time consumer \(j\) spends
-foraging in a habitat where only resource \(j\) is found (i.e.~habitat
-preference). Note that since \(w_{ji}\) is a proportion that
-\(w_{jj}=1-w_{ji}\). As with attack rates, I assume that consumer
-habitat preferences are mirror images of each other
-(\(w=w_{11}=w_{22}\)). In Lawlor and Smith (1976)'s detailed
-evolutionary analysis of this model, they again always observe divergent
-character displacement.
+A key determinant of resource abundance in this scenario is the
+consumer's total attack rate, \(a_{1,j}+a_{2,j}\). Therefore, the effect
+of character displacement on food-web dynamics depends on how the shape
+of the tradeoff function influences the evolution of consumer attack
+rates.
 
-To examine the ecological consequences of this model, I again derive the
-equilibrium solution when both consumers and both resources are present.
-Again, resource densities are equivalent at this equilibrium
-(\(\hat R = \hat R_1 = \hat R_2\) when both consumers and resources
-present), but are now determined by the following equation (derivation
-in supplementary material):
+I found that the shape of the tradeoff function qualitatively affects
+the relationship between character displacement and resource abundances
+in this scenario (fig. \ref{fig:plot_fig3} middle left). For example, if
+consumer's are constrained by a linear tradeoff (blue lines), then there
+is no net change in total attack rate (fig. \ref{fig:plot_fig3} top
+left) and character displacement has no effect on resource abundances
+(fig. \ref{fig:plot_fig3} middle left). If the tradeoff is concave down
+(green lines), then resource abundances can actually increase under
+character displacement (fig. \ref{fig:plot_fig3} middle left). This is
+because the total attack rate of consumers is maximized at intermediate
+values (\(a_{1,j}=a_{2,j}\)) and decreases as consumers diverge (fig.
+\ref{fig:plot_fig3} top left). When the tradeoff is concave up (orange
+lines), character displacement suppresses resource abundances due to the
+increase in total attack rates (fig. \ref{fig:plot_fig3} top left).
+Although the equation I derived for resource abundances was for the
+scenario where both consumers and both resources were present, it
+accurately predicts the abundance of resources when a single consumer
+reaches its evolutionary stable strategy (ESS; triangles on respective
+colored lines in fig. \ref{fig:plot_fig3} middle left). This is because
+a single consumer evolves to be a generalist that has equal attack rates
+on each resource (triangles at 0.5 along x-axis in fig.
+\ref{fig:plot_fig3} left panels), resulting in equivalent resource
+abundances.
 
-\[\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e}\]
+The effect of character displacement on resources corresponds to its
+impact on food-web stability. For example, when character displacement
+decreases resource abundances (orange points in fig. \ref{fig:plot_fig3}
+middle left), there is a decrease in food-web stability (fig.
+\ref{fig:plot_fig3} bottom left). Character divergence may not affect or
+even increase food-web stability (blue and green lines in fig.
+\ref{fig:plot_fig3} bottom left); however, evolution does not favor
+strong divergence in these scenarios (blue and green points in fig.
+\ref{fig:plot_fig3}), which dampens these contingent effects. Note that
+the dip in stability occurs when both consumers evolve to be
+generalists, a situation that is not favored in any of the foraging
+scenarios we examined (fig. \ref{fig:plot_fig3} bottom).
+
+\subsection{Resources occur in different
+habitats}\label{resources-occur-in-different-habitats}
+
+In the second foraging scenario (equation \ref{eq:3}), I again see that
+resource abundances are equivalent when both consumers and resources are
+present (\(\hat R = \hat R_1 = \hat R_2\)), but are now determined by
+the following equation (derived in Appendix S3):
+
+\begin{equation} \label{eq:6}
+  \hat{R}=\frac{1}{w_{1,j}a_{1,j}+w_{2,j}a_{2,j}}\cdot\frac{m}{e}
+\end{equation}
 
 This equation implies that if consumers evolve to become specialists on
-resources that occur in their preferred habitat (e.g. \(w>0.5\) and
-\(a_{ii}>a_{ij}\)), then the effective attack rate of consumers
-(\(wa_{ii}+(1-w)a_{ij}\)) will always increase, regardless of the
-tradeoff (Fig. \ref{fig:LS_Resources}A). Thus, character displacement
-always results in resource suppression if consumers are competing for
-resources that occur in different habitats (Fig.
-\ref{fig:LS_Resources}B). Note that the shape of the tradeoff can modify
-the effect of character displacement. This is not so much due to the
-tradeoff affecting the magnitude of displacement (it does, but the
-effect is minor), but because the form of the tradeoff affects resource
-abundances when a single consumer has reached its ESS (triangles in Fig.
-\ref{fig:LS_Resources}B). In contrast, resource densities reach a
-similar value when consumers evolve in the presence of a competitor
-(circles in Fig. \ref{fig:LS_Resources}B), because character
-displacement tends to reach a constraint of complete specialization. It
-is worth noting that in this foraging scenario, resource densities are
+resources that occur in their preferred habitat (e.g. \(w_{1,j}>0.5\)
+and \(a_{1,j}>a_{2,j}\)), then the effective attack rate of consumers
+(\(w_{1,j}a_{1,j}+w_{2,j}a_{2,j}\)) will always increase, regardless of
+the tradeoff (fig. \ref{fig:plot_fig3} top right). Thus, character
+divergence always results in resource suppression (fig.
+\ref{fig:plot_fig3} middle right). Note that the shape of the tradeoff
+can modify the effect of character displacement. This is not so much due
+to the tradeoff affecting the magnitude of displacement (it does, but
+the effect is minor), but because the form of the tradeoff affects
+resource abundances when a single consumer has reached an ESS (triangles
+in fig. \ref{fig:plot_fig3} middle right). In contrast, resource
+abundances reach a similar value when consumers evolve in the presence
+of a competitor (circles in fig. \ref{fig:plot_fig3} middle right),
+because character displacement tends to reach a constraint of complete
+specialization. It is worth noting that resource abundances are
 consistently higher at the single consumer ESS compared to the
 predictions I derived for when both consumers are present (deviation of
-triangles from respective colored lines in Fig.
-\ref{fig:LS_Resources}B). This is likely because consumers actually
-evolve to be slightly specialized on the resources that occur in their
-non-preferred habitat (deviation of triangles from 0.5 along x-axis in
-Fig. \ref{fig:LS_Resources}B).
+triangles from respective colored lines in fig. \ref{fig:plot_fig3}
+middle right). This is because consumers actually evolve to be slightly
+specialized on the resources that occur in their non-preferred habitat
+(deviation of triangles from 0.5 along x-axis in fig.
+\ref{fig:plot_fig3} right panels).
 
-The effect of character displacement on resource densities qualitatively
-matches its negative effect of food-web stability (Fig.
-\ref{fig:Stability}). This is not simply a consequence of having an
-additional consumer in the system, but emerges from the eco-evolutionary
-feedback between character displacement and resource suppression
-(compare initial (small points) with evolved strategies (large points),
-Fig. \ref{fig:Stability}). For example, when the tradeoff is concave up
-(orange), then the four-species (small circle) food web actually starts
-off as less stable than the three-species one (small triangle), but the
-eco-evolutionary dynamics cause a switch in the pattern once they reach
-their evolutionary stable strategies (large points).
+As seen previously, the effect of character displacement on resource
+abundances qualitatively corresponds to its effect on food-web stability
+(fig. \ref{fig:plot_fig3}). Specifically, character divergence decreases
+food-web stability, regardless of the tradeoff in attack rates. This is
+not simply a consequence of having an additional consumer in the system,
+but emerges from the eco-evolutionary feedback between character
+displacement and resource suppression (fig. \ref{fig:plot_fig3}). For
+example, when the tradeoff is concave up (orange), the initial
+two-consumer food web (small circle) is more stable than when there is
+only one consumer (small triangle); however, this pattern switches by
+the end of the eco-evolutionary simulation (large points).
 
@@
 \centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/LS_Resources-1.pdf}
-\caption{\label{fig:LS_Resources}Effect of character displacement on
-effective attack rates (A) and resource densities (B) when consumers
-cannot forage for resources simultaneously.}
+\includegraphics{Figure_3.pdf}
+\caption{\label{fig:plot_fig3}\textbf{Effect of character displacement
+on food-web dynamics under different evolutionary tradeoffs and foraging
+scenarios.} Lines show predicted values when both consumers and
+resources are present. Different line colors correspond to different
+tradeoffs in attack rates (green, \emph{n} = 1.15; blue, \emph{n} = 1;
+orange, \emph{n} = 0.85). Large circles (two consumers) and triangles
+(one consumer) correspond to the end points of the eco-evolutionary
+simulation for \emph{C}\textsubscript{1}, whereas as small shapes
+correspond to the starting points (only in stability panels). In both
+foraging scenarios, feeding rates increase linearly with resource
+abundance, but the equation for effective attack rate is different.}
 \end{figure}
 
-\includegraphics{Figure_ECD_LS.pdf}
+\subsection{Adding a more realistic functional
+response}\label{adding-a-more-realistic-functional-response}
 
-\subsection{Adding A More Realistic Functional
-Response}\label{adding-a-more-realistic-functional-response}
+In the third foraging scenario (equation \ref{eq:4}), I observed the
+same general effect of character divergence as the previous scenario
+(resources in different habitats, but linear functional response). This
+is because resource abundances at equilibrium are governed by a similar
+dynamic (derived in Appendix S1):
 
-Thus far I have consider a simple type I functional response of
-consumers foraging for resources in either the same or distinct
-habitats.\\
-For example, even if I consider a more realistic consumer functional
-response (derived by K S McCann, Rasmussen, and Umbanhowar (2005)):
+\begin{equation} \label{eq:7}
+  \hat{R}=\frac{1}{w_{1,j}a_{1,j}+w_{2,j}a_{2,j}}\cdot\frac{m}{e-hm}
+\end{equation}
 
-\[F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}hW_{ii}R_i+a_{ij}hW_{ij}R_j}\]
+And since evolution favors consumer divergence onto their preferred
+resources (see Appendix S1), the effective attack rate of consumers
+(\(w_{1,j}a_{1,j}+w_{2,j}a_{2,j}\)) will always increase, resulting in
+lower resource abundances and decreased food-web stability (Appendix
+S4).
 
-where consumer feeding rates on resource \(i\) are influenced by both
-resource densities; saturate as resource densities increase (due to
-handling time \(h\)); and consumer habitat preferences are modified by
-relative resource densities, such that:
+In the first two foraging scenarios, character displacement influences
+food-web stability, but all of the food webs ultimately return to a
+stable equilibrium (because \(-\lambda_{max}>0\), see Appendix S2-3). In
+this more realistic model, however, whether the food web is locally
+stable depends on consumer and resource parameters. Specifically, I
+found that the two-consumer food web will transition from having a
+locally stable equilibrium to a limit cycle under the following
+conditions (derived using Routh-Hurwitz criteria in Appendix S1):
 
-\[W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}\]
+\begin{equation} \label{eq:8}
+  w_{1,j}a_{1,j}+w_{2,j}a_{2,j} > \frac{e+hm}{hK(e-hm)}
+\end{equation}
 
-I still observe the same qualitative relationship between character
-displacement and resource suppression (Fig. \ref{fig:McCann_Resources}
-in Appendix). This is because resource competition always leads to
-character divergence and equilibrium resource densities are governed by
-similar dynamics (\emph{Mathematica} file with derivations available
-upon request):
+Thus, character displacement always pushes the food web toward an
+unstable structure in this more realistic foraging scenario (fig.
+\ref{fig:bifur_plot}). Note that I stopped the simulation in the
+four-species food web once it became locally unstable. I do not simulate
+beyond this point as this would require making assumptions about the
+dynamics of mutant consumers in variable environments, which is beyond
+the scope of this work.
 
-\[\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e-hm}\]
+\begin{figure}
+\centering
+\includegraphics{bifurcation_plot.pdf}
+\caption{\label{fig:bifur_plot}\textbf{Character divergence creates an
+unstable food web}. Lines illustrate the effects of character divergence
+across the range of specialization, while the points are the results of
+an eco-evolutionary simulation. Note that I increased the total
+investment in attack rates (\emph{A} = 3.3) to create a scenario that
+could result in an unstable food web.}
+\end{figure}
 
-I conducted the previous simulations on a scenario where the food web
-remains locally stable (\(\operatorname{Re}(\lambda_{max})<0\)). Is it
-possible that character displacement would ever result in a locally
-unstable food web? To gain analytical insight to this question, I
-examined the conditions governing the local stability of the
-four-species food web when consumers exhibit a more realistic functional
-response (K S McCann, Rasmussen, and Umbanhowar (2005)). I show that the
-four-species food web will transition from having a locally stable
-equilibrium to a limit cycle under the following conditions
-(\emph{Mathematica} file with derivation available upon request):
+\subsection{Robustness to consumer
+asymmetry}\label{robustness-to-consumer-asymmetry}
 
-\[wa_{ii}+(1-w)a_{ij} > \frac{e+hm}{hK(e-hm)}\]
+The previous analytical results and simulations make a strong assumption
+that competing consumers start off as perfect mirror images of each
+other (i.e.~symmetry). Yet, theory indicates a predictable asymmetry
+between initial consumer attack rates. This predictable asymmetry
+emerges from a process of community assembly where a single consumer
+invades a system, reaches an ESS as a generalist, followed by the
+invasion of a second, more specialized, consumer. This theoretical
+scenario has been hypothesized as the sequence of events leading to
+character displacement in threespine stickleback in small coastal lakes
+of British Columbia \citep{Schluter1992, Schluter2000}.
 
-Previously, I showed that character displacement always increases the
-effective attack rate of consumers (\(wa_{ii}+(1-w)a_{ij}\)), regardless
-of the the shape of the tradeoff (Fig. \ref{fig:LS_Resources}A). Thus,
-character displacement is capable of pushing food webs to a point where
-they are locally unstable. (it simple depends on the choice of
-parameter, I could put a situation where evolution pushes it toward
-local instability, or even the hopf bifurcation.)
+To test whether my results were robust to this asymmetry, I used the
+evolved attack rates at the end of the simulations with one consumer as
+the starting values for one of the two consumers. I did this for all
+foraging scenarios and tradeoffs previously examined. I found that my
+previous inferences are robust to including consumer asymmetry across
+different foraging scenarios and tradeoffs (Appendix S4).
 
@@
 
-I find that the foraging scenario qualitatively affects the relationship
-between character displacement and resource densities.
+\subsection{Resource Abundances}\label{resource-abundances}
 
-Most models of character displacement have assumed a scenario where
-consumers can forage for both resources simultaneously (e.g. Taper and
-Chase 1985; P A Abrams 1986). This assumption may be valid for some
-foraging scenarios (e.g.~Darwin's finches foraging for seeds); however,
-in many situations consumers forage for resources that occur in
-different habitats, and thus cannot forage for both simultaneously (K S
-McCann, Rasmussen, and Umbanhowar 2005).
+One of the criteria used to demonstrate character displacement is that
+``sites of sympatry {[}two consumers{]} and allopatry {[}one consumer{]}
+should not differ greatly in food, \ldots{}'' \citep{Schluter1992}. In
+contrast, my results suggest that ecological character displacement
+causes predictable differences in resource abundances. In fact, the
+ecological and evolutionary scenarios that favored the greatest
+character divergence always decreased the abundance of resources. For
+example, the magnitude of character displacement was highest when mobile
+consumers competed for resources that occur in different habitats. An
+empirical example of this is threespine stickleback that have diverged
+into limnetic and benthic species with specialized traits to forage on
+zooplankton and benthic invertebrates, respectively
+\citep{Schluter1992, Schluter2000}. These two resource types occur in
+distinct zones of the lake (pelagic and littoral), therefore stickleback
+must move between these zones when foraging for resources.
+Interestingly, a disproportionate number of the documented cases of
+character displacement involve carnivores \citep{Schluter2000} that are
+larger, and likely more mobile, than their resources \citep{McCann2005},
+suggesting many cases of ecological character displacement may result in
+lower resource availability.
 
-The effect of the foraging scenario on the relationship between
-character displacement and resource densities appears to be quite
-general.
+Similarly, the evolutionary tradeoff that favored character divergence
+resulted in lower resources across all foraging scenarios. Although data
+on the shape of the tradeoff in consumer foraging traits is scarce, two
+classic examples of character displacement, Darwin's finches and
+threespine stickleback, both appear to exhibit a tradeoff where extreme
+trait values increase the net foraging rate of consumers
+\citep{Schluter1985, Arnegard2014}. While it is theoretically possible
+that character displacement does not alter (or even increase) resource
+abundances, this was limited to the simplest, and arguably least
+realistic, foraging scenario and under tradeoffs that did not favor
+large displacements, and thus less likely to detect in nature. Taken
+together, my results call for empirical work to test these clear
+theoretical predictions and suggest a revision is needed for one of the
+criteria used to demonstrate character displacement.
 
-Note that detecting this effect in nature would likely be difficult,
-since a concave-down tradeoff results in relatively small displacement
-relative to other tradeoff shapes (Lawlor and Smith (1976); green points
-in Fig. \ref{fig:MacArthur_Resources}A,B).
+\subsection{Food-web Stability}\label{food-web-stability}
 
-Previous work has shown that consumers undergo divergent character
-displacement regardless of the shape of the tradeoff (Lawlor and Smith
-(1976); P A Abrams (1986)).
+My most striking result was that character divergence made food webs
+less resilient to perturbations. In fact, under the most realistic
+foraging scenario, character divergence can even result in an unstable
+food web. The mechanism underlying this destabilization is quite
+general. Character divergence generally increases the strength of
+consumer-resource interactions, but does not alter the strength of
+intraspecific interactions. This relative increase in interspecific
+interactions, combined with the natural oscillatory tendency of
+consumer-resource dynamics \citep{Lotka1925, Volterra1926}, creates a
+food-web structure that is less resilient to perturbations
+\citep{Chesson2008, Rip2011, McCann2011}.
 
-Important point: Criteria 5 in D Schluter and McPhail (1992) does not
-make sense since character displacement usually alters resource
-distributions. Think about this more.
+Interestingly, the ecological conditions that favor character divergence
+are those that are already the least resilient to perturbations. For
+example, \citet{McPeek2019} showed that character divergence is favored
+in food webs that are either highly productive, easy to find and capture
+resources, or under weak abiotic stress. This corresponds to higher
+values of \(K\) (productivity) or \(A\) (investment in attack rates), or
+lower values of \(m\) (abiotic stress). Each of these corresponding
+changes decrease food-web resilience, as they increase the strength of
+consumer-resource interactions relative to intraspecific interactions.
+For example, increasing productivity reduces intraspecific competition
+in resource populations while increasing the flux of energy to
+consumers, resulting in the paradox of enrichment
+\citep{Rosenzweig1971}. Similarly, higher feeding rates or lower
+consumer mortality both increase the relative strength of
+consumer-resource interactions, which predictably destabilizes food webs
+\citep{Rip2011, McCann2011}. This suggests that the most dramatic
+examples of character divergence will not only occur in, but also cause,
+the least stable food-web structures.
 
-Lawlor and Smith (1976) found that consumers still underwent divergent
-character displacement regardless of whether consumers could or could
-not forage for resources simultaneously.
+A handful of empirical patterns support the hypothesis that character
+divergence decreases food-web resilience. For example, a single species
+of threespine stickleback lives in hundreds of small coastal lakes of
+British Columbia, but the species pair, where character divergence has
+resulted in specialized limnetic and benthic species, are only known
+from six lakes in four independent watersheds
+\citep{Schluter1992, Schluter2000}. Perhaps many lakes had a species
+pair in the past, but have lost a species due to a less resilient
+food-web structure \citep{Borrelli2015a, Borrelli2015b}. The species
+pair are known to be vulnerable to perturbations, as they have gone
+extinct in two of the six lakes after the introduction of nonnative
+species \citep{Hatfield2001, Taylor2006, Rudman2016}. The vulnerability
+of the stickleback system also corresponds with the fact that aquatic
+food webs have several properties that make them less resilient to
+perturbations, such as higher productivity and more efficient energy
+transfer to consumers \citep{Rip2011}. Detecting the ghost of
+competition past \citep{Connell1980} may be quite difficult, but it
+could be possible with recent advances in genomics. For example,
+\citet{Feulner2018} detected genomic signatures of hybridization in
+sympatric whitefish species following periods of eutrophication. Perhaps
+solitary stickleback in some lakes retain genomic signatures of having
+been a habitat specialist in the past.
 
-These results do lead to an interesting prediction that could be tested
-either experimentally or with field data. If consumers are competing for
-resources that occur in different habitats, then greater character
-displacement corresponds to a greater decrease in resource densities
-(line trajectories in Fig. \ref{fig:LS_Resources}B). This prediction
-likely only applies to comparisons within species where there is likely
-little difference in the shape of the tradeoff among populations.
+My results contrast, but do not necessarily contradict, the notion from
+coexistence theory that character displacement contributes to species
+coexistence \citep{Lawlor1976}. Rather than studying resilience,
+coexistence theory usually studies the mutual ability of consumers with
+different phenotypes to invade when rare \citep[mutual
+invasibility,][]{Chesson2000}. In the context of character displacement,
+a shortcoming of this mutual invasibility measure is that it does not
+allow a comparison between food webs with and without a competing
+consumer. Such comparisons are necessary for inferring the effects of
+character displacement, a point that has been made clear in the criteria
+to demonstrate character displacement
+\citep{Schluter1992, Schluter2000}. Although the addition of a consumer
+to a food web can decrease its resilience in the absence of evolution
+\citep{May1973}, my results are primarily driven by an eco-evolutionary
+feedback between consumer evolution and resource abundances.
 
-In general, I find that resource supression goes hand-in-hand with local
-stability, a tendency that has been noted by others (Murdoch, Briggs,
-and Nisbet (2003); Kevin S McCann (2011)).
+\subsection{Caveats}\label{caveats}
 
-The scenario where the relationship between character displacement and
-resource suppression is the strongest is when the tradeoff in attack
-rates is concave-up and consumers are competing for resources that occur
-in different habitats (orange points in Fig. \ref{fig:LS_Resources}B and
-Fig. \ref{fig:McCann_Resources}B). Prior work in threespine stickleback
-suggests that the evolution of stickleback attack rates are constrained
-by a concave-up tradeoff (Dolph Schluter (1993); Arnegard et al. (2014))
-and that stickleback ecotypes are competing for resources that occur in
-different habitats (benthic vs.~limnetic, D Schluter and McPhail
-(1992)). Thus, I predict that ecological character displacement in
-sticklebacks would suppress resource densities and decrease local
-stability.
+Although I model the indirect effects of coevolution between consumers,
+I do not account for potential coevolution between consumers and
+resources. In the context of my model, I would expect prey to evolve
+traits that reduce consumer attack rates. Thus, prey evolution would act
+to counter the effects of character displacement on resource abundance
+and food-web stability. Note that this does not negate my general
+conclusion that ecological character displacement decreases resource
+abundances and stability; however, this process may itself create
+another eco-evolutionary feedback between consumers and resources. This
+may actually help maintain dramatic examples of character divergence and
+prevent them from destabilizing systems, because it allows consumer
+traits to become decoupled from their attack rate. Examining this
+decoupling would be ideal in a quantitative genetic model that can
+explicitly track trait dynamics, but it would not fundamentally change
+the conclusions presented here.
 
-While measuring resource densities is relatively straightforward,
-measuring stability is notoriously difficult. One of the more easy to
-acquire empirical metrics of stability is the coefficient of variation
-(\(CV=\frac{\sigma}{\mu}\)) as this only requires measurements of
-resource (or consumer) densities over time. Interestingly, there is a
-close correspondence between the coefficient of variation and the local
-stability of a food web (Gellner, McCann, and Hastings (2016)). Thus, I
-predict that character displacement in threespine stickleback will
-increase the \(CV\) of resource (or consumer) densities.
+Another potential caveat is that I explored my model in a setting that
+makes many assumptions about resource and consumer symmetry (but see
+consumer asymmetry section). Prior work has shown that allowing for
+resource asymmetry, for example, may decrease the magnitude of character
+displacement \citep{Abrams1986}. This is because this asymmetry creates
+an asymmetry in resource abundances, which dampens the effects of
+character divergence. Again though, this may dampen the effects of
+divergence, but should not change the qualitative relationship we
+observed of ecological character displacement decreasing resource
+abundances and food-web stability.
 
-Consumer and resource densities in these dynamical models can be
-interpreted from either a population or biomass perspective (Yodzis and
-Innes (1992); Murdoch, Briggs, and Nisbet (2003); Kevin S McCann
-(2011)). For the stickleback system, you have data on seasonal dynamics
-of resources in terms of both abundances and biomass. Within a season, I
-do not expect stickleback abundances to be coupled to the abundances of
-zooplankton and benthic invertebrates, given that stickleback reproduce
-annually whereas zooplankton and benthic invertebrates can reproduce
-many times within a season. Instead, I expect that stickleback biomass
-will be coupled to the biomass of zooplankton and benthic invertebrates
-within a season. This is because there will be lots of young stickleback
-at the beginning of the season, and they will eat a lot of resources and
-grow in size as the season progresses. If this is true, I also expect to
-see the largest negative effect of stickleback on resource densities
-late in the season.
+I studied this eco-evolutionary feedback between consumers and resources
+using an Adaptive Dynamics approach. A strength of this approach is that
+it enabled me to gain analytical insight to the effects of character
+divergence in a more realistic foraging scenario. This is much less
+tractable in quantitative genetic \citep{Taper1985, McPeek2017} or
+explicit genetic \citep{Doebeli1996} models of character displacement,
+which is why the foraging scenarios previously examined have been
+limited \citep[but see][]{McPeek2017}. A weakness, however, is that I
+assume a separation of time scales between ecological and evolutionary
+dynamics, an assumption that is becoming less tenable
+\citep{Hairston2005, Hendry2016}. I also do not explicitly model an
+underlying phenotypic trait for consumer attack rates nor do I allow for
+intraspecific variation. That being said, my theoretical predictions are
+likely robust to these assumptions. This is because models that
+explicitly include resource dynamics inevitably show that resource
+competition results in character divergence, regardless of whether a
+quantitative genetic or Adaptive Dynamics approach is used
+\citep{Lawlor1976, Taper1985}. A quantitative genetic model may
+certainly show differences in the pace of character divergence, but this
+should not qualitatively change its effect on food-web dynamics. It is
+important to note that my conclusions only apply to food webs with
+biotic resources that are nutritionally substitutable. It would be
+interesting to extend these current analyses to non-substitutable
+resources where character convergence is expected
+\citep{Abrams1987, Fox2008}.
 
-Although I used an Adaptive Dynamics approach here, I expect that these
-theoretical predictions would apply to models that assumed a different
-genetic architecture of attack rates (e.g.~quantitative genetics). I
-expect this because, time and time again, models that explicitly include
-resource dynamics inevitably show that resource competition results in
-divergent character displacement, regardless of the genetic architecture
-of traits (see synthesis by Taper and Chase (1985)). Finally, my
-conclusions only apply to biological resources that are nutritionally
-substitutable. It would be interesting to extend these current analyses
-to non-substitutable resources where convergent character displacement
-is expected (Peter A Abrams (1987); Fox and Vasseur (2008)).
+\subsection{Conclusions}\label{conclusions}
+
+Here, I show that an adaptive process that generates phenotypic
+diversity generally makes that diversity more susceptible to future
+extinctions. This destabilizing effect emerges from an eco-evolutionary
+feedback involving direct and indirect interactions between species in a
+food-web context. This result contrasts with the current notion that
+patterns of phenotypic diversity are solely the result of evolutionary
+constraints imposed by mutation, natural selection, gene flow, and
+genetic drift. In particular, my result supports the recent suggestion
+that food-web stability can impose an ecological constraint on
+phenotypic diversity that is agnostic to these evolutionary processes
+\citep{Borrelli2015b}. I expect that identifying when and where this
+ecological constraint arises will yield novel insight to the patterns of
+biodiversity we see in nature.
 
@@
 
-The content of this paper was greatly enhanced by numerous conversations
-with Seth Rudman. Sally Otto provided much help in early analyses of the
-mathematical model and I would not have been able to take the theory as
-far as I did without her guidance and her pushing me.
+This work was inspired by discussions with Seth Rudman, Dolph Schluter,
+Ben Gilbert, and Kevin McCann. Sally Otto provided much help in early
+analyses of the mathematical model. I would not have been able to take
+the theory as far as I did without her guidance and encouragement. For
+funding support, I thank the University of British Columbia (Four-Year
+Fellowship to M.A.~Barbour), NSERC (Discovery grant to Greg Crutsinger),
+and the Swiss National Science Foundation (grant 31003A\_160671 to Jordi
+Bascompte).
 
-Thank Ben Gilbert for his comment during a discussion of the model that
-pushed me to take a comparative approach that looked across different
-model structures to examine the generality of these results.
-
-\section{References}\label{references}
-
-\hypertarget{refs}{}
-\hypertarget{ref-Abrams1986}{}
-Abrams, P A. 1986. ``Character Displacement and Niche Shift Analyzed
-Using Consumer-Resource Models of Competition.'' \emph{Theor. Popul.
-Biol.} 29 (1): 107--60.
-
-\hypertarget{ref-Abrams1987}{}
-Abrams, Peter A. 1987. ``Alternative Models of Character Displacement
-and Niche Shift. I. Adaptive Shifts in Resource Use When There Is
-Competition for Nutritionally Nonsubstitutable Resources.''
-\emph{Evolution} 41 (3). Wiley Online Library: 651--61.
-
-\hypertarget{ref-Arnegard2014}{}
-Arnegard, Matthew E, Matthew D McGee, Blake Matthews, Kerry B Marchinko,
-Gina L Conte, Sahriar Kabir, Nicole Bedford, et al. 2014. ``Genetics of
-Ecological Divergence During Speciation.'' \emph{Nature} 511 (7509):
-307--11.
-
-\hypertarget{ref-Dayan2005}{}
-Dayan, Tamar, and Daniel Simberloff. 2005. ``Ecological and
-Community-Wide Character Displacement: The Next Generation.''
-\emph{Ecol. Lett.} 8 (8). Blackwell Science Ltd: 875--94.
-
-\hypertarget{ref-Doebeli1996}{}
-Doebeli, Michael. 1996. ``An Explicit Genetic Model for Ecological
-Character Displacement.'' \emph{Ecology} 77 (2). Ecological Society of
-America: 510--20.
-
-\hypertarget{ref-Fox2008}{}
-Fox, Jeremy W, and David A Vasseur. 2008. ``Character Convergence Under
-Competition for Nutritionally Essential Resources.'' \emph{The American
-Naturalist} 172 (5). The University of Chicago Press: 667--80.
-
-\hypertarget{ref-Gellner2016}{}
-Gellner, Gabriel, Kevin S McCann, and Alan Hastings. 2016. ``The Duality
-of Stability: Towards a Stochastic Theory of Species Interactions.''
-\emph{Theor. Ecol.} 9 (4): 477--85.
-
-\hypertarget{ref-Lawlor1976}{}
-Lawlor, Lawrence R, and J Maynard Smith. 1976. ``The Coevolution and
-Stability of Competing Species.'' \emph{Am. Nat.} 110 (971).
-{[}University of Chicago Press, American Society of Naturalists{]}:
-79--99.
-
-\hypertarget{ref-MacArthur1972}{}
-MacArthur, R H. 1972. \emph{Geographical Ecology: Patterns in the
-Distribution of Species}. Biology / {[}Princeton University Press{]}.
-Princeton University Press.
-
-\hypertarget{ref-McCann2005}{}
-McCann, K S, J B Rasmussen, and J Umbanhowar. 2005. ``The Dynamics of
-Spatially Coupled Food Webs.'' \emph{Ecol. Lett.} 8 (5). Blackwell
-Science Ltd: 513--23.
-
-\hypertarget{ref-McCann2011}{}
-McCann, Kevin S. 2011. \emph{Food Webs (Mpb-50)}. Princeton University
-Press.
-
-\hypertarget{ref-Murdoch2003}{}
-Murdoch, W W, C J Briggs, and R M Nisbet. 2003. \emph{Consumer-Resource
-Dynamics}. Monographs in Population Biology. Princeton University Press.
-
-\hypertarget{ref-Otto2007}{}
-Otto, Sarah P, and Troy Day. 2007. \emph{A Biologist's Guide to
-Mathematical Modeling in Ecology and Evolution}. Princeton University
-Press.
-
-\hypertarget{ref-Pfennig2010}{}
-Pfennig, David W, and Karin S Pfennig. 2010. ``Character Displacement
-and the Origins of Diversity.'' \emph{Am. Nat.} 176 Suppl 1 (December):
-S26--44.
-
-\hypertarget{ref-Sargent2006}{}
-Sargent, Risa D, and Sarah P Otto. 2006. ``The Role of Local Species
-Abundance in the Evolution of Pollinator Attraction in Flowering
-Plants.'' \emph{Am. Nat.} 167 (1): 67--80.
-
-\hypertarget{ref-Schluter1992}{}
-Schluter, D, and J D McPhail. 1992. ``Ecological Character Displacement
-and Speciation in Sticklebacks.'' \emph{Am. Nat.} 140 (1): 85--108.
-
-\hypertarget{ref-Schluter1993}{}
-Schluter, Dolph. 1993. ``Adaptive Radiation in Sticklebacks: Size,
-Shape, and Habitat Use Efficiency.'' \emph{Ecology} 74 (3): 699--709.
-
-\hypertarget{ref-Schluter2000}{}
----------. 2000. ``Ecological Character Displacement in Adaptive
-Radiation.'' \emph{Am. Nat.} 156 (S4): S4--S16.
-
-\hypertarget{ref-Smith1973}{}
-Smith, J Maynard, and G R Price. 1973. ``The Logic of Animal Conflict.''
-\emph{Nature} 246 (November). Nature Publishing Group: 15.
-
-\hypertarget{ref-Stuart2013}{}
-Stuart, Yoel E, and Jonathan B Losos. 2013. ``Ecological Character
-Displacement: Glass Half Full or Half Empty?'' \emph{Trends Ecol. Evol.}
-28 (7): 402--8.
-
-\hypertarget{ref-Taper1985}{}
-Taper, Mark L, and Ted J Chase. 1985. ``Quantitative Genetic Models for
-the Coevolution of Character Displacement.'' \emph{Ecology} 66 (2).
-Ecological Society of America: 355--71.
-
-\hypertarget{ref-Yodzis1992}{}
-Yodzis, P, and S Innes. 1992. ``Body Size and Consumer-Resource
-Dynamics.'' \emph{Am. Nat.} 139 (6). {[}University of Chicago Press,
-American Society of Naturalists{]}: 1151--75.
-
-\newpage
-
-\section{Appendix: More realistic functional
-response}\label{appendix-more-realistic-functional-response}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/McCann_Resources-1.pdf}
-\caption{\label{fig:McCann_Resources}Effect of character displacement on
-total attack rates (A) and resource densities (B) when consumers exhibit
-a more realistic functional response.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/Stability_McCann-1.pdf}
-\caption{\label{fig:Stability_McCann}Effect of character displacement on
-local stability when consumers exhibit a more realistic functional
-response.}
-\end{figure}
+\bibliography{references}
 

### 2019-10-30 ESS.R

@@ -1,0 +1,760 @@
+
+## Calculate Evolutionary Stable Strategies (ESS)
+# Evolutionary optimum, where a mutant is no longer able to invade (first derivative = 0 and second derivative is negative)
+
+## Load required libraries
+library(rootSolve)
+library(tidyverse)
+
+## Source in required functions
+source('identify_steady_state.R')
+source('eco_evo_sim_function.R')
+source('Consumer_Resource_Functions.R')
+source('CR_dynamics.R')
+
+## Custom functions
+
+# calculate attack rate on alternative resource
+get_a12 <- function(x, A, n) A*(1-(x/A)^n)^(1/n)
+
+## MacArthur model
+C1_ESS_2C_2R_MacArthur <- function(x, A, n, m, e, dx){
+  
+  # calculate a12 in terms of a11 (x; tradeoff)
+  a11 = x
+  a12 = A*(1-(a11/A)^n)^(1/n)
+  
+  # calculate Resource abundances at equilibrium
+  eq.R1 = 1/(a11 + a12) * m/e
+  eq.R2 = 1/(a11 + a12) * m/e
+  
+  # calculate mutant attack rates
+  a11m = x + dx
+  a12m = A*(1-(a11m/A)^n)^(1/n)
+  
+  # calculate mutant growth rate of C1 when rare
+  e*a11m*eq.R1 + e*a12m*eq.R2 - m
+}
+
+C1_ESS_1C_2R_MacArthur <- function(x, A, n, m, e, dx, r, K, R1, R2, C1){
+  
+  # calculate a12 in terms of a11 (x; tradeoff)
+  a11 <- x
+  a12 <- A*(1-(a11/A)^n)^(1/n)
+  
+  # calculate Resource abundances at equilibrium
+  get_steady <- safe.runsteady(y = c(R1 = R1, R2 = R2, C1 = C1), 
+                               func = MacArthur_1C_2R, 
+                               parms = c(a11 = a11, a12 = a12, 
+                                         r1 = r, r2 = r,
+                                         K1 = K, K2 = K,
+                                         e11 = e, e12 = e,
+                                         m1 = m)) 
+  
+  eq.R1 = get_steady$y["R1"][[1]]
+  eq.R2 = get_steady$y["R2"][[1]]
+  
[…]

### 2019-10-30 Appendix_S4.Rmd
finalize edits for Appendix S4 for bioRxiv submission

--- 2019-03-22 supplementary_ECD_model.Rmd
@@
     citation_package: natbib 
-#    includes:  
+    #includes:  
     #  in_header: Am-Nat-preamble-latex.tex
 header-includes:
+  - \usepackage{fullpage}
+  - \linespread{1.5} 
+  - \usepackage{lineno}
   - \usepackage{caption}
-- \captionsetup[figure]{labelformat=empty}
-- \usepackage{fullpage}
-- \linespread{1.7}
+  - \captionsetup[figure]{font=small,labelformat=empty}
+#- \usepackage{caption}
+#- \captionsetup[figure]{labelformat=empty}
+#- \usepackage{fullpage}
+#- \linespread{1.7}
 bibliography: references
@@
 
-*Elements*: Figures S1 and S2.
+*Elements*: Figures S1, S2, and S3.
 
@@
 
-(ignore this blank page for the moment. Rmarkdown can be a pain in the ass.)
+<!--(ignore this blank page for the moment. Rmarkdown can be a pain in the ass.)-->
 
 ```{r McCann_Plot, echo=FALSE, fig.pos="H", fig.cap="Figure S1: **Effect of character displacement when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in stability panel).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~, whereas as small shapes correspond to the starting points (only in stability panel).
 This figure illustrates how, regardless of the tradeoff, character displacement increases the effective attack rate of consumers.
@@
 
-```{r All asymmetric figs, echo=FALSE, fig.pos="H", fig.cap=""}
+```{r All asymmetric figs, echo=FALSE, fig.pos="H", fig.cap="Figure S2: **Robustness to consumer asymmetry**.
+Lines show predicted values when both consumers and resources are present.
+Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~, whereas as small shapes correspond to the starting points (only in stability panels).
+In both foraging scenarios, feeding rates increase linearly with resource abundance, but the equation for effective attack rate is different.
+The similarity with fig.
+3 (main text) shows that adding an asymmetry in consumer attack rates does not alter the conclusions reported in the main text.
+If anything, it is more likely that an asymmetry will push the system toward the boundary of stability (note position of large circles in lower left stability panel).
+Details of the consumer asymmetry simulation are given in the main text."}
 knitr::include_graphics("asymm_figs.pdf")
 ```
-
-Figure S2: **Robustness to consumer asymmetry**.
-Scenarios correspond to: (I) resources occur in the same habitat and consumer(s) exhibits a linear functional response; (II) resources occur in different habitats and consumer(s) exhibits a linear functional response; (III) resources occur in different habitats and consumer(s) exhibits a more realistic functional response (details in main text).
-Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in stability panels).
-The similarity with fig.
-S1 and fig.
-2 (main text) show that adding an asymmetry in consumer attack rates does not alter the conclusions reported in the main text.
-If anything, it is more likely that an asymmetry will push the system toward the boundary of stability (note position of large circles in stability panel of Scenario I).
-Details of the consumer asymmetry simulation are given in the main text.
 
@@
 
-```{r McCann_Plot_asymm, eval=FALSE, echo=FALSE, fig.cap="\\label{fig:McCann_Plot_asymm}**Effect of character displacement when consumers exhibit a more realistic functional response and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
+```{r McCann_Plot_asymm, echo=FALSE, fig.cap="Figure S3: **Robustness to consumer asymmetry when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~, whereas as small shapes correspond to the starting points.
 Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of consumers.
 This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.
-We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation.
-We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
+The similarity with fig.
+S1 shows that adding an asymmetry in consumer attack rates does not alter the conclusions reported in the main text.
+Details of the consumer asymmetry simulation are given in the main text."}
 knitr::include_graphics("Figure_ECD_McCann_asymm.pdf")

### 2019-10-30 Appendix_S4.tex
finalize edits for Appendix S4 for bioRxiv submission

--- 2019-01-23 manuscript_ECD_model.tex
@@
 
-  \title{Predicting the effects of character displacement on food-web dynamics}
-    \author{Matthew A. Barbour\(^{1,2,\ast}\)}
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{}
+    \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+    \author{}
+    \preauthor{}\postauthor{}
     \date{}
+    \predate{}\postdate{}
 
-%%% From AmNat_MS_template.tex
-%\documentclass[11pt]{article}
-%\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
 \usepackage{fullpage}
-%\usepackage[authoryear,sectionbib,sort]{natbib}
-\linespread{1.7}
-%\usepackage[utf8]{inputenc}
+\linespread{1.5}
 \usepackage{lineno}
-
-%%%%%%% Sections not currently working properly
-%\usepackage{titlesec}
-%\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
-%\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
-%\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
-%\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}
-
-%%% My added packages
-%\usepackage{amsmath} % for matrices
-%\usepackage{graphicx} % for organizing figures
-%\usepackage{lmodern} % for tilde
-%\usepackage[T1]{fontenc} % for tilde
-
-%%%%%%%%% Previous - likely useful though
-%\usepackage{setspace}\doublespacing
-%\usepackage{float}
-%\let\origfigure\figure
-%\let\endorigfigure\endfigure
-%\renewenvironment{figure}[1][2] {
-%    \expandafter\origfigure\expandafter[H]
-%} {
-%    \endorigfigure
-%}
+\usepackage{caption}
+\captionsetup[figure]{font=small,labelformat=empty}
 
@@
 
-\maketitle
+\vspace*{0.1cm}
 
-\noindent 1. University of British Columbia, Department of Zoology,
+\begin{center} \LARGE Appendix S4: \\ Ecological character displacement destabilizes food webs \end{center}
+
+\bigskip
+
+\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}
+
+\bigskip
+
+\noindent 1. University of Zurich, Department of Evolutionary Biology
+and Environmental Studies, 8057 Zurich, Switzerland;
+
+\noindent 2. University of British Columbia, Department of Zoology,
 Vancouver, BC V6T 1Z4, Canada;
-
-\noindent 2. University of Zurich, Department of Evolutionary Biology
-and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich,
-Switzerland;
 
@@
 
-\emph{Manuscript elements}: Figure 1, figure 2, figure 3, figure 4. All
-figures should be printed in color.
-
-\bigskip
-
-\emph{Keywords}: competition; eco-evolutionary dynamics;
-consumer-resource interactions; adaptive radiation; ecological community
-
-\bigskip
-
-\emph{Manuscript type}: Note.
-
-\bigskip
-
-\footnotesize Prepared using an \emph{Am. Nat.} inspired \LaTeX{}
-template for Rmarkdown. \normalsize
-
-\linenumbers{} \modulolinenumbers[3]
+\emph{Elements}: Figures S1, S2, and S3.
 
@@
 
-\section{Abstract}\label{abstract}
+\begin{figure}
+\centering
+\includegraphics{Figure_ECD_McCann.pdf}
+\caption{Figure S1: \textbf{Effect of character displacement when
+consumers exhibit a more realistic functional response.} Different line
+colors correspond to different tradeoff values (green, \emph{n} = 1.15;
+blue, \emph{n} = 1; orange, \emph{n} = 0.85). Large circles (two
+consumers) and triangles (one consumer) correspond to the end points of
+the eco-evolutionary simulation for \emph{C}\textsubscript{1}, whereas
+as small shapes correspond to the starting points (only in stability
+panel). This figure illustrates how, regardless of the tradeoff,
+character displacement increases the effective attack rate of consumers.
+This always resulted in a suppression of resource abundances and
+concomitant decrease in food-web stability. Initial parameter and state
+variables are given in the main text.}
+\end{figure}
 
@@
 
-\section{Introduction}\label{introduction}
-
-Ecological character displacement is thought to be a key evolutionary
-process generating biodiversity (Dolph Schluter 2000; D. W. Pfennig and
-Pfennig 2010; but see Stuart and Losos 2013). Ecological character
-displacement describes the ``process of phenotypic evolution in a
-species generated or maintained by {[}exploitative{]} resource
-competition with one or more coexisting species'' (Dolph Schluter 2000).
-Over the past 40 years, a large body of theoretical (e.g. Lawlor and
-Smith 1976; P A Abrams 1986; Doebeli 1996; Taper and Chase 1985) and
-empirical (reviewed in: Dolph Schluter 2000; Dayan and Simberloff 2005;
-Stuart and Losos 2013) work has been generated to understand the
-scenarios under which exploitative competition for resources leads to
-the divergence of consumer foraging traits. The general result emerging
-from these studies is that, if resources are nutritionally substitutable
-(Peter A Abrams 1987; Fox and Vasseur 2008) and there is no other strong
-source of density dependence acting on consumers (P A Abrams 1986), then
-resource competition drives the divergence of consumer foraging traits
-{[}Lawlor and Smith (1976); Taper and Chase (1985)). This process is not
-simply driven by ecological differences, but creates an eco-evolutionary
-feedback that drives further differentiation. This key insight was made
-by theoretical models that explicitly included resource dynamics as a
-mediator of competition in driving evolutionary change (Lawlor and Smith
-1976; P A Abrams 1986; Taper and Chase 1985).
-
-Although explicit models of resources gave key insights to the evolution
-of character displacement, the ecological feedback onto
-consumer-resource dynamics have received surprisingly little attention.
-This is likely because character displacement has been primarily studied
-through the lens of competition and coexistence theory {[}Lawlor1976;
-Germain2018{]}. For example, early theoretical work by Lawlor and Smith
-(1976) showed that ecological character displacement promotes
-coexistence by favoring specialized consumers that experience reduced
-interspecific competition. On the other hand, specialization often
-increases a consumer's ability to capture a specific resource. A well
-known consequence of increased foraging efficiency is resource
-suppression, which if sufficient enough, can generate oscillations, and
-hence less stable, consumer-resource dynamics {[}Rosenzweig 1971;
-Murdoch et al. 2003; McCann 2011{]}. NEED A LAST SENTENCE TO WRAP IT UP,
-HIGHLIGHTING HOW A FOOD-WEB PERSPECTIVE MAY YIELD NEW ECOLOGICAL
-INSIGHTS AND GIVE A CONTRASTING PERSPECTIVE TO COEXISTENCE THEORY.
-PERHAPS A I NEED TO INCLUDE A SENTENCE MID-WAY THROUGH SUGGESTING THAT A
-FOOD-WEB PERSPECTIVE MAY GIVE QUALITATIVELY DIFFERENT INSIGHTS. IT MAY
-BE NICE TO BRING IN THE SPATIAL COUPLING WORK OF MCCANN SUGGESTING THAT
-SPECIALIZATION CAN COUPLE FOOD-WEBS, BUT THIS WAS ONLY WITH ONE TOP
-PREDATOR (I.E. NO CONSUMERS\ldots{}LOOK AT OLD TEXT FOR WORDING
-SUGGESTIONS.)
-
-Here, I address this knowledge gap by building a mathematical model to
-examine how ecological character displacement affects consumer-resource
-dynamics in a food-web context. I address two questions: (1) How does
-ecological character displacement affect resource abundances? (2) How
-does character displacement affect food-web stability? To test the
-generality of these ecological effects, I examined three different
-foraging scenarios: (i) resources occur in the same habitat and
-consumers exhibit a linear functional response (e.g. MacArthur (1972);
-Lawlor and Smith (1976)); (ii) resources occur in different habitats and
-consumers exhibit a linear functional response (e.g. Lawlor and Smith
-(1976)); (iii) resources occur in distinct habitats and consumers
-exhibit a more realistic functional response (K S McCann, Rasmussen, and
-Umbanhowar 2005). I also explore the contingency of these effects to
-different assumptions about evolutionary trade-offs in consumer foraging
-traits. I focus on the ecological side of these dynamics because the
-evolutionary aspects have been well characterized (Lawlor and Smith
-1976; Taper and Chase 1985; P A Abrams 1986). I show that the
-consequences of character displacement on food-web dynamics are not
-always intuitive and depend on the foraging scenario and tradeoff in
-consumer traits.
-
-\section{Methods and Results}\label{methods-and-results}
-
-\subsection{Underlying Food-web
-Dynamics}\label{underlying-food-web-dynamics}
-
-To examine how ecological character displacement affects
-consumer-resource interactions and food-web dynamics, I analyzed a
-continuous-time model of two consumers (\(C_1,C_2\)) competing for two
-shared resources (\(R_1,R_2\)):
-
-\[\frac{dR_1}{dt}=r_1K_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{21}(R_1)C_2,\]
-\[\frac{dR_2}{dt}=r_2K_2(1-\frac{R_2}{K_2})-F_{12}(R_2)C_1-F_{22}(R_2)C_2,\]
-\[\frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{12}F_{12}(R_2)C_1-m_1C_1,\]
-\[\frac{dC_2}{dt}=e_{21}F_{21}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2,\]
-
-where \(r_i\) represents the intrinsic growth rate of resource \(i\),
-\(K_i\) represents the carrying capacity of resource \(i\), \(e_{ji}\)
-represents the conversion efficiency of resource \(i\) into consumer
-\(j\), and \(m_j\) represents the mortality rate of consumer \(j\).
-\(F_{ji}(R_i)\) represents consumer \(j\)'s feeding rate
-(i.e.~functional response) on resource \(i\). This model is a useful
-characterization of a scenario where consumers compete for two distinct
-resources (e.g.~zooplankton and benthic invertebrates) rather than if
-resources are better characterized by a continuous trait distribution
-(e.g.~seed size, see Taper and Chase (1985) for an example). In the
-scenario where consumer feeding rates increase linearly with resource
-density, such that \(F_{ji}(R_i)=a_{ji}R_i\) where \(a_{ji}\) is the
-attack rate of consumer \(j\) on resource \(i\), then this model becomes
-MacArthur's model of resource competition (MacArthur (1972)). The
-evolution of consumer attack rates in this model (and several
-extensions) have been analyzed in detail (Lawlor and Smith (1976); P A
-Abrams (1986)), with the general result being divergent character
-displacement. I say consumers have undergone divergent character
-displacement if there evolved attack rates are more specialized, defined
-as \(|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|\), when evolving with a
-competing consumer. I focus here on the ecological consequences of this
-adaptive divergence.
-
-To investigate the ecological consequences, I studied the effect of
-adaptive divergence in consumers on resource densities and the stability
-of the entire food web at equilibrium. An equilibrium is reached when
-the four rates of change in the above equations are 0, and solving the
-system at this point gives equilibrium densities for each resource
-(\(\hat R_1,\hat R_2\)) and consumer (\(\hat C_1,\hat C_2\)). At the
-coexistence equilibrium (i.e., where
-\(\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0\)), I calculated resource
-densities and performed a linear stability analysis (Otto and Day 2007).
-This stability analysis derives the dominant (largest in absolute value)
-eigenvalue, \(\lambda\), which determines whether (and how readily) the
-food web, when perturbed a small amount, will return to equilibrium (see
-supplementary material for derivation). Together, these two measures
-indicate how ecological character displacement affects resource
-densities and food-web stability.
-
-I examine the effects of this inevitable divergence in attack rates on
-food-web dynamics by (i) deriving analytical expressions for the
-relationship between attack rates and food-web dynamics and (ii)
-simulating the effects of competition on the eco-evolutionary dynamics
-of consumers and resources. To gain analytical insight, I assume that
-resources are equivalent (\(r=r_1=r_2\) and \(K=K_1=K_2\)) as well as
-consumers (\(e=e_{11}=e_{12}=e_{21}=e_{22}\); \(m=m_1=m_2\)), except
-that consumer attack rates are mirror images of each other
-(\(a_{ii}=a_{11}=a_{22}\); \(a_{ij}=a_{12}=a_{21}\)). While this
-scenario is arguably simplistic, it allows me to isolate the effects of
-ecological character displacement.
-
-To simulate eco-evolutionary dynamics, I used an Adaptive Dynamics
-approach. At each evolutionary time step, I created a mutant consumer by
-randomly choosing a consumer and modifying its attack rate on one
-resource by either subtracting or adding a small constant (0.001 in the
-following simulations) with equal probability. The mutant's attack rate
-on the other resource was determined by a tradeoff, such that
-\(\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1\), where \(A\) is the
-total investment in attack rates and \(n\) describes the shape of the
-tradeoff (Sargent and Otto 2006). This function has the useful property
-that it differentiates between cases where intermediate combinations of
-\(a_{ii}\) and \(a_{ij}\) are higher, on average, than the extremes
-(when \(n>1\), green line in Fig. \ref{fig:Tradeoff}) or, conversely,
-where the two extremes are higher, on average, than intermediate
-investments (when \(n<1\), orange line Fig. \ref{fig:Tradeoff}). When
-\(n=1\), the tradeoff function is linear, and all combinations of
-\(a_{ii}\) and \(a_{ij}\) have the same total attack rate (blue line in
-Fig. \ref{fig:Tradeoff}). Assuming the mutant consumer was rare, I then
-determined whether the mutant had higher relative fitness than the
-resident consumer, and thus could invade and replace the resident
-consumer. If the mutant was able to invade, I updated the attack rate of
-the resident consumer to the mutant attack rate and allowed consumer and
-resource dynamics to reach a steady state. I then repeated the
-simulation up to 10,000 times, which was sufficient for consumers to
-either reach an evolutionary stable strategy (ESS, Smith and Price 1973)
-or an evolutionary limit (e.g. \(\frac{a_{ii}}{a_{ii}+a_{ij}}\) is
-constrained to a maximum of 1 and minimum of 0).
-
 \begin{figure}
 \centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/Tradeoff-1.pdf}
-\caption{\label{fig:Tradeoff}Tradeoff forms for consumer attack rates.}
-\end{figure}
-
-\subsection{Consumers Forage For Both Resources
-Simultaneously}\label{consumers-forage-for-both-resources-simultaneously}
-
-When both consumers and resources are present, the density of resources
-at equilibrium are equivalent (\(\hat R = \hat R_1 = \hat R_2\)) and are
-determined by the following equation (derivation in supplementary
-material):
-
-\[\hat{R}=\frac{1}{a_{ii}+a_{ij}}\cdot\frac{m}{e}\] A key determinant of
-resource density in this model is the consumer's total attack rate,
-\(a_{ii}+a_{ij}\). Therefore, the ecological consequences of character
-displacement depend on how the tradeoff influences the evolution of
-consumer attack rates.
-
-My simulations show that the shape of the attack-rate tradeoff
-qualitatively affects the relationship between character displacement
-and resource density (Fig. \ref{fig:MacArthur_Resources}). For example,
-if consumer's are constrained by a linear tradeoff (blue lines), then
-there is no net change in total attack rate (Fig.
-\ref{fig:MacArthur_Resources}A) and character displacement has no effect
-on resource densities (Fig. \ref{fig:MacArthur_Resources}B). If the
-tradeoff is concave down (green lines), then resource abundances can
-actually increase under character displacement (Fig.
-\ref{fig:MacArthur_Resources}B). This is because the total attack rate
-of consumers is maximized at intermediate values (\(a_{ii}=a_{ij}\)) and
-decreases as consumers diverge (Fig. \ref{fig:MacArthur_Resources}A).
-When the tradeoff is concave up (orange lines), character displacement
-suppresses resource densities due to the increase in total attack rates
-(Fig. \ref{fig:MacArthur_Resources}A,B). Although the equation I derived
-for resource density was for the scenario where both consumers and both
-resources are present, it accurately predicts the density of resources
-when a single consumer reaches its ESS (triangles on respective colored
-lines in Fig. \ref{fig:MacArthur_Resources}B). This is because a single
-consumer evolves to be a generalist that has equal attack rates on each
-resource (triangles at 0.5 along x-axis in Fig.
-\ref{fig:MacArthur_Resources}A), resulting in equivalent resource
-densities.
-
-The effect of character displacement on resource density closely matches
-its impact on food-web stability (Fig. \ref{fig:Stability}). For
-example, when character displacement results in resource suppression
-(orange), there is a corresponding decrease in stability. Similarly, if
-character displacement does not influence resource densities (blue),
-then there is no corresponding effect on stability. Curiously, this
-match between impact on resources and stability breakdowns with the
-concave down tradeoff (green).
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/MacArthur_Resources-1.pdf}
-\caption{\label{fig:MacArthur_Resources}Effect of character displacement
-on total attack rates (A) and resource densities (B) when consumers can
-forage for both resources simultaneously.}
-\end{figure}
-
-\includegraphics{Figure_ECD_MacArthur.pdf}
-
-\subsection{\texorpdfstring{Consumers \emph{cannot} forage for both
-resources
-simulanteously}{Consumers cannot forage for both resources simulanteously}}\label{consumers-cannot-forage-for-both-resources-simulanteously}
-
-The only character displacement model that I am aware of that modeled
-resources in different habitats was one examined by Lawlor and Smith
-(1976). This model takes the same form as the previous model, except now
-the consumer's feeding rate takes the form:
-
-\[F_{ji}(R_i)=w_{ji}a_{ji}R_i\]
-
-where \(w_{ji}\) represents the proportion of time consumer \(j\) spends
-foraging in a habitat where only resource \(j\) is found (i.e.~habitat
-preference). Note that since \(w_{ji}\) is a proportion that
-\(w_{jj}=1-w_{ji}\). As with attack rates, I assume that consumer
-habitat preferences are mirror images of each other
-(\(w=w_{11}=w_{22}\)). In Lawlor and Smith (1976)'s detailed
-evolutionary analysis of this model, they again always observe divergent
-character displacement.
-
-To examine the ecological consequences of this model, I again derive the
-equilibrium solution when both consumers and both resources are present.
-Again, resource densities are equivalent at this equilibrium
-(\(\hat R = \hat R_1 = \hat R_2\) when both consumers and resources
-present), but are now determined by the following equation (derivation
-in supplementary material):
-
-\[\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e}\]
-
-This equation implies that if consumers evolve to become specialists on
-resources that occur in their preferred habitat (e.g. \(w>0.5\) and
-\(a_{ii}>a_{ij}\)), then the effective attack rate of consumers
-(\(wa_{ii}+(1-w)a_{ij}\)) will always increase, regardless of the
-tradeoff (Fig. \ref{fig:LS_Resources}A). Thus, character displacement
-always results in resource suppression if consumers are competing for
-resources that occur in different habitats (Fig.
-\ref{fig:LS_Resources}B). Note that the shape of the tradeoff can modify
-the effect of character displacement. This is not so much due to the
-tradeoff affecting the magnitude of displacement (it does, but the
-effect is minor), but because the form of the tradeoff affects resource
-abundances when a single consumer has reached its ESS (triangles in Fig.
-\ref{fig:LS_Resources}B). In contrast, resource densities reach a
-similar value when consumers evolve in the presence of a competitor
-(circles in Fig. \ref{fig:LS_Resources}B), because character
-displacement tends to reach a constraint of complete specialization. It
-is worth noting that in this foraging scenario, resource densities are
-consistently higher at the single consumer ESS compared to the
-predictions I derived for when both consumers are present (deviation of
-triangles from respective colored lines in Fig.
-\ref{fig:LS_Resources}B). This is likely because consumers actually
-evolve to be slightly specialized on the resources that occur in their
-non-preferred habitat (deviation of triangles from 0.5 along x-axis in
-Fig. \ref{fig:LS_Resources}B).
-
-The effect of character displacement on resource densities qualitatively
-matches its negative effect of food-web stability (Fig.
-\ref{fig:Stability}). This is not simply a consequence of having an
-additional consumer in the system, but emerges from the eco-evolutionary
-feedback between character displacement and resource suppression
-(compare initial (small points) with evolved strategies (large points),
-Fig. \ref{fig:Stability}). For example, when the tradeoff is concave up
-(orange), then the four-species (small circle) food web actually starts
-off as less stable than the three-species one (small triangle), but the
-eco-evolutionary dynamics cause a switch in the pattern once they reach
-their evolutionary stable strategies (large points).
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/LS_Resources-1.pdf}
-\caption{\label{fig:LS_Resources}Effect of character displacement on
-effective attack rates (A) and resource densities (B) when consumers
-cannot forage for resources simultaneously.}
-\end{figure}
-
-\includegraphics{Figure_ECD_LS.pdf}
-
-\subsection{Adding A More Realistic Functional
-Response}\label{adding-a-more-realistic-functional-response}
-
-Thus far I have consider a simple type I functional response of
-consumers foraging for resources in either the same or distinct
-habitats.\\
-For example, even if I consider a more realistic consumer functional
-response (derived by K S McCann, Rasmussen, and Umbanhowar (2005)):
-
-\[F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}hW_{ii}R_i+a_{ij}hW_{ij}R_j}\]
-
-where consumer feeding rates on resource \(i\) are influenced by both
-resource densities; saturate as resource densities increase (due to
-handling time \(h\)); and consumer habitat preferences are modified by
-relative resource densities, such that:
-
-\[W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}\]
-
-I still observe the same qualitative relationship between character
-displacement and resource suppression (Fig. \ref{fig:McCann_Resources}
-in Appendix). This is because resource competition always leads to
-character divergence and equilibrium resource densities are governed by
-similar dynamics (\emph{Mathematica} file with derivations available
-upon request):
-
-\[\hat{R}=\frac{1}{wa_{ii}+(1-w)a_{ij}}\cdot\frac{m}{e-hm}\]
-
-I conducted the previous simulations on a scenario where the food web
-remains locally stable (\(\operatorname{Re}(\lambda_{max})<0\)). Is it
-possible that character displacement would ever result in a locally
-unstable food web? To gain analytical insight to this question, I
-examined the conditions governing the local stability of the
-four-species food web when consumers exhibit a more realistic functional
-response (K S McCann, Rasmussen, and Umbanhowar (2005)). I show that the
-four-species food web will transition from having a locally stable
-equilibrium to a limit cycle under the following conditions
-(\emph{Mathematica} file with derivation available upon request):
-
-\[wa_{ii}+(1-w)a_{ij} > \frac{e+hm}{hK(e-hm)}\]
-
-Previously, I showed that character displacement always increases the
-effective attack rate of consumers (\(wa_{ii}+(1-w)a_{ij}\)), regardless
-of the the shape of the tradeoff (Fig. \ref{fig:LS_Resources}A). Thus,
-character displacement is capable of pushing food webs to a point where
-they are locally unstable. (it simple depends on the choice of
-parameter, I could put a situation where evolution pushes it toward
-local instability, or even the hopf bifurcation.)
-
-\section{Discussion}\label{discussion}
-
-I find that the foraging scenario qualitatively affects the relationship
-between character displacement and resource densities.
-
-Most models of character displacement have assumed a scenario where
-consumers can forage for both resources simultaneously (e.g. Taper and
-Chase 1985; P A Abrams 1986). This assumption may be valid for some
-foraging scenarios (e.g.~Darwin's finches foraging for seeds); however,
-in many situations consumers forage for resources that occur in
-different habitats, and thus cannot forage for both simultaneously (K S
-McCann, Rasmussen, and Umbanhowar 2005).
-
-The effect of the foraging scenario on the relationship between
-character displacement and resource densities appears to be quite
-general.
-
-Note that detecting this effect in nature would likely be difficult,
-since a concave-down tradeoff results in relatively small displacement
-relative to other tradeoff shapes (Lawlor and Smith (1976); green points
-in Fig. \ref{fig:MacArthur_Resources}A,B).
-
-Previous work has shown that consumers undergo divergent character
-displacement regardless of the shape of the tradeoff (Lawlor and Smith
-(1976); P A Abrams (1986)).
-
-Important point: Criteria 5 in D Schluter and McPhail (1992) does not
-make sense since character displacement usually alters resource
-distributions. Think about this more.
-
-Lawlor and Smith (1976) found that consumers still underwent divergent
-character displacement regardless of whether consumers could or could
-not forage for resources simultaneously.
-
-These results do lead to an interesting prediction that could be tested
-either experimentally or with field data. If consumers are competing for
-resources that occur in different habitats, then greater character
-displacement corresponds to a greater decrease in resource densities
-(line trajectories in Fig. \ref{fig:LS_Resources}B). This prediction
-likely only applies to comparisons within species where there is likely
-little difference in the shape of the tradeoff among populations.
-
-In general, I find that resource supression goes hand-in-hand with local
-stability, a tendency that has been noted by others (Murdoch, Briggs,
-and Nisbet (2003); Kevin S McCann (2011)).
-
-The scenario where the relationship between character displacement and
-resource suppression is the strongest is when the tradeoff in attack
-rates is concave-up and consumers are competing for resources that occur
-in different habitats (orange points in Fig. \ref{fig:LS_Resources}B and
-Fig. \ref{fig:McCann_Resources}B). Prior work in threespine stickleback
-suggests that the evolution of stickleback attack rates are constrained
-by a concave-up tradeoff (Dolph Schluter (1993); Arnegard et al. (2014))
-and that stickleback ecotypes are competing for resources that occur in
-different habitats (benthic vs.~limnetic, D Schluter and McPhail
-(1992)). Thus, I predict that ecological character displacement in
-sticklebacks would suppress resource densities and decrease local
-stability.
-
-While measuring resource densities is relatively straightforward,
-measuring stability is notoriously difficult. One of the more easy to
-acquire empirical metrics of stability is the coefficient of variation
-(\(CV=\frac{\sigma}{\mu}\)) as this only requires measurements of
-resource (or consumer) densities over time. Interestingly, there is a
-close correspondence between the coefficient of variation and the local
-stability of a food web (Gellner, McCann, and Hastings (2016)). Thus, I
-predict that character displacement in threespine stickleback will
-increase the \(CV\) of resource (or consumer) densities.
-
-Consumer and resource densities in these dynamical models can be
-interpreted from either a population or biomass perspective (Yodzis and
-Innes (1992); Murdoch, Briggs, and Nisbet (2003); Kevin S McCann
-(2011)). For the stickleback system, you have data on seasonal dynamics
-of resources in terms of both abundances and biomass. Within a season, I
-do not expect stickleback abundances to be coupled to the abundances of
-zooplankton and benthic invertebrates, given that stickleback reproduce
-annually whereas zooplankton and benthic invertebrates can reproduce
-many times within a season. Instead, I expect that stickleback biomass
-will be coupled to the biomass of zooplankton and benthic invertebrates
-within a season. This is because there will be lots of young stickleback
-at the beginning of the season, and they will eat a lot of resources and
-grow in size as the season progresses. If this is true, I also expect to
-see the largest negative effect of stickleback on resource densities
-late in the season.
-
-Although I used an Adaptive Dynamics approach here, I expect that these
-theoretical predictions would apply to models that assumed a different
-genetic architecture of attack rates (e.g.~quantitative genetics). I
-expect this because, time and time again, models that explicitly include
-resource dynamics inevitably show that resource competition results in
-divergent character displacement, regardless of the genetic architecture
-of traits (see synthesis by Taper and Chase (1985)). Finally, my
-conclusions only apply to biological resources that are nutritionally
-substitutable. It would be interesting to extend these current analyses
-to non-substitutable resources where convergent character displacement
-is expected (Peter A Abrams (1987); Fox and Vasseur (2008)).
-
-\section{Acknowledgements}\label{acknowledgements}
-
-The content of this paper was greatly enhanced by numerous conversations
-with Seth Rudman. Sally Otto provided much help in early analyses of the
-mathematical model and I would not have been able to take the theory as
-far as I did without her guidance and her pushing me.
-
-Thank Ben Gilbert for his comment during a discussion of the model that
-pushed me to take a comparative approach that looked across different
-model structures to examine the generality of these results.
-
-\section{References}\label{references}
-
-\hypertarget{refs}{}
-\hypertarget{ref-Abrams1986}{}
-Abrams, P A. 1986. ``Character Displacement and Niche Shift Analyzed
-Using Consumer-Resource Models of Competition.'' \emph{Theor. Popul.
-Biol.} 29 (1): 107--60.
-
-\hypertarget{ref-Abrams1987}{}
-Abrams, Peter A. 1987. ``Alternative Models of Character Displacement
-and Niche Shift. I. Adaptive Shifts in Resource Use When There Is
-Competition for Nutritionally Nonsubstitutable Resources.''
-\emph{Evolution} 41 (3). Wiley Online Library: 651--61.
-
-\hypertarget{ref-Arnegard2014}{}
-Arnegard, Matthew E, Matthew D McGee, Blake Matthews, Kerry B Marchinko,
-Gina L Conte, Sahriar Kabir, Nicole Bedford, et al. 2014. ``Genetics of
-Ecological Divergence During Speciation.'' \emph{Nature} 511 (7509):
-307--11.
-
-\hypertarget{ref-Dayan2005}{}
-Dayan, Tamar, and Daniel Simberloff. 2005. ``Ecological and
-Community-Wide Character Displacement: The Next Generation.''
-\emph{Ecol. Lett.} 8 (8). Blackwell Science Ltd: 875--94.
-
-\hypertarget{ref-Doebeli1996}{}
-Doebeli, Michael. 1996. ``An Explicit Genetic Model for Ecological
-Character Displacement.'' \emph{Ecology} 77 (2). Ecological Society of
-America: 510--20.
-
-\hypertarget{ref-Fox2008}{}
-Fox, Jeremy W, and David A Vasseur. 2008. ``Character Convergence Under
-Competition for Nutritionally Essential Resources.'' \emph{The American
-Naturalist} 172 (5). The University of Chicago Press: 667--80.
-
-\hypertarget{ref-Gellner2016}{}
-Gellner, Gabriel, Kevin S McCann, and Alan Hastings. 2016. ``The Duality
-of Stability: Towards a Stochastic Theory of Species Interactions.''
-\emph{Theor. Ecol.} 9 (4): 477--85.
-
-\hypertarget{ref-Lawlor1976}{}
-Lawlor, Lawrence R, and J Maynard Smith. 1976. ``The Coevolution and
-Stability of Competing Species.'' \emph{Am. Nat.} 110 (971).
-{[}University of Chicago Press, American Society of Naturalists{]}:
-79--99.
-
-\hypertarget{ref-MacArthur1972}{}
-MacArthur, R H. 1972. \emph{Geographical Ecology: Patterns in the
-Distribution of Species}. Biology / {[}Princeton University Press{]}.
-Princeton University Press.
-
-\hypertarget{ref-McCann2005}{}
-McCann, K S, J B Rasmussen, and J Umbanhowar. 2005. ``The Dynamics of
-Spatially Coupled Food Webs.'' \emph{Ecol. Lett.} 8 (5). Blackwell
-Science Ltd: 513--23.
-
-\hypertarget{ref-McCann2011}{}
-McCann, Kevin S. 2011. \emph{Food Webs (Mpb-50)}. Princeton University
-Press.
-
-\hypertarget{ref-Murdoch2003}{}
-Murdoch, W W, C J Briggs, and R M Nisbet. 2003. \emph{Consumer-Resource
-Dynamics}. Monographs in Population Biology. Princeton University Press.
-
-\hypertarget{ref-Otto2007}{}
-Otto, Sarah P, and Troy Day. 2007. \emph{A Biologist's Guide to
-Mathematical Modeling in Ecology and Evolution}. Princeton University
-Press.
-
-\hypertarget{ref-Pfennig2010}{}
-Pfennig, David W, and Karin S Pfennig. 2010. ``Character Displacement
-and the Origins of Diversity.'' \emph{Am. Nat.} 176 Suppl 1 (December):
-S26--44.
-
-\hypertarget{ref-Sargent2006}{}
-Sargent, Risa D, and Sarah P Otto. 2006. ``The Role of Local Species
-Abundance in the Evolution of Pollinator Attraction in Flowering
-Plants.'' \emph{Am. Nat.} 167 (1): 67--80.
-
-\hypertarget{ref-Schluter1992}{}
-Schluter, D, and J D McPhail. 1992. ``Ecological Character Displacement
-and Speciation in Sticklebacks.'' \emph{Am. Nat.} 140 (1): 85--108.
-
-\hypertarget{ref-Schluter1993}{}
-Schluter, Dolph. 1993. ``Adaptive Radiation in Sticklebacks: Size,
-Shape, and Habitat Use Efficiency.'' \emph{Ecology} 74 (3): 699--709.
-
-\hypertarget{ref-Schluter2000}{}
----------. 2000. ``Ecological Character Displacement in Adaptive
-Radiation.'' \emph{Am. Nat.} 156 (S4): S4--S16.
-
-\hypertarget{ref-Smith1973}{}
-Smith, J Maynard, and G R Price. 1973. ``The Logic of Animal Conflict.''
-\emph{Nature} 246 (November). Nature Publishing Group: 15.
-
-\hypertarget{ref-Stuart2013}{}
-Stuart, Yoel E, and Jonathan B Losos. 2013. ``Ecological Character
-Displacement: Glass Half Full or Half Empty?'' \emph{Trends Ecol. Evol.}
-28 (7): 402--8.
-
-\hypertarget{ref-Taper1985}{}
-Taper, Mark L, and Ted J Chase. 1985. ``Quantitative Genetic Models for
-the Coevolution of Character Displacement.'' \emph{Ecology} 66 (2).
-Ecological Society of America: 355--71.
-
-\hypertarget{ref-Yodzis1992}{}
-Yodzis, P, and S Innes. 1992. ``Body Size and Consumer-Resource
-Dynamics.'' \emph{Am. Nat.} 139 (6). {[}University of Chicago Press,
-American Society of Naturalists{]}: 1151--75.
-
-\newpage
-
-\section{Appendix: More realistic functional
-response}\label{appendix-more-realistic-functional-response}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/McCann_Resources-1.pdf}
-\caption{\label{fig:McCann_Resources}Effect of character displacement on
-total attack rates (A) and resource densities (B) when consumers exhibit
-a more realistic functional response.}
+\includegraphics{asymm_figs.pdf}
+\caption{Figure S2: \textbf{Robustness to consumer asymmetry}. Lines
+show predicted values when both consumers and resources are present.
+Different line colors correspond to different tradeoffs in attack rates
+(green, \emph{n} = 1.15; blue, \emph{n} = 1; orange, \emph{n} = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to
+the end points of the eco-evolutionary simulation for
+\emph{C}\textsubscript{1}, whereas as small shapes correspond to the
+starting points (only in stability panels). In both foraging scenarios,
+feeding rates increase linearly with resource abundance, but the
+equation for effective attack rate is different. The similarity with
+fig. 3 (main text) shows that adding an asymmetry in consumer attack
+rates does not alter the conclusions reported in the main text. If
+anything, it is more likely that an asymmetry will push the system
+toward the boundary of stability (note position of large circles in
+lower left stability panel). Details of the consumer asymmetry
+simulation are given in the main text.}
 \end{figure}
@@
 \centering
-\includegraphics{manuscript_ECD_model_files/figure-latex/Stability_McCann-1.pdf}
-\caption{\label{fig:Stability_McCann}Effect of character displacement on
-local stability when consumers exhibit a more realistic functional
-response.}
+\includegraphics{Figure_ECD_McCann_asymm.pdf}
+\caption{Figure S3: \textbf{Robustness to consumer asymmetry when
+consumers exhibit a more realistic functional response.} Different line
+colors correspond to different tradeoff values (green, \emph{n} = 1.15;
+blue, \emph{n} = 1; orange, \emph{n} = 0.85). Large circles (two
+consumers) and triangles (one consumer) correspond to the end points of
+the eco-evolutionary simulation for \emph{C}\textsubscript{1}, whereas
+as small shapes correspond to the starting points. Panel (A) shows how,
+regardless of the tradeoff, character displacement increases the
+effective attack rate of consumers. This always resulted in a
+suppression of resource abundances (B) and concomitant decrease in
+food-web stability (C). The similarity with fig. S1 shows that adding an
+asymmetry in consumer attack rates does not alter the conclusions
+reported in the main text. Details of the consumer asymmetry simulation
+are given in the main text.}
 \end{figure}
 
+\bibliography{references}
+
 \end{document}

### 2019-10-30 SupplementaryMaterialECDModel.Rmd
start of summarizing analytical expressions for eco-evo dynamics (following Matt Osmond's suggestion).

---
title: "Supplementary Material"
author: "Matt Barbour"
date: "6/18/2019"
output:
  pdf_document:
    keep_tex: true
    toc: false
    number_sections: false
    df_print: paged
    fig_caption: true
    citation_package: natbib 
    includes:  
      in_header: Am-Nat-preamble-latex.tex
bibliography: references
biblio-style: amnat
fontsize: 11pt
fontfamily: mathpazo
documentclass: article
linkcolor: black
urlcolor: black
citecolor: black
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Resources occur in same habitat

## Assuming symmetry

In Appendix X, we show that in the absence of competition, consumer $i$ will reach an evolutionary stable strategy when:

$$a_{i,1}=a_{i,2}$$
This indicates that consumers will evolve to be complete generalists.

In the presence of a competitor though, the evolution of attack rates depends on the shape of the tradeoff. 
Specifically, when $n=1$, then the initial attack rates determine the ESS; therefore, the magnitude of character displacement depends on the initial attack rates.
When $n<1$, then consumers will evolve to be completely specialized on the resource that they are pre-adapted to. 
For example, if initial $$a_{i,1}>a_{i,2}$$, then consumer $i$ will evolve until $2a_{i,1}$ and $a_{i,2}=0$, and thus maximal character displacement.
In contrast, when $n>1$, then consumers will evolve to be complete generalists, $a_{i,1}=a_{i,2}$, and thus there will be no character displacement.

In the absence of competition, the ESS in attack rates results in equal resource abundances, which are determined by:

$$R=\frac{m}{2a_{i,1}e}$$
The effects of character displacement on resource abundances depends on the shape of the tradeoff. 
When $n\geq1$, then resource abundances are equivalent for both one and two consumer communities. 
However, when $n<1$, then

# Resources occur in different habitats

## Evolutionary Stable Strategy

In Appendix X, we show that a consumer (we arbitrarily chose $C_1$) will reach an evolutionary stable strategy when:

$$e_{11}R_1w_{11}\delta_{11} + e_{12}R_2w_{12}\delta_{12}=0$$

If there were no constraints, consumers would evolve higher attack rates on both resources. 
In consideration of this, we assume there is a trade-off in attack rates so that $\delta_{12}$ is a negative function of $delta_{11}$, $\delta_{12}=-f(\delta_{11})$.
Therefore, we can rewrite the above expression as

$$e_{11}R_1w_{11}\delta_{11} = e_{12}R_2w_{12}f(\delta_{11})$$

## Resource equilibrium

One consumer:

Two consumers:

## Local stability

One consumer:

Two consumers:

# Adding a more realistic functional response

### 2019-10-30 Playground.R

@@ -1,0 +1,2 @@
+n = 1.5
+(1/2)^(1/n)

### 2019-11-07 ESS.R
move files to MaybeUseful folder outside of repo and delete unnecessary material

@@ -1,760 +1,0 @@
-
-## Calculate Evolutionary Stable Strategies (ESS)
-# Evolutionary optimum, where a mutant is no longer able to invade (first derivative = 0 and second derivative is negative)
-
-## Load required libraries
-library(rootSolve)
-library(tidyverse)
-
-## Source in required functions
-source('identify_steady_state.R')
-source('eco_evo_sim_function.R')
-source('Consumer_Resource_Functions.R')
-source('CR_dynamics.R')
-
-## Custom functions
-
-# calculate attack rate on alternative resource
-get_a12 <- function(x, A, n) A*(1-(x/A)^n)^(1/n)
-
-## MacArthur model
-C1_ESS_2C_2R_MacArthur <- function(x, A, n, m, e, dx){
-  
-  # calculate a12 in terms of a11 (x; tradeoff)
-  a11 = x
-  a12 = A*(1-(a11/A)^n)^(1/n)
-  
-  # calculate Resource abundances at equilibrium
-  eq.R1 = 1/(a11 + a12) * m/e
-  eq.R2 = 1/(a11 + a12) * m/e
-  
-  # calculate mutant attack rates
-  a11m = x + dx
-  a12m = A*(1-(a11m/A)^n)^(1/n)
-  
-  # calculate mutant growth rate of C1 when rare
-  e*a11m*eq.R1 + e*a12m*eq.R2 - m
-}
-
-C1_ESS_1C_2R_MacArthur <- function(x, A, n, m, e, dx, r, K, R1, R2, C1){
-  
-  # calculate a12 in terms of a11 (x; tradeoff)
-  a11 <- x
-  a12 <- A*(1-(a11/A)^n)^(1/n)
-  
-  # calculate Resource abundances at equilibrium
-  get_steady <- safe.runsteady(y = c(R1 = R1, R2 = R2, C1 = C1), 
-                               func = MacArthur_1C_2R, 
-                               parms = c(a11 = a11, a12 = a12, 
-                                         r1 = r, r2 = r,
-                                         K1 = K, K2 = K,
-                                         e11 = e, e12 = e,
-                                         m1 = m)) 
-  
-  eq.R1 = get_steady$y["R1"][[1]]
-  eq.R2 = get_steady$y["R2"][[1]]
-  
[…]

### 2019-11-07 Playground.R

@@ -1,2 +1,0 @@
-n = 1.5
-(1/2)^(1/n)

### 2019-11-07 confirm_asymmetry_effects.R

@@ -1,9 +1,0 @@
-test <- read.csv("asymmetric_displacement_data.csv")
-
-# inspecting the data shows that the starting values are being appropriately switched. Also, we see the same qualitative pattern of displacement in the species that starts as a generalist (becomes a specialist).
-select(test, Model, Competitor, sequence, Time, n, a11:a22, w11:w22, R1:C2, max.Re.eigen, max.Im.eigen) %>%
-  mutate(a11=round(a11,3),
-         a12=round(a12,3),
-         a21=round(a21,3),
-         a22=round(a22,3)) %>%
-  arrange(Model, n, Competitor, Time)

### 2019-11-07 inspect_model_effects.R

@@ -1,17 +1,0 @@
-library(tidyverse)
-test <- read.csv("symmetric_displacement_data.csv")
-
-select(test, Model, Competitor, Time, sequence, feasibility, n, a11:a22, w11:w22, R1:C2) %>%
-  mutate(a11=round(a11,3),
-         a12=round(a12,3),
-         a21=round(a21,3),
-         a22=round(a22,3),
-         R1=round(R1,2),
-         R2=round(R2,2),
-         C1=round(C1,2),
-         C2=round(C2,2)) %>%
-  arrange(Model, n, Competitor, Time)
-
-filter(test, Model=="MacArthur")
-filter(test, Model=="LawlorSmith")
-filter(test, Model=="McCann")

### 2019-11-07 morgue/CopyOfparameterized_CR_model.R

@@ -1,206 +1,0 @@
-
-##############################################################################################################
-## Estimate stickleback consumer-resource dynamics based on allometric scaling and literature.
-## Author: Matt Barbour
-## Email: matthew.a.barbour@gmail.com
-##############################################################################################################
-
-
-## LOAD REQUIRE FUNCTIONS ----
-
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-
-mass_scaling <- function(W, intercept, exponent) {
-  log_rate <- intercept + exponent * log(W)
-  return(exp(log_rate))
-}
-
-
-## BODY SIZE (g) ----
-
-g_zoop <- 0.004/1000            # covert 0.004 mg to g
-g_benthic_invert <- 0.25/1000   # convert 0.25 mg to g
-
-g_limnetic <- 0.81              # from Schluter 1993, Ecology Table 1
-g_benthic <- 1.48               # from Schluter 1993, Ecology Table 1
-g_solitary <- 1.12              # from Schluter 1993, Ecology Table 1
-
-
-## INTRINSIC GROWTH RATE (r) ----
-# r = birth rate per day
-# proportional to M^(-1/4) from Rip and McCann 2011 (they cite Brown et al. 2004)
-# alternatively use scaling exponent = -0.27377 for heterotherm metazoa from Fenchel 1974, Oecologia
-r_exponent <- -0.25             # theoretical scaling.
-r_scale <- 0.3                  # estimate of r from Daphnia pulex in Table 1 of Fenchel 1974. Original study is Frank et al. 1957       
-r_intercept <- log(r_scale) - r_exponent * log(g_zoop) # scale based on zooplankton size
-
-r_benthos_estimate <- mass_scaling(W = g_benthic_invert, intercept = r_intercept, exponent = r_exponent) 
-
-
-#### CARRYING CAPACITY (K) ----
-# proportional to M^(-3/4) from Rip and McCann 2011 (they cite Enquist et al. 1998 and Brown et al. 2004)
-# alternatively use empirical exponent = -0.77 from Brown et al. 2004
-K_exponent <- -0.75 
-K_scale <- 5          # admittedly, this is a bit of an arbitrary number. It does correspond to zooplankton biomass density (g/500 L) from fishless ponds
-K_intercept <- log(K_scale) - K_exponent * log(g_zoop) # scale based on zooplankton size
-
-K_benthos_estimate <- mass_scaling(W = g_benthic_invert, intercept = K_intercept, exponent = K_exponent)
-
-
[…]

### 2019-11-07 morgue/LSC.sym.evo.4sp.R

@@ -1,659 +1,0 @@
-
-#### Description: 4 species model with consumer coevolution simulation,
-####              but this is only for one parameter set where resource
-####              parameters are equivalent and consumers are mirror
-####              images of each other.
-
-## source general parameters and functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-source('Lawlor_Smith_Coarse.R')
-
-trait.thres <- 1e-10
-j.step <- 100 # for other simulations # had at 1000 for some others...
-
-## general parameters for simulations
-r <- 1
-K <- 4
-e <- 0.8
-#h <- 0.4
-aii <- 1.9 #1.93
-aij <- 1.2
-#a <- 2
-w <- 0.5
-m <- 1
-R <- 2
-C <- 1
-
-
-## 4 species simulation ----
-
-## create initial data
-sym4.init.df <- data.frame(r1 = r, r2 = r, K1 = K, K2 = K,
-                           e11 = e, e12 = e, e21 = e, e22 = e,
-                           #h11 = h, h12 = h, h21 = h, h22 = h,
-                           B11 = aii, B12 = aij, B21 = aij, B22 = aii,
-                           w11 = w, w22 = w, 
-                           m1 = m, m2 = m,
-                           R1 = R, R2 = R, C1 = C, C2 = C)
-
-set.temp <- as.matrix(sym4.init.df[1, ])
-
-set.steady <- safe.runsteady(set.temp[1,c("R1","R2","C1","C2")], func = ECD_model_LSC, parms = set.temp[1,c("r1","r2","K1","K2","e11","e12","e21","e22","B11","B12","B21","B22","w11","w22","m1","m2")])
-
-sym4.eq.jac <- jacobian.full(y = set.steady$y, func = ECD_model_LSC, parms = set.temp[1,c("r1","r2","K1","K2","e11","e12","e21","e22","B11","B12","B21","B22","w11","w22","m1","m2")])
-
-## update with new steady state and eigenvalues after the initial run
-sym4.init.df[ ,c("R1","R2","C1","C2")] <- set.steady$y
-sym4.init.df$max.Re.eigen <- max(Re(eigen(sym4.eq.jac)$values))
-sym4.init.df$max.Im.eigen <- max(Im(eigen(sym4.eq.jac)$values))
-
-
-## setup and run simulation
-df <- sym4.init.df
-sim.list.4 <- list()
-for(i in 1:dim(df)[1]){
-  
-  # choose parameter set to examine and calculate the maximum real eigenvalue
[…]

### 2019-11-07 morgue/Lawlor_Smith_Coarse.R

@@ -1,304 +1,0 @@
-######################################################
-## Description: Dynamical model and other useful functions for checking dynamics
-## Code author(s): Matt Barbour
-## Email: barbour@zoology.ubc.ca
-######################################################
-
-
-## Consumer-resource dynamics ----
-# general assumptions:
-# (1) logistic growth for the resource(s)
-# (2) multispecies, spatially-implicit type 2 functional response of consumer 
-# (3) density-independent mortality for consumer 
-# function is in the necessary format "function(Time,State,Pars)" for solving used the ode() function in the R package "deSolve"
-
-## 4 species model: 2 consumers, 2 resources
-ECD_model_LSC <- function(Time,State,Pars) {
-  # t,y,p is the necessary format for solving using the ode() function in R
-  with(as.list(c(State,Pars)), {
-    
-    a11 <- w11 * B11
-    a12 <- w12 * B12
-    a21 <- w21 * B21
-    a22 <- w22 * B22
-    
-    # consumer functional responses
-    C1R1fxn <- a11 * R1 # (W11 * a11 * R1)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
-    C1R2fxn <- a12 * R2 # (W12 * a12 * R2)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
-    C2R1fxn <- a21 * R1 # (W21 * a21 * R1)/(1 + W21 * a21 * h21 * R1 + W22 * a22 * h22 * R2)
-    C2R2fxn <- a22 * R2 # (W22 * a22 * R2)/(1 + W21 * a21 * h21 * R1 + W22 * a22 * h22 * R2)
-    
-    # dynamical equations
-    dR1.dt <- r1 * R1 * (1 - R1 / K1) - C1 * C1R1fxn - C2 * C2R1fxn
[…]

### 2019-11-07 morgue/Pachepsky_within-year_dynamics.R

@@ -1,80 +1,0 @@
-
-
-## source general parameters and functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-source('code/00_Pachepsky_fxns.R')
-
-## general parameters for simulations
-r <- 1
-K <- 4
-#e <- 0.8
-h <- 0.4
-aii <- 10
-aij <- 1
-w <- 0.8
-m <- 0.001
-R <- 2
-C <- 1
-
-## Species-pair lakes
-species_pair.init.states <- c(R1=1, R2=1, C1=0.25, C2=0.25) # note that equilibrium dynamics appear to be independent of initial state variables
-species_pair.parms <- data.frame(r1 = r, r2 = r, K1 = K, K2 = K,
-                                 #e11 = e, e12 = e, e21 = e, e22 = e,
-                                 h11 = h, h12 = h, h21 = h, h22 = h,
-                                 a11 = aii, a12 = aij, a21 = aij, a22 = aii,
-                                 w11 = w, w22 = w, m1 = m, m2 = m)
-
-species_pair_dynamics <- safe.runsteady(y = species_pair.init.states, func = Pachepsky_model, parms = species_pair.parms, stol = 1e-5)
-species_pair_dynamics$y # equilibrium densities
-
-# plot the dynamics
-matplot.0D(ode(species_pair.init.states, 1:100, Pachepsky_model, species_pair.parms), ylim=c(0,K))
-
-
-## Solitary lakes
-w_solitary <- 0.5
-aii_solitary <- (aii + aij)/2
-aij_solitary <- aii_solitary
-
-solitary.init.states <- c(R1=1, R2=1, C1=0.5) # note that equilibrium dynamics appear to be independent of initial state variables
-solitary.parms <- data.frame(r1 = r, r2 = r, K1 = K, K2 = K,
-                             #e11 = e, e12 = e, e21 = e, e22 = e,
-                             h11 = h, h12 = h,
-                             a11 = aii_solitary, a12 = aij_solitary,
-                             w11 = w_solitary, m1 = m) 
-
-solitary_dynamics <- safe.runsteady(y = solitary.init.states, func = Pachepsky_model.C1.3sp, parms = solitary.parms, stol = 1e-5)
-solitary_dynamics$y # equilibrium densities
-
-# plot the dynamics
[…]

### 2019-11-07 morgue/attack_rate_trade_off_function.R

@@ -1,20 +1,0 @@
-# from Sargent and Otto 2006
-# (a11/A)^n + (a12/A)^n = 1
-
-## using Mathematica
-a_ij_tradeoff <- function(A, a_ii, n) {
-  A * (1 - (a_ii/A)^n)^(1/n)
-}
-a_ij_tradeoff(A=2, a_ii=1.5, n=1)
-
-# from Abrams 1986
-# ki*C1i^2 + (1-ki)*C2i^2 = ki*(1-ki)
-a_ij_Abrams <- function(trait, k, n) {
-  ((k * (1 - k) - k * trait^n)/(1 - k))^(1/n)
-}
-df <- data.frame(B11 = 1)
-a_ij_Abrams(trait = 1, k=0.5, n=1)
-
-a_ii_seq <- seq(0,2,0.01)
-
-plot(a_ij_Abrams(ki=0.7, a_ii=a_ii_seq, n=1) ~ a_ii_seq, type='l')

### 2019-11-07 morgue/code/00_MacArthur_fxns.R

@@ -1,255 +1,0 @@
-######################################################
-## Description: Dynamical model and other useful functions for checking dynamics
-## Code author(s): Matt Barbour
-## Email: barbour@zoology.ubc.ca
-######################################################
-
-
-## Consumer-resource dynamics ----
-# general assumptions:
-# (1) logistic growth for the resource(s)
-# (2) multispecies, spatially-implicit type 2 functional response of consumer 
-# (3) density-independent mortality for consumer 
-# function is in the necessary format "function(Time,State,Pars)" for solving used the ode() function in the R package "deSolve"
-
-## 4 species model: 2 consumers, 2 resources
-ECD_model_Mac <- function(Time,State,Pars) {
-  # t,y,p is the necessary format for solving using the ode() function in R
-  with(as.list(c(State,Pars)), {
-    
-    # consumer functional responses
-    C1R1fxn <- a11 * R1 # (W11 * a11 * R1)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
-    C1R2fxn <- a12 * R2 # (W12 * a12 * R2)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
-    C2R1fxn <- a21 * R1 # (W21 * a21 * R1)/(1 + W21 * a21 * h21 * R1 + W22 * a22 * h22 * R2)
-    C2R2fxn <- a22 * R2 # (W22 * a22 * R2)/(1 + W21 * a21 * h21 * R1 + W22 * a22 * h22 * R2)
-    
-    # dynamical equations
-    dR1.dt <- r1 * R1 * (1 - R1 / K1) - C1 * C1R1fxn - C2 * C2R1fxn
-    dR2.dt <- r2 * R2 * (1 - R2 / K2) - C1 * C1R2fxn - C2 * C2R2fxn
-    dC1.dt <- C1 * (e11 * C1R1fxn + e12 * C1R2fxn - m1)
[…]

### 2019-11-07 morgue/code/00_Pachepsky_fxns.R

@@ -1,69 +1,0 @@
-######################################################
-## Description: Dynamical model and other useful functions for checking dynamics
-## Code author(s): Matt Barbour
-## Email: barbour@zoology.ubc.ca
-######################################################
-
-## 4 species model: 2 consumers, 2 resources
-Pachepsky_model <- function(Time,State,Pars) {
-  # t,y,p is the necessary format for solving using the ode() function in R
-  with(as.list(c(State,Pars)), {
-    
-    # constraints on consumer preference
-    w12 <- (1 - w11)
-    w21 <- (1 - w22)
-      
-    # consumer preference functions
-    W11 <- (w11 * R1)/(w11 * R1 + w12 * R2)
-    W12 <- (w12 * R2)/(w11 * R1 + w12 * R2)
-    W21 <- (w21 * R1)/(w21 * R1 + w22 * R2)
-    W22 <- (w22 * R2)/(w21 * R1 + w22 * R2)
-
-    
-    # consumer functional responses
-    C1R1fxn <- (W11 * a11 * R1)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
-    C1R2fxn <- (W12 * a12 * R2)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
-    C2R1fxn <- (W21 * a21 * R1)/(1 + W21 * a21 * h21 * R1 + W22 * a22 * h22 * R2)
-    C2R2fxn <- (W22 * a22 * R2)/(1 + W21 * a21 * h21 * R1 + W22 * a22 * h22 * R2)
-    
-    # dynamical equations
-    dR1.dt <- r1 * R1 * (1 - R1 / K1) - C1 * C1R1fxn - C2 * C2R1fxn
-    dR2.dt <- r2 * R2 * (1 - R2 / K2) - C1 * C1R2fxn - C2 * C2R2fxn
-    dC1.dt <- -m1 * C1 # C1 * (e11 * C1R1fxn + e12 * C1R2fxn - m1)
-    dC2.dt <- -m2 * C2 # C2 * (e21 * C2R1fxn + e22 * C2R2fxn - m2)
-    
-    return(list(c(dR1.dt, dR2.dt, dC1.dt, dC2.dt)))
-  })
-}
-
-## 3 species model: C1, R1, R2
[…]

### 2019-11-07 morgue/code/00_general_parameters.R

@@ -1,26 +1,0 @@
-
-## load required libraries
-library(deSolve) # numerical integration library
-library(rootSolve) # for runsteady integration function
-library(pse) # for latin hypercube
-library(plyr) # for ldply
-library(dplyr) # for manipulating data
-library(tidyr) # for tidying data
-library(cowplot) # for better base, ggplot graphics
-
-## Abundance threshold for feasible equilibriums
-#abund.thres <- 1e-2 
-
-## Trait change threshold for identifying ESS
-#trait.thres <- 1e-4
-
-## Number of mutation iterations
-#mut.reps <- 10000 
-
-## Mutation size
-#mut.size <- 0.01
-
-## All runsteady simulations go for 1000 time steps with a steady state tolerance of 1e-4. 
-runsteady.ps <- function(y, func = func, parms = parms, times = c(0,1000), stol = 1e-4){
-  runsteady(y = y, times = times, func = func, parms = parms, stol = stol)
-}

### 2019-11-07 morgue/code/00_numerical_fxns.R

@@ -1,387 +1,0 @@
-######################################################
-## Description: Dynamical model and other useful functions for checking dynamics
-## Code author(s): Matt Barbour
-## Email: barbour@zoology.ubc.ca
-######################################################
-
-
-## Consumer-resource dynamics ----
-# general assumptions:
-# (1) logistic growth for the resource(s)
-# (2) multispecies, spatially-implicit type 2 functional response of consumer 
-# (3) density-independent mortality for consumer 
-# function is in the necessary format "function(Time,State,Pars)" for solving used the ode() function in the R package "deSolve"
-
-## 4 species model: 2 consumers, 2 resources
-ECD_model <- function(Time,State,Pars,MacArthur=FALSE) {
-  # t,y,p is the necessary format for solving using the ode() function in R
-  with(as.list(c(State,Pars)), {
-    
-    if(MacArthur==FALSE){
-      # constraints on consumer preference
-      w12 <- (1 - w11)
-      w21 <- (1 - w22)
-      
-      # consumer preference functions
-      W11 <- (w11 * R1)/(w11 * R1 + w12 * R2)
-      W12 <- (w12 * R2)/(w11 * R1 + w12 * R2)
-      W21 <- (w21 * R1)/(w21 * R1 + w22 * R2)
-      W22 <- (w22 * R2)/(w21 * R1 + w22 * R2)
-    } else {
-      W11 <- 1
-      W12 <- 1
-      W21 <- 1
-      W22 <- 1
-    }
-    
-    
-    # consumer functional responses
-    C1R1fxn <- (W11 * a11 * R1)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
-    C1R2fxn <- (W12 * a12 * R2)/(1 + W11 * a11 * h11 * R1 + W12 * a12 * h12 * R2)
[…]

### 2019-11-07 morgue/code/01_get_parameters.R

@@ -1,98 +1,0 @@
-
-#### Description: Use latin hypercube to generate possible parameters that will result in coexistence equilibriums for 4 species.
-
-## load required functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-## Setup simulation ----
-
-## Specificy the number of parameter sets to generate
-n.p <- 5000 # 95 hrs to finish on my computer (almost 4 days)
-
-## Create a function that gets the model parameters and time sequence and outputs the final size of populations
-oneRun <- function(R1_0, R2_0, C1_0, C2_0,
-                   r1, r2, K1, K2,
-                   e11, e12, e21, e22,
-                   h11, h12, h21, h22,
-                   a11, a12, a21, a22,
-                   m1, m2, w11, w22){  
-  # give parameters, initial state variables, and simulation length
-  parameters <- c(r1 = r1, r2 = r2, K1 = K1, K2 = K2,
-                  e11 = e11, e12 = e12, e21 = e21, e22 = e22,
-                  h11 = h11, h12 = h12, h21 = h21, h22 = h22,
-                  a11 = a11, a12 = a12, a21 = a21, a22 = a22,
-                  m1 = m1, m2 = m2, w11 = w11, w22 = w22)
-  state <- c(R1 = R1_0, R2 = R2_0, C1 = C1_0, C2 = C2_0)
-  times <- c(0, 1000) # length of simulation
-
-  # Use ode integration and index to get final state values. Will check later whether 
-  
-  
-  # using failwith() to make function give output even with a fatal error
-  # runsteady() identifies the steady-state of the model.
-  #safe.runsteady <- failwith(default = c(R1 = NA, R2 = NA, C1 = NA, C2 = NA, steady = 0), runsteady)
-  
-  out <- safe.runsteady(y = state, func = ECD_model, parms = parameters)
-  #out <- safe.stode(y = state, func = ECD_model, parms = parameters, positive = TRUE) #times = times, 
-  
-  # return steady states
-  if(class(out) == "list"){
-    return(c(out$y, steady = attr(out, "steady")))
-  }
-  if(class(out) == "numeric"){
-    return(out)
-  }
-}
-
-## Create a function with mapply that gets a matrix of parameter combinations (combinations are lines and parameters are columns) and returns the output of the function above for each parameter combination.
-modelRun <- function (my.pars) {
-  return(mapply(oneRun, 
[…]

### 2019-11-07 morgue/code/02_check_feasibility_stability.R

@@ -1,127 +1,0 @@
-
-## Description: Ensure feasibility and stability of parameters for both 4 species and 3 species models
-
-## source general parameters and functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-## load data ----
-df.4sp.feas <- read.csv("data/hypercube.4sp.csv") %>% 
-  tbl_df() %>% 
-  select(-X) %>%
-  filter(R1 > abund.thres, R2 > abund.thres, C1 > abund.thres, C2 > abund.thres, steady == 1)
-df.4sp.feas # 525 feasible equilibriums where all 4 species coexist
-
-## double-check stability and dynamics of 4 species equilibrium (negative real eigenvalue and presence of imaginary eigenvalue) ----
-stable.4sp <- c()
-spin.4sp <- c()
-for(i in 1:dim(df.4sp.feas)[1]){
-  set <- as.matrix(df.4sp.feas[i,5:28])
-  eq.jac.set <- jacobian.full(y = set[1,c("R1","R2","C1","C2")], func = ECD_model, parms = set[1,1:20])
-  stable.4sp[i] <- max(Re(eigen(eq.jac.set)$values))
-  spin.4sp[i] <- max(Im(eigen(eq.jac.set)$values))
-}
-
-df.4sp.feas$max.Re.eigen <- stable.4sp
-df.4sp.feas$max.Im.eigen <- spin.4sp
-
-length(which(spin.4sp == 0)) # 8 of them appear to have monotonic dynamics.
-length(which(stable.4sp < 0)) # 440 feasible and stable equilibriums with 4 sp. coexisting
-
-df.4sp.feas.stab <- filter(df.4sp.feas, max.Re.eigen < 0)
-
-## determine which parameter sets also have feasible equilibriums if one consumer species is removed ----
-df <- df.4sp.feas.stab
-feas.3sp.list <- list()
-for(i in 1:dim(df)[1]){
-  
-  ## specify the simulation data to use
-  set <- as.matrix(df[i, ]) 
-  
-  ## use initial state values
-  C1.states <- set[1,c("R1_0","R2_0","C1_0")]
-  names(C1.states) <- c("R1","R2","C1")
-  C2.states <- set[1,c("R1_0","R2_0","C2_0")]
[…]

### 2019-11-07 morgue/code/03_sim_3sp_C1_evo.R

@@ -1,131 +1,0 @@
-
-## NEED TO RERUN SIMULATION WITH MUT.SIZE! CHECK OTHER SIMS!
-
-#### Description: 3 species model with C1 evolution simulation
-
-## source required functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-## load data ----
-C1.all.feas.stable.df <- read.csv('data/all.feas.stable.df.csv') %>% 
-  tbl_df() %>%
-  select(r1:K2, e11:e12, h11:h12, a11:a12, m1, w11, C1.R1:C1.C1, C1.max.eigen) %>%
-  rename(R1 = C1.R1, R2 = C1.R2, C1 = C1.C1)
-
-## setup and run simulation ----
-df <- C1.all.feas.stable.df
-sim.list.C1 <- list()
-for(i in 1:dim(df)[1]){
-  
-  ## choose parameter set to examine and calculate the maximum real eigenvalue
-  set <- as.matrix(df[i, ])
-  
-  ## setup and run sub-simulations
-  sub.sim.list.3 <- list()
-  
-  # input first data vector
-  sub.sim.list.3[[1]] <- c(set[1, ], feas.3sp = 1, steady.3sp = 1, 
-                           species = NA, mut.suc = NA, p.try = NA,
-                           sim.number = i, sequence = 1) 
- 
-  # run simulation
-  for(j in 2:mut.reps){
-    
-    test <- sub.sim.list.3 %>% ldply()
-    
-    # if the prior simulation results in the exclusion of a species, stop the simulation
-    if(sub.sim.list.3[[j-1]]["feas.3sp"] == 0) break 
-    
-    # if the prior simulation results in an unstable system, stop the simulation
-    if(sub.sim.list.3[[j-1]]["C1.max.eigen"] > 0) break  
-    
-    # if we are no longer observing substantial trait change after 100 mutations step, stop the simulation
-    if(j < 100){
-      delta.a11 <- 1 # arbitrary number greater than threshold value
-    }
-    if(j > 100){
-      delta.a11 <- mean(diff(test[(j-100):(j-1),"a11"]))
-    } 
-    if(delta.a11 < trait.thres) break 
-    
-    # if the prior simulation is at a feasible and steady state, proceed with simulation
-    if(sub.sim.list.3[[j-1]]["feas.3sp"] == 1 & 
-       sub.sim.list.3[[j-1]]["C1.max.eigen"] < 0){
[…]

### 2019-11-07 morgue/code/03_sim_3sp_C2_evo.R

@@ -1,130 +1,0 @@
-### NEED TO RERUN WITH MUT.SIZE!!! 
-
-#### Description: 3 species model with C2 evolution simulation
-
-## source required functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-## load data ----
-C2.all.feas.stable.df <- read.csv('data/all.feas.stable.df.csv') %>% 
-  tbl_df() %>%
-  select(r1:K2, e22:e21, h22:h21, a22:a21, m2, w22, C2.R1:C2.C2, C2.max.eigen) %>%
-  rename(R1 = C2.R1, R2 = C2.R2, C2 = C2.C2)
-
-## setup and run simulation ----
-df <- C2.all.feas.stable.df
-sim.list.C2 <- list()
-for(i in 1:dim(df)[1]){
-  
-  ## choose parameter set to examine and calculate the maximum real eigenvalue
-  set <- as.matrix(df[i, ])
-  
-  ## setup and run sub-simulations
-  sub.sim.list.3 <- list()
-  
-  # input first data vector
-  sub.sim.list.3[[1]] <- c(set[1, ], feas.3sp = 1, steady.3sp = 1, 
-                           species = NA, mut.suc = NA, p.try = NA,
-                           sim.number = i, sequence = 1) # input first data vector
-  
-  for(j in 2:mut.reps){
-    
-    test <- sub.sim.list.3 %>% ldply()
-    
-    # if the prior simulation results in the exclusion of a species, stop the simulation
-    if(sub.sim.list.3[[j-1]]["feas.3sp"] == 0) break 
-    
-    # if the prior simulation results in an unstable system, stop the simulation
-    if(sub.sim.list.3[[j-1]]["C2.max.eigen"] > 0) break  
-    
-    # if we are no longer observing substantial trait change after 100 mutations step, stop the simulation
-    if(j < 100){
-      delta.a22 <- 1 # arbitrary number greater than threshold value
-    }
-    if(j > 100){
-      delta.a22 <- mean(diff(test[(j-100):(j-1),"a22"]))
-    } 
-    if(delta.a22 < trait.thres) break 
-    
-    # if the prior simulation is at a feasible and steady state, proceed with simulation
-    if(sub.sim.list.3[[j-1]]["feas.3sp"] == 1 & 
-       sub.sim.list.3[[j-1]]["C2.max.eigen"] < 0){
-      
-      # extract parameters and state variables
[…]

### 2019-11-07 morgue/code/03_sim_4sp_C1evo.R

@@ -1,134 +1,0 @@
-#### Description: 4 species model with only one consumer (C1) evolving simulation
-
-## source general parameters and functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-## load data ----
-all.feas.stable.df <- read.csv('data/all.feas.stable.df.csv') %>% 
-  tbl_df() %>%
-  select(r1:C2, max.Re.eigen, max.Im.eigen)
-
-
-## setup and run simulation ----
-df <- all.feas.stable.df
-sim.list.4 <- list()
-for(i in 1:dim(df)[1]){
-  
-  # choose parameter set to examine and calculate the maximum real eigenvalue
-  set <- as.matrix(df[i, ])
-  
-  ## setup and run sub-simulations
-  sub.sim.list.4 <- list()
-  
-  # input first data vector
-  sub.sim.list.4[[1]] <- c(set[1, ], feas.4sp = 1, steady.4sp = 1, 
-                           species = NA, mut.suc = NA, p.try = NA, 
-                           sim.number = i, sequence = 1) 
-  
-  # simulation
-  for(j in 2:mut.reps){
-    
-    test <- sub.sim.list.4 %>% ldply()
-    
-    # if the prior simulation results in the exclusion of a species, stop the simulation
-    if(sub.sim.list.4[[j-1]]["feas.4sp"] == 0) break 
-    
-    # if the prior simulation results in an unstable system, stop the simulation
-    if(sub.sim.list.4[[j-1]]["max.Re.eigen"] > 0) break  
-    
-    # if we are no longer observing substantial trait change after 100 mutations step, stop the simulation
-    if(j < 100){
-      delta.a11 <- 1 # arbitrary number greater than threshold value
-      delta.a22 <- 1 # arbitrary number greater than threshold value
-    }
-    if(j > 100){
-      delta.a11 <- mean(diff(test[(j-100):(j-1),"a11"]))
-      delta.a22 <- mean(diff(test[(j-100):(j-1),"a22"]))
-    } 
-    if(delta.a11 < trait.thres & delta.a22 < trait.thres) break 
-    
-    # if the prior simulation is at a feasible and steady state, proceed with simulation
-    if(sub.sim.list.4[[j-1]]["feas.4sp"] == 1 & 
-       sub.sim.list.4[[j-1]]["max.Re.eigen"] < 0){
-      
-      # extract parameters and state variables
[…]

### 2019-11-07 morgue/code/03_sim_4sp_coevo.R

@@ -1,146 +1,0 @@
-
-#### Description: 4 species model with consumer coevolution simulation
-
-## source general parameters and functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-## load data ----
-all.feas.stable.df <- read.csv('data/all.feas.stable.df.csv') %>% 
-  tbl_df() %>%
-  select(r1:C2, max.Re.eigen, max.Im.eigen)
-  
-
-## setup and run simulation ----
-df <- all.feas.stable.df
-sim.list.4 <- list()
-for(i in 1:dim(df)[1]){
-  
-  # choose parameter set to examine and calculate the maximum real eigenvalue
-  set <- as.matrix(df[i, ])
-  
-  ## setup and run sub-simulations
-  sub.sim.list.4 <- list()
-  
-  # input first data vector
-  sub.sim.list.4[[1]] <- c(set[1, ], feas.4sp = 1, steady.4sp = 1, 
-                           species = NA, mut.suc = NA, p.try = NA, 
-                           sim.number = i, sequence = 1) 
-  
-  # simulation
-  for(j in 2:mut.reps){
-    
-    test <- sub.sim.list.4 %>% ldply()
-    
-    # if the prior simulation results in the exclusion of a species, stop the simulation
-    if(sub.sim.list.4[[j-1]]["feas.4sp"] == 0) break 
-      
-    # if the prior simulation results in an unstable system, stop the simulation
-    if(sub.sim.list.4[[j-1]]["max.Re.eigen"] > 0) break  
-  
-    # if we are no longer observing substantial trait change after 100 mutations step, stop the simulation
-    if(j < 100){
-      delta.a11 <- 1 # arbitrary number greater than threshold value
-      delta.a22 <- 1 # arbitrary number greater than threshold value
-    }
-    if(j > 100){
-     delta.a11 <- mean(diff(test[(j-100):(j-1),"a11"]))
-     delta.a22 <- mean(diff(test[(j-100):(j-1),"a22"]))
-    } 
-    if(delta.a11 < trait.thres & delta.a22 < trait.thres) break 
-    
-    # if the prior simulation is at a feasible and steady state, proceed with simulation
-    if(sub.sim.list.4[[j-1]]["feas.4sp"] == 1 & 
-      sub.sim.list.4[[j-1]]["max.Re.eigen"] < 0){
-        
-      # extract parameters and state variables
[…]

### 2019-11-07 morgue/code/04_sim_analyses.R

@@ -1,211 +1,0 @@
-
-## Description: Analyze effects of evolution and community context on character displacement and ecological stability
-
-## source general parameters and functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-## load and manipulate required data ----
-evol.sim.df.4 <- read.csv('data/evol.sim.df.4sp.csv') %>% 
-  tbl_df() %>% select(-X) %>% 
-  mutate(special.C1 = abs((a11/(a11 + a12)) - 0.5),
-         special.C2 = abs((a22/(a21 + a22)) - 0.5),
-         effective.C1 = a11*w11 + a12*(1-w11), # effective attack rate of consumer 1
-         effective.C2 = a22*w22 + a21*(1-w22)) # effective attack rate of consumer 2
-
-#evol.sim.df.4.1evo <- read.csv('data/evol.sim.df.4sp.1evo.csv') %>% 
-#  tbl_df() %>% select(-X) %>% 
-#  mutate(special.C1 = abs((a11/(a11 + a12)) - 0.5),
-#         effective.C1 = a11*w11 + a12*(1-w11)) # effective attack rate of consumer 1
-
-evol.sim.df.C1 <- read.csv('data/evol.sim.df.C1.3sp.csv') %>% 
-  tbl_df() %>% select(-X) %>% 
-  mutate(special.C1 = abs((a11/(a11 + a12)) - 0.5),
-         effective.C1 = a11*w11 + a12*(1-w11)) # effective attack rate of consumer 1
-
-evol.sim.df.C2 <- read.csv('data/evol.sim.df.C2.3sp.csv') %>% 
-  tbl_df() %>% select(-X) %>% 
-  mutate(special.C2 = abs((a22/(a21 + a22)) - 0.5),
-         effective.C2 = a22*w22 + a21*(1-w22)) # effective attack rate of consumer 2
-
-## Summarize simulation results ----
-
-# 4 species - coevolution
-summary.4 <- list()
-for(i in 1:max(evol.sim.df.4$sim.number)){
-  temp.df <- filter(evol.sim.df.4, sim.number == i)
-  end.seq <- max(temp.df$sequence)
-  
-  summary.4[[i]] <- data.frame(
-    sim.number = i,
-    end.seq.4 = end.seq,
-    R1.end.4 = as.numeric(temp.df[end.seq,"R1"]),
-    R2.end.4 = as.numeric(temp.df[end.seq,"R2"]),
[…]

### 2019-11-07 morgue/code/MacArthur.sym.evo.4sp.R

@@ -1,657 +1,0 @@
-
-#### Description: 4 species model with consumer coevolution simulation,
-####              but this is only for one parameter set where resource
-####              parameters are equivalent and consumers are mirror
-####              images of each other.
-
-## source general parameters and functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-source('code/00_MacArthur_fxns.R')
-
-trait.thres <- 1e-10
-j.step <- 1000 # 100 for other simulations
-
-## general parameters for simulations
-r <- 1
-K <- 4
-e <- 0.8
-#h <- 0.4
-aii <- 1.4 #1.93
-aij <- 1.2
-#a <- 2
-#w <- 0.6
-m <- 1
-R <- 2
-C <- 1
-
-
-## 4 species simulation ----
-
-## create initial data
-sym4.init.df <- data.frame(r1 = r, r2 = r, K1 = K, K2 = K,
-                           e11 = e, e12 = e, e21 = e, e22 = e,
-                           #h11 = h, h12 = h, h21 = h, h22 = h,
-                           a11 = aii, a12 = aij, a21 = aij, a22 = aii,
-                           #w11 = w, w22 = w, 
-                           m1 = m, m2 = m,
-                           R1 = R, R2 = R, C1 = C, C2 = C)
-
-set.temp <- as.matrix(sym4.init.df[1, ])
-
-set.steady <- safe.runsteady(set.temp[1,c("R1","R2","C1","C2")], func = ECD_model_Mac, parms = set.temp[1,c("r1","r2","K1","K2","e11","e12","e21","e22","a11","a12","a21","a22","m1","m2")])
-
-sym4.eq.jac <- jacobian.full(y = set.steady$y, func = ECD_model_Mac, parms = set.temp[1,c("r1","r2","K1","K2","e11","e12","e21","e22","a11","a12","a21","a22","m1","m2")])
-
-## update with new steady state and eigenvalues after the initial run
-sym4.init.df[ ,c("R1","R2","C1","C2")] <- set.steady$y
-sym4.init.df$max.Re.eigen <- max(Re(eigen(sym4.eq.jac)$values))
-sym4.init.df$max.Im.eigen <- max(Im(eigen(sym4.eq.jac)$values))
-
-
-## setup and run simulation
-df <- sym4.init.df
-sim.list.4 <- list()
-for(i in 1:dim(df)[1]){
-  
[…]

### 2019-11-07 morgue/code/MacArthur.sym.evo.4sp.TradeOffs.R

@@ -1,665 +1,0 @@
-
-#### Description: 4 species model with consumer coevolution simulation,
-####              but this is only for one parameter set where resource
-####              parameters are equivalent and consumers are mirror
-####              images of each other.
-
-## source general parameters and functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-source('code/00_MacArthur_fxns.R')
-source('attack_rate_trade_off_function.R')
-
-trait.thres <- 1e-10
-j.step <- 1000 # 100 for other simulations
-
-## Trade-Off ----
-A_invest <- 2.5
-a11_seq <- seq(0,A_invest,0.01)
-
-linear <- 1
-concave <- 0.95
-convex <- 1.05
-
-a_trade_offs <- data.frame(a11 = a11_seq, a12 = a_ij_tradeoff(A=A_invest, a_ii=a11_seq, n=linear),  n_shape = linear) %>%
-  bind_rows(., data.frame(a11 = a11_seq, a12 = a_ij_tradeoff(A=A_invest, a_ii=a11_seq, n=concave), n_shape = concave)) %>%
-  bind_rows(., data.frame(a11 = a11_seq, a12 = a_ij_tradeoff(A=A_invest, a_ii=a11_seq, n=convex), n_shape = convex)) %>%
-  mutate(aT = a11 + a12)
-
-# visualize the trade-off between a11 and a12
-ggplot(a_trade_offs, aes(x=a11, y=a12, color=factor(n_shape))) +
-  geom_line()
-
-ggplot(a_trade_offs, aes(x=a11, y=aT, color=factor(n_shape))) +
-  geom_line()
-
-## general parameters for simulations
-n_shape <- convex
-r <- 1
-K <- 4
-e <- 0.8
-aii <- A_invest*0.6#0.51
-aij <- a_ij_tradeoff(A = A_invest, a_ii = aii, n = n_shape)
-m <- 1
-R <- 2
-C <- 1
-
-
-## 4 species simulation ----
-
-## create initial data
-sym4.init.df <- data.frame(r1 = r, r2 = r, K1 = K, K2 = K,
-                           e11 = e, e12 = e, e21 = e, e22 = e,
-                           #h11 = h, h12 = h, h21 = h, h22 = h,
-                           a11 = aii, a12 = aij, a21 = aij, a22 = aii,
-                           #w11 = w, w22 = w, 
-                           m1 = m, m2 = m,
[…]

### 2019-11-07 morgue/code/bifurcation_plot_symmetry.R

@@ -1,167 +1,0 @@
-## load required libraries
-library(tidyverse)
-library(cowplot)
-
-## load required data ----
-C1.ESS.df <- read.csv('data/evol.symmetry.df.C1.3sp.csv')
-
-## source required functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-#### Setup the initial state variables and global parameters for all of the models ----
-
-## set state variables for all models
-R1 <- 2
-R2 <- 2
-C1 <- 1
-C2 <- 1
-i.state.2C_2R <- c(R1 = R1, R2 = R2, C1 = C1, C2 = C2) 
-
-## set duration of simulations
-Time <- 1000
-
-## set up general code for bifurcation plot. Maintaining 'w' as the key parameter to manipulate
-aii <- C1.ESS.df$a11[1]
-aij <- C1.ESS.df$a12[1]
-
-aii_seq <- seq(aii, aii*2, by = 0.01)[1:75]
-aij_seq <- seq(aij, 0, by = -0.01)[1:75]
-
-w11 <- w22 <- C1.ESS.df$w11[1]
-#a_alt <- a_start
-aECD <- w11*aii_seq + (1-w11)*aij_seq
-
-## Bifurcation plot code
-# this code loops over the sequence of different parameter values
-out <- list()
-for (i in 1:length(aii_seq)) {
-  # general parameters. Taken from Figure 3 in McCann et al. 2005. 
-  r1 <- r2 <- 1
-  K1 <- K2 <- 3.25
-  e11 <- e12 <- e21 <- e22 <- 0.8
-  h11 <- h12 <- h21 <- h22 <- 0.4
-  m1 <- m2 <- 1
-  
-  # parameters related to aECD
-  #w <- 0.8
-  #a_prime <- param.seq.a_prime[i]
-  #a_alt <- 2
-  
-  # set params 
-  param.loop <- c(r1 = r1, r2 = r2, K1 = K1, K2 = K2,
-                  e11 = e11, e12 = e12, e21 = e21, e22 = e22,
-                  h11 = h11, h12 = h12, h21 = h21, h22 = h22,
-                  a11 = aii_seq[i], a12 = aij_seq[i], a21 = aij_seq[i], a22 = aii_seq[i],
-                  m1 = m1, m2 = m2, w11 = w11, w22 = w22)
-  
-  # Run the experiment.
-  init <- ode(i.state.2C_2R, 
-              1:Time, ECD_model, param.loop)
[…]

### 2019-11-07 morgue/code/find_ESS_3sp.R

@@ -1,139 +1,0 @@
-
-#### Description: Find ESS for C1 3 sp model when resources have equivalent parameters.
-
-## source required functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-## create initial data ----
-C1.init.df <- data.frame(r1 = 1, r2 = 1, K1 = 3.25, K2 = 3.25,
-                         e11 = 0.8, e12 = 0.8, h11 = 0.4, h12 = 0.4,
-                         a11 = 2, a12 = 2, m1 = 1, w11 = 0.8, 
-                         R1 = 2, R2 = 2, C1 = 1)
-
-set.temp <- as.matrix(C1.init.df[1, ])
-
-set.steady <- safe.runsteady(set.temp[1,c("R1","R2","C1")], func = ECD_model.C1.3sp, parms = set.temp[1,1:12])
-
-C1.eq.jac <- jacobian.full(y = set.steady$y, 
-                           func = ECD_model.C1.3sp, parms = set.temp[1,1:12])
-
-C1.init.df[ ,c("R1","R2","C1")] <- set.steady$y
-C1.init.df$C1.max.eigen <- max(Re(eigen(C1.eq.jac)$values))
-
-## setup and run simulation ----
-df <- C1.init.df
-sim.list.C1 <- list()
-for(i in 1:dim(df)[1]){
-  
-  ## choose parameter set to examine and calculate the maximum real eigenvalue
-  set <- as.matrix(df[i, ])
-  
-  ## setup and run sub-simulations
-  sub.sim.list.3 <- list()
-  
-  # input first data vector
-  sub.sim.list.3[[1]] <- c(set[1, ], feas.3sp = 1, steady.3sp = 1, 
-                           species = NA, mut.suc = NA, p.try = NA,
-                           sim.number = i, sequence = 1) 
-  
-  # run simulation
-  for(j in 2:mut.reps){
-    
-    test <- sub.sim.list.3 %>% ldply()
-    
-    # if the prior simulation results in the exclusion of a species, stop the simulation
-    if(sub.sim.list.3[[j-1]]["feas.3sp"] == 0) break 
-    
-    # if the prior simulation results in an unstable system, stop the simulation
-    if(sub.sim.list.3[[j-1]]["C1.max.eigen"] > 0) break  
-    
-    # if we are no longer observing substantial trait change after 100 mutations step, stop the simulation
-    if(j < 100){
[…]

### 2019-11-07 morgue/code/global_stability_figure.R

@@ -1,134 +1,0 @@
-## load required data ----
-C1.ESS.df <- read.csv('data/evol.symmetry.df.C1.3sp.csv')
-sym.4sp.df <- read.csv('data/evol.symmetry.df.4sp.csv')
-
-## load required functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-#### Setup the initial state variables and global parameters for all of the models ----
-
-## set state variables for all models
-# These are different from the equilbrium values to illustrate the dynamic trajectories back to equilibrium.
-R1 <- 2
-R2 <- 2
-C1 <- 1
-C2 <- 1
-i.state.2C_2R <- c(R1 = R1, R2 = R2, C1 = C1, C2 = C2) 
-
-## set duration of simulations
-Time <- 1000
-
-## set up general code for the plot. Maintaining 'w' as the key parameter to manipulate
-#aii <- C1.ESS.df$a11[1]
-#aij <- C1.ESS.df$a12[1]
-
-#aii_seq <- seq(aii, aii*2, by = 0.01)[1:75]
-#aij_seq <- seq(aij, 0, by = -0.01)[1:75]
-
-#w11 <- w22 <- C1.ESS.df$w11[1]
-#a_alt <- a_start
-#aECD <- w11*aii_seq + (1-w11)*aij_seq
-
-# general parameters. Taken from Figure 3 in McCann et al. 2005. 
-#r1 <- r2 <- 1
-#K1 <- K2 <- 3.25
-#e11 <- e12 <- e21 <- e22 <- 0.8
-#h11 <- h12 <- h21 <- h22 <- 0.4
-#m1 <- m2 <- 1
-#w11 <- w22 <- C1.ESS.df$w11[1]
-
-#aii <- C1.ESS.df$a11[1]
-#aij <- C1.ESS.df$a12[1]
-
-#aii_seq <- seq(aii, aii*2, by = 0.01)[1:75]
-#aij_seq <- seq(aij, 0, by = -0.01)[1:75]
-
-p.set <- 1
-
-p1 <- sym.4sp.df[p.set, 2:21]
-
-#param.1 <- c(r1 = r1, r2 = r2, K1 = K1, K2 = K2,
-#             e11 = e11, e12 = e12, e21 = e21, e22 = e22,
-#             h11 = h11, h12 = h12, h21 = h21, h22 = h22,
[…]

### 2019-11-07 morgue/code/morgue.R

@@ -1,923 +1,0 @@
-## Chracter displacement figure
-
-seq.num <- 500#dim(evo.3)[1]
-
-plot(0:1, 0:1, type = "n", 
-     xlab = "Attack Rate \nSpecialization on R1", 
-     ylab = "Attack Rate \nSpecialization on R2")
-arrows(evo.3$special.C1R1[evo.3$sequence], 
-       evo.3$special.C1R2[evo.3$sequence],
-       evo.3$special.C1R1[evo.3$sequence+1], 
-       evo.3$special.C1R2[evo.3$sequence+1], 
-       col = gray.colors(seq.num, start = 0.9, end = 0.1), length = 0.1, lwd = 5)
-points(x = evo.3$special.C1R1[1], 
-       y = evo.3$special.C1R2[1], 
-       col = "black", bg = "white", pch = 21, cex = 5)
-points(x = evo.3$special.C1R1[seq.num], 
-       y = evo.3$special.C1R2[seq.num], 
-       col = "black", bg = "white", pch = 21, cex = 5)
-text(x = evo.3$special.C1R1[1], 
-     y = evo.3$special.C1R2[1], 
-     labels = "C1")
-text(x = evo.3$special.C1R1[seq.num], 
-     y = evo.3$special.C1R2[seq.num],
-     labels = "C1*")
-
-
-## Create a plot of how mutations proceed over time 
-# Using shaded arrows to draw the eye into the direction of mutations offer time.
-# consider turning into ggplot
-seq.num <- dim(evo.4)[1]
-
-plot(0:1, 0:1, type = "n", 
-     xlab = "Attack Rate \nSpecialization on R1", 
-     ylab = "Attack Rate \nSpecialization on R2")
-arrows(evo.4$special.C1R1[evo.4$sequence], 
-       evo.4$special.C1R2[evo.4$sequence],
-       evo.4$special.C1R1[evo.4$sequence+1], 
-       evo.4$special.C1R2[evo.4$sequence+1], 
-       col = gray.colors(seq.num, start = 0.9, end = 0.1), length = 0.1, lwd = 5)
-arrows(evo.4$special.C2R1[evo.4$sequence], 
-       evo.4$special.C2R2[evo.4$sequence],
-       evo.4$special.C2R1[evo.4$sequence+1], 
-       evo.4$special.C2R2[evo.4$sequence+1], 
-       col = gray.colors(seq.num, start = 0.9, end = 0.1), length = 0.1, lwd = 5)
-points(x = evo.4$special.C1R1[1], 
[…]

### 2019-11-07 morgue/code/parameterized_Pachepsky_within-year.R

@@ -1,205 +1,0 @@
-
-##############################################################################################################
-## Estimate stickleback consumer-resource dynamics based on allometric scaling and literature.
-## Author: Matt Barbour
-## Email: matthew.a.barbour@gmail.com
-##############################################################################################################
-
-
-## LOAD REQUIRE FUNCTIONS ----
-
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-source('code/00_Pachepsky_fxns.R')
-
-
-mass_scaling <- function(W, intercept, exponent) {
-  log_rate <- intercept + exponent * log(W)
-  return(exp(log_rate))
-}
-
-
-## BODY SIZE (g) ----
-
-g_zoop <- 0.004/1000            # covert 0.004 mg to g
-g_benthic_invert <- 0.25/1000   # convert 0.25 mg to g
-
-g_limnetic <- 0.81              # from Schluter 1993, Ecology Table 1
-g_benthic <- 1.48               # from Schluter 1993, Ecology Table 1
-g_solitary <- 1.12              # from Schluter 1993, Ecology Table 1
-
-
-## INTRINSIC GROWTH RATE (r) ----
-# r = birth rate per day
-# proportional to M^(-1/4) from Rip and McCann 2011 (they cite Brown et al. 2004)
-# alternatively use scaling exponent = -0.27377 for heterotherm metazoa from Fenchel 1974, Oecologia
-r_exponent <- -0.25             # theoretical scaling.
-r_scale <- 0.3                  # estimate of r from Daphnia pulex in Table 1 of Fenchel 1974. Original study is Frank et al. 1957       
-r_intercept <- log(r_scale) - r_exponent * log(g_zoop) # scale based on zooplankton size
-
-r_benthos_estimate <- mass_scaling(W = g_benthic_invert, intercept = r_intercept, exponent = r_exponent) 
-
-
-#### CARRYING CAPACITY (K) ----
-# proportional to M^(-3/4) from Rip and McCann 2011 (they cite Enquist et al. 1998 and Brown et al. 2004)
-# alternatively use empirical exponent = -0.77 from Brown et al. 2004
-K_exponent <- -0.75 
-K_scale <- 5          # admittedly, this is a bit of an arbitrary number. It does correspond to zooplankton biomass density (g/500 L) from fishless ponds
-K_intercept <- log(K_scale) - K_exponent * log(g_zoop) # scale based on zooplankton size
-
[…]

### 2019-11-07 morgue/code/sym.evo.4sp.R

@@ -1,636 +1,0 @@
-
-#### Description: 4 species model with consumer coevolution simulation,
-####              but this is only for one parameter set where resource
-####              parameters are equivalent and consumers are mirror
-####              images of each other.
-
-## source general parameters and functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-trait.thres <- 1e-10
-j.step <- 1000 # 100 for other simulations
-
-## general parameters for simulations
-r <- 1
-K <- 4
-e <- 0.8
-h <- 0.4
-aii <- 1.93#2
-aij <- 1.2
-#a <- 2
-w <- 0.6
-m <- 1
-R <- 2
-C <- 1
-
-
-## 4 species simulation ----
-
-## create initial data
-sym4.init.df <- data.frame(r1 = r, r2 = r, K1 = K, K2 = K,
-                         e11 = e, e12 = e, e21 = e, e22 = e,
-                         h11 = h, h12 = h, h21 = h, h22 = h,
-                         a11 = aii, a12 = aij, a21 = aij, a22 = aii,
-                         w11 = w, w22 = w, m1 = m, m2 = m,
-                         R1 = R, R2 = R, C1 = C, C2 = C)
-
-set.temp <- as.matrix(sym4.init.df[1, ])
-
-set.steady <- safe.runsteady(set.temp[1,c("R1","R2","C1","C2")], func = ECD_model, parms = set.temp[1,1:20])
-
-sym4.eq.jac <- jacobian.full(y = set.steady$y, 
-                           func = ECD_model, parms = set.temp[1,1:20])
-
-## update with new steady state and eigenvalues after the initial run
-sym4.init.df[ ,c("R1","R2","C1","C2")] <- set.steady$y
-sym4.init.df$max.Re.eigen <- max(Re(eigen(sym4.eq.jac)$values))
-sym4.init.df$max.Im.eigen <- max(Im(eigen(sym4.eq.jac)$values))
-
-
-## setup and run simulation
-df <- sym4.init.df
-sim.list.4 <- list()
-for(i in 1:dim(df)[1]){
-  
-  # choose parameter set to examine and calculate the maximum real eigenvalue
-  set <- as.matrix(df[i, ])
-  
-  ## setup and run sub-simulations
-  sub.sim.list.4 <- list()
-  
-  # input first data vector
[…]

### 2019-11-07 morgue/eco_evo_simulations.R

@@ -1,603 +1,0 @@
-
-#### PRELIMINARIES ####
-
-## Source in required functions
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-source('eco_evo_sim_function.R')
-source('Consumer_Resource_Functions.R')
-source('CR_dynamics.R')
-
-#### GENERAL PARAMETERS AND STATE VARIABLES #### 
-
-## Intrinsic growth rates
-r <- 1
-r1 <- r 
-r2 <- r 
-
-## Carrying capacity
-K <- 4
-K1 <- K
-K2 <- K
-
-## Conversion efficiency
-e <- 0.8
-e11 <- e 
-e12 <- e 
-e21 <- e 
-e22 <- e 
-
-## Mortality rate
-m <- 1
-m1 <- m
-m2 <- m
-
-## General, non-evolving parameter set for 2C,2R model
-general_parameters <- data.frame(r1 = r1, r2 = r2, K1 = K1, K2 = K2, e11 = e11, e12 = e12, e21 = e21, e22 = e22, m1 = m1, m2 = m2)
-
-## Attack rate
-A <- 2
-aii <- A*0.6
-
-## Habitat preference (for LawlorSmith and McCann models)
-w <- 0.6
-wii <- w
-wij <- (1 - w)
-
-## Handling time (only for McCann model)
-h <- 0.4
-h11 <- h
-h12 <- h
-h21 <- h
-h22 <- h
-
-## State variables
-R <- 2
-R1 <- R
-R2 <- R
-C <- 1
-C1 <- C
-C2 <- C
-
-states_2C_2R <- data.frame(R1 = R1, R2 = R2, C1 = C1, C2 = C2)
-states_1C_2R <- data.frame(R1 = R1, R2 = R2, C1 = C1)
-
-## Trade-offs. Borrowed from Sargent and Otto, 2006 Am Nat
-a12_tradeoff <- expression(A*(1-(a11m/A)^n)^(1/n)) 
-a21_tradeoff <- expression(A*(1-(a22m/A)^n)^(1/n)) 
-
-n_concaveDN <- 1.15
-n_linear <- 1
-n_concaveUP <- 0.85
-
-concaveDN_df <- data.frame(n = n_concaveDN, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_concaveDN)^(1/n_concaveDN))
[…]

### 2019-11-07 morgue/hypercube_sim.R

@@ -1,123 +1,0 @@
-######################################################
-## Description: Identify feasible and stable values for parameters and state variables
-## Code author(s): Matt Barbour
-## Email: barbour@zoology.ubc.ca
-######################################################
-
-## load required libraries ----
-library(deSolve)
-library(rootSolve)
-library(plyr)
-library(dplyr)
-library(assertr)
-library(pse) # for latin hypercube
-
-## load required functions ----
-source('Google Drive/ms_ecd/numerical_fxns.R')
-source('Google Drive/ms_ecd/local_stability.R')
-
-## setup latin hypercube sampling for parameter sets ----
-n.p <- 30 # number of parameter sets
-factors <- c("R1_0","R2_0","C1_0","C2_0",
-             "r1","r2","K1","K2",
-             "e11","e12","e21","e22",
-             "h11","h12","h21","h22",
-             "a11","a12","a21","a22",
-             "m1","m2","w11","w22")
-q <- replicate(24, "qunif")
-q.arg <- list(list(min= 0.5, max = 2),list(min = 0.5, max = 2),list(min = 0.5, max = 2),list(min = 0.5, max = 2),
-           list(min = 0.5, max = 2),list(min = 0.5, max = 2),list(min = 2, max = 5),list(min = 2, max = 5),
-           list(min = 0.5, max = 0.9),list(min = 0.5, max = 0.9),list(min = 0.5, max = 0.9),list(min = 0.5, max = 0.9),
-           list(min = 0.1, max = 0.6),list(min = 0.1, max = 0.6),list(min = 0.1, max = 0.6),list(min = 0.1, max = 0.6),
-           list(min = 0.1, max = 4),list(min = 0.1, max = 4),list(min = 0.1, max = 4),list(min = 0.1, max = 4),
-           list(min = 0.1, max = 1),list(min = 0.1, max = 1),list(min = 0.05, max = 0.95),list(min = 0.05, max = 0.95))
-
-LHS.4 <- LHS(model=NULL, factors=factors, N=n.p, q=q, q.arg=q.arg, nboot=0) # use nboot=0 if do not need partial correlations
-
-LHS.4.ps <- get.data(LHS.4) 
-
[…]

### 2019-11-07 morgue/local_stability.R

@@ -1,117 +1,0 @@
-######################################################
-## Description: 
-## Code author(s): Matt Barbour
-## Email: barbour@zoology.ubc.ca
-######################################################
-
-## load required libraries ----
-library(deSolve)
-library(plyr)
-library(dplyr)
-
-## expressions for consumer-resource dynamics WITH MUTANTS ----
-    
-# dynamical equations
-dR1.dt <- expression(r1 * R1 * (1 - R1 / K1) - C1 * ((((w11 * R1)/(w11 * R1 + (1 - w11) * R2)) * a11 * R1)/(1 + ((w11 * R1)/(w11 * R1 + (1 - w11) * R2)) * a11 * h11 * R1 + (((1 - w11) * R2)/(w11 * R1 + (1 - w11) * R2)) * a12 * h12 * R2)) - C2 * (((((1 - w22) * R1)/((1 - w22) * R1 + w22 * R2)) * a21 * R1)/(1 + (((1 - w22) * R1)/((1 - w22) * R1 + w22 * R2)) * a21 * h21 * R1 + ((w22 * R2)/((1 - w22) * R1 + w22 * R2)) * a22 * h22 * R2)) - mC1 * ((((w11 * R1)/(w11 * R1 + (1 - w11) * R2)) * a11m * R1)/(1 + ((w11 * R1)/(w11 * R1 + (1 - w11) * R2)) * a11m * h11 * R1 + (((1 - w11) * R2)/(w11 * R1 + (1 - w11) * R2)) * a12m * h12 * R2)) - mC2 * (((((1 - w22) * R1)/((1 - w22) * R1 + w22 * R2)) * a21m * R1)/(1 + (((1 - w22) * R1)/((1 - w22) * R1 + w22 * R2)) * a21m * h21 * R1 + ((w22 * R2)/((1 - w22) * R1 + w22 * R2)) * a22m * h22 * R2)))
-
[…]

### 2019-11-07 morgue/maybe_useful_morgue.R

@@ -1,198 +1,0 @@
-```{r}
-# Numerous mathematical models have been developed to examine how resource competition influences the evolution of consumer foraging traits (reviewed in Taper and Case 1985). The general result emerging from these models is that competition for shared (substitutable) resources results in the evolution of consumers that specialize on different resources relative to the generalists that evolve in the absence of competition. The only scenarios in which character divergence will not occur is when density dependence in intra- or interspecific interactions due to other factors are sufficiently strong (Abrams 1986), although there remains little empirical evidence of this to date (Schluter 2000). Despite the many models that have been developed to examine the ecological scenarios resulting in character displacement, scant attention has been paid to the ecological consequences of such character divergence on resource abundances and food-web stability (but see Lawlor and Smith 1976 for an assessment of stability). 
-
-# In terms of resources, perhaps the intuitive expectation is that character divergence would suppress resource abundances. The rationale being that character divergence will result in specialized consumers that are more effective at attacking a particular resource. Below, we demonstrate that this is not necessarily true. In fact, the effect of character displacement on resource abundances depends on both the foraging scenario and tradeoff in consumer feeding rates.
-```
-
-We note, however, that our analytical insights to the effects of consumer divergence on resource dynamics has considered a simplistic scenario (resources are equivalent and consumers have equivalent conversion efficiences, but have mirror image attack rates on each resource). To explore the generality of our predictions, we examined whether our conclusions still held under a foraging scenario that was designed to mimic the stickleback lake food web. To do this, we used empirical relationships between body size and the parameters that govern consumer-resource dynamics to parameterize an eco-evolutionary model. Specifically, we used consumer and resource biomasses 
-
-```{r Allometric Scaling of Stickleback Food Web}
-
-mass_scaling <- function(W, intercept, exponent) {
-  log_rate <- intercept + exponent * log(W)
-  return(exp(log_rate))
-}
-
-
[…]

### 2019-11-07 morgue/numerical_sims.R

@@ -1,125 +1,0 @@
-######################################################
-## Description: 
-## Code author(s): Matt Barbour
-## Email: barbour@zoology.ubc.ca
-######################################################
-
-## load required libraries ----
-library(deSolve)
-library(primer)
-library(plyr)
-library(dplyr)
-library(assertr)
-
-## source required functions ----
-source('Google Drive/ms_ecd/numerical_fxns.R')
-source('Google Drive/ms_ecd/local_stability.R')
-
-## load data
-df.4sp <- read.csv("Google Drive/ms_ecd/hypercube.4sp.csv") %>% tbl_df()
-df.4sp.feas <- df.4sp %>% select(-X) %>% filter(R1>1e-3, R2>1e-3, C1>1e-3, C2>1e-3, steady == 1) # keep feasible and stable equilibriums
-
-## choose parameter set to examine
-set <- as.matrix(df.4sp.feas[1, ])[1, ]
-
-## plot dynamics
-#dynamic_matplot(param.vector = set[5:24],
- #               init.state = c(R1 = R1 <- set[1], 
-  #                             R2 = R2 <- set[2], 
-   #                            C1 = C1 <- set[3], 
-    #                           C2 = C2 <- set[4]),
-     #           sim.length = 1000,
-      #          model = ECD_model, 
-       #         ylim = c(0,max(set[c("K1","K2")])))
-
-## setup simulation
-sim.length <- 1000
-reps <- 50 # number of mutation iterations
-sim.list.4 <- list() 
-sim.list.4[[1]] <- c(set, eq.max.eigen = NA) # input first data vector
-
-## run simulation
-for(i in 2:reps){
-  
-  # extract parameters and state variables
-  ps <- sim.list.4[[i-1]][1:20]
-  i.state <- sim.list.4[[i-1]][23:26]
-  
-  # input parameter values and state variable equilibriums and run the simulation
-  out1 <- ode(parms = ps, y = i.state, times = 1:sim.length, func = ECD_model, rootfun = rootfun) %>% data.frame()
-  
-  # assign mutation in attack rate to one of the species
-  sp <- rbinom(1,1,0.5)
-  mut <- ifelse(rbinom(1,1,0.5) == 0, -0.1, 0.1)
-  
[…]

### 2019-11-07 morgue/parameterized_CR_model.R

@@ -1,201 +1,0 @@
-
-## load required functions ----
-source('code/00_general_parameters.R')
-source('code/00_numerical_fxns.R')
-
-library(rfishbase) # for estimate of stickleback mortality
-library(tidyverse)
-
-mass_scaling <- function(W, intercept, exponent) {
-  log_rate <- intercept + exponent * log(W)
-  return(exp(log_rate))
-}
-
-## BODY SIZE (g) ----
-
-g_Daphnia_pulex <- 1.6 * 10e-4  # from Table 1 in Fenchel 1974 (they cite Frank et al. 1957)
-g_Chironomid <- 1.6 * 10e-3 # very rough estimate, need to confirm
-
-g_limnetic <- 0.81 # from Schluter 1993, Ecology Table 1
-g_benthic <- 1.48  # from Schluter 1993, Ecology Table 1
-g_solitary <- 1.12 # from Schluter 1993, Ecology Table 1
-
-
-## INTRINSIC GROWTH RATE (r) ----
-# r = birth rate per day
-# proportional to M^(-1/4) from Rip and McCann 2011 (they cite Brown et al. 2004)
-# using relationship for heterotherm metazoa from Fenchel 1974, Oecologia
-
-r_Daphnia_pulex <- 0.3          # from Table 1 in Fenchel 1974 (they cite Frank et al. 1957)
-
-r_zoops_estimate <- mass_scaling(W = g_Daphnia_pulex, intercept = -1.6391, exponent = -0.2738) 
-r_zoops_estimate/r_Daphnia_pulex # much higher than Frank et al. 1957 estimate...I'm going to stick with the mass-based estimates though to get the relative differences accurate between zooplankton and benthic inverts
-
-r_benthos_estimate <- mass_scaling(W = g_Chironomid, intercept = -1.6391, exponent = -0.2738) 
-
-
-#### CARRYING CAPACITY (K) ----
-# proportional to M^(-3/4) from Rip and McCann 2011 (they cite Enquist et al. 1998 and Brown et al. 2004)
-K_exponent <- -0.77 # from Brown et al. 2004
-
-K_zoops_estimate <- 5.23 * 1e6 # g per 500 L * individuals per gram assuming average zooplankton size of 1 microgram (based on figure in ms)
-K_intercept <- log(K_zoops_estimate) - K_exponent * log(g_Daphnia_pulex)
-
[…]

### 2019-11-07 morgue/sim_data.R

@@ -1,120 +1,0 @@
-######################################################
-## Description: Identify feasible and stable values for parameters and state variables
-## Code author(s): Matt Barbour
-## Email: barbour@zoology.ubc.ca
-######################################################
-
-## load required libraries ----
-library(deSolve)
-library(rootSolve)
-library(plyr)
-library(dplyr)
-library(assertr)
-library(pse) # for latin hypercube
-
-## load required functions ----
-source('Google Drive/ms_ecd/numerical_fxns.R')
-source('Google Drive/ms_ecd/local_stability.R')
-
-## maximum and minimum values for parameters ----
-rmax <- 5; rmin <- 1; Kmax <- 10; Kmin <- 2; emin <- 0.5; emax <- 0.9; 
-hmax <- 0.6; hmin <- 0.1; amax <- 5; amin <- 1; mmax <- 1; mmin <- 0.1; 
-wmax <- 0.95; wmin <- 0.05
-
-C.init <- 1 # initial value for consumer(s)
-reps <- 50 # of parameter sets to generate
-sim.length <- 1000 # duration of simulations on each parameter set
-
-LHS.4 <- LHS(model=NULL, factors=factors, N=50, q=q, q.arg=q.arg, nboot=0) # use nboot=0 if do not need partial correlations
-LHS.4.ps <- get.data(LHS.4)
-
-LHS.4.ps <- LHS.4.ps %>% mutate(eigen = NA)
-
-## generate for-loop for 4-species community ----
-p.list.4 <- list()
-for(i in 1:reps){
-  
-  # randomly generate parameters, define initial state values, and set simulation length 
-  ps <- c(r1 = r1 <- runif(1, min = rmin, max = rmax), 
-          r2 = r2 <- runif(1, min = rmin, max = rmax), 
-          K1 = K1 <- runif(1, min = Kmin, max = Kmax), 
-          K2 = K2 <- runif(1, min = Kmin, max = Kmax), 
-          e11 = e11 <- runif(1, min = emin, max = emax), 
-          e12 = e12 <- runif(1, min = emin, max = emax), 
-          e21 = e21 <- runif(1, min = emin, max = emax), 
-          e22 = e22 <- runif(1, min = emin, max = emax),  
-          h11 = h11 <- runif(1, min = hmin, max = hmax),
[…]

### 2019-11-07 Appendix_S4.Rmd
tidy up and clarify files in repo

@@
 ---
-#title: |
-#   | Predicting the effects of character displacement on food-web dynamics
-#   | Food-web dynamics under ecological character displacement
-#author: 
-#- Matthew A. Barbour$^{1,2,\ast}$
-# potentially use for 'date' for affiliations, as recommended by https://github.com/rstudio/bookdown/issues/348 for cross-format compatibility, with | to create new sections.
-#date: | 
-#    | 1. Affiliation
 output:
   pdf_document:
-    keep_tex: true
+    keep_tex: false
     toc: false
@@
     fig_caption: true
-    citation_package: natbib 
-    #includes:  
-    #  in_header: Am-Nat-preamble-latex.tex
 header-includes:
@@
   - \captionsetup[figure]{font=small,labelformat=empty}
-#- \usepackage{caption}
-#- \captionsetup[figure]{labelformat=empty}
-#- \usepackage{fullpage}
-#- \linespread{1.7}
-bibliography: references
-biblio-style: amnat
 fontsize: 11pt
@@
 \begin{center} \LARGE Appendix S4: \\ Ecological character displacement destabilizes food webs \end{center}
-<!--\begin{center} \LARGE Food-web dynamics under ecological character displacement \end{center}-->
 
@@
 
-<!-- Remove author info for double-blind peer review; Keep for bioRxiv submission -->
 \begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}
@@
 
-<!-- End title page formatting
-
-\linenumbers{}
-\modulolinenumbers[3]
--->
-
 \newpage
@@
 
-<!--(ignore this blank page for the moment. Rmarkdown can be a pain in the ass.)-->
-
 ```{r McCann_Plot, echo=FALSE, fig.pos="H", fig.cap="Figure S1: **Effect of character displacement when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
@@
 Initial parameter and state variables are given in the main text."}
-knitr::include_graphics("Figure_ECD_McCann.pdf")
+knitr::include_graphics("Appendix_S4_Fig_S1_McCann.pdf")
 ```
@@
 Details of the consumer asymmetry simulation are given in the main text."}
-knitr::include_graphics("asymm_figs.pdf")
+knitr::include_graphics("Appendix_S4_Fig_S2_Asymmetry_MacArthur_LawlorSmith.pdf")
 ```
 
-```{r MacArthur_Plot_asymm, echo=FALSE, eval=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}**Effect of character displacement when consumers can forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
-Panel (A) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
-Panels (B) and (C) show how different tradeoffs indirectly affect total resource abundances (B) and food-web stability (C).
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
-We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation.
-We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
-knitr::include_graphics("Figure_ECD_MacArthur_asymm.pdf")
-```
-
-```{r LS_Plot_asymm, eval=FALSE, echo=FALSE, fig.cap="\\label{fig:LS_Plot_asymm}**Effect of character displacement when consumers *cannot* forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
-Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
-This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6, *w~ij~*=0.4.
-We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation.
-We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
-knitr::include_graphics("Figure_ECD_LS_asymm.pdf")
-```
+\newpage
 
@@
 Details of the consumer asymmetry simulation are given in the main text."}
-knitr::include_graphics("Figure_ECD_McCann_asymm.pdf")
+knitr::include_graphics("Appendix_S4_Fig_S3_Asymmetry_McCann.pdf")
 ```

### 2019-11-07 manuscript_ECD_model.Rmd
added people who gave feedback on MS to acknowledgements.
tidy up and clarify files in repo

@@
 ---
-#title: |
-#   | Predicting the effects of character displacement on food-web dynamics
-#   | Food-web dynamics under ecological character displacement
-#author: 
-#- Matthew A. Barbour$^{1,2,\ast}$
-# potentially use for 'date' for affiliations, as recommended by https://github.com/rstudio/bookdown/issues/348 for cross-format compatibility, with | to create new sections.
-#date: | 
-#    | 1. Affiliation
 output:
@@
     citation_package: natbib 
-    includes:  
-      in_header: Am-Nat-preamble-latex.tex
-bibliography: references
-biblio-style: amnat
+header-includes:
+  - \usepackage{fullpage}
+  - \linespread{1.5} 
+  - \usepackage{lineno}
+  - \usepackage{caption}
+  - \captionsetup[figure]{font=small}
+bibliography: ReferencesAbbrev
+biblio-style: EcolLett
 fontsize: 11pt
@@
 \begin{center} \LARGE Ecological character displacement destabilizes food webs \end{center}
-<!--\begin{center} \LARGE Food-web dynamics under ecological character displacement \end{center}-->
 
@@
 
-\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, 8057 Zurich, Switzerland;
-
-\noindent 2. University of British Columbia, Department of Zoology, Vancouver, BC V6T 1Z4, Canada;
-
-$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch<!--, phone: +41 44 635 57 11-->
+\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;
+
+\noindent 2. University of British Columbia, Department of Zoology, 2212 Main Mall, Vancouver, BC V6T 1Z4, Canada;
+
+$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch, phone: +41 44 635 61 63
 
@@
 
+*Statement of authorship*: MAB conceived the study, derived analytical solutions, performed simulations, and wrote the paper.
+
+\bigskip
+
+*Data accessibility*: No new data were used.
+All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model) and have been archived with Zenodo (INSERT DOI).
+
+\bigskip
+
 *Running title*: Character divergence destabilizes food webs
@@
 
-<!--
+<!-- Remove rest of title page for bioRxiv submission -->
 \bigskip
@@
 
-*Number of words in abstract*:
-
-*Number of words in main text*:
-
-*Number of references*:
-
-*Number of figures*:
-
-End title page formatting -->
+*Number of words in abstract*: 143; 
+*Number of words in main text*: `r wordcountaddin:::word_count("manuscript_ECD_model.Rmd") - 143`;
+*Number of words in text boxes*: NA;
+*Number of references*: 45;
+*Number of figures*: 4;
+*Number of tables*: 0;
+*Number of text boxes*: 0
+
+<!-- End title page formatting -->
 
@@
 Specialization of consumer $j$ on resource $1$ is measured as $\frac{a_{1,j}}{a_{1,j}+a_{2,j}}$, where a value of 0.5 is a complete generalist ($a_{1,j}=a_{2,j}$), and a value of 1 is a complete specialist ($a_{2,j}=0$).
-<!--on resource 1, and a value of 0 is a complete specialist on resource 2.-->
 Values less than 0.5 indicate specialization on the other resource.
@@
 I also relax this assumption and consider a more realistic functional response when resources occurred in different habitats."}
-knitr::include_graphics("ECD_model_foraging_scenarios")
+knitr::include_graphics("Fig_1_ForagingScenarios")
 ```
@@
 An equilibrium is reached when the rates of change in equation \ref{eq:1} are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_i$) and each consumer ($\hat C_j$).
-<!--Importantly, inferences about ecological character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992].
-Therefore, I compared resource abundances between food webs with ($\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$) and without ($\hat R_1,\hat R_2, \hat C_1 > 0$) a competing consumer.-->
 I also compared the local stability of these food webs using standard methods [@Otto2007]. 
@@
 If $-\lambda_{max}<0$, then the food web is not locally stable.
-<!--This dominant eigenvalue determines whether (and how readily) the food web will return to equilibrium after a small perturbation.-->
 
@@
 To gain insight to the eco-evolutionary feedback generated by character displacement, I conducted simulations using an Adaptive Dynamics approach. 
-<!--
-The evolutionary dynamics of character displacement have been well characterized in some of the ecological scenarios examined here [@Lawlor1976; @Abrams1986] and are not the focus of this work. 
-Still, 
-Although divergent character displacement occurs in all scenarios we examined, the magnitude of this divergence may be context dependent.
-Therefore, I simulated the eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
--->
 Specifically, after letting consumer and resource abundances reach a steady state, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability.
@@
 I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
-All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model).
+All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model) and have been archived with Zenodo (INSERT DOI).
 
@@
   ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
-  geom_line() + # size=1.5
-  geom_point(data = initial_points, aes(x=a11, y=a12), size=3, show.legend=FALSE) + # size=5
-  xlab(expression(bold(a[11]))) + # a[ii]
-  ylab(expression(bold(a[21]))) + # a[ij]
-  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette) #+
-  #theme_cowplot(font_size = 30, line_size=2)# +
-  #theme(axis.title = element_text(size = 30),
-  #      axis.text = element_text(size = 20),
-  #      legend.title = element_text(size = 25),
-  #      legend.text = element_text(size = 20),
-  #      legend.spacing = unit(1.0,'cm'))
-
-save_plot(filename = "attack_tradeoffs.pdf", plot = tradeoff_plot, 
-          base_height = 4, # 7
-          base_width = 6) # 7
+  geom_line() + 
+  geom_point(data = initial_points, aes(x=a11, y=a12), size=3, show.legend=FALSE) + 
+  xlab(expression(bold(a[11]))) +
+  ylab(expression(bold(a[21]))) + 
+  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette)
+
+save_plot(filename = "Fig_2_Tradeoffs.pdf", plot = tradeoff_plot, 
+          base_height = 4,
+          base_width = 6) 
 ```
@@
 Points corresponding to attack rates at the beginning of the simulation for *C*~1~, which was pre-adapted to *R*~1~ (*a*~11~ > *a*~12~)."}
-knitr::include_graphics("attack_tradeoffs.pdf")
+knitr::include_graphics("Fig_2_Tradeoffs.pdf")
 ```
@@
 ## Resources occur in same habitat
-
-<!-- I haven't made clear that this explores a specific scenario, MacArthur's model
-
-In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972].
-The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
-I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
--->
 
@@
 
-<!-- My simulations show that the shape of the tradeoff qualitatively affects the relationship between character displacement and resource abundances (fig. \ref{fig:plot_fig3}).-->
 I found that the shape of the tradeoff function qualitatively affects the relationship between character displacement and resource abundances in this scenario (fig. \ref{fig:plot_fig3} middle left).
@@
   geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(a[ii]+a[ij])) +
-  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
   scale_shape_manual(values=c(24,21)) +
@@
   geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
-  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
-  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
   scale_shape_manual(values=c(24,21)) +
@@
 Mac_plots
+```
+
+```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource abundances (B) when consumers cannot forage for resources simultaneously."}
+LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
+  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+LS_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color=factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(R[i]+R[j])) +
+  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), LS_resources, 
+                      ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
+LS_plots
 ```
@@
   geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + 
   geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
- #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  #geom_hline(yintercept=0, color="grey") +
+  ylab(expression("Stability "(-lambda[max]))) + 
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
   guides(fill = guide_legend(override.aes=list(shape=21)))
@@
   geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  #geom_hline(yintercept=0, color="grey") +
+  ylab(expression("Stability "(-lambda[max]))) + 
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
   guides(fill = guide_legend(override.aes=list(shape=21)))
@@
 
-```{r MacArthur Model, include=FALSE}
-plot_macarthur_model <- plot_grid(
-  tradeoff_plot + theme(legend.position = "none"), 
-  Mac_attack  + theme(legend.position = "none"), 
-  Mac_resources  + theme(legend.position = "none"),
-  Stability_Mac  + theme(legend.position = "none"), ncol=2, align='hv', labels = "AUTO")
-
-save_plot(filename="Figure_ECD_MacArthur.pdf", plot=plot_macarthur_model, base_height = 6, base_width = 6)
-```
-
-```{r MacArthur_Plot, include=FALSE, eval=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}**Effect of character displacement when consumers can forage for both resources simultaneously.** Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Panel (B) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation.
-Panels (C) and (D) show how different tradeoffs indirectly affect total resource abundances (C) and food-web stability (D).
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
-We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-We set initial consumer and resource abundances as: *R~i~* = 2; *C~j~* = 1."}
-knitr::include_graphics("Figure_ECD_MacArthur.pdf")
-```
-
 ## Resources occur in different habitats
-
-<!--
-The previous model, although the starting point for many models of competition, does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in distinct habitats [@McCann2005].
-The only character displacement model that I am aware of that examined the effect of resources in different habitats was studied by @Lawlor1976.
-This model is the same as before, except now the consumer's feeding rate takes the form:
-
-$$F_{ji}(R_i)=w_{ji}a_{ji}R_i$$
-
-where $w_{ji}$ represents the proportion of time consumer $j$ spends foraging in a habitat where only resource $j$ is found (i.e.
-habitat preference).
-Note that since $w_{ji}$ is a proportion that $w_{jj}=1-w_{ji}$.
-As with attack rates, I assume that consumer habitat preferences are mirror images of each other ($w=w_{11}=w_{22}$).
-In @Lawlor1976's detailed evolutionary analysis of this model, they always observed divergent character displacement.
-
-To examine the ecological consequences of this divergence, I again derive the equilibrium solution when both consumers and both resources are present. 
-As before, resource abundances are equivalent at this equilibrium ($\hat R = \hat R_1 = \hat R_2$ when both consumers and resources present), but are now determined by the following equation (derivation in supplementary material):
--->
 
@@
 \ref{fig:plot_fig3} right panels).
-<!--This is because the ESS for a single consumer is to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in fig.
-\ref{fig:plot_fig3} right panels), resulting in higher resource abundances compared to if it evolved to be a complete generalist.-->
 
@@
 
-```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource abundances (B) when consumers cannot forage for resources simultaneously."}
-LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
-  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-LS_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color=factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(R[i]+R[j])) +
-  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), LS_resources, 
-                      ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
-LS_plots
-```
-
-```{r LS Model, include=FALSE}
-plot_LS_model <- plot_grid(
-  LS_attack  + theme(legend.position = "none"), 
-  LS_resources  + theme(legend.position = "none"),
-  Stability_LS  + theme(legend.position = "none"), ncol=2, align='hv', labels = "AUTO")
-
-save_plot(filename="Figure_ECD_LS.pdf", plot=plot_LS_model, base_height = 6, base_width = 6)
-```
-
-```{r LS_Plot, include=FALSE, eval=FALSE, fig.cap="\\label{fig:LS_Plot}**Effect of character displacement when consumers *cannot* forage for both resources simultaneously.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points (only in panel C).
-Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of the consumer community.
-This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *w~ii~*=0.6; *w~ij~*=0.4.
-We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
-knitr::include_graphics("Figure_ECD_LS.pdf")
-```
-
 ```{r Fig 3, include=FALSE}
-# plot.margin = unit(c(0, 0, 0, 0), "cm") to shrink space between plots
-
 Macs <- plot_grid(
-  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), #  + scale_y_continuous(breaks=c(1.8,2.0,2.2))
-  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)), # + scale_y_continuous(breaks=c(1.0,1.2,1.4), limits=c(1.0,1.45))
-  Stability_Mac + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)), # + scale_y_continuous(breaks=c(0,0.1))
-  ncol=1, align='v'#, labels = "AUTO"
-) # c(0,0.02,0.04,0.06,0.08,0.1)
+  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)),
+  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)),
+  Stability_Mac + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)),
+  ncol=1, align='v'
+)
 
@@
 
-save_plot(filename="Figure_3.pdf", plot=plot_fig3_sub, base_height = 6.5, base_width = 6)
+save_plot(filename="Fig_3_MacArthur_LawlorSmith.pdf", plot=plot_fig3_sub, base_height = 6.5, base_width = 6)
 ```
@@
 In both foraging scenarios, feeding rates increase linearly with resource abundance, but the equation for effective attack rate is different."}
-knitr::include_graphics("Figure_3.pdf")
+knitr::include_graphics("Fig_3_MacArthur_LawlorSmith.pdf")
 ```
@@
 
-<!--
-So far I have only considered a simple functional response, where consumer feeding rates increase linearly with resource abundances when foraging in either the same or distinct habitats. 
-We know, however, that consumer feeding rates often saturate at high resource abundances [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011]. 
-We also know that consumers do not usually spend a fixed proportion of time in a particular habitat, but their habitat preferences may be modified by resource abundances [@McCann2005]. 
-To account for these biological realities, I consider a more realistic functional response [derived by @McCann2005]:
-
-$$F_{ii}(R_i,R_j)=\frac{a_{ii}W_{ii}R_i}{1+a_{ii}hW_{ii}R_i+a_{ij}hW_{ij}R_j}$$
-
-where consumer $i$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling time $h$); and consumer habitat preferences are modified by the relative abundance of resources, such that:
-
-$$W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}$$
+<!-- Keeping text in case reviewers want me to explain evolutionary dynamics of this model in the text
 
@@
 In the third foraging scenario (equation \ref{eq:4}), I observed the same general effect of character divergence as the previous scenario (resources in different habitats, but linear functional response).
-<!--Divergent character displacement in this foraging scenario has the same general effect on food-web dynamics as the previous scenario (resources in different habitats, but linear functional response).
-<!--In terms of food-web dynamics, I observe the same general effect of ecological character displacement as before when resources occurred in distinct habitats, but consumers had a simpler functional response.
--->
 This is because resource abundances at equilibrium are governed by a similar dynamic (derived in Appendix S1):
-<!--
-Even after adding more biological realism, I still observe the same qualitative relationship between character displacement and resource suppression (fig.
-\ref{fig:McCann_Plot} in Appendix).
-This is because resource competition always leads to character divergence and equilibrium resource abundances are governed by similar dynamics  (derivation in supplementary material):
--->
 
@@
 
-<!--Previously, I showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff (fig.
-\ref{fig:LS_Plot}A).-->
 Thus, character displacement always pushes the food web toward an unstable structure in this more realistic foraging scenario (fig.
@@
   geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(bquote(atop("Effective attack rate","("*w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j]*")"))) +
-  #ylab(expression(w*a[ii]+(1-w)*a[ij])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
   geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) +
-  #ylab(expression(R[i]+R[j])) +
   xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
@@
   geom_line(aes(color=factor(n))) +
-  #geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend = FALSE) +
-  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   geom_point(data = McCann_displacements, aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor, size=Time), color="white", inherit.aes = FALSE) +
@@
   ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) +
-  #ylab(expression("Stability "(-lambda[max]))) + 
   scale_shape_manual(values=c(24,21))  +
@@
   scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  #geom_hline(yintercept=0, color="grey") +
   guides(fill = guide_legend(override.aes=list(shape=21)))
@@
   McCann_resources    + theme(legend.position = "none") + xlab(""),
-  Stability_McCann    + theme(legend.position = "none") + xlab(""), ncol=2, align='hv') # , labels = "AUTO"
+  Stability_McCann    + theme(legend.position = "none") + xlab(""), ncol=2, align='hv')
 
@@
 
-#plot_McCann_model <- ggdraw(plot_McCann_model) + draw_grob(McCann_legend, x = 0.7, y=-0.2)
-
-save_plot(filename="Figure_ECD_McCann.pdf", plot=plot_McCann_model, base_height = 6, base_width = 6)
-# moved to supplementary information
+# plotted in Appendix S4, Fig. S1
+save_plot(filename="Appendix_S4_Fig_S1_McCann.pdf", plot=plot_McCann_model, base_height = 6, base_width = 6)
 ```
@@
 
-# when r = 2.5, A = 3.3 and concaveUp=0.85, we get consmers evolving toward instability; we don't get this when n=n_linear but we do if we increase mut.size=0.01 (order of magnitude) and even when r=1!
-
-## IDEAL PARAMETERS RIGHT NOW: A=3.3, r=1, mut.size=0.01, n=n_linear (excellent looking figure with bifurcation just after0.8)
-
 ## set up general code for bifurcation plot; increase to get bifurcation
-# when A = 2.5 get strong divergence
-# at 2.9, consumer doesn't divergence, C1 becomes generalist though
-# at 2.8, there is less divergence...
 A <- 3.3
-# K <- K+0.5 # increasing K to 4.5 creates a bifurcation at extreme specialization, but the consumer no longer evolves...
-#r <- r #2.5 # r=1.5, now the consumer is still diverging when A = 2.9... at 3.3 though, there is no longer divergence in two consumer situation; get divergence again when r=2
-#aii <- 2.4
-#aij <- 2.0
-
-# assuming linear trade-off in attack rates for simulation; sometimes, the consumer doesn't evolve toward asymmetry though...
-#linear_df <- data.frame(n = n_linear, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_linear)^(1/n_linear))
-aii_seq <- seq(A/2, A, by = 0.01)#[1:75] # seq(aii, A, by = 0.01)[1:75]
-aij_seq <- (A*(1-(aii_seq/A)^n_linear)^(1/n_linear)) # n_linear #[1:75] #seq(A/2, 0, by = -0.01)[1:75]
+
+# assuming linear trade-off in attack rates for simulation
+aii_seq <- seq(A/2, A, by = 0.01)
+aij_seq <- (A*(1-(aii_seq/A)^n_linear)^(1/n_linear))
 
@@
 
-# consumer abundances
-#bind_rows(mutate(sim.df, size = 4), mutate(sim.df_3, size = 3)) %>%
-#  filter(species == "C") %>%
-#  ggplot(., aes(x = a11/(a11+a12), y = value, group=interaction(max_min, size), color=as.factor(size))) +
-#  geom_line() +
-#  ylab("Consumer abundance (max and min)")
-
-#get_stability_boundary <- bind_rows(mutate(sim.df, Competitor = "Yes"), mutate(sim.df_3, Competitor = "No")) %>%
-#  filter(species == "R") %>%
-#  spread(max_min, value)
-
 # with eco-evo begin and end points
@@
   geom_point(data = McCann_displacements_bifur, aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor, size=Time), color=cbPalette[2], fill=cbPalette[2], inherit.aes = FALSE) +
-  #geom_point(data = filter(McCann_displacements_bifur, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor), color="white", fill=cbPalette[2], size=6, inherit.aes = FALSE) +
-  #geom_point(data = filter(McCann_displacements_bifur, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor), color="white", fill=cbPalette[2], size=3, inherit.aes = FALSE) +
   scale_shape_manual(values=c(24,21))  +
@@
   xlab(expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")")))) +
-  #guides(fill = guide_legend(override.aes=list(shape=21))) +
   geom_segment(x=0.815, xend=0.815, y=0, yend=2.65, arrow = arrow(length = unit(0.03, "npc")), color="grey", fill="grey") +
@@
   theme_cowplot(font_size = 18)
-  #geom_vline(xintercept = 0.82, linetype="dotted", color="grey")
-
-save_plot(filename = "bifurcation_plot.pdf", bifurcation_plot, base_width = 8.5, base_height = 6)
+
+save_plot(filename = "Fig_4_Bifurcation.pdf", bifurcation_plot, base_width = 8.5, base_height = 6)
 ```
@@
 Note that I increased the total investment in attack rates (*A* = 3.3) to create a scenario that could result in an unstable food web."}
-knitr::include_graphics("bifurcation_plot.pdf")
+knitr::include_graphics("Fig_4_Bifurcation.pdf")
 ```
@@
 I found that my previous inferences are robust to including consumer asymmetry across different foraging scenarios and tradeoffs (Appendix S4).
-<!--A\ref{fig:MacArthur_Plot_asymm}, A\ref{fig:LS_Plot_asymm}, A\ref{fig:McCann_Plot_asymm}).-->
 
 # Discussion
+
+<!-- original opening paragraph of discussion; I think it provides too good of a summary though to put at the beginning. It makes subsequent discussion sound redundant
+
+My results show that character displacement leads to predictable differences in resource abundances and the resilience of food webs to perturbations.
+Specifically, I found that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource abundances and less resilient food webs. 
+Character displacement always increased attack rates whenever consumers were competing for resources that occured in distinct habitats. 
+This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
+Although character displacement did not always increase effective attack rates, these contingent effects only occurred under the simplest and (arguably) least realistic foraging scenario. 
+My study takes advantage of past work showing that adaptive divergence in consumer attack rates is always favored under the scenarios examined here [@LawlorSmith1976; @Abrams1986].
+Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations. 
+-->
 
@@
 
-<!--
-The simplest foraging scenario examined here (linear functional response on resources occuring in the same habitat) is the foundation of mechanistic models of resource competition [@MacArthur1972; @Lawlor1976; @Abrams1986].
-Under this scenario, it is theoretically possible that character displacement results in increased resource abundances [green lines in \ref{fig:MacArthur_Plot}].
-This only occurs, however, under an evolutionary tradeoff that constrains the magnitude of character displacement.
-This suggests that it would be unlikely to ever observe divergent character displacement and increases in resource abundances in a real system. 
-In contrast, the evolutionary tradeoff that favors character displacement the most decreased resource abundances. 
-Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the foraging rate of consumers (orange lines in \ref{fig:MacArthur_Plot}A,B).
-This is evident by the disruptive selection observed for beak size/shape in Darwin's finches [@Schluter1985] and benthic/limnetic ecotypes for stickleback [@Arnegard2014] in resource-limited environments.
-<!-- key point of the paragraph 
-Taken together, my model predicts that even in the simplest foraging scenario, the most likely relationship we would observe in nature is for food webs where consumers have undergone character displacement to have lower total resource densites.
-
-A general result emerging from my analysis was that, when consumers were competing for resources in distinct habitats, character displacement always supressed resource abundances. 
-<!-- show that this is common scenario 
-Consumers are generally larger, and as a result, more mobile than their resources; thus, their foraging behavior acts to couple resource dynamics across traditional habitat boundaries [@McCann2005].
-An example of character displacement that matches this foraging scenario is threespine stickleback in postglacial lakes of British Columbia, Canada [@Schluter1992; @Schluter2000].
-In these lakes, stickleback have undergone character displacement into limnetic and benthic species that have evolved specialized traits to forage on zooplankton and benthic invertebrates, respectively.
-These resources occur in distinct zones of the lake (pelagic and littoral) and stickleback move between these zones when foraging for resources. 
-<!-- good closing sentences --
-A disproporationate number of the documented cases of character displacement involve carnivores that are larger, and likely more mobile, than their resources [@Schluter2000],
-Therefore, my predictions may apply to many cases of ecological character displacement.
--->
-
 <!-- consider removing entirely. This is a prediction I could test with Seth in another paper...
@@
 Detecting the ghost of competition past [@Connell1980] may be quite difficult, but it could be possible with recent advances in genomics.
-<!--Testing this hypothesis ---that some lakes with a single species had a species pair in the past --- may be quite difficult, but it could be possible with recent advances in genomics.-->
 For example, @Feulner2018 detected genomic signatures of hybridization in sympatric whitefish species following periods of eutrophication.
 Perhaps solitary stickleback in some lakes retain genomic signatures of having been a habitat specialist in the past.
-
-<!-- interesting, but tighten up 
-A plausible explanation for this rarity is that the food-web structure makes the species pair more vulnerable to extinction, and many lakes that used to have a species pair have  single species of stickleback are more stable
-A plausible explanation for this rarity is that nonadaptive selection results in a more ecologically stable food-web topology [@Borrelli2015a; @Borrelli2015b].
-Testing this hypothesis ---some lakes with a single species had two in the past --- may be quite difficult, but it may be feasible with recent advances in genomics.
-For example, @Feulner2018 detected genomic signatures of hybridization in sympatric whitefish species following periods of eutrophication.
-Perhaps stickleback in some lakes retain genomic signatures of having been a habitat specialist in the past.
-Other evidence also suggests that the species pair of stickleback are vulnerable to perturbations.
-For example, the introduction of nonnative species into two of the six lakes have resulted in the extinction of the species pair [@Hatfield2001; @Taylor2006; @Rudman2016].
--->
 
@@
 Note that this does not negate my general conclusion that ecological character displacement decreases resource abundances and stability; however, this process may itself create another eco-evolutionary feedback between consumers and resources.
-<!--Therefore, the prey's evolutionary response would act to stabilize food webs, rather than coevolution *per se*.-->
 This may actually help maintain dramatic examples of character divergence and prevent them from destabilizing systems, because it allows consumer traits to become decoupled from their attack rate.
@@
 This is because models that explicitly include resource dynamics inevitably show that resource competition results in character divergence, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
-A quantitative genetic model may certainly show differences in the <!--magnitude-->pace of character divergence, but this should not qualitatively change its effect on food-web dynamics.
-<!-- In fact, recent work by @McPeek2017 has shown that divergence in a quantitative genetics model can lead to decreased resource abundances and unstable dynamics.
--->It is important to note that my conclusions only apply to food webs with biotic resources that are nutritionally substitutable.
+A quantitative genetic model may certainly show differences in the pace of character divergence, but this should not qualitatively change its effect on food-web dynamics. 
+It is important to note that my conclusions only apply to food webs with biotic resources that are nutritionally substitutable. 
 It would be interesting to extend these current analyses to non-substitutable resources where character convergence is expected [@Abrams1987; @Fox2008].
@@
 This destabilizing effect emerges from an eco-evolutionary feedback involving direct and indirect interactions between species in a food-web context.
-<!--This highlights how the process of adaptation, when considered in a food-web context, may generate feedbacks that destabilize the interconnected parts of the system.
-This suggests that observed patterns of biodiversity may not simply be a result of evolutionary constraints, but also ecological constraints
-To what extent are current patterns in biodiversity a result of ecological constraints on the persistence of evolved phenotypes? -->
 This result contrasts with the current notion that patterns of phenotypic diversity are solely the result of evolutionary constraints imposed by mutation, natural selection, gene flow, and genetic drift.
@@
 I expect that identifying when and where this ecological constraint arises will yield novel insight to the patterns of biodiversity we see in nature.
-<!--Identifying when and where we expect this ecological constraint to arise may yield novel insight to the patterns of biodiversity we see in nature. -->
-
-<!-- Omitting conclusions section for now. I will wait for feedback from friendly reviews
-
-## Conclusions
-
-<!-- originally opening paragraph of discussion; I think it provides too good of a summary though to put at the beginning. It makes subsequent discussion sound redundant
-My results show that character displacement leads to predictable differences in resource abundances and the resilience of food webs to perturbations.
-Specifically, I found that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource abundances and less resilient food webs. 
-Character displacement always increased attack rates whenever consumers were competing for resources that occured in distinct habitats. 
-This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
-Although character displacement did not always increase effective attack rates, these contingent effects only occurred under the simplest and (arguably) least realistic foraging scenario. 
-My study takes advantage of past work showing that adaptive divergence in consumer attack rates is always favored under the scenarios examined here [@LawlorSmith1976; @Abrams1986].
-Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations.
-
-<!--
-## Conclusions
-- needs revision
-- do this after I rewrite/edit the rest of the Discussion
-
-Competition theory suggested that character displacement was a stabilizing force in ecological communities [@Lawlor1976]. 
-My results show the exact opposite, when a food-web perspective is taken.
-Importantly, a food-web approach permits comparisons between communities in the presence of absence of a competitor, allowing a fair comparison of the effects of character displacement on food-web dynamics.
-These results appear to be quite general, as they characterize the general foraging scenario where consumers are more mobile than their resources. 
-These results also suggest that the very evolutionary process that generates biodiversity, creates an ecological scenario that makes that diversity fragile to loss. 
-These results highlight how evolution, when considered in a food-web context, may actually be a destabilizing force in ecological communities. 
-This may generally be true of consumer evolution, as was pointed out decades ago [@Rosenzweig1963].
-These results also highlight a tension between adaptive process (evolution of consumers to increase feeding efficiency) and a non-adaptive process (food-web stability).
-These results suggest that in systems with evidence of ecological character displacement, why do they persist? 
-These results open up a new avenue for research on ecological character displacement, one that explores in more detail the dynamical consequences of consumer evolution, rather than just static effects that assume resources are persistent and stable.
-It also suggest that actually observing evidence of ecological character displacement may be difficult ...
-It also highlights how evolution of key species in a food web may make these systems more susceptible to environmental change. 
-It's been argued that "inferences about the evolution of community structure cannot be based on ecological arguments about community stability or likely coexistence" [@McPeek2019].
-In contrast, our results suggest that current patterns in food-web structure may be the result of non-adaptive processes selecting for more stable food-web configurations.
-Our results also call for a revision of one of the criteria of character displacement, since it is clear that this process can result in different resource dynamics.
-In fact, this resource similarity hypothesis has not been tested in natural systems, although it has been qualitatively inferred, and some quantitative estimates based on physical properties of systems have been measured to try and say these habitats are similar.
-(NEED TO CHECK SCHLUTER 2000 BOOK, THE AM NAT STUDY SUGGESTS THAT IT IS A COMMONLY FILLED CRITERIA.)
-More generally, our results highlight how variation in food-web structure can alter eco-evolutionary feedbacks within communities [@McPeek2017], that also have consequences for their resilience to environmental disturbances.
-This calls for food-web ecologists to more criticaly consider the role of evolutionary processes in shaping food-web structure and dynamics.
--->
 
@@
 I would not have been able to take the theory as far as I did without her guidance and encouragement.
+Matt Osmond, Jean Gibert, and Seth Rudman provided critical feedback on an earlier draft.
 For funding support, I thank the University of British Columbia (Four-Year Fellowship to M.A.
@@
 ```{r MacArthur_asymm, include=FALSE}
-# replaced MacArthur_displacements with MacArthur_displacements_asymm
-
 Mac_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
@@
   geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(a[ii]+a[ij])) +
@@
   geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
-  #geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
 
-save_plot(filename="Figure_ECD_MacArthur_asymm.pdf", plot=Mac_plots_asymm, base_height = 6, base_width = 6)
-```
-
-```{r MacArthur_Plot_asymm, include=FALSE, eval=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot_asymm}**Effect of character displacement when consumers can forage for both resources simultaneously and have an asymmetry in initial attack rates.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation, whereas as small shapes correspond to the starting points.
-Panel (A) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community.
-Panels (B) and (C) show how different tradeoffs indirectly affect total resource abundances (B) and food-web stability (C).
-These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2.
-We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*.
-To create an asymmetry in initial attack rates, we used the ending *a~ii~* and *a~ij~* values for the one consumer simulation as the starting values for *C~2~* in the two consumer simulation.
-We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
-knitr::include_graphics("Figure_ECD_MacArthur_asymm.pdf")
+# Plot with LawlorSmith asymm in Appendix S4 (see "all asymmetry" chunk below)
+# save_plot(filename="Figure_ECD_MacArthur_asymm.pdf", plot=Mac_plots_asymm, base_height = 6, base_width = 6)
 ```
@@
 ```{r LawlorSmith_asymm, include=FALSE}
-# replaced LawlorSmith_displacements with LawlorSmith_displacements_asymm
-
 LS_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
@@
   geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(w*a[ii]+(1-w)*a[ij])) +
@@
   geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(expression(R[i]+R[j])) +
@@
 
-save_plot(filename="Figure_ECD_LS_asymm.pdf", plot=LS_plots_asymm, base_height = 6, base_width = 6)
-# moved to supplementary information
+# Plot with MacArthur asymm in Appendix S4 (see "all asymmetry" chunk below)
+# save_plot(filename="Figure_ECD_LS_asymm.pdf", plot=LS_plots_asymm, base_height = 6, base_width = 6)
+```
+
+```{r all asymmetry, echo=FALSE}
+Macs_asymm <- plot_grid(
+  Mac_attack_asymm + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))),
+  Mac_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))),
+  Stability_Mac_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))),
+  ncol=1, align='v'
+)
+
+LSs_asymm <- plot_grid(
+  LS_attack_asymm  + ggtitle("Resources: Different Habitats\n") +  theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)) + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])),
+  LS_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x=element_blank()) + xlab("") + ylab(""),
+  Stability_LS_asymm   + theme(legend.position = "none", axis.text = element_text(size=10)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(""),
+  ncol=1, align='v'
+)
+
+all_asymm_plots <- plot_grid(Macs_asymm, LSs_asymm, ncol=2, rel_widths = c(1,0.9)) 
+all_asymm_plots <- add_sub(all_asymm_plots, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
+
+# plotted in Appendix S4, Fig. S2
+save_plot(filename="Appendix_S4_Fig_S2_Asymmetry_MacArthur_LawlorSmith.pdf", plot=all_asymm_plots, base_height = 6.5, base_width = 6)
 ```
@@
 ```{r McCann_asymm, include=FALSE}
-# replaced McCann_displacements with McCann_displacements_asymm
-
 McCann_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
@@
   geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(bquote(atop("Effective attack rate","("*w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j]*")"))) +
@@
   geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
   ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) +
@@
   xlab(expression(frac(a[ij],a[1*","*j]+a[2*","*j]))) +
-  ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  #geom_hline(yintercept=0, color="grey") +
+  ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + 
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
   guides(fill = guide_legend(override.aes=list(shape=21)))
@@
 
-save_plot(filename="Figure_ECD_McCann_asymm.pdf", plot=plot_McCann_asymm, base_height = 6, base_width = 6)
-# moved to supplementary information
-```
-
-```{r all asymmetry, echo=FALSE}
-
-Macs_asymm <- plot_grid(
-  Mac_attack_asymm + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))),
-  Mac_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))),
-  Stability_Mac_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))),
-  ncol=1, align='v'#, rel_widths = c(1,0.9,0.9)
-)
-
-LSs_asymm <- plot_grid(
-  LS_attack_asymm  + ggtitle("Resources: Different Habitats\n") +  theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)) + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])),
-  LS_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x=element_blank()) + xlab("") + ylab(""),
-  Stability_LS_asymm   + theme(legend.position = "none", axis.text = element_text(size=10)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(""),
-  ncol=1, align='v'#, rel_widths = c(1,0.9,0.9)
-)
-
-all_asymm_plots <- plot_grid(Macs_asymm, LSs_asymm, ncol=2, rel_widths = c(1,0.9)) 
-all_asymm_plots <- add_sub(all_asymm_plots, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
-
-save_plot(filename="asymm_figs.pdf", plot=all_asymm_plots, base_height = 6.5, base_width = 6)
-```
+# plotted in Appendix S4, Fig. S3
+save_plot(filename="Appendix_S4_Fig_S3_Asymmetry_McCann.pdf", plot=plot_McCann_asymm, base_height = 6, base_width = 6)
+```

### 2019-11-07 manuscript_ECD_model.tex
tidy up and clarify files in repo

@@
 \usepackage{natbib}
-\bibliographystyle{amnat}
+\bibliographystyle{EcolLett}
 \usepackage{graphicx,grffile}
@@
 
-%%% From AmNat_MS_template.tex
-%\documentclass[11pt]{article}
-%\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
 \usepackage{fullpage}
-%\usepackage[authoryear,sectionbib,sort]{natbib}
-\linespread{1.5} % 1.7
-%\usepackage[utf8]{inputenc}
+\linespread{1.5}
 \usepackage{lineno}
-
-% controlling text size in figure captions
 \usepackage{caption}
 \captionsetup[figure]{font=small}
-
-%%%%%%% Sections not currently working properly
-%\usepackage{titlesec}
-%\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
-%\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
-%\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
-%\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}
-
-%%% My added packages
-%\usepackage{amsmath} % for matrices
-%\usepackage{graphicx} % for organizing figures
-%\usepackage{lmodern} % for tilde
-%\usepackage[T1]{fontenc} % for tilde
-
-%%%%%%%%% Previous - likely useful though
-%\usepackage{setspace}\doublespacing
-%\usepackage{float}
-%\let\origfigure\figure
-%\let\endorigfigure\endfigure
-%\renewenvironment{figure}[1][2] {
-%    \expandafter\origfigure\expandafter[H]
-%} {
-%    \endorigfigure
-%}
 
@@
 \noindent 1. University of Zurich, Department of Evolutionary Biology
-and Environmental Studies, 8057 Zurich, Switzerland;
+and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich,
+Switzerland;
 
-\noindent 2. University of British Columbia, Department of Zoology,
-Vancouver, BC V6T 1Z4, Canada;
+\noindent 2. University of British Columbia, Department of Zoology, 2212
+Main Mall, Vancouver, BC V6T 1Z4, Canada;
 
 \(^\ast\) Corresponding author; e-mail:
-\href{mailto:matthew.barbour@ieu.uzh.ch}{\nolinkurl{matthew.barbour@ieu.uzh.ch}}
+\href{mailto:matthew.barbour@ieu.uzh.ch}{\nolinkurl{matthew.barbour@ieu.uzh.ch}},
+phone: +41 44 635 61 63
+
+\bigskip
+
+\emph{Statement of authorship}: MAB conceived the study, derived
+analytical solutions, performed simulations, and wrote the paper.
+
+\bigskip
+
+\emph{Data accessibility}: No new data were used. All code used to
+reproduce this manuscript and simulations within are publically
+available on GitHub (\url{https://github.com/mabarbour/ECD_model}) and
+have been archived with Zenodo (INSERT DOI).
 
@@
 consumer-resource interactions; adaptation; community stability
+
+\bigskip
+
+\emph{Manuscript type}: Letter.
+
+\bigskip
+
+\emph{Number of words in abstract}: 143; \emph{Number of words in main
+text}: 4086; \emph{Number of words in text boxes}: NA; \emph{Number of
+references}: 45; \emph{Number of figures}: 4; \emph{Number of tables}:
+0; \emph{Number of text boxes}: 0
 
@@
 \centering
-\includegraphics{ECD_model_foraging_scenarios}
+\includegraphics{Fig_1_ForagingScenarios}
 \caption{\label{fig:foraging_scenarios}\textbf{Ecological foraging
@@
 in R \citep{R} and the code to reproduce these simulations is publically
-available on GitHub (\url{https://github.com/mabarbour/ECD_model}).
+available on GitHub (\url{https://github.com/mabarbour/ECD_model}) and
+have been archived with Zenodo (INSERT DOI).
 
@@
 \centering
-\includegraphics{attack_tradeoffs.pdf}
+\includegraphics{Fig_2_Tradeoffs.pdf}
 \caption{\label{fig:tradeoff}\textbf{Evolutionary tradeoffs in consumer
@@
 \centering
-\includegraphics{Figure_3.pdf}
+\includegraphics{Fig_3_MacArthur_LawlorSmith.pdf}
 \caption{\label{fig:plot_fig3}\textbf{Effect of character displacement
@@
 \centering
-\includegraphics{bifurcation_plot.pdf}
+\includegraphics{Fig_4_Bifurcation.pdf}
 \caption{\label{fig:bifur_plot}\textbf{Character divergence creates an
@@
 analyses of the mathematical model. I would not have been able to take
-the theory as far as I did without her guidance and encouragement. For
-funding support, I thank the University of British Columbia (Four-Year
-Fellowship to M.A.~Barbour), NSERC (Discovery grant to Greg Crutsinger),
-and the Swiss National Science Foundation (grant 31003A\_160671 to Jordi
-Bascompte).
+the theory as far as I did without her guidance and encouragement. Matt
+Osmond, Jean Gibert, and Seth Rudman provided critical feedback on an
+earlier draft. For funding support, I thank the University of British
+Columbia (Four-Year Fellowship to M.A.~Barbour), NSERC (Discovery grant
+to Greg Crutsinger), and the Swiss National Science Foundation (grant
+31003A\_160671 to Jordi Bascompte).
 
-\bibliography{references}
+\bibliography{ReferencesAbbrev}
 

### 2019-11-18 Appendix_S4.Rmd
incorporate final (minor) edits to main and supplementary text.

@@
   - \usepackage{lineno}
-  - \usepackage{caption}
+  - \usepackage{caption,setspace} # sets spread to 1 within captions
   - \captionsetup[figure]{font=small,labelformat=empty}
@@
 
-\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, 8057 Zurich, Switzerland;
+\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;
 
-\noindent 2. University of British Columbia, Department of Zoology, Vancouver, BC V6T 1Z4, Canada;
+\noindent 2. University of British Columbia, Department of Zoology, 2212 Main Mall, Vancouver, BC V6T 1Z4, Canada;
 
-$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch
+$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch, phone: +41 44 635 61 63
 
@@
 
-```{r McCann_Plot, echo=FALSE, fig.pos="H", fig.cap="Figure S1: **Effect of character displacement when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~, whereas as small shapes correspond to the starting points (only in stability panel).
+```{r McCann_Plot, echo=FALSE, fig.pos="H", fig.cap="Figure S1: **Effect of character displacement when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~ (choice to display *C*~1~ was arbitrary), whereas as small shapes correspond to the starting points (only in stability panel).
 This figure illustrates how, regardless of the tradeoff, character displacement increases the effective attack rate of consumers.
@@
 Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~, whereas as small shapes correspond to the starting points (only in stability panels).
-In both foraging scenarios, feeding rates increase linearly with resource abundance, but the equation for effective attack rate is different.
-The similarity with fig.
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~ (choice to display *C*~1~ was arbitrary), whereas as small shapes correspond to the starting points (only in stability panels).
+In both foraging scenarios, feeding rates increase linearly with resource abundance, but the equation for the effective attack rate is different.
+The similarity with Fig.
 3 (main text) shows that adding an asymmetry in consumer attack rates does not alter the conclusions reported in the main text.
@@
 
-```{r McCann_Plot_asymm, echo=FALSE, fig.cap="Figure S3: **Robustness to consumer asymmetry when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoff values (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~, whereas as small shapes correspond to the starting points.
-Panel (A) shows how, regardless of the tradeoff, character displacement increases the effective attack rate of consumers.
-This always resulted in a suppression of resource abundances (B) and concomitant decrease in food-web stability (C).
-The similarity with fig.
+```{r McCann_Plot_asymm, echo=FALSE, fig.cap="Figure S3: **Robustness to consumer asymmetry when consumers exhibit a more realistic functional response.** Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~ (choice to display *C*~1~ was arbitrary), whereas as small shapes correspond to the starting points.
+Regardless of the tradeoff, character displacement increases the effective attack rate of consumers (a).
+This always resulted in a suppression of resource abundances (b) and concomitant decrease in food-web stability (c).
+The similarity with Fig.
 S1 shows that adding an asymmetry in consumer attack rates does not alter the conclusions reported in the main text.

### 2019-11-18 manuscript_ECD_model.Rmd
incorporate final (minor) edits to main and supplementary text.

@@
   - \usepackage{lineno}
-  - \usepackage{caption}
+  - \usepackage{caption,setspace}
   - \captionsetup[figure]{font=small}
@@
 *Data accessibility*: No new data were used.
-All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model) and have been archived with Zenodo (INSERT DOI).
+All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model)<!-- and have been archived with Zenodo (INSERT DOI)-->.
 
@@
 
-*Running title*: Character divergence destabilizes food webs
+*Running title*: Character displacement destabilizes food webs
 
@@
 
-<!-- Remove rest of title page for bioRxiv submission -->
+<!-- Remove rest of title page for bioRxiv submission
 \bigskip
@@
 Ecological character displacement is an adaptive process that generally increases phenotypic diversity.
-Despite the fact that this diversification is due to an eco-evolutionary feedback between consumers competing for shared resources, its consequences for food-web dynamics have not been examined.
+Despite the fact that this diversification is due to an eco-evolutionary feedback between consumers competing for shared resources, its consequences for food-web dynamics have received little attention.
 Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumer attack rates affects resource abundances and the resilience of food webs to perturbations.
@@
 
-Although models that included resources led to insights to the evolution of character displacement, the ecological feedback onto consumer-resource dynamics has received surprisingly little attention. 
+Although models that included resources led to insights about the evolution of character displacement, the ecological feedback onto consumer-resource dynamics has received surprisingly little attention.
 This is likely because the ecological feedback has been primarily studied through the lens of coexistence theory [@Lawlor1976; @Germain2018; @Bassar2017; @McPeek2019]. 
-For example, early theoretical work showed that ecological character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition [@Lawlor1976].
+For example, early theoretical work showed that character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition [@Lawlor1976]. 
 Yet, this reduction in interspecific competition may, at the same time, increase interspecific interactions between specialized consumers and their resources. 
@@
 Here, I address this knowledge gap by studying a mathematical model that examines how ecological character displacement affects consumer-resource dynamics in a food-web context.
-I address two questions: (1) How does ecological character displacement affect resource abundances? (2) How does character displacement affect food-web stability? 
+Specifically, I sought to answer the question: how does character displacement in consumer attack rates affect resource abundances and food-web stability?
 To test the generality of these effects, I explored different ecological foraging scenarios and evolutionary tradeoffs in consumer attack rates. 
@@
 
-## Underlying Consumer-Resource Dynamics
+## Underlying consumer-resource dynamics
 
@@
 where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ij}$ represents the conversion efficiency of resource $i$ into consumer $j$, and $m_j$ represents the mortality rate of consumer $j$.
-$F_{ij}(R_i)$ represents consumer $j$'s feeding rate on resource $i$ (i.e.
-functional response).
+$F_{ij}(R_i)$ represents consumer $j$'s feeding rate on resource $i$ (i.e., its functional response).
 This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g.
-zooplankton and benthic invertebrates in lakes) rather than a scenario where resources are better characterized by a continuous trait distribution (e.g.
-seed size, see @Taper1985 for an example).
-Importantly, inferences about ecological character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992].
+zooplankton and benthic invertebrates in lakes) rather than a scenario where resources are better characterized by a continuous trait distribution (e.g., seed size; see @Taper1985 for an example).
+Importantly, inferences about character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992].
 Therefore, I arbitrarily set $C_2=0$ to create a food-web without a competing consumer for these comparisons.
 
-## Foraging Scenarios
+## Foraging scenarios
 
 I studied three different foraging scenarios. 
-In the first, I assume that consumers can forage for both resources simultaneously (fig. \ref{fig:foraging_scenarios} top) and their feeding rates increase linearly with resource abundance, such that:
+In the first, I assume that consumers can forage for both resources simultaneously (Fig. \ref{fig:foraging_scenarios}a) and their feeding rate increases linearly with resource abundance, such that:
 
@@
 This first scenario is the starting point for many models of resource competition [@MacArthur1972]; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005].
-The second scenario accounts for this spatial context (fig. \ref{fig:foraging_scenarios} bottom) and takes the form:
+The second scenario accounts for this spatial context (Fig. \ref{fig:foraging_scenarios}b) and takes the form:
 
@@
 
-where $w_{ij}$ represents the proportion of time consumer $j$ spends foraging in a habitat where only resource $i$ is found (i.e. habitat preference). 
+where $w_{ij}$ represents the proportion of time consumer $j$ spends foraging in a habitat where only resource $i$ is found (i.e., its habitat preference). 
 Note that since $w_{ij}$ is a proportion that $w_{1,j}=1-w_{2,j}$. <!-- $\sum_jw_{ij}=1$.<!-- $w_{ij}=1-w_{ii}$. -->
@@
 
-where consumer $j$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling times $h_{ij}$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ij}=\frac{w_{ij}R_i}{w_{1,j}R_1+w_{2,j}R_2}$.
-
-Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach, with the general result being character divergence [@Lawlor1976; @Abrams1986].
-I also used an Adaptive Dynamics approach to analyze character displacement in the third foraging scenario, and I too observed character divergence (detailed analysis in Appendix S1 of Supplementary Information).
-I say consumers have undergone character divergence if their evolved attack rates are more specialized when evolving with vs. without a competing consumer. 
+where consumer $j$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling time $h_{ij}$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ij}=\frac{w_{ij}R_i}{w_{1,j}R_1+w_{2,j}R_2}$.
+
+Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach, with the general result being divergent character displacement [@Lawlor1976; @Abrams1986].
+I also used an Adaptive Dynamics approach to analyze the evolution of consumer attack rates in the third foraging scenario, and I too observed divergent character displacement (detailed analysis given in Appendix S1).
+I say consumers have undergone divergent character displacement if their evolved attack rates are more specialized when evolving with vs. without a competing consumer. 
 Specialization of consumer $j$ on resource $1$ is measured as $\frac{a_{1,j}}{a_{1,j}+a_{2,j}}$, where a value of 0.5 is a complete generalist ($a_{1,j}=a_{2,j}$), and a value of 1 is a complete specialist ($a_{2,j}=0$).
 Values less than 0.5 indicate specialization on the other resource. 
-
-```{r foraging_scenarios, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:foraging_scenarios}**Ecological foraging scenarios.** I examined whether the effect of ecological character displacement on food-web dynamics depended on whether consumers competed for resources that occurred in the same vs.
-different habitats.
-The width of each arrow corresponds to initial attack rates (*a*~ij~) of consumer *j* on resource *i*.
-Note that *C*~1~ was pre-adapted to *R*~1~ (*a*~11~ > *a*~21~), while *C*~2~ was a mirror image, being pre-adapted to *R~2~* (*a*~22~ > *a*~12~).
+Since I did not observe convergent character displacement in any of the foraging scenarios I analyzed, I refer to divergent character displacement as simply (ecological) character displacement throughout the rest of the text.
+
+```{r foraging_scenarios, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:foraging_scenarios}**Ecological foraging scenarios.** I examined whether the effect of ecological character displacement on food-web dynamics depended on whether consumers competed for resources that occurred in the same (a) vs.
+different habitats (b).
+Note that inferences about character displacement can only be made by comparing food webs with (right) and without (left) a competing consumer, so I arbitrarily set *C*~2~ = 0 for these comparisons.
+The width of each arrow corresponds to the initial attack rate (*a*~ij~) of consumer *j* on resource *i*.
+Note that *C*~1~ was pre-adapted to *R*~1~ (*a*~11~ > *a*~21~), while *C*~2~ was a mirror image, being pre-adapted to *R~2~* (*a*~22~ = *a*~11~; *a*~12~ = *a*~21~).
 In each scenario, I assumed consumer feeding rates increased linearly with resource abundance.
-I also relax this assumption and consider a more realistic functional response when resources occurred in different habitats."}
+I also relax this assumption and consider a more realistic functional response when resources occurred in different habitats (b)."}
 knitr::include_graphics("Fig_1_ForagingScenarios")
@@
 
-## Food-web Dynamics
-
-Given that character divergence occurred across these foraging scenarios, I focus here on its consequences for food-web dynamics.
+## Food-web dynamics
+
+Given that character displacement occurred across these foraging scenarios, I focus here on its consequences for food-web dynamics.
 To do this, I analyzed differences in resource abundances and food-web stability at equilibrium. 
-An equilibrium is reached when the rates of change in equation \ref{eq:1} are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_i$) and each consumer ($\hat C_j$).
+An equilibrium is reached when there is no change in the population growth rates of consumers and resources (i.e., the rates of change in equation \ref{eq:1} are 0), and solving the system at this point gives equilibrium abundances for each resource ($\hat R_i$) and each consumer ($\hat C_j$).
 I also compared the local stability of these food webs using standard methods [@Otto2007]. 
-This stability analysis derives the dominant eigenvalue, $\lambda_{\text{max}}$, of the matrix of partial derivatives of all differential equations with respect to all variables evaluated at equilibrium.
-If $-\lambda_{max}>0$, then the food web will return to equilibrium after a small perturbation (i.e. locally stable), with more positive values indicating a faster return time.
+This stability analysis derives the dominant eigenvalue, $\lambda_{\text{max}}$, of the matrix of partial derivatives of each species' population growth rate (given by equation \ref{eq:1}) with respect to each species' abundance evaluated at equilibrium.
+If $-\lambda_{max}>0$, then the food web will return to equilibrium after a small perturbation (i.e., it is locally stable), with more positive values indicating a faster return time.
 If $-\lambda_{max}<0$, then the food web is not locally stable.
@@
 Note that I arbitrarily set $C_1$ as being pre-adapted to $R_1$ ($a_{11} > a_{21}$; $w_{11} > 0.5$), and therefore $C_2$ was pre-adapted to $R_2$.
-Controlling for other sources of variability allowed me to isolate the general effects of character divergence.
-All mathematical derivations were conducted in Mathematica [@Mathematica] and are provided in the Supplementary Information (Appendix S1-3).
+Controlling for other sources of variability allowed me to isolate the general effects of character displacement.
+All mathematical derivations were conducted in Mathematica [@Mathematica] and are provided in Appendices S1-3.
 
@@
 The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006].
-This function has the useful property that it differentiates between cases where intermediate combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ are higher than the extremes (when $n>1$, green line in fig.
-\ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line fig.
+This function has the useful property that it differentiates between cases where intermediate combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ are higher than the extremes (when $n>1$, green line in Fig.
+\ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line in Fig.
 \ref{fig:tradeoff}).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ have the same total attack rate (blue line in fig. \ref{fig:tradeoff}). 
+When $n=1$, the tradeoff function is linear, and all combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ have the same total attack rate (blue line in Fig. \ref{fig:tradeoff}). 
 Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
 If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource abundances to reach a steady state. 
-I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g.
-$\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0).
+I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g., $\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0).
 Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and $w_{11} = w_{22} = 0.6$. 
@@
 I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
-All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model) and have been archived with Zenodo (INSERT DOI).
+All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model)<!-- and has been archived with Zenodo (INSERT DOI)-->.
 
@@
 ```{r tradeoff, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: intermediate combinations of attack rates (*a*~1,*j*~, *a*~2,*j*~) are higher than the extremes (green line, *n* > 1); extreme combinations of attack rates are higher than intermediate investments (orange line, *n* < 1); and all combinations of attack rates have the same total attack rate (blue line, *n* = 1).
-Points corresponding to attack rates at the beginning of the simulation for *C*~1~, which was pre-adapted to *R*~1~ (*a*~11~ > *a*~12~)."}
+Points corresponding to attack rates at the beginning of the simulation for *C*~1~, which was pre-adapted to *R*~1~ (*a*~11~ > *a*~12~).
+Note that *C*~2~ was a mirror image of *C*~1~, being pre-adapted to *R*~2~ (*a*~22~ = *a*~11~; *a*~12~ = *a*~21~)."}
 knitr::include_graphics("Fig_2_Tradeoffs.pdf")
@@
 
-## Resources occur in same habitat
+## Resources occur in the same habitat
 
@@
 
-I found that the shape of the tradeoff function qualitatively affects the relationship between character displacement and resource abundances in this scenario (fig. \ref{fig:plot_fig3} middle left).
-For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (fig.
-\ref{fig:plot_fig3} top left) and character displacement has no effect on resource abundances (fig.
-\ref{fig:plot_fig3} middle left).
-If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (fig. \ref{fig:plot_fig3} middle left). 
-This is because the total attack rate of consumers is maximized at intermediate values ($a_{1,j}=a_{2,j}$) and decreases as consumers diverge (fig. \ref{fig:plot_fig3} top left). 
-When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (fig. \ref{fig:plot_fig3} top left). 
-Although the equation I derived for resource abundances was for the scenario where both consumers and both resources were present, it accurately predicts the abundance of resources when a single consumer reaches its evolutionary stable strategy (ESS; triangles on respective colored lines in fig.
-\ref{fig:plot_fig3} middle left).
-This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in fig.
-\ref{fig:plot_fig3} left panels), resulting in equivalent resource abundances.
+I found that the shape of the tradeoff function qualitatively affects the relationship between character displacement and resource abundances in this scenario (Fig. \ref{fig:plot_fig3}a,b).
+For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (Fig.
+\ref{fig:plot_fig3}a) and character displacement has no effect on resource abundances (Fig.
+\ref{fig:plot_fig3}b).
+If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (Fig. \ref{fig:plot_fig3}b). 
+This is because the total attack rate of consumers is maximized at intermediate values ($a_{1,j}=a_{2,j}$) and decreases as consumers diverge (Fig. \ref{fig:plot_fig3}a). 
+When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (Fig. \ref{fig:plot_fig3}a,b). 
+Although the equation I derived for resource abundances was for the scenario where both consumers and both resources were present, it accurately predicts the abundance of resources when a single consumer reaches its evolutionary stable strategy (ESS; triangles on respective colored lines in Fig.
+\ref{fig:plot_fig3}b).
+This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in Fig.
+\ref{fig:plot_fig3}a), resulting in equivalent resource abundances.
 
 The effect of character displacement on resources corresponds to its impact on food-web stability. 
-For example, when character displacement decreases resource abundances (orange points in fig.
-\ref{fig:plot_fig3} middle left), there is a decrease in food-web stability (fig.
-\ref{fig:plot_fig3} bottom left).
-Character divergence may not affect or even increase food-web stability (blue and green lines in fig.
-\ref{fig:plot_fig3} bottom left); however, evolution does not favor strong divergence in these scenarios (blue and green points in fig.
-\ref{fig:plot_fig3}), which dampens these contingent effects.
-Note that the dip in stability occurs when both consumers evolve to be generalists, a situation that is not favored in any of the foraging scenarios we examined (fig. \ref{fig:plot_fig3} bottom).
+For example, when character displacement decreases resource abundances (orange points in Fig. \ref{fig:plot_fig3}b), there is also a decrease in food-web stability (Fig. \ref{fig:plot_fig3}c). 
+Character displacement may not affect or even increase food-web stability (blue and green lines in Fig.
+\ref{fig:plot_fig3}c); however, evolution does not favor strong divergence in these scenarios (blue and green points in Fig.
+\ref{fig:plot_fig3}a), which dampens these contingent effects.
+Note that the dip in stability occurs when both consumers evolve to be generalists, a situation that is not favored in any of the foraging scenarios I examined (Fig. \ref{fig:plot_fig3}c).
 
@@
 
-This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g.
-$w_{1,j}>0.5$ and $a_{1,j}>a_{2,j}$), then the effective attack rate of consumers ($w_{1,j}a_{1,j}+w_{2,j}a_{2,j}$) will always increase, regardless of the tradeoff (fig.
-\ref{fig:plot_fig3} top right).
-Thus, character divergence always results in resource suppression (fig. \ref{fig:plot_fig3} middle right). 
+This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g., $w_{1,j}>0.5$ and $a_{1,j}>a_{2,j}$), then the effective attack rate of consumers ($w_{1,j}a_{1,j}+w_{2,j}a_{2,j}$) will always increase, regardless of the tradeoff (Fig.
+\ref{fig:plot_fig3}d).
+Thus, character displacement always results in resource suppression (Fig. \ref{fig:plot_fig3}e). 
 Note that the shape of the tradeoff can modify the effect of character displacement. 
-This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached an ESS (triangles in fig.
-\ref{fig:plot_fig3} middle right).
-In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in fig.
-\ref{fig:plot_fig3} middle right), because character displacement tends to reach a constraint of complete specialization.
-It is worth noting that resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in fig.
-\ref{fig:plot_fig3} middle right).
-This is because consumers actually evolve to be slightly specialized on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in fig.
-\ref{fig:plot_fig3} right panels).
-
-As seen previously, the effect of character displacement on resource abundances qualitatively corresponds to its effect on food-web stability (fig. \ref{fig:plot_fig3}). 
-Specifically, character divergence decreases food-web stability, regardless of the tradeoff in attack rates.
-This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (fig.
-\ref{fig:plot_fig3}).
+This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached an ESS (triangles in Fig.
+\ref{fig:plot_fig3}e).
+In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
+\ref{fig:plot_fig3}e), because character displacement tends to reach a constraint of complete specialization.
+It is worth noting that resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
+\ref{fig:plot_fig3}e).
+This is because consumers actually evolve slightly specialized attack rates on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in Fig.
+\ref{fig:plot_fig3}d).
+
+As seen previously, the effect of character displacement on resource abundances qualitatively corresponds to its effect on food-web stability (Fig. \ref{fig:plot_fig3}f). 
+Specifically, character displacement decreases food-web stability, regardless of the tradeoff in attack rates.
+This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (Fig.
+\ref{fig:plot_fig3}c).
 For example, when the tradeoff is concave up (orange), the initial two-consumer food web (small circle) is more stable than when there is only one consumer (small triangle); however, this pattern switches by the end of the eco-evolutionary simulation (large points).
@@
 Macs <- plot_grid(
-  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)),
+  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) +
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), 
   Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)),
   Stability_Mac + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)),
-  ncol=1, align='v'
+  ncol=1, align='v', labels = c("","b)","c)"), hjust = -4, vjust = 0.5, label_size = 10
 )
+Macs_title <- ggdraw() + draw_label("Resources: Same Habitat", fontface = 'bold', size = 10, hjust = 0.2)
+plot_Macs <- plot_grid(Macs_title, Macs, ncol = 1, rel_heights = c(0.1, 1))
 
 LSs <- plot_grid(
-  LS_attack  + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])) + ggtitle("Resources: Different Habitats\n") +  theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), #+ scale_y_continuous(breaks=c(0.8,1.0,1.2))  # "("*1-w[1,j]*")"
-  LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()), #+ scale_y_continuous(breaks=c(2.0,2.5,3.0), limits=c(2.0,3.2))
-  Stability_LS + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10)), # + scale_y_continuous(breaks=c(0,0.1,0.2))
-  ncol=1, align='v'
+  LS_attack  + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])) + 
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), 
+  LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()),   
+  Stability_LS + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10)), 
+  ncol=1, align='v', labels = c("","e)","f)"), hjust = -1.5, vjust = 0.5, label_size = 10
 )
-
-plot_fig3 <- plot_grid(Macs, LSs, ncol=2, rel_widths = c(1,0.9)) 
-plot_fig3_sub <- add_sub(plot_fig3, expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","* j]),")"))), vjust = 0, size=12)
+LSs_title <- ggdraw() + draw_label("Resources: Different Habitat", fontface = 'bold', size = 10, hjust = 0.3)
+plot_LSs <- plot_grid(LSs_title, LSs, ncol = 1, rel_heights = c(0.1, 1))
+
+plot_fig3 <- plot_grid(plot_Macs, plot_LSs, ncol=2, rel_widths = c(1,0.9)) 
+plot_fig3_label <- plot_fig3 + 
+  draw_label("a)", x = 0.098, y = 0.925, fontface = 'bold', size = 10) + 
+  draw_label("d)", x = 0.566, y = 0.925, fontface = 'bold', size = 10)
+plot_fig3_sub <- add_sub(plot_fig3_label, expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","* j]),")"))), vjust = 0, size=12)
 
@@
 Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~, whereas as small shapes correspond to the starting points (only in stability panels).
-In both foraging scenarios, feeding rates increase linearly with resource abundance, but the equation for effective attack rate is different."}
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~ (the choice to display *C*~1~ was arbitrary), whereas as small shapes correspond to the starting points (only in stability panels).
+In both foraging scenarios, feeding rates increase linearly with resource abundance, but the equation for the effective attack rate is different."}
 knitr::include_graphics("Fig_3_MacArthur_LawlorSmith.pdf")
@@
 
-Assuming consumers have a preferred resource (i.e. $w_{ii}>0.5$), then consumers will only reach an ESS when $|\delta_{ii}|<|\delta_{ij}|$ (e.g. green line, fig. \ref{MacArthur_Plot}A). 
+Assuming consumers have a preferred resource (i.e. $w_{ii}>0.5$), then consumers will only reach an ESS when $|\delta_{ii}|<|\delta_{ij}|$ (e.g. green line, Fig. \ref{MacArthur_Plot}A). 
 Otherwise, they will evolve to be complete specialists on their preferred resource.
@@
 
-In the third foraging scenario (equation \ref{eq:4}), I observed the same general effect of character divergence as the previous scenario (resources in different habitats, but linear functional response).
+In the third foraging scenario (equation \ref{eq:4}), I observed the same general effect of character displacement as the previous scenario (resources in different habitats, but linear functional response).
 This is because resource abundances at equilibrium are governed by a similar dynamic (derived in Appendix S1):
@@
 
-And since evolution favors consumer divergence onto their preferred resources (see Appendix S1), the effective attack rate of consumers ($w_{1,j}a_{1,j}+w_{2,j}a_{2,j}$) will always increase, resulting in lower resource abundances and decreased food-web stability (Appendix S4).
+And since evolution favors character displacement toward their preferred resources (see Appendix S1), the effective attack rate of consumers ($w_{1,j}a_{1,j}+w_{2,j}a_{2,j}$) will always increase, resulting in lower resource abundances and decreased food-web stability (Appendix S4, Fig.
+S1).
 
@@
 
-Thus, character displacement always pushes the food web toward an unstable structure in this more realistic foraging scenario (fig.
+This inequality indicates that character displacement always pushes the food web toward an unstable structure in this more realistic foraging scenario (Fig.
 \ref{fig:bifur_plot}).
-Note that I stopped the simulation in the four-species food web once it became locally unstable.
+Note that I stopped the simulation in the two-consumer food web once it became locally unstable.
 I do not simulate beyond this point as this would require making assumptions about the dynamics of mutant consumers in variable environments, which is beyond the scope of this work.
@@
   McCann_resources    + theme(legend.position = "none") + xlab(""),
-  Stability_McCann    + theme(legend.position = "none") + xlab(""), ncol=2, align='hv')
+  Stability_McCann    + theme(legend.position = "none") + xlab(""), 
+  ncol=2, align='hv', labels = c("a)","b)","c)"), hjust = -4, vjust = 1, label_size = 10)
 
@@
 
-```{r bifur_plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Character divergence creates an unstable food web**.
-Lines illustrate the effects of character divergence across the range of specialization, while the points are the results of an eco-evolutionary simulation.
-Note that I increased the total investment in attack rates (*A* = 3.3) to create a scenario that could result in an unstable food web."}
+```{r bifur_plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Character displacement creates an unstable food web**.
+Lines illustrate the effect of character displacement across the range of specialization for *C*~1~ (the choice to display *C*~1~ was arbitrary), while the points are the results of an eco-evolutionary simulation.
+Note that I increased the total investment in attack rates (*A* = 3.3) to create a scenario that could result in an unstable food web.
+Although I specified a linear tradeoff in attack rates for this simulation, different tradeoff shapes do not qualitatively alter these results (see Appendix S4, Fig.
+S1)."}
 knitr::include_graphics("Fig_4_Bifurcation.pdf")
@@
 
-The previous analytical results and simulations make a strong assumption that competing consumers start off as perfect mirror images of each other (i.e. symmetry). 
+The previous analytical results and simulations make a strong assumption that competing consumers start off as perfect mirror images of each other (i.e., there is symmetry). 
 Yet, theory indicates a predictable asymmetry between initial consumer attack rates. 
-This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, reaches an ESS as a generalist, followed by the invasion of a second, more specialized, consumer.
+This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, evolves to be a generalist that can equally attack both resources, followed by the invasion of a second, more specialized, consumer.
 This theoretical scenario has been hypothesized as the sequence of events leading to character displacement in threespine stickleback in small coastal lakes of British Columbia [@Schluter1992; @Schluter2000].
@@
 I did this for all foraging scenarios and tradeoffs previously examined. 
-I found that my previous inferences are robust to including consumer asymmetry across different foraging scenarios and tradeoffs (Appendix S4).
+I found that my previous inferences are robust to including consumer asymmetry across different foraging scenarios and tradeoffs (Appendix S4, Fig. S2-3).
 
@@
 
-## Resource Abundances
-
-One of the criteria used to demonstrate character displacement is that "sites of sympatry [two consumers] and allopatry [one consumer] should not differ greatly in food, ..." [@Schluter1992].
-In contrast, my results suggest that ecological character displacement causes predictable differences in resource abundances.
-In fact, the ecological and evolutionary scenarios that favored the greatest character divergence always decreased the abundance of resources. <!-- build paragraph around this topic sentence -->
-For example, the magnitude of character displacement was highest when mobile consumers competed for resources that occur in different habitats.
+## Resource abundances
+
+One of the criteria used to demonstrate ecological character displacement is that ``sites of sympatry [two consumers] and allopatry [one consumer] should not differ greatly in food $\textbf{[}$resource abundances$\textbf{]}$'' [@Schluter1992].
+In contrast, my results indicate that character displacement causes predictable differences in resource abundances.
+In fact, the ecological and evolutionary scenarios that favored the largest character displacement always decreased the relative abundance of resources.
+<!-- build paragraph around this topic sentence -->
+For example, if mobile consumers compete for resources that occur in different habitats, then character displacement always resulted in lower resource abundances.<!--the magnitude of character displacement was highest when mobile consumers competed for resources that occur in different habitats.-->
+Threespine stickleback, one of the classic examples of character displacement, exemplify this foraging scenario [@Schluter1992; @Schluter2000]. 
+Stickleback must move between the pelagic and littoral zones of a lake when foraging for zooplankton and benthic invertebrates, respectively.
+The theory developed here predicts that resource abundances will be lower in lakes where competing stickleback have undergone character displacement compared to lakes with only a single species of stickleback.
+Interestingly, a disproportionate number of the documented cases of character displacement involve carnivores [@Schluter2000] that are larger, and likely more mobile, than their resources [@McCann2005], suggesting that many cases of ecological character displacement may result in lower resource availability.
+
+<!-- older version: Threespine stickleback forage for two resource types, zooplankton and benthic invertebrates, which occur in distinct zones of the lake (pelagic and littoral)
 An empirical example of this is threespine stickleback that have diverged into limnetic and benthic species with specialized traits to forage on zooplankton and benthic invertebrates, respectively [@Schluter1992; @Schluter2000].
 These two resource types occur in distinct zones of the lake (pelagic and littoral), therefore stickleback must move between these zones when foraging for resources. 
-Interestingly, a disproportionate number of the documented cases of character displacement involve carnivores [@Schluter2000] that are larger, and likely more mobile, than their resources [@McCann2005], suggesting many cases of ecological character displacement may result in lower resource availability.
-
-Similarly, the evolutionary tradeoff that favored character divergence resulted in lower resources across all foraging scenarios.
+The theory developed here predicts that resource abundances will be lower in lakes with compe-->
+
+Similarly, the evolutionary tradeoff that favored character displacement decreased resource availability across all foraging scenarios.
 Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the net foraging rate of consumers [@Schluter1985; @Arnegard2014].
@@
 
-## Food-web Stability
-
-My most striking result was that character divergence made food webs less resilient to perturbations. 
-In fact, under the most realistic foraging scenario, character divergence can even result in an unstable food web. 
+## Food-web stability
+
+My most striking result was that ecological character displacement made food webs less resilient to perturbations. 
+In fact, under the most realistic foraging scenario, character displacement can even result in an unstable food web. 
 The mechanism underlying this destabilization is quite general.
-Character divergence generally increases the strength of consumer-resource interactions, but does not alter the strength of intraspecific interactions.
+Character displacement generally increases the strength of consumer-resource interactions, but does not alter the strength of intraspecific interactions.
 This relative increase in interspecific interactions, combined with the natural oscillatory tendency of consumer-resource dynamics [@Lotka1925; @Volterra1926], creates a food-web structure that is less resilient to perturbations [@Chesson2008; @Rip2011; @McCann2011].
 
-Interestingly, the ecological conditions that favor character divergence are those that are already the least resilient to perturbations.
-For example, @McPeek2019 showed that character divergence is favored in food webs that are either highly productive, easy to find and capture resources, or under weak abiotic stress.
+Interestingly, the ecological conditions that favor character displacement are those that are already the least resilient to perturbations.
+For example, @McPeek2019 showed that character displacement is favored in food webs that are either highly productive, easy to find and capture resources, or under weak abiotic stress.
 This corresponds to higher values of $K$ (productivity) or $A$ (investment in attack rates), or lower values of $m$ (abiotic stress).
@@
 Similarly, higher feeding rates or lower consumer mortality both increase the relative strength of consumer-resource interactions, which predictably destabilizes food webs [@Rip2011; @McCann2011].
-This suggests that the most dramatic examples of character divergence will not only occur in, but also cause, the least stable food-web structures.
-
-A handful of empirical patterns support the hypothesis that character divergence decreases food-web resilience.
-For example, a single species of threespine stickleback lives in hundreds of small coastal lakes of British Columbia, but the species pair, where character divergence has resulted in specialized limnetic and benthic species, are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
+This suggests that the most dramatic examples of character displacement will not only occur in, but also cause, the least stable food-web structures.
+
+A handful of empirical patterns support the hypothesis that character displacement decreases food-web resilience.
+For example, a single species of threespine stickleback lives in hundreds of small coastal lakes in British Columbia, but the species pair, where character displacement has resulted in specialized limnetic and benthic species, are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
 Perhaps many lakes had a species pair in the past, but have lost a species due to a less resilient food-web structure [@Borrelli2015a; @Borrelli2015b].
@@
 My results contrast, but do not necessarily contradict, the notion from coexistence theory that character displacement contributes to species coexistence [@Lawlor1976]. 
-Rather than studying resilience, coexistence theory usually studies the mutual ability of consumers with different phenotypes to invade when rare [mutual invasibility, @Chesson2000].
+Rather than studying resilience, coexistence theory usually studies the mutual ability of consumers with different phenotypes to invade when rare [i.e., mutual invasibility; @Chesson2000].
 In the context of character displacement, a shortcoming of this mutual invasibility measure is that it does not allow a comparison between food webs with and without a competing consumer. 
@@
 Note that this does not negate my general conclusion that ecological character displacement decreases resource abundances and stability; however, this process may itself create another eco-evolutionary feedback between consumers and resources.
-This may actually help maintain dramatic examples of character divergence and prevent them from destabilizing systems, because it allows consumer traits to become decoupled from their attack rate.
-Examining this decoupling would be ideal in a quantitative genetic model that can explicitly track trait dynamics, but it would not fundamentally change the conclusions presented here.
+This may actually help maintain dramatic examples of character displacement and prevent them from destabilizing systems, because it allows consumer traits to become decoupled from their attack rate.
+Examining this decoupling would be ideal in a quantitative genetic model that explicitly tracks trait dynamics, but it would not fundamentally change the conclusions presented here.
 
@@
 Prior work has shown that allowing for resource asymmetry, for example, may decrease the magnitude of character displacement [@Abrams1986]. 
-This is because this asymmetry creates an asymmetry in resource abundances, which dampens the effects of character divergence.
-Again though, this may dampen the effects of divergence, but should not change the qualitative relationship we observed of ecological character displacement decreasing resource abundances and food-web stability.
+<!--This is because this asymmetry creates an asymmetry in resource abundances, which dampens the amount of character displacement.-->
+While this may dampen the amount of divergence, it should not qualitatively change the relationship I observed.<!-- of ecological character displacement decreasing resource abundances and food-web stability.-->
 
 I studied this eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
-A strength of this approach is that it enabled me to gain analytical insight to the effects of character divergence in a more realistic foraging scenario.
+A strength of this approach is that it enabled me to gain analytical insight to the effects of character displacement in a more realistic foraging scenario.
 This is much less tractable in quantitative genetic [@Taper1985; @McPeek2017] or explicit genetic [@Doebeli1996] models of character displacement, which is why the foraging scenarios previously examined have been limited [but see @McPeek2017].
@@
 That being said, my theoretical predictions are likely robust to these assumptions. 
-This is because models that explicitly include resource dynamics inevitably show that resource competition results in character divergence, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
-A quantitative genetic model may certainly show differences in the pace of character divergence, but this should not qualitatively change its effect on food-web dynamics. 
+This is because models that explicitly include resource dynamics inevitably show that resource competition results in character displacement, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
+A quantitative genetic model may certainly show differences in the pace of character displacement, but this should not qualitatively change its effect on food-web dynamics. 
 It is important to note that my conclusions only apply to food webs with biotic resources that are nutritionally substitutable. 
-It would be interesting to extend these current analyses to non-substitutable resources where character convergence is expected [@Abrams1987; @Fox2008].
+It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
 
@@
 Macs_asymm <- plot_grid(
-  Mac_attack_asymm + ggtitle("Resources: Same Habitat\n") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))),
+  Mac_attack_asymm + #ggtitle("Resources: Same Habitat\n") + 
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))),
   Mac_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))),
   Stability_Mac_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))),
-  ncol=1, align='v'
+  ncol=1, align='v', labels = c("","b)","c)"), hjust = -4, vjust = 0.5, label_size = 10
 )
+Macs_asymm_title <- ggdraw() + draw_label("Resources: Same Habitat", fontface = 'bold', size = 10, hjust = 0.2)
+plot_Macs_asymm <- plot_grid(Macs_asymm_title, Macs_asymm, ncol = 1, rel_heights = c(0.1, 1))
 
 LSs_asymm <- plot_grid(
-  LS_attack_asymm  + ggtitle("Resources: Different Habitats\n") +  theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)) + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])),
+  LS_attack_asymm  + #ggtitle("Resources: Different Habitats\n") +  
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)) + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])),
   LS_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x=element_blank()) + xlab("") + ylab(""),
   Stability_LS_asymm   + theme(legend.position = "none", axis.text = element_text(size=10)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(""),
-  ncol=1, align='v'
+  ncol=1, align='v', labels = c("","e)","f)"), hjust = -1.5, vjust = 0.5, label_size = 10
 )
-
-all_asymm_plots <- plot_grid(Macs_asymm, LSs_asymm, ncol=2, rel_widths = c(1,0.9)) 
-all_asymm_plots <- add_sub(all_asymm_plots, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
+LSs_asymm_title <- ggdraw() + draw_label("Resources: Different Habitat", fontface = 'bold', size = 10, hjust = 0.3)
+plot_LSs_asymm <- plot_grid(LSs_asymm_title, LSs_asymm, ncol = 1, rel_heights = c(0.1, 1))
+
+all_asymm_temp <- plot_grid(plot_Macs_asymm, plot_LSs_asymm, ncol=2, rel_widths = c(1,0.9)) 
+all_asymm_labels <- all_asymm_temp +
+  draw_label("a)", x = 0.098, y = 0.925, fontface = 'bold', size = 10) + 
+  draw_label("d)", x = 0.566, y = 0.925, fontface = 'bold', size = 10)
+all_asymm_plots <- add_sub(all_asymm_labels, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
 
@@
                              Stability_McCann_asymm   + theme(legend.position = "none") + xlab(""), 
-                       ncol=2, align = 'hv')
+                       ncol=2, align = 'hv', labels = c("a)","b)","c)"), hjust = -4, vjust = 1, label_size = 10)
 McCann_plots_asymm

### 2019-11-18 manuscript_ECD_model.tex
incorporate final (minor) edits to main and supplementary text.

@@
 \usepackage{lineno}
-\usepackage{caption}
+\usepackage{caption,setspace}
 \captionsetup[figure]{font=small}
@@
 reproduce this manuscript and simulations within are publically
-available on GitHub (\url{https://github.com/mabarbour/ECD_model}) and
-have been archived with Zenodo (INSERT DOI).
+available on GitHub (\url{https://github.com/mabarbour/ECD_model}).
 
@@
 
-\emph{Running title}: Character divergence destabilizes food webs
+\emph{Running title}: Character displacement destabilizes food webs
 
@@
 consumer-resource interactions; adaptation; community stability
-
-\bigskip
-
-\emph{Manuscript type}: Letter.
-
-\bigskip
-
-\emph{Number of words in abstract}: 143; \emph{Number of words in main
-text}: 4086; \emph{Number of words in text boxes}: NA; \emph{Number of
-references}: 45; \emph{Number of figures}: 4; \emph{Number of tables}:
-0; \emph{Number of text boxes}: 0
 
@@
 competing for shared resources, its consequences for food-web dynamics
-have not been examined. Here, I study a model of two consumers competing
-for two shared resources to examine how character displacement in
-consumer attack rates affects resource abundances and the resilience of
-food webs to perturbations. I found that character displacement always
-strengthened consumer-resource interactions whenever consumers competed
-for resources that occurred in different habitats. This increase in
-interaction strength resulted in lower resource abundances and less
-resilient food webs. This occurred under different evolutionary
+have received little attention. Here, I study a model of two consumers
+competing for two shared resources to examine how character displacement
+in consumer attack rates affects resource abundances and the resilience
+of food webs to perturbations. I found that character displacement
+always strengthened consumer-resource interactions whenever consumers
+competed for resources that occurred in different habitats. This
+increase in interaction strength resulted in lower resource abundances
+and less resilient food webs. This occurred under different evolutionary
 tradeoffs and in both simple and more realistic foraging scenarios.
@@
 
-Although models that included resources led to insights to the evolution
-of character displacement, the ecological feedback onto
+Although models that included resources led to insights about the
+evolution of character displacement, the ecological feedback onto
 consumer-resource dynamics has received surprisingly little attention.
@@
 \citep{Lawlor1976, Germain2018, Bassar2017, McPeek2019}. For example,
-early theoretical work showed that ecological character displacement
-promotes coexistence by favoring specialized consumers that experience
-reduced interspecific competition \citep{Lawlor1976}. Yet, this
-reduction in interspecific competition may, at the same time, increase
-interspecific interactions between specialized consumers and their
-resources. Both food-web theory and empirical studies have shown that
-increasing the strength of consumer-resource interactions often
-suppresses the abundance of resources, which if sufficient enough, can
-generate oscillations and less stable consumer-resource dynamics
+early theoretical work showed that character displacement promotes
+coexistence by favoring specialized consumers that experience reduced
+interspecific competition \citep{Lawlor1976}. Yet, this reduction in
+interspecific competition may, at the same time, increase interspecific
+interactions between specialized consumers and their resources. Both
+food-web theory and empirical studies have shown that increasing the
+strength of consumer-resource interactions often suppresses the
+abundance of resources, which if sufficient enough, can generate
+oscillations and less stable consumer-resource dynamics
 \citep{Rosenzweig1971, Luckinbill1973, Murdoch2002, Murdoch2003, McCann2011}.
@@
 examines how ecological character displacement affects consumer-resource
-dynamics in a food-web context. I address two questions: (1) How does
-ecological character displacement affect resource abundances? (2) How
-does character displacement affect food-web stability? To test the
+dynamics in a food-web context. Specifically, I sought to answer the
+question: how does character displacement in consumer attack rates
+affect resource abundances and food-web stability? To test the
 generality of these effects, I explored different ecological foraging
@@
 
-\subsection{Underlying Consumer-Resource
-Dynamics}\label{underlying-consumer-resource-dynamics}
+\subsection{Underlying consumer-resource
+dynamics}\label{underlying-consumer-resource-dynamics}
 
@@
 \(F_{ij}(R_i)\) represents consumer \(j\)'s feeding rate on resource
-\(i\) (i.e.~functional response). This model is a useful
+\(i\) (i.e., its functional response). This model is a useful
 characterization of a scenario where consumers compete for two distinct
@@
 than a scenario where resources are better characterized by a continuous
-trait distribution (e.g.~seed size, see \citet{Taper1985} for an
-example). Importantly, inferences about ecological character
-displacement can only be made by comparing food webs with and without a
-competing consumer \citep{Schluter1992}. Therefore, I arbitrarily set
-\(C_2=0\) to create a food-web without a competing consumer for these
-comparisons.
+trait distribution (e.g., seed size; see \citet{Taper1985} for an
+example). Importantly, inferences about character displacement can only
+be made by comparing food webs with and without a competing consumer
+\citep{Schluter1992}. Therefore, I arbitrarily set \(C_2=0\) to create a
+food-web without a competing consumer for these comparisons.
 
-\subsection{Foraging Scenarios}\label{foraging-scenarios}
+\subsection{Foraging scenarios}\label{foraging-scenarios}
 
 I studied three different foraging scenarios. In the first, I assume
-that consumers can forage for both resources simultaneously (fig.
-\ref{fig:foraging_scenarios} top) and their feeding rates increase
-linearly with resource abundance, such that:
+that consumers can forage for both resources simultaneously (Fig.
+\ref{fig:foraging_scenarios}a) and their feeding rate increases linearly
+with resource abundance, such that:
 
@@
 resources that occur in different habitats \citep{McCann2005}. The
-second scenario accounts for this spatial context (fig.
-\ref{fig:foraging_scenarios} bottom) and takes the form:
+second scenario accounts for this spatial context (Fig.
+\ref{fig:foraging_scenarios}b) and takes the form:
 
@@
 where \(w_{ij}\) represents the proportion of time consumer \(j\) spends
-foraging in a habitat where only resource \(i\) is found (i.e.~habitat
-preference). Note that since \(w_{ij}\) is a proportion that
+foraging in a habitat where only resource \(i\) is found (i.e., its
+habitat preference). Note that since \(w_{ij}\) is a proportion that
 \(w_{1,j}=1-w_{2,j}\). Finally, it is well known that consumer feeding
@@
 the abundance of each resource; saturates as resource abundances
-increase (due to handling times \(h_{ij}\)); and consumer habitat
+increase (due to handling time \(h_{ij}\)); and consumer habitat
 preferences are modified by the relative abundance of resources, such
@@
 the first two foraging scenarios using an Adaptive Dynamics approach,
-with the general result being character divergence
+with the general result being divergent character displacement
 \citep{Lawlor1976, Abrams1986}. I also used an Adaptive Dynamics
-approach to analyze character displacement in the third foraging
-scenario, and I too observed character divergence (detailed analysis in
-Appendix S1 of Supplementary Information). I say consumers have
-undergone character divergence if their evolved attack rates are more
+approach to analyze the evolution of consumer attack rates in the third
+foraging scenario, and I too observed divergent character displacement
+(detailed analysis given in Appendix S1). I say consumers have undergone
+divergent character displacement if their evolved attack rates are more
 specialized when evolving with vs.~without a competing consumer.
@@
 specialist (\(a_{2,j}=0\)). Values less than 0.5 indicate specialization
-on the other resource.
+on the other resource. Since I did not observe convergent character
+displacement in any of the foraging scenarios I analyzed, I refer to
+divergent character displacement as simply (ecological) character
+displacement throughout the rest of the text.
 
@@
 displacement on food-web dynamics depended on whether consumers competed
-for resources that occurred in the same vs.~different habitats. The
-width of each arrow corresponds to initial attack rates
-(\emph{a}\textsubscript{ij}) of consumer \emph{j} on resource \emph{i}.
-Note that \emph{C}\textsubscript{1} was pre-adapted to
+for resources that occurred in the same (a) vs.~different habitats (b).
+Note that inferences about character displacement can only be made by
+comparing food webs with (right) and without (left) a competing
+consumer, so I arbitrarily set \emph{C}\textsubscript{2} = 0 for these
+comparisons. The width of each arrow corresponds to the initial attack
+rate (\emph{a}\textsubscript{ij}) of consumer \emph{j} on resource
+\emph{i}. Note that \emph{C}\textsubscript{1} was pre-adapted to
 \emph{R}\textsubscript{1} (\emph{a}\textsubscript{11} \textgreater{}
@@
 mirror image, being pre-adapted to \emph{R\textsubscript{2}}
-(\emph{a}\textsubscript{22} \textgreater{} \emph{a}\textsubscript{12}).
-In each scenario, I assumed consumer feeding rates increased linearly
-with resource abundance. I also relax this assumption and consider a
-more realistic functional response when resources occurred in different
-habitats.}
+(\emph{a}\textsubscript{22} = \emph{a}\textsubscript{11};
+\emph{a}\textsubscript{12} = \emph{a}\textsubscript{21}). In each
+scenario, I assumed consumer feeding rates increased linearly with
+resource abundance. I also relax this assumption and consider a more
+realistic functional response when resources occurred in different
+habitats (b).}
 \end{figure}
 
-\subsection{Food-web Dynamics}\label{food-web-dynamics}
+\subsection{Food-web dynamics}\label{food-web-dynamics}
 
-Given that character divergence occurred across these foraging
+Given that character displacement occurred across these foraging
 scenarios, I focus here on its consequences for food-web dynamics. To do
 this, I analyzed differences in resource abundances and food-web
-stability at equilibrium. An equilibrium is reached when the rates of
-change in equation \ref{eq:1} are 0, and solving the system at this
-point gives equilibrium abundances for each resource (\(\hat R_i\)) and
-each consumer (\(\hat C_j\)). I also compared the local stability of
-these food webs using standard methods \citep{Otto2007}. This stability
-analysis derives the dominant eigenvalue, \(\lambda_{\text{max}}\), of
-the matrix of partial derivatives of all differential equations with
-respect to all variables evaluated at equilibrium. If
-\(-\lambda_{max}>0\), then the food web will return to equilibrium after
-a small perturbation (i.e.~locally stable), with more positive values
-indicating a faster return time. If \(-\lambda_{max}<0\), then the food
-web is not locally stable.
+stability at equilibrium. An equilibrium is reached when there is no
+change in the population growth rates of consumers and resources (i.e.,
+the rates of change in equation \ref{eq:1} are 0), and solving the
+system at this point gives equilibrium abundances for each resource
+(\(\hat R_i\)) and each consumer (\(\hat C_j\)). I also compared the
+local stability of these food webs using standard methods
+\citep{Otto2007}. This stability analysis derives the dominant
+eigenvalue, \(\lambda_{\text{max}}\), of the matrix of partial
+derivatives of each species' population growth rate (given by equation
+\ref{eq:1}) with respect to each species' abundance evaluated at
+equilibrium. If \(-\lambda_{max}>0\), then the food web will return to
+equilibrium after a small perturbation (i.e., it is locally stable),
+with more positive values indicating a faster return time. If
+\(-\lambda_{max}<0\), then the food web is not locally stable.
 
@@
 pre-adapted to \(R_2\). Controlling for other sources of variability
-allowed me to isolate the general effects of character divergence. All
+allowed me to isolate the general effects of character displacement. All
 mathematical derivations were conducted in Mathematica
-\citep{Mathematica} and are provided in the Supplementary Information
-(Appendix S1-3).
+\citep{Mathematica} and are provided in Appendices S1-3.
 
@@
 combinations of \(a_{\text{1},j}\) and \(a_{\text{2},j}\) are higher
-than the extremes (when \(n>1\), green line in fig. \ref{fig:tradeoff})
+than the extremes (when \(n>1\), green line in Fig. \ref{fig:tradeoff})
 or, conversely, where the two extremes are higher than intermediate
-investments (when \(n<1\), orange line fig. \ref{fig:tradeoff}). When
+investments (when \(n<1\), orange line in Fig. \ref{fig:tradeoff}). When
 \(n=1\), the tradeoff function is linear, and all combinations of
 \(a_{\text{1},j}\) and \(a_{\text{2},j}\) have the same total attack
-rate (blue line in fig. \ref{fig:tradeoff}). Assuming the mutant
+rate (blue line in Fig. \ref{fig:tradeoff}). Assuming the mutant
 consumer was rare, I then determined whether the mutant had higher
@@
 sufficient for consumers to either reach an evolutionary stable strategy
-\citep[ESS,][]{Smith1973} or an evolutionary limit (e.g.
+\citep[ESS,][]{Smith1973} or an evolutionary limit (e.g.,
 \(\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}\) is constrained to a
@@
 in R \citep{R} and the code to reproduce these simulations is publically
-available on GitHub (\url{https://github.com/mabarbour/ECD_model}) and
-have been archived with Zenodo (INSERT DOI).
+available on GitHub (\url{https://github.com/mabarbour/ECD_model}).
 
@@
 \emph{R}\textsubscript{1} (\emph{a}\textsubscript{11} \textgreater{}
-\emph{a}\textsubscript{12}).}
+\emph{a}\textsubscript{12}). Note that \emph{C}\textsubscript{2} was a
+mirror image of \emph{C}\textsubscript{1}, being pre-adapted to
+\emph{R}\textsubscript{2} (\emph{a}\textsubscript{22} =
+\emph{a}\textsubscript{11}; \emph{a}\textsubscript{12} =
+\emph{a}\textsubscript{21}).}
 \end{figure}
@@
 
-\subsection{Resources occur in same
-habitat}\label{resources-occur-in-same-habitat}
+\subsection{Resources occur in the same
+habitat}\label{resources-occur-in-the-same-habitat}
 
@@
 the relationship between character displacement and resource abundances
-in this scenario (fig. \ref{fig:plot_fig3} middle left). For example, if
+in this scenario (Fig. \ref{fig:plot_fig3}a,b). For example, if
 consumer's are constrained by a linear tradeoff (blue lines), then there
-is no net change in total attack rate (fig. \ref{fig:plot_fig3} top
-left) and character displacement has no effect on resource abundances
-(fig. \ref{fig:plot_fig3} middle left). If the tradeoff is concave down
-(green lines), then resource abundances can actually increase under
-character displacement (fig. \ref{fig:plot_fig3} middle left). This is
-because the total attack rate of consumers is maximized at intermediate
-values (\(a_{1,j}=a_{2,j}\)) and decreases as consumers diverge (fig.
-\ref{fig:plot_fig3} top left). When the tradeoff is concave up (orange
-lines), character displacement suppresses resource abundances due to the
-increase in total attack rates (fig. \ref{fig:plot_fig3} top left).
-Although the equation I derived for resource abundances was for the
-scenario where both consumers and both resources were present, it
-accurately predicts the abundance of resources when a single consumer
-reaches its evolutionary stable strategy (ESS; triangles on respective
-colored lines in fig. \ref{fig:plot_fig3} middle left). This is because
-a single consumer evolves to be a generalist that has equal attack rates
-on each resource (triangles at 0.5 along x-axis in fig.
-\ref{fig:plot_fig3} left panels), resulting in equivalent resource
-abundances.
+is no net change in total attack rate (Fig. \ref{fig:plot_fig3}a) and
+character displacement has no effect on resource abundances (Fig.
+\ref{fig:plot_fig3}b). If the tradeoff is concave down (green lines),
+then resource abundances can actually increase under character
+displacement (Fig. \ref{fig:plot_fig3}b). This is because the total
+attack rate of consumers is maximized at intermediate values
+(\(a_{1,j}=a_{2,j}\)) and decreases as consumers diverge (Fig.
+\ref{fig:plot_fig3}a). When the tradeoff is concave up (orange lines),
+character displacement suppresses resource abundances due to the
+increase in total attack rates (Fig. \ref{fig:plot_fig3}a,b). Although
+the equation I derived for resource abundances was for the scenario
+where both consumers and both resources were present, it accurately
+predicts the abundance of resources when a single consumer reaches its
+evolutionary stable strategy (ESS; triangles on respective colored lines
+in Fig. \ref{fig:plot_fig3}b). This is because a single consumer evolves
+to be a generalist that has equal attack rates on each resource
+(triangles at 0.5 along x-axis in Fig. \ref{fig:plot_fig3}a), resulting
+in equivalent resource abundances.
 
@@
 impact on food-web stability. For example, when character displacement
-decreases resource abundances (orange points in fig. \ref{fig:plot_fig3}
-middle left), there is a decrease in food-web stability (fig.
-\ref{fig:plot_fig3} bottom left). Character divergence may not affect or
-even increase food-web stability (blue and green lines in fig.
-\ref{fig:plot_fig3} bottom left); however, evolution does not favor
-strong divergence in these scenarios (blue and green points in fig.
-\ref{fig:plot_fig3}), which dampens these contingent effects. Note that
+decreases resource abundances (orange points in Fig.
+\ref{fig:plot_fig3}b), there is also a decrease in food-web stability
+(Fig. \ref{fig:plot_fig3}c). Character displacement may not affect or
+even increase food-web stability (blue and green lines in Fig.
+\ref{fig:plot_fig3}c); however, evolution does not favor strong
+divergence in these scenarios (blue and green points in Fig.
+\ref{fig:plot_fig3}a), which dampens these contingent effects. Note that
 the dip in stability occurs when both consumers evolve to be
 generalists, a situation that is not favored in any of the foraging
-scenarios we examined (fig. \ref{fig:plot_fig3} bottom).
+scenarios I examined (Fig. \ref{fig:plot_fig3}c).
 
@@
 This equation implies that if consumers evolve to become specialists on
-resources that occur in their preferred habitat (e.g. \(w_{1,j}>0.5\)
+resources that occur in their preferred habitat (e.g., \(w_{1,j}>0.5\)
 and \(a_{1,j}>a_{2,j}\)), then the effective attack rate of consumers
 (\(w_{1,j}a_{1,j}+w_{2,j}a_{2,j}\)) will always increase, regardless of
-the tradeoff (fig. \ref{fig:plot_fig3} top right). Thus, character
-divergence always results in resource suppression (fig.
-\ref{fig:plot_fig3} middle right). Note that the shape of the tradeoff
-can modify the effect of character displacement. This is not so much due
-to the tradeoff affecting the magnitude of displacement (it does, but
-the effect is minor), but because the form of the tradeoff affects
-resource abundances when a single consumer has reached an ESS (triangles
-in fig. \ref{fig:plot_fig3} middle right). In contrast, resource
-abundances reach a similar value when consumers evolve in the presence
-of a competitor (circles in fig. \ref{fig:plot_fig3} middle right),
-because character displacement tends to reach a constraint of complete
-specialization. It is worth noting that resource abundances are
-consistently higher at the single consumer ESS compared to the
-predictions I derived for when both consumers are present (deviation of
-triangles from respective colored lines in fig. \ref{fig:plot_fig3}
-middle right). This is because consumers actually evolve to be slightly
-specialized on the resources that occur in their non-preferred habitat
-(deviation of triangles from 0.5 along x-axis in fig.
-\ref{fig:plot_fig3} right panels).
+the tradeoff (Fig. \ref{fig:plot_fig3}d). Thus, character displacement
+always results in resource suppression (Fig. \ref{fig:plot_fig3}e). Note
+that the shape of the tradeoff can modify the effect of character
+displacement. This is not so much due to the tradeoff affecting the
+magnitude of displacement (it does, but the effect is minor), but
+because the form of the tradeoff affects resource abundances when a
+single consumer has reached an ESS (triangles in Fig.
+\ref{fig:plot_fig3}e). In contrast, resource abundances reach a similar
+value when consumers evolve in the presence of a competitor (circles in
+Fig. \ref{fig:plot_fig3}e), because character displacement tends to
+reach a constraint of complete specialization. It is worth noting that
+resource abundances are consistently higher at the single consumer ESS
+compared to the predictions I derived for when both consumers are
+present (deviation of triangles from respective colored lines in Fig.
+\ref{fig:plot_fig3}e). This is because consumers actually evolve
+slightly specialized attack rates on the resources that occur in their
+non-preferred habitat (deviation of triangles from 0.5 along x-axis in
+Fig. \ref{fig:plot_fig3}d).
 
@@
 abundances qualitatively corresponds to its effect on food-web stability
-(fig. \ref{fig:plot_fig3}). Specifically, character divergence decreases
-food-web stability, regardless of the tradeoff in attack rates. This is
-not simply a consequence of having an additional consumer in the system,
-but emerges from the eco-evolutionary feedback between character
-displacement and resource suppression (fig. \ref{fig:plot_fig3}). For
-example, when the tradeoff is concave up (orange), the initial
-two-consumer food web (small circle) is more stable than when there is
-only one consumer (small triangle); however, this pattern switches by
-the end of the eco-evolutionary simulation (large points).
+(Fig. \ref{fig:plot_fig3}f). Specifically, character displacement
+decreases food-web stability, regardless of the tradeoff in attack
+rates. This is not simply a consequence of having an additional consumer
+in the system, but emerges from the eco-evolutionary feedback between
+character displacement and resource suppression (Fig.
+\ref{fig:plot_fig3}c). For example, when the tradeoff is concave up
+(orange), the initial two-consumer food web (small circle) is more
+stable than when there is only one consumer (small triangle); however,
+this pattern switches by the end of the eco-evolutionary simulation
+(large points).
 
@@
 (one consumer) correspond to the end points of the eco-evolutionary
-simulation for \emph{C}\textsubscript{1}, whereas as small shapes
+simulation for \emph{C}\textsubscript{1} (the choice to display
+\emph{C}\textsubscript{1} was arbitrary), whereas as small shapes
 correspond to the starting points (only in stability panels). In both
 foraging scenarios, feeding rates increase linearly with resource
-abundance, but the equation for effective attack rate is different.}
+abundance, but the equation for the effective attack rate is different.}
 \end{figure}
@@
 In the third foraging scenario (equation \ref{eq:4}), I observed the
-same general effect of character divergence as the previous scenario
+same general effect of character displacement as the previous scenario
 (resources in different habitats, but linear functional response). This
@@
 
-And since evolution favors consumer divergence onto their preferred
+And since evolution favors character displacement toward their preferred
 resources (see Appendix S1), the effective attack rate of consumers
 (\(w_{1,j}a_{1,j}+w_{2,j}a_{2,j}\)) will always increase, resulting in
-lower resource abundances and decreased food-web stability (Appendix
-S4).
+lower resource abundances and decreased food-web stability (Appendix S4,
+Fig. S1).
 
@@
 
-Thus, character displacement always pushes the food web toward an
-unstable structure in this more realistic foraging scenario (fig.
-\ref{fig:bifur_plot}). Note that I stopped the simulation in the
-four-species food web once it became locally unstable. I do not simulate
-beyond this point as this would require making assumptions about the
-dynamics of mutant consumers in variable environments, which is beyond
-the scope of this work.
+This inequality indicates that character displacement always pushes the
+food web toward an unstable structure in this more realistic foraging
+scenario (Fig. \ref{fig:bifur_plot}). Note that I stopped the simulation
+in the two-consumer food web once it became locally unstable. I do not
+simulate beyond this point as this would require making assumptions
+about the dynamics of mutant consumers in variable environments, which
+is beyond the scope of this work.
 
@@
 \includegraphics{Fig_4_Bifurcation.pdf}
-\caption{\label{fig:bifur_plot}\textbf{Character divergence creates an
-unstable food web}. Lines illustrate the effects of character divergence
-across the range of specialization, while the points are the results of
-an eco-evolutionary simulation. Note that I increased the total
-investment in attack rates (\emph{A} = 3.3) to create a scenario that
-could result in an unstable food web.}
+\caption{\label{fig:bifur_plot}\textbf{Character displacement creates an
+unstable food web}. Lines illustrate the effect of character
+displacement across the range of specialization for
+\emph{C}\textsubscript{1} (the choice to display
+\emph{C}\textsubscript{1} was arbitrary), while the points are the
+results of an eco-evolutionary simulation. Note that I increased the
+total investment in attack rates (\emph{A} = 3.3) to create a scenario
+that could result in an unstable food web. Although I specified a linear
+tradeoff in attack rates for this simulation, different tradeoff shapes
+do not qualitatively alter these results (see Appendix S4, Fig. S1).}
 \end{figure}
@@
 that competing consumers start off as perfect mirror images of each
-other (i.e.~symmetry). Yet, theory indicates a predictable asymmetry
-between initial consumer attack rates. This predictable asymmetry
-emerges from a process of community assembly where a single consumer
-invades a system, reaches an ESS as a generalist, followed by the
-invasion of a second, more specialized, consumer. This theoretical
-scenario has been hypothesized as the sequence of events leading to
-character displacement in threespine stickleback in small coastal lakes
-of British Columbia \citep{Schluter1992, Schluter2000}.
+other (i.e., there is symmetry). Yet, theory indicates a predictable
+asymmetry between initial consumer attack rates. This predictable
+asymmetry emerges from a process of community assembly where a single
+consumer invades a system, evolves to be a generalist that can equally
+attack both resources, followed by the invasion of a second, more
+specialized, consumer. This theoretical scenario has been hypothesized
+as the sequence of events leading to character displacement in
+threespine stickleback in small coastal lakes of British Columbia
+\citep{Schluter1992, Schluter2000}.
 
@@
 previous inferences are robust to including consumer asymmetry across
-different foraging scenarios and tradeoffs (Appendix S4).
+different foraging scenarios and tradeoffs (Appendix S4, Fig. S2-3).
 
@@
 
-\subsection{Resource Abundances}\label{resource-abundances}
+\subsection{Resource abundances}\label{resource-abundances}
 
-One of the criteria used to demonstrate character displacement is that
-``sites of sympatry {[}two consumers{]} and allopatry {[}one consumer{]}
-should not differ greatly in food, \ldots{}'' \citep{Schluter1992}. In
-contrast, my results suggest that ecological character displacement
-causes predictable differences in resource abundances. In fact, the
-ecological and evolutionary scenarios that favored the greatest
-character divergence always decreased the abundance of resources. For
-example, the magnitude of character displacement was highest when mobile
-consumers competed for resources that occur in different habitats. An
-empirical example of this is threespine stickleback that have diverged
-into limnetic and benthic species with specialized traits to forage on
-zooplankton and benthic invertebrates, respectively
-\citep{Schluter1992, Schluter2000}. These two resource types occur in
-distinct zones of the lake (pelagic and littoral), therefore stickleback
-must move between these zones when foraging for resources.
-Interestingly, a disproportionate number of the documented cases of
-character displacement involve carnivores \citep{Schluter2000} that are
-larger, and likely more mobile, than their resources \citep{McCann2005},
-suggesting many cases of ecological character displacement may result in
-lower resource availability.
+One of the criteria used to demonstrate ecological character
+displacement is that ``sites of sympatry {[}two consumers{]} and
+allopatry {[}one consumer{]} should not differ greatly in food
+\(\textbf{[}\)resource abundances\(\textbf{]}\)'' \citep{Schluter1992}.
+In contrast, my results indicate that character displacement causes
+predictable differences in resource abundances. In fact, the ecological
+and evolutionary scenarios that favored the largest character
+displacement always decreased the relative abundance of resources. For
+example, if mobile consumers compete for resources that occur in
+different habitats, then character displacement always resulted in lower
+resource abundances. Threespine stickleback, one of the classic examples
+of character displacement, exemplify this foraging scenario
+\citep{Schluter1992, Schluter2000}. Stickleback must move between the
+pelagic and littoral zones of a lake when foraging for zooplankton and
+benthic invertebrates, respectively. The theory developed here predicts
+that resource abundances will be lower in lakes where competing
+stickleback have undergone character displacement compared to lakes with
+only a single species of stickleback. Interestingly, a disproportionate
+number of the documented cases of character displacement involve
+carnivores \citep{Schluter2000} that are larger, and likely more mobile,
+than their resources \citep{McCann2005}, suggesting that many cases of
+ecological character displacement may result in lower resource
+availability.
 
-Similarly, the evolutionary tradeoff that favored character divergence
-resulted in lower resources across all foraging scenarios. Although data
-on the shape of the tradeoff in consumer foraging traits is scarce, two
-classic examples of character displacement, Darwin's finches and
+Similarly, the evolutionary tradeoff that favored character displacement
+decreased resource availability across all foraging scenarios. Although
+data on the shape of the tradeoff in consumer foraging traits is scarce,
+two classic examples of character displacement, Darwin's finches and
 threespine stickleback, both appear to exhibit a tradeoff where extreme
@@
 
-\subsection{Food-web Stability}\label{food-web-stability}
+\subsection{Food-web stability}\label{food-web-stability}
 
-My most striking result was that character divergence made food webs
-less resilient to perturbations. In fact, under the most realistic
-foraging scenario, character divergence can even result in an unstable
-food web. The mechanism underlying this destabilization is quite
-general. Character divergence generally increases the strength of
-consumer-resource interactions, but does not alter the strength of
+My most striking result was that ecological character displacement made
+food webs less resilient to perturbations. In fact, under the most
+realistic foraging scenario, character displacement can even result in
+an unstable food web. The mechanism underlying this destabilization is
+quite general. Character displacement generally increases the strength
+of consumer-resource interactions, but does not alter the strength of
 intraspecific interactions. This relative increase in interspecific
@@
 
-Interestingly, the ecological conditions that favor character divergence
-are those that are already the least resilient to perturbations. For
-example, \citet{McPeek2019} showed that character divergence is favored
-in food webs that are either highly productive, easy to find and capture
-resources, or under weak abiotic stress. This corresponds to higher
-values of \(K\) (productivity) or \(A\) (investment in attack rates), or
-lower values of \(m\) (abiotic stress). Each of these corresponding
-changes decrease food-web resilience, as they increase the strength of
-consumer-resource interactions relative to intraspecific interactions.
-For example, increasing productivity reduces intraspecific competition
-in resource populations while increasing the flux of energy to
-consumers, resulting in the paradox of enrichment
+Interestingly, the ecological conditions that favor character
+displacement are those that are already the least resilient to
+perturbations. For example, \citet{McPeek2019} showed that character
+displacement is favored in food webs that are either highly productive,
+easy to find and capture resources, or under weak abiotic stress. This
+corresponds to higher values of \(K\) (productivity) or \(A\)
+(investment in attack rates), or lower values of \(m\) (abiotic stress).
+Each of these corresponding changes decrease food-web resilience, as
+they increase the strength of consumer-resource interactions relative to
+intraspecific interactions. For example, increasing productivity reduces
+intraspecific competition in resource populations while increasing the
+flux of energy to consumers, resulting in the paradox of enrichment
 \citep{Rosenzweig1971}. Similarly, higher feeding rates or lower
@@
 \citep{Rip2011, McCann2011}. This suggests that the most dramatic
-examples of character divergence will not only occur in, but also cause,
-the least stable food-web structures.
+examples of character displacement will not only occur in, but also
+cause, the least stable food-web structures.
 
 A handful of empirical patterns support the hypothesis that character
-divergence decreases food-web resilience. For example, a single species
-of threespine stickleback lives in hundreds of small coastal lakes of
-British Columbia, but the species pair, where character divergence has
-resulted in specialized limnetic and benthic species, are only known
-from six lakes in four independent watersheds
+displacement decreases food-web resilience. For example, a single
+species of threespine stickleback lives in hundreds of small coastal
+lakes in British Columbia, but the species pair, where character
+displacement has resulted in specialized limnetic and benthic species,
+are only known from six lakes in four independent watersheds
 \citep{Schluter1992, Schluter2000}. Perhaps many lakes had a species
@@
 coexistence theory usually studies the mutual ability of consumers with
-different phenotypes to invade when rare \citep[mutual
-invasibility,][]{Chesson2000}. In the context of character displacement,
+different phenotypes to invade when rare \citep[i.e., mutual
+invasibility;][]{Chesson2000}. In the context of character displacement,
 a shortcoming of this mutual invasibility measure is that it does not
@@
 another eco-evolutionary feedback between consumers and resources. This
-may actually help maintain dramatic examples of character divergence and
-prevent them from destabilizing systems, because it allows consumer
+may actually help maintain dramatic examples of character displacement
+and prevent them from destabilizing systems, because it allows consumer
 traits to become decoupled from their attack rate. Examining this
-decoupling would be ideal in a quantitative genetic model that can
-explicitly track trait dynamics, but it would not fundamentally change
+decoupling would be ideal in a quantitative genetic model that
+explicitly tracks trait dynamics, but it would not fundamentally change
 the conclusions presented here.
@@
 resource asymmetry, for example, may decrease the magnitude of character
-displacement \citep{Abrams1986}. This is because this asymmetry creates
-an asymmetry in resource abundances, which dampens the effects of
-character divergence. Again though, this may dampen the effects of
-divergence, but should not change the qualitative relationship we
-observed of ecological character displacement decreasing resource
-abundances and food-web stability.
+displacement \citep{Abrams1986}. While this may dampen the amount of
+divergence, it should not qualitatively change the relationship I
+observed.
 
@@
 it enabled me to gain analytical insight to the effects of character
-divergence in a more realistic foraging scenario. This is much less
+displacement in a more realistic foraging scenario. This is much less
 tractable in quantitative genetic \citep{Taper1985, McPeek2017} or
@@
 explicitly include resource dynamics inevitably show that resource
-competition results in character divergence, regardless of whether a
+competition results in character displacement, regardless of whether a
 quantitative genetic or Adaptive Dynamics approach is used
 \citep{Lawlor1976, Taper1985}. A quantitative genetic model may
-certainly show differences in the pace of character divergence, but this
-should not qualitatively change its effect on food-web dynamics. It is
-important to note that my conclusions only apply to food webs with
+certainly show differences in the pace of character displacement, but
+this should not qualitatively change its effect on food-web dynamics. It
+is important to note that my conclusions only apply to food webs with
 biotic resources that are nutritionally substitutable. It would be
 interesting to extend these current analyses to non-substitutable
-resources where character convergence is expected
+resources where convergent character displacement is expected
 \citep{Abrams1987, Fox2008}.

### 2019-11-18 manuscript_ECD_model.Rmd
added Zenodo DOI to manuscript.

@@
 *Data accessibility*: No new data were used.
-All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model)<!-- and have been archived with Zenodo (INSERT DOI)-->.
+All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model) and has been archived with Zenodo (https://zenodo.org/badge/latestdoi/79349192).
 
@@
 I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
-All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model)<!-- and has been archived with Zenodo (INSERT DOI)-->.
+All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model) and has been archived with Zenodo (https://zenodo.org/badge/latestdoi/79349192).
 

### 2019-11-18 manuscript_ECD_model.tex
added Zenodo DOI to manuscript.

@@
 reproduce this manuscript and simulations within are publically
-available on GitHub (\url{https://github.com/mabarbour/ECD_model}).
+available on GitHub (\url{https://github.com/mabarbour/ECD_model}) and
+has been archived with Zenodo
+(\url{https://zenodo.org/badge/latestdoi/79349192}).
 
@@
 in R \citep{R} and the code to reproduce these simulations is publically
-available on GitHub (\url{https://github.com/mabarbour/ECD_model}).
+available on GitHub (\url{https://github.com/mabarbour/ECD_model}) and
+has been archived with Zenodo
+(\url{https://zenodo.org/badge/latestdoi/79349192}).
 

### 2019-11-18 manuscript_ECD_model.Rmd
added info on title page and prevented figures from being embedded in main text to follow Ecology Letters submission guidelines.

@@
 
-<!-- Remove rest of title page for bioRxiv submission
+<!-- Remove rest of title page for bioRxiv submission-->
 \bigskip
@@
 
-```{r foraging_scenarios, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:foraging_scenarios}**Ecological foraging scenarios.** I examined whether the effect of ecological character displacement on food-web dynamics depended on whether consumers competed for resources that occurred in the same (a) vs.
+```{r foraging_scenarios, include=FALSE, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:foraging_scenarios}**Ecological foraging scenarios.** I examined whether the effect of ecological character displacement on food-web dynamics depended on whether consumers competed for resources that occurred in the same (a) vs.
 different habitats (b).
@@
 
-```{r tradeoff, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: intermediate combinations of attack rates (*a*~1,*j*~, *a*~2,*j*~) are higher than the extremes (green line, *n* > 1); extreme combinations of attack rates are higher than intermediate investments (orange line, *n* < 1); and all combinations of attack rates have the same total attack rate (blue line, *n* = 1).
+```{r tradeoff, include=FALSE, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: intermediate combinations of attack rates (*a*~1,*j*~, *a*~2,*j*~) are higher than the extremes (green line, *n* > 1); extreme combinations of attack rates are higher than intermediate investments (orange line, *n* < 1); and all combinations of attack rates have the same total attack rate (blue line, *n* = 1).
 Points corresponding to attack rates at the beginning of the simulation for *C*~1~, which was pre-adapted to *R*~1~ (*a*~11~ > *a*~12~).
@@
 
-```{r plot_fig3, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:plot_fig3}**Effect of character displacement on food-web dynamics under different evolutionary tradeoffs and foraging scenarios.** Lines show predicted values when both consumers and resources are present.
+```{r plot_fig3, include=FALSE, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:plot_fig3}**Effect of character displacement on food-web dynamics under different evolutionary tradeoffs and foraging scenarios.** Lines show predicted values when both consumers and resources are present.
 Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
@@
 
-```{r bifur_plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Character displacement creates an unstable food web**.
+```{r bifur_plot, include=FALSE, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Character displacement creates an unstable food web**.
 Lines illustrate the effect of character displacement across the range of specialization for *C*~1~ (the choice to display *C*~1~ was arbitrary), while the points are the results of an eco-evolutionary simulation.

### 2019-11-18 manuscript_ECD_model.tex
added info on title page and prevented figures from being embedded in main text to follow Ecology Letters submission guidelines.

@@
 consumer-resource interactions; adaptation; community stability
+
+\bigskip
+
+\emph{Manuscript type}: Letter.
+
+\bigskip
+
+\emph{Number of words in abstract}: 143; \emph{Number of words in main
+text}: 4151; \emph{Number of words in text boxes}: NA; \emph{Number of
+references}: 45; \emph{Number of figures}: 4; \emph{Number of tables}:
+0; \emph{Number of text boxes}: 0
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{Fig_1_ForagingScenarios}
-\caption{\label{fig:foraging_scenarios}\textbf{Ecological foraging
-scenarios.} I examined whether the effect of ecological character
-displacement on food-web dynamics depended on whether consumers competed
-for resources that occurred in the same (a) vs.~different habitats (b).
-Note that inferences about character displacement can only be made by
-comparing food webs with (right) and without (left) a competing
-consumer, so I arbitrarily set \emph{C}\textsubscript{2} = 0 for these
-comparisons. The width of each arrow corresponds to the initial attack
-rate (\emph{a}\textsubscript{ij}) of consumer \emph{j} on resource
-\emph{i}. Note that \emph{C}\textsubscript{1} was pre-adapted to
-\emph{R}\textsubscript{1} (\emph{a}\textsubscript{11} \textgreater{}
-\emph{a}\textsubscript{21}), while \emph{C}\textsubscript{2} was a
-mirror image, being pre-adapted to \emph{R\textsubscript{2}}
-(\emph{a}\textsubscript{22} = \emph{a}\textsubscript{11};
-\emph{a}\textsubscript{12} = \emph{a}\textsubscript{21}). In each
-scenario, I assumed consumer feeding rates increased linearly with
-resource abundance. I also relax this assumption and consider a more
-realistic functional response when resources occurred in different
-habitats (b).}
-\end{figure}
-
 \subsection{Food-web dynamics}\label{food-web-dynamics}
@@
 (\url{https://zenodo.org/badge/latestdoi/79349192}).
-
-\begin{figure}
-\centering
-\includegraphics{Fig_2_Tradeoffs.pdf}
-\caption{\label{fig:tradeoff}\textbf{Evolutionary tradeoffs in consumer
-attack rates.} In each foraging scenario, I explored the effects of
-three different tradeoffs: intermediate combinations of attack rates
-(\emph{a}\textsubscript{1,\emph{j}}, \emph{a}\textsubscript{2,\emph{j}})
-are higher than the extremes (green line, \emph{n} \textgreater{} 1);
-extreme combinations of attack rates are higher than intermediate
-investments (orange line, \emph{n} \textless{} 1); and all combinations
-of attack rates have the same total attack rate (blue line, \emph{n} =
-1). Points corresponding to attack rates at the beginning of the
-simulation for \emph{C}\textsubscript{1}, which was pre-adapted to
-\emph{R}\textsubscript{1} (\emph{a}\textsubscript{11} \textgreater{}
-\emph{a}\textsubscript{12}). Note that \emph{C}\textsubscript{2} was a
-mirror image of \emph{C}\textsubscript{1}, being pre-adapted to
-\emph{R}\textsubscript{2} (\emph{a}\textsubscript{22} =
-\emph{a}\textsubscript{11}; \emph{a}\textsubscript{12} =
-\emph{a}\textsubscript{21}).}
-\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{Fig_3_MacArthur_LawlorSmith.pdf}
-\caption{\label{fig:plot_fig3}\textbf{Effect of character displacement
-on food-web dynamics under different evolutionary tradeoffs and foraging
-scenarios.} Lines show predicted values when both consumers and
-resources are present. Different line colors correspond to different
-tradeoffs in attack rates (green, \emph{n} = 1.15; blue, \emph{n} = 1;
-orange, \emph{n} = 0.85). Large circles (two consumers) and triangles
-(one consumer) correspond to the end points of the eco-evolutionary
-simulation for \emph{C}\textsubscript{1} (the choice to display
-\emph{C}\textsubscript{1} was arbitrary), whereas as small shapes
-correspond to the starting points (only in stability panels). In both
-foraging scenarios, feeding rates increase linearly with resource
-abundance, but the equation for the effective attack rate is different.}
-\end{figure}
-
 \subsection{Adding a more realistic functional
@@
 is beyond the scope of this work.
-
-\begin{figure}
-\centering
-\includegraphics{Fig_4_Bifurcation.pdf}
-\caption{\label{fig:bifur_plot}\textbf{Character displacement creates an
-unstable food web}. Lines illustrate the effect of character
-displacement across the range of specialization for
-\emph{C}\textsubscript{1} (the choice to display
-\emph{C}\textsubscript{1} was arbitrary), while the points are the
-results of an eco-evolutionary simulation. Note that I increased the
-total investment in attack rates (\emph{A} = 3.3) to create a scenario
-that could result in an unstable food web. Although I specified a linear
-tradeoff in attack rates for this simulation, different tradeoff shapes
-do not qualitatively alter these results (see Appendix S4, Fig. S1).}
-\end{figure}
 

### 2020-01-30 Appendix_S4.Rmd
omitted personal identifying information for double-blind peer review at Am Nat

@@
   - \usepackage{fullpage}
-  - \linespread{1.5} 
+  - \linespread{1.7} 
   - \usepackage{lineno}
@@
 
+<!-- remove for double-blind peer review
 \begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}
@@
 \bigskip
+-->
 

### 2020-01-30 Figures.Rmd
omitted personal identifying information for double-blind peer review at Am Nat

--- 2019-11-18 manuscript_ECD_model.Rmd
@@
       keep_tex: true
-    toc: false
-    number_sections: false
-    df_print: paged
       fig_caption: true
-    citation_package: natbib 
+fontsize: 11pt
+fontfamily: mathpazo
+documentclass: article
 header-includes:
@@
   - \linespread{1.5} 
-  - \usepackage{lineno}
-  - \usepackage{caption,setspace}
+  #- \usepackage{lineno}
+  - \usepackage{caption} # ,setspace
   - \captionsetup[figure]{font=small}
-bibliography: ReferencesAbbrev
-biblio-style: EcolLett
-fontsize: 11pt
-fontfamily: mathpazo
-documentclass: article
-linkcolor: black
-urlcolor: black
-citecolor: black
 ---
 
-<!-- Begin title page formatting -->
-
-\vspace*{0.1cm} 
-\begin{center} \LARGE Ecological character displacement destabilizes food webs \end{center}
-
-\bigskip
-
-\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}
-
-\bigskip
-
-\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;
-
-\noindent 2. University of British Columbia, Department of Zoology, 2212 Main Mall, Vancouver, BC V6T 1Z4, Canada;
-
-$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch, phone: +41 44 635 61 63
-
-\bigskip
-
-*Statement of authorship*: MAB conceived the study, derived analytical solutions, performed simulations, and wrote the paper.
-
-\bigskip
-
-*Data accessibility*: No new data were used.
-All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model)<!-- and have been archived with Zenodo (INSERT DOI)-->.
-
-\bigskip
-
-*Running title*: Character displacement destabilizes food webs
-
-\bigskip
-
-*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptation; community stability
-
-<!-- Remove rest of title page for bioRxiv submission
-\bigskip
-
-*Manuscript type*: Letter.
-
-\bigskip
-
-*Number of words in abstract*: 143; 
-*Number of words in main text*: `r wordcountaddin:::word_count("manuscript_ECD_model.Rmd") - 143`;
-*Number of words in text boxes*: NA;
-*Number of references*: 45;
-*Number of figures*: 4;
-*Number of tables*: 0;
-*Number of text boxes*: 0
-
-<!-- End title page formatting -->
-
-\linenumbers{}
-\modulolinenumbers[3]
-
-\newpage
-
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE)
-
-## Load required libraries
-library(deSolve) # numerical integration library
-library(rootSolve) # for runsteady integration function
-library(pse) # for latin hypercube
-library(plyr) # for ldply
-library(dplyr) # for manipulating data
-library(tidyr) # for tidying data
-library(cowplot) # for better base, ggplot graphics
-
-## color palette
-cbPalette <- c("#D55E00", "#0072B2", "#009E73")
+knitr::opts_chunk$set(echo = TRUE)
 ```
-
-```{r Simulations, include=FALSE, messages=TRUE, cache=TRUE}
-
-#### GET REQUIRED FUNCTIONS ####
-
-source('identify_steady_state.R')
-source('eco_evo_sim_function.R')
-source('Consumer_Resource_Functions.R')
-source('CR_dynamics.R')
-
-#### GENERAL PARAMETERS AND STATE VARIABLES ####
-
-## Intrinsic growth rates
-r <- 1
-r1 <- r 
-r2 <- r
-
-## Carrying capacity
-K <- 4
-K1 <- K
-K2 <- K
-
-## Conversion efficiency
-e <- 0.8
-e11 <- e 
-e12 <- e 
-e21 <- e 
-e22 <- e
-
-## Mortality rate
-m <- 1
-m1 <- m
-m2 <- m
-
-## General, non-evolving parameter set for 2C,2R model
-general_parameters <- data.frame(r1 = r1, r2 = r2, K1 = K1, K2 = K2, e11 = e11, e12 = e12, e21 = e21, e22 = e22, m1 = m1, m2 = m2)
-
-## Attack rate
-A <- 2
-aii <- A*0.61
-
-## Habitat preference (for LawlorSmith and McCann models)
-w <- 0.6
-wii <- w
-wij <- (1 - w)
-
-## Handling time (only for McCann model)
-h <- 0.4
-h11 <- h
-h12 <- h
-h21 <- h
-h22 <- h
-
-## State variables
-R <- 2
-R1 <- R
-R2 <- R
-C <- 1
-C1 <- C
-C2 <- C
-
-states_2C_2R <- data.frame(R1 = R1, R2 = R2, C1 = C1, C2 = C2)
-states_1C_2R <- data.frame(R1 = R1, R2 = R2, C1 = C1)
-
-## Trade-offs. Borrowed from Sargent and Otto, 2006 Am Nat
-a12_tradeoff <- expression(A*(1-(a11m/A)^n)^(1/n)) 
-a21_tradeoff <- expression(A*(1-(a22m/A)^n)^(1/n))
-
-n_concaveDN <- 1.15
-n_linear <- 1
-n_concaveUP <- 0.85
-
-concaveDN_df <- data.frame(n = n_concaveDN, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_concaveDN)^(1/n_concaveDN))
-linear_df <- data.frame(n = n_linear, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_linear)^(1/n_linear))
-concaveUP_df <- data.frame(n = n_concaveUP, aii = seq(0,A,0.01)) %>% mutate(aij = A*(1-(aii/A)^n_concaveUP)^(1/n_concaveUP))
-
-bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
-  ggplot(., aes(x=aii, y=aij, color = factor(n))) + geom_line() + ggtitle("Attack rate trade-offs")
-
-bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + geom_line() + ggtitle("Relationship between specialization and total attack rate")
-
-######################################### MACARTHUR MODEL #######################################################
-
-## Mutant growth rates
-mC1_MacArthur <- expression(e11*a11m*R1 + e12*a12m*R2 - m1)
-mC2_MacArthur <- expression(e21*a21m*R1 + e22*a22m*R2 - m2)
-
-#### PARAMETERS ####
-
-## Linear
-params_2C_2R_MacArthur_linear <- data.frame(general_parameters,
-                                            a11 = aii, 
-                                            a12 = eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_linear)),                     
-                                            a21 = eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_linear)), 
-                                            a22 = aii)                      
-params_1C_2R_MacArthur_linear <- params_2C_2R_MacArthur_linear[c("r1","r2","K1","K2","e11","e12","a11","a12","m1")]
-
-## concaveDN
-params_2C_2R_MacArthur_concaveDN <- params_2C_2R_MacArthur_linear
-params_2C_2R_MacArthur_concaveDN["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveDN))
-params_2C_2R_MacArthur_concaveDN["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveDN))
-params_1C_2R_MacArthur_concaveDN <- params_2C_2R_MacArthur_concaveDN[c("r1","r2","K1","K2","e11","e12","a11","a12","m1")]
-
-## concaveUP
-params_2C_2R_MacArthur_concaveUP <- params_2C_2R_MacArthur_linear
-params_2C_2R_MacArthur_concaveUP["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveUP))
-params_2C_2R_MacArthur_concaveUP["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveUP))
-params_1C_2R_MacArthur_concaveUP <- params_2C_2R_MacArthur_concaveUP[c("r1","r2","K1","K2","e11","e12","a11","a12","m1")]
-
-#### SIMULATIONS ####
-
-## Linear
-sim_2C_2R_MacArthur_linear <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_linear, 
-                                init_states = states_2C_2R, 
-                                eco_CR_model = MacArthur_2C_2R, 
-                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_linear))
-sim_1C_2R_MacArthur_linear <- eco_evo_CR(init_parameters = params_1C_2R_MacArthur_linear, 
-                                init_states = states_1C_2R, 
-                                eco_CR_model = MacArthur_1C_2R, 
-                                AD_CR_models = c(mC1_MacArthur),
-                                species_traits = c(1), 
-                                eco_param_names = c("a11"), 
-                                eco_tradeoff_names = c("a12"),  
-                                evo_param_names = c("a11m"),  
-                                evo_tradeoff_names = c("a12m"), 
-                                tradeoff_functions = c(a12_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_linear))
-# asymmetric starting points
-params_2C_2R_MacArthur_linear_asymm <- params_2C_2R_MacArthur_linear
-params_2C_2R_MacArthur_linear_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_linear[nrow(sim_1C_2R_MacArthur_linear), c("a12","a11")] # replace C2's attack rates with the ending point of the C1 generalist simulation.
-This will create an asymmetry in the attack rates and follows the logic (and our results) that a likely scenerio with asymmetry is a specialist invading a generalist.
-# just altering the parameters, everything else stays the same
-sim_2C_2R_MacArthur_linear_asymm <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_linear_asymm,
-                                init_states = states_2C_2R, 
-                                eco_CR_model = MacArthur_2C_2R, 
-                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_linear))
-
-## concaveDN
-sim_2C_2R_MacArthur_concaveDN <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_concaveDN, 
-                                init_states = states_2C_2R, 
-                                eco_CR_model = MacArthur_2C_2R, 
-                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
-sim_1C_2R_MacArthur_concaveDN <- eco_evo_CR(init_parameters = params_1C_2R_MacArthur_concaveDN, 
-                                init_states = states_1C_2R, 
-                                eco_CR_model = MacArthur_1C_2R, 
-                                AD_CR_models = c(mC1_MacArthur),
-                                species_traits = c(1), 
-                                eco_param_names = c("a11"), 
-                                eco_tradeoff_names = c("a12"),  
-                                evo_param_names = c("a11m"),  
-                                evo_tradeoff_names = c("a12m"), 
-                                tradeoff_functions = c(a12_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
-# asymmetric starting points
-params_2C_2R_MacArthur_concaveDN_asymm <- params_2C_2R_MacArthur_concaveDN
-params_2C_2R_MacArthur_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveDN[nrow(sim_1C_2R_MacArthur_concaveDN), c("a12","a11")] 
-sim_2C_2R_MacArthur_concaveDN_asymm <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_concaveDN_asymm,
-                                init_states = states_2C_2R, 
-                                eco_CR_model = MacArthur_2C_2R, 
-                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
-
-## concaveUP
-sim_2C_2R_MacArthur_concaveUP <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_concaveUP, 
-                                       init_states = states_2C_2R, 
-                                       eco_CR_model = MacArthur_2C_2R, 
-                                       AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
-                                       species_traits = c(1,2), 
-                                       eco_param_names = c("a11","a22"), 
-                                       eco_tradeoff_names = c("a12","a21"),  
-                                       evo_param_names = c("a11m","a22m"),  
-                                       evo_tradeoff_names = c("a12m","a21m"), 
-                                       tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                       extra_tradeoff_params = c(A=A, n=n_concaveUP))
-sim_1C_2R_MacArthur_concaveUP <- eco_evo_CR(init_parameters = params_1C_2R_MacArthur_concaveUP, 
-                                       init_states = states_1C_2R, 
-                                       eco_CR_model = MacArthur_1C_2R, 
-                                       AD_CR_models = c(mC1_MacArthur),
-                                       species_traits = c(1), 
-                                       eco_param_names = c("a11"), 
-                                       eco_tradeoff_names = c("a12"),  
-                                       evo_param_names = c("a11m"),  
-                                       evo_tradeoff_names = c("a12m"), 
-                                       tradeoff_functions = c(a12_tradeoff),
-                                       extra_tradeoff_params = c(A=A, n=n_concaveUP))
-# asymmetric starting points
-params_2C_2R_MacArthur_concaveUP_asymm <- params_2C_2R_MacArthur_concaveUP
-params_2C_2R_MacArthur_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_MacArthur_concaveUP[nrow(sim_1C_2R_MacArthur_concaveUP), c("a12","a11")]
-# max iterations reached!
-sim_2C_2R_MacArthur_concaveUP_asymm <- eco_evo_CR(init_parameters = params_2C_2R_MacArthur_concaveUP_asymm,
-                                init_states = states_2C_2R, 
-                                eco_CR_model = MacArthur_2C_2R, 
-                                AD_CR_models = c(mC1_MacArthur, mC2_MacArthur),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_concaveUP))
-
-######################################### LAWLOR SMITH COARSE-GRAIN MODEL #######################################################
-
-## Mutant growth rates
-mC1_LawlorSmith <- expression(e11*w11*a11m*R1 + e12*w12*a12m*R2 - m1)
-mC2_LawlorSmith <- expression(e21*w21*a21m*R1 + e22*w22*a22m*R2 - m2)
-
-#### PARAMETERS ####
-
-## Linear
-params_2C_2R_LawlorSmith_linear <- data.frame(general_parameters,
-                                              a11 = aii,
-                                              a12 = eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_linear)), 
-                                              a21 = eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_linear)), 
-                                              a22 = aii,
-                                              w11 = wii, 
-                                              w12 = wij, 
-                                              w21 = wij,
-                                              w22 = wii)
-params_1C_2R_LawlorSmith_linear <- params_2C_2R_LawlorSmith_linear[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22")]
-
-## concaveDN
-params_2C_2R_LawlorSmith_concaveDN <- params_2C_2R_LawlorSmith_linear
-params_2C_2R_LawlorSmith_concaveDN["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveDN))
-params_2C_2R_LawlorSmith_concaveDN["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveDN))
-params_1C_2R_LawlorSmith_concaveDN <- params_2C_2R_LawlorSmith_concaveDN[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22")]
-
-## concaveUP
-params_2C_2R_LawlorSmith_concaveUP <- params_2C_2R_LawlorSmith_linear
-params_2C_2R_LawlorSmith_concaveUP["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveUP))
-params_2C_2R_LawlorSmith_concaveUP["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveUP))
-params_1C_2R_LawlorSmith_concaveUP <- params_2C_2R_LawlorSmith_concaveUP[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22")]
-
-#### SIMULATIONS ####
-
-## Linear
-sim_2C_2R_LawlorSmith_linear <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_linear, 
-                                       init_states = states_2C_2R, 
-                                       eco_CR_model = LawlorSmith_2C_2R, 
-                                       AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
-                                       species_traits = c(1,2), 
-                                       eco_param_names = c("a11","a22"), 
-                                       eco_tradeoff_names = c("a12","a21"),  
-                                       evo_param_names = c("a11m","a22m"),  
-                                       evo_tradeoff_names = c("a12m","a21m"), 
-                                       tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                       extra_tradeoff_params = c(A=A, n=n_linear))
-sim_1C_2R_LawlorSmith_linear <- eco_evo_CR(init_parameters = params_1C_2R_LawlorSmith_linear, 
-                                       init_states = states_1C_2R, 
-                                       eco_CR_model = LawlorSmith_1C_2R, 
-                                       AD_CR_models = c(mC1_LawlorSmith),
-                                       species_traits = c(1), 
-                                       eco_param_names = c("a11"), 
-                                       eco_tradeoff_names = c("a12"),  
-                                       evo_param_names = c("a11m"),  
-                                       evo_tradeoff_names = c("a12m"), 
-                                       tradeoff_functions = c(a12_tradeoff),
-                                       extra_tradeoff_params = c(A=A, n=n_linear))
-# asymmetric starting points
-params_2C_2R_LawlorSmith_linear_asymm <- params_2C_2R_LawlorSmith_linear
-params_2C_2R_LawlorSmith_linear_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_linear[nrow(sim_1C_2R_LawlorSmith_linear), c("a12","a11")] 
-sim_2C_2R_LawlorSmith_linear_asymm <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_linear_asymm,
-                                init_states = states_2C_2R, 
-                                eco_CR_model = LawlorSmith_2C_2R, 
-                                AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_linear))
-
-## concaveDN
-# max mut iter reach!
-sim_2C_2R_LawlorSmith_concaveDN <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveDN, 
-                                       init_states = states_2C_2R, 
-                                       eco_CR_model = LawlorSmith_2C_2R, 
-                                       AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
-                                       species_traits = c(1,2), 
-                                       eco_param_names = c("a11","a22"), 
-                                       eco_tradeoff_names = c("a12","a21"),  
-                                       evo_param_names = c("a11m","a22m"),  
-                                       evo_tradeoff_names = c("a12m","a21m"), 
-                                       tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                       extra_tradeoff_params = c(A=A, n=n_concaveDN))
-# note that although the max mutation iterations are reached, the evolving parameters appear to oscillate near high values, i.e. they are effectively at an ESS
-plot(sim_2C_2R_LawlorSmith_concaveDN[,c("a11","a22")]) 
-plot(sim_2C_2R_LawlorSmith_concaveDN[,c("a11","a22")], xlim=c(1.85,1.95))
-
-sim_1C_2R_LawlorSmith_concaveDN <- eco_evo_CR(init_parameters = params_1C_2R_LawlorSmith_concaveDN, 
-                                       init_states = states_1C_2R, 
-                                       eco_CR_model = LawlorSmith_1C_2R, 
-                                       AD_CR_models = c(mC1_LawlorSmith),
-                                       species_traits = c(1), 
-                                       eco_param_names = c("a11"), 
-                                       eco_tradeoff_names = c("a12"),  
-                                       evo_param_names = c("a11m"),  
-                                       evo_tradeoff_names = c("a12m"), 
-                                       tradeoff_functions = c(a12_tradeoff),
-                                       extra_tradeoff_params = c(A=A, n=n_concaveDN))
-# asymmetric starting points
-params_2C_2R_LawlorSmith_concaveDN_asymm <- params_2C_2R_LawlorSmith_concaveDN
-params_2C_2R_LawlorSmith_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_concaveDN[nrow(sim_1C_2R_LawlorSmith_concaveDN), c("a12","a11")] 
-sim_2C_2R_LawlorSmith_concaveDN_asymm <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveDN_asymm,
-                                init_states = states_2C_2R, 
-                                eco_CR_model = LawlorSmith_2C_2R, 
-                                AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
-
-## concaveUP
-sim_2C_2R_LawlorSmith_concaveUP <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveUP, 
-                                        init_states = states_2C_2R, 
-                                        eco_CR_model = LawlorSmith_2C_2R, 
-                                        AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
-                                        species_traits = c(1,2), 
-                                        eco_param_names = c("a11","a22"), 
-                                        eco_tradeoff_names = c("a12","a21"),  
-                                        evo_param_names = c("a11m","a22m"),  
-                                        evo_tradeoff_names = c("a12m","a21m"), 
-                                        tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                        extra_tradeoff_params = c(A=A, n=n_concaveUP))
-sim_1C_2R_LawlorSmith_concaveUP <- eco_evo_CR(init_parameters = params_1C_2R_LawlorSmith_concaveUP, 
-                                        init_states = states_1C_2R, 
-                                        eco_CR_model = LawlorSmith_1C_2R, 
-                                        AD_CR_models = c(mC1_LawlorSmith),
-                                        species_traits = c(1), 
-                                        eco_param_names = c("a11"), 
-                                        eco_tradeoff_names = c("a12"),  
-                                        evo_param_names = c("a11m"),  
-                                        evo_tradeoff_names = c("a12m"), 
-                                        tradeoff_functions = c(a12_tradeoff),
-                                        extra_tradeoff_params = c(A=A, n=n_concaveUP))
-# asymmetric starting points
-params_2C_2R_LawlorSmith_concaveUP_asymm <- params_2C_2R_LawlorSmith_concaveUP
-params_2C_2R_LawlorSmith_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_LawlorSmith_concaveUP[nrow(sim_1C_2R_LawlorSmith_concaveUP), c("a12","a11")] 
-sim_2C_2R_LawlorSmith_concaveUP_asymm <- eco_evo_CR(init_parameters = params_2C_2R_LawlorSmith_concaveUP_asymm,
-                                init_states = states_2C_2R, 
-                                eco_CR_model = LawlorSmith_2C_2R, 
-                                AD_CR_models = c(mC1_LawlorSmith, mC2_LawlorSmith),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_concaveUP))
-
-######################################### MCCANN MODEL #######################################################
-
-## Mutant growth rates
-mC1_McCann <- expression(e11*(((w11 * R1)/(w11 * R1 + w12 * R2)) * a11m * R1)/(1 + ((w11 * R1)/(w11 * R1 + w12 * R2)) * a11m * h11 * R1 + ((w12 * R2)/(w11 * R1 + w12 * R2)) * a12m * h12 * R2) + e12*(((w12 * R2)/(w11 * R1 + w12 * R2)) * a12m * R2)/(1 + ((w11 * R1)/(w11 * R1 + w12 * R2)) * a11m * h11 * R1 + ((w12 * R2)/(w11 * R1 + w12 * R2)) * a12m * h12 * R2) - m1)
-mC2_McCann <- expression(e21*(((w21 * R1)/(w21 * R1 + w22 * R2)) * a21m * R1)/(1 + ((w21 * R1)/(w21 * R1 + w22 * R2)) * a21m * h21 * R1 + ((w22 * R2)/(w21 * R1 + w22 * R2)) * a22m * h22 * R2) + e22*(((w22 * R2)/(w21 * R1 + w22 * R2)) * a22m * R2)/(1 + ((w21 * R1)/(w21 * R1 + w22 * R2)) * a21m * h21 * R1 + ((w22 * R2)/(w21 * R1 + w22 * R2)) * a22m * h22 * R2) - m2)
-
-#### PARAMETERS ####
-
-## Linear
-params_2C_2R_McCann_linear <- data.frame(general_parameters,
-                                         a11 = aii,
-                                         a12 = eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_linear)), 
-                                         a21 = eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_linear)), 
-                                         a22 = aii,
-                                         w11 = wii, 
-                                         w12 = wij, 
-                                         w21 = wij,
-                                         w22 = wii,
-                                         h11 = h11,
-                                         h12 = h12,
-                                         h21 = h21,
-                                         h22 = h22)
-
-params_1C_2R_McCann_linear <- params_2C_2R_McCann_linear[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22","h11","h12","h21","h22")]
-
-## concaveDN
-params_2C_2R_McCann_concaveDN <- params_2C_2R_McCann_linear
-params_2C_2R_McCann_concaveDN["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveDN))
-params_2C_2R_McCann_concaveDN["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveDN))
-params_1C_2R_McCann_concaveDN <- params_2C_2R_McCann_concaveDN[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22","h11","h12","h21","h22")]
-
-## concaveUP
-params_2C_2R_McCann_concaveUP <- params_2C_2R_McCann_linear
-params_2C_2R_McCann_concaveUP["a12"] <- eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_concaveUP))
-params_2C_2R_McCann_concaveUP["a21"] <- eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_concaveUP))
-params_1C_2R_McCann_concaveUP <- params_2C_2R_McCann_concaveUP[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","w21","w22","h11","h12","h21","h22")]
-
-#### SIMULATIONS ####
-
-## Linear
-sim_2C_2R_McCann_linear <- eco_evo_CR(init_parameters = params_2C_2R_McCann_linear, 
-                                         init_states = states_2C_2R, 
-                                         eco_CR_model = McCann_2C_2R, 
-                                         AD_CR_models = c(mC1_McCann, mC2_McCann),
-                                         species_traits = c(1,2), 
-                                         eco_param_names = c("a11","a22"), 
-                                         eco_tradeoff_names = c("a12","a21"),  
-                                         evo_param_names = c("a11m","a22m"),  
-                                         evo_tradeoff_names = c("a12m","a21m"), 
-                                         tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                         extra_tradeoff_params = c(A=A, n=n_linear))
-sim_1C_2R_McCann_linear <- eco_evo_CR(init_parameters = params_1C_2R_McCann_linear, 
-                                         init_states = states_1C_2R, 
-                                         eco_CR_model = McCann_1C_2R, 
-                                         AD_CR_models = c(mC1_McCann),
-                                         species_traits = c(1), 
-                                         eco_param_names = c("a11"), 
-                                         eco_tradeoff_names = c("a12"),  
-                                         evo_param_names = c("a11m"),  
-                                         evo_tradeoff_names = c("a12m"), 
-                                         tradeoff_functions = c(a12_tradeoff),
-                                         extra_tradeoff_params = c(A=A, n=n_linear))
-# I'm thinking these initial asymmetric simulations don't quite make sense (at least for LawlorSmith and McCann).
-This is because I didn't switch the pre-adapted habitat preferences as well...This is important because C1 in the first simulation becomes C2 in the 2C-2R sim.
-This isn't an issue in the symmetry model, because I assume they are pre-adapted on different resources.
-# asymmetric starting points
-params_2C_2R_McCann_linear_asymm <- params_2C_2R_McCann_linear
-params_2C_2R_McCann_linear_asymm[c("a21","a22")] <- sim_1C_2R_McCann_linear[nrow(sim_1C_2R_McCann_linear), c("a12","a11")] # make sure a_ij and a_ii are in correct order.
-Note that habitat preference (w) is already set in the appropriate direction
-sim_2C_2R_McCann_linear_asymm <- eco_evo_CR(init_parameters = params_2C_2R_McCann_linear_asymm,
-                                init_states = states_2C_2R, 
-                                eco_CR_model = McCann_2C_2R, 
-                                AD_CR_models = c(mC1_McCann, mC2_McCann),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_linear))
-
-## concaveDN
-# max mut iter
-sim_2C_2R_McCann_concaveDN <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveDN, 
-                                         init_states = states_2C_2R, 
-                                         eco_CR_model = McCann_2C_2R, 
-                                         AD_CR_models = c(mC1_McCann, mC2_McCann),
-                                         species_traits = c(1,2), 
-                                         eco_param_names = c("a11","a22"), 
-                                         eco_tradeoff_names = c("a12","a21"),  
-                                         evo_param_names = c("a11m","a22m"),  
-                                         evo_tradeoff_names = c("a12m","a21m"), 
-                                         tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                         extra_tradeoff_params = c(A=A, n=n_concaveDN))
-# note that although the max mutation iterations are reached, the evolving parameters appear to oscillate near high values. i.e. they are effectively at an ESS
-plot(sim_2C_2R_McCann_concaveDN[,c("a11","a22")]) 
-plot(sim_2C_2R_McCann_concaveDN[,c("a11","a22")], xlim=c(1.85,1.95))
-
-sim_1C_2R_McCann_concaveDN <- eco_evo_CR(init_parameters = params_1C_2R_McCann_concaveDN, 
-                                         init_states = states_1C_2R, 
-                                         eco_CR_model = McCann_1C_2R, 
-                                         AD_CR_models = c(mC1_McCann),
-                                         species_traits = c(1), 
-                                         eco_param_names = c("a11"), 
-                                         eco_tradeoff_names = c("a12"),  
-                                         evo_param_names = c("a11m"),  
-                                         evo_tradeoff_names = c("a12m"), 
-                                         tradeoff_functions = c(a12_tradeoff),
-                                         extra_tradeoff_params = c(A=A, n=n_concaveDN))
-# asymmetric starting points
-params_2C_2R_McCann_concaveDN_asymm <- params_2C_2R_McCann_concaveDN
-params_2C_2R_McCann_concaveDN_asymm[c("a21","a22")] <- sim_1C_2R_McCann_concaveDN[nrow(sim_1C_2R_McCann_concaveDN), c("a12","a11")] 
-sim_2C_2R_McCann_concaveDN_asymm <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveDN_asymm,
-                                init_states = states_2C_2R, 
-                                eco_CR_model = McCann_2C_2R, 
-                                AD_CR_models = c(mC1_McCann, mC2_McCann),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_concaveDN))
-
-## concaveUP
-sim_2C_2R_McCann_concaveUP <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveUP, 
-                                          init_states = states_2C_2R, 
-                                          eco_CR_model = McCann_2C_2R, 
-                                          AD_CR_models = c(mC1_McCann, mC2_McCann),
-                                          species_traits = c(1,2), 
-                                          eco_param_names = c("a11","a22"), 
-                                          eco_tradeoff_names = c("a12","a21"),  
-                                          evo_param_names = c("a11m","a22m"),  
-                                          evo_tradeoff_names = c("a12m","a21m"), 
-                                          tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                          extra_tradeoff_params = c(A=A, n=n_concaveUP))
-sim_1C_2R_McCann_concaveUP <- eco_evo_CR(init_parameters = params_1C_2R_McCann_concaveUP, 
-                                          init_states = states_1C_2R, 
-                                          eco_CR_model = McCann_1C_2R, 
-                                          AD_CR_models = c(mC1_McCann),
-                                          species_traits = c(1), 
-                                          eco_param_names = c("a11"), 
-                                          eco_tradeoff_names = c("a12"),  
-                                          evo_param_names = c("a11m"),  
-                                          evo_tradeoff_names = c("a12m"), 
-                                          tradeoff_functions = c(a12_tradeoff),
-                                          extra_tradeoff_params = c(A=A, n=n_concaveUP))
-# asymmetric starting points
-params_2C_2R_McCann_concaveUP_asymm <- params_2C_2R_McCann_concaveUP
-params_2C_2R_McCann_concaveUP_asymm[c("a21","a22")] <- sim_1C_2R_McCann_concaveUP[nrow(sim_1C_2R_McCann_concaveUP), c("a12","a11")] 
-sim_2C_2R_McCann_concaveUP_asymm <- eco_evo_CR(init_parameters = params_2C_2R_McCann_concaveUP_asymm,
-                                init_states = states_2C_2R, 
-                                eco_CR_model = McCann_2C_2R, 
-                                AD_CR_models = c(mC1_McCann, mC2_McCann),
-                                species_traits = c(1,2), 
-                                eco_param_names = c("a11","a22"), 
-                                eco_tradeoff_names = c("a12","a21"),  
-                                evo_param_names = c("a11m","a22m"),  
-                                evo_tradeoff_names = c("a12m","a21m"), 
-                                tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                extra_tradeoff_params = c(A=A, n=n_concaveUP))
-```
-
-```{r Organize_Simulation_Data, include=FALSE, cache=TRUE}
-
-## Gather data. Note different formula for effective attack rate of consumer 1 "a_eff_C1". This accounts for the differences between fine- and coarse-grained environments
-sim_data <- bind_rows(
-  mutate(sim_2C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "Yes", a_eff_C1 = a11+a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "No", a_eff_C1 = a11+a12, Ctot = C1),
-  mutate(sim_2C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "Yes", a_eff_C1 = a11+a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "No", a_eff_C1 = a11+a12, Ctot = C1),
-  mutate(sim_2C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "Yes", a_eff_C1 = a11+a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "No", a_eff_C1 = a11+a12, Ctot = C1),
-  mutate(sim_2C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
-  mutate(sim_2C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
-  mutate(sim_2C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
-  mutate(sim_2C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
-  mutate(sim_2C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1),
-  mutate(sim_2C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "Yes", a_eff_C1 = w11*a11+w12*a12, Ctot = C1+C2), 
-  mutate(sim_1C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "No", a_eff_C1 = w11*a11+w12*a12, Ctot = C1))
-
-#### PREDICTING EFFECTS OF CHARACTER DISPLACEMENT ON C-R DYNAMICS ####
-
-tradeoff_df <- bind_rows(mutate(concaveDN_df, n = n_concaveDN), 
-                         mutate(linear_df, n = n_linear), 
-                         mutate(concaveUP_df, n = n_concaveUP)) %>%
-  data.frame(., general_parameters) %>%
-  mutate(a11 = aii, a12 = aij, a21 = aij, a22 = aii,
-         w11 = wii, w12 = wij, w21 = wij, w22 = wii)
-
-## MacArthur
-MacArthur_displacements <- bind_rows(
-  mutate(sim_2C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_linear)), ], 
-  mutate(sim_1C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_linear)), ],
-  mutate(sim_2C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveDN)), ], 
-  mutate(sim_1C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveDN)), ],
-  mutate(sim_2C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveUP)), ], 
-  mutate(sim_1C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveUP)), ]) 
-MacArthur_displacements$Time <- rep(c("Begin","End"),6)
-
-MacArthur_2C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:a22), init_states = states_2C_2R, eco_CR_model = MacArthur_2C_2R)
-MacArthur_1C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:a12), init_states = states_1C_2R, eco_CR_model = MacArthur_1C_2R)
-
-## MacArthur asymmetry. Note that all 1C_2R models are from the original simulations, as these were the starting points for the asymmetric sims (and therefore the appropriate comparison)
-MacArthur_displacements_asymm <- bind_rows(
-  mutate(sim_2C_2R_MacArthur_linear_asymm, Model = "MacArthur", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_linear_asymm)), ], 
-  mutate(sim_1C_2R_MacArthur_linear, Model = "MacArthur", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_linear)), ],
-  mutate(sim_2C_2R_MacArthur_concaveDN_asymm, Model = "MacArthur", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveDN_asymm)), ], 
-  mutate(sim_1C_2R_MacArthur_concaveDN, Model = "MacArthur", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveDN)), ],
-  mutate(sim_2C_2R_MacArthur_concaveUP_asymm, Model = "MacArthur", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_MacArthur_concaveUP_asymm)), ], 
-  mutate(sim_1C_2R_MacArthur_concaveUP, Model = "MacArthur", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_MacArthur_concaveUP)), ]) 
-MacArthur_displacements_asymm$Time <- rep(c("Begin","End"),6)
-
-## Lawlor Smith
-LawlorSmith_displacements <- bind_rows(
-  mutate(sim_2C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_linear)), ], 
-  mutate(sim_1C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_linear)), ],
-  mutate(sim_2C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_concaveDN)), ], 
-  mutate(sim_1C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_concaveDN)), ],
-  mutate(sim_2C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_concaveUP)), ], 
-  mutate(sim_1C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_concaveUP)), ]) 
-LawlorSmith_displacements$Time <- rep(c("Begin","End"),6)
-
-LawlorSmith_2C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:w22), init_states = states_2C_2R, eco_CR_model = LawlorSmith_2C_2R)
-LawlorSmith_1C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:a12, w11, w12), init_states = states_1C_2R, eco_CR_model = LawlorSmith_1C_2R)
-
-# asymmetry
-LawlorSmith_displacements_asymm <- bind_rows(
-  mutate(sim_2C_2R_LawlorSmith_linear_asymm, Model = "LawlorSmith", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_linear_asymm)), ], 
-  mutate(sim_1C_2R_LawlorSmith_linear, Model = "LawlorSmith", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_linear)), ],
-  mutate(sim_2C_2R_LawlorSmith_concaveDN_asymm, Model = "LawlorSmith", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_concaveDN_asymm)), ], 
-  mutate(sim_1C_2R_LawlorSmith_concaveDN, Model = "LawlorSmith", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_concaveDN)), ],
-  mutate(sim_2C_2R_LawlorSmith_concaveUP_asymm, Model = "LawlorSmith", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_LawlorSmith_concaveUP_asymm)), ], 
-  mutate(sim_1C_2R_LawlorSmith_concaveUP, Model = "LawlorSmith", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_LawlorSmith_concaveUP)), ]) 
-LawlorSmith_displacements_asymm$Time <- rep(c("Begin","End"),6)
-
-## McCann
-McCann_displacements <- bind_rows(
-  mutate(sim_2C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_linear)), ], 
-  mutate(sim_1C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_linear)), ],
-  mutate(sim_2C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_concaveDN)), ], 
-  mutate(sim_1C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_concaveDN)), ],
-  mutate(sim_2C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_concaveUP)), ], 
-  mutate(sim_1C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_concaveUP)), ]) 
-McCann_displacements$Time <- rep(c("Begin","End"),6)
-
-McCann_2C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:w22), init_states = states_2C_2R, eco_CR_model = McCann_2C_2R)
-McCann_1C_2R_tradeoff_dynamics <- CR_dynamics(init_parameters = select(tradeoff_df, r1:a12, w11, w12), init_states = states_1C_2R, eco_CR_model = McCann_1C_2R)
-
-# asymmetry
-McCann_displacements_asymm <- bind_rows(
-  mutate(sim_2C_2R_McCann_linear_asymm, Model = "McCann", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_linear_asymm)), ], 
-  mutate(sim_1C_2R_McCann_linear, Model = "McCann", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_linear)), ],
-  mutate(sim_2C_2R_McCann_concaveDN_asymm, Model = "McCann", n = n_concaveDN, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_concaveDN_asymm)), ], 
-  mutate(sim_1C_2R_McCann_concaveDN, Model = "McCann", n = n_concaveDN, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_concaveDN)), ],
-  mutate(sim_2C_2R_McCann_concaveUP_asymm, Model = "McCann", n = n_concaveUP, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_concaveUP_asymm)), ], 
-  mutate(sim_1C_2R_McCann_concaveUP, Model = "McCann", n = n_concaveUP, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_concaveUP)), ]) 
-McCann_displacements_asymm$Time <- rep(c("Begin","End"),6)
-
-# write out data for exploring later and avoiding rerunning the simulations
-write.csv(bind_rows(MacArthur_displacements, LawlorSmith_displacements, McCann_displacements), file = "symmetric_displacement_data.csv")
-write.csv(bind_rows(MacArthur_displacements_asymm, LawlorSmith_displacements_asymm, McCann_displacements_asymm), file = "asymmetric_displacement_data.csv")
-```
-
-# Abstract
-
-Ecological character displacement is an adaptive process that generally increases phenotypic diversity.
-Despite the fact that this diversification is due to an eco-evolutionary feedback between consumers competing for shared resources, its consequences for food-web dynamics have received little attention.
-Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumer attack rates affects resource abundances and the resilience of food webs to perturbations.
-I found that character displacement always strengthened consumer-resource interactions whenever consumers competed for resources that occurred in different habitats.
-This increase in interaction strength resulted in lower resource abundances and less resilient food webs.
-This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
-Taken together, my results show that the adaptive process of character displacement may come with the ecological cost of decreasing food-web resilience.
-
-\newpage
-
-# Introduction
-
-Ecological character displacement is an important adaptive process in generating biodiversity [@Schluter2000; @Pfennig2010]. 
-This process is due to "phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000]. 
-A large body of theoretical [e.g.
-@Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has examined which scenarios lead to phenotypic divergence or convergence of competing consumers.
-The general conclusion has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of density dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985].
-This adaptive process is not simply a response to static differences in resource distributions, but creates an eco-evolutionary feedback that drives further differentiation. 
-This crucial insight was made by theoretical models that explicitly included resource dynamics as a mediator of competition in driving evolutionary change [@Lawlor1976; @Abrams1986; @Taper1985].
-
-Although models that included resources led to insights about the evolution of character displacement, the ecological feedback onto consumer-resource dynamics has received surprisingly little attention.
-This is likely because the ecological feedback has been primarily studied through the lens of coexistence theory [@Lawlor1976; @Germain2018; @Bassar2017; @McPeek2019]. 
-For example, early theoretical work showed that character displacement promotes coexistence by favoring specialized consumers that experience reduced interspecific competition [@Lawlor1976]. 
-Yet, this reduction in interspecific competition may, at the same time, increase interspecific interactions between specialized consumers and their resources. 
-Both food-web theory and empirical studies have shown that increasing the strength of consumer-resource interactions often suppresses the abundance of resources, which if sufficient enough, can generate oscillations and less stable consumer-resource dynamics [@Rosenzweig1971; @Luckinbill1973; @Murdoch2002; @Murdoch2003; @McCann2011].
-Thus, a food-web perspective, which accounts for both the direct and indirect effects of consumer-resource interactions, may yield new insight to the ecological consequences of character displacement.
-
-Here, I address this knowledge gap by studying a mathematical model that examines how ecological character displacement affects consumer-resource dynamics in a food-web context.
-Specifically, I sought to answer the question: how does character displacement in consumer attack rates affect resource abundances and food-web stability?
-To test the generality of these effects, I explored different ecological foraging scenarios and evolutionary tradeoffs in consumer attack rates. 
-I found that the adaptive process of character displacement often comes with an ecological cost; resulting in food webs with lower resource availability and that are less resilient to perturbations.
-
-# Material and methods
-
-## Underlying consumer-resource dynamics
-
-To examine how ecological character displacement affects resource abundances and food-web stability, I analyzed a continuous-time model of two consumers ($C_{j=1,2}$) competing for two shared resources ($R_{i=1,2}$):
-
-\begin{equation} \label{eq:1}
-  \begin{split}
-     & \frac{dR_1}{dt}=r_1R_1(1-\frac{R_1}{K_1})-F_{11}(R_1)C_1-F_{12}(R_1)C_2 \\
-     & \frac{dR_2}{dt}=r_2R_2(1-\frac{R_2}{K_2})-F_{21}(R_2)C_1-F_{22}(R_2)C_2 \\
-     & \frac{dC_1}{dt}=e_{11}F_{11}(R_1)C_1+e_{21}F_{21}(R_2)C_1-m_1C_1 \\
-     & \frac{dC_2}{dt}=e_{12}F_{12}(R_1)C_2+e_{22}F_{22}(R_2)C_2-m_2C_2 \\
-  \end{split}
-\end{equation}
-
-where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ij}$ represents the conversion efficiency of resource $i$ into consumer $j$, and $m_j$ represents the mortality rate of consumer $j$.
-$F_{ij}(R_i)$ represents consumer $j$'s feeding rate on resource $i$ (i.e., its functional response).
-This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g.
-zooplankton and benthic invertebrates in lakes) rather than a scenario where resources are better characterized by a continuous trait distribution (e.g., seed size; see @Taper1985 for an example).
-Importantly, inferences about character displacement can only be made by comparing food webs with and without a competing consumer [@Schluter1992].
-Therefore, I arbitrarily set $C_2=0$ to create a food-web without a competing consumer for these comparisons.
-
-## Foraging scenarios
-
-I studied three different foraging scenarios. 
-In the first, I assume that consumers can forage for both resources simultaneously (Fig. \ref{fig:foraging_scenarios}a) and their feeding rate increases linearly with resource abundance, such that:
-
-\begin{equation} \label{eq:2}
-  F_{ij}(R_i)=a_{ij}R_i
-\end{equation}
-
-where $a_{ij}$ is the attack rate of consumer $j$ on resource $i$. 
-This first scenario is the starting point for many models of resource competition [@MacArthur1972]; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005].
-The second scenario accounts for this spatial context (Fig. \ref{fig:foraging_scenarios}b) and takes the form:
-
-\begin{equation} \label{eq:3}
-  F_{ij}(R_i)=w_{ij}a_{ij}R_i
-\end{equation}
-
-where $w_{ij}$ represents the proportion of time consumer $j$ spends foraging in a habitat where only resource $i$ is found (i.e., its habitat preference). 
-Note that since $w_{ij}$ is a proportion that $w_{1,j}=1-w_{2,j}$. <!-- $\sum_jw_{ij}=1$.<!-- $w_{ij}=1-w_{ii}$. -->
-Finally, it is well known that consumer feeding rates often saturate at high resource abundances [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011] and that consumers do not usually spend a fixed proportion of time in a particular habitat [@McCann2005].
-The third scenario accounts for these biological realities and takes the form [derived by @McCann2005]:     
-\begin{equation} \label{eq:4}
-  F_{ij}(R_i)=\frac{a_{ij}W_{ij}R_i}{1+a_{1,j}h_{1,j}W_{1,j}R_1+a_{2,j}h_{2,j}W_{2,j}R_2}
-\end{equation}
-
-where consumer $j$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling time $h_{ij}$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ij}=\frac{w_{ij}R_i}{w_{1,j}R_1+w_{2,j}R_2}$.
-
-Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach, with the general result being divergent character displacement [@Lawlor1976; @Abrams1986].
-I also used an Adaptive Dynamics approach to analyze the evolution of consumer attack rates in the third foraging scenario, and I too observed divergent character displacement (detailed analysis given in Appendix S1).
-I say consumers have undergone divergent character displacement if their evolved attack rates are more specialized when evolving with vs. without a competing consumer. 
-Specialization of consumer $j$ on resource $1$ is measured as $\frac{a_{1,j}}{a_{1,j}+a_{2,j}}$, where a value of 0.5 is a complete generalist ($a_{1,j}=a_{2,j}$), and a value of 1 is a complete specialist ($a_{2,j}=0$).
-Values less than 0.5 indicate specialization on the other resource. 
-Since I did not observe convergent character displacement in any of the foraging scenarios I analyzed, I refer to divergent character displacement as simply (ecological) character displacement throughout the rest of the text.
 
@@
 
-## Food-web dynamics
-
-Given that character displacement occurred across these foraging scenarios, I focus here on its consequences for food-web dynamics.
-To do this, I analyzed differences in resource abundances and food-web stability at equilibrium. 
-An equilibrium is reached when there is no change in the population growth rates of consumers and resources (i.e., the rates of change in equation \ref{eq:1} are 0), and solving the system at this point gives equilibrium abundances for each resource ($\hat R_i$) and each consumer ($\hat C_j$).
-I also compared the local stability of these food webs using standard methods [@Otto2007]. 
-This stability analysis derives the dominant eigenvalue, $\lambda_{\text{max}}$, of the matrix of partial derivatives of each species' population growth rate (given by equation \ref{eq:1}) with respect to each species' abundance evaluated at equilibrium.
-If $-\lambda_{max}>0$, then the food web will return to equilibrium after a small perturbation (i.e., it is locally stable), with more positive values indicating a faster return time.
-If $-\lambda_{max}<0$, then the food web is not locally stable.
-
-When possible, I derived analytical expressions for the relationship between consumer attack rates and food-web dynamics. 
-To do this, I simplified the model by assuming that resources are equivalent ($r=r_i$ and $K=K_i$) as well as consumers ($e=e_{ij}$; $h=h_{ij}$; $m=m_j$), except that consumer attack rates and their habitat preferences (if present) are mirror images of each other ($a_{11}=a_{22}$; $a_{12}=a_{21}$; $w_{11}=w_{22}$).
-Note that I arbitrarily set $C_1$ as being pre-adapted to $R_1$ ($a_{11} > a_{21}$; $w_{11} > 0.5$), and therefore $C_2$ was pre-adapted to $R_2$.
-Controlling for other sources of variability allowed me to isolate the general effects of character displacement.
-All mathematical derivations were conducted in Mathematica [@Mathematica] and are provided in Appendices S1-3.
-
-To gain insight to the eco-evolutionary feedback generated by character displacement, I conducted simulations using an Adaptive Dynamics approach. 
-Specifically, after letting consumer and resource abundances reach a steady state, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability.
-The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006].
-This function has the useful property that it differentiates between cases where intermediate combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ are higher than the extremes (when $n>1$, green line in Fig.
-\ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line in Fig.
-\ref{fig:tradeoff}).
-When $n=1$, the tradeoff function is linear, and all combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ have the same total attack rate (blue line in Fig. \ref{fig:tradeoff}). 
-Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
-If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource abundances to reach a steady state. 
-I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g., $\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0).
-Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and $w_{11} = w_{22} = 0.6$. 
-I set an initial value of $a_{11} = a_{22} = 1.2$, while $a_{12}$ and $a_{21}$ depended on the value of *n*. 
-I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
-All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model)<!-- and has been archived with Zenodo (INSERT DOI)-->.
-
-```{r plot tradeoff, include=FALSE}
-initial_points <- bind_rows(params_2C_2R_MacArthur_linear, params_2C_2R_MacArthur_concaveDN, params_2C_2R_MacArthur_concaveUP) %>% mutate(n = c(n_linear, n_concaveDN, n_concaveUP))
-
-tradeoff_plot <- bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
-  ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
-  geom_line() + 
-  geom_point(data = initial_points, aes(x=a11, y=a12), size=3, show.legend=FALSE) + 
-  xlab(expression(bold(a[11]))) +
-  ylab(expression(bold(a[21]))) + 
-  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette)
-
-save_plot(filename = "Fig_2_Tradeoffs.pdf", plot = tradeoff_plot, 
-          base_height = 4,
-          base_width = 6) 
-```
-
 ```{r tradeoff, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:tradeoff}**Evolutionary tradeoffs in consumer attack rates.** In each foraging scenario, I explored the effects of three different tradeoffs: intermediate combinations of attack rates (*a*~1,*j*~, *a*~2,*j*~) are higher than the extremes (green line, *n* > 1); extreme combinations of attack rates are higher than intermediate investments (orange line, *n* < 1); and all combinations of attack rates have the same total attack rate (blue line, *n* = 1).
@@
 knitr::include_graphics("Fig_2_Tradeoffs.pdf")
-```
-
-# Results
-
-## Resources occur in the same habitat
-
-In this first scenario (equation \ref{eq:2}), the abundance of resources at equilibrium are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), and are determined by the following equation (derived in Appendix S2):
-
-\begin{equation} \label{eq:5}
-  \hat{R}=\frac{1}{a_{1,j}+a_{2,j}}\cdot\frac{m}{e}
-\end{equation}
-
-A key determinant of resource abundance in this scenario is the consumer's total attack rate, $a_{1,j}+a_{2,j}$. 
-Therefore, the effect of character displacement on food-web dynamics depends on how the shape of the tradeoff function influences the evolution of consumer attack rates.
-
-I found that the shape of the tradeoff function qualitatively affects the relationship between character displacement and resource abundances in this scenario (Fig. \ref{fig:plot_fig3}a,b).
-For example, if consumer's are constrained by a linear tradeoff (blue lines), then there is no net change in total attack rate (Fig.
-\ref{fig:plot_fig3}a) and character displacement has no effect on resource abundances (Fig.
-\ref{fig:plot_fig3}b).
-If the tradeoff is concave down (green lines), then resource abundances can actually increase under character displacement (Fig. \ref{fig:plot_fig3}b). 
-This is because the total attack rate of consumers is maximized at intermediate values ($a_{1,j}=a_{2,j}$) and decreases as consumers diverge (Fig. \ref{fig:plot_fig3}a). 
-When the tradeoff is concave up (orange lines), character displacement suppresses resource abundances due to the increase in total attack rates (Fig. \ref{fig:plot_fig3}a,b). 
-Although the equation I derived for resource abundances was for the scenario where both consumers and both resources were present, it accurately predicts the abundance of resources when a single consumer reaches its evolutionary stable strategy (ESS; triangles on respective colored lines in Fig.
-\ref{fig:plot_fig3}b).
-This is because a single consumer evolves to be a generalist that has equal attack rates on each resource (triangles at 0.5 along x-axis in Fig.
-\ref{fig:plot_fig3}a), resulting in equivalent resource abundances.
-
-The effect of character displacement on resources corresponds to its impact on food-web stability. 
-For example, when character displacement decreases resource abundances (orange points in Fig. \ref{fig:plot_fig3}b), there is also a decrease in food-web stability (Fig. \ref{fig:plot_fig3}c). 
-Character displacement may not affect or even increase food-web stability (blue and green lines in Fig.
-\ref{fig:plot_fig3}c); however, evolution does not favor strong divergence in these scenarios (blue and green points in Fig.
-\ref{fig:plot_fig3}a), which dampens these contingent effects.
-Note that the dip in stability occurs when both consumers evolve to be generalists, a situation that is not favored in any of the foraging scenarios I examined (Fig. \ref{fig:plot_fig3}c).
-
-```{r MacArthur_Resources, include=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers can forage for both resources simultaneously."}
-Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  ylab(expression(a[ii]+a[ij])) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Mac_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
-  ylab(expression(R[i]+R[j])) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Mac_plots <- plot_grid(Mac_attack + theme(legend.position = "none"), Mac_resources, 
-                       ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
-Mac_plots
-```
-
-```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource abundances (B) when consumers cannot forage for resources simultaneously."}
-LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
-  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-LS_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color=factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression(R[i]+R[j])) +
-  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-LS_plots <- plot_grid(LS_attack + theme(legend.position = "none"), LS_resources, 
-                      ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
-LS_plots
-```
-
-```{r Stability, include=FALSE, fig.cap="\\label{fig:Stability}Effect of character displacement on local stability when consumers can (A) and cannot (B) forage for both resources simultaneously.
-Small and large points correspond to initial and evolved strategies, respectively."}
-Stability_Mac <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + 
-  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression("Stability "(-lambda[max]))) + 
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Stability_LS <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) +
-  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  ylab(expression("Stability "(-lambda[max]))) + 
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Stability_plots <- plot_grid(Stability_Mac + theme(legend.position = "none"), Stability_LS + theme( axis.title.y = element_blank()),
-                             ncol = 2, labels = "AUTO", rel_widths = c(0.75,1))
-Stability_plots
-```
-
-## Resources occur in different habitats
-
-In the second foraging scenario (equation \ref{eq:3}), I again see that resource abundances are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), but are now determined by the following equation (derived in Appendix S3):
-
-\begin{equation} \label{eq:6}
-  \hat{R}=\frac{1}{w_{1,j}a_{1,j}+w_{2,j}a_{2,j}}\cdot\frac{m}{e}
-\end{equation}
-
-This equation implies that if consumers evolve to become specialists on resources that occur in their preferred habitat (e.g., $w_{1,j}>0.5$ and $a_{1,j}>a_{2,j}$), then the effective attack rate of consumers ($w_{1,j}a_{1,j}+w_{2,j}a_{2,j}$) will always increase, regardless of the tradeoff (Fig.
-\ref{fig:plot_fig3}d).
-Thus, character displacement always results in resource suppression (Fig. \ref{fig:plot_fig3}e). 
-Note that the shape of the tradeoff can modify the effect of character displacement. 
-This is not so much due to the tradeoff affecting the magnitude of displacement (it does, but the effect is minor), but because the form of the tradeoff affects resource abundances when a single consumer has reached an ESS (triangles in Fig.
-\ref{fig:plot_fig3}e).
-In contrast, resource abundances reach a similar value when consumers evolve in the presence of a competitor (circles in Fig.
-\ref{fig:plot_fig3}e), because character displacement tends to reach a constraint of complete specialization.
-It is worth noting that resource abundances are consistently higher at the single consumer ESS compared to the predictions I derived for when both consumers are present (deviation of triangles from respective colored lines in Fig.
-\ref{fig:plot_fig3}e).
-This is because consumers actually evolve slightly specialized attack rates on the resources that occur in their non-preferred habitat (deviation of triangles from 0.5 along x-axis in Fig.
-\ref{fig:plot_fig3}d).
-
-As seen previously, the effect of character displacement on resource abundances qualitatively corresponds to its effect on food-web stability (Fig. \ref{fig:plot_fig3}f). 
-Specifically, character displacement decreases food-web stability, regardless of the tradeoff in attack rates.
-This is not simply a consequence of having an additional consumer in the system, but emerges from the eco-evolutionary feedback between character displacement and resource suppression (Fig.
-\ref{fig:plot_fig3}c).
-For example, when the tradeoff is concave up (orange), the initial two-consumer food web (small circle) is more stable than when there is only one consumer (small triangle); however, this pattern switches by the end of the eco-evolutionary simulation (large points).
-
-```{r Fig 3, include=FALSE}
-Macs <- plot_grid(
-  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) +
-    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), 
-  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)),
-  Stability_Mac + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)),
-  ncol=1, align='v', labels = c("","b)","c)"), hjust = -4, vjust = 0.5, label_size = 10
-)
-Macs_title <- ggdraw() + draw_label("Resources: Same Habitat", fontface = 'bold', size = 10, hjust = 0.2)
-plot_Macs <- plot_grid(Macs_title, Macs, ncol = 1, rel_heights = c(0.1, 1))
-
-LSs <- plot_grid(
-  LS_attack  + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])) + 
-    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), 
-  LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()),   
-  Stability_LS + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10)), 
-  ncol=1, align='v', labels = c("","e)","f)"), hjust = -1.5, vjust = 0.5, label_size = 10
-)
-LSs_title <- ggdraw() + draw_label("Resources: Different Habitat", fontface = 'bold', size = 10, hjust = 0.3)
-plot_LSs <- plot_grid(LSs_title, LSs, ncol = 1, rel_heights = c(0.1, 1))
-
-plot_fig3 <- plot_grid(plot_Macs, plot_LSs, ncol=2, rel_widths = c(1,0.9)) 
-plot_fig3_label <- plot_fig3 + 
-  draw_label("a)", x = 0.098, y = 0.925, fontface = 'bold', size = 10) + 
-  draw_label("d)", x = 0.566, y = 0.925, fontface = 'bold', size = 10)
-plot_fig3_sub <- add_sub(plot_fig3_label, expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","* j]),")"))), vjust = 0, size=12)
-
-save_plot(filename="Fig_3_MacArthur_LawlorSmith.pdf", plot=plot_fig3_sub, base_height = 6.5, base_width = 6)
 ```
@@
 
-## Adding a more realistic functional response
-
-<!-- Keeping text in case reviewers want me to explain evolutionary dynamics of this model in the text
-
-The evolutionary dynamics of consumer attack rates have not been examined in this model, but as with the other models [@LawlorSmith1976], I find that divergent character displacement is inevitable. 
-This is because, in the absence of competition, consumer evolution is determined by both resource abundances and their habitat preferences, and reach an evolutionary stable strategy (ESS) when:
-
-$$w_{ii}R_i^2\delta_{ii}=w_{ij}R_j^2\delta_{ij}$$ 
-where $\delta_{ii}$ represents a small mutation in $a_{ii}$ and we assume a negative tradeoff in attack rates, such that $\delta_{ij}=-f(\delta_{ii})$. 
-The square-terms for resource abundances indicate that they largely govern the evolutionary dynamics. 
-This scenario favors the evolution of a generalist phenotype, as resource specialization will decrease that resource's abundance, thus favoring evolution toward the other resource until an ESS is reached.
-In contrast, in the presence of a competitor, resource abundances will be equivalent at equilibrium (derivation in Appendix S2) and consumers reach an ESS when:
-
-$$w_{ii}\delta_{ii}=w_{ij}\delta_{ij}$$
-
-Assuming consumers have a preferred resource (i.e. $w_{ii}>0.5$), then consumers will only reach an ESS when $|\delta_{ii}|<|\delta_{ij}|$ (e.g. green line, Fig. \ref{MacArthur_Plot}A). 
-Otherwise, they will evolve to be complete specialists on their preferred resource.
--->
-
-In the third foraging scenario (equation \ref{eq:4}), I observed the same general effect of character displacement as the previous scenario (resources in different habitats, but linear functional response).
-This is because resource abundances at equilibrium are governed by a similar dynamic (derived in Appendix S1):
-
-\begin{equation} \label{eq:7}
-  \hat{R}=\frac{1}{w_{1,j}a_{1,j}+w_{2,j}a_{2,j}}\cdot\frac{m}{e-hm}
-\end{equation}
-
-And since evolution favors character displacement toward their preferred resources (see Appendix S1), the effective attack rate of consumers ($w_{1,j}a_{1,j}+w_{2,j}a_{2,j}$) will always increase, resulting in lower resource abundances and decreased food-web stability (Appendix S4, Fig.
-S1).
-
-In the first two foraging scenarios, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equilibrium (because $-\lambda_{max}>0$, see Appendix S2-3).
-In this more realistic model, however, whether the food web is locally stable depends on consumer and resource parameters. 
-Specifically, I found that the two-consumer food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derived using Routh-Hurwitz criteria in Appendix S1):
-
-\begin{equation} \label{eq:8}
-  w_{1,j}a_{1,j}+w_{2,j}a_{2,j} > \frac{e+hm}{hK(e-hm)}
-\end{equation}
-
-This inequality indicates that character displacement always pushes the food web toward an unstable structure in this more realistic foraging scenario (Fig.
-\ref{fig:bifur_plot}).
-Note that I stopped the simulation in the two-consumer food web once it became locally unstable.
-I do not simulate beyond this point as this would require making assumptions about the dynamics of mutant consumers in variable environments, which is beyond the scope of this work.
-
-```{r McCann_Resources, include=FALSE, fig.cap="\\label{fig:McCann_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers exhibit a more realistic functional response."}
-McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  ylab(bquote(atop("Effective attack rate","("*w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j]*")"))) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-McCann_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-McCann_plots <- plot_grid(McCann_attack + theme(legend.position = "none"), McCann_resources, 
-                          ncol=2, labels = "AUTO", rel_widths = c(0.65,1))
-McCann_plots
-```
-
-```{r Stability_McCann, include=FALSE, fig.cap="\\label{fig:Stability_McCann}Effect of character displacement on local stability when consumers exhibit a more realistic functional response."}
-Stability_McCann <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen)) + 
-  geom_line(aes(color=factor(n))) +
-  geom_point(data = McCann_displacements, aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor, size=Time), color="white", inherit.aes = FALSE) +
-  xlab(expression(frac(a[ij],a[1*","*j]+a[2*","*j]))) +
-  ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) +
-  scale_shape_manual(values=c(24,21))  +
-  scale_size_manual(values=c(1.5,3)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Stability_McCann
-```
-
-```{r McCann Model, include=FALSE}
-McCann_legend <- get_legend(Stability_McCann)
-
-plot_McCann_model <- plot_grid(
-  McCann_attack    + theme(legend.position = "none") + xlab(""), 
-  McCann_resources    + theme(legend.position = "none") + xlab(""),
-  Stability_McCann    + theme(legend.position = "none") + xlab(""), 
-  ncol=2, align='hv', labels = c("a)","b)","c)"), hjust = -4, vjust = 1, label_size = 10)
-
-plot_McCann_model <- add_sub(plot_McCann_model, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
-
-# plotted in Appendix S4, Fig. S1
-save_plot(filename="Appendix_S4_Fig_S1_McCann.pdf", plot=plot_McCann_model, base_height = 6, base_width = 6)
-```
-
-```{r bifurcation, echo=FALSE, cache=TRUE}
-
-## set state variables; same as previous models but in vector form
-i.state.2C_2R <- c(R1 = R1, R2 = R2, C1 = C1, C2 = C2) 
-i.state.1C_2R <- c(R1 = R1, R2 = R2, C1 = C1)
-
-## set duration of simulations
-Time <- 1000
-
-## set up general code for bifurcation plot; increase to get bifurcation
-A <- 3.3
-
-# assuming linear trade-off in attack rates for simulation
-aii_seq <- seq(A/2, A, by = 0.01)
-aij_seq <- (A*(1-(aii_seq/A)^n_linear)^(1/n_linear))
-
-## Bifurcation plot code
-# this code loops over the sequence of different parameter values
-out <- list()
-for (i in 1:length(aii_seq)) {
-
-  r1 <- r2 <- r 
-  K1 <- K2 <- K 
-  e11 <- e12 <- e21 <- e22 <- e 
-  h11 <- h12 <- h21 <- h22 <- h 
-  m1 <- m2 <- m 
-  w11 <- w22 <- wii
-  w12 <- w21 <- wij
-
-  # set params 
-  param.loop <- c(r1 = r1, r2 = r2, K1 = K1, K2 = K2,
-                  e11 = e11, e12 = e12, e21 = e21, e22 = e22,
-                  h11 = h11, h12 = h12, h21 = h21, h22 = h22,
-                  a11 = aii_seq[i], a12 = aij_seq[i], a21 = aij_seq[i], a22 = aii_seq[i],
-                  m1 = m1, m2 = m2, w11 = w11, w22 = w22, w12 = w12, w21 = w21)
-
-  # Run the experiment.
-  init <- ode(i.state.2C_2R, 
-              1:Time, McCann_2C_2R, param.loop)
-  # Rerun the experiment with the state variables at the end of the initial simulation. This will enable me to determine the range of state variables around the equilibrium.
-  temp <- ode(init[Time,-1], 1:Time, McCann_2C_2R, param.loop)[,-1]
-  out[[i]] <- cbind(temp, Rt = temp[,"R1"] + temp[,"R2"], Ct = temp[,"C1"] + temp[ ,"C2"])
-}
-
-sim.list <- list()
-for (i in 1:length(aii_seq)) {
-  sim.list[[i]] <- data.frame(rep = i,
-                              R_max = max(out[[i]][900:1000,"Rt"]),
-                              R_min = min(out[[i]][900:1000,"Rt"]),
-                              C_max = max(out[[i]][900:1000,"Ct"]),
-                              C_min = min(out[[i]][900:1000,"Ct"]))
-}
-
-sim.df <- ldply(sim.list) %>%
-  mutate(a11 = aii_seq,
-         a12 = aij_seq) %>%
-  gather(state, value, -rep, -a11, -a12) %>%
-  separate(state, into = c("species","max_min"))
-
-## Simulate 3 species food web ----
-
-out_3 <- list()
-for (i in 1:length(aii_seq)) {
-
-  r1 <- r2 <- r
-  K1 <- K2 <- K
-  e11 <- e12 <- e21 <- e22 <- e
-  h11 <- h12 <- h21 <- h22 <- h
-  m1 <- m2 <- m
-
-  # set params 
-  param.loop <- c(r1 = r1, r2 = r2, K1 = K1, K2 = K2,
-                  e11 = e11, e12 = e12, 
-                  h11 = h11, h12 = h12,
-                  a11 = aii_seq[i], a12 = aij_seq[i], 
-                  m1 = m1, w11 = w11, w12 = w12)
-
-  # Run the experiment.
-  init <- ode(i.state.1C_2R, 
-              1:Time, McCann_1C_2R, param.loop)
-  # Rerun the experiment with the state variables at the end of the initial simulation; this will enable me to determine the range of state variables around the equilibrium.
-  temp <- ode(init[Time,-1], 1:Time, McCann_1C_2R, param.loop)[,-1]   
-  out_3[[i]] <- cbind(temp, Rt = temp[,"R1"] + temp[,"R2"])
-}
-
-sim.list_3 <- list()
-for (i in 1:length(aii_seq)) {
-  sim.list_3[[i]] <- data.frame(rep = i,
-                                R_max = max(out_3[[i]][900:1000,"Rt"]),
-                                R_min = min(out_3[[i]][900:1000,"Rt"]),
-                                C_max = max(out_3[[i]][900:1000,"C1"]),
-                                C_min = min(out_3[[i]][900:1000,"C1"]))
-}
-
-sim.df_3 <- ldply(sim.list_3) %>%
-  mutate(a11 = aii_seq,
-         a12 = aij_seq) %>% 
-  gather(state, value, -rep, -a11, -a12) %>%
-  separate(state, into = c("species","max_min"))
-```
-
-```{r bifurcation simulation, include=FALSE, cache=TRUE}
-## eco-evo sims
-
-## respecify initial attack rates based on new A-value
-aii <- A*0.61
-
-# params
-params_2C_2R_McCann_linear_bifur <- data.frame(general_parameters,
-                                               a11 = aii,
-                                               a12 = eval(a12_tradeoff, data.frame(A=A, a11m=aii, n=n_linear)), 
-                                               a21 = eval(a21_tradeoff, data.frame(A=A, a22m=aii, n=n_linear)), 
-                                               a22 = aii,
-                                               w11 = wii,
-                                               w12 = wij,
-                                               w21 = wij,
-                                               w22 = wii,
-                                               h11 = h,
-                                               h12 = h,
-                                               h21 = h,
-                                               h22 = h)
-
-params_1C_2R_McCann_linear_bifur <- params_2C_2R_McCann_linear_bifur[c("r1","r2","K1","K2","e11","e12","a11","a12","m1","w11","w12","h11","h12")]
-
-# eco-evo simulations
-sim_2C_2R_McCann_linear_bifur <- eco_evo_CR(init_parameters = params_2C_2R_McCann_linear_bifur, 
-                                             init_states = states_2C_2R, # same as previous models
-                                             eco_CR_model = McCann_2C_2R, 
-                                             AD_CR_models = c(mC1_McCann, mC2_McCann),
-                                             species_traits = c(1,2), 
-                                             eco_param_names = c("a11","a22"), 
-                                             eco_tradeoff_names = c("a12","a21"),  
-                                             evo_param_names = c("a11m","a22m"),  
-                                             evo_tradeoff_names = c("a12m","a21m"), 
-                                             tradeoff_functions = c(a12_tradeoff, a21_tradeoff),
-                                             extra_tradeoff_params = c(A=A, n=n_linear)) 
-sim_1C_2R_McCann_linear_bifur <- eco_evo_CR(init_parameters = params_1C_2R_McCann_linear_bifur, 
-                                             init_states = states_1C_2R, # same as previous models
-                                             eco_CR_model = McCann_1C_2R, 
-                                             AD_CR_models = c(mC1_McCann),
-                                             species_traits = c(1), 
-                                             eco_param_names = c("a11"), 
-                                             eco_tradeoff_names = c("a12"),  
-                                             evo_param_names = c("a11m"),  
-                                             evo_tradeoff_names = c("a12m"), 
-                                             tradeoff_functions = c(a12_tradeoff),
-                                             extra_tradeoff_params = c(A=A, n=n_linear))
-
-# save simulation output 
-write.csv(sim_2C_2R_McCann_linear_bifur, "sim_2C_2R_McCann_linear_bifur.csv")
-write.csv(sim_1C_2R_McCann_linear_bifur, "sim_1C_2R_McCann_linear_bifur.csv")
-```
-
-```{r bifur evo plot, echo=FALSE}
-McCann_displacements_bifur <- bind_rows(
-  mutate(sim_2C_2R_McCann_linear_bifur, Model = "McCann", n = n_linear, Competitor = "Yes")[c(1,nrow(sim_2C_2R_McCann_linear_bifur)), ], 
-  mutate(sim_1C_2R_McCann_linear_bifur, Model = "McCann", n = n_linear, Competitor = "No")[c(1,nrow(sim_1C_2R_McCann_linear_bifur)), ]) 
-McCann_displacements_bifur$Time <- rep(c("Begin","End"),2)
-
-## put it all together
-
-# with eco-evo begin and end points
-bifurcation_plot <- bind_rows(mutate(sim.df, Competitor = "Yes"), mutate(sim.df_3, Competitor = "No")) %>%
-  filter(species == "R") %>%
-  ggplot(., aes(x = a11/(a11+a12), y = value, group=interaction(max_min, Competitor))) +
-  geom_line(aes(linetype=Competitor), color=cbPalette[2]) +
-  geom_point(data = McCann_displacements_bifur, aes(x = a11/(a11+a12), y=R1+R2, shape=Competitor, size=Time), color=cbPalette[2], fill=cbPalette[2], inherit.aes = FALSE) +
-  scale_shape_manual(values=c(24,21))  +
-  scale_size_manual(values=c(3,6)) +
-  scale_linetype_manual(values=c("dotted","solid")) +
-  ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*": max and min)"))) +
-  xlab(expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")")))) +
-  geom_segment(x=0.815, xend=0.815, y=0, yend=2.65, arrow = arrow(length = unit(0.03, "npc")), color="grey", fill="grey") +
-  geom_text(x=0.77, y=2, label="Stability\nboundary", color="grey") +
-  theme_cowplot(font_size = 18)
-
-save_plot(filename = "Fig_4_Bifurcation.pdf", bifurcation_plot, base_width = 8.5, base_height = 6)
-```
-
 ```{r bifur_plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:bifur_plot}**Character displacement creates an unstable food web**.
@@
 ```
-
-## Robustness to consumer asymmetry
-
-The previous analytical results and simulations make a strong assumption that competing consumers start off as perfect mirror images of each other (i.e., there is symmetry). 
-Yet, theory indicates a predictable asymmetry between initial consumer attack rates. 
-This predictable asymmetry emerges from a process of community assembly where a single consumer invades a system, evolves to be a generalist that can equally attack both resources, followed by the invasion of a second, more specialized, consumer.
-This theoretical scenario has been hypothesized as the sequence of events leading to character displacement in threespine stickleback in small coastal lakes of British Columbia [@Schluter1992; @Schluter2000].
-
-To test whether my results were robust to this asymmetry, I used the evolved attack rates at the end of the simulations with one consumer as the starting values for one of the two consumers. 
-I did this for all foraging scenarios and tradeoffs previously examined. 
-I found that my previous inferences are robust to including consumer asymmetry across different foraging scenarios and tradeoffs (Appendix S4, Fig. S2-3).
-
-# Discussion
-
-<!-- original opening paragraph of discussion; I think it provides too good of a summary though to put at the beginning. It makes subsequent discussion sound redundant
-
-My results show that character displacement leads to predictable differences in resource abundances and the resilience of food webs to perturbations.
-Specifically, I found that, whenever character displacement increased the effective attack rate of consumers, it resulted in lower resource abundances and less resilient food webs. 
-Character displacement always increased attack rates whenever consumers were competing for resources that occured in distinct habitats. 
-This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
-Although character displacement did not always increase effective attack rates, these contingent effects only occurred under the simplest and (arguably) least realistic foraging scenario. 
-My study takes advantage of past work showing that adaptive divergence in consumer attack rates is always favored under the scenarios examined here [@LawlorSmith1976; @Abrams1986].
-Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations. 
--->
-
-## Resource abundances
-
-One of the criteria used to demonstrate ecological character displacement is that ``sites of sympatry [two consumers] and allopatry [one consumer] should not differ greatly in food $\textbf{[}$resource abundances$\textbf{]}$'' [@Schluter1992].
-In contrast, my results indicate that character displacement causes predictable differences in resource abundances.
-In fact, the ecological and evolutionary scenarios that favored the largest character displacement always decreased the relative abundance of resources.
-<!-- build paragraph around this topic sentence -->
-For example, if mobile consumers compete for resources that occur in different habitats, then character displacement always resulted in lower resource abundances.<!--the magnitude of character displacement was highest when mobile consumers competed for resources that occur in different habitats.-->
-Threespine stickleback, one of the classic examples of character displacement, exemplify this foraging scenario [@Schluter1992; @Schluter2000]. 
-Stickleback must move between the pelagic and littoral zones of a lake when foraging for zooplankton and benthic invertebrates, respectively.
-The theory developed here predicts that resource abundances will be lower in lakes where competing stickleback have undergone character displacement compared to lakes with only a single species of stickleback.
-Interestingly, a disproportionate number of the documented cases of character displacement involve carnivores [@Schluter2000] that are larger, and likely more mobile, than their resources [@McCann2005], suggesting that many cases of ecological character displacement may result in lower resource availability.
-
-<!-- older version: Threespine stickleback forage for two resource types, zooplankton and benthic invertebrates, which occur in distinct zones of the lake (pelagic and littoral)
-An empirical example of this is threespine stickleback that have diverged into limnetic and benthic species with specialized traits to forage on zooplankton and benthic invertebrates, respectively [@Schluter1992; @Schluter2000].
-These two resource types occur in distinct zones of the lake (pelagic and littoral), therefore stickleback must move between these zones when foraging for resources. 
-The theory developed here predicts that resource abundances will be lower in lakes with compe-->
-
-Similarly, the evolutionary tradeoff that favored character displacement decreased resource availability across all foraging scenarios.
-Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the net foraging rate of consumers [@Schluter1985; @Arnegard2014].
-While it is theoretically possible that character displacement does not alter (or even increase) resource abundances, this was limited to the simplest, and arguably least realistic, foraging scenario and under tradeoffs that did not favor large displacements, and thus less likely to detect in nature.
-Taken together, my results call for empirical work to test these clear theoretical predictions and suggest a revision is needed for one of the criteria used to demonstrate character displacement.
-
-<!-- consider removing entirely. This is a prediction I could test with Seth in another paper...
-Character displacement is far from an all-or-nothing process, but may vary in its magnitude [@Schluter2000]. 
-My results suggest that the magnitude of character displacement will be negatively correlated with total resource abundances, at least when resources occur in distinct habitats (line trajectories in \ref{fig:LS_Plot}B, \ref{fig:McCann_Plot}B).
-This prediction could be tested either experimentally, by putting populations that vary in divergence in a common resource setting, or with field data that statistically controls for other factors affecting resources.
-Note though, this prediction will only apply to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
--->
-
-## Food-web stability
-
-My most striking result was that ecological character displacement made food webs less resilient to perturbations. 
-In fact, under the most realistic foraging scenario, character displacement can even result in an unstable food web. 
-The mechanism underlying this destabilization is quite general.
-Character displacement generally increases the strength of consumer-resource interactions, but does not alter the strength of intraspecific interactions.
-This relative increase in interspecific interactions, combined with the natural oscillatory tendency of consumer-resource dynamics [@Lotka1925; @Volterra1926], creates a food-web structure that is less resilient to perturbations [@Chesson2008; @Rip2011; @McCann2011].
-
-Interestingly, the ecological conditions that favor character displacement are those that are already the least resilient to perturbations.
-For example, @McPeek2019 showed that character displacement is favored in food webs that are either highly productive, easy to find and capture resources, or under weak abiotic stress.
-This corresponds to higher values of $K$ (productivity) or $A$ (investment in attack rates), or lower values of $m$ (abiotic stress).
-Each of these corresponding changes decrease food-web resilience, as they increase the strength of consumer-resource interactions relative to intraspecific interactions.
-For example, increasing productivity reduces intraspecific competition in resource populations while increasing the flux of energy to consumers, resulting in the paradox of enrichment [@Rosenzweig1971].
-Similarly, higher feeding rates or lower consumer mortality both increase the relative strength of consumer-resource interactions, which predictably destabilizes food webs [@Rip2011; @McCann2011].
-This suggests that the most dramatic examples of character displacement will not only occur in, but also cause, the least stable food-web structures.
-
-A handful of empirical patterns support the hypothesis that character displacement decreases food-web resilience.
-For example, a single species of threespine stickleback lives in hundreds of small coastal lakes in British Columbia, but the species pair, where character displacement has resulted in specialized limnetic and benthic species, are only known from six lakes in four independent watersheds [@Schluter1992; @Schluter2000].
-Perhaps many lakes had a species pair in the past, but have lost a species due to a less resilient food-web structure [@Borrelli2015a; @Borrelli2015b].
-The species pair are known to be vulnerable to perturbations, as they have gone extinct in two of the six lakes after the introduction of nonnative species [@Hatfield2001; @Taylor2006; @Rudman2016].
-The vulnerability of the stickleback system also corresponds with the fact that aquatic food webs have several properties that make them less resilient to perturbations, such as higher productivity and more efficient energy transfer to consumers [@Rip2011].
-Detecting the ghost of competition past [@Connell1980] may be quite difficult, but it could be possible with recent advances in genomics.
-For example, @Feulner2018 detected genomic signatures of hybridization in sympatric whitefish species following periods of eutrophication.
-Perhaps solitary stickleback in some lakes retain genomic signatures of having been a habitat specialist in the past.
-
-My results contrast, but do not necessarily contradict, the notion from coexistence theory that character displacement contributes to species coexistence [@Lawlor1976]. 
-Rather than studying resilience, coexistence theory usually studies the mutual ability of consumers with different phenotypes to invade when rare [i.e., mutual invasibility; @Chesson2000].
-In the context of character displacement, a shortcoming of this mutual invasibility measure is that it does not allow a comparison between food webs with and without a competing consumer. 
-Such comparisons are necessary for inferring the effects of character displacement, a point that has been made clear in the criteria to demonstrate character displacement [@Schluter1992; @Schluter2000].
-Although the addition of a consumer to a food web can decrease its resilience in the absence of evolution [@May1973], my results are primarily driven by an eco-evolutionary feedback between consumer evolution and resource abundances.
-
-## Caveats
-
-Although I model the indirect effects of coevolution between consumers, I do not account for potential coevolution between consumers and resources. 
-In the context of my model, I would expect prey to evolve traits that reduce consumer attack rates. 
-Thus, prey evolution would act to counter the effects of character displacement on resource abundance and food-web stability. 
-Note that this does not negate my general conclusion that ecological character displacement decreases resource abundances and stability; however, this process may itself create another eco-evolutionary feedback between consumers and resources.
-This may actually help maintain dramatic examples of character displacement and prevent them from destabilizing systems, because it allows consumer traits to become decoupled from their attack rate.
-Examining this decoupling would be ideal in a quantitative genetic model that explicitly tracks trait dynamics, but it would not fundamentally change the conclusions presented here.
-
-Another potential caveat is that I explored my model in a setting that makes many assumptions about resource and consumer symmetry (but see consumer asymmetry section). 
-Prior work has shown that allowing for resource asymmetry, for example, may decrease the magnitude of character displacement [@Abrams1986]. 
-<!--This is because this asymmetry creates an asymmetry in resource abundances, which dampens the amount of character displacement.-->
-While this may dampen the amount of divergence, it should not qualitatively change the relationship I observed.<!-- of ecological character displacement decreasing resource abundances and food-web stability.-->
-
-I studied this eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
-A strength of this approach is that it enabled me to gain analytical insight to the effects of character displacement in a more realistic foraging scenario.
-This is much less tractable in quantitative genetic [@Taper1985; @McPeek2017] or explicit genetic [@Doebeli1996] models of character displacement, which is why the foraging scenarios previously examined have been limited [but see @McPeek2017].
-A weakness, however, is that I assume a separation of time scales between ecological and evolutionary dynamics, an assumption that is becoming less tenable [@Hairston2005; @Hendry2016].
-I also do not explicitly model an underlying phenotypic trait for consumer attack rates nor do I allow for intraspecific variation. 
-That being said, my theoretical predictions are likely robust to these assumptions. 
-This is because models that explicitly include resource dynamics inevitably show that resource competition results in character displacement, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
-A quantitative genetic model may certainly show differences in the pace of character displacement, but this should not qualitatively change its effect on food-web dynamics. 
-It is important to note that my conclusions only apply to food webs with biotic resources that are nutritionally substitutable. 
-It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
-
-## Conclusions
-
-Here, I show that an adaptive process that generates phenotypic diversity generally makes that diversity more susceptible to future extinctions.
-This destabilizing effect emerges from an eco-evolutionary feedback involving direct and indirect interactions between species in a food-web context.
-This result contrasts with the current notion that patterns of phenotypic diversity are solely the result of evolutionary constraints imposed by mutation, natural selection, gene flow, and genetic drift.
-In particular, my result supports the recent suggestion that food-web stability can impose an ecological constraint on phenotypic diversity that is agnostic to these evolutionary processes [@Borrelli2015b].
-I expect that identifying when and where this ecological constraint arises will yield novel insight to the patterns of biodiversity we see in nature.
-
-# Acknowledgements
-
-This work was inspired by discussions with Seth Rudman, Dolph Schluter, Ben Gilbert, and Kevin McCann.
-Sally Otto provided much help in early analyses of the mathematical model. 
-I would not have been able to take the theory as far as I did without her guidance and encouragement.
-Matt Osmond, Jean Gibert, and Seth Rudman provided critical feedback on an earlier draft.
-For funding support, I thank the University of British Columbia (Four-Year Fellowship to M.A.
-Barbour), NSERC (Discovery grant to Greg Crutsinger), and the Swiss National Science Foundation (grant 31003A_160671 to Jordi Bascompte).
-
-```{r MacArthur_asymm, include=FALSE}
-Mac_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  ylab(expression(a[ii]+a[ij])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Mac_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
-  ylab(expression(R[i]+R[j])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Stability_Mac_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + 
-  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability "(-lambda[max]))) + 
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Mac_plots_asymm <- plot_grid(Mac_attack_asymm + theme(legend.position = "none"), 
-                             Mac_resources_asymm   + theme(legend.position = "none"), 
-                             Stability_Mac_asymm   + theme(legend.position = "none"), 
-                       ncol=2, labels = "AUTO", align = 'hv')
-Mac_plots_asymm
-
-# Plot with LawlorSmith asymm in Appendix S4 (see "all asymmetry" chunk below)
-# save_plot(filename="Figure_ECD_MacArthur_asymm.pdf", plot=Mac_plots_asymm, base_height = 6, base_width = 6)
-```
-
-```{r LawlorSmith_asymm, include=FALSE}
-LS_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-LS_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  ylab(expression(R[i]+R[j])) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Stability_LS_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + 
-  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  ylab(expression("Stability "(-lambda[max]))) + 
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-LS_plots_asymm <- plot_grid(LS_attack_asymm   + theme(legend.position = "none"), 
-                             LS_resources_asymm   + theme(legend.position = "none"), 
-                             Stability_LS_asymm   + theme(legend.position = "none"), 
-                       ncol=2, labels = "AUTO", align='hv')
-LS_plots_asymm
-
-# Plot with MacArthur asymm in Appendix S4 (see "all asymmetry" chunk below)
-# save_plot(filename="Figure_ECD_LS_asymm.pdf", plot=LS_plots_asymm, base_height = 6, base_width = 6)
-```
-
-```{r all asymmetry, echo=FALSE}
-Macs_asymm <- plot_grid(
-  Mac_attack_asymm + #ggtitle("Resources: Same Habitat\n") + 
-    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))),
-  Mac_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)) + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))),
-  Stability_Mac_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))),
-  ncol=1, align='v', labels = c("","b)","c)"), hjust = -4, vjust = 0.5, label_size = 10
-)
-Macs_asymm_title <- ggdraw() + draw_label("Resources: Same Habitat", fontface = 'bold', size = 10, hjust = 0.2)
-plot_Macs_asymm <- plot_grid(Macs_asymm_title, Macs_asymm, ncol = 1, rel_heights = c(0.1, 1))
-
-LSs_asymm <- plot_grid(
-  LS_attack_asymm  + #ggtitle("Resources: Different Habitats\n") +  
-    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)) + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])),
-  LS_resources_asymm   + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x=element_blank()) + xlab("") + ylab(""),
-  Stability_LS_asymm   + theme(legend.position = "none", axis.text = element_text(size=10)) + scale_x_continuous(labels=c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(""),
-  ncol=1, align='v', labels = c("","e)","f)"), hjust = -1.5, vjust = 0.5, label_size = 10
-)
-LSs_asymm_title <- ggdraw() + draw_label("Resources: Different Habitat", fontface = 'bold', size = 10, hjust = 0.3)
-plot_LSs_asymm <- plot_grid(LSs_asymm_title, LSs_asymm, ncol = 1, rel_heights = c(0.1, 1))
-
-all_asymm_temp <- plot_grid(plot_Macs_asymm, plot_LSs_asymm, ncol=2, rel_widths = c(1,0.9)) 
-all_asymm_labels <- all_asymm_temp +
-  draw_label("a)", x = 0.098, y = 0.925, fontface = 'bold', size = 10) + 
-  draw_label("d)", x = 0.566, y = 0.925, fontface = 'bold', size = 10)
-all_asymm_plots <- add_sub(all_asymm_labels, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
-
-# plotted in Appendix S4, Fig. S2
-save_plot(filename="Appendix_S4_Fig_S2_Asymmetry_MacArthur_LawlorSmith.pdf", plot=all_asymm_plots, base_height = 6.5, base_width = 6)
-```
-
-```{r McCann_asymm, include=FALSE}
-McCann_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  ylab(bquote(atop("Effective attack rate","("*w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j]*")"))) +
-  xlab(expression(frac(a[ii],a[1*","*j]+a[2*","*j]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-McCann_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
-  ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) +
-  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
-  scale_shape_manual(values=c(24,21)) +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-Stability_McCann_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
-  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
-  geom_line() +
-  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + 
-  geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
-  xlab(expression(frac(a[ij],a[1*","*j]+a[2*","*j]))) +
-  ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + 
-  scale_shape_manual(values=c(24,21))  +
-  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
-  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
-  guides(fill = guide_legend(override.aes=list(shape=21)))
-
-McCann_plots_asymm <- plot_grid(McCann_attack_asymm   + theme(legend.position = "none") + xlab(""), 
-                             McCann_resources_asymm   + theme(legend.position = "none") + xlab(""), 
-                             Stability_McCann_asymm   + theme(legend.position = "none") + xlab(""), 
-                       ncol=2, align = 'hv', labels = c("a)","b)","c)"), hjust = -4, vjust = 1, label_size = 10)
-McCann_plots_asymm
-
-plot_McCann_asymm <- add_sub(McCann_plots_asymm, label = expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","*j]),")"))))
-
-# plotted in Appendix S4, Fig. S3
-save_plot(filename="Appendix_S4_Fig_S3_Asymmetry_McCann.pdf", plot=plot_McCann_asymm, base_height = 6, base_width = 6)
-```

### 2020-01-30 Figures_Edited.tex
omitted personal identifying information for double-blind peer review at Am Nat

% Edited: Removed code before and after figure section
% that was automatically knitted from the .Rmd
\begin{figure}
\centering
\includegraphics{Fig_1_ForagingScenarios}
\caption{\label{fig:foraging_scenarios}\textbf{Ecological foraging
scenarios.} I examined whether the effect of ecological character
displacement on food-web dynamics depended on whether consumers competed
for resources that occurred in the same (a) vs.~different habitats (b).
Note that inferences about character displacement can only be made by
comparing food webs with (right) and without (left) a competing
consumer, so I arbitrarily set \emph{C}\textsubscript{2} = 0 for these
comparisons. The width of each arrow corresponds to the initial attack
rate (\emph{a}\textsubscript{ij}) of consumer \emph{j} on resource
\emph{i}. Note that \emph{C}\textsubscript{1} was pre-adapted to
\emph{R}\textsubscript{1} (\emph{a}\textsubscript{11} \textgreater{}
\emph{a}\textsubscript{21}), while \emph{C}\textsubscript{2} was a
mirror image, being pre-adapted to \emph{R\textsubscript{2}}
(\emph{a}\textsubscript{22} = \emph{a}\textsubscript{11};
\emph{a}\textsubscript{12} = \emph{a}\textsubscript{21}). In each
scenario, I assumed consumer feeding rates increased linearly with
resource abundance. I also relax this assumption and consider a more
realistic functional response when resources occurred in different
habitats (b).}
\end{figure}

\begin{figure}
\centering
\includegraphics{Fig_2_Tradeoffs.pdf}
\caption{\label{fig:tradeoff}\textbf{Evolutionary tradeoffs in consumer
attack rates.} In each foraging scenario, I explored the effects of
three different tradeoffs: intermediate combinations of attack rates
(\emph{a}\textsubscript{1,\emph{j}}, \emph{a}\textsubscript{2,\emph{j}})
are higher than the extremes (green line, \emph{n} \textgreater{} 1);
extreme combinations of attack rates are higher than intermediate
investments (orange line, \emph{n} \textless{} 1); and all combinations
of attack rates have the same total attack rate (blue line, \emph{n} =
1). Points corresponding to attack rates at the beginning of the
simulation for \emph{C}\textsubscript{1}, which was pre-adapted to
\emph{R}\textsubscript{1} (\emph{a}\textsubscript{11} \textgreater{}
\emph{a}\textsubscript{12}). Note that \emph{C}\textsubscript{2} was a
mirror image of \emph{C}\textsubscript{1}, being pre-adapted to
\emph{R}\textsubscript{2} (\emph{a}\textsubscript{22} =
\emph{a}\textsubscript{11}; \emph{a}\textsubscript{12} =
\emph{a}\textsubscript{21}).}
\end{figure}

\begin{figure}
\centering
\includegraphics{Fig_3_MacArthur_LawlorSmith.pdf}
\caption{\label{fig:plot_fig3}\textbf{Effect of character displacement
on food-web dynamics under different evolutionary tradeoffs and foraging
scenarios.} Lines show predicted values when both consumers and
resources are present. Different line colors correspond to different
tradeoffs in attack rates (green, \emph{n} = 1.15; blue, \emph{n} = 1;
orange, \emph{n} = 0.85). Large circles (two consumers) and triangles
(one consumer) correspond to the end points of the eco-evolutionary
simulation for \emph{C}\textsubscript{1} (the choice to display
\emph{C}\textsubscript{1} was arbitrary), whereas as small shapes
correspond to the starting points (only in stability panels). In both
foraging scenarios, feeding rates increase linearly with resource
abundance, but the equation for the effective attack rate is different.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Fig_4_Bifurcation.pdf}
\caption{\label{fig:bifur_plot}\textbf{Character displacement creates an
unstable food web}. Lines illustrate the effect of character
displacement across the range of specialization for
\emph{C}\textsubscript{1} (the choice to display
\emph{C}\textsubscript{1} was arbitrary), while the points are the
results of an eco-evolutionary simulation. Note that I increased the
total investment in attack rates (\emph{A} = 3.3) to create a scenario
that could result in an unstable food web. Although I specified a linear
tradeoff in attack rates for this simulation, different tradeoff shapes
do not qualitatively alter these results (see Appendix S4, Fig. S1).}
\end{figure}

### 2020-01-30 manuscript_ECD_model.Rmd
omitted personal identifying information for double-blind peer review at Am Nat

--- 2019-11-18 manuscript_ECD_model.Rmd
@@
     citation_package: natbib 
+    #includes:
+    #  after_body: Figures_Edited.tex
 header-includes:
   - \usepackage{fullpage}
-  - \linespread{1.5} 
+  - \linespread{1.7} 
   - \usepackage{lineno}
-  - \usepackage{caption,setspace}
+  - \usepackage{caption}
   - \captionsetup[figure]{font=small}
-bibliography: ReferencesAbbrev
-biblio-style: EcolLett
+bibliography: ReferencesAbbrev.bib
+biblio-style: amnat
 fontsize: 11pt
@@
 
+<!-- remove for double-blind peer-review
 \bigskip
@@
 
-*Running title*: Character displacement destabilizes food webs
+<!--*Running title*: Character displacement destabilizes food webs-->
 
@@
 
+*Manuscript type*: E-Article.
+
+\bigskip
+
+*Manuscript elements*: Figures 1-4; Online appendices S1-4 (including figures S1-3 in appendix S4).
+
+\bigskip
+
 *Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptation; community stability
 
-<!-- Remove rest of title page for bioRxiv submission
-\bigskip
-
-*Manuscript type*: Letter.
-
-\bigskip
-
-*Number of words in abstract*: 143; 
-*Number of words in main text*: `r wordcountaddin:::word_count("manuscript_ECD_model.Rmd") - 143`;
-*Number of words in text boxes*: NA;
-*Number of references*: 45;
-*Number of figures*: 4;
-*Number of tables*: 0;
-*Number of text boxes*: 0
-
 <!-- End title page formatting -->
@@
 \linenumbers{}
-\modulolinenumbers[3]
-
-\newpage
+\modulolinenumbers[2]
 
@@
 cbPalette <- c("#D55E00", "#0072B2", "#009E73")
+
+## set plot theme
+theme_set(theme_cowplot())
 ```
@@
 I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
-All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model) and has been archived with Zenodo (https://zenodo.org/badge/latestdoi/79349192).
+All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub <!--(https://github.com/mabarbour/ECD_model) -->and has been archived with Zenodo <!-- (https://zenodo.org/badge/latestdoi/79349192)-->(please contact journal office for links).
 
@@
   ylab(expression(bold(a[21]))) + 
-  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette)
+  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette) + theme_cowplot()
 
@@
   Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) +
+    theme_cowplot() +
     theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), 
-  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)),
-  Stability_Mac + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)),
+  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + 
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)), 
+  Stability_Mac + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + 
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)), 
   ncol=1, align='v', labels = c("","b)","c)"), hjust = -4, vjust = 0.5, label_size = 10
@@
   LS_attack  + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])) + 
+    theme_cowplot() +
     theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), 
-  LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()),   
-  Stability_LS + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10)), 
+  LS_resources + xlab("") + ylab("") +
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()),   
+  Stability_LS + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab("") + 
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10)), 
   ncol=1, align='v', labels = c("","e)","f)"), hjust = -1.5, vjust = 0.5, label_size = 10
@@
 
+<!-- Remove for double-blind review
 # Acknowledgements
@@
 Barbour), NSERC (Discovery grant to Greg Crutsinger), and the Swiss National Science Foundation (grant 31003A_160671 to Jordi Bascompte).
+-->
 

### 2020-01-30 manuscript_ECD_model.tex
omitted personal identifying information for double-blind peer review at Am Nat

--- 2019-11-18 manuscript_ECD_model.tex
@@
 \IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage{hyperref}
+\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
+\usepackage[unicode=true]{hyperref}
 \PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\hypersetup{unicode=true,
+\hypersetup{
             colorlinks=true,
@@
 \usepackage{natbib}
-\bibliographystyle{EcolLett}
+\bibliographystyle{amnat}
 \usepackage{graphicx,grffile}
@@
 
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{}
-    \pretitle{\vspace{\droptitle}}
-  \posttitle{}
-    \author{}
-    \preauthor{}\postauthor{}
-    \date{}
-    \predate{}\postdate{}
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
 \usepackage{fullpage}
-\linespread{1.5}
+\linespread{1.7}
 \usepackage{lineno}
-\usepackage{caption,setspace}
+\usepackage{caption}
 \captionsetup[figure]{font=small}
+
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 
-\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}
+\emph{Manuscript type}: E-Article.
 
@@
 
-\noindent 1. University of Zurich, Department of Evolutionary Biology
-and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich,
-Switzerland;
-
-\noindent 2. University of British Columbia, Department of Zoology, 2212
-Main Mall, Vancouver, BC V6T 1Z4, Canada;
-
-\(^\ast\) Corresponding author; e-mail:
-\href{mailto:matthew.barbour@ieu.uzh.ch}{\nolinkurl{matthew.barbour@ieu.uzh.ch}},
-phone: +41 44 635 61 63
-
-\bigskip
-
-\emph{Statement of authorship}: MAB conceived the study, derived
-analytical solutions, performed simulations, and wrote the paper.
-
-\bigskip
-
-\emph{Data accessibility}: No new data were used. All code used to
-reproduce this manuscript and simulations within are publically
-available on GitHub (\url{https://github.com/mabarbour/ECD_model}).
-
-\bigskip
-
-\emph{Running title}: Character displacement destabilizes food webs
+\emph{Manuscript elements}: Figures 1-4; Online appendices S1-4
+(including figures S1-3 in appendix S4).
 
@@
 
-\linenumbers{} \modulolinenumbers[3]
-
-\newpage
+\linenumbers{} \modulolinenumbers[2]
 
@@
 in R \citep{R} and the code to reproduce these simulations is publically
-available on GitHub (\url{https://github.com/mabarbour/ECD_model}).
+available on GitHub and has been archived with Zenodo (please contact
+journal office for links).
 
@@
 
-\section{Acknowledgements}\label{acknowledgements}
-
-This work was inspired by discussions with Seth Rudman, Dolph Schluter,
-Ben Gilbert, and Kevin McCann. Sally Otto provided much help in early
-analyses of the mathematical model. I would not have been able to take
-the theory as far as I did without her guidance and encouragement. Matt
-Osmond, Jean Gibert, and Seth Rudman provided critical feedback on an
-earlier draft. For funding support, I thank the University of British
-Columbia (Four-Year Fellowship to M.A.~Barbour), NSERC (Discovery grant
-to Greg Crutsinger), and the Swiss National Science Foundation (grant
-31003A\_160671 to Jordi Bascompte).
-
-\bibliography{ReferencesAbbrev}
+\bibliography{ReferencesAbbrev.bib}
 

### 2020-07-02 manuscript_ECD_model.Rmd
incorporate reviewer and editor comments.

@@
 where $a_{ij}$ is the attack rate of consumer $j$ on resource $i$. 
-This first scenario is the starting point for many models of resource competition [@MacArthur1972]; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005].
+This first scenario is the starting point for many models of resource competition [@MacArthur1972]; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@Holt1984].
 The second scenario accounts for this spatial context (Fig. \ref{fig:foraging_scenarios}b) and takes the form:
@@
 This stability analysis derives the dominant eigenvalue, $\lambda_{\text{max}}$, of the matrix of partial derivatives of each species' population growth rate (given by equation \ref{eq:1}) with respect to each species' abundance evaluated at equilibrium.
-If $-\lambda_{max}>0$, then the food web will return to equilibrium after a small perturbation (i.e., it is locally stable), with more positive values indicating a faster return time.
-If $-\lambda_{max}<0$, then the food web is not locally stable.
+All eigenvalues must be negative for the food web to be locally stable; however, the dominant eigenvalue, $\lambda_{\text{max}}$, is the one closest to positive infinity and ultimately determines food-web stability.
+If $\lambda_{\text{max}}>0$, then the food web is not locally stable. 
+If $\lambda_{\text{max}}<0$, then the food web will return to equilibrium after a small perturbation (i.e., it is locally stable), with more negative values indicating a faster return time (i.e., greater stability).
+I chose to plot $-\lambda_{\text{max}}$ rather than $\lambda_{\text{max}}$, so that increasing values would indicate greater stability, which I felt was a more intuitive way to present the results.
+<!-- old test: If $-\lambda_{max}>0$, then the food web will return to equilibrium after a small perturbation (i.e., it is locally stable), with more positive values indicating a faster return time.
+If $-\lambda_{max}<0$, then the food web is not locally stable.-->
 
@@
 I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g., $\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0).
+
+<!--As with any model, the Adaptive Dynamics approach makes certain assumptions.
+First, it assumes a separation of time scales between ecological and evolutionary dynamics, an assumption that is becoming less tenable [@Hairston2005; @Hendry2016].
+Second, it assumes there is no intraspecific variation within interacting species, despite the fact that intraspecific variation is a common feature of natural populations and often has important ecological consequences [@Bolnick2011; @Des_Roches2018].-->
+
+The Adaptive Dynamics approach I use here is one of several possible approaches for studying the eco-evolutionary dynamics of interacting species.
+A strength of this approach is that it enables me to gain analytical insight to the effects of character displacement in more realistic foraging scenarios.
+This is much less tractable in quantitative genetic [@Taper1985; @McPeek2017] or explicit genetic [@Doebeli1996] models of character displacement, which is why the foraging scenarios previously examined have been limited [but see @McPeek2017].
+A weakness, however, is that I assume a separation of time scales between ecological and evolutionary dynamics, an assumption that is becoming less tenable [@Hairston2005; @Hendry2016]. 
+In addition, I assume that there is no intraspecific variation within interacting species, despite the fact that intraspecific variation is a common feature of natural populations and often has important ecological consequences [@Bolnick2011; @Des_Roches2018].
+Still, models that explicitly include resource dynamics inevitably show that resource competition results in character displacement, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
+Therefore, it should still serve as a good starting point for studying the ecological consequences of character displacement (but see \textbf{Caveats} section in the discussion).
+
 Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and $w_{11} = w_{22} = 0.6$. 
@@
 \ref{fig:plot_fig3}a), which dampens these contingent effects.
+Note that the dip in stability toward a dominant eigenvalue of zero is an artifact of the two consumers becoming ecologically identical ($a_{11}=a_{21}=a_{12}=a_{22}$) on identical resources, and effectively become neutral species [@McPeek2019Neutral].
+Although neutrality is an ecological possibility, it is never favored by evolution in these models since character displacement occurred across all the foraging scenarios I examined.<!-- never occurs this situation is not favored in any of the foraging scenarios I examined.
 Note that the dip in stability occurs when both consumers evolve to be generalists, a situation that is not favored in any of the foraging scenarios I examined (Fig. \ref{fig:plot_fig3}c).
+In fact, if the consumers were ecologically identical ($a_{1,j}=a_{2,j}$), then they are neutral species, as indicated by the dominant eigenvalue becoming zero [@McPeek2019Neutral]. 
+This only occurs -->
 
@@
 For example, when the tradeoff is concave up (orange), the initial two-consumer food web (small circle) is more stable than when there is only one consumer (small triangle); however, this pattern switches by the end of the eco-evolutionary simulation (large points).
+As with the previous model, the dip in stability corresponds to the situation where consumers are effectively identical ($w_{11}a_{11}=w_{21}a_{21}=w_{12}a_{12}=w_{22}a_{22}$) and become neutral species, which is never favored in these models.
+<!--Note that consumer neutrality only becomes possible with certain specific combinations of parameters in these models (e.g., $a_{1,j}=a_{2,j}$ or $w_{1,j}a_{1,j}=w_{2,j}a_{2,j}$).-->
+
+It is worth noting that the scales are different between the left and right columns of Figure \ref{fig:plot_fig3}.
+When resources occur in different habitats (relative to the same habitat), the consumer's effective attack rate is lower because it is multiplied by the proportion $w_{ij}$.
+This results in relatively higher resource abundances and more stable food webs compared to when resources occur in the same habitat, and corresponds to the inherent stabilizing role of spatial heterogeneity in food webs [@Holt1984; @McCann2005].
 
@@
 Different line colors correspond to different tradeoffs in attack rates (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85).
-Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~ (the choice to display *C*~1~ was arbitrary), whereas as small shapes correspond to the starting points (only in stability panels).
+Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation for *C*~1~ (the choice to display *C*~1~ was arbitrary), whereas small shapes correspond to the starting points (only in stability panels).
 In both foraging scenarios, feeding rates increase linearly with resource abundance, but the equation for the effective attack rate is different."}
 knitr::include_graphics("Fig_3_MacArthur_LawlorSmith.pdf")
+```
+
+```{r normalize-fig-3, include=FALSE}
+## setup normalization
+# from https://www.r-bloggers.com/custom-axis-transformations-in-ggplot2/
+require(scales) # trans_new() is in the scales library
+norm_Mac <- bind_rows(MacArthur_2C_2R_tradeoff_dynamics, MacArthur_1C_2R_tradeoff_dynamics) %>%
+  mutate(eff_a = a11 + a12,
+         RT = R1 + R2,
+         Stability = -1*max.Re.eigen)
+# easiest way to manually do this in the aesthetics
+
+norm_LS <- bind_rows(LawlorSmith_2C_2R_tradeoff_dynamics, LawlorSmith_1C_2R_tradeoff_dynamics) %>%
+  mutate(eff_a = w11*a11 + w12*a12,
+         RT = R1 + R2,
+         Stability = -1*max.Re.eigen)
+
+## normalize plots MacArthur
+Mac_attackNorm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=(aii+aij - min(norm_Mac$eff_a))/(max(norm_Mac$eff_a) - min(norm_Mac$eff_a)), color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=(a11+a12 - min(norm_Mac$eff_a))/(max(norm_Mac$eff_a) - min(norm_Mac$eff_a)), fill=factor(n), shape=Competitor), size=3, color="white") +
+  ylab(expression(a[ii]+a[ij])) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Mac_resourcesNorm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=(R1+R2 - min(norm_Mac$RT))/(max(norm_Mac$RT) - min(norm_Mac$RT)), color = factor(n))) + 
+  geom_line() +
+  geom_line(data = cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_1C_2R_tradeoff_dynamics), alpha = 0.1) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=(R1+R2 - min(norm_Mac$RT))/(max(norm_Mac$RT) - min(norm_Mac$RT)), fill=factor(n), shape=Competitor), size=3, color="white") +
+  ylab(expression(R[i]+R[j])) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_MacNorm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=(-1*max.Re.eigen - min(norm_Mac$Stability))/(max(norm_Mac$Stability) - min(norm_Mac$Stability)), color = factor(n))) + 
+  geom_line() +
+  geom_line(data = cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_1C_2R_tradeoff_dynamics), alpha = 0.1) +
+  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=(-1*max.Re.eigen - min(norm_Mac$Stability))/(max(norm_Mac$Stability) - min(norm_Mac$Stability)), fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) +
+  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=(-1*max.Re.eigen - min(norm_Mac$Stability))/(max(norm_Mac$Stability) - min(norm_Mac$Stability)), fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression("Stability "(-lambda[max]))) + 
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+MacsNorm <- plot_grid(
+  Mac_attackNorm + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) +
+    expand_limits(y = c(0,1)) +
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), 
+  Mac_resourcesNorm + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + 
+    expand_limits(y = c(0,1)) +
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)), 
+  Stability_MacNorm + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + 
+    expand_limits(y = c(0,1)) +
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)), 
+  ncol=1, align='v', labels = c("","b)","c)"), hjust = -4, vjust = 0.5, label_size = 10
+)
+#Macs_title <- ggdraw() + draw_label("Resources: Same Habitat", fontface = 'bold', size = 10, hjust = 0.2)
+plot_MacsNorm <- plot_grid(Macs_title, MacsNorm, ncol = 1, rel_heights = c(0.1, 1))
+
+## LawlorSmith
+LS_attackNorm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=(wii*aii+wij*aij - min(norm_LS$eff_a))/(max(norm_LS$eff_a) - min(norm_LS$eff_a)), color = factor(n))) + 
+  geom_line() +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=(w11*a11+w12*a12 - min(norm_LS$eff_a))/(max(norm_LS$eff_a) - min(norm_LS$eff_a)), fill=factor(n), shape=Competitor), size=3, color="white") +
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
+  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+LS_resourcesNorm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=(R1+R2 - min(norm_LS$RT))/(max(norm_LS$RT) - min(norm_LS$RT)), color=factor(n))) + 
+  geom_line() +
+  geom_line(data = cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_1C_2R_tradeoff_dynamics), alpha = 0.1) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=(R1+R2 - min(norm_LS$RT))/(max(norm_LS$RT) - min(norm_LS$RT)), fill=factor(n), shape=Competitor), size=3, color="white") +
+  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression(R[i]+R[j])) +
+  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
+  scale_shape_manual(values=c(24,21)) +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+Stability_LSNorm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
+  ggplot(., aes(x=aii/(aii+aij), y=(-1*max.Re.eigen - min(norm_LS$Stability))/(max(norm_LS$Stability) - min(norm_LS$Stability)), color = factor(n))) + 
+  geom_line() +
+  geom_line(data = cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_1C_2R_tradeoff_dynamics), alpha = 0.1) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=(-1*max.Re.eigen - min(norm_LS$Stability))/(max(norm_LS$Stability) - min(norm_LS$Stability)), fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) +
+  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=(-1*max.Re.eigen - min(norm_LS$Stability))/(max(norm_LS$Stability) - min(norm_LS$Stability)), fill=factor(n), shape=Competitor), size=1.5, color="white") +
+  ylab(expression("Stability "(-lambda[max]))) + 
+  scale_shape_manual(values=c(24,21))  +
+  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
+  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
+  guides(fill = guide_legend(override.aes=list(shape=21)))
+
+LSsNorm <- plot_grid(
+  LS_attackNorm  + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])) + 
+    expand_limits(y = c(0,1)) +
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), 
+  LS_resourcesNorm + xlab("") + ylab("") +
+    expand_limits(y = c(0,1)) +
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()),   
+  Stability_LSNorm + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab("") + 
+    expand_limits(y = c(0,1)) +
+    theme_cowplot() +
+    theme(legend.position = "none", axis.text = element_text(size=10)), 
+  ncol=1, align='v', labels = c("","e)","f)"), hjust = -1.5, vjust = 0.5, label_size = 10
+)
+#LSs_title <- ggdraw() + draw_label("Resources: Different Habitat", fontface = 'bold', size = 10, hjust = 0.3)
+plot_LSsNorm <- plot_grid(LSs_title, LSsNorm, ncol = 1, rel_heights = c(0.1, 1))
+
+plot_fig3Norm <- plot_grid(plot_MacsNorm, plot_LSsNorm, ncol=2, rel_widths = c(1,0.9)) 
+plot_fig3_labelNorm <- plot_fig3Norm + 
+  draw_label("a)", x = 0.098, y = 0.925, fontface = 'bold', size = 10) + 
+  draw_label("d)", x = 0.566, y = 0.925, fontface = 'bold', size = 10)
+plot_fig3_subNorm <- add_sub(plot_fig3_labelNorm, expression(paste("Specialization ", bgroup("(",frac(a[ij],a[1*","*j]+a[2*","* j]),")"))), vjust = 0, size=12)
+
+save_plot(filename="Fig_3_MacArthur_LawlorSmith_Normalized.pdf", plot=plot_fig3_subNorm, base_height = 6.5, base_width = 6)
 ```
@@
 In this more realistic model, however, whether the food web is locally stable depends on consumer and resource parameters. 
-Specifically, I found that the two-consumer food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derived using Routh-Hurwitz criteria in Appendix S1):
+Specifically, I found that the two-consumer food web will become unstable<!--transition from having a locally stable equilibrium to a limit cycle--> under the following conditions (derived using Routh-Hurwitz criteria in Appendix S1):
 
@@
 save_plot(filename="Appendix_S4_Fig_S1_McCann.pdf", plot=plot_McCann_model, base_height = 6, base_width = 6)
+```
+
+```{r check-negative-eigenvalues, include=FALSE}
+# MacArthur
+filter(MacArthur_2C_2R_tradeoff_dynamics, a11/(a11+a12) == 0.5) %>%
+  select(a11,a12,max.Re.eigen) %>%
+  knitr::kable()
+
+filter(MacArthur_2C_2R_tradeoff_dynamics, max.Re.eigen == 0) %>%
+  select(a11,a12,max.Re.eigen) %>%
+  knitr::kable()
+
+with(MacArthur_2C_2R_tradeoff_dynamics, table(max.Im.eigen > 0)) %>% knitr::kable()
+
+with(MacArthur_1C_2R_tradeoff_dynamics, table(max.Im.eigen > 0)) %>% knitr::kable()
+
+# LawlorSmith
+filter(LawlorSmith_2C_2R_tradeoff_dynamics, w11*a11/(w11*a11+w12*a12) == 0.5) %>%
+  select(a11,a12,w11,w12,max.Re.eigen) %>%
+  knitr::kable()
+
+filter(LawlorSmith_2C_2R_tradeoff_dynamics, max.Re.eigen == 0) %>%
+  select(a11,a12,w11,w12,max.Re.eigen) %>%
+  knitr::kable()
+
+with(LawlorSmith_2C_2R_tradeoff_dynamics, table(max.Im.eigen > 0)) %>% knitr::kable()
+
+with(LawlorSmith_1C_2R_tradeoff_dynamics, table(max.Im.eigen > 0)) %>% knitr::kable()
+
+with(McCann_2C_2R_tradeoff_dynamics, table(max.Im.eigen > 0)) %>% knitr::kable()
+
+with(McCann_1C_2R_tradeoff_dynamics, table(max.Im.eigen > 0)) %>% knitr::kable()
 ```
@@
 
+<!-- stickleback paper about no differences in habitat: @Ormond2011.
+Interestingly, the species pair appears to be associated with lakes with low fish diversity...not clear with the trophic linkages are there -->
+
 <!-- original opening paragraph of discussion; I think it provides too good of a summary though to put at the beginning. It makes subsequent discussion sound redundant
@@
 
-One of the criteria used to demonstrate ecological character displacement is that ``sites of sympatry [two consumers] and allopatry [one consumer] should not differ greatly in food $\textbf{[}$resource abundances$\textbf{]}$'' [@Schluter1992].
+One of the criteria used to demonstrate ecological character displacement is that ``sites of sympatry [two consumers] and allopatry [one consumer] should not differ greatly in food $\textbf{[}$resource abundances$\textbf{]}$, climate, or other environmental features affecting the phenotype'' [@Schluter1992].
+<!-- added full criteria -->
 In contrast, my results indicate that character displacement causes predictable differences in resource abundances.
@@
 While it is theoretically possible that character displacement does not alter (or even increase) resource abundances, this was limited to the simplest, and arguably least realistic, foraging scenario and under tradeoffs that did not favor large displacements, and thus less likely to detect in nature.
-Taken together, my results call for empirical work to test these clear theoretical predictions and suggest a revision is needed for one of the criteria used to demonstrate character displacement.
-
-<!-- consider removing entirely. This is a prediction I could test with Seth in another paper...
-Character displacement is far from an all-or-nothing process, but may vary in its magnitude [@Schluter2000]. 
-My results suggest that the magnitude of character displacement will be negatively correlated with total resource abundances, at least when resources occur in distinct habitats (line trajectories in \ref{fig:LS_Plot}B, \ref{fig:McCann_Plot}B).
-This prediction could be tested either experimentally, by putting populations that vary in divergence in a common resource setting, or with field data that statistically controls for other factors affecting resources.
-Note though, this prediction will only apply to comparisons within species where there is likely little difference in the shape of the tradeoff among populations.
--->
+<!--Taken together, my results call for empirical work to test these clear theoretical predictions and suggest a revision is needed for one of the criteria used to demonstrate character displacement.-->
+
+<!--Testing the prediction that character displacement decreases resource abundances could be done either experimentally, by putting consumers that have or have not undergone character displacement in a common resource setting, or with field data that statistically controls for other factors affecting resources.
+
+Testing this prediction 
+The most straightforward prediction to test is that character displacement decreases resource abundances.-->
+
+My predictions could be tested with an experiment that puts consumers that have and have not undergone character displacement in a common resource setting and comparing their effect on resource abundances.
+One experiment with threespine stickleback has actually done this, although they focused more on the community composition of zooplankton and benthic invertebrates as well as ecosystem properties [@Harmon2009].
+Their reported data on average resource abundances support my prediction that character displacement in stickleback would decrease total resource abundances [compare totals for "BL" vs.
+"G" in Table 1 of @Harmon2009].
+For future tests, I want to emphasize that it is important to measure resource abundances over time, as I also predict that abundances will be more variable (see \textbf{Food-web stability} section).
+I specifically predict that character displacement would decrease time-averaged resource abundances. 
+An even more explicit test would be to apply a first-order multivariate autoregressive model to obtain equilibrium estimates of resource abundances [see equation 15 in @Ives2003].
+An advantage of using this type of statistical model is that one can also estimate the effective attack rate of consumers, which I predict to be increased by character displacement in most foraging scenarios.
+In designing an experiment to test this theory, it is important to keep in mind the amount of available space for consumers.
+For example, a mesocosm experiment may effectively constrain resources to be in the same habitat for the consumer, in which case I predict that the shape of the tradeoff in consumer attack rates will determine the impact of character displacement (see Fig.
+\ref{fig:plot_fig3}a-c).
+
+Although a well-designed experiment would be an ideal test, the same methods described above could be applied to observational data from the field. 
+In fact, field data may already be available that could be used to test the effects of character displacement [@Schluter2000]. 
+For example, the criterion that my theory questions is the third most frequently assessed of the six criteria [see "Environmental controls" in Figure 1 of @Schluter2000]. 
+The data from studies that measured resources could be used to test my prediction that character displacement decreases resource abundances. 
+Although my theory calls into question part of this criterion, I agree with its intention, which was to ensure that character displacement was not due to species evolving independently to different environments [@Schluter1992].
+<!--However, as I predict, resource abundances will likely differ considerably under character displacement.
+Perhaps the easiest revision to this criterion is to simply delete the word "food".-->
 
@@
 
+My results contrast, but do not necessarily contradict, the notion from coexistence theory that character displacement contributes to species coexistence [@Lawlor1976]. 
+Rather than studying resilience, coexistence theory usually studies the mutual ability of consumers with different phenotypes to invade when rare [i.e., mutual invasibility; @Chesson2000].
+In the context of character displacement, a shortcoming of this mutual invasibility measure is that it does not allow a comparison between food webs with and without a competing consumer. 
+Such comparisons are necessary for inferring the effects of character displacement, a point that has been made clear in the criteria to demonstrate character displacement [@Schluter1992; @Schluter2000].
+Although the addition of a consumer to a food web can decrease its resilience in the absence of evolution [@May1973], my results are primarily driven by an eco-evolutionary feedback between consumer evolution and resource abundances.
+
+<!--
+## Testable predictions
+
+<!--This theory is most applicable to situations where consumers compete for two discrete resources. Although there is certainly diversity within each resource type in natural systems,--> 
+<!--
+The theory presented here provides specific testable predictions for how character displacement will shape resource abundances and food-web stability.
+<!--First, we predict that changes in resource abundances will correspond with changes in stability.
+This is consistent with what we have observed here, and is a common feature of consumer-resource systems [@Murdoch2003].
+Character displacement is far from an all-or-nothing process, but may vary in its magnitude [@Schluter2000].-->
+<!--My results suggest that the magnitude of character displacement will be negatively correlated with total resource abundances, at least when resources occur in distinct habitats (line trajectories in Fig.
+\ref{fig:plot_fig3} and Fig.
+S1 in Appendix S4).
+These predictions could be tested either experimentally, by putting populations that vary in divergence in a common resource setting, or with field data that statistically controls for other factors affecting resources.
+Note though, this prediction will only apply to comparisons within species where there is likely little difference in the shape of the tradeoff among populations. -->
+
+Testing the effect of character displacement on food-web stability will be inherently more difficult than for resource abundances. 
+Fortunately, a common empirical metric of stability, the coefficient of variation in abundance (CV), corresponds to changes in the dominant eigenvalue of consumer-resource interactions [@Gellner2016TheorEcol].
+The coefficient of variation is an indicator of temporal variability and requires data on the mean ($\mu$) and standard deviation ($\sigma$) of population abundance over time ($\text{CV}=\sigma/\mu$).
+I predict that character displacement would increase the coefficient of variation (across most foraging scenarios) in each consumer and resource population. 
+As with resource abundances, a more explicit test <!--reliable, but also more data hungry approach, -->would be to apply multivariate autoregressive models to time series of consumer and resource abundances in order to directly estimate the dominant eigenvalue of the food web [see equation 22 in @Ives2003].
+
+<!-- It is worth noting that that there could be systematic differences in consumer abundance as well due to character displacement.
+In fact, character displacement will likely decrease consumer abundances (ref.
+Fig.
+), which is common in consumer-resource systems [@Abrams2019].
+However, I expect that character displacement will increase consumer-resource ratios, which may be an indicator of a less stable system [].-->
+<!--
+In addition, I expect that the magnitude of character displacement will be positively correlated with temporal variability in consumer and resource abundances.
+There are two measures of temporal variability that explicitly align with the metric of food-web stability used here ($-\lambda_{max}$)--- dominant eigenvalue of a stochastic system and the coefficient of variation in consumer and resource abundances.
+First, the dominant eigenvalue of a food web can be estimated directly from time series of interacting populations [@Ives2003].
+The dominant eigenvalue has a similar interpretation for both deterministic and stochastic models [@Ives2003], and thus could be used to test the theory.
+Second, changes in the coefficient of variation match deterministic changes in the dominant eigenvalue of a food web.
+Therefore, we expect a larger coefficient of variation in abundances with increasing character displacement.
+Note though, the changes in the coefficient of variation can be due to changes in both the mean and variability of species abundances, both of which we expect character displacement to alter.
+This theory explicitly predicts, however, that the increased variability will be driven by increased variance.
+In fact, we expect mean driven changes in stability to not occur in the context of character displacement (regions where stability dips toward zero in \ref{fig:plot_fig3}).
+This increased variability increases the risk that species will reach extremely low densities where they may collapse.
+Apparently, the "environmental controls" criteria is the third most frequently assessed of the six criteria [see Figure 1 in @Schluter2000].
+For those studies that measured resources, these data could be used to test my prediction that character displacement decreases resource abundances.
+I do agree with the initial spirit of this criteria, which was to ensure that character displacement was not due to species evolving independently to different environments [@Schluter1992].<!-- Perhaps the easiest revision to this criterion is to delete "food".-->
+
+<!--
+Also, detecting these changes may be easier for examples of character displacement in aquatic systems, like with threespine stickleback, given that aquatic food webs have several properties that make them less resilient to perturbations, such as higher productivity and more efficient energy transfer to consumers [@Rip2011].
+
+<!--
+In two of the three models we studied, food webs always remained locally stable.
+If the food-web is locally stable, then what does a less negative eigenvalue mean biologically?
+I argue that this comes into play in particular for stochastic systems.
+In stochastic systems, populations are always vulnerable to extinction, whereas this is not the case in deterministic models.
+Recent theory developed for food webs has shown a nice parallel between theory developed for deterministic and stochastic models in food webs.
+In particular, they have linked the classic checkmark pattern in stability, which we similarly observe, corresponds with the coefficient of variation in species abundances [@Gellner2016TheorEcol].
+The coefficient of variation is a common empirical metric of stability, and it is nice to see that it has this close correspondence with the dominant eigenvalue of a consumer-resource system, and potentially even more complex food webs [@Gellner2016NatComm].
+Therefore, we also expect that instability due to character displacement will be driven by variability in species abundances, rather than the mean.
+Both contribute to extinction risk at different parts of the check-mark pattern in stability [@Gellner2016TheorEcol].
+However, it is also possible to estimate dominant eigenvalue of a system directly through analyses of community time series of interacting populations [@Ives2003].
+Both approaches could be used to test our theoretical predictions.
+
+testable predictions:
+- instability due to character displacement will be driven by variability in species abundances, rather than the mean.
+Both contribute to extinction risk at different parts of the check-mark pattern in stability [@Gellner2016TheorEcol].
+- in stickleback: we would expect greater temporal variability in both consumer and resource (zooplankton and benthic invertebrates) abundances, relative to its mean, in lakes where stickleback have undergone character displacements compared to those where it has not.
+The stickleback system would be particularly nice test, given that lakes with and without character displacement are nearby in proximity and do not different from each other in edaphic characteristics [CITE].
+- We also expect that the magnitude of character displacement will correspondence to the magnitude of instability. 
+-->
+
+<!-- Or do I keep?
 A handful of empirical patterns support the hypothesis that character displacement decreases food-web resilience.
@@
 Perhaps solitary stickleback in some lakes retain genomic signatures of having been a habitat specialist in the past.
-
-My results contrast, but do not necessarily contradict, the notion from coexistence theory that character displacement contributes to species coexistence [@Lawlor1976]. 
-Rather than studying resilience, coexistence theory usually studies the mutual ability of consumers with different phenotypes to invade when rare [i.e., mutual invasibility; @Chesson2000].
-In the context of character displacement, a shortcoming of this mutual invasibility measure is that it does not allow a comparison between food webs with and without a competing consumer. 
-Such comparisons are necessary for inferring the effects of character displacement, a point that has been made clear in the criteria to demonstrate character displacement [@Schluter1992; @Schluter2000].
-Although the addition of a consumer to a food web can decrease its resilience in the absence of evolution [@May1973], my results are primarily driven by an eco-evolutionary feedback between consumer evolution and resource abundances.
+-->
 
@@
 
+<!--
 I studied this eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
@@
 It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
+-->
+<!--
+I studied this eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
+A strength of this approach is that it enabled me to gain analytical insight to the effects of character displacement in a more realistic foraging scenario.
+This is much less tractable in quantitative genetic [@Taper1985; @McPeek2017] or explicit genetic [@Doebeli1996] models of character displacement, which is why the foraging scenarios previously examined have been limited [but see @McPeek2017].
+A weakness, however, is that I assume a separation of time scales between ecological and evolutionary dynamics, an assumption that is becoming less tenable [@Hairston2005; @Hendry2016].
+Still, models that explicitly include resource dynamics inevitably show that resource competition results in character displacement, regardless of whether a quantitative genetic or Adaptive Dynamics approach is used [@Lawlor1976; @Taper1985].
+-->
+
+<!-- addressing Dan's critique -->
+One assumption that warrants further consideration is that I do not allow for intraspecific variation.
+Models of character displacement that allow trait variation to evolve actually predict that consumers will exhibit more trait variation in the absence of a competitor [@Taper1985].
+On the one hand, increased trait variation would impose a greater fitness load for the single consumer, which would weaken its top-down effect on resources [@Schreiber2011]. 
+This would act to magnify my predicted effects of character displacement on resource abundances and food-web stability. 
+On the other hand, if the increased trait variance has a large heritable component, then rapid evolution may generate oscillatory or chaotic dynamics in the single consumer food web [@Schreiber2011], which could negate my predicted effects of character displacement on food-web stability.
+Finally, it is important to note that my conclusions only apply to food webs with biotic resources that are nutritionally substitutable. 
+It would be interesting to extend these current analyses to non-substitutable resources where convergent character displacement is expected [@Abrams1987; @Fox2008].
 

### 2020-07-02 manuscript_ECD_model.tex
incorporate reviewer and editor comments.

@@
 food webs where consumers are mobile and their foraging behavior links
-resources that occur in different habitats \citep{McCann2005}. The
-second scenario accounts for this spatial context (Fig.
+resources that occur in different habitats \citep{Holt1984}. The second
+scenario accounts for this spatial context (Fig.
 \ref{fig:foraging_scenarios}b) and takes the form:
@@
 \ref{eq:1}) with respect to each species' abundance evaluated at
-equilibrium. If \(-\lambda_{max}>0\), then the food web will return to
-equilibrium after a small perturbation (i.e., it is locally stable),
-with more positive values indicating a faster return time. If
-\(-\lambda_{max}<0\), then the food web is not locally stable.
+equilibrium. All eigenvalues must be negative for the food web to be
+locally stable; however, the dominant eigenvalue,
+\(\lambda_{\text{max}}\), is the one closest to positive infinity and
+ultimately determines food-web stability. If \(\lambda_{\text{max}}>0\),
+then the food web is not locally stable. If \(\lambda_{\text{max}}<0\),
+then the food web will return to equilibrium after a small perturbation
+(i.e., it is locally stable), with more negative values indicating a
+faster return time (i.e., greater stability). I chose to plot
+\(-\lambda_{\text{max}}\) rather than \(\lambda_{\text{max}}\), so that
+increasing values would indicate greater stability, which I felt was a
+more intuitive way to present the results.
 
@@
 \(\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}\) is constrained to a
-maximum of 1 and minimum of 0). Unless otherwise noted, I conducted
-simulations with the following parameter values: \emph{r} = 1; \emph{K}
-= 4; \emph{e} = 0.8; \emph{m} = 1; \emph{A} = 2; \emph{h} = 0.4; and
-\(w_{11} = w_{22} = 0.6\). I set an initial value of
-\(a_{11} = a_{22} = 1.2\), while \(a_{12}\) and \(a_{21}\) depended on
-the value of \emph{n}. I set initial consumer and resource abundances
-to: \(R_1 = R_2 = 2\); \(C_1 = C_2 = 1\). All simulations were conducted
-in R \citep{R} and the code to reproduce these simulations is publically
-available on GitHub and has been archived with Zenodo (please contact
-journal office for links).
+maximum of 1 and minimum of 0).
+
+The Adaptive Dynamics approach I use here is one of several possible
+approaches for studying the eco-evolutionary dynamics of interacting
+species. A strength of this approach is that it enables me to gain
+analytical insight to the effects of character displacement in more
+realistic foraging scenarios. This is much less tractable in
+quantitative genetic \citep{Taper1985, McPeek2017} or explicit genetic
+\citep{Doebeli1996} models of character displacement, which is why the
+foraging scenarios previously examined have been limited \citep[but
+see][]{McPeek2017}. A weakness, however, is that I assume a separation
+of time scales between ecological and evolutionary dynamics, an
+assumption that is becoming less tenable
+\citep{Hairston2005, Hendry2016}. In addition, I assume that there is no
+intraspecific variation within interacting species, despite the fact
+that intraspecific variation is a common feature of natural populations
+and often has important ecological consequences
+\citep{Bolnick2011, Des_Roches2018}. Still, models that explicitly
+include resource dynamics inevitably show that resource competition
+results in character displacement, regardless of whether a quantitative
+genetic or Adaptive Dynamics approach is used
+\citep{Lawlor1976, Taper1985}. Therefore, it should still serve as a
+good starting point for studying the ecological consequences of
+character displacement (but see \textbf{Caveats} section in the
+discussion).
+
+Unless otherwise noted, I conducted simulations with the following
+parameter values: \emph{r} = 1; \emph{K} = 4; \emph{e} = 0.8; \emph{m} =
+1; \emph{A} = 2; \emph{h} = 0.4; and \(w_{11} = w_{22} = 0.6\). I set an
+initial value of \(a_{11} = a_{22} = 1.2\), while \(a_{12}\) and
+\(a_{21}\) depended on the value of \emph{n}. I set initial consumer and
+resource abundances to: \(R_1 = R_2 = 2\); \(C_1 = C_2 = 1\). All
+simulations were conducted in R \citep{R} and the code to reproduce
+these simulations is publically available on GitHub and has been
+archived with Zenodo (please contact journal office for links).
 
@@
 \ref{fig:plot_fig3}a), which dampens these contingent effects. Note that
-the dip in stability occurs when both consumers evolve to be
-generalists, a situation that is not favored in any of the foraging
-scenarios I examined (Fig. \ref{fig:plot_fig3}c).
+the dip in stability toward a dominant eigenvalue of zero is an artifact
+of the two consumers becoming ecologically identical
+(\(a_{11}=a_{21}=a_{12}=a_{22}\)) on identical resources, and
+effectively become neutral species \citep{McPeek2019Neutral}. Although
+neutrality is an ecological possibility, it is never favored by
+evolution in these models since character displacement occurred across
+all the foraging scenarios I examined.
 
@@
 this pattern switches by the end of the eco-evolutionary simulation
-(large points).
+(large points). As with the previous model, the dip in stability
+corresponds to the situation where consumers are effectively identical
+(\(w_{11}a_{11}=w_{21}a_{21}=w_{12}a_{12}=w_{22}a_{22}\)) and become
+neutral species, which is never favored in these models.
+
+It is worth noting that the scales are different between the left and
+right columns of Figure \ref{fig:plot_fig3}. When resources occur in
+different habitats (relative to the same habitat), the consumer's
+effective attack rate is lower because it is multiplied by the
+proportion \(w_{ij}\). This results in relatively higher resource
+abundances and more stable food webs compared to when resources occur in
+the same habitat, and corresponds to the inherent stabilizing role of
+spatial heterogeneity in food webs \citep{Holt1984, McCann2005}.
 
@@
 simulation for \emph{C}\textsubscript{1} (the choice to display
-\emph{C}\textsubscript{1} was arbitrary), whereas as small shapes
+\emph{C}\textsubscript{1} was arbitrary), whereas small shapes
 correspond to the starting points (only in stability panels). In both
@@
 stable depends on consumer and resource parameters. Specifically, I
-found that the two-consumer food web will transition from having a
-locally stable equilibrium to a limit cycle under the following
-conditions (derived using Routh-Hurwitz criteria in Appendix S1):
+found that the two-consumer food web will become unstable under the
+following conditions (derived using Routh-Hurwitz criteria in Appendix
+S1):
 
@@
 allopatry {[}one consumer{]} should not differ greatly in food
-\(\textbf{[}\)resource abundances\(\textbf{]}\)'' \citep{Schluter1992}.
+\(\textbf{[}\)resource abundances\(\textbf{]}\), climate, or other
+environmental features affecting the phenotype'' \citep{Schluter1992}.
 In contrast, my results indicate that character displacement causes
@@
 realistic, foraging scenario and under tradeoffs that did not favor
-large displacements, and thus less likely to detect in nature. Taken
-together, my results call for empirical work to test these clear
-theoretical predictions and suggest a revision is needed for one of the
-criteria used to demonstrate character displacement.
+large displacements, and thus less likely to detect in nature.
+
+My predictions could be tested with an experiment that puts consumers
+that have and have not undergone character displacement in a common
+resource setting and comparing their effect on resource abundances. One
+experiment with threespine stickleback has actually done this, although
+they focused more on the community composition of zooplankton and
+benthic invertebrates as well as ecosystem properties
+\citep{Harmon2009}. Their reported data on average resource abundances
+support my prediction that character displacement in stickleback would
+decrease total resource abundances \citep[compare totals for ``BL'' vs.
+``G'' in Table 1 of][]{Harmon2009}. For future tests, I want to
+emphasize that it is important to measure resource abundances over time,
+as I also predict that abundances will be more variable (see
+\textbf{Food-web stability} section). I specifically predict that
+character displacement would decrease time-averaged resource abundances.
+An even more explicit test would be to apply a first-order multivariate
+autoregressive model to obtain equilibrium estimates of resource
+abundances \citep[see equation 15 in][]{Ives2003}. An advantage of using
+this type of statistical model is that one can also estimate the
+effective attack rate of consumers, which I predict to be increased by
+character displacement in most foraging scenarios. In designing an
+experiment to test this theory, it is important to keep in mind the
+amount of available space for consumers. For example, a mesocosm
+experiment may effectively constrain resources to be in the same habitat
+for the consumer, in which case I predict that the shape of the tradeoff
+in consumer attack rates will determine the impact of character
+displacement (see Fig. \ref{fig:plot_fig3}a-c).
+
+Although a well-designed experiment would be an ideal test, the same
+methods described above could be applied to observational data from the
+field. In fact, field data may already be available that could be used
+to test the effects of character displacement \citep{Schluter2000}. For
+example, the criterion that my theory questions is the third most
+frequently assessed of the six criteria \citep[see ``Environmental
+controls'' in Figure 1 of][]{Schluter2000}. The data from studies that
+measured resources could be used to test my prediction that character
+displacement decreases resource abundances. Although my theory calls
+into question part of this criterion, I agree with its intention, which
+was to ensure that character displacement was not due to species
+evolving independently to different environments \citep{Schluter1992}.
 
@@
 
-A handful of empirical patterns support the hypothesis that character
-displacement decreases food-web resilience. For example, a single
-species of threespine stickleback lives in hundreds of small coastal
-lakes in British Columbia, but the species pair, where character
-displacement has resulted in specialized limnetic and benthic species,
-are only known from six lakes in four independent watersheds
-\citep{Schluter1992, Schluter2000}. Perhaps many lakes had a species
-pair in the past, but have lost a species due to a less resilient
-food-web structure \citep{Borrelli2015a, Borrelli2015b}. The species
-pair are known to be vulnerable to perturbations, as they have gone
-extinct in two of the six lakes after the introduction of nonnative
-species \citep{Hatfield2001, Taylor2006, Rudman2016}. The vulnerability
-of the stickleback system also corresponds with the fact that aquatic
-food webs have several properties that make them less resilient to
-perturbations, such as higher productivity and more efficient energy
-transfer to consumers \citep{Rip2011}. Detecting the ghost of
-competition past \citep{Connell1980} may be quite difficult, but it
-could be possible with recent advances in genomics. For example,
-\citet{Feulner2018} detected genomic signatures of hybridization in
-sympatric whitefish species following periods of eutrophication. Perhaps
-solitary stickleback in some lakes retain genomic signatures of having
-been a habitat specialist in the past.
-
 My results contrast, but do not necessarily contradict, the notion from
@@
 feedback between consumer evolution and resource abundances.
+
+Testing the effect of character displacement on food-web stability will
+be inherently more difficult than for resource abundances. Fortunately,
+a common empirical metric of stability, the coefficient of variation in
+abundance (CV), corresponds to changes in the dominant eigenvalue of
+consumer-resource interactions \citep{Gellner2016TheorEcol}. The
+coefficient of variation is an indicator of temporal variability and
+requires data on the mean (\(\mu\)) and standard deviation (\(\sigma\))
+of population abundance over time (\(\text{CV}=\sigma/\mu\)). I predict
+that character displacement would increase the coefficient of variation
+(across most foraging scenarios) in each consumer and resource
+population. As with resource abundances, a more explicit test would be
+to apply multivariate autoregressive models to time series of consumer
+and resource abundances in order to directly estimate the dominant
+eigenvalue of the food web \citep[see equation 22 in][]{Ives2003}.
 
@@
 
-I studied this eco-evolutionary feedback between consumers and resources
-using an Adaptive Dynamics approach. A strength of this approach is that
-it enabled me to gain analytical insight to the effects of character
-displacement in a more realistic foraging scenario. This is much less
-tractable in quantitative genetic \citep{Taper1985, McPeek2017} or
-explicit genetic \citep{Doebeli1996} models of character displacement,
-which is why the foraging scenarios previously examined have been
-limited \citep[but see][]{McPeek2017}. A weakness, however, is that I
-assume a separation of time scales between ecological and evolutionary
-dynamics, an assumption that is becoming less tenable
-\citep{Hairston2005, Hendry2016}. I also do not explicitly model an
-underlying phenotypic trait for consumer attack rates nor do I allow for
-intraspecific variation. That being said, my theoretical predictions are
-likely robust to these assumptions. This is because models that
-explicitly include resource dynamics inevitably show that resource
-competition results in character displacement, regardless of whether a
-quantitative genetic or Adaptive Dynamics approach is used
-\citep{Lawlor1976, Taper1985}. A quantitative genetic model may
-certainly show differences in the pace of character displacement, but
-this should not qualitatively change its effect on food-web dynamics. It
-is important to note that my conclusions only apply to food webs with
+One assumption that warrants further consideration is that I do not
+allow for intraspecific variation. Models of character displacement that
+allow trait variation to evolve actually predict that consumers will
+exhibit more trait variation in the absence of a competitor
+\citep{Taper1985}. On the one hand, increased trait variation would
+impose a greater fitness load for the single consumer, which would
+weaken its top-down effect on resources \citep{Schreiber2011}. This
+would act to magnify my predicted effects of character displacement on
+resource abundances and food-web stability. On the other hand, if the
+increased trait variance has a large heritable component, then rapid
+evolution may generate oscillatory or chaotic dynamics in the single
+consumer food web \citep{Schreiber2011}, which could negate my predicted
+effects of character displacement on food-web stability. Finally, it is
+important to note that my conclusions only apply to food webs with
 biotic resources that are nutritionally substitutable. It would be

### 2020-07-02 including_wij_in_aij.R

@@ -1,0 +1,328 @@
+## Response to reviewer #2
+
+# In this script, I address reviewer #2's question:
+# How would Fig. 3(d,e,f) change if eq. (2) was used with 
+# tilde(a)_ij=w_ij a_ij (and the trade-off applied to tilde(a)_ij?).
+
+
+## Load required libraries
+library(deSolve) # numerical integration library
+library(rootSolve) # for runsteady integration function
+library(pse) # for latin hypercube
+library(plyr) # for ldply
+library(dplyr) # for manipulating data
+library(tidyr) # for tidying data
+library(cowplot) # for better base, ggplot graphics
+
+## color palette
+cbPalette <- c("#D55E00", "#0072B2", "#009E73")
+
+## set plot theme
+theme_set(theme_cowplot())
+
+#### GET REQUIRED FUNCTIONS ####
+
+source('identify_steady_state.R')
+source('eco_evo_sim_function.R')
+source('Consumer_Resource_Functions.R')
+source('CR_dynamics.R')
+
+#### GENERAL PARAMETERS AND STATE VARIABLES #### 
+
+## Intrinsic growth rates
+r <- 1
+r1 <- r + 0.1
+r2 <- r 
+
+## Carrying capacity
+K <- 4
+K1 <- K - 0.1
+K2 <- K
+
+## Conversion efficiency
+e <- 0.8
+e11 <- e + 0.1
+e12 <- e - 0.1
+e21 <- e 
+e22 <- e 
+
+## Mortality rate
+m <- 1
+m1 <- m
+m2 <- m
+
+## Habitat preference (for LawlorSmith and McCann models)
+w <- 0.61 # adjusted to 0.61 from 0.6 to maintain consistency with previous figure
+wii <- w
+wij <- (1 - w)
+
+## General, non-evolving parameter set for 2C,2R model
+general_parameters <- data.frame(r1 = r1, r2 = r2, K1 = K1, K2 = K2, e11 = e11, e12 = e12, e21 = e21, e22 = e22, m1 = m1, m2 = m2)
+
+## Attack rate
+A <- 2
+aii <- A # this will be multiplied by wii above. This helps maintain consistency with previous figure 3.
+
+## confirmed that I don't need to adjusted A (total attack rate investment)
[…]

### 2020-07-02 manuscript_ECD_model.Rmd
add author info, acknowledgements, unstack inline equations, and final edit from AE.

@@
   - \captionsetup[figure]{font=small}
-bibliography: ReferencesAbbrev.bib
+bibliography: References.bib
 biblio-style: amnat
@@
 
-<!-- remove for double-blind peer-review
 \bigskip
@@
 *Data accessibility*: No new data were used.
-All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model) and has been archived with Zenodo (https://zenodo.org/badge/latestdoi/79349192).
+All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model) and have been archived with Zenodo (https://zenodo.org/badge/latestdoi/79349192).
 
@@
 
-where consumer $j$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling time $h_{ij}$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ij}=\frac{w_{ij}R_i}{w_{1,j}R_1+w_{2,j}R_2}$.
+where consumer $j$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling time $h_{ij}$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ij}=w_{ij}R_i/(w_{1,j}R_1+w_{2,j}R_2)$.<!--$W_{ij}=\frac{w_{ij}R_i}{w_{1,j}R_1+w_{2,j}R_2}$.-->
 
@@
 I say consumers have undergone divergent character displacement if their evolved attack rates are more specialized when evolving with vs. without a competing consumer. 
-Specialization of consumer $j$ on resource $1$ is measured as $\frac{a_{1,j}}{a_{1,j}+a_{2,j}}$, where a value of 0.5 is a complete generalist ($a_{1,j}=a_{2,j}$), and a value of 1 is a complete specialist ($a_{2,j}=0$).
+Specialization of consumer $j$ on resource $1$ is measured as $a_{1,j}/(a_{1,j}+a_{2,j})$<!--$\frac{a_{1,j}}{a_{1,j}+a_{2,j}}$-->, where a value of 0.5 is a complete generalist ($a_{1,j}=a_{2,j}$), and a value of 1 is a complete specialist ($a_{2,j}=0$).
 Values less than 0.5 indicate specialization on the other resource. 
@@
 If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource abundances to reach a steady state. 
-I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g., $\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0).
+I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g., $a_{ij}/(a_{\text{1},j}+a_{\text{2},j})$<!--$\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$--> is constrained to a maximum of 1 and minimum of 0).
 
@@
 Although data on the shape of the tradeoff in consumer foraging traits is scarce, two classic examples of character displacement, Darwin's finches and threespine stickleback, both appear to exhibit a tradeoff where extreme trait values increase the net foraging rate of consumers [@Schluter1985; @Arnegard2014].
-While it is theoretically possible that character displacement does not alter (or even increase) resource abundances, this was limited to the simplest, and arguably least realistic, foraging scenario and under tradeoffs that did not favor large displacements, and thus less likely to detect in nature.
+While it is theoretically possible that character displacement does not alter (or even increase) resource abundances, this was limited to the simplest, and arguably least realistic, foraging scenario and under tradeoffs that did not favor large displacements, and is thus less likely to detect in nature.
 <!--Taken together, my results call for empirical work to test these clear theoretical predictions and suggest a revision is needed for one of the criteria used to demonstrate character displacement.-->
@@
 
-<!-- Remove for double-blind review
 # Acknowledgements
 
-This work was inspired by discussions with Seth Rudman, Dolph Schluter, Ben Gilbert, and Kevin McCann.
-Sally Otto provided much help in early analyses of the mathematical model. 
+This work was inspired by discussions with Seth Rudman, Dolph Schluter, Benjamin Gilbert, and Kevin McCann.
+Sarah Otto provided much help in early analyses of the mathematical model. 
 I would not have been able to take the theory as far as I did without her guidance and encouragement.
-Matt Osmond, Jean Gibert, and Seth Rudman provided critical feedback on an earlier draft.
+Matthew Osmond, Jean Philippe Gibert, and Seth Rudman provided critical feedback on an earlier draft.
 For funding support, I thank the University of British Columbia (Four-Year Fellowship to M.A.
 Barbour), NSERC (Discovery grant to Greg Crutsinger), and the Swiss National Science Foundation (grant 31003A_160671 to Jordi Bascompte).
--->
 

### 2020-07-02 manuscript_ECD_model.tex
add author info, acknowledgements, unstack inline equations, and final edit from AE.

@@
 \begin{center} \LARGE Ecological character displacement destabilizes food webs \end{center}
+
+\bigskip
+
+\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}
+
+\bigskip
+
+\noindent 1. University of Zurich, Department of Evolutionary Biology
+and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich,
+Switzerland;
+
+\noindent 2. University of British Columbia, Department of Zoology, 2212
+Main Mall, Vancouver, BC V6T 1Z4, Canada;
+
+\(^\ast\) Corresponding author; e-mail:
+\href{mailto:matthew.barbour@ieu.uzh.ch}{\nolinkurl{matthew.barbour@ieu.uzh.ch}},
+phone: +41 44 635 61 63
+
+\bigskip
+
+\emph{Statement of authorship}: MAB conceived the study, derived
+analytical solutions, performed simulations, and wrote the paper.
+
+\bigskip
+
+\emph{Data accessibility}: No new data were used. All code used to
+reproduce this manuscript and simulations within are publically
+available on GitHub (\url{https://github.com/mabarbour/ECD_model}) and
+have been archived with Zenodo
+(\url{https://zenodo.org/badge/latestdoi/79349192}).
+
+\bigskip
 
@@
 preferences are modified by the relative abundance of resources, such
-that: \(W_{ij}=\frac{w_{ij}R_i}{w_{1,j}R_1+w_{2,j}R_2}\).
+that: \(W_{ij}=w_{ij}R_i/(w_{1,j}R_1+w_{2,j}R_2)\).
 
@@
 Specialization of consumer \(j\) on resource \(1\) is measured as
-\(\frac{a_{1,j}}{a_{1,j}+a_{2,j}}\), where a value of 0.5 is a complete
+\(a_{1,j}/(a_{1,j}+a_{2,j})\), where a value of 0.5 is a complete
 generalist (\(a_{1,j}=a_{2,j}\)), and a value of 1 is a complete
@@
 \citep[ESS,][]{Smith1973} or an evolutionary limit (e.g.,
-\(\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}\) is constrained to a
-maximum of 1 and minimum of 0).
+\(a_{ij}/(a_{\text{1},j}+a_{\text{2},j})\) is constrained to a maximum
+of 1 and minimum of 0).
 
@@
 realistic, foraging scenario and under tradeoffs that did not favor
-large displacements, and thus less likely to detect in nature.
+large displacements, and is thus less likely to detect in nature.
 
@@
 
-\bibliography{ReferencesAbbrev.bib}
+\section{Acknowledgements}\label{acknowledgements}
+
+This work was inspired by discussions with Seth Rudman, Dolph Schluter,
+Benjamin Gilbert, and Kevin McCann. Sarah Otto provided much help in
+early analyses of the mathematical model. I would not have been able to
+take the theory as far as I did without her guidance and encouragement.
+Matthew Osmond, Jean Philippe Gibert, and Seth Rudman provided critical
+feedback on an earlier draft. For funding support, I thank the
+University of British Columbia (Four-Year Fellowship to M.A.~Barbour),
+NSERC (Discovery grant to Greg Crutsinger), and the Swiss National
+Science Foundation (grant 31003A\_160671 to Jordi Bascompte).
+
+\bibliography{References.bib}
 
